### 2017-11-28 paper/lagrangeSpontaneousSimilarity2017.tex
demo and data

% -----------------------------------------------
% Template for ISMIR Papers
% 2017 version, based on previous ISMIR templates

% Requirements :
% * 6+n page length maximum
% * 4MB maximum file size
% * Copyright note must appear in the bottom left corner of first page
% * Clearer statement about citing own work in anonymized submission
% (see conference website for additional details)
% -----------------------------------------------

\documentclass{article}
\usepackage{ismir,amsmath,cite,url}
\usepackage{graphicx}
\usepackage{color}

% Title.
% ------
\title{Spontaneous Similarity Judgments of Musical Tones with various Playing Techniques}

% Note: Please do NOT use \thanks or a \footnote in any of the author markup

% Single address
% To use with only one author or several with the same address
% ---------------
%\oneauthor
% {Names should be omitted for double-blind reviewing}
% {Affiliations should be omitted for double-blind reviewing}

% Two addresses
% --------------
%\twoauthors
%  {First author} {School \\ Department}
%  {Second author} {Company \\ Address}

%% To make customize author list in Creative Common license, uncomment and customize the next line
%  \def\authorname{First Author, Second Author}

% Three addresses
% --------------
\threeauthors
  {Mathieu Lagrange} {Affiliation1 \\ {\tt author1@ismir.edu}}
  {Mathias Rossignol} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
  {Grégoire Lafay} {Affiliation3 \\ {\tt author3@ismir.edu}}

%% To make customize author list in Creative Common license, uncomment and customize the next line
%  \def\authorname{First Author, Second Author, Third Author}

% Four or more addresses
% OR alternative format for large number of co-authors
% ------------
%\multauthor
%{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
%  $^1$ Department of Computer Science, University , Country\\
%$^2$ International Laboratories, City, Country\\
%$^3$  Company, Address\\
%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
%}
%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}

\sloppy % please retain sloppy command for improved formatting

\begin{document}

%
\maketitle
%
\begin{abstract}
The abstract should be placed at the top left column and should contain about 150-200 words.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

\section{}\label{sec:}

% For bibtex users:
\bibliography{bib}

% For non bibtex users:
%\begin{thebibliography}{citations}
%
%\bibitem {Author:00}
%E. Author.
%``The Title of the Conference Paper,''
%{\it Proceedings of the International Symposium
%on Music Information Retrieval}, pp.~000--111, 2000.
%
%\bibitem{Someone:10}
%A. Someone, B. Someone, and C. Someone.
%``The Title of the Journal Paper,''
%{\it Journal of New Music Research},
%Vol.~A, No.~B, pp.~111--222, 2010.
%
%\bibitem{Someone:04} X. Someone and Y. Someone. {\it Title of the Book},
%    Editorial Acme, Porto, 2012.
%
%\end{thebibliography}

\end{document}

### 2017-11-28 demo/Sample.js
demo and data

@@ -1,0 +1,61 @@
+function Sample (url, callback, audioContext) {
+    this.ac = audioContext;
+    this.callback = callback;
+    url = url.replace("#", "%23");
+    if (this.ac) {
+	var request = new XMLHttpRequest();
+	request.open('GET', url, true);
+	request.responseType = 'arraybuffer';
+	request.onload = function (request) {
+	    this.ac.decodeAudioData(request.target.response, Sample.prototype.bufferInit.bind(this),
+				    function () {alert("Problème de décodage du fichier audio\n"+url);});
+	}.bind(this);
+	request.send();
+    } else {
+	this.audio = new Audio(url);
+	this.audio.autoplay = false;
+	var checkInterval = setInterval(function(){
+	    if (this.audio.readyState == HTMLMediaElement.HAVE_ENOUGH_DATA) {
+		clearInterval(checkInterval);
+		this.callback(this);
+	    }
+	}.bind(this), 100);
+    }
+}
+
+Sample.prototype.bufferInit = function(buffer) {
+    this.buffer = buffer;
+    this.bufferLength = this.buffer.getChannelData(0).length;
+    this.duration = this.bufferLength/this.buffer.sampleRate;
+    if (this.callback) this.callback(this);
+}
+
+Sample.prototype.play = function (amp) {
+    if (!amp) amp = 1;
+    if (this.ac) {
+	this.gainNode = this.ac.createGain();
+	this.gainNode.gain.value = amp;
+	this.gainNode.connect(this.ac.destination);
+	this.bufferNode = this.ac.createBufferSource();
+	this.bufferNode.buffer = this.buffer;
+	this.bufferNode.connect(this.gainNode);
+	this.bufferNode.start(this.ac.currentTime+.01);
+    } else {
+	this.audio.volume = amp;
+	this.audio.currentTime = 0;
+	this.audio.play();
+    }
+}
+
+Sample.prototype.stop = function () {
+    if (this.bufferNode) {
+	var time = this.ac.currentTime;
+	this.gainNode.gain.setValueAtTime(this.gainNode.gain.value, time);
+	this.gainNode.gain.linearRampToValueAtTime(0, time+.1);
+	this.bufferNode.stop(time+.2);
+	this.bufferNode = this.gainNode = null;
+    } else if (this.audio) {
+	this.audio.pause();
+    }
+    this.playing = false;
+}

### 2017-11-28 demo/app.js

@@ -1,0 +1,6 @@
+app = angular.module('grouping', []);
+
+app.config(['$compileProvider',
+	    function($compileProvider) {   
+		$compileProvider.aHrefSanitizationWhitelist(/^\s*(https?|ftp|mailto|blob|file):/);
+	    }]);

### 2017-11-28 demo/filter.js

@@ -1,0 +1,24 @@
+app = angular.module('filter', []);
+
+app.controller("Filter", function ($scope, $http) {
+
+    $scope.toJson = function(o){return angular.toJson(o, 2);};
+    $scope.result = [];
+
+    $http.get("ticelxp1_subject_mix_tresh_11.json").then(function(response) {
+	response.data.objects.forEach(function(o) {
+	    if (o.sounds.some(function(s){return s.match(/C4/);})) {
+		var s = o.sounds.find(function(s){return s.match(/\-mf/) && s.match(/C4/);})
+		if (s) {
+		    $scope.result.push({instrument:o.instrument, mode:o.mode, hasC4:true, hasMf:true, path:o.path+"/"+s});
+		} else {
+		    var sounds = [];
+		    s = o.sounds.forEach(function(s){if(s.match(/C4/)) sounds.push(s);})
+		    $scope.result.push({instrument:o.instrument, mode:o.mode, hasC4:true, hasMf:false, path:o.path+"/"+sounds.join(" ")});
+		}
+	    } else {
+		$scope.result.push({instrument:o.instrument, mode:o.mode, hasC4:false, hasMf:false, path:o.path+"/"+o.sounds.join(" ")});
+	    }
+	});
+    });
+});

### 2017-11-28 demo/multilingualGroupSoundsNodeApp.js

@@ -1,0 +1,146 @@
+var server = require('http').createServer();
+var io = require('socket.io').listen(server);
+var md5 = require('js-md5');
+var parser = require('ua-parser-js');
+const fs = require('fs');
+
+var port = 3007;
+
+// Load base sound data
+var soundData = require("./soundList.json");
+soundData.sounds.forEach(function(s,i) {
+    var col = Math.floor(i/6);
+    var row = i%6;
+    s.x = col/13 + .5*Math.random()/13;
+    s.y = row/6 + .5*Math.random()/6;
+});
+
+console.log("Loaded sound data, current version = "+soundData.version);
+
+var nodemailer = require('nodemailer');
+
+// create reusable transporter object using the default SMTP transport
+var transporter = nodemailer.createTransport({
+     service: 'gmail', // no need to set host or port etc.
+     auth: {
+	 user: "mathias.rossignol@gmail.com",
+	 pass: "inekmiggycpuqjmf"
+     }
+});
+
+
+io.on('connection', function(socket) {
+    console.log("User connected");
+    var userEmail = "";
+    var fileName = "";
+    var language = "fr";
+    var userData = null;
+    var finalized = false;
+    var acceptContact = false;
+    var browsers = [];
+
+    socket.on('checkUser', function(data) {
+	var email = data.email;
+	if (data.hasOwnProperty("language")) language = data.language;
+	// If user exists, loads json data containing last state of organization saved for this user into "userData" and sends it as response
+	// Else notifies the client that user doesn't exist
+	userEmail = email;
+	var emailEnc = md5(email);
+	fileName = "userData/"+emailEnc+".json";
+	fs.stat(fileName, function(err, stats) {
+	    socket.emit('userCheckResult', {'exists':err==null});
+	    if (!err) {
+		console.log("Known user "+userEmail);
+		fs.readFile(fileName, function(err, uData) {
+		    var dat = JSON.parse(uData);
+		    if (!data.hasOwnProperty("language") && dat.hasOwnProperty("language"))
+			language = dat.language;
+		    userData = dat.organization;
+		    finalized = dat.finalized;
+		    acceptContact = dat.acceptContact;
+		    if (dat.browsers) browsers = dat.browsers;
+		    browsers.push(parser(socket.request.headers['user-agent']));
+		    socket.emit('userData', userData);
+		});
+	    }
+	});
+    });
+
+    socket.on('newUser', function(data) {
+	var email = data.email;
[…]

### 2017-11-28 demo/organizer.js

@@ -1,0 +1,204 @@
+app.controller("Organizer", function($scope, $http, $timeout, $log) {
+
+    var ac = getAudioContext();
+    $scope.points = [];
+    $scope.soundsToLoad = 0;
+    $scope.loadedSounds = 0;
+
+    $scope.playFieldWidth = 600;
+    $scope.playFieldHeight = 600;
+    $scope.playFieldTop = 50;
+    $scope.playFieldLeft = 50;
+    $scope.pointHalfSize = 10;
+
+    $scope.colors = colorPalette;
+
+    var soundType = "webm";
+    var uap = new UAParser();
+    $log.log(JSON.stringify(uap.getResult()));
+    if (uap.getBrowser().name.match(/IE/) || uap.getBrowser().name.match(/Edge/)) soundType = "mp3";
+
+
+    var displayX = function(x) {return $scope.pointHalfSize+Math.floor(x*($scope.playFieldWidth-2*$scope.pointHalfSize));}
+    var displayY = function(y) {return $scope.pointHalfSize+Math.floor(y*($scope.playFieldHeight-2*$scope.pointHalfSize));}
+    var displayW = function(w) {return Math.floor(w*($scope.playFieldWidth-2*$scope.pointHalfSize));}
+    var displayH = function(h) {return Math.floor(h*($scope.playFieldHeight-2*$scope.pointHalfSize));}
+
+    var minHPH = 100;
+    var maxHPH = 320;
+    $scope.hPanelHeight = 100;
+    $scope.hPanelWidth = 100;
+    var minVPW = 250;
+    var maxVPW = 400;
+    $scope.vPanelHeight = 200;
+    $scope.vPanelWidth = 200;
+    $scope.verticalLayout = false;
+
+    var reScale = function () {
+	var w = window.innerWidth;
+	if (w<700) w=700;
+	var h = window.innerHeight;
+	if (h<500) h=500;
+	if (h-minHPH > w-minVPW) {
+	    $scope.verticalLayout = true;
+	    $scope.hPanelHeight = Math.min(maxHPH, Math.max(minHPH, h-w-10));
+	    $scope.hPanelWidth = w-20;
+	    $scope.playFieldHeight = $scope.playFieldWidth = Math.min(w-20, h-$scope.hPanelHeight-30);
+	    $scope.playFieldTop = $scope.hPanelHeight+10+Math.round((h-$scope.playFieldHeight-$scope.hPanelHeight-10)/2);
[…]

### 2017-11-28 demo/rgbcolor.js

@@ -1,0 +1,240 @@
+/**
+ * A class to parse color values
+ * @author Stoyan Stefanov <sstoo@gmail.com>
+ * @link   http://www.phpied.com/rgb-color-parser-in-javascript/
+ * @license Use it if you like it
+ * 
+ * Some modification by Mathias Rossignol, 2012/11/16
+ */
+function RGBColor(src) {
+    if (!src) src = "#888";
+    if (src instanceof RGBColor) {
+	this.r = src.r;
+	this.g = src.g;
+	this.b = src.b;
+	this.ok = src.ok;
+    } else if (src instanceof Array) {
+	if (src[0]>1 || src[1]>1 || src[2]>1) {
+	    this.r = Math.round(src[0]);
+	    this.g = Math.round(src[1]);
+	    this.b = Math.round(src[2]);
+	} else {
+	    this.r = Math.round(255*src[0]);
+	    this.g = Math.round(255*src[1]);
+	    this.b = Math.round(255*src[2]);
+	}
+	this.ok = true;
+    } else {
+	this.ok = false;
+	var color_string = src;
+	// strip any leading #
+	if (color_string.charAt(0) == '#') { // remove # if any
+            color_string = color_string.substr(1,6);
+	}
+
+	color_string = color_string.replace(/ /g,'');
+	color_string = color_string.toLowerCase();
+
+	// before getting into regexps, try simple matches
+	// and overwrite the input
+	var simple_colors = {
+            aliceblue: 'f0f8ff', antiquewhite: 'faebd7', aqua:
+            '00ffff', aquamarine: '7fffd4', azure: 'f0ffff', beige:
+            'f5f5dc', bisque: 'ffe4c4', black: '000000',
+            blanchedalmond: 'ffebcd', blue: '0000ff', blueviolet:
+            '8a2be2', brown: 'a52a2a', burlywood: 'deb887', cadetblue:
+            '5f9ea0', chartreuse: '7fff00', chocolate: 'd2691e',
+            coral: 'ff7f50', cornflowerblue: '6495ed', cornsilk:
+            'fff8dc', crimson: 'dc143c', cyan: '00ffff', darkblue:
+            '00008b', darkcyan: '008b8b', darkgoldenrod: 'b8860b',
+            darkgray: 'a9a9a9', darkgreen: '006400', darkkhaki:
+            'bdb76b', darkmagenta: '8b008b', darkolivegreen: '556b2f',
[…]

### 2017-11-28 demo/sounds.js

@@ -1,0 +1,548 @@
+sounds = [
+    {
+	"instrument": "Harp",
+	"mode": "xylophonic",
+	"hasC4": true,
+	"hasMf": false,
+	"file": "Hp-xyl-C4-p"
+    },
+    {
+	"instrument": "Alto-Sax",
+	"mode": "key-click",
+	"hasC4": true,
+	"hasMf": false,
+	"file": "ASax-key-cl-C4-p"
+    },
+    {
+	"instrument": "Alto-Sax",
+	"mode": "slap-pitched",
+	"hasC4": true,
+	"hasMf": true,
+	"file": "ASax-slap-C4-mf"
+    },
+    {
+	"instrument": "Guitar",
+	"mode": "ordinario-high-register",
+	"hasC4": false,
+	"hasMf": false,
+	"file": "Gtr-ord-hi-reg-E5-mf-3c"
+    },
+    {
+	"instrument": "Guitar",
+	"mode": "pizzicato",
+	"hasC4": true,
+	"hasMf": true,
+	"file": "Gtr-pizz-C4-mf-2c"
+    },
+    {
+	"instrument": "Harp",
+	"mode": "harmonic-fingering",
+	"hasC4": true,
+	"hasMf": false,
+	"file": "Hp-harm-fngr-C4-f"
+    },
+    {
+	"instrument": "Harp",
+	"mode": "ordinario",
+	"hasC4": true,
+	"hasMf": true,
+	"file": "Hp-ord-C4-mf"
+    },
+    {
+	"instrument": "Harp",
+	"mode": "pizzicato-bartok",
+	"hasC4": true,
+	"hasMf": false,
+	"file": "Hp-pizz-bartok-C4-ff"
+    },
+    {
+	"instrument": "Horn",
+	"mode": "ordinario",
+	"hasC4": true,
+	"hasMf": true,
+	"file": "Hn-ord-C4-mf"
+    },
+    {
+	"instrument": "Horn",
+	"mode": "slap-pitched",
+	"hasC4": true,
+	"hasMf": true,
+	"file": "Hn-slap-C4-mf"
+    },
+    {
+	"instrument": "Oboe",
+	"mode": "key-click",
+	"hasC4": true,
+	"hasMf": false,
+	"file": "Ob-key-cl-C4-pp"
+    },
+    {
+	"instrument": "Tenor-Trombone",
+	"mode": "cup-ordinario",
+	"hasC4": true,
[…]

### 2017-11-28 demo/uatest.js

@@ -1,0 +1,12 @@
+var http = require('http');
+var parser = require('ua-parser-js');
+
+http.createServer(function (req, res) {
+    // get user-agent header
+    var ua = parser(req.headers['user-agent']);
+    // write the result as response
+    res.end(JSON.stringify(ua, null, '  '));
+})
+.listen(1337, '127.0.0.1');
+
+console.log('Server running at http://127.0.0.1:1337/');

### 2017-11-28 demo/utils.js

@@ -1,0 +1,137 @@
+if (!Array.prototype.multiplyBy) {
+    Array.prototype.multiplyBy = function(f) {
+	var a = f; 
+	if (!a.length) a = [a];
+	var l = a.length;
+	if (l>this.length) l = this.length;
+	for (var i=0; i<l; i++) this[i] *= a[i];
+	for (var i=l; i<this.length; i++) this[i] *= a[l-1];
+    }
+}
+
+if (!Array.prototype.add) {
+    Array.prototype.add = function(f) {
+	var a = f; 
+	if (!a.length) a = [a];
+	var l = a.length;
+	if (l>this.length) l = this.length;
+	for (var i=0; i<l; i++) this[i] += a[i];
+	for (var i=l; i<this.length; i++) this[i] += a[l-1];
+    }
+}
+
+if (!Array.prototype.copy) {
+    Array.prototype.copy = function() {
+	var n = [];
+	for (var i=0; i<this.length; i++) n[i] = this[i];
+	return n;
+    }
+}
+
+
+
+if (!Function.prototype.curry) {
+    Function.prototype.curry = function() {
+	if (arguments.length<1) return this;
+	var __method = this;
+	var args = Array.prototype.slice.call(arguments);
+	return function() {
+	    return __method.apply(this, args.concat(Array.prototype.slice.call(arguments)));
+	}
+    }
+}
+
+if (!Function.prototype.bind) {
+    Function.prototype.bind = function(object) {
+	var __method = this;
+	return function() {
+	    return __method.apply(object, arguments);
+	}
+    }
+}
+
+if (!Object.prototype.clone) {
+    Object.prototype.clone = function () {
+	var target = {};
+	for (var i in this) {
+	    if (this.hasOwnProperty(i)) {
+		target[i] = obj[i];
+	    }
+	}
+	return target;
+    }
+}
+
+if (!Object.prototype.inherit) {
+    Object.prototype.inherit = function(baseConstructor) {
+	this.prototype = baseConstructor.prototype.clone();
+	this.prototype.constructor = this;
+    };
+}
+
+// Defining a subclass, cheat sheet :
+
+// function Subclass (sameParams) {
+//     Superclass.call(this, sameParams);
+//     this.otherStuff = "Hi there!";
+// }
+// Subclass.inherit(Superclass);
+
+// Subclass.prototype.newFun = function () {
+//     console.log(this.otherStuff);
+// }
+
[…]

### 2017-12-04 paper/lagrangeSpontaneousSimilarity2017.tex
abs ok, start intro

@@
 \documentclass{article}
-\usepackage{ismir,amsmath,cite,url}
+\usepackage{amsmath,cite,url}
 \usepackage{graphicx}
@@
 % ------
-\title{Spontaneous Similarity Judgments of Musical Tones with various Playing Techniques}
+\title{Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques}
 
@@
 % --------------
-\threeauthors
-  {Mathieu Lagrange} {Affiliation1 \\ {\tt author1@ismir.edu}}
-  {Mathias Rossignol} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-  {Grégoire Lafay} {Affiliation3 \\ {\tt author3@ismir.edu}}
+\author{Mathieu Lagrange, Christian EL-Hajj, Mathias Rossignol, Grégoire Lafay}
 
@@
 \begin{document}
-
 %
@@
 \begin{abstract}
-The abstract should be placed at the top left column and should contain about 150-200 words.
+
+Musical timbre is a multi-faceted notion that have been extensively studied, mostly focusing on its spectral aspect.
+By studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques, we aim in this paper at better understanding the influence of the change of the playing technique on the perception of timbre.
+
+A first experiment is conducted by music experts to identify which couple of instrument and playing technique in a given dataset is worth comparing to other couples of instrument / playing technique.
+In a second experiment, spontaneous similarity judgments among the retained couples are collected using a canonic free sorting task experiment design.
+
+To further study the outcomes of this experiment, we assume a two-step computational model of human perception, where the acoustic signal is fed to a statically designed processing unit that accounts for frequency and temporal modulations.
+The resulting features are then projected using a supervised technique which considers as input the perceptual similarity judgments gathered in the second experiment.
+Numerical experiments show that the induced perceptual space is able to approximate perceptual data with satisfying accuracy.
+
 \end{abstract}
@@
 
-\section{}\label{sec:}
+Spontaneous similarity judgments
+
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational of timbre perception that account well for the gathered human judgments.
+
+\section{Dimensions of Timbre}\label{sec:}
+
+\section{Data}\label{sec:}
+
+\section{Experiment 1}\label{sec:}
+
+Problématique
+
+Dans le cadre du projet TICEL, nous allons construire un algorithme permettant de mesurer les similarités entre différents modes de jeu.
+Par mode de jeu, nous entendons ici un couple "instrument + mode de jeu".
+
+Afin de mesurer les performances de l'algorithme, nous avons besoins d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.
+
+Or il n'est actuellement pas envisageable de monter une expérience permettant d'établir cette vérité terrain. Le nombre de modes de jeu à comparer est trop important.
+
+C'est pourquoi nous proposons une expérience intermédiaire ayant pour but de réduire le nombre de modes de jeu.
+
+Objectif
+
+Pour chaque mode de jeu, vous répondrez à la question suivante : Est il intéressant de comparer ce mode de jeu à un/pls autre(s) mode(s) d'un autre instrument
+
+Procédure
+
+Nous vous demandons de noter chaque mode de jeu avec un nombre d’étoiles.
+Plus le nombre d'étoiles est important, plus le mode de jeu est susceptible d'être associé de manière "analogique" à un autre mode de jeu d'un autre instrument.
+Pour exemple :
+
+Une étoile : ce mode de jeu est singulier, il n’est pas utile de le comparer avec un autre mode de jeu d'un autre instrument.
+Quatre étoiles : il y a bien une proximité sur un aspect du son entre ce mode de jeu et un autre mode d'un autre instrument, mais cette dernière n'est ni décisive ni évidente.
+Sept étoiles : il a une grande proximité entre ce mode de jeu et un autre mode d'un autre instrument, il est donc intéressant de le conserver.
+Déroulement
+
+Vous pouvez réaliser l'expérience en plusieurs fois, mais veillez à noter l'ensemble des modes de jeu pour tous les instruments.
+
+Usage
+
+Votre notation doit prendre en compte l'ensemble des modes de jeu présents dans la base, bien que par souci de clarté, les modes de jeu soient présentés par instruments
+Les modes de jeu déjà notés apparaîtront en vert
+Pour chaque mode de jeu, vous pouvez écouter plusieurs enregistrements à différentes hauteurs et nuances
+
+\section{Experiment 2}\label{sec:}
+
+\section{Modeling}\label{sec:}
 

### 2017-12-05 paper/lagrangeSpontaneousSimilarity2017.tex
working on intro

@@
 
+Musical timbre is an interesting notion to study as it is perceptively defined.
+Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse -- but somewhat more controled than other stimuli such as speech or environemntal sounds -- set of physical stimuli.
+
+A popular view of the kind of stimuli we consider is that they can be described by a spread of energy that evolve through time and across frequency.
+By following this trend, most researchers translate the negation of the ASA definition of timbre into a more relative definition : musical timbre is about the distribution of intensity across the time / frequency plane.
+
+A great deal of litterature focuses on the instaneous spectrum, \textit{i.e.} the timbre can be defined without considering evolution of the energy distribution trough time.
+Other studies demonstrated the importance of the onset in the recognition of some musical instruments.
+Those studies on a rather short term model where the waveform is observed in a range that do not exceed 100 ms.
+
+Interestingly, modern physiological models of the primary mamalian auditory system  .
+Dau's model acount for frequency and rate of modulation.
+Shamma's model account for frequency, rate and bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret.
+Though, most of the numerical experiments done using this model did not fully demonstrated the usefullness of this third dimension.
+
+Thus, by taking a signal processing view of the matter, one could consider that a first stage of the auditory system proceeds to a rather fixed set of convolutions across time and frequency, projecting the signal into a large set of descriptive features organized across frequency and rate of modulations.
+
+Thus the second step would be rather opportunistic, in the sense that, from this large set of features, one can pick the ones that are the most relevant for the task at hand.
+
+Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider a notion of dissimilarity and not the
+
+Following .
+
 Spontaneous similarity judgments
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational of timbre perception that account well for the gathered human judgments.
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper.
 

### 2017-12-05 paper/lagrangeSpontaneousSimilarity2017.tex
xp1

@@
 
-Interestingly, modern physiological models of the primary mamalian auditory system  .
+Interestingly, modern physiological models of the primary mamalian auditory system also considers the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
 Dau's model acount for frequency and rate of modulation.
@@
 
-Thus the second step would be rather opportunistic, in the sense that, from this large set of features, one can pick the ones that are the most relevant for the task at hand.
+What the higher level of the auditory cortex does remains largely unkown, but we can assume that the level of plasticity is very high, as any living entity has to adapt their percepts to a wide diversity of tasks.
+Considering the lack of information about this stage, we will assume in this paper that this second step is rather opportunistic, in the sense that, from this large set of features, it aims at multiplexing in some way the ones that are the most relevant for the task at hand.
 
-Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider a notion of dissimilarity and not the
-
-Following .
-
-Spontaneous similarity judgments
+Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts.
+A notion that have not been extensively studied in the litterature.
+It also permits to consider the notion of timbral similarity at finer grain of detail.
 
@@
 
-\section{Dimensions of Timbre}\label{sec:}
+\section{Experiment 1}\label{sec:}
 
-\section{Data}\label{sec:}
+The dataset considered in this study has audio recording of 16 different  musical instruments () played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to 25444 samples.
+The way the dataset is segmented is explained in more details in Appendix \ref{sec:dataset}.
 
-\section{Experiment 1}\label{sec:}
+Studies about perceptual similarity are plagued with a dimensional problem.
+When considering $n$ items, a complete exploration of the similarity space requires the filling of a $n^2$ matrix.
+Assuming a symmetric similarity (the similarity of A to B is equl to the similarity of B to A) reduces the number but not by much.
+
+In an attempt to reduce the number of items, we decided to perform a selection experiment where the subject is asked to give his opinion on which \ipt is relevant to study.
+An \ipt is said to be relevant to study if it is likely to be associated to another \ipt.
+Interest is rated on a 7 ticks scale.
+
+Examples are given :
+\begin{itemize}
+  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt.
+  \item Four stars: there is a proximity on one aspect of the sound between this \ipt and of an \ipt of another instrument, but this one is neither decisive nor obvious.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipn of another instrument, it is interesting to keep it.
+\end{itemize}
+
+Select which couple (instrument, playing technique) are interesting to consider.
+Question: Is it interesting to compare this couple to another couple of a different instrument ?
+The question is asked for each couple ()
+The subject can listen to all the recordings of each couple
+
+Two musical composition professors of the CNSMDP, a renowned French school of music did the test. All the couples with a grade higher than three are selected, bringing 78 couples to study.
+
+The aim of experiment 1
 
@@
 
-Afin de mesurer les performances de l'algorithme, nous avons besoins d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.
+Afin de mesurer les performances de l'algorithme, nous avons besoin d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.
 
@@
 
-\section{Modeling}\label{sec:}
+\section{Analysis}\label{sec:}
+
+\section{Model}\label{sec:}
+
+\section{Results}\label{sec:}
+
+\section{Discussion}\label{sec:}
 
@@
 \bibliography{bib}
+
+\Appendix
+
+\section{Dataset}
 

### 2017-12-06 paper/lagrangeSpontaneousSimilarity2017.tex
adding figures

@@
 \usepackage{color}
+
+\newcommand{\ipt}{IPT}
+
+\newcommand{\ml}[1]{\textcolor{blue}{ML : #1}}
 
@@
 Interest is rated on a 7 ticks scale.
+The subject can listen to all the different nuance and pitch samples of each \ipt.
+The experiment is over when all the \ipt have been rated.
 
-Examples are given :
+Rating examples are given :
 \begin{itemize}
@@
   \item Four stars: there is a proximity on one aspect of the sound between this \ipt and of an \ipt of another instrument, but this one is neither decisive nor obvious.
-  \item Seven stars: there is a large similarity between this \ipt and an \ipn of another instrument, it is interesting to keep it.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument, it is interesting to keep it.
 \end{itemize}
 
-Select which couple (instrument, playing technique) are interesting to consider.
-Question: Is it interesting to compare this couple to another couple of a different instrument ?
-The question is asked for each couple ()
-The subject can listen to all the recordings of each couple
-
-Two musical composition professors of the CNSMDP, a renowned French school of music did the test. All the couples with a grade higher than three are selected, bringing 78 couples to study.
-
-The aim of experiment 1
-
-Problématique
-
-Dans le cadre du projet TICEL, nous allons construire un algorithme permettant de mesurer les similarités entre différents modes de jeu.
-Par mode de jeu, nous entendons ici un couple "instrument + mode de jeu".
-
-Afin de mesurer les performances de l'algorithme, nous avons besoin d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.
-
-Or il n'est actuellement pas envisageable de monter une expérience permettant d'établir cette vérité terrain. Le nombre de modes de jeu à comparer est trop important.
-
-C'est pourquoi nous proposons une expérience intermédiaire ayant pour but de réduire le nombre de modes de jeu.
-
-Objectif
-
-Pour chaque mode de jeu, vous répondrez à la question suivante : Est il intéressant de comparer ce mode de jeu à un/pls autre(s) mode(s) d'un autre instrument
-
-Procédure
-
-Nous vous demandons de noter chaque mode de jeu avec un nombre d’étoiles.
-Plus le nombre d'étoiles est important, plus le mode de jeu est susceptible d'être associé de manière "analogique" à un autre mode de jeu d'un autre instrument.
-Pour exemple :
-
-Une étoile : ce mode de jeu est singulier, il n’est pas utile de le comparer avec un autre mode de jeu d'un autre instrument.
-Quatre étoiles : il y a bien une proximité sur un aspect du son entre ce mode de jeu et un autre mode d'un autre instrument, mais cette dernière n'est ni décisive ni évidente.
-Sept étoiles : il a une grande proximité entre ce mode de jeu et un autre mode d'un autre instrument, il est donc intéressant de le conserver.
-Déroulement
-
-Vous pouvez réaliser l'expérience en plusieurs fois, mais veillez à noter l'ensemble des modes de jeu pour tous les instruments.
-
-Usage
-
-Votre notation doit prendre en compte l'ensemble des modes de jeu présents dans la base, bien que par souci de clarté, les modes de jeu soient présentés par instruments
-Les modes de jeu déjà notés apparaîtront en vert
-Pour chaque mode de jeu, vous pouvez écouter plusieurs enregistrements à différentes hauteurs et nuances
+Two musical composition professors of the CNSMDP, a renowned French school of music performed the test. All the couples with a grade higher than three are selected, bringing 78 \ipt to study.
 
@@
 
+The second experiment aims at gathering spontaneous judgment of similarity.
+We chose a canonical free sorting task  for several reasons.
+First, it gives the ability to the subject to listen to the whole set of \ipt before performing the task.
+Second, it gives more freedom to the subject thus potentially more interest to perform the task than a lengthy series of triplet dissimilarity ratings.
+
+The subjects are thus asked to organize the 78 \ipt displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their acoustic similarity.
+The \ml{consigne} is deliberately vague in order to allow the subject to \ml{use its own representation of similarity}.
+The experiment is over when the sujbect assigned a color to all the dots.
+The initial location of the dots on the plane is randomized for each subject and can be changed during the experiment by the subject for convenience.
+Eventhough the location of the dots is recorded, this data is not considered during the analysis.
+
+The experiment was available online for two months and requests to perform the test have been sent to internal mailing lists of the CNSMDP and international research mailing with a focus on audio and music processing.
+The experiment have been performed on a volountary basis without gratification.
+
 \section{Analysis}\label{sec:}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/xp2example.png}
+\caption{Spatial and group organization provided by subject 1.}
+\label{fig:xp2display}
+\end{figure}
+
+31 subjects completed the experiment.
+An example of resulting organisation of \ipt is given on Figure \ref{fig:xp2display}.
+The resulting organisation for each subject are available for reader's inspection using the full feature interface provided to the subjects for performing the experiment\footnote{https://mathieulagrange.github.io/paperSpontaneousSimilarity/demo}.
+
+As discussed in the previous section, only the grouping done using the colors is considered.
+The spatial organisation could provide information about the notion of similarity, but implicitely force the fact that the similarity space is two dimensional.
+We therefore now study the properties of the groupings performed by the subject prior converting them into an overall similarity matrix.
+
+Among the 20 colors available, the subjects used on average \ml{fill}.
+The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbGroup}.
+The size of the groups is on average \ml{fill}.
+The distribution of the size of groups can be seen on Figure \ref{fig:xp2sizeGroup}.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/xp2example.png}
+\caption{Histogram of the number of groups.}
+\label{fig:xp2nbGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/xp2example.png}
+\caption{Histogram of the size of groups.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/allSim.png}
+\caption{Binary similarity matrix corresponding to the grouping of subject 1.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/allSim.png}
+\caption{Average similarity matrix corresponding to the grouping of the complete cohort.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/allLinkage.png}
+\caption{Dendrogram computed using the average similarity matrix.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
+\caption{Dendrogram computed using the average similarity matrix.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/groupModes.png}
+\caption{Clustering \ipt using average perceptual similarity into 5 clusters (a) and playing technique labels (b).
+The colors of clusters displayed on top corresponds to the playing technique label that is dominant in this cluster.}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/groupInstruments.png}
+\caption{Clustering \ipt using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
+The colors of clusters displayed on top corresponds to the playing technique label that is dominant in this cluster.}
+\label{fig:xp2sizeGroup}
+\end{figure}
 
@@
 \bibliography{bib}
-
-\Appendix
 

### 2017-12-06 paper/lagrangeSpontaneousSimilarity2017.tex
more on xp2

@@
 The second experiment aims at gathering spontaneous judgment of similarity.
-We chose a canonical free sorting task  for several reasons.
+To do so, we chose a canonical free sorting task  for several reasons.
 First, it gives the ability to the subject to listen to the whole set of \ipt before performing the task.
@@
 
-The subjects are thus asked to organize the 78 \ipt displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their acoustic similarity.
-The \ml{consigne} is deliberately vague in order to allow the subject to \ml{use its own representation of similarity}.
+The subjects are thus asked to organize the 78 \ipt displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
+The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to \ml{use its own representation of similarity}.
+
+The \ipt can be listened to by hovering the mouse on the corresponding dot.
 The experiment is over when the sujbect assigned a color to all the dots.
@@
 
-As discussed in the previous section, only the grouping done using the colors is considered.
-The spatial organisation could provide information about the notion of similarity, but implicitely force the fact that the similarity space is two dimensional.
-We therefore now study the properties of the groupings performed by the subject prior converting them into an overall similarity matrix.
+As discussed in the previous section, only the grouping given by the subjects using the color labels is considered to predict the spontaneous judgments of similarity among \ipt.
+The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the similarity space is two dimensional.
+
+We therefore now only study the properties of the groupings performed by the subject prior converting them into an overall similarity matrix.
 
@@
 
+A grouping can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
+The Figure \ref{fig:xp2bsim} show the binary similarity matrix that correponds to the grouping displayed on Figure \ref{fig:xp2display}.
+By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that correponds to our measure of the spontaneous similarity between \ipt.
+
+A cluster analysis is next performed using an agglomerative hierarchical clustering using the weighted average algorithm for computing distance between clusters.
+The resulting dendrogram is displayed on Figure \ref{fig:xp2sizeGroup}.
+
+A partition of $n$ clusters can be constructed from the agglomerative hierarchical cluster tree by finding the smallest height at which a horizontal cut through the tree leaves $n$ clusters.
+In order to
+
 \begin{figure}
@@
 \caption{Binary similarity matrix corresponding to the grouping of subject 1.}
-\label{fig:xp2sizeGroup}
+\label{fig:xp2bsim}
 \end{figure}
@@
 \caption{Average similarity matrix corresponding to the grouping of the complete cohort.}
-\label{fig:xp2sizeGroup}
+\label{fig:xp2asim}
 \end{figure}
@@
 \caption{Clustering \ipt using average perceptual similarity into 5 clusters (a) and playing technique labels (b).
-The colors of clusters displayed on top corresponds to the playing technique label that is dominant in this cluster.}
+The colors of clusters displayed on top corresponds to the playing technique that is dominant in this cluster.}
 \label{fig:xp2sizeGroup}
@@
 \caption{Clustering \ipt using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
-The colors of clusters displayed on top corresponds to the playing technique label that is dominant in this cluster.}
+The colors of clusters displayed on top corresponds to the intrument that is dominant in this cluster.}
 \label{fig:xp2sizeGroup}
@@
 \section{Dataset}
+
+distribution of instruments and modes
+
+\subsection{Naming}
+
+abbreviations 4 instruments and modes.
 

### 2017-12-08 paper/lagrangeSpontaneousSimilarity2017.tex
dendrogram

@@
 A notion that have not been extensively studied in the litterature.
-It also permits to consider the notion of timbral similarity at finer grain of detail.
+Compared to instrument class, the reference that is usually taken in timbre studies, it also permits to consider the notion of timbral similarity at finer grain of detail.
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper.
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, and 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper.
 
-\section{Experiment 1}\label{sec:}
+\section{Experiment 1}\label{sec:xp1}
 
@@
 
-\section{Experiment 2}\label{sec:}
+\section{Experiment 2}\label{sec:xp2}
 
@@
 
-\section{Analysis}\label{sec:}
+\section{Analysis}\label{sec:analysis}
 
@@
 
-We therefore now only study the properties of the groupings performed by the subject prior converting them into an overall similarity matrix.
+We therefore now only study the properties of the groupings performed by the subject prior to converting them into an overall similarity matrix.
 
-Among the 20 colors available, the subjects used on average \ml{fill}.
+Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipt.
 The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbGroup}.
-The size of the groups is on average \ml{fill}.
+The size of the groups is on average $7.7 \pm   7.2$.
 The distribution of the size of groups can be seen on Figure \ref{fig:xp2sizeGroup}.
+The large number of small groups indicates that many subjects considered that a few of the \ipt were very different from all the remaining \ipt.
 
@@
 \center
-\includegraphics[width = \textwidth]{figures/xp2example.png}
+\includegraphics[width = \textwidth]{figures/nbc.png}
 \caption{Histogram of the number of groups.}
@@
 \center
-\includegraphics[width = \textwidth]{figures/xp2example.png}
+\includegraphics[width = \textwidth]{figures/sbc.png}
 \caption{Histogram of the size of groups.}
@@
 
-\section{Model}\label{sec:}
+\section{Model}\label{sec:model}
 
-\section{Results}\label{sec:}
+\section{Results}\label{sec:results}
 
-\section{Discussion}\label{sec:}
+\section{Discussion}\label{sec:discussion}
 

### 2017-12-08 code/clusterAnalysis.m

@@ -8,8 +8,9 @@ addpath('evaluation')
 l = linkage(dc, 'weighted');
 
 dendrogram(l, 78,'Orientation','right', 'Labels', names)
-set(gca, 'xtick', [])
-saveas(gcf, ['figures/allLinkage'], 'png')
+set(gca, 'xtick', [], 'fontSize', 15)
+axis tight
+saveas(gcf, '../paper/figures/allLinkage', 'png')
 
 
 for k=2:15

### 2017-12-08 code/judgmentsDistribution.m

@@ -1,6 +1,8 @@
 
 load('judgments.mat')
 
+figPath = '../paper/figures/';
+
 nbc=[];
 sic=[];
 for k=1:size(ci, 1)
@@ -9,9 +11,20 @@ for k=1:size(ci, 1)
     sic = [sic hist(ci(k, :), nb)]; 
 end
 
-subplot 211
+disp('number of groups : ')
+disp([mean(nbc) std(nbc)])
+disp('size of groups : ')
+disp([mean(sic) std(sic)])
+
 hist(nbc, 1:20)
 axis([1 20 0 5])
-subplot 212
+xlabel('Number of groups')
+set(gca, 'fontSize', 20)
+saveas(gcf, [figPath 'nbc.png']);
+
+
 hist(sic, 1:78)
 axis([1 78 0 50])
+xlabel('Size of groups')
+set(gca, 'fontSize', 20)
+saveas(gcf, [figPath 'sbc.png']);

### 2017-12-08 code/ticelSimilarity.m

@@ -101,7 +101,9 @@ end
 for k=1:size(x, 1)
     imagesc(squareform(dc(k, :)))
     colormap gray
-    axis off
+    set(gca, 'xtick', 0)
+set(gca, 'ytick', 1:length(names))
+set(gca, 'yticklabel', names)
     saveas(gcf, ['figures/' num2str(k) 'Sim'], 'png')
 end
 

### 2017-12-08 code/clusterAnalysis.m
more code

@@ -1,6 +1,4 @@
 %function clusterAnalysis()
-
-
 modes = 0;
 load simAll
 addpath('evaluation')

### 2017-12-08 code/getMousePositionOnImage.m

@@ -1,0 +1,38 @@
+function getMousePositionOnImage(src, event)
+
+childs = get(src, 'Children');
+loc = get(childs(2), 'CurrentPoint');
+
+id = round(loc(1, 1));
+dat = guidata(src);
+dat.names{id}
+files = dir(['../sounds/' dat.names{id} '*']);
+[x, fs] = audioread(['../sounds/' files(1).name]);
+
+stop(dat.ap);
+ap = audioplayer(x, fs);
+dat.ap=ap;
+guidata(src, dat)
+playblocking(ap);
+
+% have to get the sorted list of file trough guidata
+
+
+return
+
+handles = guidata(src);
+
+cursorPoint = get(handles.axesImage, 'CurrentPoint');
+curX = cursorPoint(1,1);
+curY = cursorPoint(1,2);
+
+xLimits = get(handles.axes1, 'xlim');
+yLimits = get(handles.axes1, 'ylim');
+
+if (curX > min(xLimits) && curX < max(xLimits) && curY > min(yLimits) && curY < max(yLimits))
+disp(['Cursor coordinates are (' num2str(curX) ', ' num2str(curY) ').']);
+else
+disp('Cursor is outside bounds of image.');
+end
+
+end

### 2017-12-08 code/groupAnalysis.m

@@ -1,0 +1,85 @@
+function groupAnalysis()
+
+modes = 0;
+load clusters
+
+modeClasses = unique(modes);
+instrumentClasses = unique(instruments);
+
+cl = c(5, :);
+ncl = length(unique(cl));
+
+[scl, icl] = sort(cl);
+
+colormap(jet(15))
+
+ptx=0;
+ptl=0;
+for k=1:ncl
+    cc = length(im(cl==k));
+    ptl(k+1) = ptl(k)+cc;
+    ptx(k+1) = ptl(k)+cc/3;
+    mi = mode(im(cl==k));
+%     scl(scl==k) = mi;
+   clusterNames{k} = modeClasses(mi);
+end
+ptx(1)=[];
+
+clusterNames{end} = [clusterNames{end} 'art'];
+clusterNames{2} = [clusterNames{2} 'key'];
+
+figure(1)
+image([scl/max(scl)*15; im(icl)']);
+colorbar('Ticks', (1:length(modeClasses))+.5, 'TickLabels', modeClasses, 'TickLength', 0)
+text(ptx, ones(1, ncl), clusterNames, 'fontSize', 12);
+axis off
+set(gcf, 'WindowButtonDownFcn', @getMousePositionOnImage);
+dat.names = names(icl);
+dat.ap = audioplayer(0, 44100);
+guidata(gcf, dat);
+title('Modes')
+saveas(gcf, 'figures/groupModes', 'png')
+savefig(gcf, 'figures/groupModes')
+
+%% instruments
+
+cl = c(8, :);
+ncl = length(unique(cl));
+
+[scl, icl] = sort(cl);
+
+colormap(jet(15))
+
+ptx=0;
+ptl=0;
+clusterNames = {};
+for k=1:ncl
+    cc = length(ii(cl==k));
+    ptl(k+1) = ptl(k)+cc;
+    ptx(k+1) = ptl(k)+cc/3;
+    mi = mode(ii(cl==k));
+%     scl(scl==k) = mi;
+   clusterNames{k} = instrumentClasses(mi);
+end
+ptx(1)=[];
+pty = ones(1, ncl)-.025;
+pty(1:2:end) = pty(1:2:end)+.05;
+% clusterNames{end} = [clusterNames{end} 'art'];
+% clusterNames{2} = [clusterNames{2} 'key'];
+
+figure(2)
+image([scl/max(scl)*15; ii(icl)']);
+
+
+instrumentClasses = unique(instruments);
[…]

### 2017-12-08 code/groupAnalysis.m
cool figures

@@ -11,7 +11,7 @@ ncl = length(unique(cl));
 
 [scl, icl] = sort(cl);
 
-colormap(jet(15))
+% colormap(jet(15))
 
 ptx=0;
 ptl=0;
@@ -22,22 +22,33 @@ for k=1:ncl
     mi = mode(im(cl==k));
 %     scl(scl==k) = mi;
    clusterNames{k} = modeClasses(mi);
+   smcl(scl==k) = mi;
 end
 ptx(1)=[];
+
 
 clusterNames{end} = [clusterNames{end} 'art'];
 clusterNames{2} = [clusterNames{2} 'key'];
 
 figure(1)
-image([scl/max(scl)*15; im(icl)']);
+colormap jet
+image([smcl; im(icl)']); %scl/max(scl)*15
+l1 = line([0 79], [1.5 1.5]); l1.Color = 'k';
+for k=2:ncl   
+    l1 = line([1 1]*ptl(k)+.5, [.5 1.5]); l1.Color = 'k';
+end
 colorbar('Ticks', (1:length(modeClasses))+.5, 'TickLabels', modeClasses, 'TickLength', 0)
-text(ptx, ones(1, ncl), clusterNames, 'fontSize', 12);
-axis off
+%text(ptx, ones(1, ncl), clusterNames, 'fontSize', 12);
+% axis off
+   set(gca, 'xtick', 0)
 set(gcf, 'WindowButtonDownFcn', @getMousePositionOnImage);
 dat.names = names(icl);
 dat.ap = audioplayer(0, 44100);
 guidata(gcf, dat);
-title('Modes')
+% title('Modes')
+set(gca, 'ytick', 1:2)
+set(gca, 'yticklabel', {'Judg.', 'Ref.'});
+% ytickangle(90)
 saveas(gcf, 'figures/groupModes', 'png')
 savefig(gcf, 'figures/groupModes')
 
@@ -48,7 +59,7 @@ ncl = length(unique(cl));
 
 [scl, icl] = sort(cl);
 
-colormap(jet(15))
+% colormap(jet(15))
 
 ptx=0;
 ptl=0;
@@ -60,7 +71,9 @@ for k=1:ncl
     mi = mode(ii(cl==k));
 %     scl(scl==k) = mi;
    clusterNames{k} = instrumentClasses(mi);
+    smcl(scl==k) = mi;
 end
+smcl
 ptx(1)=[];
 pty = ones(1, ncl)-.025;
[…]

### 2017-12-08 paper/lagrangeSpontaneousSimilarity2017.tex
more analysis xp2

@@
 A partition of $n$ clusters can be constructed from the agglomerative hierarchical cluster tree by finding the smallest height at which a horizontal cut through the tree leaves $n$ clusters.
-In order to
+In order to have a better understanding of the nature of the similarity space induced by the judgments of the subjects, a cluster analysis is now performed.
 
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/allSim.png}
-\caption{Binary similarity matrix corresponding to the grouping of subject 1.}
-\label{fig:xp2bsim}
-\end{figure}
+One question that arises is whether the similarity is influenced primarily by the instrument or by the playing technique.
+To answer this question, each partition provided by cutting the dendrogram at successive levels is compared to two reference partitions.
+One is the partition of the \ipt when labeled with the instrument used (I) and the other if the partition of the \ipt when labeled with the playing technique used (Pt).
 
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/allSim.png}
-\caption{Average similarity matrix corresponding to the grouping of the complete cohort.}
-\label{fig:xp2asim}
-\end{figure}
+There are many metrics to compare two partitions by evaluating their degree of correspondance.
+The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondance analysis and the two partition are not required to have the same number of groups.
+Figure \ref{fig:clusters} shows the evolution of the NMI when comparing the partitions provided by the cluster analysis of the dendrogram produced using the human judgment of similarity and respectively the instrument partition ($j/I$) or the playing technique partition ($j/Pt$).
 
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/allLinkage.png}
-\caption{Dendrogram computed using the average similarity matrix.}
-\label{fig:xp2sizeGroup}
-\end{figure}
+There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipt.
+Also, for both curves, a global trend of increase NMI with respect to the number of clusters in the $j$ partition can be observed.
+This is a well documented fact, which simply put, means that it is easier to match a given partition with more clusters at hand.
+To compensate this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
+The resulting averaged NMI for 100 random partitions are plotted on Figure \ref{fig:clusters}.
+Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the null NMIs:
+\begin{equnarray}
+  nj/I &=& j/I - nullI  \\
+  nj/Pt &=& j/Pt - nullPt  \\
+\end{equnarray}
+This normalization is useful for identifying more easily the number of clusters for which the comparison is most profitable.
+As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference.
+For the latter, a smaller number down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
+
+\ml{change legend}
 
@@
 \includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
-\caption{Dendrogram computed using the average similarity matrix.}
-\label{fig:xp2sizeGroup}
+\caption{Cluster analysis of the similarity judgments.
+$j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
+Normalizing factors $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.}
+\label{fig:clusters}
 \end{figure}
@@
 \caption{Clustering \ipt using average perceptual similarity into 5 clusters (a) and playing technique labels (b).
-The colors of clusters displayed on top corresponds to the playing technique that is dominant in this cluster.}
-\label{fig:xp2sizeGroup}
+The color chosen to display the groups (on top) corresponds to the playing technique that is dominant in this cluster.}
+\label{fig:gm}
 \end{figure}
@@
 \caption{Clustering \ipt using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
-The colors of clusters displayed on top corresponds to the intrument that is dominant in this cluster.}
-\label{fig:xp2sizeGroup}
+The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group.}
+\label{fig:gi}
 \end{figure}

### 2017-12-11 paper/lagrangeSpontaneousSimilarity2017.tex
done with data presentation

@@
 The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondance analysis and the two partition are not required to have the same number of groups.
+It also by design bounded between 0 and 1, 1 meaning perfect match.
 Figure \ref{fig:clusters} shows the evolution of the NMI when comparing the partitions provided by the cluster analysis of the dendrogram produced using the human judgment of similarity and respectively the instrument partition ($j/I$) or the playing technique partition ($j/Pt$).
@@
 This is a well documented fact, which simply put, means that it is easier to match a given partition with more clusters at hand.
-To compensate this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
-The resulting averaged NMI for 100 random partitions are plotted on Figure \ref{fig:clusters}.
-Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the null NMIs:
+To compensate for this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
+The resulting averaged NMI for 100 randomly generated partitions are displayed on Figure \ref{fig:clusters}.
+Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the so)called null NMIs:
 \begin{equnarray}
@@
 \end{equnarray}
-This normalization is useful for identifying more easily the number of clusters for which the comparison is most profitable.
-As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference.
+This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
+As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$.
 For the latter, a smaller number down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
+Even by considering the null normalization, the matching level is much higher for the playing technique refereence.
 
-\ml{change legend}
+\ml{need the dendrogram}
+
+Once the number of clusters that lead to a partition that optimally match the reference, be it the instrument or playing technique labels are set, we can now study in more detailes the correspondance between those partitions and the reference ones.
+For this purpose, Figure \ref{fig:gi} displays two partitions.
+At the bottom is the partition that corresponds to the instruments labels, whose color code is given by the color bar on the right.
+The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
+The ordering of \ipt on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
+The partition on top is the result of the clustering with 8 clusters.
+For visualization purposes, we arbitrarily set colors to clusters of this partition to the color code of the instrument label that is most often encountered in this cluster.
+If some some occurs equally often, the instrument of the first \ipt of the cluster is chosen.
+The emerging instrument classes are: Saxophone, Clarinet, Trumpet, Flute and Cello.
+
+Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \textit{i.e.} a very diverse set of instruments labels for the \ipt they belong to.
+Also, some clusters could be merged without stong changes in the overall organisation.
+This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
+
+As can be seen on Figure \ref{fig:gm}, the correspondance between the playing technique reference and the resulting partition with 5 clusters is much higher.
+Most clusters have a high level of purity and for groups emerge : pizzicato, slap, ordinario, and ponticello.
+
+To conclude this section, the process of gathering data about the perceptual similarity of \ipt is found to be successful. The main organizationtal factor is found to be the playing technique.
 
@@
 $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
-Normalizing factors $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.}
+Normalizing factors $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
+\ml{change legend}}
 \label{fig:clusters}
@@
 \includegraphics[width = \textwidth]{figures/groupModes.png}
-\caption{Clustering \ipt using average perceptual similarity into 5 clusters (a) and playing technique labels (b).
-The color chosen to display the groups (on top) corresponds to the playing technique that is dominant in this cluster.}
+\caption{Clustering \ipt using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
+The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster.\ml{change colormap}}
 \label{fig:gm}
@@
 
-\section{Results}\label{sec:results}
+Once the perceptual data is gathered and controled, our aim is to evaluate the ability of the proposed perceptual model that is now introcued in details.
+
+\section{Validation}\label{sec:validation}
 

### 2017-12-11 paper/lagrangeSpontaneousSimilarity2017.tex
starting with model

@@
 
-Once the perceptual data is gathered and controled, our aim is to evaluate the ability of the proposed perceptual model that is now introcued in details.
+Once the perceptual data is gathered and controled, our aim is to evaluate the ability of the proposed perceptual model that is now introduced in more details.
+The model is composed of two processing stages, respectively called the primary and the secondary stage, see Figure \ref{fig:model}.
+The primary stage is designed to rely on very few meta parameters set arbitrarily by the experimenter to mimic the early stages of the mamalian auditory system.
+On contrary, the secondary stage rely largely on the data it has to model and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/groupInstruments.png}
+\caption{Computational model \ml{TODO}}
+\label{fig:model}
+\end{figure}
+
+\subsection{Primary stage}
+
+In order to model the ealry stages of the mamalian auditory system, the time scattering transform, obtained by applying a one dimensional
+wavelet transform in time to a time-frequency wavelet scalogram.
+This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important.
+
+This model is chosen over more complex alternatives for the sake of simplicity.
+Deep convolutional networks could have been considered, but the learning stage inherent in this kind of methods did not seem relevant for this study.
+On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system that do not change significantly among subjects.
+Also, more complex versions of the scattering like the joint time-frequency scattering and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are useful to model the primary stage of the auditory system \cite{dau}.
+
+\subsection{Secondary stage}
 
 \section{Validation}\label{sec:validation}
+
+\subsection{Methods}
+
+\subsection{Data}
+
+\subsection{Results}
 

### 2017-12-11 paper/lagrangeSpontaneousSimilarity2017.tex
starting with model

@@
 On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system that do not change significantly among subjects.
-Also, more complex versions of the scattering like the joint time-frequency scattering and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are useful to model the primary stage of the auditory system \cite{dau}.
+Also, more complex versions of the scattering like the joint time-frequency scattering \cite{waspaa} and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are useful to model the primary stage of the auditory system \cite{dau}.
 

### 2017-12-11 paper/lagrangeSpontaneousSimilarity2017.tex
more stuff on last sections

--- 2017-12-11 paper/lagrangeSpontaneousSimilarity2017.tex
@@
-% -----------------------------------------------
-% Template for ISMIR Papers
-% 2017 version, based on previous ISMIR templates
-
-% Requirements :
-% * 6+n page length maximum
-% * 4MB maximum file size
-% * Copyright note must appear in the bottom left corner of first page
-% * Clearer statement about citing own work in anonymized submission
-% (see conference website for additional details)
-% -----------------------------------------------
-
 \documentclass{article}
@@
 
-% Note: Please do NOT use \thanks or a \footnote in any of the author markup
-
-% Single address
-% To use with only one author or several with the same address
-% ---------------
-%\oneauthor
-% {Names should be omitted for double-blind reviewing}
-% {Affiliations should be omitted for double-blind reviewing}
-
-% Two addresses
-% --------------
-%\twoauthors
-%  {First author} {School \\ Department}
-%  {Second author} {Company \\ Address}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author}
-
-% Three addresses
-% --------------
-\author{Mathieu Lagrange, Christian EL-Hajj, Mathias Rossignol, Grégoire Lafay}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author, Third Author}
-
-% Four or more addresses
-% OR alternative format for large number of co-authors
-% ------------
-%\multauthor
-%{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
-%  $^1$ Department of Computer Science, University , Country\\
-%$^2$ International Laboratories, City, Country\\
-%$^3$  Company, Address\\
-%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
-%}
-%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}
+\author{Christian El-Hajj, Mathias Rossignol, Grégoire Lafay, Mathieu Lagrange}
 
@@
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, and 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper.
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to \ipt similarity search in the usage scenario of \ipt database browsing.
 
@@
 Deep convolutional networks could have been considered, but the learning stage inherent in this kind of methods did not seem relevant for this study.
-On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system that do not change significantly among subjects.
-Also, more complex versions of the scattering like the joint time-frequency scattering and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are useful to model the primary stage of the auditory system \cite{dau}.
+On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system.
+Those parameters do not change significantly among subjects.
+Also, more complex versions of the scattering like the joint time-frequency scattering \cite{waspaa} and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are mandatory to model the primary stage of the auditory system \cite{dau}.
 
 \subsection{Secondary stage}
+
+The secondary stage aims at providing a metric space that closely match the one resulting from the data gathering by experiment 2.
+For the reason, it involves a learning based process block that weight the features outputed by the primary stage.
+
+Indeed, if the properties of the first processing stages of the human auditory system can reasonably be considered as similar for every subjects despite potential hearing losses for some, the later ones is expected to vary more drastically due to his education, tastes and experiences with stimuli similar to the \ipt that are considered in this study.
 
@@
 
+For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}.
+For convenience, the features (scattering and mfccs) are alos provided, but the software can be used to process from scratch the SOL database.
+
 \section{Discussion}\label{sec:discussion}
+
+Thanks to the digitization capabilities available, gathering large amount of audio recordings of \ipt is achievable and several databases are now available on the market.
+Browsing of such databases can be achieved by keyword based search.
+It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the user has a precise understanding of this ontology, and 3) that the user a precise idea of the kind of timbre he is looking to find the \ipt he is searching for.
+
+An interesting alternative type of search is the similarity content based search.
+Being solely based on sound, this kind of search can be used when querying by example.
+The user provides an example of the target sound to find \ipt that match the timbre description of the query, be it an onomatepea, a tone from another instrument or any kind of environmental sound.
+
+The computational model proposed in this paper is well suited for this kind of search, as it can model complex perceptual timbre spaces and can scale to large databases.
+Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{lsh}.
+The secondary stage being learnt, it can easily adapt to user tastes and usage history by reweighting the scattering features for improved search relevance.
 

### 2017-12-12 code/clusterAnalysis.m
dendro

@@ -5,8 +5,10 @@ addpath('evaluation')
 
 l = linkage(dc, 'weighted');
 
+% names{2} = [names{2} '     .'];
+
 dendrogram(l, 78,'Orientation','right', 'Labels', names)
-set(gca, 'xtick', [], 'fontSize', 15)
+set(gca, 'xtick', [], 'fontSize', 9)
 axis tight
 saveas(gcf, '../paper/figures/allLinkage', 'png')
 

### 2017-12-12 code/groupAnalysis.m

@@ -31,7 +31,8 @@ clusterNames{end} = [clusterNames{end} 'art'];
 clusterNames{2} = [clusterNames{2} 'key'];
 
 figure(1)
-colormap jet
+colormap(lines)
+% colormap jet
 image([smcl; im(icl)']); %scl/max(scl)*15
 l1 = line([0 79], [1.5 1.5]); l1.Color = 'k';
 for k=2:ncl   
@@ -81,7 +82,7 @@ pty(1:2:end) = pty(1:2:end)+.05;
 % clusterNames{2} = [clusterNames{2} 'key'];
 
 figure(2)
-colormap parula
+colormap lines
 image([smcl; ii(icl)']);
 for k=2:ncl   
     l1 = line([1 1]*ptl(k)+.5, [.5 1.5]); l1.Color = 'k';

### 2017-12-12 paper/lagrangeSpontaneousSimilarity2017.tex
work in progress

@@
 
-Musical timbre is an interesting notion to study as it is perceptively defined.
-Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse -- but somewhat more controled than other stimuli such as speech or environemntal sounds -- set of physical stimuli.
+Musical timbre is an interesting notion to study as it is perceptively defined \cite{grey1977multidimensional}.
+Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse set of physical stimuli.
+However this diversity can be more easily more controlled than other stimuli such as speech or environmental sounds.
+
+The focus of this paper is to study the influence of the playing technique on the perception of timbre.
 
 A popular view of the kind of stimuli we consider is that they can be described by a spread of energy that evolve through time and across frequency.
-By following this trend, most researchers translate the negation of the ASA definition of timbre into a more relative definition : musical timbre is about the distribution of intensity across the time / frequency plane.
+By following this trend, most researchers translate the exclusion definition of timbre \cite{marozeau2003dependency} into a more specified definition.
+The exclusion definition states that timbre is "the fourth component of sound quality, the first three being pitch, loudness and duration".
+In an attempt to specify more this notion of timbre, we will loosely consider in this paper that musical timbre is the influence on the human auditory system of the distribution of intensity across the time / frequency plane.
 
-A great deal of litterature focuses on the instaneous spectrum, \textit{i.e.} the timbre can be defined without considering evolution of the energy distribution trough time.
-Other studies demonstrated the importance of the onset in the recognition of some musical instruments.
-Those studies on a rather short term model where the waveform is observed in a range that do not exceed 100 ms.
+A great deal of litterature focuses on the instantaneous spectrum \cite{grey1978perceptual}.
+Other studies also demonstrated the importance of the onset in the recognition of some musical instruments.
+All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common porcessing scenario that is very commonly used in computational recognition models \cite{tzanetakis2002musical}.
+
+Agus \& al effectively showed that a great deal of instrument recognition can be achieved with very little listening time \cite{agus2012fast}.
+Though, instrument recognition is only part of musical timbre perception.
+We propose in this paper to study the influence of the playing technique on musical timbre percept.
+Evidently, the influence of the playing technique can be measured in terms of energy distribution along the frequency axis but also with respect to time on long time scales.
 
 Interestingly, modern physiological models of the primary mamalian auditory system also considers the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
-Dau's model acount for frequency and rate of modulation.
-Shamma's model account for frequency, rate and bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret.
-Though, most of the numerical experiments done using this model did not fully demonstrated the usefullness of this third dimension.
+Dau's model acounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
+Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
+Though, some numerical applications of this model did not demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
 
-Thus, by taking a signal processing view of the matter, one could consider that a first stage of the auditory system proceeds to a rather fixed set of convolutions across time and frequency, projecting the signal into a large set of descriptive features organized across frequency and rate of modulations.
+Thus, by taking a signal processing view of the matter, one could consider that a first stage of the auditory system proceeds to a rather fixed set of convolutions across time and frequency, projecting the signal into a large set of descriptive features organized across frequency and rate of modulations \cite{anden2014deep}.
 
-What the higher level of the auditory cortex does remains largely unkown, but we can assume that the level of plasticity is very high, as any living entity has to adapt their percepts to a wide diversity of tasks.
-Considering the lack of information about this stage, we will assume in this paper that this second step is rather opportunistic, in the sense that, from this large set of features, it aims at multiplexing in some way the ones that are the most relevant for the task at hand.
+What the higher levels of the auditory cortex does remains largely unkown, but we can assume that the degree of freedom of the underlying model is very large contrary to the early stages.
+We will assume in this paper that this second step is rather opportunistic.
+From the large set of features it recieves, this second stage aims at multiplexing in some way the ones that are the most relevant for the task at hand.
+Related work includes the use of the auditory model proposed by Prof.
+Shamma for studying musical timbre \cite{patil2012music}.
 
 Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts.
-A notion that have not been extensively studied in the litterature.
-Compared to instrument class, the reference that is usually taken in timbre studies, it also permits to consider the notion of timbral similarity at finer grain of detail.
+A notion that have not been extensively studied in the litterature \cite{burred2010dynamic}.
+Compared the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, considering the playing technique allows us to consider the notion of timbral similarity at finer grain of detail.
 
@@
 
-The dataset considered in this study has audio recording of 16 different  musical instruments () played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+The dataset considered in this study has audio recording of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 samples.
@@
 Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the so)called null NMIs:
-\begin{equnarray}
+\begin{eqnarray}
   nj/I &=& j/I - nullI  \\
   nj/Pt &=& j/Pt - nullPt  \\
-\end{equnarray}
+\end{eqnarray}
 This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
@@
 
-\section{Dataset}
+\section{Dataset} \label{sec:dataset}
 
-distribution of instruments and modes
+The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to 25444 audio samples.
 
-\subsection{Naming}
+The recorded instruments belongs to the wind and string classes.
+For winds, the instruments are: accordion (Acc), alto, tuba (Tb), bassoon, clarinet (Cl), flute (Ft), horn (Hn), oboe (Ob), trombone, and trumpet (Tp).
+For the string class, the instruments are: guitar (Gt), harp, viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
 
-abbreviations 4 instruments and modes.
+Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+
+\section{Dendrogram}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/allLinkage.png}
+\caption{Dendrogram achiebed using the agglomerative clustering algorithm.}
+\label{fig:dendrogram}
+\end{figure}
 

### 2017-12-12 code/clusterAnalysis.m
new dendrogram

@@ -6,11 +6,16 @@ addpath('evaluation')
 l = linkage(dc, 'weighted');
 
 % names{2} = [names{2} '     .'];
-
+h=figure(1);
 dendrogram(l, 78,'Orientation','right', 'Labels', names)
 set(gca, 'xtick', [], 'fontSize', 9)
-axis tight
-saveas(gcf, '../paper/figures/allLinkage', 'png')
+% axis tight
+% saveas(gcf, '../paper/figures/allLinkage', 'png')
+
+set(h,'Units','Inches');
+pos = get(h,'Position');
+set(h,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
+saveas(gcf, '../paper/figures/dendrogram', 'pdf')
 
 
 for k=2:15

### 2017-12-13 paper/lagrangeSpontaneousSimilarity2017.tex
new dendrogram
more refs

@@
 Other studies also demonstrated the importance of the onset in the recognition of some musical instruments.
-All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common porcessing scenario that is very commonly used in computational recognition models \cite{tzanetakis2002musical}.
+All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common porcessing scenario that is very commonly used in computational recognition models of timbre \cite{tzanetakis2002musical}.
 
@@
 We propose in this paper to study the influence of the playing technique on musical timbre percept.
-Evidently, the influence of the playing technique can be measured in terms of energy distribution along the frequency axis but also with respect to time on long time scales.
+Evidently, the influence of the playing technique can be measured in terms of energy distribution along the frequency axis but also with respect to time on longer time scales than usually considered.
 
-Interestingly, modern physiological models of the primary mamalian auditory system also considers the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
+Interestingly, modern physiological models of the primary mamalian auditory system also finds that the evolution of the energy distribution at longer time scales matters, \textit{i.e.} how the energy modulates through time for given frequencies.
 Dau's model acounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
 Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
-Though, some numerical applications of this model did not demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
+Though, the use of numerical implementations of this model did not fully demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
 
@@
 A notion that have not been extensively studied in the litterature \cite{burred2010dynamic}.
-Compared the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, considering the playing technique allows us to consider the notion of timbral similarity at finer grain of detail.
+Compared the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at finer grain of detail.
 
@@
 
-The dataset considered in this study has audio recording of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+The dataset considered in this study is taken from the studio-on-line music library  \cite{peeters2000instrument}.
+It has audio recordings of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 samples.
@@
 
-Rating examples are given :
+Rating guidelines are provided :
 \begin{itemize}
@@
 
-Two musical composition professors of the CNSMDP, a renowned French school of music performed the test. All the couples with a grade higher than three are selected, bringing 78 \ipt to study.
+Two music composition professors of the CNSMDP, a renowned French school of music, performed the test. All the couples with a grade higher than three are selected, bringing 78 \ipt to study.
 
@@
 
-The second experiment aims at gathering spontaneous judgment of similarity.
-To do so, we chose a canonical free sorting task  for several reasons.
-First, it gives the ability to the subject to listen to the whole set of \ipt before performing the task.
-Second, it gives more freedom to the subject thus potentially more interest to perform the task than a lengthy series of triplet dissimilarity ratings.
+The second experiment aims at gathering spontaneous judgment of similarity for the 78 \ipt selected in the previous experiment.
+To do so, we chose a canonical free sorting task for two main reasons.
+First, this kind of task gives the ability to the subject to listen to the whole set of \ipt before performing the task.
+Second, it gives more freedom to the subject, thus potentially more interest to perform the task than a lengthy series of triplet dissimilarity ratings.
 
 The subjects are thus asked to organize the 78 \ipt displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
-The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to \ml{use its own representation of similarity}.
+The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to follow its own judgment.
 
@@
 
-As discussed in the previous section, only the grouping given by the subjects using the color labels is considered to predict the spontaneous judgments of similarity among \ipt.
-The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the similarity space is two dimensional.
+As discussed in the previous section, only the grouping given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
 
@@
 A grouping can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
-The Figure \ref{fig:xp2bsim} show the binary similarity matrix that correponds to the grouping displayed on Figure \ref{fig:xp2display}.
-By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that correponds to our measure of the spontaneous similarity between \ipt.
+By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that correponds to our measure of the spontaneous perceptual similarity between \ipt.
 
-A cluster analysis is next performed using an agglomerative hierarchical clustering using the weighted average algorithm for computing distance between clusters.
-The resulting dendrogram is displayed on Figure \ref{fig:xp2sizeGroup}.
-
+In order to gather information about the properties of the data, a cluster analysis is next performed using an agglomerative hierarchical clustering \cite{gordon1987review} using the weighted average algorithm for computing distance between clusters.
+The resulting dendrogram is displayed on Figure \ref{fig:dendrogram}.
 A partition of $n$ clusters can be constructed from the agglomerative hierarchical cluster tree by finding the smallest height at which a horizontal cut through the tree leaves $n$ clusters.
-In order to have a better understanding of the nature of the similarity space induced by the judgments of the subjects, a cluster analysis is now performed.
 
@@
 To answer this question, each partition provided by cutting the dendrogram at successive levels is compared to two reference partitions.
-One is the partition of the \ipt when labeled with the instrument used (I) and the other if the partition of the \ipt when labeled with the playing technique used (Pt).
+One reference is the partition of the \ipt when labeled with the instrument used (I) and the other if the partition of the \ipt when labeled with the playing technique used (Pt).
 
-There are many metrics to compare two partitions by evaluating their degree of correspondance.
+There are many metrics to compare two partitions by evaluating their degree of correspondance \cite{wagner2007comparing}.
 The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondance analysis and the two partition are not required to have the same number of groups.
 It also by design bounded between 0 and 1, 1 meaning perfect match.
+We consider the normalization proposed in \cite{strehl2002cluster}.
 Figure \ref{fig:clusters} shows the evolution of the NMI when comparing the partitions provided by the cluster analysis of the dendrogram produced using the human judgment of similarity and respectively the instrument partition ($j/I$) or the playing technique partition ($j/Pt$).
-
-There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipt.
-Also, for both curves, a global trend of increase NMI with respect to the number of clusters in the $j$ partition can be observed.
-This is a well documented fact, which simply put, means that it is easier to match a given partition with more clusters at hand.
-To compensate for this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
-The resulting averaged NMI for 100 randomly generated partitions are displayed on Figure \ref{fig:clusters}.
-Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the so)called null NMIs:
-\begin{eqnarray}
-  nj/I &=& j/I - nullI  \\
-  nj/Pt &=& j/Pt - nullPt  \\
-\end{eqnarray}
-This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
-As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$.
-For the latter, a smaller number down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
-Even by considering the null normalization, the matching level is much higher for the playing technique refereence.
-
-\ml{need the dendrogram}
-
-Once the number of clusters that lead to a partition that optimally match the reference, be it the instrument or playing technique labels are set, we can now study in more detailes the correspondance between those partitions and the reference ones.
-For this purpose, Figure \ref{fig:gi} displays two partitions.
-At the bottom is the partition that corresponds to the instruments labels, whose color code is given by the color bar on the right.
-The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
-The ordering of \ipt on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
-The partition on top is the result of the clustering with 8 clusters.
-For visualization purposes, we arbitrarily set colors to clusters of this partition to the color code of the instrument label that is most often encountered in this cluster.
-If some some occurs equally often, the instrument of the first \ipt of the cluster is chosen.
-The emerging instrument classes are: Saxophone, Clarinet, Trumpet, Flute and Cello.
-
-Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \textit{i.e.} a very diverse set of instruments labels for the \ipt they belong to.
-Also, some clusters could be merged without stong changes in the overall organisation.
-This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
-
-As can be seen on Figure \ref{fig:gm}, the correspondance between the playing technique reference and the resulting partition with 5 clusters is much higher.
-Most clusters have a high level of purity and for groups emerge : pizzicato, slap, ordinario, and ponticello.
-
-To conclude this section, the process of gathering data about the perceptual similarity of \ipt is found to be successful. The main organizationtal factor is found to be the playing technique.
 
@@
 \end{figure}
+
+There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipt.
+Also, for both curves, a global trend of increase with respect to the number of clusters in the $j$ partition can be observed.
+This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}.
+To compensate for this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
+The resulting averaged NMI for 100 randomly generated partitions are displayed on Figure \ref{fig:clusters}.
+Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the so-called null NMIs:
+\begin{eqnarray}
+  nj/I &=& j/I - nullI  \\
+  nj/Pt &=& j/Pt - nullPt  \\
+\end{eqnarray}
+This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
+As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$.
+For the latter, reducing the number of clusters down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
+Even by considering the null normalization, the matching level is much higher for the playing technique reference.
+
+The correspondances between those partitions and the reference ones are now studied.
+For this purpose, Figure \ref{fig:gi} displays two partitions.
+At the bottom is the partition that corresponds to the instruments labels, whose color code is given by the color bar on the right.
+The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
+The ordering of \ipt on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
+The partition on top is the result of the clustering with 8 clusters.
+For visualization purposes, we arbitrarily set colors to clusters of this partition to the color code of the instrument label that is most often encountered in this cluster.
+If some labels occur equally often, the instrument of the first \ipt of the cluster is chosen.
+The emerging instrument classes are: Saxophone, Clarinet, Trumpet, Flute and Cello.
+
+Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \textit{i.e.} a very diverse set of instruments labels for the \ipt they belong to.
+Also, some clusters could be merged without stong changes in the overall organisation.
+This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
+
+As can be seen on Figure \ref{fig:gm}, the correspondance between the playing technique reference and the resulting partition with 5 clusters is much higher.
+Most clusters have a high level of purity and for groups emerge : Pizzicato, Slap, Ordinario, and Ponticello.
+
+To conclude this section, the process of gathering data about the perceptual similarity of \ipt is found to be successful as the subjects used their degree of freedom in a consistent manner and no inconsistency in the data has been detected during the analysis process.
+The main result is that the playing technique is an important factor when modeling the spontaneous judgment of similarity gathered in the second experiment.
 
@@
 
-In order to model the ealry stages of the mamalian auditory system, the time scattering transform, obtained by applying a one dimensional
+In order to model the early stages of the mamalian auditory system, the time scattering transform \cite{anden2014deep}, obtained by applying a one dimensional
 wavelet transform in time to a time-frequency wavelet scalogram.
-This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important.
+This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important \cite{dau1997modeling}.
 
 This model is chosen over more complex alternatives for the sake of simplicity.
-Deep convolutional networks could have been considered, but the learning stage inherent in this kind of methods did not seem relevant for this study.
-On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system.
-Those parameters do not change significantly among subjects.
-Also, more complex versions of the scattering like the joint time-frequency scattering \cite{waspaa} and the spiral scattering are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are mandatory to model the primary stage of the auditory system \cite{dau}.
+Deep convolutional networks could have been considered \cite{lee2009unsupervised}, but the learning stage inherent in this kind of methods did not seem relevant for this study.
+On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system as those parameters are not expected to change significantly among subjects.
+Also, more complex versions of the scattering like the joint time-frequency scattering \cite{anden2015joint} and the spiral scattering \cite{lostanlen2016wavelet} are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are mandatory to model the primary stage of the auditory system.
 
@@
 The computational model proposed in this paper is well suited for this kind of search, as it can model complex perceptual timbre spaces and can scale to large databases.
-Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{lsh}.
+Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{pauleve2010locality}.
 The secondary stage being learnt, it can easily adapt to user tastes and usage history by reweighting the scattering features for improved search relevance.
 
-% For bibtex users:
+\bibliographystyle{alpha}
 \bibliography{bib}
@@
 The recorded instruments belongs to the wind and string classes.
-For winds, the instruments are: accordion (Acc), alto, tuba (Tb), bassoon, clarinet (Cl), flute (Ft), horn (Hn), oboe (Ob), trombone, and trumpet (Tp).
-For the string class, the instruments are: guitar (Gt), harp, viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+For winds, the instruments are: accordion (Acc), saxophone alto (aSax), tuba (Tb), bassoon (Bn), clarinet (Cl), flute (Ft), horn (Hn), oboe (Ob), trombone, and trumpet (Tp).
+For the string class, the instruments are: guitar (Gt), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+
+Some instruments can be complemented with sordina (S), wha (W). The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina
 
@@
 \center
-\includegraphics[width = \textwidth]{figures/allLinkage.png}
-\caption{Dendrogram achiebed using the agglomerative clustering algorithm.}
+%\includegraphics[width = \textwidth]{figures/dendrogram.pdf}
+\caption{Dendrogram obtained after processing the similarity judgment matrix using the agglomerative clustering algorithm.}
 \label{fig:dendrogram}

### 2017-12-13 paper/lagrangeSpontaneousSimilarity2017.tex
naming pts

@@
 The recorded instruments belongs to the wind and string classes.
-For winds, the instruments are: accordion (Acc), saxophone alto (aSax), tuba (Tb), bassoon (Bn), clarinet (Cl), flute (Ft), horn (Hn), oboe (Ob), trombone, and trumpet (Tp).
-For the string class, the instruments are: guitar (Gt), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
 
-Some instruments can be complemented with sordina (S), wha (W). The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina
+Some instruments can be complemented with sordina (S), wha (W).
+The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
 
 Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+For  the second experiment described in this paper, the playing techniques considered are: artificial-harmonic (art), blow (blow), exploding slap  (explo), harmonic (harm), key-click (key), col legno   (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
 
@@
 
-% For non bibtex users:
-%\begin{thebibliography}{citations}
-%
-%\bibitem {Author:00}
-%E. Author.
-%``The Title of the Conference Paper,''
-%{\it Proceedings of the International Symposium
-%on Music Information Retrieval}, pp.~000--111, 2000.
-%
-%\bibitem{Someone:10}
-%A. Someone, B. Someone, and C. Someone.
-%``The Title of the Journal Paper,''
-%{\it Journal of New Music Research},
-%Vol.~A, No.~B, pp.~111--222, 2010.
-%
-%\bibitem{Someone:04} X. Someone and Y. Someone. {\it Title of the Book},
-%    Editorial Acme, Porto, 2012.
-%
-%\end{thebibliography}
-
 \end{document}

### 2017-12-14 paper/lagrangeSpontaneousSimilarity2017.tex
better figures

@@
 \caption{Clustering \ipt using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
-The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster.\ml{change colormap}}
+The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.}
 \label{fig:gm}
@@
 \caption{Clustering \ipt using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
-The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group.}
+The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.}
 \label{fig:gi}
@@
 Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
-For  the second experiment described in this paper, the playing techniques considered are: artificial-harmonic (art), blow (blow), exploding slap  (explo), harmonic (harm), key-click (key), col legno   (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+For  the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap  (explo), harmonic (harm), key-click (key), col legno   (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
 
@@
 \center
-%\includegraphics[width = \textwidth]{figures/dendrogram.pdf}
+\includegraphics[width = \textwidth]{figures/dendrogram.pdf}
 \caption{Dendrogram obtained after processing the similarity judgment matrix using the agglomerative clustering algorithm.}

### 2017-12-14 code/groupAnalysis.m

@@ -30,28 +30,35 @@ ptx(1)=[];
 clusterNames{end} = [clusterNames{end} 'art'];
 clusterNames{2} = [clusterNames{2} 'key'];
 
+vec = [smcl; im(icl)'];
+
 figure(1)
-colormap(lines)
+% cm = colormap(parula);
+cm = jet(max(vec(:)));
+cm = cm(randperm(length(cm)), :);
+colormap(cm);
+
 % colormap jet
-image([smcl; im(icl)']); %scl/max(scl)*15
+image(vec); %scl/max(scl)*15
 l1 = line([0 79], [1.5 1.5]); l1.Color = 'k';
 for k=2:ncl   
     l1 = line([1 1]*ptl(k)+.5, [.5 1.5]); l1.Color = 'k';
 end
 colorbar('Ticks', (1:length(modeClasses))+.5, 'TickLabels', modeClasses, 'TickLength', 0)
+[clusterNames{:}]
 %text(ptx, ones(1, ncl), clusterNames, 'fontSize', 12);
 % axis off
    set(gca, 'xtick', 0)
-set(gcf, 'WindowButtonDownFcn', @getMousePositionOnImage);
-dat.names = names(icl);
-dat.ap = audioplayer(0, 44100);
-guidata(gcf, dat);
+% set(gcf, 'WindowButtonDownFcn', @getMousePositionOnImage);
+% dat.names = names(icl);
+% dat.ap = audioplayer(0, 44100);
+% guidata(gcf, dat);
 % title('Modes')
 set(gca, 'ytick', 1:2)
 set(gca, 'yticklabel', {'Judg.', 'Ref.'});
 % ytickangle(90)
-saveas(gcf, 'figures/groupModes', 'png')
-savefig(gcf, 'figures/groupModes')
+saveas(gcf, '../paper/figures/groupModes', 'png')
+savefig(gcf, '../paper/figures/groupModes')
 
 %% instruments
 
@@ -74,7 +81,7 @@ for k=1:ncl
    clusterNames{k} = instrumentClasses(mi);
     smcl(scl==k) = mi;
 end
-smcl
+% smcl
 ptx(1)=[];
 pty = ones(1, ncl)-.025;
 pty(1:2:end) = pty(1:2:end)+.05;
@@ -82,24 +89,28 @@ pty(1:2:end) = pty(1:2:end)+.05;
 % clusterNames{2} = [clusterNames{2} 'key'];
 
 figure(2)
-colormap lines
[…]

### 2017-12-14 experiments/timbralSimilaritySol/report/tex/exposeTmp.tex
addition of experiment

\begin{frame}\frametitle{Factors flow graph}

\begin{center}

\begin{figure}

\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{../figures/factors.pdf}

\label{factorFlowGraph}

\end{figure}

\end{center}

\end{frame}

\begin{frame}\frametitle{sct: 1000, split: none, reference: judgments, randomize: 0, expand: 0, cut: 1, median: 1, compress: 1, standardize: 1}

\begin{table} 
\begin{center} 
\ 
 \setlength{\tabcolsep}{.16667em} 
\begin{tabular}{llllc} 
features & projection & averageJudgment & separateJudgment & p (\%) \\ 
\hline 
mfcc & lmnn & 0 & 0 &  86.31 $\pm$5.91 \\ 
mfcc & lmnn & 0 & 1 &  86.18 $\pm$6.05 \\ 
mfcc & lmnn & 1 & 0 &  86.22 $\pm$5.92 \\ 
mfcc & none & 1 &  &  85.07 $\pm$6.19 \\ 
mfcc & lda & 1 &  &  81.50 $\pm$7.65 \\ 
scat & lmnn & 0 & 0 &  93.31 $\pm$3.92 \\ 
scat & lmnn & 0 & 1 & \textbf{\textcolor{red}{ 98.09 $\pm$1.28}} \\ 
scat & lmnn & 1 & 0 &  94.80 $\pm$3.26 \\ 
scat & none & 1 &  &  87.01 $\pm$5.81 \\ 
scat & lda & 1 &  & 80.95 $\pm$10.37 \\ 
\end{tabular} 
\end{center} 
\label{sc1000SpnoRejuRa0Ex0Cu1Me1Co1St1} 
\end{table}

\end{frame}

\begin{frame}\frametitle{\small split: none, reference: judgments, randomize: 0, expand: 0, cut: 1, median: 1, compress: 1, standardize: 1} 
\begin{center} 
\begin{figure} 
\centering 
\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{./figures/Fig160.pdf} 
\label{spnoRejuRa0Ex0Cu1Me1Co1St1} 
\end{figure} 
\end{center}

\end{frame}

### 2017-12-14 paper/lagrangeSpontaneousSimilarity2017.tex
addition of experiment

@@
 
-Indeed, if the properties of the first processing stages of the human auditory system can reasonably be considered as similar for every subjects despite potential hearing losses for some, the later ones is expected to vary more drastically due to his education, tastes and experiences with stimuli similar to the \ipt that are considered in this study.
+Indeed, if the properties of the first processing stages of the human auditory system can reasonably be considered as similar for every subjects despite potential hearing losses for some, the later ones is expected to vary more drastically due to their respective education, tastes and experiences with stimuli similar to the \ipt that are considered in this study.
+We therefore consider a supervised setting in order to optimize the representational space provided by the scattering transform.
+
+To do so, several
 

### 2017-12-14 experiments/timbralSimilaritySol/clusterEnsemble/ClusterBasics-V2.0/kmeans.m

@@ -1,0 +1,92 @@
+% function [centres, options, post, errlog] = kmeans(...)
+%
+% This function has been adopted from a k means implementation
+% copyright (c) Christopher M Bishop, Ian T Nabney (1996, 1997)
+
+function [centres, options, post, errlog] = kmeans(centres, data, options, measure)
+
+[ndata, data_dim] = size(data);
+[ncentres, dim] = size(centres);
+
+if dim ~= data_dim
+  error('Data dimension does not match dimension of centres')
+end
+
+if (ncentres > ndata)
+  error('More centres than data')
+end
+
+% Sort out the options
+if (options(14))
+  niters = options(14);
+else
+  % niters = 25; % AS was this for a long time
+  niters = 50;
+  niters = 100; % just for ml paper examples
+end
+
+store = 0;
+if (nargout > 3)
+  store = 1;
+  errlog = zeros(1, niters);
+end
+
+% Check if centres and posteriors need to be initialised from data
+if (options(5) == 1)
+  % Do the initialisation
+  perm = randperm(ndata);
+  perm = perm(1:ncentres);
+
+  % Assign first ncentres (permuted) data points as centres
+  centres = data(perm, :);
+end
+% Matrix to make unit vectors easy to construct
+id = eye(ncentres);
+
+% Main loop of algorithm
+for n = 1:niters
+
+  % Save old centres to check for termination
+  old_centres = centres;
+  
+  % Calculate posteriors based on existing centres
+  d2 = sqrt(-log((feval(measure,data,centres))));
+  % Assign each point to nearest centre
+  [minvals, index] = min(d2', [], 1);
+  post = id(index,:);
+
+  num_points = sum(post, 1);
+  % Adjust the centres based on new posteriors
+  for j = 1:ncentres
+    if (num_points(j) > 0)
+      centres(j,:) = sum(data(find(post(:,j)),:), 1)/num_points(j);
+    end
+  end
+
+  % Error value is total squared distance from cluster centres
+  e = sum(minvals);
+  if store
+    errlog(n) = e;
+  end
+  if options(1) > 0
+    fprintf(1, 'Cycle %4d  Error %11.6f\n', n, e);
+  end
+
+  if n > 1
+    % Test for termination
+    if max(max(abs(centres - old_centres))) < options(2) & ...
+        abs(old_e - e) < options(3)
+      options(8) = e;
+      return;
+    end
+  end
+  old_e = e;
+end
+
[…]

### 2017-12-14 experiments/timbralSimilaritySol/clusterEnsemble/ClusterBasics-V2.0/sgraph.m

@@ -1,0 +1,55 @@
+% function labels = sgraph(k,dataname)
+%
+% copyright (c) 1998-2011 by Alexander Strehl
+
+function labels = sgraph(k,dataname) 
+
+scriptfile = ['' 'partgraph' num2str(sum(dataname=='0')) num2str(sum(dataname=='1')) num2str(sum(dataname=='2')) num2str(sum(dataname=='3')) '.bat'];
+
+if ~exist('dataname'),
+      dataname = ['' 'graph0'];
+end;
+resultname = [dataname '.part.' num2str(k)];
+
+lastchar = str2num(dataname(length(dataname)));
+if (isempty(lastchar)),
+  disp('sgraph: file does not comply to name convention');
+  lastchar = 0;
+end;
+s='';
+if (lastchar<2),
+   scriptfile = ['pmetis ' dataname ' ' num2str(k)];
+else
+   ubfactor = 5;
+   scriptfile = ['shmetis ' dataname ' ' num2str(k) ' ' num2str(ubfactor)];
+end;
+
+system(scriptfile);
+
+fid = fopen(resultname,'r');
+if (fid == -1),
+  disp('sgraph: partitioning not successful due to external error');
+  fid = fopen(dataname);
+  if (fid == -1),
+    disp('sgraph: graph file is not accessible');
+  else
+    if lastchar>=2,
+      junk = fscanf(fid,'%d',1); 
+    end;
+    labels = ones(1,fscanf(fid,'%d',1));
+    if isempty(labels),
+      disp('sgraph: empty label vector - suspecting file system full');
+    end;
+    fclose(fid);
+  end;
+else
+  disp(['sgraph: ' scriptfile ' completed - loading ' resultname]);
+  labels = (fscanf(fid,'%d')+1)';
+  fclose(fid);
+end;
+
+fid = fopen(resultname,'r');
+if (fid ~= -1),
+  fclose(fid);
+  delete(resultname);
+end;

### 2017-12-14 experiments/timbralSimilaritySol/clusterEnsemble/ClusterBasics-V2.0/simxjac.m

@@ -1,0 +1,28 @@
+% function s = simxjac(a,b)
+%
+% DESCRIPTION
+%   computes extended Jaccard similarity between row objects in matrices a and b
+%
+% Copyright (c) 1998-2011 by Alexander Strehl
+
+function s = simxjac(a,b)
+
+if ~exist('b')
+  b = a;
+end;
+
+n = size(a,1);
+m = size(b,1);
+d = size(a,2);
+
+if (d~=size(b,2))
+  disp('simxjac: data dimensions do not match');
+  return;
+end;
+
+temp = a *1* b'; 
+asquared = sum((a.^2),2); 
+bsquared = sum((b.^2),2); 
+s = temp ./ ((asquared * ones(1,m)) + (ones(n,1) * bsquared') - temp);
+
+

### 2017-12-14 experiments/timbralSimilaritySol/clusterEnsemble/ClusterBasics-V2.0/wgraph.m

@@ -1,0 +1,88 @@
+% function dataname = wgraph(e,w,method,dataname)
+%
+% copyright (c) 1998-2011 by Alexander Strehl
+
+function dataname = wgraph(e,w,method,dataname)
+
+if ~exist('method')
+    method = 0;
+end;
+
+if ~exist('dataname'),
+    dataname = ['' 'graph'];
+end;
+dataname = [dataname num2str(method)];
+
+if ((method == 0)|(method == 1)),
+    e=e-diag(diag(e));
+end;
+e = ints(e);
+if ((method==1)|(method==3)),
+    w = ints(w);
+end;
+
+while exist(dataname,'file')
+    dataname = [dataname num2str(method)];
+end;
+
+fid = fopen(dataname,'w');
+
+if (fid~=-1)
+    
+    disp(['wgraph: writing ' dataname ]);
+    
+    if (method == 0),
+        fprintf(fid,'%d %d 1\n',[size(e,1) full(sum(sum(e>0)))/2]);
+    else
+        if (method == 1),
+            fprintf(fid,'%d %d 11\n',[size(e,1) full(sum(sum(e>0)))/2]);
+        else
+            validcolumns = find((sum(e,1)>0));
+            if (length(validcolumns)~=size(e,2)),
+                disp('wgraph: removing empty feature columns');
+                e = e(:,validcolumns);
+            end;
+            fprintf(fid,'%d %d 1\n',[size(e,2) size(e,1)]);
+        end;
+    end;
+    
+    if ((method == 0)|(method == 1)),
+        for i=1:size(e,1),
+            edges = find(e(i,:)>0);
+            weights = e(i,edges);
+            if method == 0,
+                interlaced = zeros(1,2*length(edges));
+                interlaced(1:2:2*length(edges)-1) = edges;
+                interlaced(2:2:2*length(edges)) = weights;
+            else
+                interlaced = zeros(1,1+2*length(edges));
+                interlaced(1) = w(i);
+                interlaced(2:2:2*length(edges)-1+1) = edges;
+                interlaced(3:2:2*length(edges)+1) = weights;
+            end;
+            fprintf(fid,'%d ',interlaced);
+            fprintf(fid,'\n');
+        end;
+    else
+        disp(['wgraph: ' num2str(size(e,1)) ' vertices and ' num2str(size(e,2)) ' non-zero hyperedges']);
+        for i=1:size(e,2),
+            edges = find(e(:,i)>0)';
+            if method==2
[…]

### 2017-12-14 experiments/timbralSimilaritySol/clusterEnsemble/combineJudgments.m

@@ -1,0 +1,27 @@
+load ticelJudgments
+addpath(genpath('.'))
+
+ensemble = clusterensemble(ci);
+
+% generate reference algorithmically
+reference = 20;
+
+for kclus=1:size(ci, 1)
+    [accuracyv, classMatching] = accuracy(ci(reference,:), ci(kclus,:));
+    
+    labeltemp=ci(kclus,:);
+    labeltemp2=ci(kclus,:);
+    for kcm=1:size(classMatching,1)
+        n=classMatching(kcm,:);
+        vcm=find(n);
+        if(~isempty(vcm))
+            
+            labeltemp(labeltemp2==kcm)=vcm;
+        end
+    end
+    ci2(kclus,:)=labeltemp;
+    
+end
+medoid = mode(ci2);
+
+save('../ticelJudgments.mat', 'medoid', 'ensemble','-append')

### 2017-12-14 experiments/timbralSimilaritySol/filterData.m

@@ -1,0 +1,13 @@
+
+function data = filterData(data, idx)
+
+data.features = data.features(idx, :);
+data.mode = data.mode(idx);
+data.modeFamily = data.modeFamily(idx);
+data.instrument = data.instrument(idx);
+data.family = data.family(idx);
+data.file = data.file(idx);
+if isfield(data, 'judgments')
+   data.judgments = data.judgments(:, idx); 
+end
+

### 2017-12-14 experiments/timbralSimilaritySol/getFeatures.m

@@ -1,0 +1,119 @@
+function data = getFeatures(config, data, setting, step)
+
+if length(data.mode)<size(data.features, 1)
+ data.features(end, :) = []; % TODO fix this
+end
+if length(data.mode)<length(data.file)
+ data.file(end) = []; % TODO fix this
+end
+ 
+
+
+switch setting.split
+    case 'none'
+        idx = 1:length(data.file);
+    case 'octave'
+        idx = octaveSelection(config, data);
+    otherwise
+        nbItems = length(data.mode);
+        idx = randi(nbItems, ceil(nbItems*str2num(setting.split)/100), 1);
+        if step==2 && setting.test
+            idx = setdiff(1:nbItems, idx);
+        end
+end
+data = filterData(data, idx);
+% 
+% switch setting.reference
+%     case 'judgments'
+%         idx = [];
+%         gt = [];
+%         d = load([config.codePath 'ticelJudgments.mat']);
+%         for k=1:length(d.names)
+%             oct = strfind(data.file, d.names{k});
+%             oct = find(~cellfun(@isempty, oct));
+%             idx = [idx oct];
+%             gt = [gt repmat(d.ci(:, k), 1, length(oct))];
+%         end
+%         [idx, ia] = unique(idx);
+%         data = filterData(data, idx);
+%         data.judgments = gt(:, ia); %TODO fix this
+% end
+idx = removeLessThan(data, setting);
+data = filterData(data, idx);
+features = data.features;
+
+switch setting.cut
+    case 1
+        switch setting.features
+            case 'mfcc'
+                features = features(:, 1:13);
+            case 'scat'
+                var=nanstd(features); % TODO remove ?
+                var=var/(sum(var));
+                [vars, I] = sort(var,2,'descend');
+                features=features(:,I);
+                sumvar=cumsum(vars);
+                 ix=find(sumvar>.83,1);
+                features=features(:,1:ix);
+        end
+end
+
+switch setting.median
+    case 1
+        features = features./repmat(median(features,1),size(features, 1),1);
+end
+
+switch setting.compress
+    case 1
+        features=log1p(features/1e-3);
+end
+
+if ~isnan(setting.expand) && setting.expand
+     features = expandFeatures(features, setting.expand);
+end
+
+switch setting.randomize
+    case 1
+     features = randn(size(features));
+end
+
[…]

### 2017-12-14 experiments/timbralSimilaritySol/handleJudgments.m

@@ -1,0 +1,25 @@
+function [data, judgments] = handleJudgments(config, data, average)
+
+if ~exist('average', 'var'), average = 0; end
+
+if length(data.mode)<length(data.file), data.file(end) = []; end
+idx = [];
+gt = [];
+mgt = [];
+egt = [];
+d = load([config.codePath 'ticelJudgments.mat']);
+for k=1:length(d.names)
+    oct = strfind(data.file, d.names{k});
+    oct = find(~cellfun(@isempty, oct));
+    idx = [idx oct];
+    gt = [gt repmat(d.ci(:, k), 1, length(oct))];
+    % mgt = [mgt repmat(d.medoid(k), 1, length(oct))];
+    egt = [egt repmat(d.ensemble(k), 1, length(oct))];
+end
+[idx, ia] = unique(idx);
+if average
+    judgments = egt(:, ia);
+else
+    judgments = gt(:, ia);
+end
+data = filterData(data, idx);

### 2017-12-14 experiments/timbralSimilaritySol/libs/lda.m

@@ -1,0 +1,91 @@
+% LDA - MATLAB subroutine to perform linear discriminant analysis
+% by Will Dwinnell and Deniz Sevis
+%
+% Use:
+% W = LDA(Input,Target,Priors)
+%
+% W       = discovered linear coefficients (first column is the constants)
+% Input   = predictor data (variables in columns, observations in rows)
+% Target  = target variable (class labels)
+% Priors  = vector of prior probabilities (optional)
+%
+% Note: discriminant coefficients are stored in W in the order of unique(Target)
+%
+% Example:
+%
+% % Generate example data: 2 groups, of 10 and 15, respectively
+% X = [randn(10,2); randn(15,2) + 1.5];  Y = [zeros(10,1); ones(15,1)];
+%
+% % Calculate linear discriminant coefficients
+% W = LDA(X,Y);
+%
+% % Calulcate linear scores for training data
+% L = [ones(25,1) X] * W';
+%
+% % Calculate class probabilities
+% P = exp(L) ./ repmat(sum(exp(L),2),[1 2]);
+%
+%
+% Last modified: Dec-11-2010
+
+
+function W = lda(Input,Target,Priors)
+
+% Determine size of input data
+[n m] = size(Input);
+
+% Discover and count unique class labels
+ClassLabel = unique(Target);
+k = length(ClassLabel);
+
+% Initialize
+nGroup     = NaN(k,1);     % Group counts
+GroupMean  = NaN(k,m);     % Group sample means
+PooledCov  = zeros(m,m);   % Pooled covariance
+W          = NaN(k,m+1);   % model coefficients
+
+if  (nargin >= 3)  PriorProb = Priors;  end
+
+% Loop over classes to perform intermediate calculations
+for i = 1:k,
+    % Establish location and size of each class
+    Group      = (Target == ClassLabel(i));
+    nGroup(i)  = sum(double(Group));
+    
+    % Calculate group mean vectors
+    GroupMean(i,:) = mean(Input(Group,:));
+    
+    % Accumulate pooled covariance information
+    PooledCov = PooledCov + ((nGroup(i) - 1) / (n - k) ).* cov(Input(Group,:));
+end
+
+% Assign prior probabilities
+if  (nargin >= 3)
+    % Use the user-supplied priors
+    PriorProb = Priors;
+else
+    % Use the sample probabilities
+    PriorProb = nGroup / n;
+end
+
[…]

### 2017-12-14 experiments/timbralSimilaritySol/libs/lmnn-fun/lmnn2/lmnn2.m

@@ -1,0 +1,610 @@
+function [L,Det]=lmnn2(x,y,varargin)
+% function [L,Det]=lmnn2(x,y,Kg,'Parameter1',Value1,'Parameter2',Value2,...);
+%
+% Input:
+%
+% x = input matrix (each column is an input vector) 
+% y = labels 
+% (*optional*)  L = initial transformation matrix (e.g eye(size(x,1)))
+% (*optional*) Kg = attract Kg nearest similar labeled vectos
+% 
+% Important Parameters:
+% diagonal = (default false) If set to true, a diagonal matrix is learned  
+% stepsize = (default 1e-09)
+% outdim = (default: size(x,1)) output dimensionality
+% maxiter = maximum number of iterations (default: 1000)
+% validation = (def 0) fraction of data used as validation 
+%               (e.g. 0.2 means 20% of the training data is used as val)
+% validationstep = (def 15) every "valcount" steps do validation
+% quiet = {0,1} surpress output (default=0)  
+% mu = must be >0 and <1 - tradeoff between loss and regularizer (default mu=0.5)
+% subsample = (default 0.1) percentage of constraints that are subsampled (set to 1 for exact solution)
+%
+% Specific parameters (for experts only):
+% correction = (def 15) how many steps between each update 
+%              The number of impostors are fixed for until next "correction"
+% factor = (def 1.1) multiplicative factor by which the
+%         "correction" gab increases
+% obj = (def 1) if 1, solver solves in L, if 0, solver solves in L'*L
+% thresho = (def 1e-9) cut off for change in objective function (if
+%           improvement is less, stop)
+% thresha = (def 1e-22) cut off for stepsize, if stepsize is
+%           smaller stop
+% scale = (def. 0) if 1, all data gets re-scaled s.t. average
+%         distance to closest neighbor is 1
+%
+%
+% Output:
+%
+% L = linear transformation xnew=L*x
[…]

### 2017-12-14 experiments/timbralSimilaritySol/libs/lmnn-fun/lmnn3/lmnnCG.m

@@ -1,0 +1,133 @@
+function [L,Det]=lmnnCG(xTr,yTr,Kg,varargin)
+% function [L,Det]=lmnnCG(x,y,Kg,'Parameter1',Value1,'Parameter2',Value2,...);
+%
+% Input:
+% x = input matrix (each column is an input vector) 
+% y = labels 
+% Kg = attract Kg nearest similar labeled vectors 
+% 
+% Important Parameters:
+% outdim = (default: size(x,1)) output dimensionality
+% maxiter = maximum number of iterations (default: 1000)
+% quiet = true/false surpress output (default=false)  
+%
+%
+% Specific parameters (for experts only):
+% initL = define initial matrix L to be passed in (default initL=[])
+% thresh = termination criterion (default=1e-05)
+% single = true/false performs most computations in sigle precision (default=true)
+% GPU = true/false performs computation on GPU (default: false)
+%
+%
+% Output:
+%
+% L = linear transformation xnew=L*x
+%    
+% Det.obj = objective function over time
+% Det.nimp = number of impostors over time
+% Det.pars = all parameters used in run
+% Det.time = time needed for computation
+% Det.iter = number of iterations
+% Det.verify = verify (results of validation - if used) 
+%  
+% Version 3.0.0
+% copyright by Kilian Q. Weinbergerr (2005-2015)
+% Cornell University
+% contact kilianweinberger@cornell.edu
+%
+
+global Gx
+global Gy
+global pars;
+
+[d,n]=size(xTr);
+% sanity tests:
+if size(yTr,2)<size(yTr,1), yTr=yTr';end;
+assert(size(yTr,2)==n,'There must be as many labels as inputs. (Each _column_ is an input.)')
+assert(size(yTr,2)>0,'Data must be non-empty.');
+
+minclass=min(diff(find(diff([min(yTr)-1 sort(yTr) max(yTr)+1]))));
+if minclass<=Kg,
+ Kg=min(Kg,minclass-1);
+ fprintf('Warning: K too high. Setting K=%i\n',minclass-1);
+end;
+
+if exist('SODWsp','file')~=2,
+	fprintf('Helperfunctions not in path ...\n')
+	if exist('setpaths3.m','file')==2,
+		fprintf('calling setpaths3.m ...\n')
+		setpaths3();
+	end;
[…]

### 2017-12-14 experiments/timbralSimilaritySol/octaveSelection.m

@@ -1,0 +1,42 @@
+ function idx = octaveSelection(config, data, octaveNumber)
+ 
+ if ~exist('octaveNumber', 'var'), octaveNumber = 4; end
+
+%  config.inputPath = '/data/databases/instruments/SOL/';
+ 
+fid = fopen([config.inputPath 'fileList.txt']);
+fileList = fscanf(fid, '%s\n');
+fclose(fid);
+fileList = regexp(fileList, '.wav', 'split');
+fileList(end)=[];
+
+D = regexp(fileList, '/', 'split');
+for k=1:length(D)-1,
+    dk = D{k};
+    instrument{k} = dk{1};
+    mode{k} = [dk{1} '-' dk{2}];
+    name{k} = dk{3};
+end
+
+oct = strfind(name, [num2str(octaveNumber) '-']);
+oct = find(~cellfun(@isempty, oct));
+
+mul = strfind(name, 'mul');
+mul = find(~cellfun(@isempty, mul));
+
+oct = setdiff(oct, mul);
+
+idx = oct;
+
+% add unpitched
+% notes = {'A', 'B', 'C', 'D', 'E', 'F', 'G'};
+% pitched = [];
+% for k=1:length(notes)
+%    c = strfind(name, notes{k});
+%    pitched = [pitched find(~cellfun(@isempty, c))];
+% end
+% 
+% unpitched = setdiff(1:length(name), pitched);
+% idx = [oct unpitched];
+
+

### 2017-12-14 experiments/timbralSimilaritySol/timbralSimilaritySol.m

@@ -1,0 +1,190 @@
+function timbralSimilaritySol(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
+% Welcome to the main entry point of timbralSimilaritySol                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'tisiso';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]

### 2017-12-14 experiments/timbralSimilaritySol/tisiso1features.m

@@ -1,0 +1,98 @@
+function [config, store, obs] = tisiso1features(config, setting, data)
+% tisiso1features FEATURES step of the expLanes experiment timbralSimilaritySol
+%    [config, store, obs] = tisiso1features(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: Mathieu Lagrange
+% Date: 09-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, timbralSimilaritySol('do', 1, 'mask', {4 1}); return; else store=[]; obs=[]; end
+
+% fid = fopen([config.inputPath 'fileList.txt']);
+% fileList = textscan(fid, '%s/%s\n');
+% fileList = fileList{1};
+% fclose(fid);
+
+fid = fopen([config.inputPath 'fileList.txt']);
+fileList = fscanf(fid, '%s\n');
+fclose(fid);
+fileList = regexp(fileList, '.wav', 'split');
+fileList(end)=[];
+
+D = regexp(fileList, '/', 'split');
+for k=1:length(D)-1,
+    dk = D{k};
+    instrument{k} = dk{1};
+    mode{k} = [dk{1} '-' dk{2}];
+    modeFamily{k} = dk{2};
+    name{k} = dk{3};
+end
+
+si = regexp(instrument, '-', 'split');
+for k=1:length(si),
+    sik = si{k};
+    family{k} = sik{1};
+end
+
+obs.nbInstrument = length(unique(instrument));
+obs.nbMode = length(unique(mode));
+obs.nbFamily = length(unique(family));
+obs.nbmodeFamily = length(unique(modeFamily));
+obs.nbFiles = length(fileList);
+
+[~, ~, gt] = unique(instrument);
+obs.clInstrument = hist(gt, (1:obs.nbInstrument));
+[~, ~, gt] = unique(mode);
+obs.clMode = hist(gt, (1:obs.nbMode));
+[~, ~, gt] = unique(family);
+obs.clFamily = hist(gt, (1:obs.nbFamily));
+[~, ~, gt] = unique(modeFamily);
+obs.clmodeFamily = hist(gt, (1:obs.nbmodeFamily));
+
+store.instrument = instrument;
+store.mode = mode;
+store.modeFamily = modeFamily;
[…]

### 2017-12-14 experiments/timbralSimilaritySol/tisiso2project.m

@@ -1,0 +1,60 @@
+function [config, store, obs] = tisiso2project(config, setting, data)
+% tisiso2project PROJECT step of the expLanes experiment timbralSimilaritySol
+%    [config, store, obs] = tisiso2project(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: Mathieu Lagrange
+% Date: 09-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, timbralSimilaritySol('do', 2, 'mask', {2 1 2 1 1 5 1 1 2 0 0 2 1 2}); return; else store=[]; obs=[]; end
+
+rng(0);
+
+% store = expLoad(config, '', 2);
+% if iscell(store.projection)
+%     for k=1:length(store.projection)
+%         projection(:, :, k) = store.projection{k};
+%     end
+%     store.projection = projection;
+% end
+% return
+if strcmp(setting.projection, 'none') || (strcmp(setting.projection, 'lda') && strcmp(setting.reference, 'judgments') && setting.averageJudgment==0)
+    store.projection = [];
+    return
+end
+
+switch setting.reference
+    case 'judgments'
+        [data, judgments] = handleJudgments(config, data, setting.averageJudgment);
+        
+        parfor k=1:size(judgments, 1)
+            ju = judgments(k, :);
+            projection(k, :, :) = process2project(store, config, data, setting, ju);
+        end      
+    otherwise
+        projection = process2project(store, config, data, setting);
+end
+
+store.projection = projection;
+
+function projection = process2project(store, config, data, setting, ju)
+
+if exist('ju', 'var'), data.(setting.reference) = ju; end
+
+data = getFeatures(config, data, setting, config.step.id);
+
+[~, ~, gt] = unique(data.(setting.reference));
+
+switch setting.projection
+    case 'lmnn'
+        projection = lmnnCG(data.features', gt', setting.neighbors, 'quiet', 1);
+    case 'lda'
+        projection = lda(data.features, gt);
+    otherwise
+        projection = [];
+end

### 2017-12-14 experiments/timbralSimilaritySol/tisiso3performance.m

@@ -1,0 +1,82 @@
+function [config, store, obs] = tisiso3performance(config, setting, data)
+% tisiso3performance PERFORMANCE step of the expLanes experiment timbralSimilaritySol
+%    [config, store, obs] = tisiso3performance(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: Mathieu Lagrange
+% Date: 09-Jan-2017
+
+% Set behavior for debug mode
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {2 1 2 1 0 5 1 0 2 2 2 2}); return; else store=[]; obs=[]; end
+
+rng(0);
+
+if ((strcmp(setting.projection, 'lda') || strcmp(setting.projection, 'none')) && strcmp(setting.reference, 'judgments') && setting.averageJudgment==0) || ...
+   (strcmp(setting.projection, 'lmnn') && strcmp(setting.reference, 'judgments') && setting.averageJudgment==1 && setting.separateJudgment==1) || ...
+   (strcmp(setting.projection, 'none') && strcmp(setting.reference, 'judgments') && setting.separateJudgment==1)
+    obs.p=NaN;
+    return
+end
+
+data1 = expLoad(config, '', 1);
+obs.p=[];
+switch setting.reference
+    case 'judgments'
+        [data1, judgments] = handleJudgments(config, data1);              
+        parfor k=1:size(judgments, 1)  % par   
+            p(k) = process3performance(config, data1, data, setting, judgments(k, :), k);
+        end
+        obs.p = p;
+    otherwise
+        obs.p = process3performance(config, data1, data, setting);
+end
+
+function p = process3performance(config, data1, data, setting, judgment, k)
+
+if exist('judgment', 'var')
+    data1.(setting.reference) = judgment;
+end
+data1 = getFeatures(config, data1, setting, config.step.id);
+
+[~, ~, gt] = unique(data1.(setting.reference));
+
+features = data1.features;
+switch setting.projection
+    case 'lmnn'
+        switch setting.reference
+            case 'judgments'
+                if setting.separateJudgment
+                    data.projection = squeeze(data.projection(k, :, :));
+                else
+                    data.projection = squeeze(mean(data.projection, 1));
[…]

### 2017-12-14 experiments/timbralSimilaritySol/tisisoInit.m

@@ -1,0 +1,10 @@
+function [config, store] = tisisoInit(config)                              
+% tisisoInit INITIALIZATION of the expLanes experiment timbralSimilaritySol
+%    [config, store] = tisisoInit(config)                                  
+%      - config : expLanes configuration state                             
+%      -- store  : processing data to be saved for the other steps         
+                                                                           
+% Copyright: Mathieu Lagrange                                              
+% Date: 09-Jan-2017                                                        
+                                                                           
+if nargin==0, timbralSimilaritySol(); return; else store=[];  end          

### 2017-12-14 experiments/timbralSimilaritySol/tisisoReport.m

@@ -1,0 +1,79 @@
+function config = tisisoReport(config)
+% tisisoReport REPORTING of the expLanes experiment timbralSimilaritySol
+%    config = tisisoInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: Mathieu Lagrange
+% Date: 09-Jan-2017
+
+if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'perceptualProjectionSlides'); return; end
+
+switch config.reportName
+    case 'debugSlides'
+        
+    case 'perceptualProjectionSlides'
+        
+%         mask = {[2  3], 0, 0, 1, 2, 6, 1, 1, 2, 2, 2, 2, 1, 2};
+%         config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'expand', 'sct', 'percent', 0, 'obs', 1, 'highlight', 0, 'uncertainty', -1);
+%         
+        mask = {[2  3], 0, [1  2], 1, 2, [5 6], 1, 1, 2, 2, 2, 2 1};
+        mask = {[2 3], 5, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
+        config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4);
+   
+        mask = {[2 3], 0, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
+        config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'expand', 'sct', 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'uncertainty', -1);
+        
+    case 'scatteringTechniqueSlides'
+        
+        % database description
+                config = expExpose(config, 't', 'obs', [6 5 7 8 9], 'step', 1, 'mask', {4, 1}, 'orientation', 'vi', 'caption', 'Sol db', 'precision', 0, 'highlight', -1);
+                config = expExpose(config, 't', 'obs', [1:4], 'step', 1, 'mask', {4, 1}, 'orientation', 'vi', 'caption', 'Sol db', 'precision', 0, 'highlight', -1);
+        
+        
+        gt = [1 4];
+        %% mfcc parametrization
+                for k=1:length(gt)
[…]

### 2017-12-14 paper/lagrangeSpontaneousSimilarity2017.tex
secondary stage

@@
 
-The secondary stage aims at providing a metric space that closely match the one resulting from the data gathering by experiment 2.
-For the reason, it involves a learning based process block that weight the features outputed by the primary stage.
+The secondary stage involves a learning based process block that will weight the features outputed by the primary stage.
 
@@
 We therefore consider a supervised setting in order to optimize the representational space provided by the scattering transform.
+To do so, we will consider as our reference the clusterings provided by the subjects of experiment 2.
+Therefore, the goal of the secondary stage is to provide a metric space that closely match the one resulting from the data gathering by experiment 2.
 
-To do so, several
+To do so, one can first consider the linear discriminant analysis (LDA) \cite{duda2000pattern} to project the data into a representational space of $C-1$ dimensions with $C$ being the number of classes of the reference.
+The projection computed using this method is such that the ratio of the between-class distance
+to the within-class distance is maximized, thus achieving maximum discrimination between elements of different class.
+
+Another solution is to compute a projection matrix that maximizes the fact that neighbors in the representational space belong to the same class in the reference clustering while items from different classes are separated by a large margin.
+To do so, the large margin nearest neighbor (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance} is considered.
+One advantage of the latter is that the resulting representational space is of the same dimensionality as the input one whereas the LDA output dimentionality is contrained by the number of classes in the reference grouping.
+
+Those methods are designed for considering only one reference grouping.
+Though, in our case, there is one reference grouping per subject.
+One approach is find consensus grouping that maximally represent the 31 reference groupings.
+
+Another approach is to combine the projection matrices issued from the processing of the projection algorithm over the 31 reference groupings.
+As the number of classes can vary between reference groupings, this method is not feasible using the LDA method.
 

### 2017-12-18 paper/lagrangeSpontaneousSimilarity2017.tex
result sec

@@
 Though, in our case, there is one reference grouping per subject.
-One approach is find consensus grouping that maximally represent the 31 reference groupings.
+One approach is to find one consensus grouping that maximally represent the 31 reference groupings.
+In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approach.
+This consensus is chosen as the one that has the best averaged NMI when compared to the 31 reference groupings \cite{strehl2002cluster}.
 
 Another approach is to combine the projection matrices issued from the processing of the projection algorithm over the 31 reference groupings.
-As the number of classes can vary between reference groupings, this method is not feasible using the LDA method.
+As the number of classes can vary between reference groupings, the combination is not feasible using the LDA method.
+On contrary, the LMNN method can outputs a projection matrix per reference grouping.
+Those projection matrices being of the same dimensionality they can safely be averaged to estimate an consensus projection.
 
@@
 
+As it will be discussed in the next Section, the main application scenario that is considered in this paper is the the recommandation of sounds based on examplar.
+This involves the definition of a similarity measure that behave similarly to the human perception, at least, that conform to the data we gathered in experiment 2.
+
+Our problem is thus casted into a ranking problem, where from several candidates, the algorithm is asked to provide a ranking list of those candidates, sorted by similarity.
+
+\subsection{Data}
+
+As explained in the first section, gathering perceptual similarity data is plagued is plagued with a size problem.
+In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound examplar we consider the whole range of nuance and pitch that is available.
+This leads to a dataset of 9346 sound recordings.
+It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken here as a reasonable assumption.
+
+\subsection{Metric}
+
+The performance of the different methods are evaluated in a ranking setting. The well known precision at rank $k$ is thus considered, with $k=5$.
+
 \subsection{Methods}
 
-\subsection{Data}
+To evaluate the performance of the proposed approach, several baselines are considered for comparison purposes.
+For the first stage, the well known Mel-Frequency Cepstral Coefficients (MFCC)s are considered \cite{rabiner1993fundamentals}.
+The MFCCs are computed from frame of 25ms, the number of mel bands is 40 and all the DCT coefficients are kept.
+The resulting features have a dimensionality of 40.
+The MFCCs are then standardized by removing the mean and dividing by the standard deviation computed over the dataset.
+
+The time scattering is computed over a range frame sizes, from 25 ms to 1 second.
+The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale.
+As the frame increase, the dimensionality increases.
+The resulting scattering features are then logarithmically compressed and normalized using the median computed over the dataset.
+
+The LDA method do not requires any setting of parameters and the LMNN method is set to optimize for a neigbors size of 5.
 

### 2017-12-18 paper/lagrangeSpontaneousSimilarity2017.tex
metric subsec

@@
 
-The performance of the different methods are evaluated in a ranking setting. The well known precision at rank $k$ is thus considered, with $k=5$.
+The performance of the different methods are evaluated in a ranking setting.
+The precision at rank $k$ is thus considered, with $k=5$.
+For one sound recording termed the query, the $k$ the nearest other sound recordings are retrieved.
+The precision is then the ratio of those sound recordings that have the same label as the query in the grouping we consider as reference.
+This process is repeated for each sound recordings taken as query, and the performance is averaged over the queries.
 

### 2017-12-19 paper/lagrangeSpontaneousSimilarity2017.tex
adding data

--- 2017-12-18 paper/lagrangeSpontaneousSimilarity2017.tex
@@
 
-In order to model the early stages of the mamalian auditory system, the time scattering transform \cite{anden2014deep}, obtained by applying a one dimensional
+In order to model the early stages of the mamalian auditory system, the time scattering transform (TST) \cite{anden2014deep}, obtained by applying a one dimensional
 wavelet transform in time to a time-frequency wavelet scalogram.
@@
 One approach is to find one consensus grouping that maximally represent the 31 reference groupings.
-In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approach.
+In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approaches.
 This consensus is chosen as the one that has the best averaged NMI when compared to the 31 reference groupings \cite{strehl2002cluster}.
@@
 
-The performance of the different methods are evaluated in a ranking setting. The well known precision at rank $k$ is thus considered, with $k=5$.
+The performance of the different methods are evaluated in a ranking setting.
+The precision at rank $k$ is thus considered, with $k=5$.
+For one sound recording taken as the query, the $k$ nearest sound recordings are retrieved from the remaining items of the database.
+The precision is the number of neighbors that have the same label as the query in the grouping we consider as reference divided by $k$.
+This process is repeated for each sound recordings taken as query, and the performance is averaged over each query.
+For example, if a sound is played by a Violin, and among the 5 nearest neighbors of this query sound, 2 are played by a Violin, the precision is $0.4$.
+
+When considering several groupings as reference, the precision is averaged per groupings. In our case, the precision is thus computed for the 31 reference groupings and averaged.
 
@@
 The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale.
-As the frame increase, the dimensionality increases.
+As the frame size increase, the dimensionality increases.
+Unless otherwise stated, the frame size is one second.
 The resulting scattering features are then logarithmically compressed and normalized using the median computed over the dataset.
@@
 
+In order to validate the proposed model, some experiments are performed using the above described methods. 3 experimental factors with 2 or 3 modalities are considered:
+\begin{itemize}
+  \item type of features: MFCCs or TST
+  \item type of projection: LDA, LMNN or no projection
+  \item type of reference for the optimization of the projection if any: consensus (single consensus grouping) or reference (the 31 reference groupings)
+\end{itemize}
+
+\begin{table}
+\begin{tabular}{lcc}
+  MFCCs & - & - &   85.07 \pm 6.19 \\
+  MFCCs & LDA &  consensus  &   81.50 \pm 7.65 \\
+MFCCs & LMNN & consensus &   86.22 \pm 5.92 \\
+MFCCs & LMNN & reference & 86.31 \pm 5.91 \\
+% MFCCs & LMNN & reference &   86.18 \pm 6.05 \\
+TST & - & - &   87.01 \pm 5.81 \\
+TST & LDA & consensus &   80.95 \pm 10.37 \\
+TST & LMNN  & reference  &   93.31 \pm 3.92 \\
+TST & LMNN & consensus &   94.80 \pm 3.26 \\
+% TST & LMNN & reference &   98.09 \pm 1.28 \\
+\end{tabular}
+\end{table}
+
+separate judgments another table
+
 For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}.
-For convenience, the features (scattering and mfccs) are alos provided, but the software can be used to process from scratch the SOL database.
+For convenience, the features (scattering and mfccs) are also provided, but the software can be used to process audio data.
 

### 2017-12-22 paper/lagrangeSpontaneousSimilarity2017.tex
conclusion

@@
   \item type of features: MFCCs or TST
-  \item type of projection: LDA, LMNN or no projection
-  \item type of reference for the optimization of the projection if any: consensus (single consensus grouping) or reference (the 31 reference groupings)
+  \item type of projection: LDA, LMNN
+  \item type of reference for the optimization of the projection if any: consensus (single consensus grouping) or multiple (the 31 reference groupings)
 \end{itemize}
+The results are shown on Table \ref{tab:res1}.
+Considering the TST over the MFCCs to implement the primary stage do not improve by much (2 \%) if no secondary stage is used.
+When the best setting of the secondary stage is considered, the gain is higher (7 \%).
+this potentially mean that the extra capabilities of description of the TST over the MFCCs need to be targeted to the task at hand in some way in order to gain the full benefits from its use.
+
+In terms of projection method to implement the secondary stage, the LMNN approach is favored over the LDA, probably due to the fact that the dimensionality reduction induced by the low number of classes relatively to the number of dimensions of the input space strongly degrades the description capabilities of the resulting projection.
+The lower performance of the LDA with respect to no projection is probably partly due to that and the fact that the metric is not computed on the consensus grouping use for computing the LDA.
+
+In this experiment, the consensus approach is favored for TST but not for MFCCs.
+This approach has the considerable advantage of being more computationaly efficicient as with only one reference to optimize from the LMMN training is much faster.
+Though, it is very likely that for a larger number of reference groupings, the multiple LMNN would be useful to better model the diversity of the references.
 
 \begin{table}
-\begin{tabular}{lcc}
-  MFCCs & - & - &   85.07 \pm 6.19 \\
-  MFCCs & LDA &  consensus  &   81.50 \pm 7.65 \\
-MFCCs & LMNN & consensus &   86.22 \pm 5.92 \\
-MFCCs & LMNN & reference & 86.31 \pm 5.91 \\
-% MFCCs & LMNN & reference &   86.18 \pm 6.05 \\
-TST & - & - &   87.01 \pm 5.81 \\
-TST & LDA & consensus &   80.95 \pm 10.37 \\
-TST & LMNN  & reference  &   93.31 \pm 3.92 \\
-TST & LMNN & consensus &   94.80 \pm 3.26 \\
-% TST & LMNN & reference &   98.09 \pm 1.28 \\
+  \caption{Precision achieved by the different settings. Numbers in bold face indicate best performance.}
+  \label{tab:res1}
+  \begin{center}
+\begin{tabular}{lllc}
+features & projection & reference & precision (\%) \\
+  \hline
+  MFCCs & - & - &   85.07 $\pm$ 6.19 \\
+  MFCCs & LDA &  consensus  &   81.50 $\pm$ 7.65 \\
+MFCCs & LMNN & multiple & 86.31 $\pm$ 5.91 \\
+MFCCs & LMNN & consensus &   86.22 $\pm$ 5.92 \\
+% MFCCs & LMNN & reference &   86.18 $\pm$ 6.05 \\
+TST & - & - &   87.01 $\pm$ 5.81 \\
+TST & LDA & consensus &   80.95 $\pm$ 10.37 \\
+TST & LMNN  & multiple  &   93.31 $\pm$ 3.92 \\
+TST & LMNN & consensus &   \textbf{94.80 $\pm$ 3.26} \\
+% TST & LMNN & reference &   98.09 $\pm$ 1.28 \\
 \end{tabular}
+\end{center}
 \end{table}
 
-separate judgments another table
+We now investigate the influence of the frame size, as it is likely that the more observation time is available, the better the description will be as several playing techniques involve strong modulations over long periods of time.
+We therefore consider a frame size for the 2 sets of features ranging from 25 ms.
+to 1s.
+The evolution of the precision with respect to the frame size is shown on Figure \ref{fig:frame}.
+Even by enlarging the frame size, the performance of the MFCCs do not increase due to the lack of modulation's modeling.
+On contrary, the TST is able to use the larger frame size to achieve better precision.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/frame.png}
+\caption{Spatial and group organization provided by subject 1.}
+\label{fig:frame}
+\end{figure}
+
+An interesting phenomenon appears, the gain of performance due to the larger frame size better improves for the TST if the secondary stage is considered.
+Again, it seems that enriching the description is most useful if the description is optimized with a purpose.
+
+We investigated the capabilities of the proposed system to model a perceptual space averaged over several subjects.
+For application purposes that will be discussed in the next Section, it can be useful to consider the reference grouping of only one user and see if the system perform well with this kind of reference.
+To do so, the evaluation protocol is slightly modified.
+First the secondary stage is optimized for the 31 reference groupings independantly and for each resulting projection the precision is computed.
+The 31 precision measures are then averaged and shown in Table \ref{tab:res2}.
+The excellent results of the TST optimized using LMNN demonstrates the usefullness of the proposed approach to model complex similarity space such as the percpetual ones gathered in Experiment 2.
+
+\begin{table}
+  \caption{Precision achieved when considered subject based reference groupings. Numbers in bold face indicate best performance.}
+  \label{tab:res2}
+  \begin{center}
+\begin{tabular}{llc}
+features & projection & precision (\%) \\
+  \hline
+MFCCs & LMNN & 86.18 $\pm$ 6.05 \\
+TST & LMNN & 98.09 $\pm$ 1.28 \\
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 Being solely based on sound, this kind of search can be used when querying by example.
-The user provides an example of the target sound to find \ipt that match the timbre description of the query, be it an onomatepea, a tone from another instrument or any kind of environmental sound.
+The user provides an example of the target sound to find \ipt that match the timbre description of the query, be it an onomatepea, a tone from another instrument or any kind of sound.
 
@@
 Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{pauleve2010locality}.
-The secondary stage being learnt, it can easily adapt to user tastes and usage history by reweighting the scattering features for improved search relevance.
+The secondary stage being learnt, it can easily adapt to user tastes and usage history by reweighting the features for improved search relevance.
+
+\section{Conclusion}
+
+The modeling of the spontaneous perceptual similarity have been studied.
+Experimental data have been collected using two perceptual tests whose outomes have been analysed to produce reference data.
+This data is then modeled computationaly using a two step model.
+
+...
+
+Computational experiments showed an interesting phenomenon that high dimsensional features spaces such as the achieved using the TST are most useful if another computation step is considered, if this second step is able to incorporate some sort of prior knowledge regarding the task at hand.
+It would be interesting to verify if this phenomenon can be observed for other types of targeted tasks, using larger datasets and with a higher level of control regarding the learning procedure to study the impact of potential overfiting or learning biases.
 

### 2017-12-22 experiments/timbralSimilaritySol/tisisoReport.m

@@ -6,7 +6,7 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'perceptualProjectionSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'r', 'reportName', 'perceptualProjectionSlides'); return; end
 
 switch config.reportName
     case 'debugSlides'
@@ -15,7 +15,8 @@ switch config.reportName
         
 %         mask = {[2  3], 0, 0, 1, 2, 6, 1, 1, 2, 2, 2, 2, 1, 2};
 %         config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'expand', 'sct', 'percent', 0, 'obs', 1, 'highlight', 0, 'uncertainty', -1);
-%         
+% 
+
         mask = {[2  3], 0, [1  2], 1, 2, [5 6], 1, 1, 2, 2, 2, 2 1};
         mask = {[2 3], 5, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
         config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4);

### 2018-01-08 paper/lagrangeSpontaneousSimilarity2017.tex
prof read intro

@@
 Musical timbre is a multi-faceted notion that have been extensively studied, mostly focusing on its spectral aspect.
-By studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques, we aim in this paper at better understanding the influence of the change of the playing technique on the perception of timbre.
+By studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques, we aim in this paper at better understanding the influence of the playing technique on the perception of timbre.
 
-A first experiment is conducted by music experts to identify which couple of instrument and playing technique in a given dataset is worth comparing to other couples of instrument / playing technique.
+A first experiment is conducted by music experts to identify which couple of instrument / playing technique in a given dataset is worth comparing to other couples of instrument / playing technique.
 In a second experiment, spontaneous similarity judgments among the retained couples are collected using a canonic free sorting task experiment design.
@@
 Musical timbre is an interesting notion to study as it is perceptively defined \cite{grey1977multidimensional}.
-Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse set of physical stimuli.
-However this diversity can be more easily more controlled than other stimuli such as speech or environmental sounds.
+Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse but also controled and well documented set of physical stimuli.
+Indeed, this diversity can be more easily controlled than other stimuli such as speech or environmental sounds.
 
 The focus of this paper is to study the influence of the playing technique on the perception of timbre.
-
-A popular view of the kind of stimuli we consider is that they can be described by a spread of energy that evolve through time and across frequency.
+A popular view of the kind of stimuli considered here is that they can be described by a distribution of energy that evolve through time and across frequency.
 By following this trend, most researchers translate the exclusion definition of timbre \cite{marozeau2003dependency} into a more specified definition.
@@
 A great deal of litterature focuses on the instantaneous spectrum \cite{grey1978perceptual}.
-Other studies also demonstrated the importance of the onset in the recognition of some musical instruments.
+Other studies also demonstrated the importance of the onset in the recognition of some musical instruments \cite{eronen2001comparison}.
 All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common porcessing scenario that is very commonly used in computational recognition models of timbre \cite{tzanetakis2002musical}.
@@
 The modeling of the spontaneous perceptual similarity have been studied.
-Experimental data have been collected using two perceptual tests whose outomes have been analysed to produce reference data.
+Experimental data have been collected using two perceptual tests whose outomes have been analysed to produce reference data of spontaneous perceptual judgments.
 This data is then modeled computationaly using a two step model.
 
-...
+The first step considers the use of the time scattering transform that is well suited for modeling acoustic signals whose spectrum is modulated over time.
+This transform is considered to induce a perceptual feature space that is transformed by the second step that consist in a learning based projection.
 
-Computational experiments showed an interesting phenomenon that high dimsensional features spaces such as the achieved using the TST are most useful if another computation step is considered, if this second step is able to incorporate some sort of prior knowledge regarding the task at hand.
-It would be interesting to verify if this phenomenon can be observed for other types of targeted tasks, using larger datasets and with a higher level of control regarding the learning procedure to study the impact of potential overfiting or learning biases.
+Computational experiments showed an interesting phenomenon that high dimensional features spaces such as the one achieved using the TST are most useful if another computation step is considered, if this second step is able to incorporate some sort of prior knowledge regarding the task at hand in order to refine the resulting representation.
+It would be interesting to verify if this phenomenon can be observed for other types of targeted tasks, using larger datasets and with a higher level of control regarding the learning procedure to study the impact of potential overfiting or learning biases of this king of techniques.
+
+\section*{Acknowlegments}
+
+The authors would like to thank PSL for funding this research, the CNSMDP for fruitfull collaboration and the IRCAM for providing access to the SOL database.
 

### 2018-01-09 paper/lagrangeSpontaneousSimilarity2017.tex
done with intro

@@
 \usepackage{color}
+\usepackage{xspace}
 
-\newcommand{\ipt}{IPT}
+\newcommand{\ipt}{IPT\xspace}
 
@@
 
-The focus of this paper is to study the influence of the playing technique on the perception of timbre.
+In order further restrict this diversity, we focus in this paper on the influence of the playing technique used for the production of a single tone on the perception of timbre.
 A popular view of the kind of stimuli considered here is that they can be described by a distribution of energy that evolve through time and across frequency.
@@
 
-A great deal of litterature focuses on the instantaneous spectrum \cite{grey1978perceptual}.
+A great deal of litterature focuses on the instantaneous spectrum \cite{grey1978perceptual} to model this distribution.
 Other studies also demonstrated the importance of the onset in the recognition of some musical instruments \cite{eronen2001comparison}.
-All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common porcessing scenario that is very commonly used in computational recognition models of timbre \cite{tzanetakis2002musical}.
+All those studies generally assume a rather short term model where the waveform is observed in a range that do not exceed 100 ms, a common processing scenario that is very commonly used in computational recognition models of timbre \cite{tzanetakis2002musical}.
 
-Agus \& al effectively showed that a great deal of instrument recognition can be achieved with very little listening time \cite{agus2012fast}.
+Agus \& al effectively showed that instrument recognition can be achieved with very little listening time \cite{agus2012fast}.
 Though, instrument recognition is only part of musical timbre perception.
@@
 
-Interestingly, modern physiological models of the primary mamalian auditory system also finds that the evolution of the energy distribution at longer time scales matters, \textit{i.e.} how the energy modulates through time for given frequencies.
+Interestingly, modern physiological models of the primary mamalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
 Dau's model acounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
 Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
-Though, the use of numerical implementations of this model did not fully demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
+Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
 
@@
 A notion that have not been extensively studied in the litterature \cite{burred2010dynamic}.
-Compared the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at finer grain of detail.
+Compared to the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to \ipt similarity search in the usage scenario of \ipt database browsing.
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to instrument / playing technique (\ipt) similarity search in the usage scenario of \ipt database browsing.
 

### 2018-01-09 paper/lagrangeSpontaneousSimilarity2017.tex
done up to model

@@
 \newcommand{\ipt}{IPT\xspace}
+\newcommand{\ipts}{IPTs\xspace}
 
@@
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to instrument / playing technique (\ipt) similarity search in the usage scenario of \ipt database browsing.
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of timbral similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to define a numerical similarity that can be considered in the usage scenario of browsing databases of intrument / playing technique couples (\ipts).
 
@@
 
-The dataset considered in this study is taken from the studio-on-line music library  \cite{peeters2000instrument}.
+The dataset considered in this study is taken from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
 It has audio recordings of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 samples.
-The way the dataset is segmented is explained in more details in Appendix \ref{sec:dataset}.
+The way the dataset is structured is explained in more details in Appendix \ref{sec:dataset}.
 
@@
 When considering $n$ items, a complete exploration of the similarity space requires the filling of a $n^2$ matrix.
-Assuming a symmetric similarity (the similarity of A to B is equl to the similarity of B to A) reduces the number but not by much.
+Assuming a symmetric similarity, \textit{i.e.} the similarity of A to B is equal to the similarity of B to A, reduces the number but not by much.
 
-In an attempt to reduce the number of items, we decided to perform a selection experiment where the subject is asked to give his opinion on which \ipt is relevant to study.
-An \ipt is said to be relevant to study if it is likely to be associated to another \ipt.
+In order to reduce the number of items, we decided to perform a selection experiment where the subject is asked to give his opinion on which \ipts is relevant to study.
+An \ipt is said to be relevant to study if it is likely to be associated to another \ipt of another instrument.
 Interest is rated on a 7 ticks scale.
 The subject can listen to all the different nuance and pitch samples of each \ipt.
-The experiment is over when all the \ipt have been rated.
+The experiment is over when all the \ipts have been rated.
 
@@
 \begin{itemize}
-  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt.
+  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
   \item Four stars: there is a proximity on one aspect of the sound between this \ipt and of an \ipt of another instrument, but this one is neither decisive nor obvious.
-  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument, it is interesting to keep it.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
 \end{itemize}
 
-Two music composition professors of the CNSMDP, a renowned French school of music, performed the test. All the couples with a grade higher than three are selected, bringing 78 \ipt to study.
+Two music composition professors of the "conservatoire national sup\'erieur de musique et de danse de Paris", a renowned French school of music, performed the test.
+All the couples with a grade higher than three are selected, bringing 78 \ipts to study.
 
@@
 
-The second experiment aims at gathering spontaneous judgment of similarity for the 78 \ipt selected in the previous experiment.
-To do so, we chose a canonical free sorting task for two main reasons.
-First, this kind of task gives the ability to the subject to listen to the whole set of \ipt before performing the task.
+The second experiment aims at gathering spontaneous judgment of similarity for the 78 \ipts selected in the previous experiment.
+To do so, we consider a canonical free sorting task for two main reasons.
+First, this kind of task gives the ability to the subject to listen to the whole set of \ipts before performing the task.
 Second, it gives more freedom to the subject, thus potentially more interest to perform the task than a lengthy series of triplet dissimilarity ratings.
 
-The subjects are thus asked to organize the 78 \ipt displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
+The subjects are thus asked to organize the 78 \ipts displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
 The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to follow its own judgment.
 
-The \ipt can be listened to by hovering the mouse on the corresponding dot.
+The plane is displayed on a computer screen using a web based interface.
+The \ipts can be listened to by hovering the mouse on the corresponding dot.
 The experiment is over when the sujbect assigned a color to all the dots.
@@
 The experiment have been performed on a volountary basis without gratification.
+Subjects were asked to use headphones at a comfortable listening level.
 
@@
 The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
-
 We therefore now only study the properties of the groupings performed by the subject prior to converting them into an overall similarity matrix.
@@
 To answer this question, each partition provided by cutting the dendrogram at successive levels is compared to two reference partitions.
-One reference is the partition of the \ipt when labeled with the instrument used (I) and the other if the partition of the \ipt when labeled with the playing technique used (Pt).
+One reference is the partition of the \ipts when labeled with the instrument used (I) and the other if the partition of the \ipts when labeled with the playing technique used (Pt).
 
-There are many metrics to compare two partitions by evaluating their degree of correspondance \cite{wagner2007comparing}.
+There are many metrics available to compare two partitions by evaluating their degree of correspondance \cite{wagner2007comparing}.
 The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondance analysis and the two partition are not required to have the same number of groups.
@@
 Normalizing factors $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
-\ml{change legend}}
+\ml{change legend, and use only one null}}
 \label{fig:clusters}
@@
 
-There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipt.
-Also, for both curves, a global trend of increase with respect to the number of clusters in the $j$ partition can be observed.
+There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipts.
+Also, for both curves, a global trend of increase in the $j/x$ partitions with respect to the number of clusters can be observed.
 This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}.
-To compensate for this bias, one can consider the NMI that would be achieved by comparing several random partitions of increasing number of clusters to the reference one, be it the instrument partition ($nullI$) or the playing technique partition ($nullPt$).
-The resulting averaged NMI for 100 randomly generated partitions are displayed on Figure \ref{fig:clusters}.
-Those curves are useful for normalizing $j/I$ and $j/Pt$ by simply substracting the so-called null NMIs:
+To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters.
+This so-called null partition is termed $null$.
+The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}.
+Normalizing $j/I$ and $j/Pt$ to reduce the impact of this bias, we simply substract the $null$ NMI:
 \begin{eqnarray}
-  nj/I &=& j/I - nullI  \\
-  nj/Pt &=& j/Pt - nullPt  \\
+  nj/I &=& j/I - null \\
+  nj/Pt &=& j/Pt - null  \\
 \end{eqnarray}
@@
 The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
-The ordering of \ipt on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
+The ordering of \ipts on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
 The partition on top is the result of the clustering with 8 clusters.
@@
 \includegraphics[width = \textwidth]{figures/groupModes.png}
-\caption{Clustering \ipt using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
+\caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
 The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.}
@@
 \includegraphics[width = \textwidth]{figures/groupInstruments.png}
-\caption{Clustering \ipt using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
+\caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
 The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.}

### 2018-01-11 paper/lagrangeSpontaneousSimilarity2017.tex
done up to results

@@
 
-As discussed in the previous section, only the grouping given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+As discussed in the previous section, only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
 The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
-We therefore now only study the properties of the groupings performed by the subject prior to converting them into an overall similarity matrix.
+We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
 
@@
 
-A grouping can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
+A clustering can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
 By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that correponds to our measure of the spontaneous perceptual similarity between \ipt.
@@
 
-Once the perceptual data is gathered and controled, our aim is to evaluate the ability of the proposed perceptual model that is now introduced in more details.
-The model is composed of two processing stages, respectively called the primary and the secondary stage, see Figure \ref{fig:model}.
+Once the perceptual data is gathered and controled, our aim is now to evaluate the ability of the proposed two stages perceptual model brifely described and motivated in the Introduction.
+The model is composed of two successive processing stages, respectively called the primary and the secondary stage, see Figure \ref{fig:model}.
 The primary stage is designed to rely on very few meta parameters set arbitrarily by the experimenter to mimic the early stages of the mamalian auditory system.
-On contrary, the secondary stage rely largely on the data it has to model and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
+On contrary, the secondary stage rely largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
 
@@
 \includegraphics[width = \textwidth]{figures/groupInstruments.png}
-\caption{Computational model \ml{TODO}}
+\caption{Computational model composed of two successive processing stages.}
 \label{fig:model}
@@
 
-In order to model the early stages of the mamalian auditory system, the time scattering transform (TST) \cite{anden2014deep}, obtained by applying a one dimensional
-wavelet transform in time to a time-frequency wavelet scalogram.
-This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important \cite{dau1997modeling}.
+In order to model the early stages of the mamalian auditory system, we use the time scattering transform (TST). This transform is obtained by applying a one dimensional
+wavelet transform in time to a time-frequency wavelet scalogram \cite{anden2014deep}.
+It is thus well suited to encode modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important \cite{dau1997modeling}.
 
-This model is chosen over more complex alternatives for the sake of simplicity.
+This model is chosen over more data driven alternatives mainly for the sake of simplicity.
 Deep convolutional networks could have been considered \cite{lee2009unsupervised}, but the learning stage inherent in this kind of methods did not seem relevant for this study.
@@
 
-The secondary stage involves a learning based process block that will weight the features outputed by the primary stage.
+The secondary stage involves a learning based process block that will weight and linearly combine the features outputed by the primary stage.
 
 Indeed, if the properties of the first processing stages of the human auditory system can reasonably be considered as similar for every subjects despite potential hearing losses for some, the later ones is expected to vary more drastically due to their respective education, tastes and experiences with stimuli similar to the \ipt that are considered in this study.
-We therefore consider a supervised setting in order to optimize the representational space provided by the scattering transform.
+We therefore consider a supervised setting in order to optimize the feature space provided by the scattering transform.
 To do so, we will consider as our reference the clusterings provided by the subjects of experiment 2.
-Therefore, the goal of the secondary stage is to provide a metric space that closely match the one resulting from the data gathering by experiment 2.
+Therefore, the goal of the secondary stage is to provide a metric space that closely match the one resulting from the data gathered by experiment 2.
 
-To do so, one can first consider the linear discriminant analysis (LDA) \cite{duda2000pattern} to project the data into a representational space of $C-1$ dimensions with $C$ being the number of classes of the reference.
-The projection computed using this method is such that the ratio of the between-class distance
-to the within-class distance is maximized, thus achieving maximum discrimination between elements of different class.
+To do so, one can first consider the linear discriminant analysis (LDA) \cite{duda2000pattern} to project the data into a representational space of $C-1$ dimensions with $C$ being the number of clusters of the reference.
+In our case study, the reference is the partition of \ipts provided by a subject during experiment 2.
+The projection computed using this method is such that the ratio of the between-class distance to the within-class distance is maximized, thus achieving maximum discrimination between elements of different clusters.
 
-Another solution is to compute a projection matrix that maximizes the fact that neighbors in the representational space belong to the same class in the reference clustering while items from different classes are separated by a large margin.
+Another solution is to compute a projection matrix that maximizes the fact that neighbors in the representational space belong to the same cluster in the reference clustering while items from different clusters are separated by a large margin.
 To do so, the large margin nearest neighbor (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance} is considered.
-One advantage of the latter is that the resulting representational space is of the same dimensionality as the input one whereas the LDA output dimentionality is contrained by the number of classes in the reference grouping.
+One advantage of the latter is that the resulting representational space is of the same dimensionality as the input one whereas the LDA output dimentionality is contrained by the number of classes in the reference clustering.
 
-Those methods are designed for considering only one reference grouping.
-Though, in our case, there is one reference grouping per subject.
-One approach is to find one consensus grouping that maximally represent the 31 reference groupings.
-In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approaches.
-This consensus is chosen as the one that has the best averaged NMI when compared to the 31 reference groupings \cite{strehl2002cluster}.
+An important issue is that those methods are designed for considering only one reference clustering.
+It is thus suited to model the perceptual space of one subject but not many.
+However, in our case, we would like to model a perceptual space that is shared among subjects.
+One approach is to find a consensus clustering that maximally represent the 31 reference clusterings.
+To implement this solution, the chosen consensus clustering is taken from 3 solutions issued from 3 different algorithmic approaches introduced in \cite{strehl2002cluster}.
+This consensus clusterings is chosen among the 3 solutions as the one that has the best averaged NMI when compared to the 31 reference clusterings.
 
-Another approach is to combine the projection matrices issued from the processing of the projection algorithm over the 31 reference groupings.
-As the number of classes can vary between reference groupings, the combination is not feasible using the LDA method.
-On contrary, the LMNN method can outputs a projection matrix per reference grouping.
+Another approach is to combine the projection matrices issued from the processing of the projection algorithm over the 31 reference clusterings.
+As the number of classes can vary between reference clusterings, the kind of combination is not feasible using the LDA method.
+On contrary, the LMNN method aloways outputs a projection matrix with equal dimensions.
 Those projection matrices being of the same dimensionality they can safely be averaged to estimate an consensus projection.
@@
 
-As it will be discussed in the next Section, the main application scenario that is considered in this paper is the the recommandation of sounds based on examplar.
-This involves the definition of a similarity measure that behave similarly to the human perception, at least, that conform to the data we gathered in experiment 2.
+As it will be discussed in the next Section, the main application scenario that is considered in this paper is the recommandation of sounds based on examplar.
+This involves the definition of a similarity measure that behave similarly to the human perception.
+In this paper, the reference used for assessing closseness to human perception is the perceptual data gathered in experiment 2.
 
@@
 As explained in the first section, gathering perceptual similarity data is plagued is plagued with a size problem.
+Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte.
 In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound examplar we consider the whole range of nuance and pitch that is available.
 This leads to a dataset of 9346 sound recordings.
-It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken here as a reasonable assumption.
+It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
 
@@
 For one sound recording taken as the query, the $k$ nearest sound recordings are retrieved from the remaining items of the database.
-The precision is the number of neighbors that have the same label as the query in the grouping we consider as reference divided by $k$.
-This process is repeated for each sound recordings taken as query, and the performance is averaged over each query.
+The precision is the number of neighbors that have the same label as the query in the clustering we consider as reference divided by $k$.
 For example, if a sound is played by a Violin, and among the 5 nearest neighbors of this query sound, 2 are played by a Violin, the precision is $0.4$.
+This process is repeated for each sound recordings taken as query, and the performance is averaged over all the queries.
 
-When considering several groupings as reference, the precision is averaged per groupings. In our case, the precision is thus computed for the 31 reference groupings and averaged.
+When considering several clusterings as reference, the precision is averaged per clusterings. In our case, the precision is thus computed for the 31 reference clusterings and averaged.
 
@@
 For the first stage, the well known Mel-Frequency Cepstral Coefficients (MFCC)s are considered \cite{rabiner1993fundamentals}.
-The MFCCs are computed from frame of 25ms, the number of mel bands is 40 and all the DCT coefficients are kept.
-The resulting features have a dimensionality of 40.
-The MFCCs are then standardized by removing the mean and dividing by the standard deviation computed over the dataset.
+The MFCCs are computed from audio frame of 25ms to 1 second, the number of mel bands is 40 and all the discrete cosine transform coefficients are kept.
+The resulting features have thus a dimensionality of 40.
+The MFCCs are then standardized by removing the mean and dividing by the standard deviation computed over the whole dataset.
 
-The time scattering is computed over a range frame sizes, from 25 ms to 1 second.
+The time scattering transform is computed over a range frame sizes, from 25 ms to 1 second.
 The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale.
@@
 Unless otherwise stated, the frame size is one second.
-The resulting scattering features are then logarithmically compressed and normalized using the median computed over the dataset.
+The resulting scattering features are then logarithmically compressed and normalized using the median computed over the whole dataset.
 
@@
 
-In order to validate the proposed model, some experiments are performed using the above described methods. 3 experimental factors with 2 or 3 modalities are considered:
+In order to validate the proposed model, experiments are performed using the above described methods. 3 experimental factors with 2 to 3 modalities are considered:
 \begin{itemize}
@@
   \item type of projection: LDA, LMNN
-  \item type of reference for the optimization of the projection if any: consensus (single consensus grouping) or multiple (the 31 reference groupings)
+  \item type of reference for the optimization of the projection if any: consensus (single consensus clustering) or multiple (the 31 reference clusterings)
 \end{itemize}
@@
 In terms of projection method to implement the secondary stage, the LMNN approach is favored over the LDA, probably due to the fact that the dimensionality reduction induced by the low number of classes relatively to the number of dimensions of the input space strongly degrades the description capabilities of the resulting projection.
-The lower performance of the LDA with respect to no projection is probably partly due to that and the fact that the metric is not computed on the consensus grouping use for computing the LDA.
+The lower performance of the LDA with respect to no projection is probably partly due to that and the fact that the metric is not computed on the consensus clustering use for computing the LDA.
 
@@
 This approach has the considerable advantage of being more computationaly efficicient as with only one reference to optimize from the LMMN training is much faster.
-Though, it is very likely that for a larger number of reference groupings, the multiple LMNN would be useful to better model the diversity of the references.
+Though, it is very likely that for a larger number of reference clusterings, the multiple LMNN would be useful to better model the diversity of the references.
 
@@
 We investigated the capabilities of the proposed system to model a perceptual space averaged over several subjects.
-For application purposes that will be discussed in the next Section, it can be useful to consider the reference grouping of only one user and see if the system perform well with this kind of reference.
+For application purposes that will be discussed in the next Section, it can be useful to consider the reference clustering of only one user and see if the system perform well with this kind of reference.
 To do so, the evaluation protocol is slightly modified.
-First the secondary stage is optimized for the 31 reference groupings independantly and for each resulting projection the precision is computed.
+First the secondary stage is optimized for the 31 reference clusterings independantly and for each resulting projection the precision is computed.
 The 31 precision measures are then averaged and shown in Table \ref{tab:res2}.
@@
 \begin{table}
-  \caption{Precision achieved when considered subject based reference groupings. Numbers in bold face indicate best performance.}
+  \caption{Precision achieved when considered subject based reference clusterings. Numbers in bold face indicate best performance.}
   \label{tab:res2}

### 2018-01-15 code/clusterAnalysis.m
updated legend

@@ -54,9 +54,10 @@ figure(1)
 plot([ai; am; ari; arm; ai-ari; am-arm]', 'LineWidth', 2)
 xlabel('Number of clusters')
 ylabel('Nmi')
-legend({'judgmentsInstruments', 'judgmentsModes', 'nullInstruments', 'nullModes', 'normalizedJudgmentsInstruments', 'normalizedJudgmentsModes'})
+legend({'j/I', 'j/Pt', 'nullI', 'nullPt', 'nj/I', 'nj/Pt'})
 set(gca, 'FontSize', 16)
 saveas(gcf, 'figures/clusterAnalysis', 'png')
+saveas(gcf, 'figures/clusterAnalysis', 'fig')
 
 
 figure(2)

### 2018-01-15 paper/lagrangeSpontaneousSimilarity2017.tex

@@
 
-Musical timbre is a multi-faceted notion that have been extensively studied, mostly focusing on its spectral aspect.
-By studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques, we aim in this paper at better understanding the influence of the playing technique on the perception of timbre.
+Musical timbre is a multi-faceted notion that have been extensively studied, mostly by focusing on its spectral aspect.
+The aim of this paper is to better understand the influence of the playing technique on the perception of musical timbre.
+We do so by studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques,
 
@@
 To further study the outcomes of this experiment, we assume a two-step computational model of human perception, where the acoustic signal is fed to a statically designed processing unit that accounts for frequency and temporal modulations.
-The resulting features are then projected using a supervised technique which considers as input the perceptual similarity judgments gathered in the second experiment.
+The resulting features are then projected in a supervised manner so that the resulting descrption space matches the perceptual similarity judgments gathered in the second experiment.
 Numerical experiments show that the induced perceptual space is able to approximate perceptual data with satisfying accuracy.
@@
 
-Musical timbre is an interesting notion to study as it is perceptively defined \cite{grey1977multidimensional}.
+Being perceptively defined, musical timbre is an interesting notion to study to better comprehend human auditory perception \cite{grey1977multidimensional}.
 Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse but also controled and well documented set of physical stimuli.
@@
 
-In order further restrict this diversity, we focus in this paper on the influence of the playing technique used for the production of a single tone on the perception of timbre.
-A popular view of the kind of stimuli considered here is that they can be described by a distribution of energy that evolve through time and across frequency.
+In order to further restrict this diversity, we focus in this paper on the influence of the playing technique used for the production of a single tone on the perception of timbre.
+A popular view of this kind of stimuli is that they can be described by a distribution of energy that evolve through time and across frequency.
 By following this trend, most researchers translate the exclusion definition of timbre \cite{marozeau2003dependency} into a more specified definition.
@@
 $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
-Normalizing factors $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
+Normalizing null partitions based NMI $null$ is used to obtain normalized measures: $nj/I$ and $nj/Pt$.
 \ml{change legend, and use only one null}}

### 2018-01-15 paper/lagrangeSpontaneousSimilarity2017.tex
good dendrogram

@@
 
-\author{Christian El-Hajj, Mathias Rossignol, Grégoire Lafay, Mathieu Lagrange}
+\author{Christian El-Hajj, Mathias Rossignol, Gr\'egoire Lafay, Mathieu Lagrange}
 
@@
 $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
-Normalizing null partitions based NMI $null$ is used to obtain normalized measures: $nj/I$ and $nj/Pt$.
-\ml{change legend, and use only one null}}
+Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.}
 \label{fig:clusters}
@@
 To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters.
-This so-called null partition is termed $null$.
 The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}.
-Normalizing $j/I$ and $j/Pt$ to reduce the impact of this bias, we simply substract the $null$ NMI:
+This so-called null partitions NMI are respectively termed $nullI$ and $nullPt$ depending on the reference used.
+To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply substract the $null$ NMI:
 \begin{eqnarray}

### 2018-01-25 paper/lagrangeSpontaneousSimilarity2017.tex
done vi1
trivial commit to test version control
comprehend -> understand

@@
 
-Being perceptively defined, musical timbre is an interesting notion to study to better comprehend human auditory perception \cite{grey1977multidimensional}.
+Being perceptively defined, musical timbre is an interesting notion to study to better understand human auditory perception \cite{grey1977multidimensional}.
 Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse but also controled and well documented set of physical stimuli.
@@
 
-\begin{figure}
+\begin{figure}[t]
 \center
-\includegraphics[width = \textwidth]{figures/groupInstruments.png}
-\caption{Computational model composed of two successive processing stages.}
+\includegraphics[width = \textwidth]{figures/spontaneousSimilarity.png}
+\caption{Computational model composed of two successive processing stages and alternative implementations.}
 \label{fig:model}

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
title capitalization
the title should be changed (played, playing)
descrption -> description

@@
 % ------
-\title{Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques}
+\title{Spontaneous Similarity Judgements of Musical Tones played with Various Playing Techniques}
 
@@
 To further study the outcomes of this experiment, we assume a two-step computational model of human perception, where the acoustic signal is fed to a statically designed processing unit that accounts for frequency and temporal modulations.
-The resulting features are then projected in a supervised manner so that the resulting descrption space matches the perceptual similarity judgments gathered in the second experiment.
+The resulting features are then projected in a supervised manner so that the resulting description space matches the perceptual similarity judgments gathered in the second experiment.
 Numerical experiments show that the induced perceptual space is able to approximate perceptual data with satisfying accuracy.

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
80 column abstract
purple comments are from me (VL).
disable sloppy temporarily
we’ll restore it at the end
purple comments correspond to me (VL)
comment on the choice of Q = 8

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% HEADER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \documentclass{article}
@@
 \newcommand{\ml}[1]{\textcolor{blue}{ML : #1}}
+\newcommand{\vl}[1]{\textcolor{purple}{VL : #1}}
 
-% Title.
-% ------
-\title{Spontaneous Similarity Judgements of Musical Tones played with Various Playing Techniques}
-
-\author{Christian El-Hajj, Mathias Rossignol, Gr\'egoire Lafay, Mathieu Lagrange}
-
-\sloppy % please retain sloppy command for improved formatting
+%\sloppy % please retain sloppy command for improved formatting
 
@@
 \maketitle
-%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
 
-Musical timbre is a multi-faceted notion that have been extensively studied, mostly by focusing on its spectral aspect.
-The aim of this paper is to better understand the influence of the playing technique on the perception of musical timbre.
-We do so by studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques,
+Musical timbre is a multi-faceted notion that have been extensively studied,
+mostly by focusing on its spectral aspect.
+The aim of this paper is to better understand the influence
+of the playing technique on the perception of musical timbre.
+We do so by studying the spontaneous judgments of similarity between several
+instruments played with a diverse set of playing techniques,
 
-A first experiment is conducted by music experts to identify which couple of instrument / playing technique in a given dataset is worth comparing to other couples of instrument / playing technique.
-In a second experiment, spontaneous similarity judgments among the retained couples are collected using a canonic free sorting task experiment design.
+A first experiment is conducted by music experts to identify
+which couple of instrument / playing technique in a given dataset
+is worth comparing to other couples of instrument / playing technique.
+In a second experiment, spontaneous similarity
+judgments among the retained couples are collected
+using a canonic free sorting task experiment design.
 
-To further study the outcomes of this experiment, we assume a two-step computational model of human perception, where the acoustic signal is fed to a statically designed processing unit that accounts for frequency and temporal modulations.
-The resulting features are then projected in a supervised manner so that the resulting description space matches the perceptual similarity judgments gathered in the second experiment.
-Numerical experiments show that the induced perceptual space is able to approximate perceptual data with satisfying accuracy.
+To further study the outcomes of this experiment, we assume a
+two-step computational model of human perception, where the acoustic signal
+is fed to a statically designed processing unit that accounts for frequency
+and temporal modulations.
+The resulting features are then projected in a supervised manner so that
+the resulting description space matches the perceptual similarity judgments
+gathered in the second experiment.
+Numerical experiments show that the induced perceptual space
+is able to approximate perceptual data with satisfying accuracy.
 
 \end{abstract}
-%
+
 \section{Introduction}\label{sec:introduction}
@@
 The time scattering transform is computed over a range frame sizes, from 25 ms to 1 second.
-The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale.
+The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale \vl{The mel scale has a varying quality factor, so I don't see how this can be right}.
 As the frame size increase, the dimensionality increases.

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
New title. Timbral similarity in extended playing techniques: the next frontier in querying instrumental databases
bis
Intro should cite Kotska

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{Timbral similarity between extended playing techniques:
+the next frontier in querying instrumental databases}
+
+\author{
+Christian El-Hajj,
+Mathias Rossignol,
+Gr\'egoire Lafay,
+Mathieu Lagrange}
+
 \begin{document}
-%
 \maketitle
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}\label{sec:introduction}
@@
 
-\section{Experiment 1}\label{sec:xp1}
+\vl{Cite Kostka, Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11, on the musicology of instrumental playing techniques. #biblio}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
Gather Sections 1 and 2 into a common section
comment on pitch invariance

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DATA COLLECTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Data collection}\label{sec:xp1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%% EXPERT JUDGMENTS OF PECULIARITY %%%%%%%%%%%%%%%%%%%%%%
+\subsection{Expert judgments of technique peculiarities}
+
 The dataset considered in this study is taken from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
@@
 
-\section{Experiment 2}\label{sec:xp2}
+%%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
+\subsection{Crowdsourced judgments of pairwise similarity}
 
@@
 It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
+\vl{It's too bad that this "reasonable" assumption is not challenged against the data.
+In music psychology, the "bandwidth for timbre invariance" is of the order of one octave (Handel and Erickson, Music Perception, 2001) or less (Steele and Williams, Music Perception, 2006).
+One way to do it without collecting new similarity judgments would be to break down the precision by pitch of query.
+Presumably we have an inverted U shape with a peak near C4.
+#research #figure #biblio}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on citing Castellengo for definitions of timbre

@@
 
-\vl{Cite Kostka, Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11, on the musicology of instrumental playing techniques. #biblio}
+\vl{On the musicology of extended playing techniques, cite:
+Kostka, S (2016). Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11. #biblio}
+
+\vl{On the two definitions of timbre, cite:
+ Castellengo, M and D Dubois (2007). ?Timbre ou timbres? Propriété
+du signal, de l?instrument, ou construction cognitive?? In: Cahiers
+de la Société québécoise de recherche en musique 9, pp. 25?38
+#biblio}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on citing McAdams for similarity judgments

@@
 
+\vl{On pairwise similarity judgments of timbre, cite:
+McAdams, S et al. (1995). Perceptual scaling of synthesized musical
+timbres: common dimensions, specificities, and latent subject
+classes. In: Psychological research 58.3, pp. 177?192}
+
 \vl{On the musicology of extended playing techniques, cite:
@@
 \vl{On the two definitions of timbre, cite:
- Castellengo, M and D Dubois (2007). ?Timbre ou timbres? Propriété
-du signal, de l?instrument, ou construction cognitive?? In: Cahiers
+ Castellengo, M and D Dubois (2007). Timbre ou timbres? Propriété
+du signal, de l?instrument, ou construction cognitive? In: Cahiers
 de la Société québécoise de recherche en musique 9, pp. 25?38

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on Hornbostel-Sachs
trivial commits

@@
 \newcommand{\ml}[1]{\textcolor{blue}{ML : #1}}
-\newcommand{\vl}[1]{\textcolor{purple}{VL : #1}}
+\newcommand{\vl}[1]{\textcolor{red}{VL : #1}}
 
@@
 timbres: common dimensions, specificities, and latent subject
-classes. In: Psychological research 58.3, pp. 177?192}
+classes. In: Psychological research 58.3, pp. 177-192}
 
 \vl{On the musicology of extended playing techniques, cite:
-Kostka, S (2016). Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11. #biblio}
+Kostka, S (2016). Materials and Techniques of Post Tonal Music, Routledge, 2016, chapter 11. @biblio}
 
 \vl{On the two definitions of timbre, cite:
- Castellengo, M and D Dubois (2007). Timbre ou timbres? Propriété
-du signal, de l?instrument, ou construction cognitive? In: Cahiers
-de la Société québécoise de recherche en musique 9, pp. 25?38
-#biblio}
+ Castellengo, M and D Dubois (2007). Timbre ou timbres ? Propri\'{e}t\'{e}
+du signal, de l'instrument, ou construction cognitive ? In: Cahiers
+de la Soci\'{e}t\'{e} qu\'{e}b\'{e}coise de recherche en musique 9, pp. 25-38
+ @biblio}
+
+\vl{We should probably say a word on the Hornbostel-Sachs taxonomy,
+which is based on player-instrument interaction, but is not outdated
+given the variety of contemporary music techniques. @biblio}
 
@@
 Presumably we have an inverted U shape with a peak near C4.
-#research #figure #biblio}
+@research @figure @biblio}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
fix some spelling mistakes in introduction

@@
 Being perceptively defined, musical timbre is an interesting notion to study to better understand human auditory perception \cite{grey1977multidimensional}.
-Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse but also controled and well documented set of physical stimuli.
+Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse but also controlled and well documented set of physical stimuli.
 Indeed, this diversity can be more easily controlled than other stimuli such as speech or environmental sounds.
@@
 
-A great deal of litterature focuses on the instantaneous spectrum \cite{grey1978perceptual} to model this distribution.
+A great deal of literature focuses on the instantaneous spectrum \cite{grey1978perceptual} to model this distribution.
 Other studies also demonstrated the importance of the onset in the recognition of some musical instruments \cite{eronen2001comparison}.
@@
 
-Interestingly, modern physiological models of the primary mamalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
-Dau's model acounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
+Interestingly, modern physiological models of the primary mammalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
+Dau's model accounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
 Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
-Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefullness of the bandwidth dimension \cite{mesgarani2006discrimination}.
+Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.
 
@@
 We will assume in this paper that this second step is rather opportunistic.
-From the large set of features it recieves, this second stage aims at multiplexing in some way the ones that are the most relevant for the task at hand.
-Related work includes the use of the auditory model proposed by Prof.
-Shamma for studying musical timbre \cite{patil2012music}.
+From the large set of features it receives, this second stage aims at multiplexing in some way the ones that are the most relevant for the task at hand.
+Related work includes the use of the auditory model proposed by Shamma for studying musical timbre \cite{patil2012music}.
 
 Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts.
-A notion that have not been extensively studied in the litterature \cite{burred2010dynamic}.
+A notion that have not been extensively studied in the literature \cite{burred2010dynamic}.
 Compared to the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of timbral similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to define a numerical similarity that can be considered in the usage scenario of browsing databases of intrument / playing technique couples (\ipts).
+The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of timbral similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to define a numerical similarity that can be considered in the usage scenario of browsing databases of instrument / playing technique couples (\ipts).
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on citing Siedenburg et al.

@@
 timbres: common dimensions, specificities, and latent subject
-classes. In: Psychological research 58.3, pp. 177-192}
+classes. In: Psychological research 58.3, pp. 177-192.
+@biblio}
 
@@
  @biblio}
+
+\vl{On the importance of spectrotemporal modulations in the description
+of timbre, and the commonalities and differences between the notions of
+"feature" in music signal processing and music psychology:
+Siedenburg, Fujinaga, and McAdams. "A comparison of
+approaches to timbre descriptors in music information retrieval
+and music psychology." In: Journal of New Music Research 45.1,
+pp. 27?41. @biblio}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
update capitalization of institutions in acknowledgments

@@
 
-The authors would like to thank PSL for funding this research, the CNSMDP for fruitfull collaboration and the IRCAM for providing access to the SOL database.
+The authors would like to thank PSL for funding this research, CNSMDP for fruitful collaboration and Ircam for providing access to the SOL database.
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
bis
comment on Dendrogram

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ACKNOWLEDGMENTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowlegments}
 
-The authors would like to thank PSL for funding this research, CNSMDP for fruitful collaboration and Ircam for providing access to the SOL database.
+The authors would like to thank
+PSL for funding this research,
+CNSMDP for fruitful collaboration,
+and Ircam for providing access to the SOL database.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\vl{This dendrogram is very hard to read. Abbreviations must be replaced by the full names of every instrument and playing tecniques.
+Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure
+}
+
 \end{document}

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on Table 1
indentation on expert ratings subsection

@@
 
-The dataset considered in this study is taken from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
-It has audio recordings of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+The dataset considered in this study is taken
+from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
+It has audio recordings of 16 different musical instruments played
+with different playing techniques,
+which leads to 143 different couples of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 samples. 
-The way the dataset is structured is explained in more details in Appendix \ref{sec:dataset}.
+The way the dataset is structured
+is explained in more details in Appendix \ref{sec:dataset}.
 
 Studies about perceptual similarity are plagued with a dimensional problem.
-When considering $n$ items, a complete exploration of the similarity space requires the filling of a $n^2$ matrix.
-Assuming a symmetric similarity, \textit{i.e.} the similarity of A to B is equal to the similarity of B to A, reduces the number but not by much.
+When considering $n$ items, a complete exploration of the similarity space
+requires the filling of a $n^2$ matrix. Assuming a symmetric similarity,
+\textit{i.e.} the similarity of A to B is equal to the similarity of B to A,
+reduces the number but not by much.
 
-In order to reduce the number of items, we decided to perform a selection experiment where the subject is asked to give his opinion on which \ipts is relevant to study.
-An \ipt is said to be relevant to study if it is likely to be associated to another \ipt of another instrument.
-Interest is rated on a 7 ticks scale.
-The subject can listen to all the different nuance and pitch samples of each \ipt.
-The experiment is over when all the \ipts have been rated.
+In order to reduce the number of items, we decided to perform a selection
+experiment where the subject is asked to give his opinion on which
+\ipts is relevant to study. An \ipt is said to be relevant to study
+if it is likely to be associated to another \ipt of another instrument.
+Interest is rated on a 7 ticks scale. The subject can listen to all the different
+nuance and pitch samples of each \ipt. The experiment is over when all the
+\ipts have been rated.
 
@@
   \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
-  \item Four stars: there is a proximity on one aspect of the sound between this \ipt and of an \ipt of another instrument, but this one is neither decisive nor obvious.
+  \item Four stars: there is a proximity on one aspect of the sound between this
+  \ipt and of an \ipt of another instrument,
+  but this one is neither decisive nor obvious.
   \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
@@
 
-Two music composition professors of the "conservatoire national sup\'erieur de musique et de danse de Paris", a renowned French school of music, performed the test.
-All the couples with a grade higher than three are selected, bringing 78 \ipts to study.
+Two music composition professors of the
+"conservatoire national sup\'erieur de musique et de danse de Paris",
+a renowned French school of music, performed the test.
+All the couples with a grade higher than three are selected,
+bringing 78 \ipts to study.
 
@@
 Presumably we have an inverted U shape with a peak near C4.
-@research @figure @biblio}
+@research @computation @figure @biblio}
 
@@
 \begin{table}
-  \caption{Precision achieved by the different settings. Numbers in bold face indicate best performance.}
+  \caption{Precision achieved by the different settings.
+Numbers in bold face indicate best performance.
+\vl{Why are the results with TST, LMNN, and "reference" commented?
+Do they correspond to same rater between training and evaluation?
+@methodology}}
   \label{tab:res1}

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
too much passive form in this subsection
this subsection needs a box-a-whisker plot of ratings
comment on similarity of the histogram of expert ratings

@@
 bringing 78 \ipts to study.
+
+\vl{Too much passive form in this subsection. @writing}
+
+\vl{This subsection needs a box-and-whisker plot of ratings, where the 143 different
+couples of instrument and playing techniques are sorted on the x-axis by average rating. @figure @ratings}
+
+\vl{Are the histogram of ratings similar for the two raters? It would be good to make a statistical test for this. @research @stats @ratings}
 

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
comment on the lack of precision on the description of SOL
A bubble plot of number of notes per instrument (x axis) and playing technique (y axis) would be extremely useful
histograms should be half width

@@
 \includegraphics[width = \textwidth]{figures/nbc.png}
-\caption{Histogram of the number of groups.}
+\caption{Histogram of the number of groups.
+\vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
 \label{fig:xp2nbGroup}
@@
 \includegraphics[width = \textwidth]{figures/sbc.png}
-\caption{Histogram of the size of groups.}
+\caption{Histogram of the size of groups.
+\vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
 \label{fig:xp2sizeGroup}
@@
 
+\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+The abbreviations can be removed. @data @writing}
+
+\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+(y axis) would be extremely useful. @data @figure}
+
 \section{Dendrogram}

### 2018-01-28 paper/lagrangeSpontaneousSimilarity2017.tex
note on the definition of IPT
comment on figure 4

@@
 
-The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of timbral similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to define a numerical similarity that can be considered in the usage scenario of browsing databases of instrument / playing technique couples (\ipts).
+The contributions of the paper are as follows:
+1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role in the spontaneous judgments of timbral similarity, 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper and 4) discusses the use of this computational model to define a numerical similarity that can be considered in the usage scenario of browsing databases of instrument / playing technique couples (\ipts).
+
+\vl{It should be made very clear that the IPT taxonomy is
+the set product of the instrument taxonomy
+and the playing technique taxonomy. @writing}
 
@@
 $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
-Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.}
+Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
+\vl{This line plot is very hard to understand.
+The legend box occludes the data on the top-right corner.
+It would be preferable to remove this legend box and label lines
+by placing text upon them,
+so that the plot would be legible to readers with color vision deficiencies.
+Besides, the choice of line labels is cryptic, as it relies on abbreviations.
+The NMI abbreviation should be expanded.
+@figure}}
 \label{fig:clusters}

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
more comments on figure 4
comment on figure 1

@@
 
+\newcommand*{\eg}{e.g.\@\xspace}
 \newcommand{\ipt}{IPT\xspace}
@@
 \includegraphics[width = \textwidth]{figures/xp2example.png}
-\caption{Spatial and group organization provided by subject 1.}
+\caption{Spatial and group organization provided by subject 1.
+\vl{It would be good to show the three stages of the human-computer interface:
+random initialization,
+geometric clustering on screen,
+and color clustering.
+Besides, the figure is not accessible to readers with
+color vision deficiencies at the moment.
+The color of the clusters should be reinforced with
+geometrical information, \eg{} Voronoi cells.
+@data @figure}}
 \label{fig:xp2display}
@@
 Besides, the choice of line labels is cryptic, as it relies on abbreviations.
+The distinction between I and Pt
+can be illustrated by two complementary colors, \eg{} blue versus orange.
+The distinction between j, nj, and null
+can be illustrated by three distinct markers
+(empty circle, filled triangle, asterisk),
+associated with
+three distinct line styles (solid, dashed, dotted).
 The NMI abbreviation should be expanded.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
"i.e." need not be italicized (IEEE). Search and replace

@@
 \newcommand*{\eg}{e.g.\@\xspace}
+\newcommand*{\ie}{i.e.\@\xspace}
 \newcommand{\ipt}{IPT\xspace}
@@
 
-Interestingly, modern physiological models of the primary mammalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.
+Interestingly, modern physiological models of the primary mammalian auditory system also raises the perceptual importance of the evolution of the energy distribution at longer time scales, \ie{} how the energy modulates through time for given frequencies.
 Dau's model accounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
@@
 A notion that have not been extensively studied in the literature \cite{burred2010dynamic}.
-Compared to the reference that is usually taken in timbre studies \textit{i.e.} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.
+Compared to the reference that is usually taken in timbre studies \ie{} the instrument class, taking into account also the playing technique allows us to consider the notion of timbral similarity at a complementary finer grain of detail.
 
@@
 requires the filling of a $n^2$ matrix. Assuming a symmetric similarity,
-\textit{i.e.} the similarity of A to B is equal to the similarity of B to A,
+\ie{} the similarity of A to B is equal to the similarity of B to A,
 reduces the number but not by much.
@@
 
-Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \textit{i.e.} a very diverse set of instruments labels for the \ipt they belong to.
+Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
 Also, some clusters could be merged without stong changes in the overall organisation.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
spell check introduction

@@
 
-What the higher levels of the auditory cortex does remains largely unkown, but we can assume that the degree of freedom of the underlying model is very large contrary to the early stages.
+What the higher levels of the auditory cortex does remains largely unknown, but we can assume that the degree of freedom of the underlying model is very large contrary to the early stages.
 We will assume in this paper that this second step is rather opportunistic.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
Naming the two stages "primary" and "secondary" is not very informative. How about "feature extraction" and "similarity learning"?

@@
 \caption{Spatial and group organization provided by subject 1.
-\vl{It would be good to show the three stages of the human-computer interface:
+\vl{
+Why have you chosen subject 1, who made 14 clusters?
+This is an unusually large number, according to the
+histogram in Figure 2.
+It would be more representative of the population to illustrate this figure
+with the data of a subject with the median number of clusters.
+It would be good to show the three stages of the human-computer interface:
 random initialization,
@@
 On contrary, the secondary stage rely largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
+
+\vl{Naming the two stages "primary" and "secondary" is not very informative.
+How about "feature extraction" and "similarity learning"? @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
organisation -> organization. search and replace.
American spelling

@@
 31 subjects completed the experiment.
-An example of resulting organisation of \ipt is given on Figure \ref{fig:xp2display}.
-The resulting organisation for each subject are available for reader's inspection using the full feature interface provided to the subjects for performing the experiment\footnote{https://mathieulagrange.github.io/paperSpontaneousSimilarity/demo}.
+An example of resulting organization of \ipt is given on Figure \ref{fig:xp2display}.
+The resulting organization for each subject are available for reader's inspection using the full feature interface provided to the subjects for performing the experiment\footnote{https://mathieulagrange.github.io/paperSpontaneousSimilarity/demo}.
 
 As discussed in the previous section, only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
-The spatial organisation of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
+The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
 We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
@@
 Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
-Also, some clusters could be merged without stong changes in the overall organisation.
+Also, some clusters could be merged without stong changes in the overall organization.
 This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
volountary -> voluntary
fix spelling errors

@@
 The \ipts can be listened to by hovering the mouse on the corresponding dot.
-The experiment is over when the sujbect assigned a color to all the dots.
+The experiment is over when the subject assigned a color to all the dots.
 The initial location of the dots on the plane is randomized for each subject and can be changed during the experiment by the subject for convenience.
-Eventhough the location of the dots is recorded, this data is not considered during the analysis.
+Event hough the location of the dots is recorded, this data is not considered during the analysis.
 
 The experiment was available online for two months and requests to perform the test have been sent to internal mailing lists of the CNSMDP and international research mailing with a focus on audio and music processing.
-The experiment have been performed on a volountary basis without gratification.
+The experiment have been performed on a voluntary basis without gratification.
 Subjects were asked to use headphones at a comfortable listening level.
@@
 As discussed in the previous section, only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
-The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitely force the fact that the timbral similarity space is two dimensional.
+The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly force the fact that the timbral similarity space is two dimensional.
 We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
@@
 A clustering can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
-By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that correponds to our measure of the spontaneous perceptual similarity between \ipt.
+By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that corresponds to our measure of the spontaneous perceptual similarity between \ipt.
 
@@
 This so-called null partitions NMI are respectively termed $nullI$ and $nullPt$ depending on the reference used.
-To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply substract the $null$ NMI:
+To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply subtract the $null$ NMI:
 \begin{eqnarray}
@@
 Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
-Also, some clusters could be merged without stong changes in the overall organization.
+Also, some clusters could be merged without strong changes in the overall organization.
 This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
@@
 
-Once the perceptual data is gathered and controled, our aim is now to evaluate the ability of the proposed two stages perceptual model brifely described and motivated in the Introduction.
+Once the perceptual data is gathered and controlled, our aim is now to evaluate the ability of the proposed two stages perceptual model brifely described and motivated in the Introduction.
 The model is composed of two successive processing stages, respectively called the primary and the secondary stage, see Figure \ref{fig:model}.
@@
 
-In order to model the early stages of the mamalian auditory system, we use the time scattering transform (TST). This transform is obtained by applying a one dimensional
+In order to model the early stages of the mammalian auditory system, we use the time scattering transform (TST). This transform is obtained by applying a one dimensional
 wavelet transform in time to a time-frequency wavelet scalogram \cite{anden2014deep}.
@@
 
-The secondary stage involves a learning based process block that will weight and linearly combine the features outputed by the primary stage.
+The secondary stage involves a learning based process block that will weight and linearly combine the features produced by the primary stage.
 
@@
 To do so, the large margin nearest neighbor (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance} is considered.
-One advantage of the latter is that the resulting representational space is of the same dimensionality as the input one whereas the LDA output dimentionality is contrained by the number of classes in the reference clustering.
+One advantage of the latter is that the resulting representational space is of the same dimensionality as the input one whereas the LDA output dimensionality is constrained by the number of classes in the reference clustering.
 
@@
 As the number of classes can vary between reference clusterings, the kind of combination is not feasible using the LDA method.
-On contrary, the LMNN method aloways outputs a projection matrix with equal dimensions.
+On contrary, the LMNN method always outputs a projection matrix with equal dimensions.
 Those projection matrices being of the same dimensionality they can safely be averaged to estimate an consensus projection.
@@
 
-As it will be discussed in the next Section, the main application scenario that is considered in this paper is the recommandation of sounds based on examplar.
+As it will be discussed in the next Section, the main application scenario that is considered in this paper is the recommendation of sounds based on exemplar.
 This involves the definition of a similarity measure that behave similarly to the human perception.
-In this paper, the reference used for assessing closseness to human perception is the perceptual data gathered in experiment 2.
+In this paper, the reference used for assessing similarity to human perception is the perceptual data gathered in experiment 2.
 
@@
 Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte.
-In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound examplar we consider the whole range of nuance and pitch that is available.
+In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound exemplar we consider the whole range of nuance and pitch that is available.
 This leads to a dataset of 9346 sound recordings.
@@
 To evaluate the performance of the proposed approach, several baselines are considered for comparison purposes.
-For the first stage, the well known Mel-Frequency Cepstral Coefficients (MFCC)s are considered \cite{rabiner1993fundamentals}.
+For the first stage, the well-known Mel-Frequency Cepstral Coefficients (MFCC)s are considered \cite{rabiner1993fundamentals}.
 The MFCCs are computed from audio frame of 25ms to 1 second, the number of mel bands is 40 and all the discrete cosine transform coefficients are kept.
@@
 
-The LDA method do not requires any setting of parameters and the LMNN method is set to optimize for a neigbors size of 5.
+The LDA method do not requires any setting of parameters and the LMNN method is set to optimize for a neighbors size of 5.
 
@@
 In this experiment, the consensus approach is favored for TST but not for MFCCs.
-This approach has the considerable advantage of being more computationaly efficicient as with only one reference to optimize from the LMMN training is much faster.
+This approach has the considerable advantage of being more computationally efficient as with only one reference to optimize from the LMMN training is much faster.
 Though, it is very likely that for a larger number of reference clusterings, the multiple LMNN would be useful to better model the diversity of the references.
@@
 To do so, the evaluation protocol is slightly modified.
-First the secondary stage is optimized for the 31 reference clusterings independantly and for each resulting projection the precision is computed.
+First the secondary stage is optimized for the 31 reference clusterings independently and for each resulting projection the precision is computed.
 The 31 precision measures are then averaged and shown in Table \ref{tab:res2}.
-The excellent results of the TST optimized using LMNN demonstrates the usefullness of the proposed approach to model complex similarity space such as the percpetual ones gathered in Experiment 2.
+The excellent results of the TST optimized using LMNN demonstrates the usefulness of the proposed approach to model complex similarity space such as the perceptual ones gathered in Experiment 2.
 
@@
 For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}.
-For convenience, the features (scattering and mfccs) are also provided, but the software can be used to process audio data.
+For convenience, the features (scattering and MFCCs) are also provided, but the software can be used to process audio data.
 
@@
 Being solely based on sound, this kind of search can be used when querying by example.
-The user provides an example of the target sound to find \ipt that match the timbre description of the query, be it an onomatepea, a tone from another instrument or any kind of sound.
+The user provides an example of the target sound to find \ipt that match the timbre description of the query, be it an onomatopoeia, a tone from another instrument or any kind of sound.
 
@@
 Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{pauleve2010locality}.
-The secondary stage being learnt, it can easily adapt to user tastes and usage history by reweighting the features for improved search relevance.
+The secondary stage being learnt, it can easily adapt to user tastes and usage history by re-weighting the features for improved search relevance.
 
@@
 The modeling of the spontaneous perceptual similarity have been studied.
-Experimental data have been collected using two perceptual tests whose outomes have been analysed to produce reference data of spontaneous perceptual judgments.
-This data is then modeled computationaly using a two step model.
+Experimental data have been collected using two perceptual tests whose outcomes have been analyzed to produce reference data of spontaneous perceptual judgments.
+This data is then modeled computationally using a two step model.
 
@@
 Computational experiments showed an interesting phenomenon that high dimensional features spaces such as the one achieved using the TST are most useful if another computation step is considered, if this second step is able to incorporate some sort of prior knowledge regarding the task at hand in order to refine the resulting representation.
-It would be interesting to verify if this phenomenon can be observed for other types of targeted tasks, using larger datasets and with a higher level of control regarding the learning procedure to study the impact of potential overfiting or learning biases of this king of techniques.
+It would be interesting to verify if this phenomenon can be observed for other types of targeted tasks, using larger datasets and with a higher level of control regarding the learning procedure to study the impact of potential overfitting or learning biases of this king of techniques.
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
correspondance -> correspondence

@@
 
-There are many metrics available to compare two partitions by evaluating their degree of correspondance \cite{wagner2007comparing}.
-The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondance analysis and the two partition are not required to have the same number of groups.
+There are many metrics available to compare two partitions by evaluating their degree of correspondence \cite{wagner2007comparing}.
+The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondence analysis and the two partition are not required to have the same number of groups.
 It also by design bounded between 0 and 1, 1 meaning perfect match.
@@
 
-The correspondances between those partitions and the reference ones are now studied.
+The correspondences between those partitions and the reference ones are now studied.
 For this purpose, Figure \ref{fig:gi} displays two partitions.
@@
 
-As can be seen on Figure \ref{fig:gm}, the correspondance between the playing technique reference and the resulting partition with 5 clusters is much higher.
+As can be seen on Figure \ref{fig:gm}, the correspondence between the playing technique reference and the resulting partition with 5 clusters is much higher.
 Most clusters have a high level of purity and for groups emerge : Pizzicato, Slap, Ordinario, and Ponticello.
@@
 The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
+\vl{This is crucial and should not be deferred to the appendix. @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
on contrary -> on the contrary
search and replace

@@
 The primary stage is designed to rely on very few meta parameters set arbitrarily by the experimenter to mimic the early stages of the mamalian auditory system.
-On contrary, the secondary stage rely largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
+On the contrary, the secondary stage rely largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
 
@@
 Deep convolutional networks could have been considered \cite{lee2009unsupervised}, but the learning stage inherent in this kind of methods did not seem relevant for this study.
-On contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system as those parameters are not expected to change significantly among subjects.
+On the contrary, the scattering transform can be conveniently parametrized to match the perceptual capabilities of the auditory system as those parameters are not expected to change significantly among subjects.
 Also, more complex versions of the scattering like the joint time-frequency scattering \cite{anden2015joint} and the spiral scattering \cite{lostanlen2016wavelet} are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are mandatory to model the primary stage of the auditory system.
@@
 As the number of classes can vary between reference clusterings, the kind of combination is not feasible using the LDA method.
-On contrary, the LMNN method always outputs a projection matrix with equal dimensions.
+On the contrary, the LMNN method always outputs a projection matrix with equal dimensions.
 Those projection matrices being of the same dimensionality they can safely be averaged to estimate an consensus projection.
@@
 Even by enlarging the frame size, the performance of the MFCCs do not increase due to the lack of modulation's modeling.
-On contrary, the TST is able to use the larger frame size to achieve better precision.
+On the contrary, the TST is able to use the larger frame size to achieve better precision.
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on figures 5 and 6

@@
 \caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
-The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.}
+The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.
+\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
 \label{fig:gm}
@@
 \caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
-The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.}
+The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.
+\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
 \label{fig:gi}
@@
 The primary stage is designed to rely on very few meta parameters set arbitrarily by the experimenter to mimic the early stages of the mamalian auditory system.
-On the contrary, the secondary stage rely largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
+On the contrary, the secondary stage relies largely on the data it has to predict and in this respect mimic the high plasticity of the later stages of the mamalian auditory system.
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on figure 7

@@
 \includegraphics[width = \textwidth]{figures/spontaneousSimilarity.png}
-\caption{Computational model composed of two successive processing stages and alternative implementations.}
+\caption{Computational model composed of two successive processing stages and alternative implementations.
+\vl{The boxes "features" and "projection" should be renamed "feature extraction" and "metric learning".
+This figure is not very useful since it does not distinguish training time from prediction time.
+@figure}}
 \label{fig:model}
@@
 Also, more complex versions of the scattering like the joint time-frequency scattering \cite{anden2015joint} and the spiral scattering \cite{lostanlen2016wavelet} are not considered in order to achieve a projection of reasonable dimensionality and for lack of evidence that those dimensions are mandatory to model the primary stage of the auditory system.
+
+\vl{I would suggest trying to avoid the invention of the "TST" acronym for "temporal scattering transform" if possible.
+This acronym does not already exist in the relevant literature. @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on the primary stage

@@
 
+\vl{The argument behind not training a deep convolutional network is a strawman argument.
+There is a also learning stage inherent to our approach: LDA or LMNN. @writing}
+
+\vl{I do believe that time-frequency scattering (or one of its variants, see McDermott or Patil or Sch\"{a}dler)
+is, in a very precise way, mandatory to model the central auditory system. It's OK to think otherwise but that
+should be motivated. @writing}
+
 \vl{I would suggest trying to avoid the invention of the "TST" acronym for "temporal scattering transform" if possible.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
rename primary and secondary stage
comment on consensus clustering

@@
 
-\subsection{Primary stage}
+\subsection{Feature extraction}
 
@@
 
-\subsection{Secondary stage}
+\subsection{Metric learning}
 
-The secondary stage involves a learning based process block that will weight and linearly combine the features produced by the primary stage.
+The secondary stage involves a learning-based process block that will weight and linearly combine the features produced by the primary stage.
 
@@
 This consensus clusterings is chosen among the 3 solutions as the one that has the best averaged NMI when compared to the 31 reference clusterings.
+
+\vl{The paragraph above is the most important in the whole paper, and it deserves more attention. @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
capitalization of "next section"

@@
 
-As it will be discussed in the next Section, the main application scenario that is considered in this paper is the recommendation of sounds based on exemplar.
+As it will be discussed in the next section, the main application scenario that is considered in this paper is the recommendation of sounds based on exemplar.
 This involves the definition of a similarity measure that behave similarly to the human perception.
@@
 We investigated the capabilities of the proposed system to model a perceptual space averaged over several subjects.
-For application purposes that will be discussed in the next Section, it can be useful to consider the reference clustering of only one user and see if the system perform well with this kind of reference.
+For application purposes that will be discussed in the next section, it can be useful to consider the reference clustering of only one user and see if the system perform well with this kind of reference.
 To do so, the evaluation protocol is slightly modified.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
casted -> cast

@@
 On the contrary, the LMNN method always outputs a projection matrix with equal dimensions.
-Those projection matrices being of the same dimensionality they can safely be averaged to estimate an consensus projection.
+Those projection matrices being of the same dimensionality they can safely be averaged to estimate a consensus projection.
+
+\vl{What does "safely" mean here? We are at the epicenter of the paper here, every word matters. @writing}
 
@@
 
-Our problem is thus casted into a ranking problem, where from several candidates, the algorithm is asked to provide a ranking list of those candidates, sorted by similarity.
+Our problem is thus cast into a ranking problem, where from several candidates, the algorithm is asked to provide a ranking list of those candidates, sorted by similarity.
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on 6.2

@@
 reduces the number but not by much.
+
+\vl{The paragraph above needs to be rewritten. It introduces mathematical
+notations that are not used later in the text. The adjective "plagued"
+is derogatory and could be replaced by "hindered". @writing}
 
@@
 
-As explained in the first section, gathering perceptual similarity data is plagued is plagued with a size problem.
+As explained in the first section, gathering perceptual similarity data is plagued with a size problem.
 Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte.
@@
 The precision is the number of neighbors that have the same label as the query in the clustering we consider as reference divided by $k$.
-For example, if a sound is played by a Violin, and among the 5 nearest neighbors of this query sound, 2 are played by a Violin, the precision is $0.4$.
+For instance, if a sound is played by a Violin, and among the 5 nearest neighbors of this query sound, 2 are played by a Violin, the precision is $0.4$.
 This process is repeated for each sound recordings taken as query, and the performance is averaged over all the queries.
+
+\vl{I'm extremely puzzled by the paragraph above. I thought that the point of the paper was precisely to liberate oneself from the taxonomy of musical instruments? @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
"Though," -> "However,"
search and replace

@@
 Agus \& al effectively showed that instrument recognition can be achieved with very little listening time \cite{agus2012fast}.
-Though, instrument recognition is only part of musical timbre perception.
+However, instrument recognition is only part of musical timbre perception.
 We propose in this paper to study the influence of the playing technique on musical timbre percept.
@@
 Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
-Though, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.
+However, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.
 
@@
 
-Though, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
+However, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
 Also, some clusters could be merged without strong changes in the overall organization.
@@
 This approach has the considerable advantage of being more computationally efficient as with only one reference to optimize from the LMMN training is much faster.
-Though, it is very likely that for a larger number of reference clusterings, the multiple LMNN would be useful to better model the diversity of the references.
+However, it is very likely that for a larger number of reference clusterings, the multiple LMNN would be useful to better model the diversity of the references.
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
more comments on writing

@@
 The time scattering transform is computed over a range frame sizes, from 25 ms to 1 second.
-The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale \vl{The mel scale has a varying quality factor, so I don't see how this can be right}.
+The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale.
 As the frame size increase, the dimensionality increases.
@@
 The resulting scattering features are then logarithmically compressed and normalized using the median computed over the whole dataset.
+
+\vl{The mel scale has a varying quality factor, so I don't see how this can be right.}
+\vl{I understand what "logarithmically compressed" and "normalized using the median" because I worked on this before but,
+for someone who is less experienced, this might not be precise enough.}
 
@@
 this potentially mean that the extra capabilities of description of the TST over the MFCCs need to be targeted to the task at hand in some way in order to gain the full benefits from its use.
+
+\vl{The paragraph above is very peremptory and at the same time quite vague.
+I could quote many tasks in which there is no need to "target in some way" the scattering transform to obtain better results than MFCC.
+What scientific conclusion are we drawing here? @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
new title \title{Learning Auditory Similarities Between Instrumental Playing Techniques}
update intro

@@
 
-\title{Timbral similarity between extended playing techniques:
-the next frontier in querying instrumental databases}
+\title{Learning Auditory Similarities Between Instrumental Playing Techniques}
 
@@
 Dau's model accounts for  rate of modulation over frequency bands of fixed bandwidth on a logarithmic frequency scale\cite{dau1997modeling}.
-Shamma's model accounts for rate  of modulations over several bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
+Shamma's model accounts for rate of modulations over several subbands to fully describe the physiological evidence gathered by studying the auditory system of the ferret \cite{yang1992auditory}.
 However, the use of numerical implementations of this model applied to several types of classification tasks of audio data did not fully demonstrated the usefulness of the bandwidth dimension \cite{mesgarani2006discrimination}.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on the use of the word "opportunistic"

@@
 
-What the higher levels of the auditory cortex does remains largely unknown, but we can assume that the degree of freedom of the underlying model is very large contrary to the early stages.
+What the higher levels of the auditory cortex does remains largely unknown, but we can assume that the degree of freedom of the underlying model is very large, unlike at the early stages.
 We will assume in this paper that this second step is rather opportunistic.
@@
 Related work includes the use of the auditory model proposed by Shamma for studying musical timbre \cite{patil2012music}.
+
+\vl{The paragraph above is extremely vague. I don't think that "opportunistic" is the word we're looking for here. @writing}
 
@@
 
-\vl{Too much passive form in this subsection. @writing}
+\vl{Too much passive form in this subsection. This last paragraph ("a renowned French school of music") is un-academic. @writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on figure 8

@@
 \includegraphics[width = \textwidth]{figures/frame.png}
-\caption{Spatial and group organization provided by subject 1.}
+\caption{Spatial and group organization provided by subject 1.
+\vl{This caption is obviously wrong, as it was copied verbatim from Figure 1. The x axis should be made logarithmic, which uneven tick marks.
+The legend is difficult to understand (acronyms, subtle difference between "multiple" and "consensus").
+Again, the legend box could be removed after placing the labels next to the lines themselves. @figure}}
 \label{fig:frame}
@@
 It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the user has a precise understanding of this ontology, and 3) that the user a precise idea of the kind of timbre he is looking to find the \ipt he is searching for.
+
+\vl{Let's use the gender-neutral singular "they" instead of "he".@writing}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on the name for the repository

@@
 
-For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}.
+For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity \vl{I would love a more memorable name.
+@writing}}.
 For convenience, the features (scattering and MFCCs) are also provided, but the software can be used to process audio data.

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
comment on discussion

@@
 Thanks to the digitization capabilities available, gathering large amount of audio recordings of \ipt is achievable and several databases are now available on the market.
-Browsing of such databases can be achieved by keyword based search.
+Browsing of such databases can be achieved by keyword-based search.
 It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the user has a precise understanding of this ontology, and 3) that the user a precise idea of the kind of timbre he is looking to find the \ipt he is searching for.
+
+\vl{Curiously, this paper does not contain a review of such databases available on the market. @biblio}
 
@@
 
-\vl{This dendrogram is very hard to read. Abbreviations must be replaced by the full names of every instrument and playing tecniques.
-Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure
-}
+\vl{This dendrogram is very hard to read. Abbreviations must be replaced by the full names of every instrument and playing techniques.
+Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
last comments on this version of the paper

@@
 
-To do so, one can first consider the linear discriminant analysis (LDA) \cite{duda2000pattern} to project the data into a representational space of $C-1$ dimensions with $C$ being the number of clusters of the reference.
+\vl{I am a bit wary about not giving proper names to the two experiments, but instead calling them "experiment 1" and "experiment 2".
+How about "expert judgments of uniqueness" versus "crowdsourced judgments of similarity"? @writing}
+
+To do so, one can first consider the linear discriminant analysis (LDA) \cite{duda2000pattern} to project the data into a representational space of $(C-1)$ dimensions, with $C$ being the number of clusters of the reference.
 In our case study, the reference is the partition of \ipts provided by a subject during experiment 2.
@@
 
-\vl{The paragraph above is the most important in the whole paper, and it deserves more attention. @writing}
+\vl{The paragraph above is the most important in the whole paper, and it deserves more attention.
+How do these three solutions differ?
+Which one was chosen in the end?
+Did it make a significant difference in NMI?
+Is the consensus clustering shown somewhere in the paper?
+Is it related to the dendrogram in Appendix B?
+I have so many questions.
+@writing}
+
+\vl{I'm afraid there is also a methodological problem here.
+We need to have a hold-out set of IPTs to evaluate the performance of the clustering algorithm after having trained the metric learning algorithm.
+@methodology @computation}
 
@@
 
-\vl{What does "safely" mean here? We are at the epicenter of the paper here, every word matters. @writing}
+\vl{What does "safely" mean here? We are at the epicenter of the paper here: every word matters. @writing}
 
@@
 
-Our problem is thus cast into a ranking problem, where from several candidates, the algorithm is asked to provide a ranking list of those candidates, sorted by similarity.
+Our problem is thus cast into a ranking problem, where from several candidates, the algorithm is asked to provide a ranked list of those candidates, sorted by similarity.
 
@@
 \vl{The mel scale has a varying quality factor, so I don't see how this can be right.}
-\vl{I understand what "logarithmically compressed" and "normalized using the median" because I worked on this before but,
+\vl{I understand what "logarithmically compressed" and "normalized using the median" because I did work on this before but,
 for someone who is less experienced, this might not be precise enough.}
@@
 Numbers in bold face indicate best performance.
-\vl{Why are the results with TST, LMNN, and "reference" commented?
+\vl{What does these variances correspond to?
+Variance across raters, across IPTs, across instances of each IPT?
+Furthermore, why are the results with TST, LMNN, and "reference" commented?
 Do they correspond to same rater between training and evaluation?
-@methodology}}
+@methodology @writing}}
   \label{tab:res1}

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
ml answers to vl comments

@@
 Christian El-Hajj,
+Vincent Lostanlen,
 Mathias Rossignol,
@@
 \vl{The paragraph above is extremely vague. I don't think that "opportunistic" is the word we're looking for here. @writing}
+\ml{ok}
 
@@
 and the playing technique taxonomy. @writing}
+\ml{ok}
 
@@
 given the variety of contemporary music techniques. @biblio}
+\ml{great, I was missing such references}
 
@@
 is derogatory and could be replaced by "hindered". @writing}
+\ml{ok}
 
-In order to reduce the number of items, we decided to perform a selection
+In order to reduce the number of items, we perform a selection
 experiment where the subject is asked to give his opinion on which
@@
 nuance and pitch samples of each \ipt. The experiment is over when all the
-\ipts have been rated.
+\ipts are rated.
 
@@
 Two music composition professors of the
-"conservatoire national sup\'erieur de musique et de danse de Paris",
-a renowned French school of music, performed the test.
+"conservatoire national sup\'erieur de musique et de danse de Paris", performed the test.
 All the couples with a grade higher than three are selected,
 bringing 78 \ipts to study.
-
-\vl{Too much passive form in this subsection. This last paragraph ("a renowned French school of music") is un-academic. @writing}
 
@@
 \vl{Are the histogram of ratings similar for the two raters? It would be good to make a statistical test for this. @research @stats @ratings}
+
+\ml{ok, je regarde si on a toujours les données}
 
@@
 @data @figure}}
+\ml{ok, tu peux utiliser le site de demo et faire une capture d'écran de l'utilisateur qui te va. Pour les autres display, je ne suis pas sur que cela soit utile.}
 \label{fig:xp2display}
@@
 \vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
+\ml{ok}
 \label{fig:xp2nbGroup}
@@
 \vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
+\ml{ok}
 \label{fig:xp2sizeGroup}
@@
 @figure}}
+\ml{ok, le .fig est dispo}
 \label{fig:clusters}
@@
 \vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
+\ml{Je demande a voir, mais je ne suis pas satisfait de cette visu de toute facon. Le code est dispo dans le rep code. Lancer ticelSim en premier.}
 \label{fig:gm}
@@
 \vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
+\ml{ok}
 \label{fig:gi}
@@
 How about "feature extraction" and "similarity learning"? @writing}
+\ml{ok}
 
@@
 There is a also learning stage inherent to our approach: LDA or LMNN. @writing}
+\ml{so argue the lack of large dataset ? Or not talk about it}
 
@@
 should be motivated. @writing}
+\ml{then you can proceed to the rewrite}
 
@@
 This acronym does not already exist in the relevant literature. @writing}
+\ml{this is not an invention, this is a mean to be compact in this paper.}
 
@@
 How about "expert judgments of uniqueness" versus "crowdsourced judgments of similarity"? @writing}
+\ml{J'ai bien peur que cela alourdisse le discours. Mais bon, visiblement on ne repete experiment 2 que 6 fois, ca devrait aller.}
 
@@
 @writing}
+\ml{ok, je refais un code sur cette partie. Le dendrogram n'est pas un clustering, il part de la matrice de similarité. Bon apres, c'est tres lié...}
 
@@
 @methodology @computation}
+\ml{Mmm, je ne suis pas sur, dans l'idée, j'ai besoin d'une projection adaptée a mon contenu, rien ne m'empeche d'optimiser sur tout le corpus.
+On peut imaginer couper en deux mais ca fait pas beaucoup de données.
+Perso, je suis pour laiser ca comme ca et le faire si un relecteur rale.
+Mais je suis ouvert a la discussion.}
 
@@
 \vl{What does "safely" mean here? We are at the epicenter of the paper here: every word matters. @writing}
+\ml{I agree, we should develop a bit.}
 
@@
 @research @computation @figure @biblio}
+\ml{Indeed, It would be interesting but would require some research time.}
 
@@
 \vl{I'm extremely puzzled by the paragraph above. I thought that the point of the paper was precisely to liberate oneself from the taxonomy of musical instruments? @writing}
+\ml{This is just an example.
+May be state better that is htis example the reference is the instrument taxonomy.
+Or even better refer to a clustering reference of one subject, but in this case the cluster denomination is more cryptic, like cluster 1, cluster 2...}
 
@@
 \vl{The mel scale has a varying quality factor, so I don't see how this can be right.}
+\ml{I don't understand}
 \vl{I understand what "logarithmically compressed" and "normalized using the median" because I did work on this before but,
 for someone who is less experienced, this might not be precise enough.}
+\ml{ok}
 
@@
 What scientific conclusion are we drawing here? @writing}
+\ml{to be discussed}
 
@@
 Do they correspond to same rater between training and evaluation?
-@methodology @writing}}
+@methodology @writing}
+  \ml{variance across raters. Do not understand the second question, to be discussed.}}
   \label{tab:res1}
@@
 \includegraphics[width = \textwidth]{figures/frame.png}
-\caption{Spatial and group organization provided by subject 1.
-\vl{This caption is obviously wrong, as it was copied verbatim from Figure 1. The x axis should be made logarithmic, which uneven tick marks.
+\caption{Evolution of the precision with respect to the frame size for the different front ends.
+\vl{The x axis should be made logarithmic, which uneven tick marks.
 The legend is difficult to understand (acronyms, subtle difference between "multiple" and "consensus").
 Again, the legend box could be removed after placing the labels next to the lines themselves. @figure}}
+\ml{Proposition ?}
 \label{fig:frame}
@@
 For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity \vl{I would love a more memorable name.
-@writing}}.
+@writing}\ml{proposition ?
+abbreviation of the new title ?}}.
 For convenience, the features (scattering and MFCCs) are also provided, but the software can be used to process audio data.
@@
 Browsing of such databases can be achieved by keyword-based search.
-It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the user has a precise understanding of this ontology, and 3) that the user a precise idea of the kind of timbre he is looking to find the \ipt he is searching for.
+It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the users have a precise understanding of this ontology, and 3) that the users a precise idea of the kind of timbre they are looking to find the \ipt they are searching for.
 
 \vl{Curiously, this paper does not contain a review of such databases available on the market. @biblio}
-
-\vl{Let's use the gender-neutral singular "they" instead of "he".@writing}
+\ml{Indeed}
 
@@
 \vl{This is crucial and should not be deferred to the appendix. @writing}
+\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
 
@@
 The abbreviations can be removed. @data @writing}
+\ml{}
 
@@
 (y axis) would be extremely useful. @data @figure}
+\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
 
@@
 Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure}
+\ml{Isomap do not show a dendrogram, right ?}
 

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
addition of analyis code of xp1

@@
 with different playing techniques,
-which leads to 143 different couples of instrument / playing technique.
+which leads to 235 different couples of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 samples.
@@
 Two music composition professors of the
-"conservatoire national sup\'erieur de musique et de danse de Paris", performed the test.
+"conservatoire national sup\'erieur de musique et de danse de Paris" performed the test.
 All the couples with a grade higher than three are selected,
@@
 
-\vl{This subsection needs a box-and-whisker plot of ratings, where the 143 different
+\vl{This subsection needs a box-and-whisker plot of ratings, where the 235 different
 couples of instrument and playing techniques are sorted on the x-axis by average rating. @figure @ratings}
@@
 
-\ml{ok, je regarde si on a toujours les données}
+\ml{les histograms sont dans la figure statxp1. Le code pour la generer est dans code/xp1 faire tourner extractData et statData. Un ttest ne rejette pas l'hypothese nulle.}
 
@@
 
-The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of 16 different musical instruments played with different playing techniques, which leads to 143 different couple of instrument / playing technique.
+The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of 16 different musical instruments played with different playing techniques, which leads to 235 different couple of instrument / playing technique.
 For each couple, the pitch and intonation is varied leading to 25444 audio samples.

### 2018-01-29 code/xp1/analyse.m

@@ -1,0 +1,51 @@
+clear all
+addpath(genpath('functions'))
+load('extractedData/results.mat');
+
+data2use={1,9,[1 9]};
+seuil={[5 6 7],[5 6 7],[11 12 13 14]};
+
+for jj=1:length(data2use)
+    
+    rating2use=rating(:,data2use{jj});
+    mode2use=mode(:,1);
+    instrument2use=instrument(:,1);
+    sounds2use=sounds(:,1);
+    path2use=path(:,1);
+    
+    rating2use=sum(rating2use,2);
+    [rating2use,ind]=sort(rating2use,'descend');
+    
+    for ll=1:length(seuil{jj})
+        ind=ind(rating2use>=seuil{jj}(ll));
+        
+        %% Instrument
+        
+        oldIntrument=unique(instrument2use);
+        newInstrument=unique(instrument2use(ind));
+        
+        isNewInstrument=cellfun(@(x) any(strcmp(x,newInstrument)),oldIntrument);
+        
+        disp(['Seuil=' num2str(seuil{jj}(ll)) '; Data2use: '  num2str(jj) ])
+        disp(['Nbr New Instruments=' num2str(sum(isNewInstrument)) '; Nbr missing Instrument=' num2str(sum(~isNewInstrument))])
+        
+        data=[];
+        objects=[];
+        for pp=1:length(ind)
+            data.objects(pp).sounds=sounds2use(ind(pp));
+            data.objects(pp).path=path2use(ind(pp));
+            data.objects(pp).instrument=instrument2use(ind(pp));
+            data.objects(pp).mode=mode2use(ind(pp));
+        end
+        
+        data.instrument_missing=oldIntrument(~isNewInstrument);
+        data.instrument_used=newInstrument;
+        
+        if length(data2use{jj})==1
+            savejson('',data,['export/ticelxp1_subject_' num2str(data2use{jj}) '_tresh_' num2str(seuil{jj}(ll)) '.json']);
+        else
+            savejson('',data,['export/ticelxp1_subject_mix_tresh_' num2str(seuil{jj}(ll)) '.json' ]);
+        end
+        disp('')
+    end
+end

### 2018-01-29 code/xp1/data/userIds.js

@@ -1,0 +1,1 @@
+[{"userEmail":"null","userId":"null"},{"userEmail":"gregoirelafay@sfr.fr","userId":"gregoirelafaysfrfr"},{"userEmail":"yanachats@gmail.com","userId":"yanachatsgmailcom"},{"userEmail":"menuet-tomtom@wanadoo.fr","userId":"menuettomtomwanadoofr"},{"userEmail":"dgmcohen@free.fr","userId":"dgmcohenfreefr"},{"userEmail":"philipp.mayer@stud.hmtm.de","userId":"philippmayerstudhmtmde"},{"userEmail":"pbrandeis@cnsmdp.fr","userId":"pbrandeiscnsmdpfr"},{"userEmail":"farnazmodarresifar@gmail.com","userId":"farnazmodarresifargmailcom"},{"userEmail":"luis.naon@bbox.fr","userId":"luisnaonbboxfr"},{"userEmail":"luisgquintana@gmail.com","userId":"luisgquintanagmailcom"},{"userEmail":"contact@fredericdurieux.com","userId":"contactfredericdurieuxcom"}]

### 2018-01-29 code/xp1/extractData.m

@@ -1,0 +1,33 @@
+clear all
+addpath(genpath('functions'));
+
+
+files=dir('results/*js');
+rating=[];
+instrument={};
+mode={};
+name={};
+path={};
+sounds={};
+
+keep = 1:length(files);
+keep = [1 9];
+
+for jj=1:length(keep)
+    name{jj}=files(jj).name;
+    disp(files(keep(jj)).name);
+    xx=1;
+    results=loadjson(['results/' files(keep(jj)).name]);
+    for ll=1:length(results)
+        for nn=1:length(results{ll}.modes)
+            rating(xx,jj)=results{ll}.modes{nn}.rating;
+            instrument{xx,jj}=results{ll}.name;
+            mode{xx,jj}=results{ll}.modes{nn}.name;
+            sounds{xx,jj}=results{ll}.modes{nn}.sounds;
+            path{xx,jj}=results{ll}.modes{nn}.path;
+            xx=xx+1;
+        end
+    end
+end
+
+save('extractedData/results','rating','instrument','mode','name','sounds','path')

### 2018-01-29 code/xp1/statData.m

@@ -1,0 +1,15 @@
+clear all
+
+load('extractedData/results.mat')
+
+a = histcounts(rating(:, 1), 1:7)
+b = histcounts(rating(:, 2), 1:7)
+
+bar([a; b]')
+xlabel('Rating')
+set(gca, 'fontsize', 14)
+
+saveas(gcf, 'statxp1.png');
+saveas(gcf, 'statxp1.fig');
+
+[h,p,ci,stats] = ttest2(rating(:, 1), rating(:, 2))

### 2018-01-30 paper/lagrangeSpontaneousSimilarity2017.tex
learnt -> learned (AmE spelling)
whitespace
fix a typo in a comment of mine

@@
 \vl{We should probably say a word on the Hornbostel-Sachs taxonomy,
-which is based on player-instrument interaction, but is not outdated
+which is based on player-instrument interaction, but is now outdated
 given the variety of contemporary music techniques. @biblio}
@@
 Once the features and the projection matrix are computed, the similarity computation is quite effective and can be further enhanced by considering fast similarity search scheme such as locally sensitive least square hashing \cite{pauleve2010locality}.
-The secondary stage being learnt, it can easily adapt to user tastes and usage history by re-weighting the features for improved search relevance.
+The secondary stage being learned, it can easily adapt to user tastes and usage history by re-weighting the features for improved search relevance.
 

### 2018-02-13 paper/lagrangeSpontaneousSimilarity2017.tex
réponse
better deal with consensus clustering and fix few other things

@@
 \center
-\includegraphics[width = \textwidth]{figures/xp2example.png}
+\includegraphics[width = \textwidth]{figures/xp2example1.png}
 \caption{Spatial and group organization provided by subject 1.
@@
 @data @figure}}
-\ml{ok, tu peux utiliser le site de demo et faire une capture d'écran de l'utilisateur qui te va. Pour les autres display, je ne suis pas sur que cela soit utile.}
+\ml{ok, le nombre de cluster median est 10, les figures xp2example1 2 et 3 sont valides. La premiere est la plus propre. Pour les autres display, je ne suis pas sur que cela soit utile.}
 \label{fig:xp2display}
@@
 
-\subsection{Metric learning}
+\subsection{Metric learning} \label{sec:consensus}
 
@@
 One approach is to find a consensus clustering that maximally represent the 31 reference clusterings.
-To implement this solution, the chosen consensus clustering is taken from 3 solutions issued from 3 different algorithmic approaches introduced in \cite{strehl2002cluster}.
-This consensus clusterings is chosen among the 3 solutions as the one that has the best averaged NMI when compared to the 31 reference clusterings.
+To implement this solution, the hypergraph partitionning approach is considered \cite{kernighan1970efficient}.
+It approaches the concensus clustering problem as by re-partitionning the data
+using the given clusters as indications of strong bonds. The cluster ensemble problem is then
+formulated as a partitioning of the hypergraph that represents the sets of clusterings by cutting a minimal number of hyperedges \cite{strehl2002cluster}.
+
+Ah hypergraph is a generalization of a graph where vertices are connected by hyperedges that can connect not only two by any sets of vertices.
+In our case, \ipts are the vertices and the clusters of the reference partitions are the hyperedges.
+
+Computation of the resulting clustering shown on Figure \ref{fig:projectionC} is done using the hypergraph partitioning package HMETIS \cite{han1997scalable} using parameters suggested in \cite{strehl2002cluster}.
 
@@
 @writing}
-\ml{ok, je refais un code sur cette partie. Le dendrogram n'est pas un clustering, il part de la matrice de similarité. Bon apres, c'est tres lié...}
+\ml{Le dendrogram n'est pas un clustering, il part de la matrice de similarité.
+Bon apres, c'est tres lié...On ne peut rejeter l'hypothese que les trois solutions sont equivalentes, du coup pour simplifier je ne presente que l'approche classique.}
 
@@
 \ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+\vl{Les donn\'{e} de l'\{e}tude, relatives aux Figures 5 et 6.\}
+
+\section{Mds projections}
+
+In order to better interpret the data gathered during experiment 2, a multidimensional scaling projection in two dimensions is done using the average similarity matrix computed over the 31 reference clusterings, see \ref{sec:analysis}.
+The projection is computed  using a nonmetric multidimensional scaling where the Euclidean distances between points  approximate a monotonic transformation of the corresponding similarities using the Kruskal's normalized stress criterion \cite{kruskal1978multidimensional}.
+Figures \ref{fig:projectionI}, \ref{fig:projectionM}, and \ref{fig:projectionC} consider the same projection with different labels, respectively the names of the instruments of each \ipt, the name of the playing technique or the labels of consensus clustering, see Section \ref{sec:consensus} for the latter.The projections linked to the sounds used as stimuli are available  online \url{}.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/AllClassInstruments33}
+\caption{Mds projection with instruments as labels.}
+\label{fig:projectionI}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/AllClassModes33}
+\caption{Mds projection with playing techniques as labels.}
+\label{fig:projectionM}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/AllClassConsensus33}
+\caption{Mds projection with consensus clustering as labels.}
+\label{fig:projectionC}
+\end{figure}
 

### 2018-02-13 code/clusterEnsemble/ClusterEnsemble-V2.0/ceevalmutual.m

@@ -3,7 +3,7 @@
 % copyright (c) 1998-2011 by Alexander Strehl
 
 
-function m = ceevalmutual(cls,cl),
+function [m, q] = ceevalmutual(cls,cl),
 
 m = 0;
 totinds = 0;

### 2018-02-13 code/clusterEnsemble/ClusterEnsemble-V2.0/clusterensemble.m

@@ -53,10 +53,10 @@ for i = 1:length(workfcts);
    else
       cl(i,:) = feval(workfct,cls,k);
    end;
-   q(i) = ceevalmutual(cls,cl(i,:));
+   [q(i) qq(i, :)] = ceevalmutual(cls,cl(i,:));
    disp(['clusterensemble: ' workfct ' at ' num2str(q(i))]);
 end;
 
 [qual, best] = max(q);
 
-cl = cl(best,:);
+cl = cl(2,:);

### 2018-02-13 code/clusterEnsemble/combineJudgments.m

@@ -1,0 +1,10 @@
+load ../ticelJudgments
+addpath(genpath('.'))
+
+PATH = getenv('PATH');
+setenv('PATH', [PATH ':./ClusterEnsemble-V2.0/']);
+
+ensemble = clusterensemble(ci);
+
+
+save('../ticelJudgments.mat', 'ensemble','-append')

### 2018-02-13 code/xp2/clusterAnalysis.m

@@ -15,8 +15,8 @@ set(gca, 'xtick', [], 'fontSize', 7)
 set(h,'Units','Inches');
 pos = get(h,'Position');
 set(h,'PaperPositionMode','Auto','PaperUnits','Inches','PaperSize',[pos(3), pos(4)])
-saveas(gcf, '../../paper/figures/dendrogram', 'pdf')
-saveas(gcf, '../../paper/figures/dendrogram', 'fig')
+% saveas(gcf, '../../paper/figures/dendrogram', 'pdf')
+% saveas(gcf, '../../paper/figures/dendrogram', 'fig')
 
 
 for k=2:15

### 2018-02-13 code/xp2/groupAnalysis.m

@@ -1,4 +1,6 @@
-function groupAnalysis()
+function groupAnalysis(useConsensus)
+
+if ~exist('useConsensus', 'var'), useConsensus = 1; end
 
 modes = 0;
 load clusters
@@ -26,6 +28,10 @@ for k=1:ncl
 end
 ptx(1)=[];
 
+if (useConsensus),
+    load ticelJudgments
+   smcl = ensemble; 
+end
 
 clusterNames{end} = [clusterNames{end} 'art'];
 clusterNames{2} = [clusterNames{2} 'key'];
@@ -57,8 +63,8 @@ colorbar('Ticks', (1:length(modeClasses))+.5, 'TickLabels', modeClasses, 'TickLe
 set(gca, 'ytick', 1:2)
 set(gca, 'yticklabel', {'Judg.', 'Ref.'});
 % ytickangle(90)
-saveas(gcf, '../paper/figures/groupModes', 'png')
-savefig(gcf, '../paper/figures/groupModes')
+% saveas(gcf, '../paper/figures/groupModes', 'png')
+% savefig(gcf, '../paper/figures/groupModes')
 
 %% instruments
 
@@ -88,6 +94,8 @@ pty(1:2:end) = pty(1:2:end)+.05;
 % clusterNames{end} = [clusterNames{end} 'art'];
 % clusterNames{2} = [clusterNames{2} 'key'];
 
+if (useConsensus), smcl = ensemble; end
+
 figure(2)
 vec = [smcl; ii(icl)'];
 cm = jet(max(vec(:)));
@@ -112,5 +120,5 @@ set(gca, 'ytick', 1:2)
 set(gca, 'yticklabel', {'Judg.', 'Ref.'});
 % ytickangle(90)
 % title('Instruments')
-saveas(gcf, '../paper/figures/groupInstruments', 'png')
-savefig(gcf, '../paper/figures/groupInstruments')
+% saveas(gcf, '../paper/figures/groupInstruments', 'png')
+% savefig(gcf, '../paper/figures/groupInstruments')

### 2018-02-13 code/xp2/judgmentsDistribution.m

@@ -1,7 +1,7 @@
 
 load('judgments.mat')
 
-figPath = '../paper/figures/';
+figPath = '../../paper/figures/';
 
 nbc=[];
 sic=[];
@@ -12,9 +12,9 @@ for k=1:size(ci, 1)
 end
 
 disp('number of groups : ')
-disp([mean(nbc) std(nbc)])
+disp([mean(nbc) std(nbc) median(nbc)])
 disp('size of groups : ')
-disp([mean(sic) std(sic)])
+disp([mean(sic) std(sic) median(sic)])
 
 hist(nbc, 1:20)
 axis([1 20 0 5])

### 2018-02-13 code/xp2/ticelSimilarity.m

@@ -100,17 +100,20 @@ for k=1:size(x, 1)
     scatter(x(k, :), y(k, :), 100, ci(k, :), 'filled');
     axis off
     saveas(gcf, ['figures/' num2str(k)], 'png')
+    if (length(unique(ci(k, :))) == 10)
+        disp(k)
+    end
 end
 
 
-for k=1:size(x, 1)
-    imagesc(squareform(dc(k, :)))
-    colormap gray
-    set(gca, 'xtick', 0)
-set(gca, 'ytick', 1:length(names))
-set(gca, 'yticklabel', names)
-    saveas(gcf, ['figures/' num2str(k) 'Sim'], 'png')
-end
+% for k=1:size(x, 1)
+%     imagesc(squareform(dc(k, :)))
+%     colormap gray
+%     set(gca, 'xtick', 0)
+% set(gca, 'ytick', 1:length(names))
+% set(gca, 'yticklabel', names)
+%     saveas(gcf, ['figures/' num2str(k) 'Sim'], 'png')
+% end
 
 
 
@@ -153,6 +156,10 @@ savejson('', dat, 'figures/allInstruments.json');
 dat.color = ii;
 savejson('', dat, 'figures/allInstrumentsRef.json');
 
+ load ticelJudgments
+dat.color = ensemble;
+savejson('', dat, 'figures/allConsensus.json');
+
 function pm = displayData(ds, type, instruments, modes, source, nb)
 
 pm = mdscale(squareform(ds), 2);
@@ -185,6 +192,23 @@ axis off
 
 title([source ' ' type ' Modes (' num2str(nb)  ')'])
 saveas(gcf, ['figures/' source type 'Modes' num2str(nb)], 'png')
+
+load ticelJudgments
+
+figure(1)
+[cClasses, ~, bi] = unique(ensemble);
+colormap(jet(length(cClasses)))
+scatter(pm(:, 1), pm(:, 2), 100, bi, 'filled')
+caxis([1 16])
+colorbar('Ticks',.5+(1:length(cClasses)),...
+         'TickLabels',cClasses, 'TickLength', 0)
+     
+% dx = 0.01; dy = 0.01; % displacement so the text does not overlay the data points
+% text(pm(:, 1)+dx, pm(:, 2)+dy,  names);
+axis off
+
+title([source ' ' type ' Consensus (' num2str(nb)  ')'])
+saveas(gcf, ['figures/' source type 'Consensus' num2str(nb)], 'png')
 
 
 

### 2018-02-13 experiments/timbralSimilaritySol/clusterEnsemble/combineJudgments.m

@@ -1,27 +1,0 @@
-load ticelJudgments
-addpath(genpath('.'))
-
-ensemble = clusterensemble(ci);
-
-% generate reference algorithmically
-reference = 20;
-
-for kclus=1:size(ci, 1)
-    [accuracyv, classMatching] = accuracy(ci(reference,:), ci(kclus,:));
-    
-    labeltemp=ci(kclus,:);
-    labeltemp2=ci(kclus,:);
-    for kcm=1:size(classMatching,1)
-        n=classMatching(kcm,:);
-        vcm=find(n);
-        if(~isempty(vcm))
-            
-            labeltemp(labeltemp2==kcm)=vcm;
-        end
-    end
-    ci2(kclus,:)=labeltemp;
-    
-end
-medoid = mode(ci2);
-
-save('../ticelJudgments.mat', 'medoid', 'ensemble','-append')

### 2019-01-08 experiments/report/tex/timbralSimilaritySolPerceptualProjectionSlides.tex
better deal with consensus clustering and fix few other things
first results for joint with sct 25ms

--- 2017-12-14 experiments/timbralSimilaritySol/report/tex/exposeTmp.tex
@@
+ 
+\documentclass{beamer}            
+\usepackage{beamerthemedefault, multimedia}            
+\useoutertheme{smoothbars}            
+\useinnertheme[shadow=true]{rounded}            
+\setbeamercovered{transparent}            
+\setbeamertemplate{navigation symbols}{}            
+\setbeamertemplate{footline}[frame number]            
+\usepackage{graphicx}            
+\usepackage{morefloats}            
+\usepackage{amsmath}            
+\usepackage{amssymb}            
+\usepackage{rotating}            
+% mcode options for matlab code insertion bw (for printing), numbered (line numbers), framed (frame around code blocks), useliterate (convert Matlab expressions to Latex ones), autolinebreaks (automatic code wraping, use it with caution
+\usepackage[literate]{mcode}            
+\graphicspath{{figures/}{tex/}{../figures/}{../../}{../}}             
+\title{timbralSimilaritySolPerceptualProjectionSlides}            
+\author{ Mathieu Lagrange }
+
+\begin{document}
+
+\maketitle
+
+% Please use this file to document your experiment            
+% You can compile the report by setting the option 'report' as detailed in your expLanes configuration file.
+
+The first table compares the plain and joint scattering without any reprojection.           
+\begin{enumerate}           
+\item median: use median renormalization for scattering features           
+\item compress: use log compression for scattering features           
+\item standardize: standardize features at the end           
+\end{enumerate}           
+Use of all post processing give best performance. The second table show the results of using some projection. separateJudgment results can be seen as upper bound results.
+
+\begin{enumerate}           
+\item averageJudgment: use cluster ensemble techniques http://strehl.com/soft.html to obtain one clustering. thus only one projection is computed           
+\item separateJudgment: can be seen as an oracle method where each projection is evaluated using the clustering it has considered and average performance           
+\end{enumerate}
+
 \begin{frame}\frametitle{Factors flow graph}
@@
 
-\begin{frame}\frametitle{sct: 1000, split: none, reference: judgments, randomize: 0, expand: 0, cut: 1, median: 1, compress: 1, standardize: 1}
+\begin{frame}\frametitle{features: tfscat, sct: 25, projection: none, split: none, reference: judgments, randomize: 0}
 
@@
 \setlength{\tabcolsep}{.16667em}  
-\begin{tabular}{llllc} 
-features & projection & averageJudgment & separateJudgment & p (\%) \\ 
+\begin{tabular}{lllccccc}  
+median & compress & standardize & p (\%) &  &  &  & p (\%) \\  
 \hline  
-mfcc & lmnn & 0 & 0 &  86.31 $\pm$5.91 \\ 
-mfcc & lmnn & 0 & 1 &  86.18 $\pm$6.05 \\ 
-mfcc & lmnn & 1 & 0 &  86.22 $\pm$5.92 \\ 
-mfcc & none & 1 &  &  85.07 $\pm$6.19 \\ 
-mfcc & lda & 1 &  &  81.50 $\pm$7.65 \\ 
-scat & lmnn & 0 & 0 &  93.31 $\pm$3.92 \\ 
-scat & lmnn & 0 & 1 & \textbf{\textcolor{red}{ 98.09 $\pm$1.28}} \\ 
-scat & lmnn & 1 & 0 &  94.80 $\pm$3.26 \\ 
-scat & none & 1 &  &  87.01 $\pm$5.81 \\ 
-scat & lda & 1 &  & 80.95 $\pm$10.37 \\ 
+0 & 0 & 0 & 69.25 $\pm$9.67 &  &  &  & 74.80 $\pm$8.67 \\  
+0 & 0 & 1 & 76.65 $\pm$8.09 &  &  &  & 85.61 $\pm$5.56 \\  
+0 & 1 & 0 & 81.92 $\pm$7.03 &  &  &  & \textbf{89.49 $\pm$4.30} \\  
+0 & 1 & 1 & 81.49 $\pm$7.24 &  &  &  & \textbf{89.95 $\pm$4.08} \\  
+1 & 0 & 0 & 76.99 $\pm$8.14 &  &  &  & 86.43 $\pm$5.36 \\  
+1 & 0 & 1 & 76.65 $\pm$8.09 &  &  &  & 85.61 $\pm$5.56 \\  
+1 & 1 & 0 & \textbf{85.70 $\pm$6.30} &  &  &  & \textbf{91.08 $\pm$3.83} \\  
+1 & 1 & 1 & \textbf{\textcolor{red}{86.39 $\pm$6.13}} &  &  &  & \textbf{\textcolor{red}{91.12 $\pm$3.80}} \\  
 \end{tabular}  
 \end{center}  
-\label{sc1000SpnoRejuRa0Ex0Cu1Me1Co1St1} 
+\label{fetfSc25PrnoSpnoRejuRa0}  
 \end{table}
@@
 \end{frame}   
+\begin{frame}\frametitle{features: tfscat, sct: 25, split: none, reference: judgments, randomize: 0, median: 1, compress: 1, standardize: 1}
 
-\begin{frame}\frametitle{\small split: none, reference: judgments, randomize: 0, expand: 0, cut: 1, median: 1, compress: 1, standardize: 1} 
+\begin{table} 
 \begin{center} 
-\begin{figure} 
-\centering 
-\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{./figures/Fig160.pdf} 
-\label{spnoRejuRa0Ex0Cu1Me1Co1St1} 
-\end{figure} 
+\ 
+ \setlength{\tabcolsep}{.16667em} 
+\begin{tabular}{lllccccc} 
+pro & avejud & sepjud & p (\%) &  &  &  & p (\%) \\ 
+\hline 
+none & 1 & 0 & 86.39 $\pm$6.13 &  &  &  & 91.12 $\pm$3.80 \\ 
+lmnn & 0 & 0 & 91.59 $\pm$4.66 &  &  &  & \textbf{96.20 $\pm$2.27} \\ 
+lmnn & 0 & 1 & \textbf{\textcolor{red}{95.22 $\pm$3.23}} &  &  &  & \textbf{\textcolor{red}{96.89 $\pm$2.08}} \\ 
+lmnn & 1 & 0 & 93.19 $\pm$3.84 &  &  &  & 95.60 $\pm$2.32 \\ 
+lda & 1 & 0 & 82.40 $\pm$7.64 &  &  &  & 81.31 $\pm$8.51 \\ 
+lda & 1 & 1 & 82.40 $\pm$7.64 &  &  &  & 81.31 $\pm$8.51 \\ 
+\end{tabular} 
 \end{center} 
+\label{fetfSc25SpnoRejuRa0Me1Co1St1} 
+\end{table}
 
 \end{frame}  
+\input{tex/exposeTmp} % expLanesInsertionFlag DO NOT CLEAR (but move it where you want the generated temporary LaTEX code to be inserted)
+
+\bibliographystyle{abbrvnat}            
+\bibliography{bib}
+
+\end{document}

### 2019-01-08 experiments/report/timbralSimilaritySolPerceptualProjectionSlides.tex
first results for joint with sct 25ms

--- 2019-01-08 experiments/report/tex/timbralSimilaritySolPerceptualProjectionSlides.tex
@@
 
-\begin{frame}\frametitle{Factors flow graph}
-
-\begin{center}
-
-\begin{figure}
-
-\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{../figures/factors.pdf}
-
-\label{factorFlowGraph}
-
-\end{figure}
-
-\end{center}
-
-\end{frame}
-
-\begin{frame}\frametitle{features: tfscat, sct: 25, projection: none, split: none, reference: judgments, randomize: 0}
-
-\begin{table}  
-\begin{center}  
-\  
-\setlength{\tabcolsep}{.16667em}  
-\begin{tabular}{lllccccc}  
-median & compress & standardize & p (\%) &  &  &  & p (\%) \\  
-\hline  
-0 & 0 & 0 & 69.25 $\pm$9.67 &  &  &  & 74.80 $\pm$8.67 \\  
-0 & 0 & 1 & 76.65 $\pm$8.09 &  &  &  & 85.61 $\pm$5.56 \\  
-0 & 1 & 0 & 81.92 $\pm$7.03 &  &  &  & \textbf{89.49 $\pm$4.30} \\  
-0 & 1 & 1 & 81.49 $\pm$7.24 &  &  &  & \textbf{89.95 $\pm$4.08} \\  
-1 & 0 & 0 & 76.99 $\pm$8.14 &  &  &  & 86.43 $\pm$5.36 \\  
-1 & 0 & 1 & 76.65 $\pm$8.09 &  &  &  & 85.61 $\pm$5.56 \\  
-1 & 1 & 0 & \textbf{85.70 $\pm$6.30} &  &  &  & \textbf{91.08 $\pm$3.83} \\  
-1 & 1 & 1 & \textbf{\textcolor{red}{86.39 $\pm$6.13}} &  &  &  & \textbf{\textcolor{red}{91.12 $\pm$3.80}} \\  
-\end{tabular}  
-\end{center}  
-\label{fetfSc25PrnoSpnoRejuRa0}  
-\end{table}
-
-\end{frame}   
-\begin{frame}\frametitle{features: tfscat, sct: 25, split: none, reference: judgments, randomize: 0, median: 1, compress: 1, standardize: 1}
-
-\begin{table} 
-\begin{center} 
-\ 
- \setlength{\tabcolsep}{.16667em} 
-\begin{tabular}{lllccccc} 
-pro & avejud & sepjud & p (\%) &  &  &  & p (\%) \\ 
-\hline 
-none & 1 & 0 & 86.39 $\pm$6.13 &  &  &  & 91.12 $\pm$3.80 \\ 
-lmnn & 0 & 0 & 91.59 $\pm$4.66 &  &  &  & \textbf{96.20 $\pm$2.27} \\ 
-lmnn & 0 & 1 & \textbf{\textcolor{red}{95.22 $\pm$3.23}} &  &  &  & \textbf{\textcolor{red}{96.89 $\pm$2.08}} \\ 
-lmnn & 1 & 0 & 93.19 $\pm$3.84 &  &  &  & 95.60 $\pm$2.32 \\ 
-lda & 1 & 0 & 82.40 $\pm$7.64 &  &  &  & 81.31 $\pm$8.51 \\ 
-lda & 1 & 1 & 82.40 $\pm$7.64 &  &  &  & 81.31 $\pm$8.51 \\ 
-\end{tabular} 
-\end{center} 
-\label{fetfSc25SpnoRejuRa0Me1Co1St1} 
-\end{table}
-
-\end{frame}  
 \input{tex/exposeTmp} % expLanesInsertionFlag DO NOT CLEAR (but move it where you want the generated temporary LaTEX code to be inserted)

### 2019-01-08 experiments/handleJudgments.m

@@ -1,6 +1,7 @@
-function [data, judgments] = handleJudgments(config, data, average)
+function [data, judgments, idx] = handleJudgments(config, data, average, filter)
 
 if ~exist('average', 'var'), average = 0; end
+if ~exist('filter', 'var'), filter = 1; end
 
 if length(data.mode)<length(data.file), data.file(end) = []; end
 idx = [];
@@ -22,4 +23,7 @@ if average
 else
     judgments = gt(:, ia);
 end
+
+if (filter)
 data = filterData(data, idx);
+end

### 2019-01-08 experiments/tisiso1features.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso1features(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 1, 'mask', {4 1}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 1, 'mask', {4 2 2 1 0 5 1 0 0 0 0 2 1 1 0}); return; else store=[]; obs=[]; end
 
 % fid = fopen([config.inputPath 'fileList.txt']);
 % fileList = textscan(fid, '%s/%s\n');
@@ -61,20 +61,18 @@ store.family = family;
 store.file = fileList;
 
 scat_opt.M = 2;
-scat_opt.oversampling = 4;
-scat_opt.path_margin = 4;
+scat_opt.oversampling = 1;
+scat_opt.path_margin = 1;
 
-% store = expLoad(config, '', 1);
-% store.file = fileList;
-% return
+%[~, ~, idx] = handleJudgments(config, store, 0, 0);
 
 if strcmp(setting.features, 'null'), return; end
 
 failed = zeros(1, length(fileList));
-length(fileList)
-for k=1:length(fileList)
+done = zeros(1, length(fileList));
+parfor k=1:length(fileList)
+    cc = 1;
     [a,sr] = audioread([config.inputPath fileList{k} '.wav']);
-    cc = 0;
     switch setting.features
         case 'mel'
             [~, cc] = melfcc(a(:,1), sr, 'wintime', setting.sct/1000, 'hoptime', setting.sct/4000);
@@ -92,28 +90,29 @@ for k=1:length(fileList)
             tm_filt_opt = struct();
             tm_filt_opt.Q = [12 1];
             tm_filt_opt.J = T_to_J(sr*setting.sct/1000, tm_filt_opt);
-
+            
             % NOTE: The parameter `fr_filt_opt.J` controls the largest scale
             % along the frequency axis as a power of two. For `J = 4`, this
             % means a largest frequency scale of `2^4 = 16`, which is equal to
             % 1.5 octaves since `Q = 12`.
             fr_filt_opt = struct();
             fr_filt_opt.J = 4;
-
+            
             Wop = joint_tf_wavelet_factory_1d(size(a, 1), tm_filt_opt, ...
                 fr_filt_opt, scat_opt);
-
+            
             S = scat(a(:,1), Wop);
-
[…]

### 2019-01-08 experiments/tisiso2project.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso2project(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 2, 'mask', {2 1 2 1 1 5 1 1 2 0 0 2 1 2}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 2, 'mask', {4, 3, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1}); return; else store=[]; obs=[]; end
 
 rng(0);
 
@@ -32,7 +32,7 @@ switch setting.reference
     case 'judgments'
         [data, judgments] = handleJudgments(config, data, setting.averageJudgment);
         
-        parfor k=1:size(judgments, 1)
+        for k=1:size(judgments, 1)
             ju = judgments(k, :);
             projection(k, :, :) = process2project(store, config, data, setting, ju);
         end      
@@ -47,6 +47,8 @@ function projection = process2project(store, config, data, setting, ju)
 if exist('ju', 'var'), data.(setting.reference) = ju; end
 
 data = getFeatures(config, data, setting, config.step.id);
+
+data.features = real(data.features);
 
 [~, ~, gt] = unique(data.(setting.reference));
 

### 2019-01-08 experiments/tisiso3performance.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {2 1 2 1 0 5 1 0 2 2 2 2}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {4 1 1 1 0 5 1 0 0 2 2 1 1 2 1}); return; else store=[]; obs=[]; end
 
 rng(0);
 
@@ -65,6 +65,9 @@ switch setting.projection
         end
 end
 % features = features(:, 1);
+
+features = real(features);
+
 
 n=knnsearch(features,features,'k',setting.neighbors+1);
 n(:, 1) = [];

### 2019-01-08 experiments/tisisoReport.m

@@ -6,7 +6,7 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcvd', 'reportName', 'perceptualProjectionSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'perceptualProjectionSlides'); return; end
 
 switch config.reportName
     case 'debugSlides'
@@ -17,12 +17,22 @@ switch config.reportName
 %         config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'expand', 'sct', 'percent', 0, 'obs', 1, 'highlight', 0, 'uncertainty', -1);
 % 
 
-        mask = {[2  3], 0, [1  2], 1, 2, [5 6], 1, 1, 2, 2, 2, 2 1};
-        mask = {[2 3], 5, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
-        config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4);
-   
-        mask = {[2 3], 0, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
-        config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'expand', 'sct', 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'uncertainty', -1);
+        mask = {[2 3], 1, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
+      mask = {[3 4], 1, 0, 1, 2, 5, 1, 1, 2, 0, 0, 0, 1};
+      
+      mask = {[3], 1, 1, 1, 2, 5, 1, 1, 2, 0, 0, 0, 1};
+      config = expExpose(config, 'l', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'visible', 0);
+      mask{1} = 4;
+      config = expExpose(config, 'l', 'step', 3, 'mergeDisplay', 'h', 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'noFactor', 1);
+  
+      mask = {[3], 1, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
[…]

### 2019-01-11 experiments/tisiso1features.m
failing file with sct 1000

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso1features(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 1, 'mask', {4 2 2 1 0 5 1 0 0 0 0 2 1 1 0}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 1, 'mask', {4 5 2 1 0 5 1 0 0 0 0 2 1 1 0}); return; else store=[]; obs=[]; end
 
 % fid = fopen([config.inputPath 'fileList.txt']);
 % fileList = textscan(fid, '%s/%s\n');
@@ -70,7 +70,7 @@ if strcmp(setting.features, 'null'), return; end
 
 failed = zeros(1, length(fileList));
 done = zeros(1, length(fileList));
-parfor k=1:length(fileList)
+for k=1554 %1:length(fileList)
     cc = 1;
     [a,sr] = audioread([config.inputPath fileList{k} '.wav']);
     switch setting.features

### 2019-01-11 experiments/tisisoReport.m

@@ -6,10 +6,13 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'perceptualProjectionSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'debugSlides'); return; end
 
 switch config.reportName
     case 'debugSlides'
+        
+      mask = {  4, 5, [2  3], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
+      config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4);
         
     case 'perceptualProjectionSlides'
         

### 2019-01-11 experiments/bugs/bug.m
failing file with sct 1000

@@ -1,0 +1,23 @@
+
+ [a,sr] = audioread('TpC-slap-G3-p.wav');
+  setting.sct = 1000;
+  addpath(genpath('../libs/scatnet'))
+scat_opt.M = 2;
+scat_opt.oversampling = 1;
+scat_opt.path_margin = 1;
+
+tm_filt_opt = struct();
+tm_filt_opt.Q = [12 1];
+tm_filt_opt.J = T_to_J(sr*setting.sct/1000, tm_filt_opt);
+
+% NOTE: The parameter `fr_filt_opt.J` controls the largest scale
+% along the frequency axis as a power of two. For `J = 4`, this
+% means a largest frequency scale of `2^4 = 16`, which is equal to
+% 1.5 octaves since `Q = 12`.
+fr_filt_opt = struct();
+fr_filt_opt.J = 4;
+
+Wop = joint_tf_wavelet_factory_1d(size(a, 1), tm_filt_opt, ...
+    fr_filt_opt, scat_opt);
+
+S = scat(a(:,1), Wop);

### 2019-01-11 experiments/tisiso1features.m

@@ -70,7 +70,7 @@ if strcmp(setting.features, 'null'), return; end
 
 failed = zeros(1, length(fileList));
 done = zeros(1, length(fileList));
-for k=1554 %1:length(fileList)
+parfor k=1:length(fileList)
     cc = 1;
     [a,sr] = audioread([config.inputPath fileList{k} '.wav']);
     switch setting.features

### 2019-01-11 experiments/tisisoReport.m

@@ -11,8 +11,8 @@ if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'debugSlides')
 switch config.reportName
     case 'debugSlides'
         
-      mask = {  4, 5, [2  3], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
-      config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4);
+      mask = {  [3 4], [1 2], [1 2], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
+      config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'expand', 'sct');
         
     case 'perceptualProjectionSlides'
         

### 2019-02-28 experiments/getFeatures.m
karcher test

@@ -1,28 +1,32 @@
 function data = getFeatures(config, data, setting, step)
 
 if length(data.mode)<size(data.features, 1)
- data.features(end, :) = []; % TODO fix this
+    data.features(end, :) = []; % TODO fix this
 end
 if length(data.mode)<length(data.file)
- data.file(end) = []; % TODO fix this
+    data.file(end) = []; % TODO fix this
 end
- 
 
 
-switch setting.split
-    case 'none'
+switch setting.reference
+    case 'judgments'
         idx = 1:length(data.file);
-    case 'octave'
-        idx = octaveSelection(config, data);
     otherwise
-        nbItems = length(data.mode);
-        idx = randi(nbItems, ceil(nbItems*str2num(setting.split)/100), 1);
-        if step==2 && setting.test
-            idx = setdiff(1:nbItems, idx);
+        switch setting.split
+            case 'none'
+                idx = 1:length(data.file);
+            case 'octave'
+                idx = octaveSelection(config, data);
+            otherwise
+                nbItems = length(data.mode);
+                idx = randi(nbItems, ceil(nbItems*str2num(setting.split)/100), 1);
+                if step==2 && setting.test
+                    idx = setdiff(1:nbItems, idx);
+                end
         end
 end
 data = filterData(data, idx);
-% 
+%
 % switch setting.reference
 %     case 'judgments'
 %         idx = [];
@@ -53,7 +57,7 @@ switch setting.cut
                 [vars, I] = sort(var,2,'descend');
                 features=features(:,I);
                 sumvar=cumsum(vars);
-                 ix=find(sumvar>.83,1);
+                ix=find(sumvar>.83,1);
                 features=features(:,1:ix);
         end
 end
@@ -69,12 +73,12 @@ switch setting.compress
 end
 
 if ~isnan(setting.expand) && setting.expand
-     features = expandFeatures(features, setting.expand);
+    features = expandFeatures(features, setting.expand);
 end
 
 switch setting.randomize
     case 1
-     features = randn(size(features));
+        features = randn(size(features));
 end
 
 switch setting.standardize
@@ -105,13 +109,13 @@ function features = expandFeatures(features, n)
 ref = features;
 ex = features;
 while size(features, 2) < n
-   for k=1:size(ref, 2)
-       for m=1:size(ex, 2)
[…]

### 2019-02-28 experiments/libs/karcher_mean.m

@@ -1,0 +1,58 @@
+% KARCHER_MEAN Compute Karcher mean of subspaces
+%
+% Usage
+%    X = karcher_mean(A);
+%
+% Input
+%    A: An array of matrices of size n-by-p-by-k, where n is the dimension of
+%       the ambient space, p is the dimension of the subspace, and k is the
+%       number of subspaces. The matrix A(:,:,ell) is an orthonormal matrix
+%       which constitutes a basis for the ellth subspace.
+%
+% Output
+%    X: An orthonormal matrix of size n-by-p whose columns form a basis for a
+%       Karcher mean of the given subspaces. This is a subspace that locally
+%       minimizes
+%
+%          sum_{ell=1}^k dist(X, A_ell)^2 ,
+%
+%       where A_ell is the subspace spanned by A(:,:,ell) and dist(X, Y)^2 is
+%       the sum of the squared principal angles between X and Y, that is
+%
+%          dist(X, Y)^2 = sum(acos(svd(X'*Y)).^2) .
+%
+%       The output X is obtained by minimizing this objective starting at
+%       A(:,:,1).
+
+function X = karcher_mean(A)
+    n = size(A, 1);
+    p = size(A, 2);
+    k = size(A, 3);
+
+    M = grassmannfactory(n, p);
+
+    problem = struct();
+    problem.M = M;
+    problem.cost = @(X)(cost(M, A, X));
+    problem.grad = @(X)(grad(M, A, X));
+    problem.hess = approxhessianFD(problem);
+
+    options = struct();
+    options.verbosity = 0;
+
+    X = trustregions(problem, A(:,:,1), options);
+end
+
+function f = cost(M, A, X)
+    f = 0;
+    for ell = 1:size(A, 3)
+        f = f + 1/2*M.dist(X, A(:,:,ell))^2;
+    end
+end
+
+function u = grad(M, A, X)
+    u = M.zerovec();
+    for ell = 1:size(A, 3)
+        u = M.lincomb(X, 1, u, -1, M.log(X, A(:,:,ell)));
+    end
+end

### 2019-02-28 experiments/libs/karcher_mean_ex.m

@@ -1,0 +1,21 @@
+addpath(genpath('manopt'))
+
+% Set parameters n (dimension of ambient subspace) and p (dimension of subspace).
+n = 8;
+p = 3;
+k = 2;
+
+% Generate random subspaces.
+A = zeros(n, p, k);
+for ell = 1:k
+    [A(:,:,ell), ~] = qr(randn(n, p), 0);
+end
+
+% Find the mean subspace.
+X = karcher_mean(A);
+
+% Find distances of mean to subspaces. For k = 2, X should be the same distance from both subspaces.
+dists = zeros(k, 1);
+for ell = 1:k
+    dists(ell) = sqrt(sum(acos(svd(X'*A(:,:,ell))).^2));
+end

### 2019-02-28 experiments/tisiso1features.m

@@ -73,6 +73,8 @@ done = zeros(1, length(fileList));
 parfor k=1:length(fileList)
     cc = 1;
     [a,sr] = audioread([config.inputPath fileList{k} '.wav']);
+    a = pad_signal(a, max(size(a, 1), sr*setting.sct/1000), 'zero', true);
+    
     switch setting.features
         case 'mel'
             [~, cc] = melfcc(a(:,1), sr, 'wintime', setting.sct/1000, 'hoptime', setting.sct/4000);

### 2019-02-28 experiments/tisiso2project.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso2project(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 2, 'mask', {4, 3, 0, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 2, 'mask', {2 3 3 5 0 5 1 1 2 0 0 2 1 2 0}); return; else store=[]; obs=[]; end
 
 rng(0);
 
@@ -23,7 +23,7 @@ rng(0);
 %     store.projection = projection;
 % end
 % return
-if strcmp(setting.projection, 'none') || (strcmp(setting.projection, 'lda') && strcmp(setting.reference, 'judgments') && setting.averageJudgment==0)
+if strcmp(setting.projection, 'none') || (strcmp(setting.projection, 'lda') && strcmp(setting.reference, 'judgments') && setting.averageJudgment==0) || (strcmp(setting.reference, 'judgments') && strcmp(setting.split, 'octave'))
     store.projection = [];
     return
 end
@@ -31,7 +31,7 @@ end
 switch setting.reference
     case 'judgments'
         [data, judgments] = handleJudgments(config, data, setting.averageJudgment);
-        
+        [data, judgments] = splitJudgments(data, judgments, str2num(setting.split));
         for k=1:size(judgments, 1)
             ju = judgments(k, :);
             projection(k, :, :) = process2project(store, config, data, setting, ju);

### 2019-02-28 experiments/tisiso3performance.m

@@ -11,13 +11,16 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {4 1 1 1 0 5 1 0 0 2 2 1 1 2 1}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 1 2 1 0 5 1 0 0 2 2 2 1 1 3}); return; else store=[]; obs=[]; end
 
 rng(0);
 
-if ((strcmp(setting.projection, 'lda') || strcmp(setting.projection, 'none')) && strcmp(setting.reference, 'judgments') && setting.averageJudgment==0) || ...
-   (strcmp(setting.projection, 'lmnn') && strcmp(setting.reference, 'judgments') && setting.averageJudgment==1 && setting.separateJudgment==1) || ...
-   (strcmp(setting.projection, 'none') && strcmp(setting.reference, 'judgments') && setting.separateJudgment==1)
+if strcmp(setting.reference, 'judgments') && (...
+        (strcmp(setting.projection, 'lda')  && (setting.averageJudgment==0 || (setting.averageJudgment==1 && setting.separateJudgment==0))) || ...
+        (strcmp(setting.projection, 'none')  && (setting.separateJudgment==1 || (setting.averageJudgment==0 && setting.separateJudgment==1))) || ...
+        (strcmp(setting.projection, 'lmnn')  && setting.averageJudgment==1 && setting.separateJudgment==1) || ...
+        strcmp(setting.split, 'octave') ...
+        )
     obs.p=NaN;
     return
 end
@@ -26,8 +29,9 @@ data1 = expLoad(config, '', 1);
 obs.p=[];
 switch setting.reference
     case 'judgments'
-        [data1, judgments] = handleJudgments(config, data1);              
-        parfor k=1:size(judgments, 1)  % par   
+        [data1, judgments] = handleJudgments(config, data1);
+        [data1, judgments] = splitJudgments(data1, judgments, str2num(setting.split), setting.test);
+        parfor k=1:size(judgments, 1)  % par
             p(k) = process3performance(config, data1, data, setting, judgments(k, :), k);
         end
         obs.p = p;
@@ -49,13 +53,22 @@ switch setting.projection
     case 'lmnn'
         switch setting.reference
             case 'judgments'
-                if setting.separateJudgment
[…]

### 2019-02-28 experiments/tisisoReport.m

@@ -6,14 +6,34 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'debugSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'karcherSlides'); return; end
 
 switch config.reportName
-    case 'debugSlides'
+    case 'karcherSlides'
+        mask = {2:4, 1, 2, 1, 0, 5, 1, 1, 2, 2, 2, 2, 1, 1};
+     config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 3, 'shortFactors', 1, 'uncertainty', -1);
+   case 'completeSlides'
         
-      mask = {  [3 4], [1 2], [1 2], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
-      config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'expand', 'sct');
-        
+      mask = {[3 4], [0], [1 2], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
+      config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'expand', 'sct', 'uncertainty', -1);
+     mask = {[2 3 4], [0], [0], 1, 0, 5, 1, 1, 2, 2, 2, 2, 1};
+     config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 3, 'expand', 'sct', 'shortFactors', 1, 'uncertainty', -1);
+     
+    case 'splitSlides'
+  
+        %none
+            mask = {[2  3  4], 1, 1, [2  3  4], 0, 5, 1, 1, 2, 2, 2, 2, 1, 1, 1};
+     % config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 3, 'uncertainty', -1, 'expand', 'test');
+
+        % lmnn
[…]

### 2019-04-23 paper/paper.tex
Start new paper outline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% HEADER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{amsmath,cite,url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{xspace}

\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\newcommand{\ipt}{IPT\xspace}
\newcommand{\ipts}{IPTs\xspace}

\newcommand{\ml}[1]{\textcolor{blue}{ML: #1}}
\newcommand{\vl}[1]{\textcolor{red}{VL: #1}}
\newcommand{\ja}[1]{\textcolor{purple}{JA: #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Learning Auditory Similarities Between Instrumental Playing Techniques}

\author{
Christian El-Hajj,
Vincent Lostanlen,
Mathias Rossignol, \\
Gr\'egoire Lafay,
and Mathieu Lagrange}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
ABSTRACT HERE
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}

\begin{itemize}
\item How to accurately model timbre perception?
\item Studies either consider stimuli that are ``too simple'' (e.g., tones, instruments with fixed playing techniques) or ``too complex'' (e.g., speech, environmental sounds).
\item Computational models do not capture enough information (MFCCs and related representations) or are hard to compute or analyze (STRFs, deep networks).
\item Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (instrumental playing techniques, or IPTs). Provides more control in modeling different aspects of timbre.
\item In addition, propose a computational model for timbral similarity: (joint) scattering plus a linear layer. Fixed representation to extract time-frequency structure with adaptivity in linear component.
\end{itemize}

\section{Results}
\label{sec:results}

\begin{itemize}
\item Subjects asked to cluster 78 IPTs into an arbitrary number of clusters (maximum 20). Resulting partition gives measure of timbral similarity.
\item Partitions do not completely agree with either instrument (I) of playing technique (PT) taxonomies, although slightly favor PTs. Perceptual measurement captures timbre across IPTs. Show through NMI or other measurements.
\item Propose computational model for these perceptual similarity measurements: scattering transform (time or time-frequency), followed by a learned linear layer (LDA or LDMM) to adapt representation to measured data.
\item Model shows good agreement with similarity measurements under different testing protocols (aggregated clusters, aggregated subspaces, cross validation).
\end{itemize}

\section{Discussion}
\label{sec:discussion}

\begin{itemize}
\item Demonstrates complex timbral structure perceived by subjects in study: neither instrument (i.e., spectral envelope) or playing technique (i.e., temporal modulation) dominates similarity judgment. Need joint to characterize time-frequency structure (see Patil et al).
\item Proposed computational model (scattering + linear) reproduces similarity judgments accurately. Easy to ``retrain'' for other data. Fixed wavelet structure allows for analysis of linear layer and guarantees invariance and stability properties.
\item Applications: query-by-example in musical sample databases.
\end{itemize}

\section{Methods}
\label{sec:methods}

\begin{itemize}
\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities. The result is a convolutional network with fixed filters. Performs well in various classification and regression tasks when augmented with a learned linear layer.
\item Linear layer is provided here by LDA or LMNN.
\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{alpha}
\bibliography{bib}

\end{document}

### 2019-09-06 paper/paper.tex
First draft of introduction

@@
 
+\textbf{Section outline:}
 \begin{itemize}
@@
 \end{itemize}
+
+One of the most important properties for identifying the source of a sound is its \emph{timbre}.
+While it is typically defined as the quality of a sound that is independent of its pitch, duration, or loudness, this definition fails to provide an explicit characterization of the attribute.
+Due to its central role in audio perception, timbre has been studied from several angles, including perceptual and psychophysical studies as well as computational methods in machine listening.
+The consensus is that timbre depends not on the purely spectral or temporal structure of a sound, but in its joint spectro--temporal structure, but beyond this, there is little agreement about its precise characterization.
+
+One difficulty that arises in studying the perceptual dimensions of timbre relates to its potential complexity---the dimensionality of the full timbral space is high.
+To combat this, many perceptual and psychophysical studies focus on simple stimuli: pure sine waves, musical instruments with fixed pitch and playing technique, and so on.
+While this has led to important insights in timbre perception, the range of timbres explored by these stimuli is necessarily limited.
+
+On the other end of the spectrum are the computational methods.
+Since their goals are typically task-oriented (recognizing speech, identifying environmental sounds, and so on), they are less concerned with establishing models of auditory perception suitable for analysis.
+While certain audio representations are simple enough to analyze, such as mel-frequency cepstral coefficients (MFCCs) and related approaches, they do not capture enough structure to adequately characterize timbre.
+Indeed, for an overwhelming majority of machine listening tasks have been these representations have been outperformed by learned representations, such as deep neural networks (DNNs).
+While DNNs have enjoyed significant success for these tasks, their relative complexity and the fact that they are learned from data make their analysis difficult, preventing us from deriving relevant insights into timbre perception.
+
+Certain audio representations have been proposed as a middle ground between expressivity and ease of analysis.
+One of the more prominent among these are the \emph{cortical representations} introduced by S. Shamma and collaborators.
+These rely on a spectrogram-like decomposition of the sound, which is then further decomposed in time and frequency using \emph{spectro--temporal receptive fields} (STRFs).
+This model is based on neurophysiological measurements on ferrets, and have been validated in several contexts, ranging from applications in machine listening and reproduction of psychophysical phenomena in auditory perception.
+However, since the original goal of the representation was to model observed electrophysiological signals, this increases its complexity, rendering mathematical anaylsis more difficult.
+
+In particular, cortical representations have been used to study timbre in the work of Patil et al. \cite{patil2012music}.
+While this demonstrates the power of the representation to characterize certain dimensions of timbre, the range of sounds was limited, consisting of single musical notes played using standard instrumental techniques.
+As a result, only a subset of the relevant timbral dimensions are explored, limiting the conclusions of the work.
+
+In this work, we consider representation similar to the cortical representation, namely the \emph{scattering transform}, which provides a mathematical foundation for the study of auditory perception in terms of invariance and stability properties.
+Instead of single musical notes played with standard techniques, we extend the range of sounds to include extended instrumental playing techniques.
+While increasing the complexity of the timbral space explored, these sounds do not include the full range of timbre exhibited, for example, by speech or environmental sounds.
+As a result, we are able to better analyze their characteristics and link them to the properties of the transform.
 

### 2019-09-09 paper/paper.tex
Write scattering subsection

@@
 \newcommand{\ipts}{IPTs\xspace}
+
+\newcommand{\patk}{p\mathrm{@}k}
 
@@
 \end{itemize}
+
+\subsection{Scattering transform}
+\label{sec:scattering}
+
+While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
+As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
+
+A closely related representation is that of the joint time--frequency scattering transform, introduced by And\'{e}n et al. \cite{anden2015joint,anden2019joint}.
+Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
+The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
+This representation contains a large amount of information, but not in a form suitable for building statistical models.
+In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may incude large changes to the scalogram.
+One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients \cite{anden2014deep}.
+
+These first-order coefficients alone cannot accurately characterize the time--frequency structure of the signal, however, since they only include spectral information.
+To augment them, we perform another wavelet decomposition, this time in two dimensions, over the scalogram.
+Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
+Again, we compute the modulus of the transform and average in time.
+The result is known as the set of second-order joint time--frequency scattering coefficients \cite{anden2015joint,anden2019joint}.
+
+The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
+On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
+The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
+Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids \cite{anden2012scattering,anden2019joint}, and beating phenomena \cite{anden2014deep}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019joint,lostanlen2018relevance,lostanlen2018extended}.
+
+To evaluate the ability of the scattering transform to characterize the timbre of an audio signal, we consider the task of classifying musical notes from a wide variety of instruments played according to a set of extended instrumental playing techniques (IPTs).
+Increasing the variability in IPTs allows us to explore a wider range of timbre, providing a more robust validation of the proposed model for modeling timbre.
+The metric used is the \emph{precision at rank $k$} ($\patk$), which is calculated by computing the $k$ nearest neighbors in the feature space and recording the proportion of the neighbors that are in the same ``class.''
+In the following, we set $k = 5$.
+
+Concatenating the first- and second-order joint scattering coefficients, we obtain a feature vector of dimension $???$ for a sound recording of duration $???$.
+Computing nearest neighbors with the standard Euclidean measure gives a $\patk$ of $87.0 \pm 5.8$, where the first value indicates the mean and the second the standard deviation.
+As a result, this simple transformation of the original signal results in a representation that lends itself naturally to discriminating timbre.
+Note that this result is obtained by simply looking at the raw feature vectors---no learning is performed to adapt the transform towards discriminating timbre.
+
+\subsection{Similarity weighting}
+\label{sec:weighting}
 

### 2019-09-09 paper/paper.tex
Start writing weighting subsection

@@
 
+While the raw scattering representation provides a powerful model for timbre, it does so without any prior information.
+Incorporating information about what distinguishes one timbre from another should improve the accuracy of the model.
+In the same was that the lower levels of auditory perception do not vary much between subjects (or between species, in the case of mammals), we expect that higher levels incorporate some amount of learning.
+Introducing a learned component into our model, however, could greatly reduce its usefulness with respect to interpretability.
+For this reason, we restrict ourselves to a simple, linear transform of the scattering coefficients.
+By optimizing the linear transform with respect to similarity measurements obtained from human subjects, we expect to obtain a more accurate model of timbre perception.
+
+These human similarity judgments are obtained from a set of $78$ instrument--IPT pairs selected by two music composition professors at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+A subject was provided with these 78 recordings and asked to cluster them in a free sorting task by arranging correspodings dots in two dimensions on a computer screen and assigning a color to each dot.
+While the location of the dots was recorded, this information was not used in our analysis and only served to help the subject organize the sounds spatially.
+The colors assigned to each dot---and, consequently, to each sound---was recorded for each subject.
+
+The above similarity judgments were crowdsourced by sending out requests to perform the experiment on internal mailing lists of the CNSMDP and international research mailing lists with focus on audio and music processing.
+We made the experiment avaialble for two months and obtained results for a total of $31$ subjects.
+
 \section{Discussion}

### 2019-09-10 paper/paper.tex
Add results on raw MFCCs

@@
 Note that this result is obtained by simply looking at the raw feature vectors---no learning is performed to adapt the transform towards discriminating timbre.
+We contrast this with the results obtained with a baseline MFCC feature vector.
+This results in a $\patk$ of $85.1 \pm 6.2$.
+We therefore conclude that, although the scattering transform provides a good indication of timbral similarity, using its nearest neighbors based on a standard Euclidan distance does not allow us to properly account for the higher-level time--frequency structure characterized by the second-order coefficients.
 

### 2019-09-10 paper/paper.tex
Fix spelling
Finish weighting section

@@
 The above similarity judgments were crowdsourced by sending out requests to perform the experiment on internal mailing lists of the CNSMDP and international research mailing lists with focus on audio and music processing.
-We made the experiment avaialble for two months and obtained results for a total of $31$ subjects.
+We made the experiment available for two months and obtained results for a total of $31$ subjects.
+
+Given the clustering assignments of each subject, we then constructed a consensus clustering using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
+Interestingly, although several sounds pairs were obtained from the same instrument (but different IPTs) and others were obtained using the same IPT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
+For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
+This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
+Running LMNN on MFCCs with the consensus clustering obtained above increases $\patk$ slightly to $86.2 \pm 5.9$.
+Consequently, there seems to be little additional timbre information in the MFCCs compared to that given by the standard Euclidean distance.
+We contrast this with LMNN applied to the joint scattering coefficients, which yields a $\patk$ of $94.8 \pm 3.3$, which is a significant increase.
+We therefore conclude that a simple reweighting of the scattering transform yields a more appropriate timbral similarity measure.
 

### 2019-09-11 paper/paper.tex
descriptions of figures
Adopt PLOS template

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% HEADER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Template for PLoS
+% Version 3.5 March 2018
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
-\documentclass{article}
-\usepackage{amsmath,cite,url}
-\usepackage{graphicx}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% our packages
+\usepackage{url}
 \usepackage{color}
 \usepackage{xspace}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
+
+%% BEGIN OUR MACROS SECTION
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Learning Auditory Similarities Between Instrumental Playing Techniques}
-
-\author{
-Christian El-Hajj,
-Vincent Lostanlen,
-Mathias Rossignol, \\
-Gr\'egoire Lafay,
-and Mathieu Lagrange}
+%% END OUR MACROS SECTION
 
 \begin{document}
+\vspace*{0.2in}
 
-\maketitle
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Learning Auditory Similarities Between Instrumental Playing Techniques} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Christian El-Hajj\textsuperscript{1},
+Vincent Lostanlen\textsuperscript{2},
+Mathias Rossignol\textsuperscript{3},
+Gr\'egoire Lafay\textsuperscript{1},
+Mathieu Lagrange\textsuperscript{2*}
+\\
+\bigskip
+\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\bigskip
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* correspondingauthor@institute.edu
 
-\begin{abstract}
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 ABSTRACT HERE
-\end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
+AUTHOR SUMMARY HERE
 
-\section{Introduction}
+\linenumbers
+
+\section*{Figures}
+
+Fig 1: strf + scattering (upsampling strf)
+
+Fig 2: one liner music players (3 ipts bass pizz, bass archo, violin archo) sound, spec, scat, isomap
+other line : perceptual clustering
+
+Fig 3: same 3 sounds spec + scatt
+
+Fig 4: tags clouds
+
+Fig last: perfs mfcc separable joint
+
+% Use "Eq" instead of "Equation" for equation citations.
+\section*{Introduction}
 \label{sec:introduction}
@@
 
-\section{Results}
-\label{sec:results}
+\section*{Materials and methods}
+\label{sec:methods}
+
+\begin{itemize}
+\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
+\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
+\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
+\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
+\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities.
+The result is a convolutional network with fixed filters.
+Performs well in various classification and regression tasks when augmented with a learned linear layer.
+\item Linear layer is provided here by LDA or LMNN.
+\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
+\end{itemize}
+
+\section*{Results}
 
@@
 
-\subsection{Scattering transform}
+\subsection*{Scattering transform}
 \label{sec:scattering}
@@
 
-\subsection{Similarity weighting}
+\subsection*{Similarity weighting}
 \label{sec:weighting}
@@
 
-\section{Discussion}
+\section*{Discussion}
 \label{sec:discussion}
@@
 
-\section{Methods}
-\label{sec:methods}
+\section*{Conclusion}
+\label{sec:conclusion}
 
-\begin{itemize}
-\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
-\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
-\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
-\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
-\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities.
-The result is a convolutional network with fixed filters.
-Performs well in various classification and regression tasks when augmented with a learned linear layer.
-\item Linear layer is provided here by LDA or LMNN.
-\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
-\end{itemize}
+\section*{Supporting information}
+\label{sec:supp}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgments}
+\label{sec:ack}
 
-\bibliographystyle{alpha}
+\nolinenumbers
+
+\bibliographystyle{plos2015}
 \bibliography{bib}

### 2019-09-14 paper/paper.tex
Move methods to end and remove conclusion
Move survey discussion to separate section

@@
 
-\section*{Materials and methods}
-\label{sec:methods}
-
-\begin{itemize}
-\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
-\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
-\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
-\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
-\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities.
-The result is a convolutional network with fixed filters.
-Performs well in various classification and regression tasks when augmented with a learned linear layer.
-\item Linear layer is provided here by LDA or LMNN.
-\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
-\end{itemize}
-
 \section*{Results}
@@
 \end{itemize}
+
+\subsection*{Timbre similarity perception}
+\label{sec:survey}
+
+To validate our timbre perception model, we first acquired similarity judgments from a survey of human subjects.
+The survey was carried out for $78$ sounds, each representing an instrument--playing technique (IPT) pair.
+These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
+
+In the survey, a subject was provided with these $78$ IPTs at the same pitch and asked to cluster them in a free sorting task.
+This was achieved by arranging correspoding dots in a two-dimensional plane on a computer screen and assigning a color to each dot.
+While the location of the dots was recorded, this information was not used in our analysis and only served to help the subject organize the sounds spatially.
+The colors assigned to each dot---and, consequently, to each sound---was recorded for each subject.
+
+The above similarity judgments were crowdsourced by sending out requests to perform the experiment on internal mailing lists of the CNSMDP and international research mailing lists with focus on audio and music processing.
+We made the experiment available for two months and obtained results for a total of $31$ subjects.
+
+Given the clustering assignments of each subject, we then constructed a consensus clustering using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
+Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
 
@@
 
-These human similarity judgments are obtained from a set of $78$ instrument--IPT pairs selected by two music composition professors at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
-A subject was provided with these 78 recordings and asked to cluster them in a free sorting task by arranging correspodings dots in two dimensions on a computer screen and assigning a color to each dot.
-While the location of the dots was recorded, this information was not used in our analysis and only served to help the subject organize the sounds spatially.
-The colors assigned to each dot---and, consequently, to each sound---was recorded for each subject.
-
-The above similarity judgments were crowdsourced by sending out requests to perform the experiment on internal mailing lists of the CNSMDP and international research mailing lists with focus on audio and music processing.
-We made the experiment available for two months and obtained results for a total of $31$ subjects.
-
-Given the clustering assignments of each subject, we then constructed a consensus clustering using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
-Interestingly, although several sounds pairs were obtained from the same instrument (but different IPTs) and others were obtained using the same IPT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
-
 We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
@@
 
-\section*{Conclusion}
-\label{sec:conclusion}
+\section*{Materials and methods}
+\label{sec:methods}
+
+\begin{itemize}
+\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
+\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
+\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
+\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
+\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities.
+The result is a convolutional network with fixed filters.
+Performs well in various classification and regression tasks when augmented with a learned linear layer.
+\item Linear layer is provided here by LDA or LMNN.
+\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
+\end{itemize}
 

### 2019-09-14 paper/paper.tex
Move outline into comments

@@
 
-\textbf{Section outline:}
-\begin{itemize}
-\item How to accurately model timbre perception?
-\item Studies either consider stimuli that are ``too simple'' (e.g., tones, instruments with fixed playing techniques) or ``too complex'' (e.g., speech, environmental sounds).
-\item Computational models do not capture enough information (MFCCs and related representations) or are hard to compute or analyze (STRFs, deep networks).
-\item Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (instrumental playing techniques, or IPTs).
+% Section outline:
+% - How to accurately model timbre perception?
+% - Studies either consider stimuli that are ``too simple'' (e.g., tones, instruments with fixed playing techniques) or ``too complex'' (e.g., speech, environmental sounds).
+% - Computational models do not capture enough information (MFCCs and related representations) or are hard to compute or analyze (STRFs, deep networks).
+% - Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (instrumental playing techniques, or IPTs).
 Provides more control in modeling different aspects of timbre.
-\item In addition, propose a computational model for timbral similarity: (joint) scattering plus a linear layer.
+% - In addition, propose a computational model for timbral similarity: (joint) scattering plus a linear layer.
 Fixed representation to extract time-frequency structure with adaptivity in linear component.
-\end{itemize}
 
@@
 
-\textbf{Section outline:}
-\begin{itemize}
-\item Subjects asked to cluster 78 IPTs into an arbitrary number of clusters (maximum 20). Resulting partition gives measure of timbral similarity.
-\item Partitions do not completely agree with either instrument (I) of playing technique (PT) taxonomies, although slightly favor PTs.
+% Section outline:
+% - Subjects asked to cluster 78 IPTs into an arbitrary number of clusters (maximum 20). Resulting partition gives measure of timbral similarity.
+% - Partitions do not completely agree with either instrument (I) of playing technique (PT) taxonomies, although slightly favor PTs.
 Perceptual measurement captures timbre across IPTs.
 Show through NMI or other measurements.
-\item Propose computational model for these perceptual similarity measurements: scattering transform (time or time-frequency), followed by a learned linear layer (LDA or LDMM) to adapt representation to measured data.
-\item Model shows good agreement with similarity measurements under different testing protocols (aggregated clusters, aggregated subspaces, cross validation).
-\end{itemize}
+% - Propose computational model for these perceptual similarity measurements: scattering transform (time or time-frequency), followed by a learned linear layer (LDA or LDMM) to adapt representation to measured data.
+% - Model shows good agreement with similarity measurements under different testing protocols (aggregated clusters, aggregated subspaces, cross validation).
 

### 2019-09-14 paper/paper.tex
Edit scattering subsection

@@
 
-A closely related representation is that of the joint time--frequency scattering transform, introduced by And\'{e}n et al. \cite{anden2015joint,anden2019joint}.
+A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
 Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
@@
 This representation contains a large amount of information, but not in a form suitable for building statistical models.
-In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may incude large changes to the scalogram.
-One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients \cite{anden2014deep}.
+In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
+It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
+One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
+These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
 
-These first-order coefficients alone cannot accurately characterize the time--frequency structure of the signal, however, since they only include spectral information.
-To augment them, we perform another wavelet decomposition, this time in two dimensions, over the scalogram.
+Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
+One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
+Again, we compute the modulus of the wavelet coefficients and average in time.
+The resulting coefficients are known as the second-order time scattering coefficients.
+These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
+
+While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
+Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
+If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
+We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
+This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
+The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
+
+This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
+To remedy this, another approach was suggested by And\'{e}n et al.
+\cite{anden2015joint,anden2019joint} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
 Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-Again, we compute the modulus of the transform and average in time.
-The result is known as the set of second-order joint time--frequency scattering coefficients \cite{anden2015joint,anden2019joint}.
+As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
 
@@
 The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids \cite{anden2012scattering,anden2019joint}, and beating phenomena \cite{anden2014deep}.
+Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019joint}, and beating phenomena \cite{anden2014deep}.
 They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019joint,lostanlen2018relevance,lostanlen2018extended}.
@@
 Increasing the variability in IPTs allows us to explore a wider range of timbre, providing a more robust validation of the proposed model for modeling timbre.
-The metric used is the \emph{precision at rank $k$} ($\patk$), which is calculated by computing the $k$ nearest neighbors in the feature space and recording the proportion of the neighbors that are in the same ``class.''
+The metric used is the \emph{precision at rank $k$} ($\patk$), which is calculated by computing the $k$ nearest neighbors in the feature space and recording the proportion of the neighbors that are in the same class.
 In the following, we set $k = 5$.
+The classes were obtained from the consensus clustering of human similarity judgments described in the previous section.
+The $\patk$ therefore provides a measure of consistency between the timbre similarity measures obtained by our model and those of the survey subjects.
 
-Concatenating the first- and second-order joint scattering coefficients, we obtain a feature vector of dimension $???$ for a sound recording of duration $???$.
-Computing nearest neighbors with the standard Euclidean measure gives a $\patk$ of $87.0 \pm 5.8$, where the first value indicates the mean and the second the standard deviation.
+Concatenating the first- and second-order separable time and frequency scattering coefficients, we obtain a feature vector of dimension $???$ for a sound recording of duration $???$.
+Computing nearest neighbors with the standard Euclidean measure gives a $\patk$ of $0.870 \pm 0.058$, where the first value indicates the mean and the second the standard deviation.
 As a result, this simple transformation of the original signal results in a representation that lends itself naturally to discriminating timbre.
 Note that this result is obtained by simply looking at the raw feature vectors---no learning is performed to adapt the transform towards discriminating timbre.
-We contrast this with the results obtained with a baseline MFCC feature vector.
-This results in a $\patk$ of $85.1 \pm 6.2$.
-We therefore conclude that, although the scattering transform provides a good indication of timbral similarity, using its nearest neighbors based on a standard Euclidan distance does not allow us to properly account for the higher-level time--frequency structure characterized by the second-order coefficients.
+Replacing the separable second-order coefficients with the joint second-order coefficients boosts the $\patk$ to $???$.
+This is evidence that accurately characterizing the joint time--frequency structure of the sound is important capturing its timbre.
+
+We performed the same evaluation with a standard MFCC representation, obtaining a $\patk$ of $0.851 \pm 0.062$.
+As a result, we conclude that a Euclidean distance over scattering transforms (separable or joint) yield little additional perceptual information compared to standard spectral descriptors.
 

### 2019-09-16 paper/paper.tex
Update section on weighting

@@
 
-While the raw scattering representation provides a powerful model for timbre, it does so without any prior information.
-Incorporating information about what distinguishes one timbre from another should improve the accuracy of the model.
-In the same was that the lower levels of auditory perception do not vary much between subjects (or between species, in the case of mammals), we expect that higher levels incorporate some amount of learning.
-Introducing a learned component into our model, however, could greatly reduce its usefulness with respect to interpretability.
-For this reason, we restrict ourselves to a simple, linear transform of the scattering coefficients.
-By optimizing the linear transform with respect to similarity measurements obtained from human subjects, we expect to obtain a more accurate model of timbre perception.
+To improve our measure, we need to replace the standard Euclidean distance on the raw scattering coefficients by something better adapted to discriminating timbre.
+Equivalently, we can apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
+Very flexible and powerful methods can be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase in complexity.
+This increased complexity of the model reduces its usefulness with respect to interpretability.
+
+For this reason, we restrict ourselves to a simple linear transform of the scattering coefficients.
+Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
+We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
 
@@
 This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-Running LMNN on MFCCs with the consensus clustering obtained above increases $\patk$ slightly to $86.2 \pm 5.9$.
+Running LMNN on MFCCs with the consensus clustering obtained above increases $\patk$ slightly to $0.862 \pm 0.059$.
 Consequently, there seems to be little additional timbre information in the MFCCs compared to that given by the standard Euclidean distance.
-We contrast this with LMNN applied to the joint scattering coefficients, which yields a $\patk$ of $94.8 \pm 3.3$, which is a significant increase.
+We contrast this with LMNN applied to the separable scattering coefficients, which yields a $\patk$ of $0.948 \pm 0.033$, which is a significant increase.
+For the joint scattering coefficients, we obtain a $\patk$ of $???$.
 We therefore conclude that a simple reweighting of the scattering transform yields a more appropriate timbral similarity measure.

### 2019-09-17 demo/organizer.js
improve demo

@@ -76,11 +76,9 @@ app.controller("Organizer", function($scope, $http, $timeout, $log) {
 
 all = [
   ' ',
-      'mdsProjectionReferenceInstruments',
-      'mdsProjectionReferencePlayingTechniques',
-    'mdsProjectionSubjectsClustering_5',
-    'mdsProjectionSubjectsClustering_8',
-    'mdsProjectionSubjectsConsensusClustering'
+  'labelByConsensusClustering',        'labelBySubjectiveClustering_5clusters',
+'labelByInstruments',   'labelBySubjectiveClustering_8clusters',
+'labelByPlayingTechniques'
   ];
 
   nums = Array.apply(null, {length: 32}).map(Number.call, Number).map(String);

### 2019-09-17 paper/paper.tex
Switch to BioMed Central template
As required by Jasmp.

@@
-% Template for PLoS
-% Version 3.5 March 2018
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file 
-% and leave only the final text of your manuscript. 
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that 
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column. 
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
+%% BioMed_Central_Tex_Template_v1.06
+%%                                      %
+%  bmc_article.tex            ver: 1.06 %
+%                                       %
 
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+%%IMPORTANT: do not delete the first line of this template
+%%It must be present to enable the BMC Submission system to
+%%recognise this template!!
 
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                     %%
+%%  LaTeX template for BioMed Central  %%
+%%     journal article submissions     %%
+%%                                     %%
+%%          <8 June 2012>              %%
+%%                                     %%
+%%                                     %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% For instructions on how to fill out this Tex template           %%
+%% document please refer to Readme.html and the instructions for   %%
+%% authors page on the biomed central website                      %%
+%% http://www.biomedcentral.com/info/authors/                      %%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%% BioMed Central currently use the MikTex distribution of         %%
+%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
+%% http://www.miktex.org                                           %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
+\documentclass{bmcart}
 
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% our packages
+\usepackage{amsthm,amsmath}
+\usepackage[utf8]{inputenc}
 \usepackage{url}
@@
 
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                             %%
+%%  If you wish to display your graphics for   %%
+%%  your own use using includegraphic or       %%
+%%  includegraphics, then comment out the      %%
+%%  following two lines of code.               %%
+%%  NB: These line *must* be included when     %%
+%%  submitting to BMC.                         %%
+%%  All figure files must be submitted as      %%
+%%  separate graphics through the BMC          %%
+%%  submission process, not included in the    %%
+%%  submitted article.                         %%
+%%                                             %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
+\def\includegraphic{}
+\def\includegraphics{}
 
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-% Remove comment for double spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-%\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-%\setlength{\headheight}{27.023pt}
-%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\today}
-
-%% Include all macros below
-
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
-%% END MACROS SECTION
-
-%% BEGIN OUR MACROS SECTION
+%%% Put your definitions there:
+\startlocaldefs
 
@@
 
-%% END OUR MACROS SECTION
+\endlocaldefs
 
+%%% Begin ...
 \begin{document}
-\vspace*{0.2in}
 
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Learning Auditory Similarities Between Instrumental Playing Techniques} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+%%% Start of article front matter
+\begin{frontmatter}
+
+\begin{fmbox}
+\dochead{Research}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the title of your article here     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{Learning Auditory Similarities Between Instrumental Playing Techniques}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the authors here                   %%
+%%                                          %%
+%% Specify information, if available,       %%
+%% in the form:                             %%
+%%   <key>={<id1>,<id2>}                    %%
+%%   <key>=                                 %%
+%% Comment or delete the keys which are     %%
+%% not used. Repeat \author command as much %%
+%% as required.                             %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\author[
+   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   corref={aff1},                       % id of corresponding address, if any
+   noteref={n1},                        % id's of article notes, if any
+   email={email@example.com}            % email address
+]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
+\author[
+   addressref={aff2},
+   email={email@example.com}
+]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
+\author[
+   addressref={aff2},
+   email={email@example.com}
+]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
+\author[
+   addressref={aff2},
+   email={email@example.com}
+]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
+\author[
+   addressref={aff2},
+   email={email@example.com}
+]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the authors' addresses here        %%
+%%                                          %%
+%% Repeat \address commands as much as      %%
+%% required.                                %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\address[id=aff1]{%                           % unique id
+  \orgname{Department of Zoology, Cambridge}, % university, etc
+  \street{Waterloo Road},                     %
+  %\postcode{}                                % post or zip code
+  \city{London},                              % city
+  \cny{UK}                                    % country
 }
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Christian El-Hajj\textsuperscript{1},
-Vincent Lostanlen\textsuperscript{2},
-Mathias Rossignol\textsuperscript{3},
-Gr\'egoire Lafay\textsuperscript{1},
-Mathieu Lagrange\textsuperscript{2*}
-\\
-\bigskip
-\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
-\\
-\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
-\\
-\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
-\\
-\bigskip
+\address[id=aff2]{%
+  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
+  \street{D\"{u}sternbrooker Weg 20},
+  \postcode{24105}
+  \city{Kiel},
+  \cny{Germany}
+}
 
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* correspondingauthor@institute.edu
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter short notes here                   %%
+%%                                          %%
+%% Short notes will be after addresses      %%
+%% on first page.                           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
-ABSTRACT HERE
+\begin{artnotes}
+\note[id=n1]{Equal contributor}
+\end{artnotes}
 
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
-\section*{Author summary}
-AUTHOR SUMMARY HERE
+\end{fmbox}
 
-\linenumbers
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The Abstract begins here                 %%
+%%                                          %%
+%% Please refer to the Instructions for     %%
+%% authors on http://www.biomedcentral.com  %%
+%% and include the section headings         %%
+%% accordingly for your article type.       %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section*{Figures}
+\begin{abstractbox}
 
-Fig 1: strf + scattering (upsampling strf)
+\begin{abstract}
+Abstract here.
+\end{abstract}
 
-Fig 2: one liner music players (3 ipts bass pizz, bass archo, violin archo) sound, spec, scat, isomap
-other line : perceptual clustering
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The keywords begin here                  %%
+%%                                          %%
+%% Put each keyword in separate \kwd{}.     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Fig 3: same 3 sounds spec + scatt
+\begin{keyword}
+\kwd{sample}
+\kwd{article}
+\kwd{author}
+\end{keyword}
 
-Fig 4: tags clouds
+\end{abstractbox}
+%
+%\end{fmbox}% uncomment this for twcolumn layout
 
-Fig last: perfs mfcc separable joint
+\end{frontmatter}
 
-% Use "Eq" instead of "Equation" for equation citations.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The Main Body begins here                %%
+%%                                          %%
+%% Please refer to the instructions for     %%
+%% authors on:                              %%
+%% http://www.biomedcentral.com/info/authors%%
+%% and include the section headings         %%
+%% accordingly for your article type.       %%
+%%                                          %%
+%% See the Results and Discussion section   %%
+%% for details on how to create sub-sections%%
+%%                                          %%
+%% use \cite{...} to cite references        %%
+%%  \cite{koon} and                         %%
+%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
+%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section*{Introduction}
-\label{sec:introduction}
+\label{sec:intro}
 
@@
 
-\section*{Results}
-
-% Section outline:
-% - Subjects asked to cluster 78 IPTs into an arbitrary number of clusters (maximum 20). Resulting partition gives measure of timbral similarity.
-% - Partitions do not completely agree with either instrument (I) of playing technique (PT) taxonomies, although slightly favor PTs.
-Perceptual measurement captures timbre across IPTs.
-Show through NMI or other measurements.
-% - Propose computational model for these perceptual similarity measurements: scattering transform (time or time-frequency), followed by a learned linear layer (LDA or LDMM) to adapt representation to measured data.
-% - Model shows good agreement with similarity measurements under different testing protocols (aggregated clusters, aggregated subspaces, cross validation).
-
-\subsection*{Timbre similarity perception}
+\section*{Timbre similarity judgments}
 \label{sec:survey}
@@
 Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+\section*{Proposed approach}
+\label{sec:method}
 
@@
 
+\section*{Results}
+\label{sec:results}
+
+\ja{The results are currently sprinkled into the above methods section.
+These need to be extracted and summarized here.}
+
 \section*{Discussion}
@@
 
-\section*{Materials and methods}
-\label{sec:methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Backmatter begins here                   %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{itemize}
-\item IPTs pre-screened by two composition teachers to select 78 ``interesting'' IPTs that are very similar to (i.e., easily confused with) other IPTs.
-\item Subjects asked to perform free sorting task on the selected IPTs according to similarity. Allowed to label with up to 20 different colors.
-\item Clustering gives similarity measurement for each subject. These measurements were aggregated in a certain way.
-\item The aggregated clustering was compared to instrument and playing technique taxonomies using NMI or other measures.
-\item The (time or time-frequency) scattering transform is composed of alternating layers of wavelet transforms and modulus nonlinearities.
-The result is a convolutional network with fixed filters.
-Performs well in various classification and regression tasks when augmented with a learned linear layer.
-\item Linear layer is provided here by LDA or LMNN.
-\item Performance is measured in a query-by-example setting with precision at $k$, where $k = 5$.
-\end{itemize}
+\begin{backmatter}
 
-\section*{Supporting information}
-\label{sec:supp}
+\section*{Competing interests}
+  The authors declare that they have no competing interests.
 
-\section*{Acknowledgments}
-\label{sec:ack}
+\section*{Author's contributions}
+    Text for this section \ldots
 
-\nolinenumbers
+\section*{Acknowledgements}
+  Text for this section \ldots
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                  The Bibliography                       %%
+%%                                                         %%
+%%  Bmc_mathpys.bst  will be used to                       %%
+%%  create a .BBL file for submission.                     %%
+%%  After submission of the .TEX file,                     %%
+%%  you will be prompted to submit your .BBL file.         %%
+%%                                                         %%
+%%                                                         %%
+%%  Note that the displayed Bibliography will not          %%
+%%  necessarily be rendered by Latex exactly as specified  %%
+%%  in the online Instructions for Authors.                %%
+%%                                                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibliographystyle{plos2015}
+\bibliographystyle{bmc-mathphys}
 \bibliography{bib}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Figures                       %%
+%%                               %%
+%% NB: this is for captions and  %%
+%% Titles. All graphics must be  %%
+%% submitted separately and NOT  %%
+%% included in the Tex document  %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Figures}
+
+\begin{figure}[h!]
+\caption{\csentence{STRF + scattering (upsampling STRF)}
+A short description of the figure content should go here.}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Model overview}
+One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
+scat, isomap.
+Second line : perceptual clustering.}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Tag clouds}
+A short description of the figure content should go here.}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Perfs MFCC, separable, joint (last)}
+A short description of the figure content should go here.}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Tables                        %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Tables}
+\begin{table}[h!]
+\caption{Sample table title. This is where the description of the table should go.}
+      \begin{tabular}{cccc}
+        \hline
+           & B1  &B2   & B3\\ \hline
+        A1 & 0.1 & 0.2 & 0.3\\
+        A2 & ... & ..  & .\\
+        A3 & ..  & .   & .\\ \hline
+      \end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Additional Files              %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section*{Additional Files}
+  \subsection*{Additional file 1 --- Sample additional file title}
+    Additional file descriptions text (including details of how to
+    view the file, if it is in a non-standard format or the file extension).  This might
+    refer to a multi-page table or a figure.
+
+  \subsection*{Additional file 2 --- Sample additional file title}
+    Additional file descriptions text.
+
+\end{backmatter}
+
 \end{document}

### 2019-09-17 paper/paper.tex
Remove unneeded comment
Rewrite introduction

@@
 \end{abstractbox}
-%
-%\end{fmbox}% uncomment this for twcolumn layout
 
@@
 % Section outline:
-% - How to accurately model timbre perception?
+% - How to accurately model timbre space? Useful for studying auditory perception, browsing databases, etc.
 % - Studies either consider stimuli that are ``too simple'' (e.g., tones, instruments with fixed playing techniques) or ``too complex'' (e.g., speech, environmental sounds).
 % - Computational models do not capture enough information (MFCCs and related representations) or are hard to compute or analyze (STRFs, deep networks).
-% - Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (instrumental playing techniques, or IPTs).
-Provides more control in modeling different aspects of timbre.
-% - In addition, propose a computational model for timbral similarity: (joint) scattering plus a linear layer.
+% - Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (PTs). Provides more control in modeling different aspects of timbre.
+% - In addition, propose a computational model for timbral similarity: scattering (separable or joint) plus a linear layer.
 Fixed representation to extract time-frequency structure with adaptivity in linear component.
@@
 While it is typically defined as the quality of a sound that is independent of its pitch, duration, or loudness, this definition fails to provide an explicit characterization of the attribute.
-Due to its central role in audio perception, timbre has been studied from several angles, including perceptual and psychophysical studies as well as computational methods in machine listening.
-The consensus is that timbre depends not on the purely spectral or temporal structure of a sound, but in its joint spectro--temporal structure, but beyond this, there is little agreement about its precise characterization.
+Due to its central role in audio perception, timbre has been studied from several angles, including perceptual and psychophysical studies.
+Since timbre extraction forms the basis of many audio-based information retrieval tasks, such as classification and determining acoustical similarity, the subject has also been studied from a machine listening perspective.
+\ja{Need references for the above statements.}
 
-One difficulty that arises in studying the perceptual dimensions of timbre relates to its potential complexity---the dimensionality of the full timbral space is high.
-To combat this, many perceptual and psychophysical studies focus on simple stimuli: pure sine waves, musical instruments with fixed pitch and playing technique, and so on.
-While this has led to important insights in timbre perception, the range of timbres explored by these stimuli is necessarily limited.
+A natural candidate for characterizing timbre is the spectral content of a signal.
+This information, however, is not rich enough to fully characterize the timbre of a sound, as evidenced by the recent move from spectral descriptors such as mel-spectrograms and mel-frequency cepstral coefficients (MFCCs) and descriptors of spectral energy distributions.
+Indeed, the success of deep neural networks (DNNs) applied to time--frequency representations (such as spectrograms or scalograms) in machine listening have demonstrated the necessity for capturing both spectral and temporal content.
+In neurophysiological, the auditory perception models proposed by S.
+Shamma and collaborators, known as \emph{spectro--temporal receptive fields} (STRFs) all characterize temporal structure in addition to purely spectral information.
+It seems, therefore, that timbre is inherently spectro--temporal.
+\ja{Again, refs.}
 
-On the other end of the spectrum are the computational methods.
-Since their goals are typically task-oriented (recognizing speech, identifying environmental sounds, and so on), they are less concerned with establishing models of auditory perception suitable for analysis.
-While certain audio representations are simple enough to analyze, such as mel-frequency cepstral coefficients (MFCCs) and related approaches, they do not capture enough structure to adequately characterize timbre.
-Indeed, for an overwhelming majority of machine listening tasks have been these representations have been outperformed by learned representations, such as deep neural networks (DNNs).
-While DNNs have enjoyed significant success for these tasks, their relative complexity and the fact that they are learned from data make their analysis difficult, preventing us from deriving relevant insights into timbre perception.
+While the above models offer some insight into the structure of timbre, their complexity makes interpretation and analysis difficult.
+The expressive power of DNNs are due to tuning a large number of parameters, so it is often hard to relate a specific component to a property of the resulting perception model.
+STRFs, on the other hand, have a simpler structure, but are ultimately constructed to model electrophysiological signal in the auditory cortex.
+As a result, they consist several adjustments that make their computation costly and their mathematical analysis difficult.
+\ja{Refs.}
 
-Certain audio representations have been proposed as a middle ground between expressivity and ease of analysis.
-One of the more prominent among these are the \emph{cortical representations} introduced by S. Shamma and collaborators.
-These rely on a spectrogram-like decomposition of the sound, which is then further decomposed in time and frequency using \emph{spectro--temporal receptive fields} (STRFs).
-This model is based on neurophysiological measurements on ferrets, and have been validated in several contexts, ranging from applications in machine listening and reproduction of psychophysical phenomena in auditory perception.
-However, since the original goal of the representation was to model observed electrophysiological signals, this increases its complexity, rendering mathematical anaylsis more difficult.
+Another difficulty in studying the perception of timbre is the choice of stimuli.
+Previous work on timbre characterization of STRFs \cite{patil2012music} has focused on instrumental sounds of isolated notes played with standard playing techniques.
+This inherently restricts the range of available timbre studied, limiting the conclusions of the work.
+On the other end of the spectrum are computational models for information retrieval tasks.
+Since these are primarily concerned with performance on real-world sounds, the possible range of timbre is very broad.
+The analysis of the resulting model is therefore all the more difficult, since we have little control over the complexity of the stimuli.
+\ja{Refs.}
 
-In particular, cortical representations have been used to study timbre in the work of Patil et al. \cite{patil2012music}.
-While this demonstrates the power of the representation to characterize certain dimensions of timbre, the range of sounds was limited, consisting of single musical notes played using standard instrumental techniques.
-As a result, only a subset of the relevant timbral dimensions are explored, limiting the conclusions of the work.
+Our contribution is twofold.
+First, we propose a simple model for timbre perception involving a fixed time--frequency descriptor, the scattering, followed by a learned linear layer optimized for the timbral space under consideration.
+Second, we validate this model on a set of instrumental sounds with extended playing techniques (PTs), allowing for a richer range of timbre compared to ordinary instrumental sounds while retaining the structure of the space necessary for interpretability.
 
-In this work, we consider representation similar to the cortical representation, namely the \emph{scattering transform}, which provides a mathematical foundation for the study of auditory perception in terms of invariance and stability properties.
-Instead of single musical notes played with standard techniques, we extend the range of sounds to include extended instrumental playing techniques.
-While increasing the complexity of the timbral space explored, these sounds do not include the full range of timbre exhibited, for example, by speech or environmental sounds.
-As a result, we are able to better analyze their characteristics and link them to the properties of the transform.
+As part of validating the model, we have collected timbral similarity judgments in a crowdsourced listening experiment.
+These are used as a ``ground truth'' dataset, which allow us to evaluate the model's ability to characterize the perceived similarity of the stimuli.
+Analysis of these experimental results also confirms the necessity of including a range of PTs, as these are shown to influence similarity judgments to the same extent as the identity of the instrument itself.
+
+We conclude with a discussion of potential applications to browsing interfaces for musical databases, where composers can easily find sounds on a query-by-example basis.
+The adaptivity of the proposed model has the additional advantag of allowing it to be adjusted to a particular listener's preferences by retraining the linear layer.
 

### 2019-09-18 paper/paper.tex
clean repo
subjective test begins

@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
-
-%%IMPORTANT: do not delete the first line of this template
-%%It must be present to enable the BMC Submission system to
-%%recognise this template!!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%          <8 June 2012>              %%
-%%                                     %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.html and the instructions for   %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass{bmcart}
@@
 \usepackage{xspace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %%
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %%
-%%  submitted article.                         %%
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In the survey, a subject was provided with these $78$ IPTs at the same pitch and asked to cluster them in a free sorting task.
-This was achieved by arranging correspoding dots in a two-dimensional plane on a computer screen and assigning a color to each dot.
+In the survey, a subject was provided with one sound recording per IPTs with the same nuance and pitch when relevant. The subject is then asked to cluster those $78$ sounds in a free sorting task.
+This was achieved by arranging corresponding dots in a two-dimensional plane on a computer screen and assigning a color to each dot.
 While the location of the dots was recorded, this information was not used in our analysis and only served to help the subject organize the sounds spatially.
@@
 
-Given the clustering assignments of each subject, we then constructed a consensus clustering using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
+The outcome of this process can be formalized as follows.
+For each subject, each pair of IPTs is given a boolean similarity value.
+If the two IPTs have been given the same color by the subject, the similarity is set to one, and zero otherwise, see Figure \ref{fig:test}.
+This information is usually called a "clustering", where each item is given a label within a closed set of labels, but those labels do not carry information by itself.
+
+\ja{equation needed ?}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/xp2example1.png}
+\caption{Spatial and group organization provided by subject 1.}
+\label{fig:xp2display}
+\end{figure}
+
+\ml{also have the resulting binary similarity matrix ?}
+
+Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
+In this case, the resulting timbre space is specialized to a given person.
+It may also be of interest to condition the timbre space using the information given by several subjects.
+Though, conditioning the timbre space using several binary similarity matrices is not trivial, and will be discussed extensively in Section \ref{sec:results}.
+
+We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
 Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+More information about the data collection process is available in Section \ref{sec:subjective}.
 
@@
 \label{sec:method}
+
+\ml{introduce the system globally, while stressing the ability of the system to be conditionned for one or many people}
 
@@
 
+\section{Dataset} \label{sec:dataset}
+
+The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of 16 different musical instruments played with different playing techniques, which leads to 235 different couple of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to 25444 audio samples.
+
+The recorded instruments belongs to the wind and string classes.
+For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+
+Some instruments can be complemented with sordina (S), wha (W).
+The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
+\vl{This is crucial and should not be deferred to the appendix. @writing}
+\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
+
+Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+For  the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap  (explo), harmonic (harm), key-click (key), col legno   (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+
+\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+The abbreviations can be removed. @data @writing}
+\ml{}
+
+\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+(y axis) would be extremely useful. @data @figure}
+\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+\vl{Les donn\'{e} de l'\{e}tude, relatives aux Figures 5 et 6.\}
+
 \section*{Discussion}
@@
 \end{itemize}
+
+\section*{Human timbre judgments}
+\label{sec:subjective}
+
+The dataset considered in this study is taken
+from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
+It has audio recordings of 16 different musical instruments played
+with different playing techniques,
+which leads to $235$ different couples of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to $25444$ recordings.
+The way the dataset is structured
+is explained in more details in Appendix \ref{sec:dataset}.
+
+Gathering the similarity judgments between each pair of samples would be a daunting task.
+Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
+
+One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
+This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
+
+\subsection{Reducing Expert judgments of technique peculiarities}
+
+Even so, considering every pairs of IPTs is not practical. A selection
+experiment is thus conducted where the subject is asked to give his opinion on which
+\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
+if it is likely to be associated to another \ipt of another instrument.
+Interest is rated on a 7 ticks scale. The subject can listen to all the different
+nuance and pitch samples of each \ipt. The experiment is over when all the
+\ipts are rated.
+
+Rating guidelines are provided :
+\begin{itemize}
+  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
+  \item Four stars: there is a proximity on one aspect of the sound between this
+  \ipt and of an \ipt of another instrument,
+  but this one is neither decisive nor obvious.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
+\end{itemize}
+
+Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
+
+\vl{This subsection needs a box-and-whisker plot of ratings, where the 235 different
+couples of instrument and playing techniques are sorted on the x-axis by average rating. @figure @ratings}
+
+\ml{il n'y que deux raters, ca ne pas pas sens de faire et ca sera illisible}
+
+\vl{Are the histogram of ratings similar for the two raters? It would be good to make a statistical test for this. @research @stats @ratings}
+
+\ml{les histograms sont dans la figure statxp1. Le code pour la generer est dans code/xp1 faire tourner extractData et statData. Un ttest rejette l'hypothese nulle.}
+
+Even if the hypothesis that the distributions of ratings comes from independent random samples from normal distributions with equal means and equal but unknown variances, is verified using the two-sample t-test, inspection of the respective histograms reveals that averaging the ratings of the two raters is reasonable.
+
+Keeping in mind that the gathering of the perceptual judgments of \ipt similarities shall remain practical, the number of \ipts considered shall be lower than $100$.
+We thus select the \ipts with an average rating higher than three, which brings us 78 \ipts to study.
+
+%%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
+\subsection{Crowdsourced judgments of pairwise similarity}
+
+The second experiment aims at gathering spontaneous judgment of similarity for those 78 \ipts using a canonical free sorting task.
+This type of task is chosen among others for two main reasons.
+First, it gives the ability to the subject to listen to the whole set of \ipts before performing the task.
+Second, it gives more freedom to the subject, thus potentially more interest to perform the task than a lengthy series of XXY dissimilarity ratings.
+
+The subjects are thus asked to organize the 78 \ipts displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
+The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to follow its own judgment.
+
+The plane is displayed on a computer screen using a web based interface\footnote{The interface is available here: http://soundthings.org/ticel/groupSounds}.
+The \ipts can be listened to by hovering the mouse on the corresponding dot.
+The experiment is over when the subject assigned a color to all the dots.
+The initial location of the dots on the plane is randomized for each subject and can be changed during the experiment by the subject for convenience.
+Event hough the location of the dots is recorded, this data is not considered during the analysis.
+
+The experiment was available online for two months and requests to perform the test have been sent to internal mailing lists of the CNSMDP and international research mailing with a focus on audio and music processing.
+The experiment have been performed on a voluntary basis without gratification.
+Subjects were asked to use headphones at a comfortable listening level.
+
+31 subjects completed the experiment.
+An example of resulting organization of \ipt is given on Figure \ref{fig:xp2display}.
+\footnote{The resulting organization for each subject are available for reader's inspection using the full feature interface provided to the subjects for performing the experiment \url{https://mathieulagrange.github.io/paperSpontaneousSimilarity/demo}}.
+
+As only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
+We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
+
+This results in $31$ clusterings that can be considered for defining a perceptual timbre space of \ipts.
+
+\subsection{Analysis}\label{sec:analysis}
+
+In order to gain information about the relevance those $31$ clusterings for defining a perceptual timbre space of \ipts, an analysis is performed.
+
+Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipts.
+The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbGroup}.
+The size of the groups is on average $7.7 \pm   7.2$.
+The distribution of the size of groups can be seen on Figure \ref{fig:xp2sizeGroup}.
+The large number of small groups indicates that many subjects considered that a few of the \ipt were very different from all the remaining \ipt.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/nbc.png}
+\caption{Histogram of the number of groups.
+\vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
+\ml{ok}
+\label{fig:xp2nbGroup}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/sbc.png}
+\caption{Histogram of the size of groups.
+\vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
+\ml{ok}
+\label{fig:xp2sizeGroup}
+\end{figure}
+
+A clustering can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
+By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that corresponds to our measure of the spontaneous perceptual similarity between \ipt.
+
+In order to gather information about the properties of the data, a cluster analysis is next performed using an agglomerative hierarchical clustering \cite{gordon1987review} using the weighted average algorithm for computing distance between clusters.
+The resulting dendrogram is displayed on Figure \ref{fig:dendrogram}.
+A partition of $n$ clusters can be constructed from the agglomerative hierarchical cluster tree by finding the smallest height at which a horizontal cut through the tree leaves $n$ clusters.
+
+One question that arises is whether the similarity is influenced primarily by the instrument or by the playing technique.
+To answer this question, each partition provided by cutting the dendrogram at successive levels is compared to two reference partitions.
+One reference is the partition of the \ipts when labeled with the instrument used (I) and the other if the partition of the \ipts when labeled with the playing technique used (Pt).
+
+There are many metrics available to compare two partitions by evaluating their degree of correspondence \cite{wagner2007comparing}.
+The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondence analysis and the two partition are not required to have the same number of groups.
+It also by design bounded between 0 and 1, 1 meaning perfect match.
+We consider the normalization proposed in \cite{strehl2002cluster}.
+Figure \ref{fig:clusters} shows the evolution of the NMI when comparing the partitions provided by the cluster analysis of the dendrogram produced using the human judgment of similarity and respectively the instrument partition ($j/I$) or the playing technique partition ($j/Pt$).
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
+\caption{Cluster analysis of the similarity judgments.
+$j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
+Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
+\vl{This line plot is very hard to understand.
+The legend box occludes the data on the top-right corner.
+It would be preferable to remove this legend box and label lines
+by placing text upon them,
+so that the plot would be legible to readers with color vision deficiencies.
+Besides, the choice of line labels is cryptic, as it relies on abbreviations.
+The distinction between I and Pt
+can be illustrated by two complementary colors, \eg{} blue versus orange.
+The distinction between j, nj, and null
+can be illustrated by three distinct markers
+(empty circle, filled triangle, asterisk),
+associated with
+three distinct line styles (solid, dashed, dotted).
+The NMI abbreviation should be expanded.
+@figure}}
+\ml{ok, le .fig est dispo}
+\label{fig:clusters}
+\end{figure}
+
+There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipts.
+Also, for both curves, a global trend of increase in the $j/x$ partitions with respect to the number of clusters can be observed.
+This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}.
+To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters.
+The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}.
+This so-called null partitions NMI are respectively termed $nullI$ and $nullPt$ depending on the reference used.
+To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply subtract the $null$ NMI:
+\begin{eqnarray}
+  nj/I &=& j/I - null \\
+  nj/Pt &=& j/Pt - null  \\
+\end{eqnarray}
+This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
+As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$.
+For the latter, reducing the number of clusters down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
+Even by considering the null normalization, the matching level is much higher for the playing technique reference.
+
+The correspondences between those partitions and the reference ones are now studied.
+For this purpose, Figure \ref{fig:gi} displays two partitions.
+At the bottom is the partition that corresponds to the instruments labels, whose color code is given by the color bar on the right.
+The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
+The ordering of \ipts on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
+The partition on top is the result of the clustering with 8 clusters.
+For visualization purposes, we arbitrarily set colors to clusters of this partition to the color code of the instrument label that is most often encountered in this cluster.
+If some labels occur equally often, the instrument of the first \ipt of the cluster is chosen.
+The emerging instrument classes are: Saxophone, Clarinet, Trumpet, Flute and Cello.
+
+However, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
+Also, some clusters could be merged without strong changes in the overall organization.
+This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
+
+As can be seen on Figure \ref{fig:gm}, the correspondence between the playing technique reference and the resulting partition with 5 clusters is much higher.
+Most clusters have a high level of purity and for groups emerge : Pizzicato, Slap, Ordinario, and Ponticello.
+
+To conclude this section, the process of gathering data about the perceptual similarity of \ipt is found to be successful as the subjects used their degree of freedom in a consistent manner and no inconsistency in the data has been detected during the analysis process.
+The main result is that the playing technique is an important factor when modeling the spontaneous judgment of similarity gathered in the second experiment.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/groupModes.png}
+\caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
+The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.
+\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
+\ml{Je demande a voir, mais je ne suis pas satisfait de cette visu de toute facon. Le code est dispo dans le rep code. Lancer ticelSim en premier.}
+\label{fig:gm}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/groupInstruments.png}
+\caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
+The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.
+\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
+\ml{ok}
+\label{fig:gi}
+\end{figure}
 

### 2019-09-18 code/xp1/statData.m

@@ -3,7 +3,7 @@ clear all
 load('extractedData/results.mat')
 
 a = histcounts(rating(:, 1), 1:7)
-b = histcounts(rating(:, 2), 1:7)
+b = histcounts(rating(:, end), 1:7)
 
 bar([a; b]')
 xlabel('Rating')

### 2019-09-19 paper/paper.tex
Update graphics redefinition
Needs to account for optional (width) argument.
Add SF to acks
Fix comment
Edit similarity judgment section

@@
 
-\def\includegraphic{}
-\def\includegraphics{}
+\def\includegraphic[#1]{}
+\def\includegraphics[#1]{}
 
@@
 
-To validate our timbre perception model, we first acquired similarity judgments from a survey of human subjects.
-The survey was carried out for $78$ sounds, each representing an instrument--playing technique (IPT) pair.
-These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
-The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
+To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
+The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch and nuance (for more information on the dataset, see the Dataset section).
+\ja{Mathieu, which pitch and nuance?}%
+In the survey, a subject asked to dots on a two-dimensional place in a free sorting task.
+Each dot corresponded to a different IPT recording and was moved by the user dragging them with the mouse.
+In addition to spatially arranging the dots, the user was also asked to assign one of $20$ colors to each dot, providing a simple clustering of the sounds.
+While both the location and color of each dot was recorded, only the color was used in our subsequent analysis---the spatial arrangement was only used to facilitate the user's ability to organize the sounds.
+More information on the data collection process is available in the Human timbre judgments section.
 
-In the survey, a subject was provided with one sound recording per IPTs with the same nuance and pitch when relevant. The subject is then asked to cluster those $78$ sounds in a free sorting task.
-This was achieved by arranging corresponding dots in a two-dimensional plane on a computer screen and assigning a color to each dot.
-While the location of the dots was recorded, this information was not used in our analysis and only served to help the subject organize the sounds spatially.
-The colors assigned to each dot---and, consequently, to each sound---was recorded for each subject.
-
-The above similarity judgments were crowdsourced by sending out requests to perform the experiment on internal mailing lists of the CNSMDP and international research mailing lists with focus on audio and music processing.
-We made the experiment available for two months and obtained results for a total of $31$ subjects.
-
-The outcome of this process can be formalized as follows.
-For each subject, each pair of IPTs is given a boolean similarity value.
-If the two IPTs have been given the same color by the subject, the similarity is set to one, and zero otherwise, see Figure \ref{fig:test}.
-This information is usually called a "clustering", where each item is given a label within a closed set of labels, but those labels do not carry information by itself.
-
-\ja{equation needed ?}
+The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
+We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
+The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fix:xp2display}.
 
@@
 
+For each subject, we constructed a matrix of size $78 \times 78$ to represent their similarity judgments for each pair of the $78$ IPTs.
+Specifically, the $(i, j)$th element of the matrix was set to one if the $i$th and $j$th IPT were assigned the same color.
+If not, it was set to zero.
+In other words, we record the clustering assignment of each subject as a binary similarity matrix.
+
+\ml{Joakim, equation needed ?}
+\ja{I don't think so.
+It should be fairly straightforward.}%
+
 \ml{also have the resulting binary similarity matrix ?}
+\ja{Yes, but I think we should have it for the consensus clustering described below.}%
 
 Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
-In this case, the resulting timbre space is specialized to a given person.
-It may also be of interest to condition the timbre space using the information given by several subjects.
-Though, conditioning the timbre space using several binary similarity matrices is not trivial, and will be discussed extensively in Section \ref{sec:results}.
+In this case, the correspondign binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
+It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
+That being said, conditioning the timbre space using several binary similarity matrices is not a trivial task, and will be discussed extensively in the Results section.
+\ja{Why discuss it extensively there?
+Don't we just settle on the partitioning algorithm below?}%
 
@@
 Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
-
-More information about the data collection process is available in Section \ref{sec:subjective}.
+\ja{We should have some figure to back this up.
+Can this be done using the tag clouds?}
 
@@
 \section*{Acknowledgements}
-  Text for this section \ldots
+The Flatiron Institute is a division of the Simons Foundation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-09-19 paper/paper.tex
Move Dataset to back matter
Edit dataset section

@@
 
-\section{Dataset} \label{sec:dataset}
-
-The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of 16 different musical instruments played with different playing techniques, which leads to 235 different couple of instrument / playing technique.
-For each couple, the pitch and intonation is varied leading to 25444 audio samples.
-
-The recorded instruments belongs to the wind and string classes.
-For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
-For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
-
-Some instruments can be complemented with sordina (S), wha (W).
-The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
-\vl{This is crucial and should not be deferred to the appendix. @writing}
-\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
-
-Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
-For  the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap  (explo), harmonic (harm), key-click (key), col legno   (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
-
-\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
-The abbreviations can be removed. @data @writing}
-\ml{}
-
-\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
-(y axis) would be extremely useful. @data @figure}
-\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
-\vl{Les donn\'{e} de l'\{e}tude, relatives aux Figures 5 et 6.\}
-
 \section*{Discussion}
@@
 
+%% Appendices
+
+\section*{Dataset}
+\label{sec:dataset}
+
+The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
+In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
+For each couple, the pitch and intonation is varied leading to 25444 audio samples.
+
+The recorded instruments belongs to the wind and string classes.
+For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+\ja{Better to display these in a table?}
+
+Some instruments can be complemented with sordina (S), wha (W).
+The addition of this kind of device modifies the shape of the instrument.
+Consequently, an alto saxophone and alto saxophone with sordina are considered as different instruments.
+\vl{This is crucial and should not be deferred to the appendix. @writing}
+\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
+
+Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+
+\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+The abbreviations can be removed. @data @writing}
+\ml{}
+
+\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+(y axis) would be extremely useful. @data @figure}
+\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+\vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-09-19 experiments/splitJudgments.m
fixing exp to new version of expLanes

@@ -1,0 +1,16 @@
+function [data, judgments] = splitJudgments(data, judgments, split, test)
+
+if ~exist('test', 'var'), test = 0; end
+
+if ~isempty(split)
+    rng(0);
+    
+    idx = randi(ceil(100/split), size(judgments, 2), 1)-1;
+    if test
+        idx = find(idx==0);
+    else
+        idx = find(idx);
+    end
+    data = filterData(data, idx);
+    judgments = judgments(:, idx);
+end

### 2019-09-19 experiments/tisiso3performance.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 1 2 1 0 5 1 0 0 2 2 2 1 1 3}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 0 1 1 2 5 1 1 2 2 2 2 1 0}); return; else store=[]; obs=[]; end
 
 rng(0);
 
@@ -25,13 +25,14 @@ if strcmp(setting.reference, 'judgments') && (...
     return
 end
 
-data1 = expLoad(config, '', 1);
+data1 = expLoad(config, '', 1, 'data');
 obs.p=[];
 switch setting.reference
     case 'judgments'
         [data1, judgments] = handleJudgments(config, data1);
         [data1, judgments] = splitJudgments(data1, judgments, str2num(setting.split), setting.test);
         parfor k=1:size(judgments, 1)  % par
+            k
             p(k) = process3performance(config, data1, data, setting, judgments(k, :), k);
         end
         obs.p = p;
@@ -60,16 +61,26 @@ switch setting.projection
                         data.projection = squeeze(mean(data.projection, 1));
                     case 2
                         for  k=1:size(data.projection, 1)
-                             subspaces(:, :, k) = squeeze(data.projection(k, :, :))^2;
-%                                 projection = squeeze(data.projection(k, :, :));
-%                             [subspaces(:, :, k), ~] = qr(projection);
+                            projMatrix = squeeze(data.projection(k, :, :));
+                            subspaces(:, :, k) = projMatrix*projMatrix';
                         end
                         meanSubSpace = psd_karcher_mean(subspaces);
+                        disp('done karcher');
                         data.projection = sqrtm(meanSubSpace);
+                        %                         for  k=1:size(data.projection, 1)
+                        %                              subspaces(:, :, k) = squeeze(data.projection(k, :, :))^2;
+                        %                                 projection = squeeze(data.projection(k, :, :));
+                        %                             [subspaces(:, :, k), ~] = qr(projection);
+                        %                         end
+                        %                         subspaces = (subspaces + permute(subspaces, [2 1 3]))/2;
+                        %                         meanSubSpace = psd_karcher_mean(subspaces);
+                        %                         data.projection = sqrtm(meanSubSpace);
                     case 3
[…]

### 2019-09-19 experiments/tisisoReport.m

@@ -6,18 +6,15 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'karcherSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'r', 'reportName', 'completeSlides'); return; end
 
 switch config.reportName
-    case 'karcherSlides'
-        mask = {2:4, 1, 2, 1, 0, 5, 1, 1, 2, 2, 2, 2, 1, 1};
-     config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 3, 'shortFactors', 1, 'uncertainty', -1);
    case 'completeSlides'
         
-      mask = {[3 4], [0], [1 2], 1, 0, 5, 1, 0, 0, 2, 2, 2, 1};
-      config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 4, 'expand', 'sct', 'uncertainty', -1);
-     mask = {[2 3 4], [0], [0], 1, 0, 5, 1, 1, 2, 2, 2, 2, 1};
-     config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 1, 'highlight', 0, 'precision', 3, 'expand', 'sct', 'shortFactors', 1, 'uncertainty', -1);
+      mask = {[3 4], [0], [1 2], 2, 2, 5, 1, 0, 0, 2, 2, 2, 1};
+      config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 4, 'expand', 'sct', 'uncertainty', -1);
+     mask = {[2 3 4], [0], [0], 2, 2, 5, 1, 1, 2, 2, 2, 2, 1};
+     config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3, 'expand', 'sct', 'shortFactors', 1, 'uncertainty', -1);
      
     case 'splitSlides'
   

### 2019-09-19 paper/paper.tex
Move human judgments section to backmatter

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Backmatter begins here                   %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{backmatter}
+
+\section*{Competing interests}
+  The authors declare that they have no competing interests.
+
+\section*{Author's contributions}
+    Text for this section \ldots
+
+\section*{Acknowledgements}
+The Flatiron Institute is a division of the Simons Foundation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                  The Bibliography                       %%
+%%                                                         %%
+%%  Bmc_mathpys.bst  will be used to                       %%
+%%  create a .BBL file for submission.                     %%
+%%  After submission of the .TEX file,                     %%
+%%  you will be prompted to submit your .BBL file.         %%
+%%                                                         %%
+%%                                                         %%
+%%  Note that the displayed Bibliography will not          %%
+%%  necessarily be rendered by Latex exactly as specified  %%
+%%  in the online Instructions for Authors.                %%
+%%                                                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bibliographystyle{bmc-mathphys}
+\bibliography{bib}
+
+%% Appendices
+
+\section*{Dataset}
+\label{sec:dataset}
+
+The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
+In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
+For each couple, the pitch and intonation is varied leading to 25444 audio samples.
+
+The recorded instruments belongs to the wind and string classes.
+For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+\ja{Better to display these in a table?}
+
+Some instruments can be complemented with sordina (S), wha (W).
+The addition of this kind of device modifies the shape of the instrument.
+Consequently, an alto saxophone and alto saxophone with sordina are considered as different instruments.
+\vl{This is crucial and should not be deferred to the appendix. @writing}
+\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
+
+Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+
+\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+The abbreviations can be removed. @data @writing}
+\ml{}
+
+\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+(y axis) would be extremely useful. @data @figure}
+\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+\vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
+
 \section*{Human timbre judgments}
@@
 
-\subsection{Reducing Expert judgments of technique peculiarities}
+\section*{Reducing Expert judgments of technique peculiarities}
+
+\ja{This needs to be summarized (possible in the previous section).
+I'm copying two sentences in here that I had in the Timbre similarity judgments section.}%
+These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
 
@@
 %%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
-\subsection{Crowdsourced judgments of pairwise similarity}
+\section*{Crowdsourced judgments of pairwise similarity}
+
+\ja{A lot of this is covered in the main text.
+Not sure what we need to add in the appendix.}
 
@@
 
-\subsection{Analysis}\label{sec:analysis}
+\section*{Analysis}\label{sec:analysis}
+
+\ja{I thought all of this stuff was going away and we were just replacing it with a brief analysis using the tag clouds (and so on).
+At any rate, the relevant analysis results should be in the main text (under Timbre similarity judgments, after we describe the experiments).}%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{Competing interests}
-  The authors declare that they have no competing interests.
-
-\section*{Author's contributions}
-    Text for this section \ldots
-
-\section*{Acknowledgements}
-The Flatiron Institute is a division of the Simons Foundation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{bmc-mathphys}
-\bibliography{bib}
-
-%% Appendices
-
-\section*{Dataset}
-\label{sec:dataset}
-
-The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
-In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
-For each couple, the pitch and intonation is varied leading to 25444 audio samples.
-
-The recorded instruments belongs to the wind and string classes.
-For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
-For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
-\ja{Better to display these in a table?}
-
-Some instruments can be complemented with sordina (S), wha (W).
-The addition of this kind of device modifies the shape of the instrument.
-Consequently, an alto saxophone and alto saxophone with sordina are considered as different instruments.
-\vl{This is crucial and should not be deferred to the appendix. @writing}
-\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
-
-Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
-For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
-
-\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
-The abbreviations can be removed. @data @writing}
-\ml{}
-
-\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
-(y axis) would be extremely useful. @data @figure}
-\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
-\vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-09-24 paper/paper.tex
directives ofr result section
refactor introduction

@@
 
-% Section outline:
-% - How to accurately model timbre space? Useful for studying auditory perception, browsing databases, etc.
-% - Studies either consider stimuli that are ``too simple'' (e.g., tones, instruments with fixed playing techniques) or ``too complex'' (e.g., speech, environmental sounds).
-% - Computational models do not capture enough information (MFCCs and related representations) or are hard to compute or analyze (STRFs, deep networks).
-% - Present a study of timbral similarity perception on diverse set of instrumental sounds with extended playing techniques (PTs). Provides more control in modeling different aspects of timbre.
-% - In addition, propose a computational model for timbral similarity: scattering (separable or joint) plus a linear layer.
-Fixed representation to extract time-frequency structure with adaptivity in linear component.
+Music information retrieval operates at either of two levels: symbolic and auditory.
+By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, or intensity at the source.
+Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis.
+As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} dry, rough, bright, round, and so forth) or by specifying an instrumental playing technique (\eg{} bowed or plucked).
 
-One of the most important properties for identifying the source of a sound is its \emph{timbre}.
-While it is typically defined as the quality of a sound that is independent of its pitch, duration, or loudness, this definition fails to provide an explicit characterization of the attribute.
-Due to its central role in audio perception, timbre has been studied from several angles, including perceptual and psychophysical studies.
-Since timbre extraction forms the basis of many audio-based information retrieval tasks, such as classification and determining acoustical similarity, the subject has also been studied from a machine listening perspective.
-\ja{Need references for the above statements.}
+Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
+On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them.
+On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
+For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in Italian.
 
-A natural candidate for characterizing timbre is the spectral content of a signal.
-This information, however, is not rich enough to fully characterize the timbre of a sound, as evidenced by the recent move from spectral descriptors such as mel-spectrograms and mel-frequency cepstral coefficients (MFCCs) and descriptors of spectral energy distributions.
-Indeed, the success of deep neural networks (DNNs) applied to time--frequency representations (such as spectrograms or scalograms) in machine listening have demonstrated the necessity for capturing both spectral and temporal content.
-In neurophysiological, the auditory perception models proposed by S.
-Shamma and collaborators, known as \emph{spectro--temporal receptive fields} (STRFs) all characterize temporal structure in addition to purely spectral information.
-It seems, therefore, that timbre is inherently spectro--temporal.
-\ja{Again, refs.}
+The prospect of retrieving timbre perception across playing techniques necessarily exceeds the symbolic domain.
+Instead, it involves a cognitive process which arises from the subjective experience of listening to music.
+From the standpoint of audio signal processing, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
+Rather than to merely discriminate instruments as mutually exclusive categories, this distance function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.
 
-While the above models offer some insight into the structure of timbre, their complexity makes interpretation and analysis difficult.
-The expressive power of DNNs are due to tuning a large number of parameters, so it is often hard to relate a specific component to a property of the resulting perception model.
-STRFs, on the other hand, have a simpler structure, but are ultimately constructed to model electrophysiological signal in the auditory cortex.
-As a result, they consist several adjustments that make their computation costly and their mathematical analysis difficult.
-\ja{Refs.}
+Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation.
+Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
+The outcome of this is an instrumental audio sample, alongside textual attributes of playing technique.
+In this context, the computer serves as a bridge from the auditory level to the symbolic level, allowing to approximate a potentially infinite realm of timbral sensations by an orchestra of priorly specified extent and technical range.
+To maximize the perceptual fidelity of such an approximation, a query-by-example search engine in the auditory domain should summarize the vastest possible vocabulary of timbral templates.
 
-Another difficulty in studying the perception of timbre is the choice of stimuli.
-Previous work on timbre characterization of STRFs \cite{patil2012music} has focused on instrumental sounds of isolated notes played with standard playing techniques.
-This inherently restricts the range of available timbre studied, limiting the conclusions of the work.
-On the other end of the spectrum are computational models for information retrieval tasks.
-Since these are primarily concerned with performance on real-world sounds, the possible range of timbre is very broad.
-The analysis of the resulting model is therefore all the more difficult, since we have little control over the complexity of the stimuli.
-\ja{Refs.}
+In this paper, we propose a machine listening model which computes the dissimilarity in timbre between any two audio samples.
+The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
+Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
+A second originality of our model is that we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
+Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
 
-Our contribution is twofold.
-First, we propose a simple model for timbre perception involving a fixed time--frequency descriptor, the scattering, followed by a learned linear layer optimized for the timbral space under consideration.
-Second, we validate this model on a set of instrumental sounds with extended playing techniques (PTs), allowing for a richer range of timbre compared to ordinary instrumental sounds while retaining the structure of the space necessary for interpretability.
+%% TODO: list Sections
 
-As part of validating the model, we have collected timbral similarity judgments in a crowdsourced listening experiment.
-These are used as a ``ground truth'' dataset, which allow us to evaluate the model's ability to characterize the perceived similarity of the stimuli.
-Analysis of these experimental results also confirms the necessity of including a range of PTs, as these are shown to influence similarity judgments to the same extent as the identity of the instrument itself.
-
-We conclude with a discussion of potential applications to browsing interfaces for musical databases, where composers can easily find sounds on a query-by-example basis.
-The adaptivity of the proposed model has the additional advantag of allowing it to be adjusted to a particular listener's preferences by retraining the linear layer.
+%% TODO: mention companion website and open-source software
 
@@
 These need to be extracted and summarized here.}
+
+post processing
+
+std consisten across features
+mfcc cut
+consistent with [previous papers] each item provides incremental gain of perforamnces
+
+table 3 valeurs erreur
+
+ablation study
+
+tfscat+lmnn
+tfscat
+scat
+
+9 pas de weighting averaging not end to end
+
+separate pas trop de perte : consensus bon
+
+perf du best tool chain on complete dataset
+
+randomize
+
+perf on 50/50
+
+lda
+
+-lmnn
+
+tf -> scat
+
+tfscat -> mfcc + etendus
 

### 2019-09-24 paper/paper.tex
clarify intro

@@
 On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
-For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in Italian.
+For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
 
-The prospect of retrieving timbre perception across playing techniques necessarily exceeds the symbolic domain.
-Instead, it involves a cognitive process which arises from the subjective experience of listening to music.
-From the standpoint of audio signal processing, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
+The prospect of modeling timbre perception necessarily exceeds the symbolic domain.
+Instead, it involves a cognitive process which arises from the subjective experience of listening.
+From an audio signal processing standpoint, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
 Rather than to merely discriminate instruments as mutually exclusive categories, this distance function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.

### 2019-09-25 paper/paper.tex
start results section, add figure to show consensus

@@
 \usepackage{xspace}
-
-\def\includegraphic[#1]{}
-\def\includegraphics[#1]{}
+\usepackage{graphicx}
+% \def\includegraphic[#1]{}
+% \def\includegraphics[#1]{}
 
@@
    email={email@example.com}            % email address
-]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
+]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
 \author[
@@
    email={email@example.com}
-]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
+]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
 \author[
@@
 ]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
+\author[
+   addressref={aff2},
+   email={email@example.com}
+]{\inits{JA}\fnm{Joakim} \snm{And\`en}}
 \author[
@@
 To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
-The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch and nuance (for more information on the dataset, see the Dataset section).
-\ja{Mathieu, which pitch and nuance?}%
-In the survey, a subject asked to dots on a two-dimensional place in a free sorting task.
-Each dot corresponded to a different IPT recording and was moved by the user dragging them with the mouse.
-In addition to spatially arranging the dots, the user was also asked to assign one of $20$ colors to each dot, providing a simple clustering of the sounds.
-While both the location and color of each dot was recorded, only the color was used in our subsequent analysis---the spatial arrangement was only used to facilitate the user's ability to organize the sounds.
-More information on the data collection process is available in the Human timbre judgments section.
+The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
+For more information on the dataset considered in this study, see Section \ref{sec:dataset}.
+In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
+Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
+In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple clustering of the sounds.
+While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
+More information on the data collection process is available in  Section \ref{sec:judgments}.
 
@@
 
-For each subject, we constructed a matrix of size $78 \times 78$ to represent their similarity judgments for each pair of the $78$ IPTs.
+For each subject, a matrix of size $78 \times 78$ is constructed to represent the similarity judgments for each pair of the $78$ IPTs.
 Specifically, the $(i, j)$th element of the matrix was set to one if the $i$th and $j$th IPT were assigned the same color.
-If not, it was set to zero.
-In other words, we record the clustering assignment of each subject as a binary similarity matrix.
+If not, it is set to zero.
+In other words, the clustering assignment of each subject as a binary similarity matrix.
+Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
+In this case, the corresponding binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
 
-\ml{Joakim, equation needed ?}
-\ja{I don't think so.
-It should be fairly straightforward.}%
+It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
+We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
+Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other, see Figure \ref{fig:consensusVsIpt}.
 
-\ml{also have the resulting binary similarity matrix ?}
-\ja{Yes, but I think we should have it for the consensus clustering described below.}%
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{../code/xp2/consensusVsI.png}
+\caption{'ASax'    'BBTb'    'BbCl'    'Bn'    'Cb'    'Fl'    'Gtr'    'Hn'    'Hp'    'Ob'    'TTbn'    'TpC'  'Va'    'Vc'}
+\label{fig:consensusVsIpt}
+\end{figure}
 
-Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
-In this case, the correspondign binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
-It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
-That being said, conditioning the timbre space using several binary similarity matrices is not a trivial task, and will be discussed extensively in the Results section.
-\ja{Why discuss it extensively there?
-Don't we just settle on the partitioning algorithm below?}%
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{../code/xp2/consensusVsPt.png}
+\caption{ 'art-harm'    'blow-no-reed'    'explo-slap'    'harm-fngr'    'key-cl'    'legno-batt'    'legno-tratto'
+    'nonvib'    'ord'    'ord-closed'    'ord-hi-reg'    'ord-open'    'pdl-tone'    'pizz'    'pizz-bartok'
+    'pizz-lv'    'pizz-sec'    'pont'    'slap'    'slap-unp'    'tasto'    'tng-ram'}
+\label{fig:consensusVsIpt}
+\end{figure}
 
-We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
-Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
 \ja{We should have some figure to back this up.
 Can this be done using the tag clouds?}
+\ml{tentative solution using similarity matrices, see fig, if ok should be improved by vl (code/xp2/consensusVsIpt.m)}
 
@@
 \label{sec:results}
+
+The processing pipeline presented in Figure \ref{fig:pipeline} addresses the problem of conditioning a timbre space using perceptual judgments of timbral similarity.
+The relevance of the many design choices that had to be made  are addressed in this section using an experimental protocol that considers a dataset of audio that will be given as input to the processing pipeline and one metric used for evaluating its performance under a given experimental setting.
+
+As gathering perceptual similarity data is plagued with a size problem.
+the perceptual data gathered during the perceptual experiment considers 78 \ipts each represented by one sound recording.
+In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is lower than the timbral similarity constraints expressed by the subjects.
+Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre, it is  thus taken in this paper as a reasonable assumption.
+Quantification of this invariance is left for further study.
+
+This result in considering for each of the 78 \ipts the whole range of nuance and pitch, leading to a dataset composed of 9346 sound recordings where two sounds of this dataset $a$ and $b$, corresponding to \ipts $I_a$ and $I_b$ are considered as similar with respect to a given reference clustering $c_i$ if the \ipts $I_a$ and $I_b$ are grouped in the same cluster.
+
+The metric is the average precision at rank $k=5$ (ap@5), that allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
+
+The reference considered here is thus the $31$ clusterings obtained with the perceptual experiment, see Section \ref{sec:judgments}.
+for each of those clusterings, the precision at rank $5$ is computed, and the $31$ p@5 are averaged to obtain the ap@5.
+
+\ml{note pour vincent, on a pas de fold, mais on a les resultats par sujet, on peut donc evaluer la dispersion de la p@5 par sujets si cela est utile.}
+
+The choice of the rank is chosen with the application in mind.
+When searching for sounds that are similar to a query, the listener is likely to listen to up to $5$ proposals, rarely more.
+For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision is not considered here.
+
+The processing pipeline is composed of 2 main building blocks. The first takes a digital audio recording as input and outputs one vector of features.
+
+The second considers one vector of features as input and another vector of project features as output.
 
@@
 
-As only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
 The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
@@
 This results in $31$ clusterings that can be considered for defining a perceptual timbre space of \ipts.
-
-\section*{Analysis}\label{sec:analysis}
-
-\ja{I thought all of this stuff was going away and we were just replacing it with a brief analysis using the tag clouds (and so on).
-At any rate, the relevant analysis results should be in the main text (under Timbre similarity judgments, after we describe the experiments).}%
-
-In order to gain information about the relevance those $31$ clusterings for defining a perceptual timbre space of \ipts, an analysis is performed.
-
 Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipts.
@@
 \label{fig:xp2sizeGroup}
-\end{figure}
-
-A clustering can be considered as a binary similarity measure, with $s(a, b) = 1$ if $a$ and $b$ belong to the same group, and $0$ otherwise.
-By averaging the 31 binary similarity matrix, an overall floating point matrix is obtained that corresponds to our measure of the spontaneous perceptual similarity between \ipt.
-
-In order to gather information about the properties of the data, a cluster analysis is next performed using an agglomerative hierarchical clustering \cite{gordon1987review} using the weighted average algorithm for computing distance between clusters.
-The resulting dendrogram is displayed on Figure \ref{fig:dendrogram}.
-A partition of $n$ clusters can be constructed from the agglomerative hierarchical cluster tree by finding the smallest height at which a horizontal cut through the tree leaves $n$ clusters.
-
-One question that arises is whether the similarity is influenced primarily by the instrument or by the playing technique.
-To answer this question, each partition provided by cutting the dendrogram at successive levels is compared to two reference partitions.
-One reference is the partition of the \ipts when labeled with the instrument used (I) and the other if the partition of the \ipts when labeled with the playing technique used (Pt).
-
-There are many metrics available to compare two partitions by evaluating their degree of correspondence \cite{wagner2007comparing}.
-The normalized mutual information (NMI) is interesting as it do not require alignement of the groups labels between the two partitions prior correspondence analysis and the two partition are not required to have the same number of groups.
-It also by design bounded between 0 and 1, 1 meaning perfect match.
-We consider the normalization proposed in \cite{strehl2002cluster}.
-Figure \ref{fig:clusters} shows the evolution of the NMI when comparing the partitions provided by the cluster analysis of the dendrogram produced using the human judgment of similarity and respectively the instrument partition ($j/I$) or the playing technique partition ($j/Pt$).
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
-\caption{Cluster analysis of the similarity judgments.
-$j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition.
-Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
-\vl{This line plot is very hard to understand.
-The legend box occludes the data on the top-right corner.
-It would be preferable to remove this legend box and label lines
-by placing text upon them,
-so that the plot would be legible to readers with color vision deficiencies.
-Besides, the choice of line labels is cryptic, as it relies on abbreviations.
-The distinction between I and Pt
-can be illustrated by two complementary colors, \eg{} blue versus orange.
-The distinction between j, nj, and null
-can be illustrated by three distinct markers
-(empty circle, filled triangle, asterisk),
-associated with
-three distinct line styles (solid, dashed, dotted).
-The NMI abbreviation should be expanded.
-@figure}}
-\ml{ok, le .fig est dispo}
-\label{fig:clusters}
-\end{figure}
-
-There is a clear gain of NMI when considering the playing technique which means that the subjects primarily used cues corresponding to the playing technique to organize the \ipts.
-Also, for both curves, a global trend of increase in the $j/x$ partitions with respect to the number of clusters can be observed.
-This is a well documented fact, which simply put, means that it is easier to match any given partition with more clusters at hand \cite{tibshirani2001estimating}.
-To compensate for this bias, one can consider the NMI that would be achieved by comparing to the reference partition several random partitions of increasing number of clusters.
-The resulting averaged NMI for 100 randomly generated partitions is displayed on Figure \ref{fig:clusters}.
-This so-called null partitions NMI are respectively termed $nullI$ and $nullPt$ depending on the reference used.
-To normalize $j/I$ and $j/Pt$ in order to reduce the impact of this bias, we simply subtract the $null$ NMI:
-\begin{eqnarray}
-  nj/I &=& j/I - null \\
-  nj/Pt &=& j/Pt - null  \\
-\end{eqnarray}
-This normalization is useful for identifying more easily the number of clusters for which the comparison is most relevant.
-As can be seen on Figure \ref{fig:clusters}, the number of clusters that leads to the highest NMI is 5 for the playing technique reference and 8 for the instrument reference, leading respectively to an NMI of $0.46$ and $0.21$.
-For the latter, reducing the number of clusters down to 6 do not decrease the NMI by much, a fact that will be latter discussed in more details.
-Even by considering the null normalization, the matching level is much higher for the playing technique reference.
-
-The correspondences between those partitions and the reference ones are now studied.
-For this purpose, Figure \ref{fig:gi} displays two partitions.
-At the bottom is the partition that corresponds to the instruments labels, whose color code is given by the color bar on the right.
-The abbreviations used for the instruments are detailed in Section \ref{sec:dataset}.
-The ordering of \ipts on the horizontal axis is the one given by the cluster analysis, see the dendrogram on Figure \ref{fig:dendrogram}.
-The partition on top is the result of the clustering with 8 clusters.
-For visualization purposes, we arbitrarily set colors to clusters of this partition to the color code of the instrument label that is most often encountered in this cluster.
-If some labels occur equally often, the instrument of the first \ipt of the cluster is chosen.
-The emerging instrument classes are: Saxophone, Clarinet, Trumpet, Flute and Cello.
-
-However, even with an optimal number of clusters, we see that the clusters have  a low purity, \ie{} a very diverse set of instruments labels for the \ipt they belong to.
-Also, some clusters could be merged without strong changes in the overall organization.
-This probably explains the small changes of NMI when reducing the number of clusters, see Figure \ref{fig:clusters}.
-
-As can be seen on Figure \ref{fig:gm}, the correspondence between the playing technique reference and the resulting partition with 5 clusters is much higher.
-Most clusters have a high level of purity and for groups emerge : Pizzicato, Slap, Ordinario, and Ponticello.
-
-To conclude this section, the process of gathering data about the perceptual similarity of \ipt is found to be successful as the subjects used their degree of freedom in a consistent manner and no inconsistency in the data has been detected during the analysis process.
-The main result is that the playing technique is an important factor when modeling the spontaneous judgment of similarity gathered in the second experiment.
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/groupModes.png}
-\caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top).
-The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.
-\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
-\ml{Je demande a voir, mais je ne suis pas satisfait de cette visu de toute facon. Le code est dispo dans le rep code. Lancer ticelSim en premier.}
-\label{fig:gm}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/groupInstruments.png}
-\caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b).
-The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.
-\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
-\ml{ok}
-\label{fig:gi}
 \end{figure}

### 2019-09-25 code/xp2/consensusVsIpt.m

@@ -1,0 +1,41 @@
+load ticelJudgments
+load clusters
+
+ins = unique(instruments);
+mod = unique(modes);
+
+for k=1:length(instruments)
+   in(k) =  find(ismember(ins, instruments{k}));
+   mo(k) =  find(ismember(mod, modes{k}));
+end
+
+[sin, iin] = sort(in);
+[smo, imo] = sort(mo);
+
+
+
+figure(1)
+colormap bone
+imagesc(squareform(pdist(ensemble(iin).', 'jaccard')))
+hold on
+sep = find(diff(sin));
+for k=1:length(sep)
+   line([0, 78], [sep(k), sep(k)]) 
+    line([sep(k), sep(k)], [0, 78]) 
+end
+hold off
+instruments(iin(sep))
+saveas(gcf, 'consensusVsI.png')
+
+figure(2)
+colormap bone
+imagesc(squareform(pdist(ensemble(imo).', 'jaccard')))
+sep = find(diff(smo));
+for k=1:length(sep)
+   line([0, 78], [sep(k), sep(k)]) 
+    line([sep(k), sep(k)], [0, 78]) 
+end
+hold off
+modes(imo(sep))
+saveas(gcf, 'consensusVsPt.png')
+

### 2019-09-26 paper/paper.tex
1/2 results done

@@
 
-As gathering perceptual similarity data is plagued with a size problem.
-the perceptual data gathered during the perceptual experiment considers 78 \ipts each represented by one sound recording.
-In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is lower than the timbral similarity constraints expressed by the subjects.
-Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre, it is  thus taken in this paper as a reasonable assumption.
+\subsection*{Data}
+
+The perceptual data gathered during the perceptual experiment considers 78 \ipts each represented by one sound recording that the subject could listen to before clustering the \ipts.
+For those 78 \ipts, 31 clusterings $C_k, 0<k<31, Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
+If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
+In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings.
+Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.
+
+In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
+Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation}, it is  thus taken in this paper as a reasonable assumption.
 Quantification of this invariance is left for further study.
 
-This result in considering for each of the 78 \ipts the whole range of nuance and pitch, leading to a dataset composed of 9346 sound recordings where two sounds of this dataset $a$ and $b$, corresponding to \ipts $I_a$ and $I_b$ are considered as similar with respect to a given reference clustering $c_i$ if the \ipts $I_a$ and $I_b$ are grouped in the same cluster.
+The whole range of nuance and pitch is thus considered for each of the 78 \ipts, leading to a dataset composed of 9346 sound recordings.
+The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, Card(D_k)=9346$ using the following scheme.
+If a sound of this dataset $s_a$ corresponds to \ipt $i$, then $D_k(a) = C_k(i), \forall k$.
 
-The metric is the average precision at rank $k=5$ (ap@5), that allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
+\subsection*{Metric}
 
-The reference considered here is thus the $31$ clusterings obtained with the perceptual experiment, see Section \ref{sec:judgments}.
-for each of those clusterings, the precision at rank $5$ is computed, and the $31$ p@5 are averaged to obtain the ap@5.
+The metric considered in this paper is the average precision at rank $k=5$ (ap@5).
+It allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
+The ap@5 is computed as follows.
+
+Considering an output of the system, that is a features vector $F_n$ per sound $s_n$, the precision at rank $5$ is computed for each of the $D_k$ clusterings, and the $31$ p@5 are averaged to obtain the ap@5.
+
+For a given reference clustering $D_k$, the p@5 is computed as follows.
+For each sound $s_q, 0<q<9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n<5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors.
+For this query, the precision is defined as:
+$$
+p@5_q = \frac{\sum_{n=5}^5 D_q==D_n}{5}
+$$
+The p@5 is the average $p@5_q$ computed over all queries.
 
@@
 
-The processing pipeline is composed of 2 main building blocks. The first takes a digital audio recording as input and outputs one vector of features.
+\subsection*{Performance analysis}
 
-The second considers one vector of features as input and another vector of project features as output.
+The processing pipeline is composed of 2 main building blocks.
+The first block $B_1$ takes a digital audio recording as input and outputs one vector of features.
+For this block, the main factors are the time support $B_1^t$ for computing the features and the type of features $B_1^f$.
+We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{mfccs}, \text{separable scattering}, \text{joint scattering} \}$.
+The mfccs are the standard Mel frequency cepstral coefficients with 40 Mel bands and selection of the 13$^\text{th}$ first discrete cosine transform (dct) coefficients.
 
-\ja{The results are currently sprinkled into the above methods section.
-These need to be extracted and summarized here.}
+\ml{provide parameters for scattering}
 
-post processing
+As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
 
-std consisten across features
-mfcc cut
-consistent with [previous papers] each item provides incremental gain of perforamnces
+For each sound, a set of features are computed for each time support. This set is averaged over time to give one feature vector per sound.
+All features are standardized with zero mean and unit variance across the dataset.
 
-table 3 valeurs erreur
+The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output.
+For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$.
+The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{}.
+For this algorithm the neighborhood is set to $5$.
 
-ablation study
+The complete experiments with all settings have been computed. For the sake of clarity, only part of the results are discussed here.
 
-tfscat+lmnn
-tfscat
-scat
+The best performing system considers the separable scattering computed over $1000$ms time support, projected using the lmnn algorithm.
+Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.5\%$ of $ap@5$.
+That means that, for a given sound dataset of \ipts for which subjective evaluation of similarity between \ipts are available, a very large number of queries would return interesting results to the user using this processing pipeline.
 
-9 pas de weighting averaging not end to end
+In order to better understand the benefits of each design choice, an ablation study \ml{vl ref ??} is now performed using a slightly different protocol.
+In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those.
+In our case, two factors can readily be identified, the variability of the audio data and the one of the perceptual judgments.
 
-separate pas trop de perte : consensus bon
+We mitigate the first by considering a random split of the dataset in two halves, one for learning the projection and one for evaluating the performance.
+The impact of the variability of the perceptual judgments and its impact on machine listening systems such as the one discussed in this paper is not in the scope of this paper and left for further research.
+By considering this protocol, the proposed system reaches a $ap@5$ of $95.5\%$.
+If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\%$, indicating a low risk of overfitting.
 
-perf du best tool chain on complete dataset
++rand
 
-randomize
-
-perf on 50/50
-
-lda
+lmnn -> lda
 
@@
 
-tf -> scat
+time support
 
-tfscat -> mfcc + etendus
+tfscat -> scat
+
+tfscat -> mfcc
+
+tfscat -> mfcc monomials
 

### 2019-09-26 experiments/getFeatures.m
needed data for result section

@@ -20,7 +20,7 @@ switch setting.reference
             otherwise
                 nbItems = length(data.mode);
                 idx = randi(nbItems, ceil(nbItems*str2num(setting.split)/100), 1);
-                if step==2 && setting.test
+                if step==3 && setting.test
                     idx = setdiff(1:nbItems, idx);
                 end
         end

### 2019-09-26 experiments/tisiso2project.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso2project(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 2, 'mask', {2 3 3 5 0 5 1 1 2 0 0 2 1 2 0}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 2, 'mask', {4 5 2 5 0 5 1 1 2 0 0 2 1 2 0}); return; else store=[]; obs=[]; end
 
 rng(0);
 
@@ -32,7 +32,7 @@ switch setting.reference
     case 'judgments'
         [data, judgments] = handleJudgments(config, data, setting.averageJudgment);
         [data, judgments] = splitJudgments(data, judgments, str2num(setting.split));
-        for k=1:size(judgments, 1)
+        parfor k=1:size(judgments, 1)
             ju = judgments(k, :);
             projection(k, :, :) = process2project(store, config, data, setting, ju);
         end      

### 2019-09-26 experiments/tisiso3performance.m

@@ -11,13 +11,13 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 0 1 1 2 5 1 1 2 2 2 2 1 0}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 5 2 1 0 5 1 1 2 2 2 2 1 2}); return; else store=[]; obs=[]; end
 
 rng(0);
 
 if strcmp(setting.reference, 'judgments') && (...
         (strcmp(setting.projection, 'lda')  && (setting.averageJudgment==0 || (setting.averageJudgment==1 && setting.separateJudgment==0))) || ...
-        (strcmp(setting.projection, 'none')  && (setting.separateJudgment==1 || (setting.averageJudgment==0 && setting.separateJudgment==1))) || ...
+        (strcmp(setting.projection, 'none')  && (setting.separateJudgment==1 || setting.averageJudgment==1 || (setting.averageJudgment==0 && setting.separateJudgment==1))) || ...
         (strcmp(setting.projection, 'lmnn')  && setting.averageJudgment==1 && setting.separateJudgment==1) || ...
         strcmp(setting.split, 'octave') ...
         )
@@ -32,7 +32,6 @@ switch setting.reference
         [data1, judgments] = handleJudgments(config, data1);
         [data1, judgments] = splitJudgments(data1, judgments, str2num(setting.split), setting.test);
         parfor k=1:size(judgments, 1)  % par
-            k
             p(k) = process3performance(config, data1, data, setting, judgments(k, :), k);
         end
         obs.p = p;

### 2019-09-26 experiments/tisisoReport.m

@@ -6,71 +6,112 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'r', 'reportName', 'completeSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'cSlides'); return; end
 
 switch config.reportName
-   case 'completeSlides'
+    case 'tt'
+    
+    case 'dg'
+    mask =   {[2  3  4], 5, 1, 1, 1, 5, 1, 1, 0, 0, 0, 0, 1, 1, 1};
+        caption = ' post processing of features';
+        config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'precision', 3, 'expand', 'split', 'uncertainty', -1, 'highlight', -1, 'caption', caption, 'save', 'cs1');
+       
+    case 'cSlides'
+        % post processing of features
+        mask =   {[2  3  4], 5, 1, 1, 1, 5, 1, 1, 0, 0, 0, 0, 1, 1, 1};
+        caption = ' post processing of features';
+        config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'precision', 3, 'expand', 'split', 'uncertainty', -1, 'highlight', -1, 'caption', caption, 'save', 'cs2');
+   
+        % expand
+        mask = {2:4, 5, [1  2], 2, 2, 5, 1, 0, 2, 2, 2, 2, 1, 1, 1};
+        caption = 'expand mfccs using monomials up to the dimensionality of the scattering';
+        config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'precision', 3, 'highlight', -1, 'caption', caption, 'save', 'cs3');    
         
-      mask = {[3 4], [0], [1 2], 2, 2, 5, 1, 0, 0, 2, 2, 2, 1};
-      config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 4, 'expand', 'sct', 'uncertainty', -1);
-     mask = {[2 3 4], [0], [0], 2, 2, 5, 1, 1, 2, 2, 2, 2, 1};
[…]

### 2019-09-26 paper/paper.tex
Fix name and affiliation for JA
Write discussion section

@@
 \author[
-   addressref={aff2},
-   email={email@example.com}
-]{\inits{JA}\fnm{Joakim} \snm{And\`en}}
+   addressref={ccm},
+   email={janden@flatironinstitute.org}
+]{\inits{JA}\fnm{Joakim} \snm{And\'en}}
 \author[
@@
   \cny{Germany}
+}
+\address[id=ccm]{%
+  \orgname{Center for Computational Mathematics, Flatiron Institute},
+  \street{162 5th Avenue},
+  \city{New York, NY 10010},
+  \cny{USA}
 }
@@
 
-\begin{itemize}
-\item Demonstrates complex timbral structure perceived by subjects in study: neither instrument (i.e., spectral envelope) or playing technique (i.e., temporal modulation) dominates similarity judgment.
-Need joint to characterize time-frequency structure (see Patil et al).
-\item Proposed computational model (scattering + linear) reproduces similarity judgments accurately.
-Easy to ``retrain'' for other data.
-Fixed wavelet structure allows for analysis of linear layer and guarantees invariance and stability properties.
-\item Applications: query-by-example in musical sample databases.
-\end{itemize}
+The success of the above system in identifying timbral similarities has immediate applications in browsing music databases.
+These are typically organized based on instrumental and PT taxonomies, with additional keywords offering a more flexible organization.
+Accessing the sounds in these databases therefore requires some knowledge of the taxonomy and keywords used.
+Furthermore, the user needs to have some idea of the particular IPT they are searching for.
+
+As an alternative, content-based searches allow the user to identify sounds based on similarity to some given query sound.
+This query-by-example approach provides an opportunity to search for sounds without having a specific instrument of PT in mind, yielding a wider range of available sounds.
+A composer with access to such a system would therefore be able to draw on a more diverse palette of musical timbre.
+
+The computational model proposed in this work is well suited to such a query-by-example task.
+We have shown that it is able to approximate well the timbral judgments of a wide range of subjects included in our study.
+Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on a reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
+This model can then be applied to new data, or, alternatively, be retrained with these new examples if the existing model proves unsatisfactory.
+
+We shall note, however, that the current model has several drawbacks.
+First, it is only applied to instrumental sounds.
+While this has the advantage of simplifying the intepretation of the results, the range of timbre under consideration is necessarily limited (although less restricted than only considering \emph{ordinario} PTs).
+This also makes applications such as \emph{query-by-humming} difficult, since we cannot guarantee that the timbral similarity measure is accurate for such sounds.
+
+That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
+Indeed, the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed and work for more general classes of sounds.
+To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
+These can again be obtained by performing a new clustering experiment with one or more human subjects.
+
+Another aspect is the granularity of the similarity judgments.
+In the above method, we have used \emph{hard} clustering assignments to build our model.
+A more nuanced similarity judgment would ask users to rate the similarity of a pair of IPTs on a more graduated scale, which would yield a finer, or \emph{soft}, assignment.
+This however, comes with additional difficulties in providing a consistent scale across subjects, but could be feasible if the goal is to only adapt the timbral space to a single individual.
+An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assigments.
 

### 2019-09-27 paper/paper.tex
first version of the results section

@@
 
-\subsection*{Performance analysis}
+\subsection*{Analysis}
 
@@
 
-\ml{provide parameters for scattering}
+\ml{provide parameters for scattering
+\texttt{
+case 'scat'
+    filt\_opt.Q = [12 1];
+    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
+case 'tfscat'
+    tm\_filt\_opt.Q = [12 1];
+    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
+    fr\_filt\_opt.J = 4;}}
 
@@
 
-+rand
+Another risk when considering reprojection techniques such as the lmnn may appear when the dimensionality of the features gets closer to the  number of items in the dataset.
+For example there exists a trivial solution for optimizing the p@1 when the dimensionality of the features is equal to the number of items in the dataset.
+In this case, no information needs to be carried by the features set to obtain perfect performance.
 
-lmnn -> lda
+In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
+In this case, the performance drops down to $23.6\%$.
+Not considering the learning block gives a $aP@5$ of $23.8\%$.
+It can be concluded from these results that the learning block is not able to perform effectively if no information is carried by the input features set.
 
--lmnn
+The slight decrease of performance when considering lmnn may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
 
-time support
+The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
+One of those is the linear discriminant analysis \ml{ja ref ?}\cite{}.
+When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.
 
-tfscat -> scat
+The lower performance of the lda may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from 1180 to 20.
 
-tfscat -> mfcc
+Considering a learning block implies some constraints.
+Performing the learning on large datasets require some computation.
+The projection has to be updated if the reference clustering or the dataset are modified significantly.
+Not considering any learning block reduces the performance to $92.9\% \pm 3$.
 
-tfscat -> mfcc monomials
+Concerning the feature computation block $B_1$, an important factor is the time support.
+When considering the commonly used duration of 25ms, the performance drop down to \ml{check value} $\%$.
+This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture long term modulations which plays an important part when considering extended playing techniques.
+
+Considering the separable scattering gives a performance of $93\% \pm 4.2$.
+
+Simplifying further the feature set by considering the mfccs gives a performance of $82.3\% \pm 7.1$.
+Considering the mfccs is equivalent to considering the separable scattering at order 1.
+Compared to the separable scattering, the information captured by the mfccs is considerably reduced.
+In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the mfccs are complemented with monomials in order to reach the dimensionality of the joint scattering.
+To do so, the 40 features $f_i$ of the mfccs are multiplied together $f_{i, j} = f_i*f_j, i \neq j$.
+The resulting features set is composed of the 40 mfccs features $f_i$ and the first 1140 $f_{i, j}$.
+The resulting aP@5 of $81.8\% \pm 7$ indicates that the increase of dimensionality is likely to have no effect, confirming the relevance of the additional information provided by the joint scattering.
+\ml{vl outro on modulations}
 

### 2019-09-27 experiments/tisisoReport.m
update results

@@ -53,7 +53,7 @@ switch config.reportName
         config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'precision', 3, 'expand', 'split', 'uncertainty', -1, 'highlight', -1, 'caption', caption);  
         
         % complete
-        mask = {[2 3 4], [0], [1 2], 1, 2, 5, 1, 1, 2, 2, 2, 2, 1};
+        mask = {[2 3 4], [0], [1 2], 2, 2, 5, 1, 1, 2, 2, 2, 2, 1};
         config = expExpose(config, 't', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3, 'expand', 'sct', 'shortFactors', 1, 'uncertainty', -1, 'save', 'cs6');
         config = expExpose(config, 'p', 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 4, 'expand', 'sct', 'uncertainty', -1);
         

### 2019-09-28 experiments/tisiso3performance.m
fix exp

@@ -11,12 +11,12 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {3 5 2 1 0 5 1 1 2 2 2 2 1 2}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {[4]    [5]    [3]    [2]    [2]    [5]    [1]    [1]    [2]    [2] [2]    [2]    [1]    [2]    [1]}); return; else store=[]; obs=[]; end
 
 rng(0);
 
 if strcmp(setting.reference, 'judgments') && (...
-        (strcmp(setting.projection, 'lda')  && (setting.averageJudgment==0 || (setting.averageJudgment==1 && setting.separateJudgment==0))) || ...
+        (strcmp(setting.projection, 'lda')  && (setting.averageJudgment==0 || (setting.averageJudgment==1 && setting.separateJudgment==1))) || ...
         (strcmp(setting.projection, 'none')  && (setting.separateJudgment==1 || setting.averageJudgment==1 || (setting.averageJudgment==0 && setting.separateJudgment==1))) || ...
         (strcmp(setting.projection, 'lmnn')  && setting.averageJudgment==1 && setting.separateJudgment==1) || ...
         strcmp(setting.split, 'octave') ...

### 2019-09-28 experiments/tisisoReport.m

@@ -6,11 +6,29 @@ function config = tisisoReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 09-Jan-2017
 
-if nargin==0, timbralSimilaritySol('report', 'rcv', 'reportName', 'cSlides'); return; end
+if nargin==0, timbralSimilaritySol('report', 'r', 'reportName', 'tt'); return; end
 
 switch config.reportName
     case 'tt'
-    
+        % full
+        mask = {4, 5, 2, 1, 2, 5, 1, 1, 2, 2, 2, 2, 1, 2, 1};
+        config = expExpose(config, 'l', 'save', ['tt0'], 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3);
+        mask = {4, 5, 2, 2, 2, 5, 1, 1, 2, 2, 2, 2, 1, 2, 1};
+        factors  = [1 7 3 3 2 1 1];
+        modalities = [4 2 3 1 5 3 2];
+        captions = {'split', 'random', 'lda', 'nolearning', '25ms', 'separable', 'mfcc', 'monomials'};
+        for k=1:length(factors)
+            k
+            captions{k}
+            mk=mask;
+            mk{factors(k)}= modalities(k)
+            config = expExpose(config, 'l', 'save', ['tt' num2str(k)], 'step', 3, 'mask', mk, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3);
+        end
+         % monomials
+         captions{end}
+        mask = {2, 5, 2, 2, 2, 5, 1, 2, 2, 2, 2, 2, 1, 2, 1};
+        config = expExpose(config, 'l', 'save', ['tt' num2str(length(factors)+1)], 'step', 3, 'mask', mask, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3);
+      
     case 'dg'
     mask =   {[2  3  4], 5, 1, 1, 1, 5, 1, 1, 0, 0, 0, 0, 1, 1, 1};
         caption = ' post processing of features';

### 2019-09-28 experiments/processResults.m
final display of results

@@ -1,0 +1,22 @@
+clear all
+captions = {'full', 'split', 'random', 'lda', 'nolearning', '25ms', 'separable', 'mfcc', 'monomials'}
+for k=1:length(captions)
+   d(k) = load(['report/tables/tt' num2str(k-1)]);
+   p(k, :) = d(k).data.rawData{1}.p;
+   x(k, :) = k*ones(1, size(p, 2));
+   c(k, :) = 1:size(p, 2);
+end
+
+mean(p')
+fig = figure(1);
+boxplot(p')
+xticklabels(captions);
+saveas(fig, 'boxRes.png')
+
+fig = figure(2);
+
+s=scatter(x(:), p(:), 'filled')
+s.CData = c(:);
+s.SizeData = 50;
+saveas(fig, 'samplesRes.png')
+

### 2019-09-28 experiments/tisiso3performance.m

@@ -11,7 +11,7 @@ function [config, store, obs] = tisiso3performance(config, setting, data)
 % Date: 09-Jan-2017
 
 % Set behavior for debug mode
-if nargin==0, timbralSimilaritySol('do', 3, 'mask', {[4]    [5]    [3]    [2]    [2]    [5]    [1]    [1]    [2]    [2] [2]    [2]    [1]    [2]    [1]}); return; else store=[]; obs=[]; end
+if nargin==0, timbralSimilaritySol('do', 3, 'mask', {  [4]    [5]    [1]    [2]    [2]    [5]    [1]    [1]    [2]    [2] [2]    [2]    [1]    [2]    [1]}); return; else store=[]; obs=[]; end
 
 rng(0);
 

### 2019-09-28 experiments/tisisoReport.m

@@ -22,6 +22,9 @@ switch config.reportName
             captions{k}
             mk=mask;
             mk{factors(k)}= modalities(k)
+            if k==4
+                mk{14}=1;
+            end
             config = expExpose(config, 'l', 'save', ['tt' num2str(k)], 'step', 3, 'mask', mk, 'percent', 0, 'obs', 'p', 'highlight', 0, 'precision', 3);
         end
          % monomials

### 2019-09-30 paper/paper.tex
fix exp
start writing a related work section

@@
 %% TODO: mention companion website and open-source software
+
+\section*{Related work}
+Timbre involves multiple time scales in conjunction, from a few microseconds for an attack transient to several seconds for a sustained tone.
+Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more.
+Mapping this input to a feature space in which distances denote timbral dissimilarities requires a data-driven stage of dimensionality reduction.
+The scientific literature exhibits a methodological divide as regards the collection of human-annotated data.
+
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem.
+In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
+Because such tags often belong to the publicly available metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
+Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
+Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign  allows an explicit verification of instrument tags.
+For instance, the Open-MIC dataset, maintained by the Community for Open and Sustainable Music Information Research (COSMIR), comprises a vast corpus of X polyphonic music recordings spanning X instruments, as a derivative of the Free Music Archive (FMA) dataset.
+Another example is the Medley-solos-DB dataset, which comprises X monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music.
+Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition.
+We refer to X for a recent review of the state of the art in this domain.
+
+On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations.
+In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
+Rather, each excerpt must be played back to multiple human listeners.
+Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, thus restricting the size of the musical corpus under study.
+Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in artificial neural networks.
+Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF).
 
@@
 Concerning the feature computation block $B_1$, an important factor is the time support.
-When considering the commonly used duration of 25ms, the performance drop down to \ml{check value} $\%$.
+When considering the commonly used duration of 25ms, the performance drop down to $92\% \pm$.
 This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture long term modulations which plays an important part when considering extended playing techniques.

### 2019-09-30 paper/paper.tex
some comments to be discussed

@@
 
-The prospect of modeling timbre perception necessarily exceeds the symbolic domain.
+The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
 Instead, it involves a cognitive process which arises from the subjective experience of listening.
@@
 
-Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation.
+Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \ml{vl : cite esling and others ?}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
@@
 In this paper, we propose a machine listening model which computes the dissimilarity in timbre between any two audio samples.
-The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
+
+\ml{decrire les contributions signal ici}
+
+The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \ml{vl cite dlfm ?}.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \ml{vl cite}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
@@
 
-On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations.
+On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \ml{vl cite}.
 In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
@@
 Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF).
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF) \ml{vl citer, dire que c'est du coup une base solide de représentation et que le modèle mathamtique que l'on utilisera ici est quasi equivalent}.
+
+\ml{parler des travaux de Mounia}
 
@@
 Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
-In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple clustering of the sounds.
+In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
 While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
@@
 
-To evaluate the ability of the scattering transform to characterize the timbre of an audio signal, we consider the task of classifying musical notes from a wide variety of instruments played according to a set of extended instrumental playing techniques (IPTs).
-Increasing the variability in IPTs allows us to explore a wider range of timbre, providing a more robust validation of the proposed model for modeling timbre.
-The metric used is the \emph{precision at rank $k$} ($\patk$), which is calculated by computing the $k$ nearest neighbors in the feature space and recording the proportion of the neighbors that are in the same class.
-In the following, we set $k = 5$.
-The classes were obtained from the consensus clustering of human similarity judgments described in the previous section.
-The $\patk$ therefore provides a measure of consistency between the timbre similarity measures obtained by our model and those of the survey subjects.
-
-Concatenating the first- and second-order separable time and frequency scattering coefficients, we obtain a feature vector of dimension $???$ for a sound recording of duration $???$.
-Computing nearest neighbors with the standard Euclidean measure gives a $\patk$ of $0.870 \pm 0.058$, where the first value indicates the mean and the second the standard deviation.
-As a result, this simple transformation of the original signal results in a representation that lends itself naturally to discriminating timbre.
-Note that this result is obtained by simply looking at the raw feature vectors---no learning is performed to adapt the transform towards discriminating timbre.
-Replacing the separable second-order coefficients with the joint second-order coefficients boosts the $\patk$ to $???$.
-This is evidence that accurately characterizing the joint time--frequency structure of the sound is important capturing its timbre.
-
-We performed the same evaluation with a standard MFCC representation, obtaining a $\patk$ of $0.851 \pm 0.062$.
-As a result, we conclude that a Euclidean distance over scattering transforms (separable or joint) yield little additional perceptual information compared to standard spectral descriptors.
-
 \subsection*{Similarity weighting}
@@
 
-To improve our measure, we need to replace the standard Euclidean distance on the raw scattering coefficients by something better adapted to discriminating timbre.
-Equivalently, we can apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
-Very flexible and powerful methods can be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase in complexity.
-This increased complexity of the model reduces its usefulness with respect to interpretability.
+The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity.
+Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
 
-For this reason, we restrict ourselves to a simple linear transform of the scattering coefficients.
+Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
+More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
+
 Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
@@
 This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-Running LMNN on MFCCs with the consensus clustering obtained above increases $\patk$ slightly to $0.862 \pm 0.059$.
-Consequently, there seems to be little additional timbre information in the MFCCs compared to that given by the standard Euclidean distance.
-We contrast this with LMNN applied to the separable scattering coefficients, which yields a $\patk$ of $0.948 \pm 0.033$, which is a significant increase.
-For the joint scattering coefficients, we obtain a $\patk$ of $???$.
-We therefore conclude that a simple reweighting of the scattering transform yields a more appropriate timbral similarity measure.
 
@@
 
-Considering the separable scattering gives a performance of $93\% \pm 4.2$.
+Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ ja write details on what is lost here ?}
 
@@
 
+\ml{ parler des contributions signal ici}
+
 The success of the above system in identifying timbral similarities has immediate applications in browsing music databases.
-These are typically organized based on instrumental and PT taxonomies, with additional keywords offering a more flexible organization.
+These are typically organized based on instrumental and playing techniques taxonomies, with additional keywords offering a more flexible organization.
 Accessing the sounds in these databases therefore requires some knowledge of the taxonomy and keywords used.
@@
 We have shown that it is able to approximate well the timbral judgments of a wide range of subjects included in our study.
-Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on a reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
+Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on the reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
 This model can then be applied to new data, or, alternatively, be retrained with these new examples if the existing model proves unsatisfactory.
@@
 First, it is only applied to instrumental sounds.
-While this has the advantage of simplifying the intepretation of the results, the range of timbre under consideration is necessarily limited (although less restricted than only considering \emph{ordinario} PTs).
+While this has the advantage of simplifying the interpretation of the results, the range of timbre under consideration is necessarily limited (although less restricted than only considering \emph{ordinario} PTs).
 This also makes applications such as \emph{query-by-humming} difficult, since we cannot guarantee that the timbral similarity measure is accurate for such sounds.
+
+\ml{vl je reecrira le par du dessus en positif fture work}
 

### 2019-10-01 experiments/processResults.m
new figs

@@ -1,22 +1,41 @@
 clear all
 captions = {'full', 'split', 'random', 'lda', 'nolearning', '25ms', 'separable', 'mfcc', 'monomials'}
+ind = 1;
+
+iaInd = 2;
+d = load(['report/tables/tt' num2str(iaInd-1)]);
+[~, ia] = sort(d.data.rawData{1}.p);
+
+
 for k=1:length(captions)
-   d(k) = load(['report/tables/tt' num2str(k-1)]);
-   p(k, :) = d(k).data.rawData{1}.p;
-   x(k, :) = k*ones(1, size(p, 2));
-   c(k, :) = 1:size(p, 2);
+    if k==3, continue;end
+    
+    d(ind) = load(['report/tables/tt' num2str(k-1)]);
+    p(ind, :) = d(ind).data.rawData{1}.p;
+    p(ind, :) = p(ind, ia);   
+    
+    x(ind, :) = ind*ones(1, size(p, 2));
+    c(ind, :) = 1:size(p, 2);
+    ind = ind+1;
 end
+
+captions(3)=[];
+
 
 mean(p')
 fig = figure(1);
-boxplot(p')
-xticklabels(captions);
+boxplot(p', 'orientation', 'horizontal')
+set(gca, 'yticklabels', captions);
+xlabel('aP@5')
 saveas(fig, 'boxRes.png')
 
 fig = figure(2);
 
-s=scatter(x(:), p(:), 'filled')
+s=scatter(p(:), x(:), 'filled');
 s.CData = c(:);
-s.SizeData = 50;
+s.SizeData = 70;
+set(gca, 'yticklabels', captions);
+axis([55 103 0.5 8.5])
+xlabel('aP@5')
 saveas(fig, 'samplesRes.png')
 

### 2019-10-01 paper/paper.tex
abstract v1

@@
 \begin{abstract}
-Abstract here.
+
+The notion of musical timbre have been extensively considered through the lens of the classic instrumentarium.
+For such a purpose, datasets of instruments played \textit{ordinario} are considered.
+On thoses datasets, computational model of timbre considering static spectral features such as Mel Frequency Cepstral coefficents displayed excellent performance.
+
+The availability of richer musical datasets with many playing techniques, allows us to consider computational modeling of musical timbre with the outcomes of recent study  that stresses the importance of long term modulations in auditory perception.
+
+We propose a timbre model that is based on a parametrized...
+
+This model is evaluated in a retrieval task using a perceptually defined retrieval objective.
+Computational experiments demonstrates 1) the importance of handling long term modulations both in time and frequency, with a gain of $14\%$ in precision compared to Mfccs  2) the relevance of a learned layer that is able provide an average gain of $3\%$ in precision with a halving of the variance.
+
 \end{abstract}

### 2019-10-01 experiments/processResults.m

@@ -23,6 +23,7 @@ captions(3)=[];
 
 
 mean(p')
+var(p')
 fig = figure(1);
 boxplot(p', 'orientation', 'horizontal')
 set(gca, 'yticklabels', captions);
@@ -35,7 +36,7 @@ s=scatter(p(:), x(:), 'filled');
 s.CData = c(:);
 s.SizeData = 70;
 set(gca, 'yticklabels', captions);
-axis([55 103 0.5 8.5])
+axis([50 100 0.5 8.5])
 xlabel('aP@5')
 saveas(fig, 'samplesRes.png')
 

### 2019-10-01 paper/paper.tex
abstract v1

@@
 This model is evaluated in a retrieval task using a perceptually defined retrieval objective.
-Computational experiments demonstrates 1) the importance of handling long term modulations both in time and frequency, with a gain of $14\%$ in precision compared to Mfccs  2) the relevance of a learned layer that is able provide an average gain of $3\%$ in precision with a halving of the variance.
+Computational experiments demonstrates 1) the importance of handling long term modulations both in time and frequency, with a gain of $14\%$ in precision compared to Mfccs  2) the relevance of a learned layer that is able provide an average gain of $3\%$ in precision with a reduction of the variance of the prediction error of a factor of 2.
 

### 2019-10-02 paper/paper.tex
Add discussion of model in introduction

@@
 In this paper, we propose a machine listening model which computes the dissimilarity in timbre between any two audio samples.
-
-\ml{decrire les contributions signal ici}
+It is composed of two stages, one fixed and one learned, the former providing a baseline model for audio perception, with the latter specifically adapting the representation for estimation of timbre similarity.
+The fixed stage is constructed using the joint time--frequency scattering transform introduced by And\'en et al. \cite{anden2015joint,anden2019joint}.
+This representation extracts the time--frequency structure of the audio signal in a manner invariant to small time-shifts and stable to time-warping deformations.
+As such, it provides a general description of the audio signal that is not necessarily tailored to measuring timbre similarity.
+The second stage consists of a learned linear transform that, applied to the scattering representation, induces a metric that more closely aligns with timbre similarity judgments extracted from human listeners.
+We see that, numerically, this learned metric almost exactly reproduces the human similarity judgments.
+Moreover, we perform an ablation study of the model showing that degrading any element of it (replacing the scattering transform by simpler spectral descriptors, for example) significantly reduces the accuracy of the fit.
 

### 2019-10-02 paper/paper.tex
Add discussion on signal contributions

@@
 
-\ml{ parler des contributions signal ici}
+We see from the ablation study conducted above that each step of the proposed model is necessary for its high performance.
+Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
+This explains the relative success over such models \cite{lostanlen2018relevance,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity \cite{bauge2013representing}.
+
+On the other hand, we note that the complexity of the learned model does not have to be high.
+Indeed, for this task, a linear model on the scattering coefficients is sufficient.
+There is no need for deep networks with large numbers of parameters to accurately represent the similarity information.
+In other words, the scattering transform parametrizes the signal structure in a way that many relevant quantities (such as timbre similarity) can be extracted through a simple linear mapping.
+This is in line with other classification results, where linear support vector machines applied to scattering coefficients have achieved significant success \cite{anden2015joint,anden2019joint}.
+
+We also see the necessity of a fully \emph{joint} time--frequency model for accessing timbre, as opposed to a purely spectral model or one that treats the two axes in a separable manner.
+This fact has also been observed in other contexts, such as the work of Patil et al. \cite{patil2012music}.
+A related observation is the need for capturing large-scale structure.
+Indeed, reducing the window size to $25~\mathrm{ms}$ means that we lose a great deal of time--frequency structure, bringing results closer to that of the separable model.
 

### 2019-10-02 paper/paper.tex
update affiliation 1

--- 2019-10-01 paper/paper.tex
@@
 \address[id=aff1]{%                           % unique id
-  \orgname{Department of Zoology, Cambridge}, % university, etc
-  \street{Waterloo Road},                     %
+  \orgname{Music and Audio Research Lab, New York University}, % university, etc
+  \street{370 Jay Street},                     %
   %\postcode{}                                % post or zip code
-  \city{London},                              % city
-  \cny{UK}                                    % country
+  \city{New York, NY},                              % city
+  \cny{USA}                                    % country
 }
@@
 \begin{abstract}
-
-The notion of musical timbre have been extensively considered through the lens of the classic instrumentarium.
-For such a purpose, datasets of instruments played \textit{ordinario} are considered.
-On thoses datasets, computational model of timbre considering static spectral features such as Mel Frequency Cepstral coefficents displayed excellent performance.
-
-The availability of richer musical datasets with many playing techniques, allows us to consider computational modeling of musical timbre with the outcomes of recent study  that stresses the importance of long term modulations in auditory perception.
-
-We propose a timbre model that is based on a parametrized...
-
-This model is evaluated in a retrieval task using a perceptually defined retrieval objective.
-Computational experiments demonstrates 1) the importance of handling long term modulations both in time and frequency, with a gain of $14\%$ in precision compared to Mfccs  2) the relevance of a learned layer that is able provide an average gain of $3\%$ in precision with a reduction of the variance of the prediction error of a factor of 2.
-
+Abstract here.
 \end{abstract}

### 2019-10-02 paper/paper.tex
Edit abstract

--- 2019-10-02 paper/paper.tex
@@
 \begin{abstract}
-
-The notion of musical timbre have been extensively considered through the lens of the classic instrumentarium.
-For such a purpose, datasets of instruments played \textit{ordinario} are considered.
-On thoses datasets, computational model of timbre considering static spectral features such as Mel Frequency Cepstral coefficents displayed excellent performance.
-
-The availability of richer musical datasets with many playing techniques, allows us to consider computational modeling of musical timbre with the outcomes of recent study  that stresses the importance of long term modulations in auditory perception.
-
-We propose a timbre model that is based on a parametrized...
-
-This model is evaluated in a retrieval task using a perceptually defined retrieval objective.
-Computational experiments demonstrates 1) the importance of handling long term modulations both in time and frequency, with a gain of $14\%$ in precision compared to Mfccs  2) the relevance of a learned layer that is able provide an average gain of $3\%$ in precision with a reduction of the variance of the prediction error of a factor of 2.
-
+The timbre of a sound is an important factor in determining its identity and source.
+One successful means of exploring the timbre space has been the construction of computational models to analyze datasets of musical instrument recordings.
+However, these datasets have typically only included a limited set of instruments played with standard playing techniques, limiting the timbral range.
+For these datasets, models based on static spectral descriptors, such as mel-frequency cepstral coefficients (MFCCs), have enjoyed excellent performance.
+However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, we need more sophisticated models which incorporate dynamic large-scale signal structure.
+To this end, we propose a simple but rich model based on joint time--frequency scattering transforms that are linearly transformed to yield a timbral descriptor.
+The linear mapping is learned from human similarity judgments obtained in a crowdsourced free sorting task.
+We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
+In addition, we perform an ablation study demonstrating the necessity of various components in this model.
 \end{abstract}

### 2019-10-05 paper/paper.tex
cite maresz 2013 cmr

@@
 
-Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \ml{vl : cite esling and others ?}.
+Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.

### 2019-10-05 paper/paper.tex
redefine \eg and \ie
simplify paragraph 2

@@
 
-\newcommand*{\eg}{e.g.\@\xspace}
-\newcommand*{\ie}{i.e.\@\xspace}
+\newcommand*{\eg}{e.g.,\@\xspace}
+\newcommand*{\ie}{i.e.,\@\xspace}
 \newcommand{\ipt}{IPT\xspace}
@@
 From an audio signal processing standpoint, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
-Rather than to merely discriminate instruments as mutually exclusive categories, this distance function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.
+Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.
 

### 2019-10-05 paper/paper.tex
fix typo in introduction

@@
 
-Behind the overarching challenge of coming up with a robust predictive challenge for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
+Behind the overarching challenge of coming up with a robust predictive model for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.

### 2019-10-05 paper/paper.tex
proofread motivation section
simplify contribution section

@@
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
-The outcome of this is an instrumental audio sample, alongside textual attributes of playing technique.
+The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
 In this context, the computer serves as a bridge from the auditory level to the symbolic level, allowing to approximate a potentially infinite realm of timbral sensations by an orchestra of priorly specified extent and technical range.
@@
 
-In this paper, we propose a machine listening model which computes the dissimilarity in timbre between any two audio samples.
-It is composed of two stages, one fixed and one learned, the former providing a baseline model for audio perception, with the latter specifically adapting the representation for estimation of timbre similarity.
-The fixed stage is constructed using the joint time--frequency scattering transform introduced by And\'en et al. \cite{anden2015joint,anden2019joint}.
-This representation extracts the time--frequency structure of the audio signal in a manner invariant to small time-shifts and stable to time-warping deformations.
-As such, it provides a general description of the audio signal that is not necessarily tailored to measuring timbre similarity.
-The second stage consists of a learned linear transform that, applied to the scattering representation, induces a metric that more closely aligns with timbre similarity judgments extracted from human listeners.
-We see that, numerically, this learned metric almost exactly reproduces the human similarity judgments.
-Moreover, we perform an ablation study of the model showing that degrading any element of it (replacing the scattering transform by simpler spectral descriptors, for example) significantly reduces the accuracy of the fit.
+In this paper, we propose a machine listening system which computes the dissimilarity in timbre between any two audio samples.
+This system composes two stages: unsupervised feature extraction and supervised metric learning.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2019joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
+It aims at encoding patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
+The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN), which recombines scattering coefficients so as to fit a training set of human judgments.
 

### 2019-10-05 paper/paper.tex
cite DLFM paper

@@
 
-The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \ml{vl cite dlfm ?}.
+The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018dlfm}.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
@@
 Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
+A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the average precision at rank 5 of our system.
 

### 2019-10-05 paper/paper.tex
bis

@@
 
-The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018dlfm}.
+The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.

### 2019-10-05 paper/paper.tex
cite lostanlen 2016 ismir
cite martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp

@@
 Behind the overarching challenge of coming up with a robust predictive model for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
-Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
+Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal \cite{antoine2018isma}.
 The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \ml{vl cite}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp,lostanlen2016ismir}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.

### 2019-10-05 paper/paper.tex
cite humphrey 2016, mcfee 2016

@@
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign  allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset, maintained by the Community for Open and Sustainable Music Information Research (COSMIR), comprises a vast corpus of X polyphonic music recordings spanning X instruments, as a derivative of the Free Music Archive (FMA) dataset.
+For instance, the Open-MIC dataset \cite{humphrey2016ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of X polyphonic music recordings spanning X instruments, as a derivative of the Free Music Archive (FMA) dataset.
 Another example is the Medley-solos-DB dataset, which comprises X monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music.

### 2019-10-05 paper/paper.tex
fill in related work section on MIR

--- 2019-10-05 paper/paper.tex
@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp,lostanlen2016ismir}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
@@
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign  allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset, maintained by the Community for Open and Sustainable Music Information Research (COSMIR), comprises a vast corpus of X polyphonic music recordings spanning X instruments, as a derivative of the Free Music Archive (FMA) dataset.
-Another example is the Medley-solos-DB dataset, which comprises X monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music.
+For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning across 20 instruments, as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
+Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
 Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition.
-We refer to X for a recent review of the state of the art in this domain.
+We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
 

### 2019-10-05 paper/paper.tex
cite bhalke

@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.

### 2019-10-05 paper/paper.tex
cite depireux

@@
 Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF) \ml{vl citer, dire que c'est du coup une base solide de représentation et que le modèle mathamtique que l'on utilisera ici est quasi equivalent}.
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF) \cite{depireux2001amphyssoc}.
 
-\ml{parler des travaux de Mounia}
+% \ml{vl citer, dire que c'est du coup une base solide de représentation et que le modèle mathamtique que l'on utilisera ici est quasi equivalent}.
+
+%\ml{parler des travaux de Mounia}
 

### 2019-10-05 paper/paper.tex
cite faure

@@
 Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
-On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them.
+On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{faure1996icmpc}.
 On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
@@
 This system composes two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2019joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
 It aims at encoding patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
@@
 Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRF) \cite{depireux2001amphyssoc}.
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
+
+Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
+Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
 

### 2019-10-05 paper/paper.tex
cite mcfee 2015 ismir

@@
 Mapping this input to a feature space in which distances denote timbral dissimilarities requires a data-driven stage of dimensionality reduction.
-The scientific literature exhibits a methodological divide as regards the collection of human-annotated data.
+The scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}.
 
@@
 Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
-Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition.
+Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir}.
 We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.

### 2019-10-05 paper/paper.tex
cite bellet, gururani, pons

@@
 The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
-It aims at encoding patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
-The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN), which recombines scattering coefficients so as to fit a training set of human judgments.
+It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
+The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{}.
+It recombines scattering coefficients so as to fit a training set of human judgments.
 
@@
 Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
-Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir}.
+Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
 We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
@@
 Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
+
+We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
 

### 2019-10-05 paper/paper.tex
cite the recent siedenburg and mcadams chapter
cite downie and mcadams, improve intro

@@
 
-Music information retrieval operates at either of two levels: symbolic and auditory.
+Music information retrieval operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
 By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, or intensity at the source.
-Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis.
-As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} dry, rough, bright, round, and so forth) or by specifying an instrumental playing technique (\eg{} bowed or plucked).
+Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis \cite{siedenburg2019chapter}.
+As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
 
 Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
-On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{faure1996icmpc}.
-On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
+On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
+On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
@@
 Behind the overarching challenge of coming up with a robust predictive model for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
-Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal \cite{antoine2018isma}.
+Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
 The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
-In this context, the computer serves as a bridge from the auditory level to the symbolic level, allowing to approximate a potentially infinite realm of timbral sensations by an orchestra of priorly specified extent and technical range.
+In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
 To maximize the perceptual fidelity of such an approximation, a query-by-example search engine in the auditory domain should summarize the vastest possible vocabulary of timbral templates.
 
-In this paper, we propose a machine listening system which computes the dissimilarity in timbre between any two audio samples.
+In this article, we propose a machine listening system which computes the dissimilarity in timbre between any two audio samples.
 This system composes two stages: unsupervised feature extraction and supervised metric learning.
@@
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
-The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{}.
-It recombines scattering coefficients so as to fit a training set of human judgments.
+The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
+It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}.
 
@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.

### 2019-10-05 paper/paper.tex
cite more mcadams

@@
 
-On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \ml{vl cite}.
+On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
 In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.

### 2019-10-07 code/xp2/evaluation/bss_eval_sources.m
done renorm figs

@@ -1,201 +1,0 @@
-function [SDR,SIR,SAR,perm]=bss_eval_sources(se,s)
-
-% BSS_EVAL_SOURCES Ordering and measurement of the separation quality for
-% estimated source signals in terms of filtered true source, interference
-% and artifacts.
-% 
-% The decomposition allows a time-inmode filter distortion of length
-% 512, as described in Section III.B of the reference below.
-%
-% [SDR,SIR,SAR,perm]=bss_eval_sources(se,s)
-%
-% Inputs:
-% se: nsrc x nsampl matrix containing estimated sources
-% s: nsrc x nsampl matrix containing true sources
-%
-% Outputs:
-% SDR: nsrc x 1 vector of Signal to Distortion Ratios
-% SIR: nsrc x 1 vector of Source to Interference Ratios
-% SAR: nsrc x 1 vector of Sources to Artifacts Ratios
-% perm: nsrc x 1 vector containing the best ordering of estimated sources
-% in the mean SIR sense (estimated source number perm(j) corresponds to
-% true source number j)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Copyright 2008 Emmanuel Vincent
-% This software is distributed under the terms of the GNU Public License
-% version 3 (http://www.gnu.org/licenses/gpl.txt)
-% If you find it useful, please cite the following reference:
-% Emmanuel Vincent, Rémi Gribonval, and Cédric Févotte, "Performance
-% measurement in blind audio source separation," IEEE Trans. on Audio,
-% Speech and Language Processing, 14(4):1462-1469, 2006.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-
-%%% Errors %%%
-if nargin<2, error('Not enough input arguments.'); end
-[nsrc,nsampl]=size(se);
-[nsrc2,nsampl2]=size(s);
-if nsrc2~=nsrc, error('The number of estimated sources and reference sources must be equal.'); end
-if nsampl2~=nsampl, error('The estimated sources and reference sources must have the same duration.'); end
-
-%%% Performance criteria %%%
[…]

### 2019-10-07 paper/paper.tex
write on STRFs + 5 more citations

@@
 In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
-To maximize the perceptual fidelity of such an approximation, a query-by-example search engine in the auditory domain should summarize the vastest possible vocabulary of timbral templates.
+%To maximize the perceptual fidelity of such an approximation, a query-by-example search engine in the auditory domain should summarize a vast  of timbres.
 
-In this article, we propose a machine listening system which computes the dissimilarity in timbre between any two audio samples.
-This system composes two stages: unsupervised feature extraction and supervised metric learning.
+This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
+Our system composes two stages: unsupervised feature extraction and supervised metric learning.
 The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
@@
 This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
+The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
+The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
+Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
+More recently, the STRF was redefined directly as a spectrographic representation of the spike-triggered average, in order to improve readability \cite{klein2000compneur}.
 

### 2019-10-08 paper/paper.tex
incorporate comments from meetings

@@
 Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
-More recently, the STRF was redefined directly as a spectrographic representation of the spike-triggered average, in order to improve readability \cite{klein2000compneur}.
+More recently, in order to improve readability \cite{klein2000compneur}, the STRF was redefined directly as a spectrographic representation of the spike-triggered average.
 
@@
 
-\ml{introduce the system globally, while stressing the ability of the system to be conditionned for one or many people}
+\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}
 
@@
 In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
-Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation}, it is  thus taken in this paper as a reasonable assumption.
+Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation.
+reformuler tout le paragraphe}, it is  thus taken in this paper as a reasonable assumption.
 Quantification of this invariance is left for further study.
@@
 
-\ml{provide parameters for scattering
+\ml{VL. provide parameters for scattering
 \texttt{
@@
 
-In order to better understand the benefits of each design choice, an ablation study \ml{vl ref ??} is now performed using a slightly different protocol.
+In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol.
 In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those.
@@
 The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
-One of those is the linear discriminant analysis \ml{ja ref ?}\cite{}.
+One of those is the linear discriminant analysis \ml{vl ref ?}\cite{}.
 When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.
@@
 
-Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ ja write details on what is lost here ?}
+Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
 
@@
 
-\ml{vl je reecrira le par du dessus en positif fture work}
+\ml{je reecrirai le par du dessus en positif future work (TODO VL).}
 
@@
 For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
-\ja{Better to display these in a table?}
-
-Some instruments can be complemented with sordina (S), wha (W).
-The addition of this kind of device modifies the shape of the instrument.
-Consequently, an alto saxophone and alto saxophone with sordina are considered as different instruments.
-\vl{This is crucial and should not be deferred to the appendix. @writing}
-\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}
+\ml{Better to display these in a table? VL todo table}
 

### 2019-10-09 paper/paper.tex
details generation of sim matrices, and organizing figures

@@
 
-% \ml{vl citer, dire que c'est du coup une base solide de représentation et que le modèle mathamtique que l'on utilisera ici est quasi equivalent}.
+\begin{figure}[h!]
+\caption{\csentence{STRF + scattering (upsampling STRF)}
+A short description of the figure content should go here.}
+\end{figure}
 
-%\ml{parler des travaux de Mounia}
+\ml{vl on fait toujours cette figure ?}
 
@@
 While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
-More information on the data collection process is available in  Section \ref{sec:judgments}.
+More information on the data collection process is available in  Section \ref{sec:subjective}.
 
@@
 We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
-The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fix:xp2display}.
+The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.
 
@@
 We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
-Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other, see Figure \ref{fig:consensusVsIpt}.
+
+In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
+For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
+The PTs similarity matrix is computed in the same way.
+
+The ordering of the instruments or the PTs on the vertical and horizontals axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
+This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.
+
+Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
 
@@
 \center
-\includegraphics[width = \textwidth]{../code/xp2/consensusVsI.png}
-\caption{'ASax'    'BBTb'    'BbCl'    'Bn'    'Cb'    'Fl'    'Gtr'    'Hn'    'Hp'    'Ob'    'TTbn'    'TpC'  'Va'    'Vc'}
-\label{fig:consensusVsIpt}
+\includegraphics[width = \textwidth]{figures/consensusVsI.png}
+\caption{}
+\label{fig:consensusVsI}
 \end{figure}
@@
 \center
-\includegraphics[width = \textwidth]{../code/xp2/consensusVsPt.png}
-\caption{ 'art-harm'    'blow-no-reed'    'explo-slap'    'harm-fngr'    'key-cl'    'legno-batt'    'legno-tratto'
-    'nonvib'    'ord'    'ord-closed'    'ord-hi-reg'    'ord-open'    'pdl-tone'    'pizz'    'pizz-bartok'
-    'pizz-lv'    'pizz-sec'    'pont'    'slap'    'slap-unp'    'tasto'    'tng-ram'}
-\label{fig:consensusVsIpt}
+\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
+\caption{ }
+\label{fig:consensusVsPt}
 \end{figure}
-
-\ja{We should have some figure to back this up.
-Can this be done using the tag clouds?}
-\ml{tentative solution using similarity matrices, see fig, if ok should be improved by vl (code/xp2/consensusVsIpt.m)}
 
@@
 \label{sec:method}
+
+\begin{figure}[h!]
+\caption{\csentence{Model overview}
+One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
+scat, isomap.
+Second line : perceptual clustering.}
+\label{fig:pipeline}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.}
+\end{figure}
 
@@
 For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$.
-The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{}.
+The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{sec:weighting}.
 For this algorithm the neighborhood is set to $5$.
@@
 
-\section*{Figures}
-
-\begin{figure}[h!]
-\caption{\csentence{STRF + scattering (upsampling STRF)}
-A short description of the figure content should go here.}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Model overview}
-One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
-scat, isomap.
-Second line : perceptual clustering.}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Tag clouds}
-A short description of the figure content should go here.}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Perfs MFCC, separable, joint (last)}
-A short description of the figure content should go here.}
-\end{figure}
+%\section*{Figures}
 
@@
 
-\section*{Tables}
-\begin{table}[h!]
-\caption{Sample table title. This is where the description of the table should go.}
-      \begin{tabular}{cccc}
-        \hline
-           & B1  &B2   & B3\\ \hline
-        A1 & 0.1 & 0.2 & 0.3\\
-        A2 & ... & ..  & .\\
-        A3 & ..  & .   & .\\ \hline
-      \end{tabular}
-\end{table}
+% \section*{Tables}
+% \begin{table}[h!]
+% \caption{Sample table title. This is where the description of the table should go.}
+%       \begin{tabular}{cccc}
+%         \hline
+%            & B1  &B2   & B3\\ \hline
+%         A1 & 0.1 & 0.2 & 0.3\\
+%         A2 & ... & ..  & .\\
+%         A3 & ..  & .   & .\\ \hline
+%       \end{tabular}
+% \end{table}
 

### 2019-10-09 paper/paper.tex
cite livshin

--- 2019-10-08 paper/paper.tex
@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
@@
 The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
-Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
-More recently, in order to improve readability \cite{klein2000compneur}, the STRF was redefined directly as a spectrographic representation of the spike-triggered average.
+Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but suffering from cross-term interference \cite{eggermont1993hearing}.
+Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
+Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than the original one, which it supersedes.
+In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
+This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
+Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model''.
+
+%
 

### 2019-10-09 paper/paper.tex

--- 2019-10-09 paper/paper.tex
@@
 The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
-Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
-More recently, in order to improve readability \cite{klein2000compneur}, the STRF was redefined directly as a spectrographic representation of the spike-triggered average.
+Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but suffering from cross-term interference \cite{eggermont1993hearing}.
+Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
+Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than the original one, which it supersedes.
+In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
+This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
+Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model''.
 

### 2019-10-09 paper/paper.tex
write about strfs for classification

@@
 Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
-Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than the original one, which it supersedes.
+Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner-Ville distribution.
 In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
@@
 Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model''.
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}. 
+Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
+These biologically inspired features have shown to outperform the state of the art, especially in the small data regime.
 
-Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
-Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
-
-We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
+% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
 
@@
 
-\ml{vl on fait toujours cette figure ?}
+Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
+Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
+
+We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
 
@@
 We also see the necessity of a fully \emph{joint} time--frequency model for accessing timbre, as opposed to a purely spectral model or one that treats the two axes in a separable manner.
-This fact has also been observed in other contexts, such as the work of Patil et al. \cite{patil2012music}.
+This fact has also been observed in other contexts, such as the work of Patil et al. \cite{patil2012ploscompbiol}.
 A related observation is the need for capturing large-scale structure.

### 2019-10-09 paper/paper.tex
write on the confusion matrices of patil

@@
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
-Because such tags often belong to the publicly available metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
+Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
 Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
@@
 Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
-Over the past decade, the surge of these large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning techniques, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
+Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
+Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer.
 We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
@@
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features have shown to outperform the state of the art, especially in the small data regime.
+These biologically inspired features have shown to outperform the state of the art, especially in the small data regime when deep learning is inapplicable. % TODO cite?
+Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
+In other words, although there is no evidence that the full cortical model consistently predicts the same instrument labels as a human listener, cross-instrument confusions do occur in the same proportions overall.
 

### 2019-10-09 paper/paper.tex
write about the scale-rate domain

@@
 Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}. 
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the scale-rate domain $(\alpha, \beta)$.
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.

### 2019-10-10 paper/paper.tex
add hemery and thoret
cite thoret2016jasa, hemery2015frontiers

--- 2019-10-09 paper/paper.tex
@@
 The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
-Instead, it involves a cognitive process which arises from the subjective experience of listening.
+Instead, it involves a cognitive process which arises from the subjective experience of listening. % TODO VL refs. 
 From an audio signal processing standpoint, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
-Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.
+Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal \cite{thoret2018jasa}.
 
@@
 In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
-%To maximize the perceptual fidelity of such an approximation, a query-by-example search engine in the auditory domain should summarize a vast  of timbres.
 
@@
 Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
-A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the average precision at rank 5 of our system.
+A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
 
@@
 Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}. 
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate-scale domain $(\alpha, \beta)$.
+
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
@@
 Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
-In other words, although there is no evidence that the full cortical model consistently predicts the same instrument labels as a human listener, cross-instrument confusions do occur in the same proportions overall.
+Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
+This is in contrast with fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for saliency map estimation. %cite 
+Lastly, beyond the realm of supervised classification, a previous publication \cite{hemery2015frontiers} has shown that query-by-example search in the feature space of STRFs allows to recover the acoustic similarity between environmental soundscapes, even after temporal integration and unsupervised dimensionality reduction.
+
+%Lastly, beyond the realm of supervised classification, % todo cite thoret, cite hemery
+%For these reasons, STRFs offer an appealing starting point towards the development of a query-by-example search engine for instrumental playing techniques.
+%Yet, the full cortical model
 
@@
 \label{sec:scattering}
+
+% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
 

### 2019-10-10 paper/paper.tex
cite joder2009taslp

@@
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis,joder2009taslp}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.

### 2019-10-10 paper/paper.tex
cite andreux + dafx paper

@@
 
-Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
+Despite their widespread use, purely linguistic references to timbre fail to convey the intention of the composer.
 On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
@@
 Our system composes two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018relevance}.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
@@
 The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
-Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but suffering from cross-term interference \cite{eggermont1993hearing}.
+Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but eliciting cross-term interferences \cite{eggermont1993hearing}.
 Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
@@
 Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
-This is in contrast with fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for saliency map estimation. %cite 
-Lastly, beyond the realm of supervised classification, a previous publication \cite{hemery2015frontiers} has shown that query-by-example search in the feature space of STRFs allows to recover the acoustic similarity between environmental soundscapes, even after temporal integration and unsupervised dimensionality reduction.
+This is unlike fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for saliency map estimation. %VL TODO cite Choi Auralization
+Lastly, beyond the realm of supervised classification, a previous publication \cite{hemery2015frontiers} has shown that query-by-example search with STRFs allows to discriminate categories of environmental soundscapes, even after temporal integration and unsupervised dimensionality reduction.
 
-%Lastly, beyond the realm of supervised classification, % todo cite thoret, cite hemery
-%For these reasons, STRFs offer an appealing starting point towards the development of a query-by-example search engine for instrumental playing techniques.
-%Yet, the full cortical model
-
-% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
-
-\begin{figure}[h!]
+\begin{figure}
 \caption{\csentence{STRF + scattering (upsampling STRF)}
@@
 
-Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
-Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
+These reasons make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
+Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
+Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
+On the contrary, joint time--frequency scattering offers a faster extraction spectrotemporal modulation while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
 
-We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
+% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
+
+%Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
+%Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
+%We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
 
@@
 Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019joint}, and beating phenomena \cite{anden2014deep}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019joint,lostanlen2018relevance,lostanlen2018extended}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019joint,lostanlen2018jasmp,lostanlen2018extended}.
 
@@
 Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
-This explains the relative success over such models \cite{lostanlen2018relevance,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity \cite{bauge2013representing}.
+This explains the relative success over such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity \cite{bauge2013representing}.
 

### 2019-10-10 paper/paper.tex
cite mallat2016philtrans

@@
 
-These reasons make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
+The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
 Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
 Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
-On the contrary, joint time--frequency scattering offers a faster extraction spectrotemporal modulation while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
+On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulation while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
+Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables.
+This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
 

### 2019-10-10 paper/paper.tex
write more on contributions

@@
 Our system composes two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015joint}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
@@
 
-The main originality of our model is that it encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
+This article gathers three novel contributions.
+First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
-A second originality of our model is that we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
+Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
 Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
 A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
+Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
+These visualizations confirm the theoretical predictions of previous research \cite{anden2015mlsp}, which relied on simple synthetic sounds such as exponential chirps.
+They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology, while offering an accelerated algorithm for scalable feature extraction.
 
@@
 Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate-scale domain $(\alpha, \beta)$.
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$.
 
@@
 
-\begin{figure}
-\caption{\csentence{STRF + scattering (upsampling STRF)}
-A short description of the figure content should go here.}
-\end{figure}
+%\begin{figure}
+%\caption{\csentence{STRF + scattering (upsampling STRF)}
+%A short description of the figure content should go here.}
+%\end{figure}
 
@@
 Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
-On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulation while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
-Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables.
+On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
+Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
 This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
+Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
+Recently, a previous publication
 
@@
 %Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
-%Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019joint}.
+%Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019tsp}.
 %We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
@@
 To remedy this, another approach was suggested by And\'{e}n et al.
-\cite{anden2015joint,anden2019joint} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
+\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
 Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
@@
 The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019joint}, and beating phenomena \cite{anden2014deep}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019joint,lostanlen2018jasmp,lostanlen2018extended}.
+Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
 
@@
 In other words, the scattering transform parametrizes the signal structure in a way that many relevant quantities (such as timbre similarity) can be extracted through a simple linear mapping.
-This is in line with other classification results, where linear support vector machines applied to scattering coefficients have achieved significant success \cite{anden2015joint,anden2019joint}.
+This is in line with other classification results, where linear support vector machines applied to scattering coefficients have achieved significant success \cite{anden2015mlsp,anden2019tsp}.
 

### 2019-10-10 paper/paper.tex
cite wang2019ismir
andreux2018ismir

@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
-Recently, a previous publication
+Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
+Despite showing experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument (here, the Chinese bamboo flute) to an entire orchestra.
+In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation in terms of its ability to reproduce subjective dissimilarity judgments, remotely from any classification task.
 
+This paper strives to fill the gap in scholarship
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).

### 2019-10-10 paper/paper.tex
edit related work section

@@
 Instead, it involves a cognitive process which arises from the subjective experience of listening. % TODO VL refs. 
-From an audio signal processing standpoint, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
+The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
 Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal \cite{thoret2018jasa}.
@@
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features have shown to outperform the state of the art, especially in the small data regime when deep learning is inapplicable. % TODO cite?
+These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable. % TODO cite?
 Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
@@
 Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
-Despite showing experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument (here, the Chinese bamboo flute) to an entire orchestra.
+Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
 In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation in terms of its ability to reproduce subjective dissimilarity judgments, remotely from any classification task.
 
-This paper strives to fill the gap in scholarship
+In summary, our paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
+From the standpoint of MIR, 
+Conversely, from the standpoint of music cognition, this paper 
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).

### 2019-10-10 paper/paper.tex
conclude the related work section

@@
 
-%\begin{figure}
-%\caption{\csentence{STRF + scattering (upsampling STRF)}
-%A short description of the figure content should go here.}
-%\end{figure}
-
 The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
@@
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
-In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation in terms of its ability to reproduce subjective dissimilarity judgments, remotely from any classification task.
+In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
 
-In summary, our paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, 
-Conversely, from the standpoint of music cognition, this paper 
+In summary, this paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
+From the standpoint of MIR, our model offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being tailored to subjective preferences.
+
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
-
-%Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
-%Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019tsp}.
-%We refer to \cite{bellet2015book} for a review of the state of the art in this domain.
+%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
 

### 2019-10-11 paper/paper.tex
rename data collection section

@@
 From the standpoint of MIR, our model offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being tailored to subjective preferences.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
 
@@
 
-\section*{Timbre similarity judgments}
+\section*{Data collection of timbre disimilarity judgments}
 \label{sec:survey}
@@
 The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:dataset}.
+For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
@@
 \section*{Dataset}
-\label{sec:dataset}
+\label{sec:appendix-dataset}
 
@@
 The way the dataset is structured
-is explained in more details in Appendix \ref{sec:dataset}.
+is explained in more details in Appendix \ref{sec:appendix-dataset}.
 

### 2019-10-11 paper/paper.tex
improve up to sec 2

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 % \def\includegraphic[#1]{}
@@
 For these datasets, models based on static spectral descriptors, such as mel-frequency cepstral coefficients (MFCCs), have enjoyed excellent performance.
-However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, we need more sophisticated models which incorporate dynamic large-scale signal structure.
-To this end, we propose a simple but rich model based on joint time--frequency scattering transforms that are linearly transformed to yield a timbral descriptor.
-The linear mapping is learned from human similarity judgments obtained in a crowdsourced free sorting task.
+However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, more sophisticated models which incorporate dynamic large-scale signal structure are needed.
+To this end, we propose a simple but rich model based on joint time--frequency scattering transforms that are linearly transformed to yield perceptively motivated timbral descriptors as the linear mapping is learned from human similarity judgments obtained in a crowdsourced free sorting task.
 We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
@@
 On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
-On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
+On the other hand, the sole mention of a playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
@@
 
-Behind the overarching challenge of coming up with a robust predictive model for human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
+Behind the overarching challenge of coming up with a robust predictive model for such a human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
@@
 These visualizations confirm the theoretical predictions of previous research \cite{anden2015mlsp}, which relied on simple synthetic sounds such as exponential chirps.
-They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology, while offering an accelerated algorithm for scalable feature extraction.
+They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.
 
-%% TODO: list Sections
-
-%% TODO: mention companion website and open-source software
+The remaining of this paper is organized as follows.
+Section \ref{sec:soa} presents an in-depth review of the state of the art in the computational modeling of musical timbre.
+The perceptual data that allows us to train and evaluate the proposed model is briefly presented in Section \ref{sec:survey}.
+Extended presentation of the gathering protocol used for collecting this data is presented in Annex \ref{sec:subjective}.
+The computational model and its core components are described in Section \ref{sec:method}.
+Their benefits are evaluated in Section \ref{sec:results} and discussed in Section \ref{sec:discussion}\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}.
 
 \section*{Related work}
+\label{sec:soa}
+
 Timbre involves multiple time scales in conjunction, from a few microseconds for an attack transient to several seconds for a sustained tone.
@@
 
-In summary, this paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, our model offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
+Contributions presented in this paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
+From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
 Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
@@
 
-\section*{Data collection of timbre disimilarity judgments}
+\section*{Timbre disimilarity judgments}
 \label{sec:survey}
 
-To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
+To construct a dataset for exploring the perception of timbre for different instruments and playing techniques, we conducted a survey on the perceived similarity of a set of sounds.
 The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.

### 2019-10-11 paper/paper.tex
cite mishra and peinkofer, write on percussion

--- 2019-10-11 paper/paper.tex
@@
 Timbre involves multiple time scales in conjunction, from a few microseconds for an attack transient to several seconds for a sustained tone.
-Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more.
+Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more \cite{joder2009taslp}.
 Mapping this input to a feature space in which distances denote timbral dissimilarities requires a data-driven stage of dimensionality reduction.
-The scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}.
+In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}.
 
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis,joder2009taslp}.
+On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
 In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
@@
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable. % TODO cite?
-Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
+These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable.Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
 Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
-This is unlike fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for saliency map estimation. %VL TODO cite Choi Auralization
+This is unlike fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for feature inversion \cite{mishra2018ismir}.
 Lastly, beyond the realm of supervised classification, a previous publication \cite{hemery2015frontiers} has shown that query-by-example search with STRFs allows to discriminate categories of environmental soundscapes, even after temporal integration and unsupervised dimensionality reduction.
@@
 \label{sec:survey}
+
+The European symphonic orchestra encompasses four families of instruments: woodwinds, brass, strings, and percussion.
+In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
+The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
+Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a review) these are rarely shared with pitched instruments.
+In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed.
 

### 2019-10-11 paper/paper.tex
list instruments

@@
 
-The European symphonic orchestra encompasses four families of instruments: woodwinds, brass, strings, and percussion.
+The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
 In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
 The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
-Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a review) these are rarely shared with pitched instruments.
-In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed.
+Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
+In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
+We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive sounds, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
+
+We consider a list of 15 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
+% TODO check tuba abbreviation
+Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
 

### 2019-10-11 paper/paper.tex

@@
 
-\section*{Data collection of timbre disimilarity judgments}
+\section*{Data collection of timbre similarity judgments}
 \label{sec:survey}

### 2019-10-11 paper/paper.tex
rework outline

@@
 
-The remaining of this paper is organized as follows.
-Section \ref{sec:soa} presents an in-depth review of the state of the art in the computational modeling of musical timbre.
-The perceptual data that allows us to train and evaluate the proposed model is briefly presented in Section \ref{sec:survey}.
-Extended presentation of the gathering protocol used for collecting this data is presented in Annex \ref{sec:subjective}.
-The computational model and its core components are described in Section \ref{sec:method}.
-Their benefits are evaluated in Section \ref{sec:results} and discussed in Section \ref{sec:discussion}\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}.
+Section \ref{sec:soa} reviews recent research in the computational modeling of musical timbre.
+Section \ref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
+Section \ref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+Section \ref{sec:results} reports the results of an ablation study of our system.
+Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}.
 

### 2019-10-11 paper/paper.tex
grammar in list of contributions

@@
 
-Contributions presented in this paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
+The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
 From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
@@
 
-\section*{Data collection of timbre similarity judgments}
+\section*{Timbre similarity judgments}
 \label{sec:survey}

### 2019-10-12 paper/paper.tex
write on mutes

@@
 The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
-Instead, it involves a cognitive process which arises from the subjective experience of listening. % TODO VL refs.
+Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
 The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
@@
 
-Behind the overarching challenge of coming up with a robust predictive model for such a human behavior, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
+Behind the overarching challenge of coming up with a robust predictive model for human listening behaviors, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
@@
 Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
+Each of these instruments affords a panel of playing techniques ranging in size between X and Y. % TODO update with numbers
+Furthermore, independently of the playing technique, some instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
 

### 2019-10-13 paper/paper.tex
add list of all stimuli in appendix

@@
 
-We consider a list of 15 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-% TODO check tuba abbreviation
+We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
 Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
-Each of these instruments affords a panel of playing techniques ranging in size between X and Y. % TODO update with numbers
-Furthermore, independently of the playing technique, some instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
+Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
+Once augmented with mutes, the list of 16 instruments grows up to X. % TODO complete
+Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between X (for the X) and Y (for the Y). % TODO update with number.
+In total, the feasible triplets of instrument, mute, and technique amount to a list of $N=235$ combinations.
+
+% TODO explain the conundrum leading to reduce the dataset size.
 
@@
 
+\begin{table}
+\caption{Full list of audio stimuli (1/2).}
+      \begin{tabular}{llll}
+          name & instrument  & mute   &  playing technique \\ \hline \hline
+        \texttt{ASax-key-cl-C4-p} & Alto saxophone & & key click \\
+        \texttt{ASax-ord-C4-mf} & Alto saxophone &  &  \\
+        \texttt{ASax-ord-hi-reg-C6-mf} & Alto saxophone &   & \\ 
+        \texttt{ASax-slap-C4-mf} & Alto saxophone & & slap tonguing \\ 
+        \texttt{ASax-slap-unp-C4-p} & Alto saxophone & & unpitched slap tonguing \\ \hline
+        \texttt{BbCl-key-cl-C4-pp} & Soprano clarinet & & key click \\
+        \texttt{BbCl-ord-hi-reg-A6-ff} & Soprano clarinet & & \\ \hline
+        \texttt{BBTb-explo-slap-C\#1-mf} & Bass tuba &  & explosive slap \\ 
+		\texttt{BBTb-pdl-tone-F1-mf} & Bass tuba & & pedal tone \\
+		\texttt{BBTb-slap-F1-mf} & Bass tuba & & slap tonguing \\
+		\texttt{BBTb-slap-unp-mf-1} & Bass tuba & & unpitched slap tonguing \\ \hline
+		\texttt{Bn-key-cl-C3-mf} & Bassoon & & key click \\
+		\texttt{Bn-ord-C4-mf} & Bassoon & & \\ \hline
+		\texttt{Cb-pizz-bartok-C4-ff-1c} & Contrabass & & pizzicato Bart\'ok (snap) \\
+		\texttt{Cb-pizz-lv-C4-mf-1c} & Contrabass & & pizzicato \& \emph{laissez vibrer} \\
+        \texttt{Cb-pizz-sec-C4-mf-1c} & Contrabass & & pizzicato secco \\
+        \texttt{Cb-pont-C4-mf-1c} & Contrabass & & sul ponticello \\ \hline
+        \texttt{Fl-key-cl-C4-f} & Flute & & key click \\
+        \texttt{Fl-ord-C4-mf} & Flute & & \\
+        \texttt{Fl-tng-ram-C4-mf} & Flute & & tongue ram \\ \hline
+        \texttt{Gtr-ord-C4-mf-2c} & Spanish guitar & & \\
+        \texttt{Gtr-ord-hi-reg-E5-mf-3c} & Spanish guitar & & \\
+        \texttt{Gtr-pizz-bartok-C4-ff-2c} & Spanish guitar & & pizzicato Bart\'ok\\
+        \texttt{Gtr-pizz-C4-mf-2c} & Spanish guitar &  & pizzicato \\ \hline
+        \texttt{Hn-ord-C4-mf} & French horn & & \\
+        \texttt{Hn-slap-C4-mf} & French horn & & slap tonguing \\ \hline
+        \texttt{Hp-harm-fngr-C4-f} & Harp & & harmonic fingering \\
+        \texttt{Hp-ord-C4-m4} & Harp & & \\
+        \texttt{Hp-pizz-bartok-C4-mf} & Harp & & pizzicato Bart\'ok \\
+        \texttt{Hp-xyl-C4-p} & Harp & & xylophonic \\ \hline
+        \texttt{Ob-blow-no-reed-C4} & Oboe & & blow without reed \\
+        \texttt{Ob-key-cl-C4-pp} & Oboe & & key click \\
+        \texttt{Ob-ord-C4-mf} & Oboe & & \\ \hline
+        \texttt{TpC-ord-C4-mf} & Trumpet in C & & \\
+        \texttt{TpC-pdl-tone-F3-mf} & Trumpet in C & & pedal tone \\
+        \texttt{TpC-slap-C4-p} & Trumpet in C & & slap tonguing \\
+        \texttt{TpC+C-ord-C4-mf} & Trumpet in C & cup & \\
+        \texttt{TpC+H-ord-C4-mf} & Trumpet in C & harmon & \\
+        \texttt{TpC+S-ord-C4-mf} & Trumpet in C & straight & \\
+        \texttt{TpC+W-ord-closed-C4-mf} & Trumpet in C & wah (closed) & \\
+        \texttt{TpC+W-ord-open-C4-mf} & Trumpet in C & wah (open) & \\ \hline
+        \texttt{TTbn-ord-C4-mf} & Tenor trombone & &  \\ 
+        \texttt{TTbn+C-ord-C4-mf} & Tenor trombone & cup & \\ 
+        \texttt{TTbn+H-ord-C4-mf} & Tenor trombone & harmon & \\
+        \texttt{TTbn+S-ord-C4-mf} & Tenor trombone & straight & \\
+        \texttt{TTbn+W-ord-closed-C4-mf} & Tenor trombone & wah (closed) & \\
+        \texttt{TTbn+W-ord-open-C4-mf} & Tenor trombone & wah (open) & \\ \hline
+        \texttt{Va-art-harm-C5-mf-4c} & Viola & & artificial harmonic \\
+        \texttt{Va-legno-batt-C4-mf-3c} & Viola & & col legno battuto \\
+        \texttt{Va-ord-C4-mf-3c} & Viola & & \\
+        \texttt{Va-pizz-bartok-C4-ff-3c} & Viola & & pizzicato Bart\'ok \\
+        \texttt{Va-pizz-lv-C4-mf-3c} & Viola & & pizzicato \& \emph{laissez vibrer} \\
+        \texttt{Va-pizz-sec-C4-mf-3c} & Viola & & pizzicato secco \\
+        \texttt{Va-pont-C4-mf-3c} & Viola & & sul ponticello \\
+        \texttt{Va+S-ord-C3-mf-3c} & Viola & sordina & \\
+        \texttt{Va+SP-ord-D4-mf-2c} & Viola & piombo & \\ \hline
+        \texttt{Vc-art-harm-C4-mf} & Cello & & artificial harmonic \\
+        \texttt{Vc-legno-batt-C4-mf-1c} & Cello & & col legno battuto \\
+        \texttt{Vc-legno-tratto-C4-mf-1c} & Cello & & col legno tratto \\
+        \texttt{Vc-nonvib-C4-mf-1c} & Cello & & nonvibrato \\
+        \texttt{Vc-ord-C4-mf-1c} & Cello & & \\
+        \texttt{Vc-pizz-bartok-C4-ff-1c} & Cello & & pizzicato Bart\'ok \\
+        \texttt{Vc-pizz-lv-C4-mf-1c} & Cello & & pizzicato \& \emph{laissez vibrer} \\
+        \texttt{Vc-pizz-sec-C4-mf-1c} & Cello & & pizzicato secco \\
+        \texttt{Vc-pont-C4-mf-2c} & Cello & & sul ponticello \\
+        \texttt{Vc-tasto-C4-mf-1c} & Cello & & sul tasto \\
+        \texttt{Vc+S-ord-C4-mf-1c} & Cello & sordina & \\
+        \texttt{Vc+SP-ord-C4-mf-1c} & Cello & piombo & \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{table*}
+\caption{Full list of audio stimuli (2/2).}
+\begin{tabular}{llll}
+          name & instrument  & mute   &  playing technique \\ \hline \hline
+          \texttt{Vn-art-harm-G5-mf-4c} & Violin & & artificial harmonic \\
+          \texttt{Vn-legno-batt-C4-mf-4c} & Violin & & col legno battuto \\
+          \texttt{Vn-nonvib-C4-mf-4c} & Violin & & nonvibrato \\
+          \texttt{Vn-ord-C4-mf-4c} & Violin & & \\
+          \texttt{Vn-pizz-bartok-C4-ff-4c} & Violin & & pizzicato Bart\'ok \\
+          \texttt{Vn-pizz-lv-C4-mf-4c} & Violin & pizzicato \& \emph{laissez vibrer} \\
+          \texttt{Vn-pizz-sec-C4-mf-4c} & Violin & pizzicato secco \\
+          \texttt{Vn-pont-C4-mf-4c} & Violin & & sul ponticello \\
+          \texttt{Vn+S-ord-C4-mf-4c} & Violin & sordina & \\
+          \texttt{Vn+SP-ord-C4-mf-4c} & Violin & piombo & \\ \hline
+\end{tabular}
+\end{table*}
+
 The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.

### 2019-10-14 paper/paper.tex
fix affiliations

@@
    addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},                       % id of corresponding address, if any
-   noteref={n1},                        % id's of article notes, if any
-   email={email@example.com}            % email address
+                        % id of corresponding address, if any
+%   noteref={n1},                        % id's of article notes, if any
+   email={vincent.lostanlen@nyu.edu}            % email address
 ]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
@@
    addressref={aff2},
-   email={email@example.com}
+   email={christian.elhajj@ls2n.fr}
 ]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
 \author[
-   addressref={aff2},
-   email={email@example.com}
+   addressref={aff3},
+   email={mathias.rossignol@lonofi.com}
 ]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
 \author[
-   addressref={aff2},
-   email={email@example.com}
+   addressref={aff3},
+   email={gregoire.lafay@lonofi.com}
 ]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
@@
 \author[
+  corref={aff2},
   addressref={aff2},
-   email={email@example.com}
+  email={mathieu.lagrange@ls2n.fr}
 ]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}
@@
 \address[id=aff2]{%
-  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
-  \street{D\"{u}sternbrooker Weg 20},
-  \postcode{24105}
-  \city{Kiel},
-  \cny{Germany}
+  \orgname{LS2N, CNRS, Centrale Nantes, Nantes University},
+  \street{1, rue de la Noe},
+  \postcode{44000},
+  \city{Nantes},
+  \cny{France}
+}
+\address[id=aff3]{%
+  \orgname{Lonofi},
+  \street{57 rue Letort},
+  \postcode{75018},
+  \city{Paris},
+  \cny{France}
 }
@@
 \begin{artnotes}
-\note[id=n1]{Equal contributor}
+%\note[id=n1]{Equal contributor}
 \end{artnotes}

### 2019-10-14 paper/paper.tex
more on figs

@@
 \includegraphics[width = \textwidth]{figures/xp2example1.png}
-\caption{Spatial and group organization provided by subject 1.}
+\caption{Spatial and group organization provided by subject 1\footnote{The organizations of the other subjects are available online \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
+For each subject, the stimuli can be played back by hovering on the dots.}.}
 \label{fig:xp2display}
@@
 
-It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
+In this study, we choose to condition the space using the aggregate similarity judgments given by several subjects.
 We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
@@
 \includegraphics[width = \textwidth]{figures/consensusVsI.png}
-\caption{}
+\caption{Normalized similarity matrix of the consensus clustering aggregated by the musical instruments used. Ordering is chosen to maximize readability.}
 \label{fig:consensusVsI}
@@
 \includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{ }
+\caption{Normalized similarity matrix of the consensus clustering aggregated by the playing techniques used. Ordering is chosen to maximize readability.}
 \label{fig:consensusVsPt}

### 2019-10-14 paper/paper.tex
introduce pipeline

@@
 
+\section*{Proposed approach}
+\label{sec:method}
+
+\begin{figure}[h!]
+\caption{\csentence{Model overview}
+One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
+scat, isomap.
+Second line : perceptual clustering.}
+\label{fig:pipeline}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.}
+\end{figure}
+
+Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
+The first component projects the audio signal of the query and the items of the database in a high dimensional timbre space using the scattering transform.
+The second component modify the timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:judgments}.
+It considers the large margin nearest neigbors to learn the weights of the reprojection.
+
+\subsection*{Scattering transform}
+\label{sec:scattering}
+
+% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
+
+While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
+As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
+
+A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
+Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
+The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
+This representation contains a large amount of information, but not in a form suitable for building statistical models.
+In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
+It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
+One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
+These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
+
+Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
+One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
+Again, we compute the modulus of the wavelet coefficients and average in time.
+The resulting coefficients are known as the second-order time scattering coefficients.
+These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
+
+While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
+Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
+If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
+We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
+This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
+The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
+
+This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
+To remedy this, another approach was suggested by And\'{e}n et al.
+\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
+Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
+As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
+
+The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
+On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
+The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
+Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+
+\subsection*{Similarity weighting}
+\label{sec:weighting}
+
+The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity.
+Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
+
+Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
+More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
+
+Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
+We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
+
+We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
+For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
+This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
+
 \section*{Timbre similarity judgments}
@@
 \end{figure}
-
-\section*{Proposed approach}
-\label{sec:method}
-
-\begin{figure}[h!]
-\caption{\csentence{Model overview}
-One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
-scat, isomap.
-Second line : perceptual clustering.}
-\label{fig:pipeline}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.}
-\end{figure}
-
-\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}
-
-\subsection*{Scattering transform}
-\label{sec:scattering}
-
-% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
-
-While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
-As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
-
-A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
-Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
-The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
-This representation contains a large amount of information, but not in a form suitable for building statistical models.
-In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
-It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
-One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
-
-Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
-One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
-Again, we compute the modulus of the wavelet coefficients and average in time.
-The resulting coefficients are known as the second-order time scattering coefficients.
-These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
-
-While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
-Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
-If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
-We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
-This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
-
-This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
-To remedy this, another approach was suggested by And\'{e}n et al.
-\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
-Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
-
-The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
-On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
-The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-
-\subsection*{Similarity weighting}
-\label{sec:weighting}
-
-The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity.
-Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
-
-Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
-More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
-
-Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
-We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
-
-We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
-For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
-This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
 

### 2019-10-14 paper/paper.tex
clean appendix

@@
 The first component projects the audio signal of the query and the items of the database in a high dimensional timbre space using the scattering transform.
-The second component modify the timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:judgments}.
+The second component further modifies the timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
 It considers the large margin nearest neigbors to learn the weights of the reprojection.
@@
 The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
+%For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
@@
 \includegraphics[width = \textwidth]{figures/xp2example1.png}
-\caption{Spatial and group organization provided by subject 1\footnote{The organizations of the other subjects are available online \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
-For each subject, the stimuli can be played back by hovering on the dots.}.}
+\caption{Spatial and group organization provided by subject 1.
+Contributions of the other subjects are available online \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
+For each subject, the stimuli can be played back by hovering on the dots.}
 \label{fig:xp2display}
@@
 
-\section*{Dataset}
-\label{sec:appendix-dataset}
+% \section*{Dataset}
+% \label{sec:appendix-dataset}
+%
+%
+% The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
+% In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
+% For each couple, the pitch and intonation is varied leading to 25444 audio samples.
+%
+% The recorded instruments belongs to the wind and string classes.
+% For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+% For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+% \ml{Better to display these in a table? VL todo table}
+%
+% Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+% For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+%
+% \vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+% The abbreviations can be removed. @data @writing}
+% \ml{}
+%
+% \vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+% (y axis) would be extremely useful. @data @figure}
+% \ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+% \vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
+
+\section*{Human timbre judgments}
+\label{sec:subjective}
+
+The dataset considered in this study is taken
+from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
+It has audio recordings of 16 different musical instruments played
+with different playing techniques,
+which leads to $235$ different couples of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to $25444$ recordings.
+%The way the dataset is structured is explained in more details in Appendix \ref{sec:appendix-dataset}.
+
+Gathering the similarity judgments between each pair of samples would be a daunting task.
+Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
+
+One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
+This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
+
+\section*{Reducing Expert judgments of technique peculiarities}
+
+\ja{This needs to be summarized (possible in the previous section).
+I'm copying two sentences in here that I had in the Timbre similarity judgments section.}%
+These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
+
+Even so, considering every pairs of IPTs is not practical. A selection
+experiment is thus conducted where the subject is asked to give his opinion on which
+\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
+if it is likely to be associated to another \ipt of another instrument.
+Interest is rated on a 7 ticks scale. The subject can listen to all the different
+nuance and pitch samples of each \ipt. The experiment is over when all the
+\ipts are rated.
+
+Rating guidelines are provided :
+\begin{itemize}
+  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
+  \item Four stars: there is a proximity on one aspect of the sound between this
+  \ipt and of an \ipt of another instrument,
+  but this one is neither decisive nor obvious.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
+\end{itemize}
+
+Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
 
 \begin{table}
+  \label{tab:list1}
 \caption{Full list of audio stimuli (1/2).}
@@
 \begin{table*}
+  \label{tab:list2}
 \caption{Full list of audio stimuli (2/2).}
@@
 
-The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
-In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
-For each couple, the pitch and intonation is varied leading to 25444 audio samples.
-
-The recorded instruments belongs to the wind and string classes.
-For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
-For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
-\ml{Better to display these in a table? VL todo table}
-
-Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
-For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
-
-\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
-The abbreviations can be removed. @data @writing}
-\ml{}
-
-\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
-(y axis) would be extremely useful. @data @figure}
-\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
-\vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
-
-\section*{Human timbre judgments}
-\label{sec:subjective}
-
-The dataset considered in this study is taken
-from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
-It has audio recordings of 16 different musical instruments played
-with different playing techniques,
-which leads to $235$ different couples of instrument / playing technique.
-For each couple, the pitch and intonation is varied leading to $25444$ recordings.
-The way the dataset is structured
-is explained in more details in Appendix \ref{sec:appendix-dataset}.
-
-Gathering the similarity judgments between each pair of samples would be a daunting task.
-Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
-
-One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
-This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
-
-\section*{Reducing Expert judgments of technique peculiarities}
-
-\ja{This needs to be summarized (possible in the previous section).
-I'm copying two sentences in here that I had in the Timbre similarity judgments section.}%
-These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
-The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
-
-Even so, considering every pairs of IPTs is not practical. A selection
-experiment is thus conducted where the subject is asked to give his opinion on which
-\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
-if it is likely to be associated to another \ipt of another instrument.
-Interest is rated on a 7 ticks scale. The subject can listen to all the different
-nuance and pitch samples of each \ipt. The experiment is over when all the
-\ipts are rated.
-
-Rating guidelines are provided :
-\begin{itemize}
-  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
-  \item Four stars: there is a proximity on one aspect of the sound between this
-  \ipt and of an \ipt of another instrument,
-  but this one is neither decisive nor obvious.
-  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
-\end{itemize}
-
-Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
-
-\vl{This subsection needs a box-and-whisker plot of ratings, where the 235 different
-couples of instrument and playing techniques are sorted on the x-axis by average rating. @figure @ratings}
-
-\ml{il n'y que deux raters, ca ne pas pas sens de faire et ca sera illisible}
-
-\vl{Are the histogram of ratings similar for the two raters? It would be good to make a statistical test for this. @research @stats @ratings}
-
-\ml{les histograms sont dans la figure statxp1. Le code pour la generer est dans code/xp1 faire tourner extractData et statData. Un ttest rejette l'hypothese nulle.}
-
 Even if the hypothesis that the distributions of ratings comes from independent random samples from normal distributions with equal means and equal but unknown variances, is verified using the two-sample t-test, inspection of the respective histograms reveals that averaging the ratings of the two raters is reasonable.
@@
 Keeping in mind that the gathering of the perceptual judgments of \ipt similarities shall remain practical, the number of \ipts considered shall be lower than $100$.
-We thus select the \ipts with an average rating higher than three, which brings us 78 \ipts to study.
+We thus select the \ipts with an average rating higher than three, which brings us 78 \ipts to study which are listed in Tables \ref{tab:list1} and \ref{tab:list2}.
 
@@
 Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipts.
-The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbGroup}.
+The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbSize}(a).
 The size of the groups is on average $7.7 \pm   7.2$.
-The distribution of the size of groups can be seen on Figure \ref{fig:xp2sizeGroup}.
+The distribution of the size of groups can be seen on Figure \ref{fig:xp2nbSize}(b).
 The large number of small groups indicates that many subjects considered that a few of the \ipt were very different from all the remaining \ipt.
@@
 \center
-\includegraphics[width = \textwidth]{figures/nbc.png}
-\caption{Histogram of the number of groups.
-\vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
-\ml{ok}
-\label{fig:xp2nbGroup}
-\end{figure}
-
-\begin{figure}
-\center
+\begin{tabular}{cc}
+  (a) & (b) \\
+  \includegraphics[width = \textwidth]{figures/nbc.png} &
   \includegraphics[width = \textwidth]{figures/sbc.png}
-\caption{Histogram of the size of groups.
-\vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
-\ml{ok}
-\label{fig:xp2sizeGroup}
+\end{tabular}
+\caption{Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects.}
+\label{fig:xp2nbSize}
 \end{figure}

### 2019-10-19 paper/paper.tex
Comments and small edits

@@
 
-\newcommand{\ml}[1]{\textcolor{blue}{ML: #1}}
-\newcommand{\vl}[1]{\textcolor{red}{VL: #1}}
-\newcommand{\ja}[1]{\textcolor{purple}{JA: #1}}
+\newcommand{\Card}{\mathrm{Card}}
+
+\newcommand{\ml}[1]{\textcolor{blue}{ML: #1}\xspace}
+\newcommand{\vl}[1]{\textcolor{red}{VL: #1}\xspace}
+\newcommand{\ja}[1]{\textcolor{purple}{JA: #1}\xspace}
 
@@
 However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, more sophisticated models which incorporate dynamic large-scale signal structure are needed.
-To this end, we propose a simple but rich model based on joint time--frequency scattering transforms that are linearly transformed to yield perceptively motivated timbral descriptors as the linear mapping is learned from human similarity judgments obtained in a crowdsourced free sorting task.
+To this end, we propose a simple but rich model based on joint time--frequency scattering transforms followed by a linear mapping.
+The linear part is learned from human similarity judgments obtained in a crowdsourced free sorting task, so the resulting descriptor provides a perceptually motivated characterization of timbre.
 We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
@@
 
-Music information retrieval operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
-By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, or intensity at the source.
-Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis \cite{siedenburg2019chapter}.
+Music information retrieval (MIR) operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
+By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
+Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
 As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
@@
 On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
-On the other hand, the sole mention of a playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
+On the other hand, the sole mention of a playing technique does not determine any intended effect; as a result, it fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
@@
 The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
-Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal \cite{thoret2018jasa}.
+Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
 
@@
 The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
-In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
+In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of predetermined range.
 
 This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
-Our system composes two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
+Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
@@
 Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
-Said otherwise, we rely on listeners, not performers, to define and evaluate the task at hand.
+In other words, we rely on listeners, not performers, to define and evaluate the task at hand.
 A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
 Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
-These visualizations confirm the theoretical predictions of previous research \cite{anden2015mlsp}, which relied on simple synthetic sounds such as exponential chirps.
+These visualizations confirm the theoretical predictions of previous research \cite{anden2019tsp}, which relied on simple synthetic sounds such as exponential chirps.
 They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.
 
+\ja{These section numbers aren't showing up since we're not numbering the sections (per the style guide).}
 Section \ref{sec:soa} reviews recent research in the computational modeling of musical timbre.
@@
 Section \ref{sec:results} reports the results of an ablation study of our system.
-Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}.
+Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data at \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}
 
@@
 On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
-In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags'', encoded as binary outputs within some predefined label space.
+In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags,'' encoded as binary outputs within some predefined label space.
 Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
@@
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning across 20 instruments, as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
-Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments, as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
+For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20000 polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
+Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
 Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
@@
 Rather, each excerpt must be played back to multiple human listeners.
-Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, thus restricting the size of the musical corpus under study.
-Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in artificial neural networks.
+Yet collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
+Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
 Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
+\ja{This is quite an abrupt transition.
+Why ``nevertheless''?}%
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
 The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
-The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
+The simplest method to compute it in practice is by reverse correlation, i.e.\ by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
 Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but eliciting cross-term interferences \cite{eggermont1993hearing}.
 Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
-Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner-Ville distribution.
+Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner--Ville distribution.
 In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
+\ja{``Spectrographic''?}
 This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
@@
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable.Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
+These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable.
+Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
 Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
@@
 This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
-Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
-Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
-Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
+Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
+Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014deep}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
+Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
 In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
+\ja{I don't understand this last sentence.
+What does ``remotely'' mean here?}%
 
 The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
+From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
 Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
+
+\ja{This section is way too long.
+It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
+This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
+We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
+}
 
@@
 Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
-The first component projects the audio signal of the query and the items of the database in a high dimensional timbre space using the scattering transform.
-The second component further modifies the timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
-It considers the large margin nearest neigbors to learn the weights of the reprojection.
+The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
+The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
+This warping, or reweighting, of the space is performed using LMNN.
+\ja{Again, section numbers are not showing up here.}%
 
@@
 They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+\ja{Need to flesh out the above with some equations.
+We're no longer talking to biologists, so we don't need to be afraid.}%
 
@@
 
-The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity.
-Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
+The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
+We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
 
@@
 The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
-In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
+\ja{European or Western?}%
+In this article, we focus on the first three, and leave the question of learning auditory similarities between percussion instruments to future research.
 The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
@@
 We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive sounds, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
+\ja{Is it really necessary to go into such detail only to explain why we're not dealing with percussive sounds?
+One sentence should be enough.}%
 
-We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
+We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), C trumpet (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
+Among this list, the first six are string instruments, the next five are woodwinds, and the last five are brass instruments.
 Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
@@
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
-Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
-In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
+Each dot corresponds to a different IPT recording and is moved by the user dragging them with the mouse.
+In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple measure of clustering of the sounds.
 While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
 More information on the data collection process is available in Section \ref{sec:subjective}.
+\ja{Missing section number.}%
 
@@
 \caption{Spatial and group organization provided by subject 1.
-Contributions of the other subjects are available online \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
+Other subjects are available at \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
 For each subject, the stimuli can be played back by hovering on the dots.}
@@
 
-In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
-For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
-The PTs similarity matrix is computed in the same way.
+In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used (see Figure \ref{fig:consensusVsI}) and one for the 23 PTs used (see Figure \ref{fig:consensusVsPt}).
+For the instrument similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that involved the two instruments.
+The PT similarity matrix is computed in the same way.
 
@@
 This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.
+\ja{Do we need to talk about this?
+It's just for visualization.
+If we want to keep it, we should provide references to the methods used.}%
 
@@
 \ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
+
+\ja{Should we loop back here and explain how the consensus clustering is used as input to the LMNN?
+Feels like we've lost the thread a bit here otherwise.}%
 
@@
 The relevance of the many design choices that had to be made are addressed in this section using an experimental protocol that considers a dataset of audio that will be given as input to the processing pipeline and one metric used for evaluating its performance under a given experimental setting.
+\ja{This last sentence is quite vague.
+What are we trying to say?}%
 
@@
 
-The perceptual data gathered during the perceptual experiment considers 78 \ipts each represented by one sound recording that the subject could listen to before clustering the \ipts.
-For those 78 \ipts, 31 clusterings $C_k, 0<k<31, Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
+The perceptual data gathered during the perceptual experiment considers 78 \ipts, each represented by one sound recording that the subject could listen to before clustering the \ipts.
+For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
+\ja{What is $C_k$ here?
+Also, should it be $0 < k \le 31$?
+Might be more natural to say that $k = 1, 2, \ldots, 31$, in that case.}%
 If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
@@
 In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
-Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation.
-reformuler tout le paragraphe}, it is  thus taken in this paper as a reasonable assumption.
+Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation.
+reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
 Quantification of this invariance is left for further study.
@@
 The whole range of nuance and pitch is thus considered for each of the 78 \ipts, leading to a dataset composed of 9346 sound recordings.
-The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, Card(D_k)=9346$ using the following scheme.
+The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, \Card(D_k)=9346$ using the following scheme.
 If a sound of this dataset $s_a$ corresponds to \ipt $i$, then $D_k(a) = C_k(i), \forall k$.
+\ja{What's $s_a$ and $a$ here?
+We need to be explicit about these definitions.}%
 
@@
 
-The metric considered in this paper is the average precision at rank $k=5$ (ap@5).
+The metric considered in this work is the average precision at rank $k=5$ (ap@5).
 It allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
@@
 $$
-p@5_q = \frac{\sum_{n=5}^5 D_q==D_n}{5}
+p@5_q = \frac{1}{5} \sum_{n=1}^5 \mathbf{1}_{D_q=D_n}
 $$
 The p@5 is the average $p@5_q$ computed over all queries.
+\ja{This is perhaps better described in reverse, starting with $p@5_q$, then $p@5$, then $ap@5$?
+Also, we define the precisions here as proportions (between zero and one), but later, we present them as percentages (between $0\%$ and $100\%$).}%
 
@@
 When searching for sounds that are similar to a query, the listener is likely to listen to up to $5$ proposals, rarely more.
-For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision is not considered here.
+For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision are not considered here.
 
@@
 
-The processing pipeline is composed of 2 main building blocks.
+The processing pipeline is composed of two main building blocks.
 The first block $B_1$ takes a digital audio recording as input and outputs one vector of features.
 For this block, the main factors are the time support $B_1^t$ for computing the features and the type of features $B_1^f$.
-We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{mfccs}, \text{separable scattering}, \text{joint scattering} \}$.
-The mfccs are the standard Mel frequency cepstral coefficients with 40 Mel bands and selection of the 13$^\text{th}$ first discrete cosine transform (dct) coefficients.
+We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{MFCCs}, \text{separable scattering}, \text{joint scattering} \}$.
+The MFCCs are the standard mel-frequency cepstral coefficients with $40$ coefficients out of which we select the first $13$.
+\ja{I'm not sure I like this notation.
+First, why do we need to label the blocks ($B_1$ and $B_2$)?
+We only have two of them.
+Isn't it more straightforward to just call them ``the unsupervised part'' and ``the learned part'' or something like that?
+We can also replace $B_1^t$ with something more straightforward, like $T$ (otherwise, it's a lot of bits to encode very little information).
+Likewise, instead of $B_1^f$, we can just have $F$.}
 
@@
 As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
+\ja{References.}
 
-For each sound, a set of features are computed for each time support. This set is averaged over time to give one feature vector per sound.
+For each recording, a set of features are computed for each time frame.
+This set is averaged over time to give one feature vector per recording.
 All features are standardized with zero mean and unit variance across the dataset.
+\ja{Is this the ``per-feature Gaussianization'' we mention in the introduction?
+Should give the reference in that case (and maybe explain a bit more in detail).}%
 
 The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output.
-For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$.
-The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{sec:weighting}.
+For this block, the main factor is the type of projection $B_2^p \in \{ \text{LDA}, \text{LMNN} \}$.
+The LDA is the linear discriminant analysis \cite{} and the LMNN is the algorithm presented in Section \ref{sec:weighting}.
 For this algorithm the neighborhood is set to $5$.
+\ja{Again, I would prefer to ditch this $B_2$ notation.}
 
-The complete experiments with all settings have been computed. For the sake of clarity, only part of the results are discussed here.
+The complete experiments with all settings have been computed.
+For the sake of clarity, only part of the results are discussed here.
+\ja{What does this mean?}%
 
-The best performing system considers the separable scattering computed over $1000$ms time support, projected using the lmnn algorithm.
+The best performing system considers the separable scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
+\ja{Separable scattering?
+Not the joint?}%
 Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.5\%$ of $ap@5$.
 That means that, for a given sound dataset of \ipts for which subjective evaluation of similarity between \ipts are available, a very large number of queries would return interesting results to the user using this processing pipeline.
+\ja{What conclusion do we draw from this?}
 
 In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol.
-In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those.
-In our case, two factors can readily be identified, the variability of the audio data and the one of the perceptual judgments.
+In order to reduce the risk of misleading performance results using learning systems due to overfitting, we identify factors of variability that could lead to such problems and mitigate those.
+\ja{Is this part of the ablation study or is this a ``meta'' consideration?}
+In our case, two factors can readily be identified, the variability of the audio data and that of the perceptual judgments.
 
@@
 If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\%$, indicating a low risk of overfitting.
+\ja{We can expect better performance by using larger folds for the training (i.e., not restricting ourselves to 2-fold cross validation).}
 
-Another risk when considering reprojection techniques such as the lmnn may appear when the dimensionality of the features gets closer to the  number of items in the dataset.
-For example there exists a trivial solution for optimizing the p@1 when the dimensionality of the features is equal to the number of items in the dataset.
+Another risk when considering reprojection techniques such as the LMNN may appear when the dimensionality of the features approaches the number of items in the dataset.
+For example, there exists a trivial solution for optimizing $p@1$ when the dimensionality of the features is equal to the number of items in the dataset.
+\ja{What is that trivial solution?}%
 In this case, no information needs to be carried by the features set to obtain perfect performance.
@@
 In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-In this case, the performance drops down to $23.6\%$.
-Not considering the learning block gives a $aP@5$ of $23.8\%$.
-It can be concluded from these results that the learning block is not able to perform effectively if no information is carried by the input features set.
+In this case, the performance drops to $23.6\%$.
+Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
+It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
+The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
 
-The slight decrease of performance when considering lmnn may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
+The LMNN algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
+One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
+When considering LDA in place of LMNN, the performance drops to $81.2\%$.
+The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
 
-The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
-One of those is the linear discriminant analysis \ml{vl ref ?}\cite{}.
-When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.
-
-The lower performance of the lda may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from 1180 to 20.
-
-Considering a learning block implies some constraints.
-Performing the learning on large datasets require some computation.
-The projection has to be updated if the reference clustering or the dataset are modified significantly.
-Not considering any learning block reduces the performance to $92.9\% \pm 3$.
+Including a learned block implies some constraints.
+In particular, learning the linear transform on large datasets requires more computational time.
+Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
+It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
+In this case, the performance reduces to $92.9\% \pm 3$.
+\ja{Why the $\pm$ here and not elsewhere?}%
 
 Concerning the feature computation block $B_1$, an important factor is the time support.
-When considering the commonly used duration of 25ms, the performance drop down to $92\% \pm$.
-This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture long term modulations which plays an important part when considering extended playing techniques.
+When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $92\% \pm$.
+\ja{$\pm$ what?}%
+This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.
+\ja{Other references?}%
 
@@
 
-Simplifying further the feature set by considering the mfccs gives a performance of $82.3\% \pm 7.1$.
-Considering the mfccs is equivalent to considering the separable scattering at order 1.
-Compared to the separable scattering, the information captured by the mfccs is considerably reduced.
-In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the mfccs are complemented with monomials in order to reach the dimensionality of the joint scattering.
-To do so, the 40 features $f_i$ of the mfccs are multiplied together $f_{i, j} = f_i*f_j, i \neq j$.
-The resulting features set is composed of the 40 mfccs features $f_i$ and the first 1140 $f_{i, j}$.
-The resulting aP@5 of $81.8\% \pm 7$ indicates that the increase of dimensionality is likely to have no effect, confirming the relevance of the additional information provided by the joint scattering.
+Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $82.3\% \pm 7.1$.
+Note that the MFCCs are equivalent to first-order separable scattering coefficients \cite{anden2014tsp}.
+Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
+In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
+\ja{Explain why this is necessary.}%
+To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i*f_j, i \neq j$.
+\ja{What are the $f_i$ here?
+How are they defined?}%
+The resulting features set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
+The resulting $ap@5$ of $81.8\% \pm 7$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
 \ml{vl outro on modulations}

### 2019-10-21 paper/paper.tex
punctuation in intro

--- 2019-10-14 paper/paper.tex
@@
    addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-                        % id of corresponding address, if any
-%   noteref={n1},                        % id's of article notes, if any
-   email={vincent.lostanlen@nyu.edu}            % email address
+   corref={aff1},                       % id of corresponding address, if any
+   noteref={n1},                        % id's of article notes, if any
+   email={email@example.com}            % email address
 ]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
@@
    addressref={aff2},
-   email={christian.elhajj@ls2n.fr}
+   email={email@example.com}
 ]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
 \author[
-   addressref={aff3},
-   email={mathias.rossignol@lonofi.com}
+   addressref={aff2},
+   email={email@example.com}
 ]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
 \author[
-   addressref={aff3},
-   email={gregoire.lafay@lonofi.com}
+   addressref={aff2},
+   email={email@example.com}
 ]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
@@
 \author[
-  corref={aff2},
    addressref={aff2},
-  email={mathieu.lagrange@ls2n.fr}
+   email={email@example.com}
 ]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}
@@
 \address[id=aff2]{%
-  \orgname{LS2N, CNRS, Centrale Nantes, Nantes University},
-  \street{1, rue de la Noe},
-  \postcode{44000},
-  \city{Nantes},
-  \cny{France}
-}
-\address[id=aff3]{%
-  \orgname{Lonofi},
-  \street{57 rue Letort},
-  \postcode{75018},
-  \city{Paris},
-  \cny{France}
+  \orgname{Marine Ecology Department, Institute of Marine Sciences Kiel},
+  \street{D\"{u}sternbrooker Weg 20},
+  \postcode{24105}
+  \city{Kiel},
+  \cny{Germany}
 }
@@
 \begin{artnotes}
-%\note[id=n1]{Equal contributor}
+\note[id=n1]{Equal contributor}
 \end{artnotes}
@@
 Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis \cite{siedenburg2019chapter}.
-As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
+As a result, symbolic representations describe timbre indirectly: either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
 
@@
 
+\section*{Timbre similarity judgments}
+\label{sec:survey}
+
+The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
+In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
+The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
+Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
+In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
+We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive sounds, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
+
+We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
+Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
+Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
+Once augmented with mutes, the list of 16 instruments grows up to X. % TODO complete
+Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between X (for the X) and Y (for the Y). % TODO update with number.
+In total, the feasible triplets of instrument, mute, and technique amount to a list of $N=235$ combinations.
+
+% TODO explain the conundrum leading to reduce the dataset size.
+
+To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
+The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
+For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
+In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
+Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
+In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
+While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
+More information on the data collection process is available in  Section \ref{sec:subjective}.
+
+The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
+We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
+The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/xp2example1.png}
+\caption{Spatial and group organization provided by subject 1.}
+\label{fig:xp2display}
+\end{figure}
+
+For each subject, a matrix of size $78 \times 78$ is constructed to represent the similarity judgments for each pair of the $78$ IPTs.
+Specifically, the $(i, j)$th element of the matrix was set to one if the $i$th and $j$th IPT were assigned the same color.
+If not, it is set to zero.
+In other words, the clustering assignment of each subject as a binary similarity matrix.
+Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
+In this case, the corresponding binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
+
+It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
+We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
+
+In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
+For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
+The PTs similarity matrix is computed in the same way.
+
+The ordering of the instruments or the PTs on the vertical and horizontals axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
+This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.
+
+Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/consensusVsI.png}
+\caption{}
+\label{fig:consensusVsI}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
+\caption{ }
+\label{fig:consensusVsPt}
+\end{figure}
+
 \section*{Proposed approach}
@@
 
-Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
-The first component projects the audio signal of the query and the items of the database in a high dimensional timbre space using the scattering transform.
-The second component modify the timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:judgments}.
-It considers the large margin nearest neigbors to learn the weights of the reprojection.
+\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}
 
@@
 This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-
-\section*{Timbre similarity judgments}
-\label{sec:survey}
-
-The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
-In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
-The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
-Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
-In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
-We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive sounds, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
-
-We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
-Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
-Once augmented with mutes, the list of 16 instruments grows up to X. % TODO complete
-Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between X (for the X) and Y (for the Y). % TODO update with number.
-In total, the feasible triplets of instrument, mute, and technique amount to a list of $N=235$ combinations.
-
-% TODO explain the conundrum leading to reduce the dataset size.
-
-To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
-The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
-In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
-Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
-In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
-While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
-More information on the data collection process is available in  Section \ref{sec:subjective}.
-
-The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
-We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
-The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/xp2example1.png}
-\caption{Spatial and group organization provided by subject 1\footnote{The organizations of the other subjects are available online \url{https://mathieulagrange.github.io/paperPlayingTechniquesSimilarityLearning/demo}.
-For each subject, the stimuli can be played back by hovering on the dots.}.}
-\label{fig:xp2display}
-\end{figure}
-
-For each subject, a matrix of size $78 \times 78$ is constructed to represent the similarity judgments for each pair of the $78$ IPTs.
-Specifically, the $(i, j)$th element of the matrix was set to one if the $i$th and $j$th IPT were assigned the same color.
-If not, it is set to zero.
-In other words, the clustering assignment of each subject as a binary similarity matrix.
-Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
-In this case, the corresponding binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
-
-In this study, we choose to condition the space using the aggregate similarity judgments given by several subjects.
-We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
-
-In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
-For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
-The PTs similarity matrix is computed in the same way.
-
-The ordering of the instruments or the PTs on the vertical and horizontals axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
-This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.
-
-Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
-
-\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/consensusVsI.png}
-\caption{Normalized similarity matrix of the consensus clustering aggregated by the musical instruments used. Ordering is chosen to maximize readability.}
-\label{fig:consensusVsI}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{Normalized similarity matrix of the consensus clustering aggregated by the playing techniques used. Ordering is chosen to maximize readability.}
-\label{fig:consensusVsPt}
-\end{figure}
 

### 2019-10-21 paper/paper.tex
todo add more refs on orchestration
proofread related work section

@@
 In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
+% TODO add more references to automatic orchestration?
+Crestel and Esling?
 
@@
 The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
-It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}.
+It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from a collective consensus.
 
@@
 Section \ref{sec:results} reports the results of an ablation study of our system.
-Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol used to produce the results discussed in this paper are available alongside the anonymized version of the perceptual data: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}}.
+Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects: \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
 
@@
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features  outperform the state of the art, especially in the small data regime when deep learning is inapplicable.Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
+These biologically inspired features  outperform the state of the art, especially in the small data regime where deep learning is inapplicable.
+Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
 Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
@@
 This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
-Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
-Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
-Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
-In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
+Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
+Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results on a dataset of continuous performance recordings of the Chinese bamboo flute, involving periodic spectrotemporal modulations such as vibratos and trills.
+Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this finding can scale up from a single instrument to an entire orchestra.
+In addition, all previous publications in time--frequency scattering, whether separable or joint, lack a human-centric evaluation, remotely from any classification task.
 
@@
 From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
-
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate \emph{in silico} for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).

### 2019-10-21 paper/paper.tex
write on data collection

@@
 In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
-We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive sounds, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
+We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
 
 We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-Among this list, the former six are strings, the next five are woodwind, and the last five are brass.
-Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'' inside in the bell of a trumpet.
+Among this list, the former six are strings, the next six are woodwind, and the last four are brass.
+Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'', also known as \emph{harmon}, inside in the bell of a trumpet.
 Once augmented with mutes, the list of 16 instruments grows up to X. % TODO complete
 Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between X (for the X) and Y (for the Y). % TODO update with number.
-In total, the feasible triplets of instrument, mute, and technique amount to a list of $N=235$ combinations.
+In the rest of this paper, we abbreviate instrument--mute--technique by means of the acronym ``IMT''.
+One example of IMT is \texttt{TpC+S-ord}, \ie{} trumpet in C with a straight mute played in the ordinary technique.
+Another example of IMT is \texttt{Vn-pont}, \ie{} violin without any mute played in the \emph{sul ponticello} (bowing near the bridge) technique.
 
-% TODO explain the conundrum leading to reduce the dataset size.
+Performers can play each IMT at various pitches, according to a tessitura that depends on the instrument but not on the choice of mute.
+Among the 16 instruments in this study, the two instruments with widest and narrowest tessituras, in their respective ordinary techniques, are the X (X semitones) and the Y (Y semitones) respectively.
+% TODO update with numbers, check the plural of tessitura
+Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzoforte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).
+However, the resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
+For example, the pitch of pedal tones in brass instruments is tied to the fundamental mode of the bore, \ie{} usually B-flat or F. % TODO check this, use special package for musical pitches
+Likewise, the intensity of key clicks in the oboe is necessarily \emph{pp}, while the intensity of snap pizzicato \emph{\`a la Bart\'ok} in plucked strings is necessarily \emph{ff}.
+
+In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
+The set product of all five variables results in a combinatorial explosion of possible timbral sensations. % TODO verify ``set product'' terminology
+The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total.
+% TODO check year in which SOL was recorded, update with number
+We should note that this dense sampling comes at the detriment of other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
+Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
+Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and efficient annotation interface.
+
+First, we purposefully disentangle categorical variables (IMTs) from continuous variables (pitch and intensity) in the SOL dataset.
+Indeed, under first approximation, the perception of timbre is invariant to pitch and intensity.
+Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C (\ie{} C4) and \emph{mf}. % TODO use pitch spelling package
+After this selection, every IMT triplet contains a single acoustic exemplar, regarded as canonical in the following.
+The number of canonical stimuli for the entire SOL dataset is equal to 235.
+We should note, however, that the proposed pitch--intensity cannot be strictly enforced across all IMTs.
+Indeed, as explained above, a fraction of IMTs can only be achieved at restricted values of pitch and intensity parameters, \eg{} pedal tones or key clicks.
+Therefore, at a small cost of consistency, we only enforce the pitch--intensity reference (\ie{} C4 and \emph{mf}) when practically feasible, and fall back to other pitches and intensities if necessary.
+
+Secondly, we reduce the number of IMTs in our study by focusing on those which are deemed to be most relevant.
+Here, we define the relevance of an IMT as the possibility of imitating it by means of another IMT from a different instrument.
+One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in strings instruments.
+To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP), and explicitly gave them our definition of relevance.
+Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
+In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
+Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments.
+Once both experts completed their annotation, we retained all IPTs whose average relevance score was judged equal to 3 or higher.
+At the end of this stage of pre-screening, the list of IMT reduces from 235 to 78.
+Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
+% TODO explain that the accordion was judged too idiosyncratic altogether. because it's a free reed with keyboard activation and handheld airflow.
+
+% one paragraph on annotation interface
+Thirdly,
+% in 30 to 60 minutes.
 
@@
 \begin{table}
-\caption{Full list of audio stimuli (1/2).}
+\caption{Full list of audio stimuli (1/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
       \begin{tabular}{llll}
@@
 		\texttt{Bn-ord-C4-mf} & Bassoon & & \\ \hline
-		\texttt{Cb-pizz-bartok-C4-ff-1c} & Contrabass & & pizzicato Bart\'ok (snap) \\
+		\texttt{Cb-pizz-bartok-C4-ff-1c} & Contrabass & & pizzicato \`a la Bart\'ok (snap) \\
 		\texttt{Cb-pizz-lv-C4-mf-1c} & Contrabass & & pizzicato \& \emph{laissez vibrer} \\
@@
 \begin{table*}
-\caption{Full list of audio stimuli (2/2).}
+\caption{Full list of audio stimuli (2/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
 \begin{tabular}{llll}

### 2019-10-28 paper/paper.tex

@@
 
-Music information retrieval operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
-By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, or intensity at the source.
-Timbre, in contrast, is a qualitative attribute of music, thus irreducible to a one-dimensional axis \cite{siedenburg2019chapter}.
-As a result, symbolic representations describe timbre indirectly: either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
+Music information retrieval (MIR) operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
+By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
+Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
+As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
 
@@
 This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
-Our system composes two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
+Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
-Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results on a dataset of continuous performance recordings of the Chinese bamboo flute, involving periodic spectrotemporal modulations such as vibratos and trills.
-Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this finding can scale up from a single instrument to an entire orchestra.
-In addition, all previous publications in time--frequency scattering, whether separable or joint, lack a human-centric evaluation, remotely from any classification task.
+Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014deep}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
+Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
+In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
+\ja{I don't understand this last sentence.
+What does ``remotely'' mean here?}%
 
 The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate \emph{in silico} for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
+From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
+
+\ja{This section is way too long.
+It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
+This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
+We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
+}
+
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
@@
 
+\section*{Proposed approach}
+\label{sec:method}
+
+\begin{figure}[h!]
+\caption{\csentence{Model overview}
+One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
+scat, isomap.
+Second line : perceptual clustering.}
+\label{fig:pipeline}
+\end{figure}
+
+\begin{figure}[h!]
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.}
+\end{figure}
+
+Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
+The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
+The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
+This warping, or reweighting, of the space is performed using LMNN.
+\ja{Again, section numbers are not showing up here.}%
+
+\subsection*{Scattering transform}
+\label{sec:scattering}
+
+% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
+
+While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
+As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
+
+A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
+Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
+The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
+This representation contains a large amount of information, but not in a form suitable for building statistical models.
+In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
+It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
+One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
+These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
+
+Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
+One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
+Again, we compute the modulus of the wavelet coefficients and average in time.
+The resulting coefficients are known as the second-order time scattering coefficients.
+These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
+
+While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
+Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
+If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
+We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
+This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
+The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
+
+This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
+To remedy this, another approach was suggested by And\'{e}n et al.
+\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
+Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
+As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
+
+The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
+On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
+The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
+Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+\ja{Need to flesh out the above with some equations.
+We're no longer talking to biologists, so we don't need to be afraid.}%
+
+\subsection*{Similarity weighting}
+\label{sec:weighting}
+
+The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
+We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
+
+Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
+More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
+
+Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
+We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
+
+We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
+For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
+This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
+
 \section*{Timbre similarity judgments}
@@
 
+\section*{Timbre similarity judgments}
+\label{sec:survey}
 The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
@@
 % one paragraph on annotation interface
+
 Thirdly,
@@
 
-\section*{Proposed approach}
-\label{sec:method}
-
-\begin{figure}[h!]
-\caption{\csentence{Model overview}
-One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
-scat, isomap.
-Second line : perceptual clustering.}
-\label{fig:pipeline}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.}
-\end{figure}
-
-\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}
-
-\subsection*{Scattering transform}
-\label{sec:scattering}
-
-% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
-
-While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
-As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
-
-A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
-Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
-The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
-This representation contains a large amount of information, but not in a form suitable for building statistical models.
-In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
-It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
-One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
-
-Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
-One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
-Again, we compute the modulus of the wavelet coefficients and average in time.
-The resulting coefficients are known as the second-order time scattering coefficients.
-These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
-
-While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
-Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
-If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
-We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
-This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
-
-This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
-To remedy this, another approach was suggested by And\'{e}n et al.
-\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
-Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
-
-The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
-On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
-The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering tranforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-
-\subsection*{Similarity weighting}
-\label{sec:weighting}
-
-The standard Euclidean distance on the raw scattering coefficients  provides some sort universal representation of timbre dissimilarity.
-Considering some sort of learning to apply further transformations to the scattering coefficients and compute the Euclidean distance on the result.
-
-Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
-More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
-
-Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
-We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
-
-We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
-For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
-This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-
 \section*{Results}
@@
 As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
-
-For each sound, a set of features are computed for each time support. This set is averaged over time to give one feature vector per sound.
+\ja{References.}
+
+For each recording, a set of features are computed for each time frame.
+This set is averaged over time to give one feature vector per recording.
 All features are standardized with zero mean and unit variance across the dataset.
+\ja{Is this the ``per-feature Gaussianization'' we mention in the introduction?
+Should give the reference in that case (and maybe explain a bit more in detail).}%
 
 The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output.
-For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$.
-The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{sec:weighting}.
+For this block, the main factor is the type of projection $B_2^p \in \{ \text{LDA}, \text{LMNN} \}$.
+The LDA is the linear discriminant analysis \cite{} and the LMNN is the algorithm presented in Section \ref{sec:weighting}.
 For this algorithm the neighborhood is set to $5$.
-
-The complete experiments with all settings have been computed. For the sake of clarity, only part of the results are discussed here.
-
-The best performing system considers the separable scattering computed over $1000$ms time support, projected using the lmnn algorithm.
+\ja{Again, I would prefer to ditch this $B_2$ notation.}
+
+The complete experiments with all settings have been computed.
+For the sake of clarity, only part of the results are discussed here.
+\ja{What does this mean?}%
+
+The best performing system considers the separable scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
+\ja{Separable scattering?
+Not the joint?}%
 Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.5\%$ of $ap@5$.
 That means that, for a given sound dataset of \ipts for which subjective evaluation of similarity between \ipts are available, a very large number of queries would return interesting results to the user using this processing pipeline.
+\ja{What conclusion do we draw from this?}
 
 In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol.
-In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those.
-In our case, two factors can readily be identified, the variability of the audio data and the one of the perceptual judgments.
+In order to reduce the risk of misleading performance results using learning systems due to overfitting, we identify factors of variability that could lead to such problems and mitigate those.
+\ja{Is this part of the ablation study or is this a ``meta'' consideration?}
+In our case, two factors can readily be identified, the variability of the audio data and that of the perceptual judgments.
 
@@
 If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\%$, indicating a low risk of overfitting.
-
-Another risk when considering reprojection techniques such as the lmnn may appear when the dimensionality of the features gets closer to the  number of items in the dataset.
-For example there exists a trivial solution for optimizing the p@1 when the dimensionality of the features is equal to the number of items in the dataset.
+\ja{We can expect better performance by using larger folds for the training (i.e., not restricting ourselves to 2-fold cross validation).}
+
+Another risk when considering reprojection techniques such as the LMNN may appear when the dimensionality of the features approaches the number of items in the dataset.
+For example, there exists a trivial solution for optimizing $p@1$ when the dimensionality of the features is equal to the number of items in the dataset.
+\ja{What is that trivial solution?}%
 In this case, no information needs to be carried by the features set to obtain perfect performance.
@@
 In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-In this case, the performance drops down to $23.6\%$.
-Not considering the learning block gives a $aP@5$ of $23.8\%$.
-It can be concluded from these results that the learning block is not able to perform effectively if no information is carried by the input features set.
-
-The slight decrease of performance when considering lmnn may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
-
-The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
-One of those is the linear discriminant analysis \ml{vl ref ?}\cite{}.
-When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.
-
-The lower performance of the lda may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from 1180 to 20.
-
-Considering a learning block implies some constraints.
-Performing the learning on large datasets require some computation.
-The projection has to be updated if the reference clustering or the dataset are modified significantly.
-Not considering any learning block reduces the performance to $92.9\% \pm 3$.
+In this case, the performance drops to $23.6\%$.
+Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
+It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
+The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
+
+The LMNN algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
+One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
+When considering LDA in place of LMNN, the performance drops to $81.2\%$.
+The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
+
+Including a learned block implies some constraints.
+In particular, learning the linear transform on large datasets requires more computational time.
+Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
+It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
+In this case, the performance reduces to $92.9\% \pm 3$.
+\ja{Why the $\pm$ here and not elsewhere?}%
 
 Concerning the feature computation block $B_1$, an important factor is the time support.
-When considering the commonly used duration of 25ms, the performance drop down to $92\% \pm$.
-This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture long term modulations which plays an important part when considering extended playing techniques.
+When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $92\% \pm$.
+\ja{$\pm$ what?}%
+This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.
+\ja{Other references?}%
 
@@
 
-\section*{Dataset}
-\label{sec:appendix-dataset}
+% \section*{Dataset}
+% \label{sec:appendix-dataset}
+%
+%
+% The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
+% In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
+% For each couple, the pitch and intonation is varied leading to 25444 audio samples.
+%
+% The recorded instruments belongs to the wind and string classes.
+% For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
+% For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
+% \ml{Better to display these in a table? VL todo table}
+%
+% Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
+% For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
+%
+% \vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
+% The abbreviations can be removed. @data @writing}
+% \ml{}
+%
+% \vl{A bubble plot of number of notes per instrument (x axis) and playing technique
+% (y axis) would be extremely useful. @data @figure}
+% \ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
+% \vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
+
+\section*{Human timbre judgments}
+\label{sec:subjective}
+
+The dataset considered in this study is taken
+from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
+It has audio recordings of 16 different musical instruments played
+with different playing techniques,
+which leads to $235$ different couples of instrument / playing technique.
+For each couple, the pitch and intonation is varied leading to $25444$ recordings.
+%The way the dataset is structured is explained in more details in Appendix \ref{sec:appendix-dataset}.
+
+Gathering the similarity judgments between each pair of samples would be a daunting task.
+Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
+
+One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
+This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
+
+\section*{Reducing Expert judgments of technique peculiarities}
+
+\ja{This needs to be summarized (possible in the previous section).
+I'm copying two sentences in here that I had in the Timbre similarity judgments section.}%
+These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
+
+Even so, considering every pairs of IPTs is not practical. A selection
+experiment is thus conducted where the subject is asked to give his opinion on which
+\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
+if it is likely to be associated to another \ipt of another instrument.
+Interest is rated on a 7 ticks scale. The subject can listen to all the different
+nuance and pitch samples of each \ipt. The experiment is over when all the
+\ipts are rated.
+
+Rating guidelines are provided :
+\begin{itemize}
+  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
+  \item Four stars: there is a proximity on one aspect of the sound between this
+  \ipt and of an \ipt of another instrument,
+  but this one is neither decisive nor obvious.
+  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
+\end{itemize}
+
+Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
 

### 2019-10-29 paper/paper.tex
address JA's comment on biological vs artificial neurons

@@
 Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
-Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-\ja{This is quite an abrupt transition.
-Why ``nevertheless''?}%
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
+Nevertheless, while training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}, thus yielding a useful form of inductive bias for downstream machine listening applications.
 The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.

### 2019-10-29 paper/paper.tex
cite theunissen

@@
 Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner--Ville distribution.
-In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain.
-\ja{``Spectrographic''?}
+In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain \cite{theunissen2000jneur}.
 This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
-Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014deep}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
+Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014tsp}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
@@
 
-A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014deep}.
+A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
 Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
@@
 One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014deep}.
+These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
 
@@
 The resulting coefficients are known as the second-order time scattering coefficients.
-These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014deep}.
+These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
 
@@
 This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014deep}.
+The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
 
@@
 The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014deep}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014deep}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014deep,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
 \ja{Need to flesh out the above with some equations.

### 2019-10-29 paper/paper.tex
cite wu2018taslp

@@
 
-\section*{Timbre similarity judgments}
-\label{sec:survey}
 The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
@@
 In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
-We refer to \cite{x} and \cite{x} for a review of recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively. % TODO VL CITE
+We refer to \cite{wu2018taslp} and \cite{x} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.
+% TODO VL CITE
 

### 2019-10-29 paper/paper.tex
cite pearce 2019

@@
 In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
-We refer to \cite{wu2018taslp} and \cite{x} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.
+We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.
 % TODO VL CITE

### 2019-10-30 paper/paper.tex
write on the accordion

@@
 We should note that this dense sampling comes at the detriment of other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
+Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.
 Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
@@
 Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
-% TODO explain that the accordion was judged too idiosyncratic altogether. because it's a free reed with keyboard activation and handheld airflow.
+It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic, regardless of playing technique, to be relevant for this experiment.
+Indeed, the accordion is the only instrument in the aforementioned list of instrument to have free reeds, keyboard-based actuation, or handheld airflow.
 

### 2019-10-31 paper/paper.tex
set product -> Cartesian product

@@
 In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
-The set product of all five variables results in a combinatorial explosion of possible timbral sensations. % TODO verify ``set product'' terminology
+The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
 The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total.

### 2019-10-31 paper/paper.tex
N=78
mezzo-forte -> mezzo forte

@@
 % TODO update with numbers, check the plural of tessitura
-Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzoforte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).
+Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzo forte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).
 However, the resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
@@
 Once both experts completed their annotation, we retained all IPTs whose average relevance score was judged equal to 3 or higher.
-At the end of this stage of pre-screening, the list of IMT reduces from 235 to 78.
+At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
 Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
@@
 To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
-The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
+The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo forte}.
 For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.

### 2019-10-31 paper/paper.tex
remotely -> independently

@@
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
-In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.
-\ja{I don't understand this last sentence.
-What does ``remotely'' mean here?}%
+In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, independently from any classification task.
 

### 2019-10-31 paper/paper.tex
fix sections

--- 2019-10-31 paper/paper.tex
@@
 In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
-The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
+The set product of all five variables results in a combinatorial explosion of possible timbral sensations. % TODO verify ``set product'' terminology
 The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total.
@@
 The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
+For more information on the dataset considered in this study, see Section \ref{sec:subjective}.
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
@@
 \caption{Full list of audio stimuli (1/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
+\label{tab:list1}
       \begin{tabular}{llll}
@@
 \caption{Full list of audio stimuli (2/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
+\label{tab:list2}
 \begin{tabular}{llll}

### 2019-10-31 paper/paper.tex
start writing on the gui

--- 2019-10-31 paper/paper.tex
@@
 Indeed, the accordion is the only instrument in the aforementioned list of instrument to have free reeds, keyboard-based actuation, or handheld airflow.
-
-% one paragraph on annotation interface
-
-Thirdly,
+%cumbersome
+
+Thirdly, we design a graphical user interface for partitioning a corpus of short audio samples.
+The need for such an interface arises from the unscalability of Likert scales in the context of pairwise similarity judgments.
+Assuming that similarity is a symmetric quantity, collecting a dense matrix of continuously valued ratings of similarity among a dataset of $N$ items would require $\frac{1}{2}(N^2-N)$ Likert scales.
+In the case of IMTs, this number would amount to about 3k horizontal sliders or sets of radio buttons.
+This number is 
 % in 30 to 60 minutes.

### 2019-10-31 paper/paper.tex
fix sections

--- 2019-10-31 paper/paper.tex
@@
 
-\ja{These section numbers aren't showing up since we're not numbering the sections (per the style guide).}
-Section \ref{sec:soa} reviews recent research in the computational modeling of musical timbre.
-Section \ref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
-Section \ref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-Section \ref{sec:results} reports the results of an ablation study of our system.
-Section \ref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+\ml{reword.}
+Section \nameref{sec:soa} reviews recent research in the computational modeling of musical timbre.
+Section \nameref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
+Section \nameref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+Section \nameref{sec:results} reports the results of an ablation study of our system.
+Section \nameref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
 

### 2019-10-31 paper/paper.tex
use verbs in outline

@@
 
-\ml{reword.}
-Section \nameref{sec:soa} reviews recent research in the computational modeling of musical timbre.
-Section \nameref{sec:survey} describes our data collection procedure for auditory similarities across playing techniques.
-Section \nameref{sec:method} presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-Section \nameref{sec:results} reports the results of an ablation study of our system.
-Section \nameref{sec:discussion} discusses the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+The \nameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
+The \nameref{sec:survey} section describes our data collection procedure for auditory similarities across playing techniques.
+The \nameref{sec:method} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+The \nameref{sec:results} section conducts an ablation study of our system.
+The \nameref{sec:discussion} section summarizes the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
 

### 2019-10-31 paper/paper.tex
lower case
address JA's comments re named section

@@
 The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
-The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in Section \ref{sec:survey}.
+The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:survey} section.
 This warping, or reweighting, of the space is performed using LMNN.
-\ja{Again, section numbers are not showing up here.}%
 

### 2019-10-31 paper/paper.tex
lower case

--- 2019-10-31 paper/paper.tex
@@
 % \def\includegraphics[#1]{}
+
+\newcommand{\lnameref}[1]{%
+\bgroup
+\let\nmu\MakeLowercase
+\nameref{#1}\egroup}
+\newcommand{\fnameref}[1]{%
+\bgroup
+\def\nmu{\let\nmu\MakeLowercase}%
+\nameref{#1}\egroup}
+
+\newcommand{\nmu}{}
 
@@
 
-The \nameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
-The \nameref{sec:survey} section describes our data collection procedure for auditory similarities across playing techniques.
-The \nameref{sec:method} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-The \nameref{sec:results} section conducts an ablation study of our system.
-The \nameref{sec:discussion} section summarizes the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
-
-\section*{Related work}
+The \lnameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
+The \lnameref{sec:survey} section describes our data collection procedure for auditory similarities across playing techniques.
+The \lnameref{sec:method} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+The \lnameref{sec:results} section conducts an ablation study of our system.
+The \lnameref{sec:discussion} section summarizes the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+
+\section*{\nmu Related work}
 \label{sec:soa}
@@
 
-\section*{Proposed approach}
+\section*{\nmu Proposed approach}
 \label{sec:method}
@@
 
-\section*{Timbre similarity judgments}
+\section*{\nmu Timbre similarity judgments}
 \label{sec:survey}
@@
 The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
+For more information on the dataset considered in this study, see Section \ref{sec:subjective}.
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
@@
 While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
-More information on the data collection process is available in  Section \ref{sec:subjective}.
+%More information on the data collection process is available in  Section \ref{sec:subjective}.
 
@@
 
-\section*{Results}
+\section*{\nmu Results}
 \label{sec:results}
@@
 
-\section*{Discussion}
+\section*{\nmu Discussion}
 \label{sec:discussion}
@@
 
-\section*{Competing interests}
+\section*{\nmu Competing interests}
   The authors declare that they have no competing interests.
 
-\section*{Author's contributions}
+\section*{\nmu Author's contributions}
     Text for this section \ldots
 
-\section*{Acknowledgements}
+\section*{\nmu Acknowledgements}
 The Flatiron Institute is a division of the Simons Foundation.
@@
 
-% \section*{Dataset}
+% \section*{\nmu Dataset}
 % \label{sec:appendix-dataset}
@@
 
-\section*{Human timbre judgments}
+\section*{\nmu Human timbre judgments}
 \label{sec:subjective}
@@
 
-\section*{Reducing Expert judgments of technique peculiarities}
+\section*{\nmu Reducing Expert judgments of technique peculiarities}
 
@@
 %%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
-\section*{Crowdsourced judgments of pairwise similarity}
+\section*{\nmu Crowdsourced judgments of pairwise similarity}
 

### 2019-10-31 paper/paper.tex
i don't think there is need for eqs in the rel work section

@@
 They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-\ja{Need to flesh out the above with some equations.
-We're no longer talking to biologists, so we don't need to be afraid.}%
+%\ja{Need to flesh out the above with some equations.
+%We're no longer talking to biologists, so we don't need to be afraid.}% i think we will put equations later, in the methods section, while making sure to re-use the notations lambda, alpha, beta. OK?
 

### 2019-10-31 paper/paper.tex
20000 -> 20k

--- 2019-10-31 paper/paper.tex
@@
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20000 polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
+For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
 Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.

### 2019-10-31 paper/paper.tex
add TODO VL's with ML after JA's comments

--- 2019-10-31 paper/paper.tex
@@
 This number is
+% TODO explain what C_k is. Equivalence class over the set of IPTs.
 % in 30 to 60 minutes.
@@
 
+% TODO VL: organize design choices.
+% (1) temporal support
+% (2) time-frequency scattering / time scattering
+% (3) learning / no learning
+
 \subsection*{Data}
@@
 For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
+%TODO VL update integer range for k.
+% Describe every k as an equivalency class
 \ja{What is $C_k$ here?
@@
 We need to be explicit about these definitions.}%
-
-\subsection*{Metric}
+% TODO VL: give more detail here. The integer a is a sample index for an IPT.
+% i is the IPT of the sound a.
+% Explain label propagation.
+
+\subsection*{Evaluation metric}
+
+% VL TODO: explain that this reflects a use case of browsing.
 
@@
 As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
+% TODO VL: ref. JASMP + DLfM
 \ja{References.}

### 2019-10-31 paper/paper.tex
update integer ranges for n and q

@@
 For a given reference clustering $D_k$, the p@5 is computed as follows.
-For each sound $s_q, 0<q<9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n<5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors.
+For each sound $s_q, 0<q\leq9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n\leq5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors.
 For this query, the precision is defined as:

### 2019-10-31 paper/paper.tex
more TODO VL's
address ja's comment

@@
 
+% TODO VL: move the following paragraph down so that we explain: (1) p@5_q, (2) p@5, (3) ap@5.
 Considering an output of the system, that is a features vector $F_n$ per sound $s_n$, the precision at rank $5$ is computed for each of the $D_k$ clusterings, and the $31$ p@5 are averaged to obtain the ap@5.
@@
 Also, we define the precisions here as proportions (between zero and one), but later, we present them as percentages (between $0\%$ and $100\%$).}%
+% TODO VL: state that the AP@5 will be reported under the form of percentages throughout the rest of the paper.
 
@@
 
+% TODO VL: make a block diagram.
 The processing pipeline is composed of two main building blocks.
@@
 
+% TODO VL: turn into active present voice.
+% We have an exhaustive benchmark on the Cartesian product of all design choices, but we report results under the form of an ablation study for the sake of conciseness.
 The complete experiments with all settings have been computed.
@@
 
-The best performing system considers the separable scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
-\ja{Separable scattering?
-Not the joint?}%
+The best performing system considers the joint time--frequency scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
 Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.5\%$ of $ap@5$.
@@
 \ja{What conclusion do we draw from this?}
+% TODO VL: draw a big-picture conclusion from this wrt the original motivation of the paper.
 

### 2019-10-31 paper/paper.tex
more TODO's

@@
 
+% TODO VL: clearly define a second part of the analysis from here onwards.
+% we test the validity of the ablation study itself.
+% we bring up alternate hypotheses and refute them one by one.
+% like i did in ISMIR 2016.
 In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol.
@@
 If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\%$, indicating a low risk of overfitting.
-\ja{We can expect better performance by using larger folds for the training (i.e., not restricting ourselves to 2-fold cross validation).}
-
-Another risk when considering reprojection techniques such as the LMNN may appear when the dimensionality of the features approaches the number of items in the dataset.
-For example, there exists a trivial solution for optimizing $p@1$ when the dimensionality of the features is equal to the number of items in the dataset.
-\ja{What is that trivial solution?}%
-In this case, no information needs to be carried by the features set to obtain perfect performance.
-
+
+%\ja{We can expect better performance by using larger folds for the training %(i.e., not restricting ourselves to 2-fold cross validation).}
+% VL: hors sujet.
+
+%Another risk when considering reprojection techniques such as the LMNN may appear when the dimensionality of the features approaches the number of items in the dataset.
+%For example, there exists a trivial solution for optimizing $p@1$ when the dimensionality of the features is equal to the number of items in the dataset.
+%\ja{What is that trivial solution?}%
+%In this case, no information needs to be carried by the features set to obtain perfect performance.
+
+% TODO VL: add an introductory sentence here. "Risk of overfitting of reprojection techniques using LMNN"
 In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.

### 2019-10-31 paper/paper.tex
more TODO's

@@
 Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
+% TODO VL: write about frequency-depedendent time shift. cite IEEE TSP
 
@@
 \ja{Explain why this is necessary.}%
-To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i*f_j, i \neq j$.
+To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
 \ja{What are the $f_i$ here?
 How are they defined?}%
-The resulting features set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
+% TODO VL: give reference to MFCCs.
+The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
 The resulting $ap@5$ of $81.8\% \pm 7$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
 \ml{vl outro on modulations}
+% TODO VL: insist on the role of modulations in playing techniques.
 

### 2019-10-31 paper/paper.tex
fix numbers in results sec

@@
 The best performing system considers the joint time--frequency scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
-Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.5\%$ of $ap@5$.
+Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.0\% \pm 0.6$ of $ap@5$.
 That means that, for a given sound dataset of \ipts for which subjective evaluation of similarity between \ipts are available, a very large number of queries would return interesting results to the user using this processing pipeline.
@@
 The impact of the variability of the perceptual judgments and its impact on machine listening systems such as the one discussed in this paper is not in the scope of this paper and left for further research.
-By considering this protocol, the proposed system reaches a $ap@5$ of $95.5\%$.
-If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\%$, indicating a low risk of overfitting.
+By considering this protocol, the proposed system reaches a $ap@5$ of $96.2\% \pm 2.0$.
+If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\% \pm 2.0$, indicating a low risk of overfitting.
 
@@
 In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-In this case, the performance drops to $23.6\%$.
-Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
+In this case, the performance drops to $23.6\% \pm 10.6$.
+%Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
 It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
@@
 One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
-When considering LDA in place of LMNN, the performance drops to $81.2\%$.
+When considering LDA in place of LMNN, the performance drops to $76.6\% \pm 10.9$.
 The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
@@
 Concerning the feature computation block $B_1$, an important factor is the time support.
-When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $92\% \pm$.
-\ja{$\pm$ what?}%
+When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $90.9\% \pm 4$.
 This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.
@@
 
-Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
+Considering the separable scattering gives a performance of $91\% \pm 4.5$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
 % TODO VL: write about frequency-depedendent time shift. cite IEEE TSP
 
-Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $82.3\% \pm 7.1$.
+Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
 Note that the MFCCs are equivalent to first-order separable scattering coefficients \cite{anden2014tsp}.
@@
 The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
-The resulting $ap@5$ of $81.8\% \pm 7$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
+The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
 \ml{vl outro on modulations}

### 2019-11-01 paper/paper.tex
cite caetano2019swarm

--- 2019-10-31 paper/paper.tex
@@
 The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
-In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range.
-% TODO add more references to automatic orchestration?
-Crestel and Esling?
+In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range \cite{caetano2019swarm}.
 

### 2019-11-11 paper/paper.tex
comment out abstract

@@
 \begin{abstract}
-The timbre of a sound is an important factor in determining its identity and source.
-One successful means of exploring the timbre space has been the construction of computational models to analyze datasets of musical instrument recordings.
-However, these datasets have typically only included a limited set of instruments played with standard playing techniques, limiting the timbral range.
-For these datasets, models based on static spectral descriptors, such as mel-frequency cepstral coefficients (MFCCs), have enjoyed excellent performance.
-However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, more sophisticated models which incorporate dynamic large-scale signal structure are needed.
-To this end, we propose a simple but rich model based on joint time--frequency scattering transforms followed by a linear mapping.
-The linear part is learned from human similarity judgments obtained in a crowdsourced free sorting task, so the resulting descriptor provides a perceptually motivated characterization of timbre.
-We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
-In addition, we perform an ablation study demonstrating the necessity of various components in this model.
+
 \end{abstract}
+
+%\begin{abstract}
+%The timbre of a sound is an important factor in determining its identity and source.
+%One successful means of exploring the timbre space has been the construction of computational models to analyze datasets of musical instrument recordings.
+%However, these datasets have typically only included a limited set of instruments played with standard playing techniques, limiting the timbral range.
+%For these datasets, models based on static spectral descriptors, such as mel-frequency cepstral coefficients (MFCCs), have enjoyed excellent performance.
+%However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, more sophisticated models which incorporate dynamic large-scale signal structure are needed.
+%To this end, we propose a simple but rich model based on joint time--frequency scattering transforms followed by a linear mapping.
+%The linear part is learned from human similarity judgments obtained in a crowdsourced free sorting task, so the resulting descriptor provides a perceptually motivated characterization of timbre.
+%We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
+%In addition, we perform an ablation study demonstrating the necessity of various components in this model.
+%\end{abstract}
 

### 2019-11-11 paper/paper.tex
listening behaviors in humans

@@
 
-Behind the overarching challenge of coming up with a robust predictive model for human listening behaviors, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
+Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
 Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.

### 2019-11-11 paper/paper.tex
avoid repetition of "relied"

@@
 Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
-These visualizations confirm the theoretical predictions of previous research \cite{anden2019tsp}, which relied on simple synthetic sounds such as exponential chirps.
+These visualizations confirm the theoretical predictions of previous research \cite{anden2019tsp}, which were founded on simple synthetic sounds such as exponential chirps.
 They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.

### 2019-11-11 paper/paper.tex
punctuation

@@
 Rather, each excerpt must be played back to multiple human listeners.
-Yet collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
+Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
 Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.

### 2019-11-11 paper/paper.tex
comment out \ja in introduction

@@
 
-\ja{This section is way too long.
-It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
-This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
-We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
-}
+%\ja{This section is way too long.
+%It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
+%This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
+%We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
+%}
 

### 2019-11-11 paper/paper.tex
change csentence in overview of the proposed model
fill in SOL stats

@@
 \begin{figure}[h!]
-\caption{\csentence{Model overview}
+\caption{\csentence{Overview of the proposed approach}
 One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
@@
 %\ja{Need to flesh out the above with some equations.
-%We're no longer talking to biologists, so we don't need to be afraid.}% i think we will put equations later, in the methods section, while making sure to re-use the notations lambda, alpha, beta. OK?
+%We're no longer talking to biologists, so we don't need to be afraid.}
 
@@
 We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.
-% TODO VL CITE
 
@@
 Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'', also known as \emph{harmon}, inside in the bell of a trumpet.
-Once augmented with mutes, the list of 16 instruments grows up to X. % TODO complete
-Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between X (for the X) and Y (for the Y). % TODO update with number.
+Once augmented with mutes, the list of 16 instruments grows up to 33.
+Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between 11 (for the accordion) and 41 (for the bass tuba).
 In the rest of this paper, we abbreviate instrument--mute--technique by means of the acronym ``IMT''.
@@
 Performers can play each IMT at various pitches, according to a tessitura that depends on the instrument but not on the choice of mute.
-Among the 16 instruments in this study, the two instruments with widest and narrowest tessituras, in their respective ordinary techniques, are the X (X semitones) and the Y (Y semitones) respectively.
-% TODO update with numbers, check the plural of tessitura
+Among the 16 instruments in this study, the two instruments with widest and narrowest tessituras, in their respective ordinary techniques, are the accordion (81 semitones) and the trumpet in C (32 semitones) respectively.
 Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzo forte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).

### 2019-11-11 paper/paper.tex
start describing cyberlioz

@@
 Indeed, the accordion is the only instrument in the aforementioned list of instrument to have free reeds, keyboard-based actuation, or handheld airflow.
-%cumbersome
 
@@
 Assuming that similarity is a symmetric quantity, collecting a dense matrix of continuously valued ratings of similarity among a dataset of $N$ items would require $\frac{1}{2}(N^2-N)$ Likert scales.
-In the case of IMTs, this number would amount to about 3k horizontal sliders or sets of radio buttons.
-This number is
+In the case of $N=78$ IMTs, the task would amount to about 3k horizontal sliders, \ie{} several hours of cumbersome work for the human annotator.
+Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of continuous ratings.
+To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all $31$ participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}.
+The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz, whose \emph{Treatise on Orchestration} (1844) marked a watershed in the history of European classical music, because of its deliberate focus on the role of timbre as a parameter for musical expression.
+Cyberlioz consists of a square panel on which is displayed a collection of circular dots, each of them corresponding to one of the $N=78$ IMTs.
+Hovering the screen pointer onto each dot results in a playback of the representative audio sample of this IMT, \ie{} C4 and \emph{mf} in most cases.
+% TODO VL: use pitch notation library to express C4
 % TODO explain what C_k is. Equivalence class over the set of IPTs.

### 2019-11-11 paper/paper.tex
write more on cyberlioz

@@
 The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz, whose \emph{Treatise on Orchestration} (1844) marked a watershed in the history of European classical music, because of its deliberate focus on the role of timbre as a parameter for musical expression.
-Cyberlioz consists of a square panel on which is displayed a collection of circular dots, each of them corresponding to one of the $N=78$ IMTs.
-Hovering the screen pointer onto each dot results in a playback of the representative audio sample of this IMT, \ie{} C4 and \emph{mf} in most cases.
-% TODO VL: use pitch notation library to express C4
+
+Cyberlioz consists of a square panel on which is displayed a collection of circular grey dots, each of them corresponding to one of the $N=78$ IMTs, and initially distributed uniformly at random.
+Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} C4 and \emph{mf} in most cases. % TODO VL: use pitch notation library to express C4
+Furthermore, each dot can be freely placed on the screen by clicking, dragging, and dropping.
+Lastly, the user can assign a color to each dot by selecting it among a palette of 20 hues.
+The goal of the Cyberlioz interface is to form clusters of timbre similarity between IMTs, expressed by sameness of color.
+In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\timesN$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
 % TODO explain what C_k is. Equivalence class over the set of IPTs.

### 2019-11-11 paper/paper.tex
update intro

@@
 On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
-On the other hand, the sole mention of a playing technique does not determine any intended effect; as a result, it fails to generalize to the entire instrumentarium \cite{kolozali2011ismir}.
+On the other hand, the sole mention of a playing technique does not determine any intended effect; as a result, it does not generalize to the entire instrumentarium \cite{kolozali2011ismir}.
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.

### 2019-11-11 paper/paper.tex
describe SOL with numbers

@@
 The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
-The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total.
-% TODO check year in which SOL was recorded, update with number
-We should note that this dense sampling comes at the detriment of other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
+The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a dense sampling of this combinatorial explosion, amounting to 25444 audio signals in total.
+We should note that this dense sampling is by no means comprehensive, as it erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
 Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.

### 2019-11-11 paper/paper.tex
write on data collection

@@
 The goal of the Cyberlioz interface is to form clusters of timbre similarity between IMTs, expressed by sameness of color.
+Here, any two audio samples are defined as similar if and only if one can be replaced by the other in a context of spectralist orchestration.
 In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\timesN$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
+
+In X 2016, we recruited volunteers to use Cyberlioz on their own computer, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
+We publicized this study on international mailing lists for research in music audio processing, such as X and Y. % TODO give mailing lists
+Within two months, 31 subjects (X men, Y women) accessed the Cyberlioz website and completed the task. % TODO break down by gender
 % TODO explain what C_k is. Equivalence class over the set of IPTs.

### 2019-11-11 paper/paper.tex
write on similarity graph

--- 2019-11-11 paper/paper.tex
@@
 The goal of the Cyberlioz interface is to form clusters of timbre similarity between IMTs, expressed by sameness of color.
+Here, any two audio samples are defined as similar if and only if one can be replaced by the other in a context of spectralist orchestration.
+
 In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\timesN$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
+In X 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
+We publicized this study on international mailing lists for research in music audio processing, such as X and Y. % TODO give mailing lists
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender?
+
+Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a similarity graph $\mathcal{G}_k$, where the integer $k$ is an arbitrary subject index between $1$ and $K$.
+For a given $k$, the graph $\mathcal{G}_k$ contains $N=78$ vertices, each representing a different IMT in the corpus.
+In $\mathcal{G}_k$, an edge connects any two vertices $i$ and $j$ if the corresponding dots in Cyberlioz have the same color.
 % TODO explain what C_k is. Equivalence class over the set of IPTs.

### 2019-11-11 paper/paper.tex
remove redundant info in appendix

@@
 
-To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
-The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo forte}.
-For more information on the dataset considered in this study, see Section \ref{sec:subjective}.
 In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
@@
 
-The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
-We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
 The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.
@@
 
-The subjects are thus asked to organize the 78 \ipts displayed as grey dots on a two dimensional plane into groups by assigning colors to each dots according to their "acoustic similarity".
-The textual instructions provided to introduce the experiment was chosen to be deliberately vague in order to allow the subject to follow its own judgment.
-
-The plane is displayed on a computer screen using a web based interface\footnote{The interface is available here: http://soundthings.org/ticel/groupSounds}.
-The \ipts can be listened to by hovering the mouse on the corresponding dot.
-The experiment is over when the subject assigned a color to all the dots.
-The initial location of the dots on the plane is randomized for each subject and can be changed during the experiment by the subject for convenience.
-Event hough the location of the dots is recorded, this data is not considered during the analysis.
-
-The experiment was available online for two months and requests to perform the test have been sent to internal mailing lists of the CNSMDP and international research mailing with a focus on audio and music processing.
-The experiment have been performed on a voluntary basis without gratification.
-Subjects were asked to use headphones at a comfortable listening level.
-
 31 subjects completed the experiment.

### 2019-11-11 paper/paper.tex
write on hypergraph partioning

@@
 
-In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\timesN$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
+In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\times N$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
 In X 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
 We publicized this study on international mailing lists for research in music audio processing, such as X and Y. % TODO give mailing lists
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender?
-
-Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a similarity graph $\mathcal{G}_k$, where the integer $k$ is an arbitrary subject index between $1$ and $K$.
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender? give age stats? musical experience?
+Figure \ref{fig:xp2display}. % TODO
+
+Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a similarity graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
 For a given $k$, the graph $\mathcal{G}_k$ contains $N=78$ vertices, each representing a different IMT in the corpus.
 In $\mathcal{G}_k$, an edge connects any two vertices $i$ and $j$ if the corresponding dots in Cyberlioz have the same color.
-% TODO explain what C_k is. Equivalence class over the set of IPTs.
-% in 30 to 60 minutes.
-
-In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
-Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
-In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
-While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
-%More information on the data collection process is available in  Section \ref{sec:subjective}.
-
-The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.
-
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/xp2example1.png}
-\caption{Spatial and group organization provided by subject 1.}
-\label{fig:xp2display}
-\end{figure}
-
-For each subject, a matrix of size $78 \times 78$ is constructed to represent the similarity judgments for each pair of the $78$ IPTs.
-Specifically, the $(i, j)$th element of the matrix was set to one if the $i$th and $j$th IPT were assigned the same color.
-If not, it is set to zero.
+Otherwise, there is no edge connecting $i$ and $j$.
+This way, $\mathcal{G}_k$ contains as many connected components as the number of similarity clusters for the subject $k$, \ie{} the number of distinct colors on the Cyberlioz interface in the response of $k$.
+
+%Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
+We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
+
+We aggregate the similarity judgments from all $K$ subjects by embedding them into a hypergraph $\mathcal{H}$, that is, a graph whose edges may connect three or more vertices at once.
+Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_k$'s among all subject $k$.
+Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
+
 In other words, the clustering assignment of each subject as a binary similarity matrix.

### 2019-11-11 paper/paper.tex
write on consensus

@@
 Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
-
-In other words, the clustering assignment of each subject as a binary similarity matrix.
-Depending on the application, it may be of interest to condition the timbre space using the information of only one subject.
-In this case, the corresponding binary similarity matrix may be used and the resulting timbre space is specialized to a given person.
-
-It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects.
-We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.
+% TODO describe the gist of hypergraph partitioning?
+While the graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants, while discarding intersubjective variability.
+Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ subjects.
+This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a newcoming subject, after less than one hour of using Cyberlioz.
 

### 2019-11-12 paper/paper.tex
start rewriting abstract

@@
 \begin{abstract}
-
+Extended playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in erudite and folk contexts.
+They affect the spectrotemporal modulations of pitched audio signals, thereby modifying timbre perception systematically.
+However, most existing approaches to music similarity retrieval inadequately reflect this effect, as they fail to describe timbre beyond the so-called ``ordinary'' technique; use instrument identity as a proxy for timbre quality; or do not allow for rapid customization to the perceptual idiosyncracies of a newcoming subject.
+%Our article fills this gap in scholarship by providing a scalable algorithm for measuring the timbre similarity between any two audio samples
 \end{abstract}
@@
 This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a newcoming subject, after less than one hour of using Cyberlioz.
+
+%TODO VL: comment on visualization of similarity matrices.
 

### 2019-12-02 paper/paper.tex
describe hypergraph partitioning

@@
 
+\subsection*{Dataset}
 We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
@@
 
+\subsection*{Disentanglement of factors}
 First, we purposefully disentangle categorical variables (IMTs) from continuous variables (pitch and intensity) in the SOL dataset.
@@
 
+\subsection*{Expert pre-screening}
 Secondly, we reduce the number of IMTs in our study by focusing on those which are deemed to be most relevant.
@@
 It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic, regardless of playing technique, to be relevant for this experiment.
-Indeed, the accordion is the only instrument in the aforementioned list of instrument to have free reeds, keyboard-based actuation, or handheld airflow.
-
+Indeed, the accordion is the only instrument in the aforementioned list of 
+instrument to have free reeds, keyboard-based actuation, or handheld airflow.
+
+\subsection*{Efficient annotation interface}
 Thirdly, we design a graphical user interface for partitioning a corpus of short audio samples.
@@
 Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender? give age stats? musical experience?
-Figure \ref{fig:xp2display}. % TODO
+%Figure \ref{fig:xp2display}. % TODO
 
@@
 % TODO describe the gist of hypergraph partitioning?
+To construct $G_0$, we select a number of clusters $C=5$ a priori, and run hypergraph partitioning to assign each vertex $i$ to one of these clusters.
+Intuitively, hypergraph partitioning optimizes a tradeoff between two objectives: first, balancing the size of all clusters in terms of their respective numbers of vertices; and secondly, keeping most hyperedges enclosed within as few distinct clusters as possible.
+
 While the graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants, while discarding intersubjective variability.

### 2019-12-02 paper/paper.tex
more subsection titles

@@
 
+\subsection*{Goal}
 This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
@@
 
+\subsection*{Contributions}
 This article gathers three novel contributions.
@@
 
+\subsection*{Outline}
 The \lnameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
@@
 
+\subsection*{Timbre modeling as music information retrieval (MIR)}
 On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
@@
 
+\subsection*{Timbre modeling as music cognition research}
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
@@
 
+\subsection*{Spectrotemporal receptive fields (STRF)}
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
@@
 
+\subsection*{Claim of originality}
 The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
@@
 
+\subsection*{Hypergraph partitioning}
 Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a similarity graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
@@
 %TODO VL: comment on visualization of similarity matrices.
+\subsection*{Inter-instrument and inter-technique similarities}
 

### 2020-01-13 paper/paper.tex
Card -> card
typo

@@
 
-\newcommand{\Card}{\mathrm{Card}}
+\newcommand{\Card}{\mathrm{card}}
 
@@
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
-In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, independently from any classification task.
+In addition, previous publications in time--frequency scattering, whether separable or joint, lack a human-centric evaluation, independently from any classification task.
 

### 2020-01-13 paper/paper.tex
define I
define m and n

@@
 instrument to have free reeds, keyboard-based actuation, or handheld airflow.
+Consequently, regardless of mute and technique, the number of instruments in our study is equal to $I=15$.
 
@@
 For a given $k$, the graph $\mathcal{G}_k$ contains $N=78$ vertices, each representing a different IMT in the corpus.
-In $\mathcal{G}_k$, an edge connects any two vertices $i$ and $j$ if the corresponding dots in Cyberlioz have the same color.
-Otherwise, there is no edge connecting $i$ and $j$.
+In $\mathcal{G}_k$, an edge connects any two vertices $m$ and $n$ if the corresponding dots in Cyberlioz have the same color.
+Otherwise, there is no edge connecting $m$ and $n$.
 This way, $\mathcal{G}_k$ contains as many connected components as the number of similarity clusters for the subject $k$, \ie{} the number of distinct colors on the Cyberlioz interface in the response of $k$.

### 2020-01-13 paper/paper.tex
write on definition of A

@@
 
-%TODO VL: comment on visualization of similarity matrices.
 \subsection*{Inter-instrument and inter-technique similarities}
 
-In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
+Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
+Specifically, we ask whether human listeners, when confronted with a diverse corpus of IMT samples, tend to cluster those samples in terms of an organological taxonomy.
+To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}$, containing $I=15$ rows and $I$ columns.
+For every instrumment--instrument pair $(i,j)$, we set the value of $\mathbf{A}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
+\[
+\mathbf{A}(i,j) = \dfrac{
+2 \Card \big\{ m, n \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+}{
+\Card \big\{n \,\vert\, \mathrm{Instrument}(n) \in \{ i, j \} \big\}
+}.
+\]
+%In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
 For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
@@
 
-The ordering of the instruments or the PTs on the vertical and horizontals axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
+%Then, we run Ward's minimum variance method to 
+%$\sigma : I \rightarrow I$ 
+%The ordering of the instruments or the PTs on the vertical and horizontal axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
 This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.

### 2020-01-14 paper/paper.tex
write on inter-technique similarity

@@
 \newcommand*{\ie}{i.e.,\@\xspace}
+\newcommand*{\vs}{vs.\@\xspace}
+
 \newcommand{\ipt}{IPT\xspace}
@@
 instrument to have free reeds, keyboard-based actuation, or handheld airflow.
-Consequently, regardless of mute and technique, the number of instruments in our study is equal to $I=15$.
+Consequently, regardless of mute and technique, the set of instruments $\mathcal{I}$ in our study contains 15 elements.
 
@@
 
-\subsection*{Inter-instrument and inter-technique similarities}
+\subsection*{Inter-instrument similarity}
 
@@
 Specifically, we ask whether human listeners, when confronted with a diverse corpus of IMT samples, tend to cluster those samples in terms of an organological taxonomy.
-To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}$, containing $I=15$ rows and $I$ columns.
-For every instrumment--instrument pair $(i,j)$, we set the value of $\mathbf{A}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
+To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
+For every instrumment--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
 \[
-\mathbf{A}(i,j) = \dfrac{
-2 \Card \big\{ m, n \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+\mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
+2\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
 }{
@@
 }.
+\label{eq:instrument-similarity}
 \]
@@
 
-%Then, we run Ward's minimum variance method to 
+Then, we run Ward's method to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
+This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$.
+Intuitively, for every instrument--instrument pair $(i, j) \in \mathcal{I}^2$, the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
+Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
+
+Interestingly, the matrix $\widetilde{\mathbf{A}}_{\mathcal{I}}$ exhibits a block diagonal structure, which roughly reflects the classical taxonomy of musical instruments: four woodwinds (BbCl, Fl, ASax, Ob) are clustered together, followed by a cluster containing one woodwind (Bn) and four brass (BBTb, Hn, TTbn, and TpC), followed by a cluster of all strings (Cb, Va, Vn, Vc, Gtr, Hp).
+Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
+
 %$\sigma : I \rightarrow I$ 
@@
 
-Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
-
-\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
+%Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
+
+%\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
 
@@
 
+\subsection*{Inter-technique similarity}
+From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
+We define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinals:
+\[
+\mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
+2\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+}{
+\Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
+},
+\]
+alike the definition of $\mathbf{A}_{\mathcal{I}}$.
+Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
+Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$ as a result of such agglomerative hierarchical clustering procedure.
+
+Contrary to inter-instrument similarity (Figure \ref{fig:consensusVsI}), inter-technique similarity does not exhibit a clear segregation of instrument families.
+For example, the first two rows in $\widetilde{\mathbf{A}}_{\mathcal{T}}$ correspond to a oboe sample (\texttt{Ob-blow-no-reed-C4}, \ie{} blowing without the reed) and a cello sample (\texttt{Vc-legno-tratto-C4-mf-1c}, \ie{} drawing the wooden part of the bow across the string) respectively, despite the fact that the oboe is a woodwind instrument whereas the cello is a string instrument.
+More generally, even though the block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less prevalent than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$, playing technique clusters appear to transcend instrument categories.
+Nevertheless, these clusters correspond to some basic attributes of qualitative timbre, such as percussive \vs sustained and soft \vs rough.
+
 \begin{figure}
@@
 \includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{ }
+\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).}
 \label{fig:consensusVsPt}

### 2020-04-04 paper/paper.tex
add fig1 (with some mockups)
write on cyberlioz

@@
 \begin{figure}[h!]
-\caption{\csentence{Overview of the proposed approach}
-One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
-scat, isomap.
-Second line : perceptual clustering.}
+\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
+\caption{\csentence{Overview of the proposed approach}}
 \label{fig:pipeline}
@@
 To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all $31$ participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}.
-The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz, whose \emph{Treatise on Orchestration} (1844) marked a watershed in the history of European classical music, because of its deliberate focus on the role of timbre as a parameter for musical expression.
+The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz.
+The choice is by no means coincidental: Berlioz is famous for having, in his \emph{Treatise on Orchestration} (1844), shed a particular focus on the role of timbre as a parameter for musical expression.
 

### 2020-04-04 paper/paper.tex
rephrase paragraph on computer-assisted orchestration

@@
 
-Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted instrumentation \cite{maresz2013cmr}.
-Following the aesthetic tradition of spectralism in contemporary music creation, the composer queries the machine with an arbitrary audio signal.
-The outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
-In this context, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range \cite{caetano2019swarm}.
+Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted orchestration \cite{maresz2013cmr}.
+In such a software environment, the composer queries the machine with an arbitrary audio signal; the outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
+Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range \cite{caetano2019swarm}.
 

### 2020-04-04 paper/paper.tex
todo caption
write on Figure 4

@@
 This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
+
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
@@
 \includegraphics[width = \textwidth]{figures/consensusVsI.png}
-\caption{}
+\caption{
+%TODO caption
+}
 \label{fig:consensusVsI}
@@
 For example, the first two rows in $\widetilde{\mathbf{A}}_{\mathcal{T}}$ correspond to a oboe sample (\texttt{Ob-blow-no-reed-C4}, \ie{} blowing without the reed) and a cello sample (\texttt{Vc-legno-tratto-C4-mf-1c}, \ie{} drawing the wooden part of the bow across the string) respectively, despite the fact that the oboe is a woodwind instrument whereas the cello is a string instrument.
+
 More generally, even though the block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less prevalent than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$, playing technique clusters appear to transcend instrument categories.
-Nevertheless, these clusters correspond to some basic attributes of qualitative timbre, such as percussive \vs sustained and soft \vs rough.
+Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
+Interestingly, the ordinary technique (\textt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
+This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{}, an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
 
@@
 \includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).}
+\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).
+%TODO add a double arrow
+% rough <-- soft --> rough
+% sustained <-- "ordinary" --> percussive
+}
 \label{fig:consensusVsPt}

### 2020-04-04 paper/paper.tex
move up figure 1
finish reworking related work section
move methods section after data section

@@
 
+Figure \ref{fig:pipeline} summarizes our experimental protocol: it illustrates how visual annotations (top) can inform feature extraction (center) to produce a nearest-neighbor search engine which is consistent with timbre similarity judgments (bottom).
+
 \subsection*{Contributions}
@@
 
+\begin{figure}[h!]
+\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
+\caption{\csentence{Overview of the proposed approach}}
+\label{fig:pipeline}
+\end{figure}
+\clearpage
+
 \subsection*{Outline}
@@
 Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
+
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
@@
 Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
+
 Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
@@
 This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}, thus yielding a useful form of inductive bias for downstream machine listening applications.
+
 The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
@@
 This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
+
 Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
@@
 
-%\ja{This section is way too long.
-%It's important that we situate ourselves with respect to the existing literature, but we don't need to spend 2+ pages doing so.
-%This currently reads more like a laundry list of various work related to timbre perception on musical instruments on one hand and a history of STRFs on the other.
-%We need to keep in mind what our central themes (wider range of timbre, similarity-based evaluation, simple scattering+linear model) are and link the existing work to them in a concise manner.
-%}
-
 % enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
 %We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
-
-\section*{\nmu Proposed approach}
-\label{sec:method}
-
-\begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
-\caption{\csentence{Overview of the proposed approach}}
-\label{fig:pipeline}
-\end{figure}
-
-\begin{figure}[h!]
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.}
-\end{figure}
-
-Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
-The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
-The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:survey} section.
-This warping, or reweighting, of the space is performed using LMNN.
-
-\subsection*{Scattering transform}
-\label{sec:scattering}
-
-% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
-
-While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
-As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
-
-A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
-Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
-The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
-This representation contains a large amount of information, but not in a form suitable for building statistical models.
-In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
-It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
-One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
-
-Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
-One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
-Again, we compute the modulus of the wavelet coefficients and average in time.
-The resulting coefficients are known as the second-order time scattering coefficients.
-These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
-
-While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
-Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
-If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
-We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
-This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
-
-This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
-To remedy this, another approach was suggested by And\'{e}n et al.
-\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
-Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
-
-The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
-On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
-The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-%\ja{Need to flesh out the above with some equations.
-%We're no longer talking to biologists, so we don't need to be afraid.}
-
-\subsection*{Similarity weighting}
-\label{sec:weighting}
-
-The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
-We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
-
-Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
-More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
-
-Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
-We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
-
-We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
-For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
-This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
 
@@
 Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments.
+
 Once both experts completed their annotation, we retained all IPTs whose average relevance score was judged equal to 3 or higher.
@@
 \]
-%In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}.
-For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments.
-The PTs similarity matrix is computed in the same way.
 
@@
 Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
-
-%$\sigma : I \rightarrow I$ 
-%The ordering of the instruments or the PTs on the vertical and horizontal axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection.
-This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.
-
-%Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.
-
-%\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}
 
@@
 Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
-Interestingly, the ordinary technique (\textt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
+Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
 This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{}, an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
@@
 
-\section*{\nmu Results}
+\section*{\nmu Proposed approach}
+\label{sec:method}
+
+\begin{figure}[h!]
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.}
+\end{figure}
+
+%Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
+%The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
+%The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:survey} section.
+%This warping, or reweighting, of the space is performed using LMNN.
+
+\subsection*{Scattering transform}
+\label{sec:scattering}
+
+% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
+
+While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
+As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
+
+A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
+Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
+The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
+This representation contains a large amount of information, but not in a form suitable for building statistical models.
+In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
+It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
+One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
+These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
+
+Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
+One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
+Again, we compute the modulus of the wavelet coefficients and average in time.
+The resulting coefficients are known as the second-order time scattering coefficients.
+These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
+
+While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
+Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
+If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
+We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
+This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
+The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
+
+This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
+To remedy this, another approach was suggested by And\'{e}n et al.
+\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
+Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
+As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
+
+The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
+On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
+The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
+Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+%\ja{Need to flesh out the above with some equations.
+%We're no longer talking to biologists, so we don't need to be afraid.}
+
+\subsection*{Similarity weighting}
+\label{sec:weighting}
+
+The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
+We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
+
+Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
+More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
+
+Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
+We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
+
+We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
+For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
+This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
+
+\section*{\nmu Machine listening experiments}
 \label{sec:results}

### 2020-04-04 paper/paper.tex
introduce methods section

@@
 \label{sec:method}
+The previous section described our protocol for collecting timbral similarity judgments between instrumental playing techniques at scale.
+In this section, we aim to recover these similarity judgments from digital audio recording, according to a paradigm of supervised metric learning.
+To this end, we present a new machine listening system, composing joint time--frequency scattering and large-margin nearest neighbors (LMNN).
 

### 2020-04-04 paper/paper.tex
provide feedback on methods section
renamed method subsections

@@
 
-\begin{figure}[h!]
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.}
-\end{figure}
-
 %Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
@@
 
-\subsection*{Scattering transform}
+\subsection*{Joint time--frequency scattering transform}
 \label{sec:scattering}
-
-% cite Lostanlen PhD dissertation. classification results on Medley-solos-DB. suggests that JTFS is competitive wrt deep learning techniques even in the large data regime (500-5000 samples per class).
 
@@
 %We're no longer talking to biologists, so we don't need to be afraid.}
-
-\subsection*{Similarity weighting}
+% TODO: avoid citing the same papers over and over
+
+\begin{figure}
+\caption{\csentence{Sample scattering}
+Same 3 sounds spec + scatt.
+% TODO figure + caption
+}
+\end{figure}
+
+\subsection*{Metric learning with large-margin nearest neighbors (LMNN)}
 \label{sec:weighting}
@@
 This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-
-\section*{\nmu Machine listening experiments}
+% TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
+
+\section*{\nmu Results}
 \label{sec:results}
 
-The processing pipeline presented in Figure \ref{fig:pipeline} addresses the problem of conditioning a timbre space using perceptual judgments of timbral similarity.
-The relevance of the many design choices that had to be made are addressed in this section using an experimental protocol that considers a dataset of audio that will be given as input to the processing pipeline and one metric used for evaluating its performance under a given experimental setting.
-\ja{This last sentence is quite vague.
-What are we trying to say?}%
-
+This sec
 % TODO VL: organize design choices.
@@
 
-\subsection*{Data}
+\subsection*{Semi-supervised label propagation}
 

### 2020-04-04 paper/paper.tex
introduce results section

@@
 
-This sec
+The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
+We now turn to apply these methods to the problem of browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
 % TODO VL: organize design choices.

### 2020-04-04 paper/paper.tex
bis

@@
 The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
-We now turn to apply these methods to the problem of browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
+We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of a timbre similarity graph $\mathcal{G}_0$.
+In practice, for training and evaluation purposes, the graph $\mathcal{G}_0$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \ref{sec:survey} section.
+% browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
+
 % TODO VL: organize design choices.

### 2020-04-04 paper/paper.tex
todo cite wang 2020 icassp
start writing on semi-supervised label propagation

@@
 Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014tsp}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
+% TODO cite wang 2020 icassp
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
@@
 
-The perceptual data gathered during the perceptual experiment considers 78 \ipts, each represented by one sound recording that the subject could listen to before clustering the \ipts.
-For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
+In music information retrieval, the core motivation behind timbre similarity modeling lies in the prospect of navigating an online corpus of digital music.
+The typical size of such a corpus ranges between $N\sim10^3$ and $N\sim10^7$, depending on the use case.
+As of today, the two largest databases of isolated musical notes with extended playing techniques are FullSOL and conTimbre, with $N\sim1.2\cdot10^5$ and $N\sim8.7\cdot10^4$ samples respectively.% TODO cite FullSOL and conTimbre
+In comparison, the instrument--mute--technique (IMT) triplets in the present study on human subjects amount to merely $N=78$ samples, \ie{} three orders of magnitude less than conTimbre.
+%even though there is expert pre-screening, etc.
+% does not reflect the typical scale at which these systems are deployed.
+
+%The perceptual data gathered during the perceptual experiment considers 78 \ipts, each represented by one sound recording that the subject could listen to before clustering the \ipts.
+%For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
 %TODO VL update integer range for k.
 % Describe every k as an equivalency class
-\ja{What is $C_k$ here?
-Also, should it be $0 < k \le 31$?
-Might be more natural to say that $k = 1, 2, \ldots, 31$, in that case.}%
-If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
-In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings.
-Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.
+%\ja{What is $C_k$ here?
+%Also, should it be $0 < k \le 31$?
+%Might be more natural to say that $k = 1, 2, \ldots, 31$, in that case.}%
+%If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
+%In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings.
+%Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.
 

### 2020-05-11 paper/paper.tex
edit first sentence
edit subtitles in RW section

@@
 
-Music information retrieval (MIR) operates at either of two levels: symbolic and auditory \cite{downie2003mir}.
+Music information retrieval (MIR) operates at two levels: symbolic and auditory \cite{downie2003mir}.
 By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
@@
 
-\subsection*{Timbre modeling as music information retrieval (MIR)}
+\subsection*{Timbre modeling in music information retrieval (MIR)}
 On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
@@
 
-\subsection*{Timbre modeling as music cognition research}
+\subsection*{Timbre modeling in music cognition research}
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
@@
 
-The standard Euclidean distance on the raw scattering coefficients provides a universal representation of timbre dissimilarity.
+The standard Euclidean distance on the raw scattering coefficients provides an instrument-agnostic representation of timbre dissimilarity.
 We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
@@
 
-In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
-Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation.
+%In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
+%Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation.
 reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
-Quantification of this invariance is left for further study.
+%Quantification of this invariance is left for further study.
 

### 2020-05-11 paper/paper.tex
typo

@@
 To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrumment--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
+For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
 \[

### 2020-05-11 paper/paper.tex
lnameref

@@
 We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of a timbre similarity graph $\mathcal{G}_0$.
-In practice, for training and evaluation purposes, the graph $\mathcal{G}_0$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \ref{sec:survey} section.
+In practice, for training and evaluation purposes, the graph $\mathcal{G}_0$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section.
 % browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.

### 2020-05-11 paper/paper.tex
write more on semi-supervised label propagation

@@
 
-In music information retrieval, the core motivation behind timbre similarity modeling lies in the prospect of navigating an online corpus of digital music.
-The typical size of such a corpus ranges between $N\sim10^3$ and $N\sim10^7$, depending on the use case.
-As of today, the two largest databases of isolated musical notes with extended playing techniques are FullSOL and conTimbre, with $N\sim1.2\cdot10^5$ and $N\sim8.7\cdot10^4$ samples respectively.% TODO cite FullSOL and conTimbre
-In comparison, the instrument--mute--technique (IMT) triplets in the present study on human subjects amount to merely $N=78$ samples, \ie{} three orders of magnitude less than conTimbre.
-%even though there is expert pre-screening, etc.
-% does not reflect the typical scale at which these systems are deployed.
-
-%The perceptual data gathered during the perceptual experiment considers 78 \ipts, each represented by one sound recording that the subject could listen to before clustering the \ipts.
-%For those 78 \ipts, 31 clusterings $C_k, 0<k<31, \Card(C_k)=78$ are available, that are our reference in terms of timbre similarity.
-%TODO VL update integer range for k.
-% Describe every k as an equivalency class
-%\ja{What is $C_k$ here?
-%Also, should it be $0 < k \le 31$?
-%Might be more natural to say that $k = 1, 2, \ldots, 31$, in that case.}%
-%If \ipt $i$ and \ipt $j$ are clustered together by subject $k$, then $C_k(i)=C_k(j)$.
-%In essence, an effective ranking system should retrieve sounds that belong to \ipts that are clustered together in each of the 31 clusterings.
-%Depending on the preferences of subjects, there may be competition between clusterings that should be mitigated when appreciating quantitatively the ranking performance.
-
-%In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects.
-%Assuming invariance to timbre with respect to pitch and nuance is in line with the ANSI definition of timbre \ml{vl, citation.
-reformuler tout le paragraphe}, it is thus taken in this work as a reasonable assumption.
-%Quantification of this invariance is left for further study.
-
-The whole range of nuance and pitch is thus considered for each of the 78 \ipts, leading to a dataset composed of 9346 sound recordings.
-The clusterings $C_k$ are propagated to this dataset to produce new reference clusterings $D_k, 0<k<31, \Card(D_k)=9346$ using the following scheme.
-If a sound of this dataset $s_a$ corresponds to \ipt $i$, then $D_k(a) = C_k(i), \forall k$.
-\ja{What's $s_a$ and $a$ here?
-We need to be explicit about these definitions.}%
-% TODO VL: give more detail here. The integer a is a sample index for an IPT.
-% i is the IPT of the sound a.
-% Explain label propagation.
+In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must withdraw from a diverge realm of instruments and techniques.
+Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.
+
+Given an audio stimulus $n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
+Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
+We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
+This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL (roughly one third of the entire dataset, i.e., $25444$ samples) for which the notion of timbre is perceptually encoded as a discrete category.
+In the case of the consensus cluster graph ($\mathcal{G}_0$), this discrete category takes one of $C=5$ distinct values.
 

### 2020-05-11 paper/paper.tex
typo
write more on label propagation

@@
 
-Once both experts completed their annotation, we retained all IPTs whose average relevance score was judged equal to 3 or higher.
+Once both experts completed their annotation, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
 At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
@@
 In the case of the consensus cluster graph ($\mathcal{G}_0$), this discrete category takes one of $C=5$ distinct values.
+However, for a particular subject $k$, this discrete category takes one of $C_k$ distinct values, where the number of timbre clusters varies between $3$ and $19$ depending on $k$.
 

### 2020-05-12 paper/paper.tex
start reworking the evaluation metric subsection

@@
 \usepackage{amsthm,amsmath}
+\usepackage{dsfont}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
+\usepackage{relsize}
 % \def\includegraphic[#1]{}
@@
 The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
-We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of a timbre similarity graph $\mathcal{G}_0$.
-In practice, for training and evaluation purposes, the graph $\mathcal{G}_0$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section.
+We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
+In practice, for training and evaluation purposes, the cluster graph $\mathcal{G}$ may stem from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
+Alternatively, it may correspond to the subjective preferences of a single user $k$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
 % browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
@@
 
-Given an audio stimulus $n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
+Given an audio stimulus $n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}$.
 Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
@@
 % VL TODO: explain that this reflects a use case of browsing.
-
-The metric considered in this work is the average precision at rank $k=5$ (ap@5).
-It allows us to evaluate if the outputs of the proposed system are likely to satisfy the user.
-The ap@5 is computed as follows.
-
-% TODO VL: move the following paragraph down so that we explain: (1) p@5_q, (2) p@5, (3) ap@5.
-Considering an output of the system, that is a features vector $F_n$ per sound $s_n$, the precision at rank $5$ is computed for each of the $D_k$ clusterings, and the $31$ p@5 are averaged to obtain the ap@5.
-
-For a given reference clustering $D_k$, the p@5 is computed as follows.
-For each sound $s_q, 0<q\leq9346$ considered as a query, its 5 nearest neighbors $s_n, 0<n\leq5$ in the timbre space under evaluation are retrieved using the Euclidean distance between corresponding features vectors.
-For this query, the precision is defined as:
-$$
-p@5_q = \frac{1}{5} \sum_{n=1}^5 \mathbf{1}_{D_q=D_n}
-$$
-The p@5 is the average $p@5_q$ computed over all queries.
-\ja{This is perhaps better described in reverse, starting with $p@5_q$, then $p@5$, then $ap@5$?
-Also, we define the precisions here as proportions (between zero and one), but later, we present them as percentages (between $0\%$ and $100\%$).}%
-% TODO VL: state that the AP@5 will be reported under the form of percentages throughout the rest of the paper.
-
-\ml{note pour vincent, on a pas de fold, mais on a les resultats par sujet, on peut donc evaluer la dispersion de la p@5 par sujets si cela est utile.}
-
-The choice of the rank is chosen with the application in mind.
-When searching for sounds that are similar to a query, the listener is likely to listen to up to $5$ proposals, rarely more.
-For the same reason, more global metrics commonly considered for ranking tasks such as the mean average precision are not considered here.
+% let n be a query. (1<= n <= N').
+% R: rank. number of neighbors. R=5. Typical for a user application. (display, attention span ...)
+% K: number of subjects.
+
+\begin{equation}
+p_{\mathbf{\Phi}}(n, k, R) =
+    \dfrac{1}{R}
+    \sum_{r=1}^{R}
+    \mathlarger{\mathds{1}}
+    \lbrace
+        \mathbf{\Phi}(\boldsymbol{x}_n, r)
+        \in
+        \mathcal{G}_k (n)
+    \rbrace
+\end{equation}
+
+\begin{equation}
+\mathrm{P}_{\mathbf{\Phi}}(k, R) =
+\dfrac{1}{N^{\prime}}
+\sum_{n=1}^{N^{\prime}}
+p_{\mathbf{\Phi}}(n, k, R)
+\end{equation}
+
+\begin{equation}
+\mathrm{AP}_{\mathbf{\Phi}}(R) =
+\dfrac{1}{K}
+\sum_{k=1}^{K}
+\mathrm{P}_{\mathbf{\Phi}}(k, R)
+\end{equation}
 

### 2020-05-12 paper/paper.tex
write more on evaluation metrics

@@
 
+Let us denote by $\boldsymbol{x}_1 \ldots \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}=9346$ audio samples associated to our annotated dataset, after semi-supervised label propagation (see subsection above).
+Given a sample $n\leq N^{\prime}$ and a human subject $k\leq K$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.
+Our machine listening system takes the waveform $\boldsymbol{x}_{n}$ as input and returns a ranked list of nearest neighbors: $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_2 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_3 (\boldsymbol{x}_n)$, and so forth.
+
+In the context of browsing an audio collection by timbre similarity, $\boldsymbol{x}_n$ is a user-defined query while the function $\mathbf{\Phi}$ plays the role of a search engine.
+We consider that the first retrieved sample, $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, to be relevant to user $k$ if and only if it belongs to the same cluster as $\boldsymbol{x}_n$ in the cluster graph $\mathcal{G}_k$; hence the Boolean condition $\mathbf{\Phi}_{1}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
+Likewise, the second retrieved sample is relevant if and only if $\mathbf{\Phi}_{2}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
+To evaluate $\mathbf{\Phi}$ on the query $\boldsymbol{x}_n$, we measure the relevance of all nearest neighbors $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ up to some fixed rank $R$, and average the result:
 \begin{equation}
@@
     \mathlarger{\mathds{1}}
-    \lbrace
-        \mathbf{\Phi}(\boldsymbol{x}_n, r)
+    \big(
+        \mathbf{\Phi}_r(\boldsymbol{x}_n)
         \in
         \mathcal{G}_k (n)
-    \rbrace
+    \big).
 \end{equation}
-
+In the equation above, the function $\mathds{1}$ converts Booleans to integers, i.e., $\mathds{1}(b)$ equals $1$ if $b$ is true and return $0$ if $b$ is false.
+Thus, the function $p_{\mathbf{\Phi}}$ takes fractional values between $0$ and $1$, which are typically expressed in percentage points.
+
+The precision at rank $R$ of the system $\mathbf{\Phi}$ is defined as the average value taken by the function $p_{\mathbf{\Phi}}$, for constant $R$, over the entire corpus of $N^{\prime}=9346$ audio samples:
 \begin{equation}
@@
 
+Lastly, the ``average precision at rank $R$'' (henceforth, AP@$R$) is the average value of $\mathrm{P}_{\mathbf{\Phi}}$, for constant $R$, across all $K=31$ subjects from our perceptual study:
 \begin{equation}
@@
 \end{equation}
+In the rest of this paper, we set the constant $R$ to $5$.
+This is in accordance with the protocol of \cite{lostanlen2018extended}, who trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
 

### 2020-05-12 paper/paper.tex
finish writing on evaluation

@@
 \end{equation}
+It appears from the above that an effective system $\mathbf{\Phi}$ should retrieve sounds whose IMT triplets are similar according to all of the $K=31$ cluster graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$.
+
 In the rest of this paper, we set the constant $R$ to $5$.
 This is in accordance with the protocol of \cite{lostanlen2018extended}, who trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
+Moreover, note that the value $R=5$ typically reflects the number of virtual panels simultaneously on display in a software environment for computer-assisted orchestration.
 

### 2020-05-12 paper/paper.tex
formulate questions in the ablation study

@@
 \usepackage{relsize}
+\usepackage{siunitx}
 % \def\includegraphic[#1]{}
@@
 
+$\mathcal{Y}_1 (\boldsymbol{x}) = \mathcal{Y} \backslash \lbrace \boldsymbol{x} \rbrace$.
+
+\begin{equation}
+\mathbf{\Phi}_1 (\boldsymbol{x}) =
+\arg \min_{\boldsymbol{y}\in \mathcal{Y}_1(\boldsymbol{x})}
+\;
+\big\Vert
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
+-
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+\big\Vert
+\end{equation}
+
+$\mathcal{Y}_{r} (\boldsymbol{x}) = \mathcal{Y}_{(r-1)}(\boldsymbol{x}) \backslash \lbrace \mathbf{\Phi}_r (\boldsymbol{x}) \rbrace$.
+
+\begin{equation}
+\mathbf{\Phi}_r (\boldsymbol{x}) =
+\arg \min_{\boldsymbol{y}\in \mathcal{Y}_r(\boldsymbol{x})}
+\;
+\big\Vert
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
+-
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+\big\Vert
+\end{equation}
+
 \section*{\nmu Results}
@@
 \subsection*{Evaluation metric}
-
-% VL TODO: explain that this reflects a use case of browsing.
-% let n be a query. (1<= n <= N').
-% R: rank. number of neighbors. R=5. Typical for a user application. (display, attention span ...)
-% K: number of subjects.
 
@@
 
-\subsection*{Analysis}
+\subsection*{Best performing system}
+Our topline system comprises four computational blocks:
+\begin{enumerate}
+\item Joint time--frequency scattering up to a maximal time scale of $T=$\SI{1000}{\milli\second},
+\item Temporal averaging at the scale of the whole musical note,
+\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended}, and
+\item Nearest neighbor search according to priorly learned non-Euclidean metric.
+\end{enumerate}
+
+\section*{Ablation study}
+\subsection*{Is the dataset large enough for evaluation?}
+
+\subsection*{Is metric learning necessary?}
+
+\subsection*{How important is the amount of temporal context $T$?}
+
+\subsection*{Is joint time--frequency scattering necessary?}
+
+\subsection*{How important is feature extraction vs. metric learning?}
+
+\section*{Analysis}
 

### 2020-05-13 paper/paper.tex
write more on the best performing system

@@
 
-\section*{\nmu Proposed approach}
-\label{sec:method}
+\section*{\nmu Methods}
+\label{sec:methods}
 The previous section described our protocol for collecting timbral similarity judgments between instrumental playing techniques at scale.
@@
 % TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
+% cite Musgrave et al. Metric learning reality check
 
@@
 We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training and evaluation purposes, the cluster graph $\mathcal{G}$ may stem from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
-Alternatively, it may correspond to the subjective preferences of a single user $k$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
+In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
+However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
 % browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
@@
 \subsection*{Best performing system}
-Our topline system comprises four computational blocks:
+Our topline system comprises five computational blocks:
 \begin{enumerate}
@@
 \item Temporal averaging at the scale of the whole musical note,
-\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended}, and
+\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended},
+\item Per-feature affine scaling so that the dataset has null mean and unit variance,
 \item Nearest neighbor search according to priorly learned non-Euclidean metric.
 \end{enumerate}
+Note that the non-Euclidean metric is learned via large-margin nearest neighbors (LMNN, see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
+Therefore, the system $\mathbf{\Phi}$ performs timbre similarity retrieval in a user-agnostic way, and can serve as a convenient default algorithm for newcoming users.
+That being said, it is conceivable to replicate the five-stage protocol above on the cluster graph $\mathcal{G}_k$ of a specific user $k$, instead of the cluster graph $\mathcal{G}_0$.
+This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity, in comparison with the default setting $\mathcal{G}_0$.
+
+For composers to personalize their search engine $\mathbf{\Phi}$, the quid pro quo is to define manually a cluster graph among $N=78$ audio samples, by means of the Cyberlioz annotation interface (see \lnameref{sec:survey} section).
+This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
 
@@
 
-% TODO VL: make a block diagram.
 The processing pipeline is composed of two main building blocks.

### 2020-05-13 paper/paper.tex
announce result
finish results section

@@
 \item Per-feature affine scaling so that the dataset has null mean and unit variance,
-\item Nearest neighbor search according to priorly learned non-Euclidean metric.
+\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
 \end{enumerate}
@@
 This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
+
+Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ subjects of the order of $1\%$.
+This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
+In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bambooo flute (\emph{dizi}).
+However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
+In particular, it does not assume the ``metadata'' (instrument, mute, technique, pitch, dynamics, and so forth) of any musical sample $\boldsymbol{x}_n$ to be observable, in part or in full, at training time.
 

### 2020-05-13 paper/paper.tex
introduce ablation study section

@@
 \section*{Ablation study}
+The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques ($N^{\prime}=9346$ isolated notes), as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
+Over a cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $99.0\% \pm 1$.
+We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
+
 \subsection*{Is the dataset large enough for evaluation?}

### 2020-05-13 paper/paper.tex
check some to-todo's, move comments around

@@
 
+\ml{VL. provide parameters for scattering
+\texttt{
+case 'scat'
+    filt\_opt.Q = [12 1];
+    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
+case 'tfscat'
+    tm\_filt\_opt.Q = [12 1];
+    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
+    fr\_filt\_opt.J = 4;}}
+
 \begin{figure}
@@
 
-The processing pipeline is composed of two main building blocks.
-The first block $B_1$ takes a digital audio recording as input and outputs one vector of features.
-For this block, the main factors are the time support $B_1^t$ for computing the features and the type of features $B_1^f$.
-We considered $B_1^t \in \{ 25, 128, 250, 500, 1000 \}$ ms and $B_1^f \in \{ \text{MFCCs}, \text{separable scattering}, \text{joint scattering} \}$.
-The MFCCs are the standard mel-frequency cepstral coefficients with $40$ coefficients out of which we select the first $13$.
-\ja{I'm not sure I like this notation.
-First, why do we need to label the blocks ($B_1$ and $B_2$)?
-We only have two of them.
-Isn't it more straightforward to just call them ``the unsupervised part'' and ``the learned part'' or something like that?
-We can also replace $B_1^t$ with something more straightforward, like $T$ (otherwise, it's a lot of bits to encode very little information).
-Likewise, instead of $B_1^f$, we can just have $F$.}
-
-\ml{VL. provide parameters for scattering
-\texttt{
-case 'scat'
-    filt\_opt.Q = [12 1];
-    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
-case 'tfscat'
-    tm\_filt\_opt.Q = [12 1];
-    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
-    fr\_filt\_opt.J = 4;}}
-
 As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
@@
 \ja{What does this mean?}%
-
-The best performing system considers the joint time--frequency scattering computed over $1000~\mathrm{ms}$ time support, projected using the LMNN algorithm.
-Evaluated over the entire dataset, that is the computation of the projection and the evaluation are performed on the whole dataset, this system reaches the satifactory performance of $99.0\% \pm 0.6$ of $ap@5$.
-That means that, for a given sound dataset of \ipts for which subjective evaluation of similarity between \ipts are available, a very large number of queries would return interesting results to the user using this processing pipeline.
-\ja{What conclusion do we draw from this?}
-% TODO VL: draw a big-picture conclusion from this wrt the original motivation of the paper.
 

### 2020-05-13 paper/paper.tex
bis

@@
 \section*{Analysis}
-
-As found in many studies, log compression and median processing are beneficial post processing steps for the scattering.
-% TODO VL: ref. JASMP + DLfM
-\ja{References.}
-
-For each recording, a set of features are computed for each time frame.
-This set is averaged over time to give one feature vector per recording.
-All features are standardized with zero mean and unit variance across the dataset.
-\ja{Is this the ``per-feature Gaussianization'' we mention in the introduction?
-Should give the reference in that case (and maybe explain a bit more in detail).}%
-
-The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output.
-For this block, the main factor is the type of projection $B_2^p \in \{ \text{LDA}, \text{LMNN} \}$.
-The LDA is the linear discriminant analysis \cite{} and the LMNN is the algorithm presented in Section \ref{sec:weighting}.
-For this algorithm the neighborhood is set to $5$.
-\ja{Again, I would prefer to ditch this $B_2$ notation.}
-
-% TODO VL: turn into active present voice.
-% We have an exhaustive benchmark on the Cartesian product of all design choices, but we report results under the form of an ablation study for the sake of conciseness.
-The complete experiments with all settings have been computed.
-For the sake of clarity, only part of the results are discussed here.
-\ja{What does this mean?}%
 

### 2020-05-13 paper/paper.tex
reorganize paragraphs for ablation study

@@
 
-\subsection*{Is the dataset large enough for evaluation?}
-
-\subsection*{Is metric learning necessary?}
-
-\subsection*{How important is the amount of temporal context $T$?}
-
-\subsection*{Is joint time--frequency scattering necessary?}
-
-\subsection*{How important is feature extraction vs. metric learning?}
-
-\section*{Analysis}
-
-% TODO VL: clearly define a second part of the analysis from here onwards.
-% we test the validity of the ablation study itself.
-% we bring up alternate hypotheses and refute them one by one.
-% like i did in ISMIR 2016.
-In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol.
-In order to reduce the risk of misleading performance results using learning systems due to overfitting, we identify factors of variability that could lead to such problems and mitigate those.
-\ja{Is this part of the ablation study or is this a ``meta'' consideration?}
-In our case, two factors can readily be identified, the variability of the audio data and that of the perceptual judgments.
+\subsection*{Is the dataset large enough for evaluation? Is the system overfitting?}
 
@@
 
-%\ja{We can expect better performance by using larger folds for the training %(i.e., not restricting ourselves to 2-fold cross validation).}
-% VL: hors sujet.
+\subsection*{Is metric learning necessary?}
+
+The LMNN algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
+One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
+When considering LDA in place of LMNN, the performance drops to $76.6\% \pm 10.9$.
+The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
+
+Including a learned block implies some constraints.
+In particular, learning the linear transform on large datasets requires more computational time.
+Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
+It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
+In this case, the performance reduces to $92.9\% \pm 3$.
+\ja{Why the $\pm$ here and not elsewhere?}%
 
@@
 
-The LMNN algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
-One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
-When considering LDA in place of LMNN, the performance drops to $76.6\% \pm 10.9$.
-The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
-
-Including a learned block implies some constraints.
-In particular, learning the linear transform on large datasets requires more computational time.
-Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
-It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
-In this case, the performance reduces to $92.9\% \pm 3$.
-\ja{Why the $\pm$ here and not elsewhere?}%
+\subsection*{How important is the amount of temporal context $T$?}
 
@@
 \ja{Other references?}%
+
+\subsection*{Is joint time--frequency scattering necessary?}
 

### 2020-05-13 paper/paper.tex
write on qbe, cite cella, carpentier, cartwright, lemaitre

@@
 In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must withdraw from a diverge realm of instruments and techniques.
-Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.
+Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity level (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.
 
@@
 \section*{Ablation study}
-The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques ($N^{\prime}=9346$ isolated notes), as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
+The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
 Over a cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $99.0\% \pm 1$.
@@
 \subsection*{Is the dataset large enough for evaluation? Is the system overfitting?}
-
-We mitigate the first by considering a random split of the dataset in two halves, one for learning the projection and one for evaluating the performance.
-The impact of the variability of the perceptual judgments and its impact on machine listening systems such as the one discussed in this paper is not in the scope of this paper and left for further research.
-By considering this protocol, the proposed system reaches a $ap@5$ of $96.2\% \pm 2.0$.
-If considering the same first half both for learning and evaluation leads to a $ap@5$ of $96.5\% \pm 2.0$, indicating a low risk of overfitting.
+First of all, it is worth noting that the presented P@5 figure of $99.0\% \pm 1$ does not abide to a conventional training set vs.
+test set paradigm, as is most often done in machine learning research.
+Rather, the LMNN algorithm is trained on all available samples ($N^{\prime}=9346$ isolated notes) with the ``consensus'' cluster graph as training objective ($\mathcal{G}=\mathcal{G}_0$).
+Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k>0$).
+The reason behind this choice is that, in the context of computer-assisted orchestration, the collection of audio samples has a fixed size, as it is shipped alongside the software itself.
+This is the case, for example, of Orchidea\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}}, which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
+
+Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:survey} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
+That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
+One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
+However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
+Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.
+
+Despite this caveat, we may adopt a ``query-by-example'' (QbE) paradigm by  partitioning the database and audio samples in half ($\frac{1}{2}N^{\prime}=4673$), training the LMNN algorithm on the first half, and querying it with samples from the other half.
+In this evaluation framework, our system reaches a P@5 of $96.2\% \pm 2$.
+Interestingly, querying the system with samples from the training set leads to a P@5 of $96.5\% \pm 2$, i.e., roughly on par with the test set.
+%Thus, it appears that
 

### 2020-05-13 paper/paper.tex
finish subsection on overfitting

@@
 Despite this caveat, we may adopt a ``query-by-example'' (QbE) paradigm by  partitioning the database and audio samples in half ($\frac{1}{2}N^{\prime}=4673$), training the LMNN algorithm on the first half, and querying it with samples from the other half.
-In this evaluation framework, our system reaches a P@5 of $96.2\% \pm 2$.
-Interestingly, querying the system with samples from the training set leads to a P@5 of $96.5\% \pm 2$, i.e., roughly on par with the test set.
-%Thus, it appears that
+In this evaluation framework, our system reaches a P@5 of $\textbf{96.2\%} \pm 2$.
+Interestingly, querying the system with samples from the training set leads to a P@5 of $\textbf{96.5\%} \pm 2$, i.e., roughly on par with the test set.
+Thus it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
 

### 2020-05-13 paper/paper.tex
write on metric learning

@@
 \subsection*{Is metric learning necessary?}
-
-The LMNN algorithm is a popular algorithm for solving such a task, but other alternatives can be considered.
-One of those is the linear discriminant analysis (LDA) \ml{vl ref ?}\cite{}.
-When considering LDA in place of LMNN, the performance drops to $76.6\% \pm 10.9$.
-The lower performance of LDA may be explained by the fact that the dimensionality of the projected feature set is set to one plus the number of clusters in the reference clustering, which lowers the dimensionality from $1180$ to $20$.
-
-Including a learned block implies some constraints.
-In particular, learning the linear transform on large datasets requires more computational time.
-Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
-It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
-In this case, the performance reduces to $92.9\% \pm 3$.
-\ja{Why the $\pm$ here and not elsewhere?}%
+Replacing the large-margin nearest neighbors (LMNN) metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
+However, we should note that, from a computational standpoint, the nearest-neighbor search $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ is considerably faster with LDA than with LMNN.
+Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $(C+1)$, where $C$ is the number of clusters in the cluster graph $\mathcal{G}$.
+% TODO complete this paragraph. 1180/20 ^2 = 59^2 = 3481
+
+Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all.
+This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
+%Including a learned block implies some constraints.
+%In particular, learning the linear transform on large datasets requires more computational time.
+%Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
+%It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
+%In this case, the performance reduces to $92.9\% \pm 3$.
 
@@
 % TODO VL: add an introductory sentence here. "Risk of overfitting of reprojection techniques using LMNN"
-In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-In this case, the performance drops to $23.6\% \pm 10.6$.
+%In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
+%In this case, the performance drops to $23.6\% \pm 10.6$.
 %Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
-It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
-The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
+%It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
+%The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
 

### 2020-05-14 paper/paper.tex
clarify that C_0 = 19. write more on metric learning

@@
 Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
-% TODO describe the gist of hypergraph partitioning?
-To construct $G_0$, we select a number of clusters $C=5$ a priori, and run hypergraph partitioning to assign each vertex $i$ to one of these clusters.
+
+For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
+Across all $K=31$ subjects, the value of $C_k$ varies between $3$ and $19$, with a median value of $10$.
+To construct $\mathcal{G}_0$, we select a number of clusters $C_0$ that is equal to the maximal value of the $C_k$'s; that is,
+\begin{equation}
+C_0 = \max_{1\leq k \leq K} C_k = 19
+\end{equation}
+Then, we run hypergraph partitioning on $\mathcal{H}$ to assign each vertex $i$ to one of the $C_0$ clusters in $\mathcal{G}_0$.
 Intuitively, hypergraph partitioning optimizes a tradeoff between two objectives: first, balancing the size of all clusters in terms of their respective numbers of vertices; and secondly, keeping most hyperedges enclosed within as few distinct clusters as possible.
@@
 This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL (roughly one third of the entire dataset, i.e., $25444$ samples) for which the notion of timbre is perceptually encoded as a discrete category.
-In the case of the consensus cluster graph ($\mathcal{G}_0$), this discrete category takes one of $C=5$ distinct values.
-However, for a particular subject $k$, this discrete category takes one of $C_k$ distinct values, where the number of timbre clusters varies between $3$ and $19$ depending on $k$.
 
@@
 However, we should note that, from a computational standpoint, the nearest-neighbor search $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ is considerably faster with LDA than with LMNN.
-Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $(C+1)$, where $C$ is the number of clusters in the cluster graph $\mathcal{G}$.
-% TODO complete this paragraph. 1180/20 ^2 = 59^2 = 3481
+Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $d_0 = (C_0+1)$, where $C_0 = 19$ is the number of clusters in the consensus cluster graph $\mathcal{G}_0$.
+Because the runtime complexity of nearest-neighbor search grows quadratically with the dimensionality, i.e., is $O(d^2)$, replacing LMNN by LDA accelerates the system by a factor $(d/d_0)^2 \approx 3 \cdot 10^3$.
 
@@
 This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
-%Including a learned block implies some constraints.
-%In particular, learning the linear transform on large datasets requires more computational time.
-%Moreover, the learned transform must be updated if the reference clustering or the dataset are change significantly.
-%It is therefore reasonable to consider the performance of the raw joint time--frequency scattering coefficients without a learned block.
-%In this case, the performance reduces to $92.9\% \pm 3$.
-
-%Another risk when considering reprojection techniques such as the LMNN may appear when the dimensionality of the features approaches the number of items in the dataset.
-%For example, there exists a trivial solution for optimizing $p@1$ when the dimensionality of the features is equal to the number of items in the dataset.
-%\ja{What is that trivial solution?}%
-%In this case, no information needs to be carried by the features set to obtain perfect performance.
+Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search.
+Over our cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
+This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.
 
@@
 
-Concerning the feature computation block $B_1$, an important factor is the time support.
-When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $90.9\% \pm 4$.
+Our 
+When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $\textbf{90.9\%} \pm 4$.
 This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.
-\ja{Other references?}%
 

### 2020-05-14 paper/paper.tex
write on the constant T

@@
 
-Our 
+Our best performing system operates with joint time--frequency scattering coefficients as spectrotemporal modulation features.
+These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
+This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
+
 When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $\textbf{90.9\%} \pm 4$.

### 2020-05-14 paper/paper.tex
recall the definition of S2

@@
 Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search.
-Over our cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
+Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
 This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.
@@
 This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
-
-When considering the commonly used duration of $25~\mathrm{ms}$, the performance drops down to $\textbf{90.9\%} \pm 4$.
-This demonstrates that the joint scattering is able to benefit from a larger time support, possibly to capture longer-term modulations which play an important part when considering extended playing techniques \cite{lostanlen2018extended}.
+Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$.
+
+As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
+Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
+This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.
 
@@
 
-Considering the separable scattering gives a performance of $91\% \pm 4.5$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}
+Lastly, we consider the impact of alternative choices for feature extraction.
+Let us recall the full definition of joint time--frequency scattering:
+\begin{equation}
+\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
+\Big\vert
+\big\vert
+\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
+\big\vert
+\overset{t}{\ast}
+\boldsymbol{\psi}_{\alpha}
+\overset{\log_2 \lambda}{\ast}
+\boldsymbol{\psi}_{\beta}
+\Big\vert
+\overset{t}{\ast} \boldsymbol{\phi}_{T}
+\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F},
+\end{equation}
+where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
+
+Considering the separable scattering gives a performance of $91\% \pm 4.5$.
 % TODO VL: write about frequency-depedendent time shift. cite IEEE TSP
@@
 Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
-Note that the MFCCs are equivalent to first-order separable scattering coefficients \cite{anden2014tsp}.
 Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
 In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
-\ja{Explain why this is necessary.}%
 To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
@@
 The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
-\ml{vl outro on modulations}
 % TODO VL: insist on the role of modulations in playing techniques.

### 2020-05-14 paper/paper.tex
write equation for separable time-frequency scattering
reword outline

@@
 The \lnameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
-The \lnameref{sec:survey} section describes our data collection procedure for auditory similarities across playing techniques.
-The \lnameref{sec:method} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-The \lnameref{sec:results} section conducts an ablation study of our system.
-The \lnameref{sec:discussion} section summarizes the implications of our findings for future research on timbral similarity.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+The \lnameref{sec:survey} section describes our computer--human interface for annotating auditory similarities across playing techniques.
+The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+The \lnameref{sec:results} section describes our experimental benchmark.
+The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
+The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
 
@@
 
-\section*{\nmu Timbre similarity judgments}
+\section*{\nmu Data collection}
 \label{sec:survey}
@@
 
-\section*{Ablation study}
+\section*{\nmu Discussion}
+\label{sec:discussion}
+
 The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
@@
 \begin{equation}
-\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
+\mathbf{S_2^{\mathrm{joint}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
+\left(
 \Big\vert
@@
 \overset{t}{\ast} \boldsymbol{\phi}_{T}
-\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F},
+\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}\right)(t),
 \end{equation}
 where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
+
+% as a point of comparison
+% complex modulus between alpha and beta
+
+\begin{equation}
+\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
+\left(
+\bigg\vert
+\Big\vert
+\big\vert
+\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
+\big\vert
+\overset{t}{\ast}
+\boldsymbol{\psi}_{\alpha}
+\Big\vert
+\overset{t}{\ast} \boldsymbol{\phi}_{T}
+\overset{\log_2 \lambda}{\ast}
+\boldsymbol{\psi}_{\beta}
+\bigg\vert
+\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}
+\right)(t),
+\end{equation}
 
@@
 
-\section*{\nmu Discussion}
-\label{sec:discussion}
+\section*{\nmu Conclusion}
+\label{sec:conclusion}
 

### 2020-05-14 paper/paper.tex
write more on separable scattering

@@
 
-% TODO VL: add an introductory sentence here. "Risk of overfitting of reprojection techniques using LMNN"
-%In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-%In this case, the performance drops to $23.6\% \pm 10.6$.
-%Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
-%It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
-%The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
-
 \subsection*{How important is the amount of temporal context $T$?}
@@
 \overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}\right)(t),
+\label{eq:joint-scattering}
 \end{equation}
 where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
-
-% as a point of comparison
-% complex modulus between alpha and beta
-
+Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
 \begin{equation}
@@
 \right)(t),
+\label{eq:separable-scattering}
 \end{equation}
-
-Considering the separable scattering gives a performance of $91\% \pm 4.5$.
+which has been applied successfully to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.
+
+In comparison with its joint counterpart, separable time--frequency scattering contains about half as many coefficients.
+This is because the temporal wavelet transform with $\boldsymbol{\psi}_\alpha$ and the frequential wavelet transform with $\boldsymbol{\psi}_\beta$ are separated by an operation of complex modulus.
+Hence, $\boldsymbol{\psi}_\beta$ operates on a real-valued input, and flipping the sign of the scale variable $\beta$ is no longer necessary.
+Moreover, separable time--frequency scattering has a lower algorithmic complexity than joint time--frequency scattering.
+Indeed, in Equation \ref{eq:separable-scattering}, the frequential wavelet transform with $\boldsymbol{\psi}_{\beta}$ operates on a tensor whose time axis is subsampled at a fixed rate $T^{-1}$, thus allowing vectorized computations.
+Conversely, in Equation \ref{eq:joint-scattering}, the frequential wavelet transform must operate on a multiresolution input, whose sample rate varies  depending on $\alpha$: it ranges between $T^{-1}$ and the sample rate of $\boldsymbol{x}$ itself.
+
+Considering the separable scattering gives a performance of $\textbf{91\%} \pm 4.5$.
 % TODO VL: write about frequency-depedendent time shift. cite IEEE TSP
-
+% TODO VL: insist on the role of modulations in playing techniques.
+
+\subsection*{What is the baseline performance?}
 Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
@@
 The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
-% TODO VL: insist on the role of modulations in playing techniques.
+
+% TODO VL: add an introductory sentence here. "Risk of overfitting of reprojection techniques using LMNN"
+%In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
+%In this case, the performance drops to $23.6\% \pm 10.6$.
+%Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
+%It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
+%The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
 
@@
 Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
-This explains the relative success over such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity \cite{bauge2013representing}.
+This explains the relative success over such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity.
 

### 2020-05-14 paper/paper.tex
finish section on separable scattering, cite schadler

@@
 
-Considering the separable scattering gives a performance of $\textbf{91\%} \pm 4.5$.
-% TODO VL: write about frequency-depedendent time shift. cite IEEE TSP
-% TODO VL: insist on the role of modulations in playing techniques.
+Yet, despite its greater simplicity, separable time--frequency scattering suffers from known weaknesses in its ability to represent spectrotemporal modulations.
+In particular, a previous publication has shown that frequency-dependent time shifts affect joint time--frequency scattering coefficients while leaving separable time--frequency scattering coefficients almost unchanged \cite{anden2019tsp}.
+For this reason, separable time--frequency scattering coefficients can only serve to recognize periodic modulations (e.g.
+vibrato), but joint time--frequency scattering is required to recognize pitch evolution-based techniques (PETs, e.g.
+glissando) \cite{wang2020icassp}.
+The same observation was made by \cite{schadler2015jasa} in the case of separable Gabor filterbank features (GBFB), which bear some resemblance with separable time--frequency scattering.
+
+Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
+This figure is noticeably worse than joint time--frequency scattering, all other things being equal.
 

### 2020-05-15 paper/paper.tex
add keywords
from the IEEE taxonomy
comment more on joint vs separable

@@
 \begin{keyword}
-\kwd{sample}
-\kwd{article}
-\kwd{author}
+\kwd{audio databases}
+\kwd{continuous wavelet transform}
+\kwd{demodulation}
+\kwd{distance learning}
+\kwd{human--computer interaction}
+\kwd{music information retrieval.}
 \end{keyword}
@@
 Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
-This figure is noticeably worse than joint time--frequency scattering, all other things being equal.
+This figure is noticeably worse than joint time--frequency scattering ($99.0\% \pm 1$), all other things being equal.
+This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:soa} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
 

### 2020-05-15 paper/paper.tex
fix typo
$k\geq 1$

@@
 In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
-However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
+However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
 % browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
@@
 
-Lastly, we consider the impact of alternative choices for feature extraction.
 Let us recall the full definition of joint time--frequency scattering:
@@
 \subsection*{What is the baseline performance?}
+Lastly, we consider the impact of alternative choices for feature extraction.
+
+$81.8\% \pm 7.2$
+
+Arguably, this figure is not directly comparable with our best performing system, since
+
+To disprove this hypothesis, 
 Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.

### 2020-05-15 paper/paper.tex
fix typo

@@
 Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
-This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{}, an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
+This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{} an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
 

### 2020-05-16 paper/paper.tex
finish discussion section

@@
 Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.

### 2020-05-16 paper/paper.tex
fix typo

@@
 
-In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must withdraw from a diverge realm of instruments and techniques.
+In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must withdraw from a diverse realm of instruments and techniques.
 Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity level (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.

### 2020-05-16 paper/paper.tex
fix typo

@@
 In the context of browsing an audio collection by timbre similarity, $\boldsymbol{x}_n$ is a user-defined query while the function $\mathbf{\Phi}$ plays the role of a search engine.
-We consider that the first retrieved sample, $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, to be relevant to user $k$ if and only if it belongs to the same cluster as $\boldsymbol{x}_n$ in the cluster graph $\mathcal{G}_k$; hence the Boolean condition $\mathbf{\Phi}_{1}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
+We consider the first retrieved sample, $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, to be relevant to user $k$ if and only if it belongs to the same cluster as $\boldsymbol{x}_n$ in the cluster graph $\mathcal{G}_k$; hence the Boolean condition $\mathbf{\Phi}_{1}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
 Likewise, the second retrieved sample is relevant if and only if $\mathbf{\Phi}_{2}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.

### 2020-05-16 paper/paper.tex
rephrase

@@
 In the rest of this paper, we set the constant $R$ to $5$.
-This is in accordance with the protocol of \cite{lostanlen2018extended}, who trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
+This is in accordance with the protocol of \cite{lostanlen2018extended}, in which the authors trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
 Moreover, note that the value $R=5$ typically reflects the number of virtual panels simultaneously on display in a software environment for computer-assisted orchestration.

### 2020-05-16 paper/paper.tex
k  \geq 1
punctuation

@@
 Rather, the LMNN algorithm is trained on all available samples ($N^{\prime}=9346$ isolated notes) with the ``consensus'' cluster graph as training objective ($\mathcal{G}=\mathcal{G}_0$).
-Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k>0$).
+Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$).
 The reason behind this choice is that, in the context of computer-assisted orchestration, the collection of audio samples has a fixed size, as it is shipped alongside the software itself.
@@
 Interestingly, querying the system with samples from the training set leads to a P@5 of $\textbf{96.5\%} \pm 2$, i.e., roughly on par with the test set.
-Thus it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
+Thus, it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
 

### 2020-05-16 paper/paper.tex
write on the mfcc baseline

@@
 \subsection*{What is the baseline performance?}
-Lastly, we consider the impact of alternative choices for feature extraction.
-
-$81.8\% \pm 7.2$
-
-Arguably, this figure is not directly comparable with our best performing system, since
-
-To disprove this hypothesis, 
-Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
-Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
-In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
-To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
-\ja{What are the $f_i$ here?
-How are they defined?}%
-% TODO VL: give reference to MFCCs.
-The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
-The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
+Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
+Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat/}}.
+This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
+Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{81.8\%} \pm 7$.
+
+Arguably, this figure is not directly comparable with our best performing system (P@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
+In order to clarify the role of feature dimensionality in our computational pipeline, we apply a feature engineering technique involving multiplicative combinations of MFCC.
+We construct the following Gram matrix:
+\begin{equation}
+\mathbf{G} \boldsymbol{x}(\alpha, \beta) =
+\int_0^{+\infty}
+\mathrm{MFCC}(\boldsymbol{x})(t, \alpha)
+\times
+\mathrm{MFCC}(\boldsymbol{x})(t, \beta)
+\;\mathrm{d}t,
+\end{equation}
+where $\alpha$ and $\beta$ represent different dimensions (``quefrencies'') of the MFCC feature vector.
+The symmetric matrix $\mathbf{G}\boldsymbol{x}$ contains $40$ rows and $40$ columns, hence $800$ unique coefficients.
+Concatenating these coefficients to the $40$ averaged MFCC features results in a feature vector of $840$ coefficients.
+This dimension is of the same order of magnitude as the dimension of our joint time--frequency scattering representation ($d=1180$, see \lnameref{sec:methods} section).
+
+Training the LMNN algorithm on this 840-dimensional representation is analogous to a ``kernel trick'' in support vector machines \cite{chang2010jmlr}.
+In our case, the implicit similarity kernel is a homogeneous quadratic kernel.
+Despite this increase in representational power, we obtain a P@5 of $81.5\% \pm 7.3$, i.e. essentially the same as MFCC under a linear kernel.
+Therefore, it appears that the gap in performance between MFCC ($81.8\%\pm 7$) and joint time--frequency scattering ($99.0\% \pm 1$) is primarily attributable to the multiscale extraction of joint spectrotemporal modulations, whereas high-dimensional embedding likely plays a minor role.
 

### 2020-05-16 paper/paper.tex
fix baugé citation

@@
 Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,bauge2013representing,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
 %\ja{Need to flesh out the above with some equations.

### 2020-05-16 paper/paper.tex
write abstract

@@
 
+%%% Put your definitions there:
+\startlocaldefs
+
+\newcommand*{\eg}{e.g.,\@\xspace}
+\newcommand*{\ie}{i.e.,\@\xspace}
+\newcommand*{\vs}{vs.\@\xspace}
+
+\newcommand{\ipt}{IPT\xspace}
+\newcommand{\ipts}{IPTs\xspace}
+
+\newcommand{\patk}{p\mathrm{@}k}
+
+\newcommand{\Card}{\mathrm{card}}
+
+\newcommand{\ml}[1]{\textcolor{blue}{ML: #1}\xspace}
+\newcommand{\vl}[1]{\textcolor{red}{VL: #1}\xspace}
+\newcommand{\ja}[1]{\textcolor{purple}{JA: #1}\xspace}
+
 \newcommand{\lnameref}[1]{%
@@
 \newcommand{\nmu}{}
-
-%%% Put your definitions there:
-\startlocaldefs
-
-\newcommand*{\eg}{e.g.,\@\xspace}
-\newcommand*{\ie}{i.e.,\@\xspace}
-\newcommand*{\vs}{vs.\@\xspace}
-
-\newcommand{\ipt}{IPT\xspace}
-\newcommand{\ipts}{IPTs\xspace}
-
-\newcommand{\patk}{p\mathrm{@}k}
-
-\newcommand{\Card}{\mathrm{card}}
-
-\newcommand{\ml}[1]{\textcolor{blue}{ML: #1}\xspace}
-\newcommand{\vl}[1]{\textcolor{red}{VL: #1}\xspace}
-\newcommand{\ja}[1]{\textcolor{purple}{JA: #1}\xspace}
 
@@
 \begin{abstract}
-Extended playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in erudite and folk contexts.
-They affect the spectrotemporal modulations of pitched audio signals, thereby modifying timbre perception systematically.
-However, most existing approaches to music similarity retrieval inadequately reflect this effect, as they fail to describe timbre beyond the so-called ``ordinary'' technique; use instrument identity as a proxy for timbre quality; or do not allow for rapid customization to the perceptual idiosyncracies of a newcoming subject.
-%Our article fills this gap in scholarship by providing a scalable algorithm for measuring the timbre similarity between any two audio samples
+Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in erudite and folk contexts.
+However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique; use instrument identity as a proxy for timbre quality; and do not allow for rapid customization to the perceptual idiosyncracies of a newcoming subject.
+In this article, we ask 31 human subjects to organize 78 isolated notes into an unconstrained number of timbre clusters.
+Analyzing their responses suggests that timbre perception operates within a more flexible taxonomy than those provided by instruments or playing techniques alone.
+In addition, we propose a machine listening model to recover the cluster graph of auditory similarities across instruments, mutes, and techniques.
+Our model relies on joint time--frequency scattering features to extract spectrotemporal modulations at multiple scales and rates.
+Furthermore, it minimizes triplet loss in the cluster graph by means of the large-margin nearest neighbor (LMNN) metric learning algorithm.
+Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of $99.0\%\pm 1$.
+An ablation study demonstrates that removing either the joint time--frequency scattering transform or the metric learning algorithm noticeably degrades performance.
 \end{abstract}
-
-%\begin{abstract}
-%The timbre of a sound is an important factor in determining its identity and source.
-%One successful means of exploring the timbre space has been the construction of computational models to analyze datasets of musical instrument recordings.
-%However, these datasets have typically only included a limited set of instruments played with standard playing techniques, limiting the timbral range.
-%For these datasets, models based on static spectral descriptors, such as mel-frequency cepstral coefficients (MFCCs), have enjoyed excellent performance.
-%However, with richer datasets containing a wider array of instruments played with a variety of playing techniques, more sophisticated models which incorporate dynamic large-scale signal structure are needed.
-%To this end, we propose a simple but rich model based on joint time--frequency scattering transforms followed by a linear mapping.
-%The linear part is learned from human similarity judgments obtained in a crowdsourced free sorting task, so the resulting descriptor provides a perceptually motivated characterization of timbre.
-%We show that this proposed computational model accurately reproduces the human similarity judgments to $99\%$ accuracy.
-%In addition, we perform an ablation study demonstrating the necessity of various components in this model.
-%\end{abstract}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
-Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014tsp}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
-% TODO cite wang 2020 icassp
+The authors of a recent publication \cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014tsp}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
 Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
@@
 % TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
-% cite Musgrave et al. Metric learning reality check
 
@@
 However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
-% browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.
-
-% TODO VL: organize design choices.
-% (1) temporal support
-% (2) time-frequency scattering / time scattering
-% (3) learning / no learning
 
@@
 
-% TODO VL: add an introductory sentence here. "Risk of overfitting of reprojection techniques using LMNN"
 %In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
@@
 
-\section*{\nmu Human timbre judgments}
-\label{sec:subjective}
-
-The dataset considered in this study is taken
-from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
-It has audio recordings of 16 different musical instruments played
-with different playing techniques,
-which leads to $235$ different couples of instrument / playing technique.
-For each couple, the pitch and intonation is varied leading to $25444$ recordings.
+%\section*{\nmu Human timbre judgments}
+%\label{sec:subjective}
+
+%The dataset considered in this study is taken
+%from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
+%It has audio recordings of 16 different musical instruments played
+%with different playing techniques,
+%which leads to $235$ different couples of instrument / playing technique.
+%For each couple, the pitch and intonation is varied leading to $25444$ recordings.
 %The way the dataset is structured is explained in more details in Appendix \ref{sec:appendix-dataset}.
 
-Gathering the similarity judgments between each pair of samples would be a daunting task.
+%Gathering the similarity judgments between each pair of samples would be a daunting task.
 Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
 
-One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
+%One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
 This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
 
-\section*{\nmu Reducing Expert judgments of technique peculiarities}
-
-\ja{This needs to be summarized (possible in the previous section).
-I'm copying two sentences in here that I had in the Timbre similarity judgments section.}%
-These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
-The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
-
-Even so, considering every pairs of IPTs is not practical. A selection
-experiment is thus conducted where the subject is asked to give his opinion on which
-\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
-if it is likely to be associated to another \ipt of another instrument.
-Interest is rated on a 7 ticks scale. The subject can listen to all the different
-nuance and pitch samples of each \ipt. The experiment is over when all the
-\ipts are rated.
-
-Rating guidelines are provided :
-\begin{itemize}
-  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
-  \item Four stars: there is a proximity on one aspect of the sound between this
-  \ipt and of an \ipt of another instrument,
-  but this one is neither decisive nor obvious.
-  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
-\end{itemize}
-
-Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
+%\section*{\nmu Reducing Expert judgments of technique peculiarities}
+
+%These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
+%The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
+
+%Even so, considering every pairs of IPTs is not practical. A selection
+%experiment is thus conducted where the subject is asked to give his opinion on which
+%\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
+%if it is likely to be associated to another \ipt of another instrument.
+%Interest is rated on a 7 ticks scale. The subject can listen to all the different
+%nuance and pitch samples of each \ipt. The experiment is over when all the
+%\ipts are rated.
+
+%Rating guidelines are provided :
+%\begin{itemize}
+%  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
+%  \item Four stars: there is a proximity on one aspect of the sound between this
+%  \ipt and of an \ipt of another instrument,
+%  but this one is neither decisive nor obvious.
+%  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
+%\end{itemize}
+
+%Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
 
@@
 
-Even if the hypothesis that the distributions of ratings comes from independent random samples from normal distributions with equal means and equal but unknown variances, is verified using the two-sample t-test, inspection of the respective histograms reveals that averaging the ratings of the two raters is reasonable.
-
-Keeping in mind that the gathering of the perceptual judgments of \ipt similarities shall remain practical, the number of \ipts considered shall be lower than $100$.
+%Even if the hypothesis that the distributions of ratings comes from independent random samples from normal distributions with equal means and equal but unknown variances, is verified using the two-sample t-test, inspection of the respective histograms reveals that averaging the ratings of the two raters is reasonable.
+
+%Keeping in mind that the gathering of the perceptual judgments of \ipt similarities shall remain practical, the number of \ipts considered shall be lower than $100$.
 We thus select the \ipts with an average rating higher than three, which brings us 78 \ipts to study which are listed in Tables \ref{tab:list1} and \ref{tab:list2}.
@@
 %%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
-\section*{\nmu Crowdsourced judgments of pairwise similarity}
-
-\ja{A lot of this is covered in the main text.
-Not sure what we need to add in the appendix.}
-
-The second experiment aims at gathering spontaneous judgment of similarity for those 78 \ipts using a canonical free sorting task.
+%\section*{\nmu Crowdsourced judgments of pairwise similarity}
+
+%\ja{A lot of this is covered in the main text.
+%Not sure what we need to add in the appendix.}
+
+%The second experiment aims at gathering spontaneous judgment of similarity for those 78 \ipts using a canonical free sorting task.
 This type of task is chosen among others for two main reasons.
@@
 
-31 subjects completed the experiment.
+%31 subjects completed the experiment.
 An example of resulting organization of \ipt is given on Figure \ref{fig:xp2display}.
@@
 
-Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
+%Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
 The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
@@
 
-This results in $31$ clusterings that can be considered for defining a perceptual timbre space of \ipts.
+%This results in $31$ clusterings that can be considered for defining a perceptual timbre space of \ipts.
 Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipts.
@@
 
-\begin{figure}
-\center
-\begin{tabular}{cc}
-  (a) & (b) \\
-  \includegraphics[width = \textwidth]{figures/nbc.png} &
-  \includegraphics[width = \textwidth]{figures/sbc.png}
-\end{tabular}
-\caption{Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects.}
-\label{fig:xp2nbSize}
-\end{figure}
+%\begin{figure}
+%\center
+%\begin{tabular}{cc}
+%  (a) & (b) \\
+%  \includegraphics[width = \textwidth]{figures/nbc.png} &
+%  \includegraphics[width = \textwidth]{figures/sbc.png}
+%\end{tabular}
+%\caption{Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects.}
+%\label{fig:xp2nbSize}
+%\end{figure}
 
@@
 
-\section*{Additional Files}
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
-
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
+%\section*{Additional Files}
+%  \subsection*{Additional file 1 --- Sample additional file title}
+%    Additional file descriptions text (including details of how to
+%    view the file, if it is in a non-standard format or the file extension).  This might
+%    refer to a multi-page table or a figure.
+
+%  \subsection*{Additional file 2 --- Sample additional file title}
+%    Additional file descriptions text.
 

### 2020-05-16 paper/paper.tex
move fig 1
shorten footnote

@@
 \usepackage{amsthm,amsmath}
+\usepackage{amssymb}
 \usepackage{dsfont}
@@
 
-\begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
-\caption{\csentence{Overview of the proposed approach}}
-\label{fig:pipeline}
-\end{figure}
-\clearpage
-
 \subsection*{Outline}
@@
 The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
-The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{Reproducible research statement: the implementation of the proposed model and the experimental protocol of this paper are available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
 
@@
 
-%\section*{Figures}
+\section*{Figures}
+
+\begin{figure}[h!]
+\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
+\caption{\csentence{Overview of the proposed approach}}
+\label{fig:pipeline}
+\end{figure}
 

### 2020-05-16 paper/paper.tex
move figures

@@
 
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/consensusVsI.png}
-\caption{
-%TODO caption
-}
-\label{fig:consensusVsI}
-\end{figure}
-
 \subsection*{Inter-technique similarity}
@@
 
-\begin{figure}
-\center
-\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).
-%TODO add a double arrow
-% rough <-- soft --> rough
-% sustained <-- "ordinary" --> percussive
-}
-\label{fig:consensusVsPt}
-\end{figure}
-
 \section*{\nmu Methods}
@@
 \texttt{
-case 'scat'
-    filt\_opt.Q = [12 1];
-    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
 case 'tfscat'
@@
     fr\_filt\_opt.J = 4;}}
-
-\begin{figure}
-\caption{\csentence{Sample scattering}
-Same 3 sounds spec + scatt.
-% TODO figure + caption
-}
-\end{figure}
 
@@
 
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/consensusVsI.png}
+\caption{
+%TODO caption
+}
+\label{fig:consensusVsI}
+\end{figure}
+
+\begin{figure}
+\center
+\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
+\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).
+%TODO add a double arrow
+% rough <-- soft --> rough
+% sustained <-- "ordinary" --> percussive
+}
+\label{fig:consensusVsPt}
+\end{figure}
+
+\begin{figure}
+\caption{\csentence{Sample scattering}
+Same 3 sounds spectrogram + scattering
+% TODO figure + caption
+}
+\label{fig:sample-scattering}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-05-16 paper/paper.tex
start writing on scattering

@@
 
-While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
-As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
-
-A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
-Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
-The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
-This representation contains a large amount of information, but not in a form suitable for building statistical models.
-In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
-It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
-One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
-
-Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
-One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
-Again, we compute the modulus of the wavelet coefficients and average in time.
-The resulting coefficients are known as the second-order time scattering coefficients.
-These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
-
-While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
-Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
-If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
-We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
-This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
-
-This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
-To remedy this, another approach was suggested by And\'{e}n et al.
+%While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
+%As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
+
+%A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
+%Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
+%The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
+%This representation contains a large amount of information, but not in a form suitable for building statistical models.
+%In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
+%It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
+%One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
+%These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
+
+%Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
+%One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
+%Again, we compute the modulus of the wavelet coefficients and average in time.
+%The resulting coefficients are known as the second-order time scattering coefficients.
+%These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
+
+%While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
+%Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
+%If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
+%We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
+%This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
+%The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
+
+%This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
+%To remedy this, another approach was suggested by And\'{e}n et al.
 \cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
-Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
-
-The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
-On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
-The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
-They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
+%Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
+%As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
+
+%The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
+%On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
+%The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
+%Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
+%They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
 %\ja{Need to flesh out the above with some equations.
@@
 % TODO: avoid citing the same papers over and over
+
+Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ a Hilbert-analytic filter with null average, unit dimensionless center frequency, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
+We define a constant-$Q$ wavelet filterbank as the family $\boldsymbol{\psi}_{\lambda} : t \mapsto \lambda \boldsymbol{\psi}(\lambda t)$.
+Each wavelet $\boldsymbol{\psi}_{\lambda}$ has a center frequency of $\lambda$, an ERB of $\lambda/Q$, and an effective receptive field of $(2\pi Q/\lambda)$ in the time domain.
+In practice, we define $\boldsymbol{\psi}$ as a Morlet wavelet:
+\begin{equation}
+\boldsymbol{\psi}:t \longmapsto \exp\left(-\dfrac{t^2}{2\sigma_{\psi}^2}\right) 
+\big[ \exp\left(2\pi \mathrm{i}t\right) - \kappa_{\psi} \big],
+\end{equation}
+where the Gaussian width $\sigma_{\psi}$ grows in proportion with the quality factor $Q$ and the corrective term $\kappa_{\psi}$ ensures that $\boldsymbol{\psi}$ has a null average over $\mathbb{R}$.
+Moreover, we discretize the frequency variable $\lambda$ according to a geometric progression of common ratio $2^{\frac{1}{Q}}$.
+
+We set the constant quality factor of the wavelet filterbank $(\boldsymbol{\psi}_\lambda)_\lambda$ to $Q=12$, thus matching twelve-tone equal temperament in music.
+Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R}$, followed by an application of the pointwise complex modulus, yields the wavelet scalogram
+\begin{equation}
+\mathbf{U_1}\boldsymbol{x}(t,\lambda) =
+\big\vert
+\boldsymbol{x}
+\ast
+\boldsymbol{\psi}_{\lambda}
+\big\vert(t)
+=
+\left\vert
+\int_{\mathbb{R}}
+\boldsymbol{x}(t^{\prime})
+\,
+\boldsymbol{\psi}_{\lambda} (t - t^{\prime})
+\;
+\mathrm{d}{t^\prime}
+\right\vert,
+\end{equation}
+indexed by time $t$ and frequency $\lambda$.
+
+\begin{equation}
+\Psi : (t,\log_2 \lambda) \longmapsto 
+\exp\left(-\dfrac{t^2+(\log_2 \lambda)^2}{2\sigma^2}\right) 
+\Big[ \exp\big(2\pi \mathrm{i} (t + \log_2 \lambda)\big) - \kappa \Big]
+\end{equation}
 

### 2020-05-16 paper/paper.tex
write more on scattering

@@
 \big[ \exp\left(2\pi \mathrm{i}t\right) - \kappa_{\psi} \big],
+\label{eq:psi}
 \end{equation}
@@
 Moreover, we discretize the frequency variable $\lambda$ according to a geometric progression of common ratio $2^{\frac{1}{Q}}$.
-
+Thus, the base-two logarithm of center frequency, denoted by $\log_2 \lambda$, follows an arithmetic progression.
 We set the constant quality factor of the wavelet filterbank $(\boldsymbol{\psi}_\lambda)_\lambda$ to $Q=12$, thus matching twelve-tone equal temperament in music.
+
 Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R}$, followed by an application of the pointwise complex modulus, yields the wavelet scalogram
 \begin{equation}
-\mathbf{U_1}\boldsymbol{x}(t,\lambda) =
+\mathbf{U_1}\boldsymbol{x}(t,\log_2\lambda) =
 \big\vert
@@
 \end{equation}
-indexed by time $t$ and frequency $\lambda$.
-
+indexed by time $t$ and log-frequency $\log_2\lambda$.
+Then, we define a two-dimensional Morlet wavelet $\boldsymbol{\Psi}$, taking a spectrotemporal pair $(t, \log_2 \lambda)$ as input:
 \begin{equation}
 \Psi : (t,\log_2 \lambda) \longmapsto 
-\exp\left(-\dfrac{t^2+(\log_2 \lambda)^2}{2\sigma^2}\right) 
-\Big[ \exp\big(2\pi \mathrm{i} (t + \log_2 \lambda)\big) - \kappa \Big]
+\exp\left(-\dfrac{t^2+(\log_2 \lambda)^2}{2\sigma_{\Psi}^2}\right) 
+\Big[ \exp\big(2\pi \mathrm{i} (t + \log_2 \lambda)\big) - \kappa_{\Psi} \Big].
 \end{equation}
-
-\ml{VL. provide parameters for scattering
-\texttt{
-case 'tfscat'
-    tm\_filt\_opt.Q = [12 1];
-    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
-    fr\_filt\_opt.J = 4;}}
+In the equation above, the Gaussian width $\sigma_{\Psi}$ is chosen so that the quality factor of the Morlet wavelet $\Psi$ is equal to $1$, both over the time dimension and over the log-frequency dimension.
+Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
 

### 2020-05-16 paper/paper.tex
bis

@@
 Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
+Following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:soa} section), we define a two-dimensional wavelet filterbank of the form:
+\begin{equation}
+\mathbf{\Psi}_{\alpha,\beta} : (t, \log_2 \lambda) \longmapsto
+\mathbf{\Psi}\big(\alpha t, \beta \log_2 \lambda \big),
+\end{equation}
+where $\alpha$ is a temporal modulation rate and $\beta$ is a temporal modulation scale.
+
+While $\alpha$ is measured in Hertz and is strictly positive, $\beta$ is measured in cycles per octaves and may take positive as well as negative values.
+Both $\alpha$ and $\beta$ are discretized by geometric progressions of common ratio equal to $2$.
+Furthermore, the edge case $\beta=0$ corresponds to $\mathbf{\Psi}_{\alpha,\beta}$ being a Gaussian low-pass filter over the log-frequency dimension, while remaining a Morlet band-pass filter of center frequency $\alpha$ over the time dimension.
+
+% TODO define U2.
+% Briefly define S2.
+
+In comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
+We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.
 

### 2020-05-16 paper/paper.tex
cite schorkhuber and klapuri, write more on scattering

@@
 \subsection*{Outline}
-The \lnameref{sec:soa} section reviews recent research in the computational modeling of musical timbre.
+The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
 The \lnameref{sec:survey} section describes our computer--human interface for annotating auditory similarities across playing techniques.
@@
 \section*{\nmu Related work}
-\label{sec:soa}
+\label{sec:related-work}
 
@@
 
-Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R}$, followed by an application of the pointwise complex modulus, yields the wavelet scalogram
-\begin{equation}
-\mathbf{U_1}\boldsymbol{x}(t,\log_2\lambda) =
+Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R})$, followed by an application of the pointwise complex modulus, yields the wavelet scalogram
+\begin{equation}
+\mathbf{U_1}\boldsymbol{x}(t,\lambda) =
 \big\vert
@@
 \end{equation}
-indexed by time $t$ and log-frequency $\log_2\lambda$.
-Then, we define a two-dimensional Morlet wavelet $\boldsymbol{\Psi}$, taking a spectrotemporal pair $(t, \log_2 \lambda)$ as input:
-\begin{equation}
-\Psi : (t,\log_2 \lambda) \longmapsto 
-\exp\left(-\dfrac{t^2+(\log_2 \lambda)^2}{2\sigma_{\Psi}^2}\right) 
-\Big[ \exp\big(2\pi \mathrm{i} (t + \log_2 \lambda)\big) - \kappa_{\Psi} \Big].
-\end{equation}
+which is discretized according to a dyadic multirate scheme, similarly to the constant-$Q$ transform of \cite{schorkhuber2010smc}.
+Then, we define a two-dimensional Morlet wavelet $\boldsymbol{\Psi}$ of the form
+\begin{equation}
+\Psi : (t, u) \longmapsto 
+\exp\left(-\dfrac{t^2+u^2}{2\sigma_{\Psi}^2}\right) 
+\Big[ \exp\big(2\pi \mathrm{i} (t + u)\big) - \kappa_{\Psi} \Big],
+\end{equation}
+taking two real variables $t$ and $u$ as input.
+The former is the time variable while the latter is the base-two logarithm of frequency: $u = \log_2 \lambda$.
+Note that $u$ roughly corresponds to the human perception of relative pitch \cite{lostanlen2020icassp}.
+
 In the equation above, the Gaussian width $\sigma_{\Psi}$ is chosen so that the quality factor of the Morlet wavelet $\Psi$ is equal to $1$, both over the time dimension and over the log-frequency dimension.
 Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
-Following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:soa} section), we define a two-dimensional wavelet filterbank of the form:
+From $\mathbf{\Psi}$, we define a two-dimensional wavelet filterbank of the form:
 \begin{equation}
 \mathbf{\Psi}_{\alpha,\beta} : (t, \log_2 \lambda) \longmapsto
-\mathbf{\Psi}\big(\alpha t, \beta \log_2 \lambda \big),
-\end{equation}
-where $\alpha$ is a temporal modulation rate and $\beta$ is a temporal modulation scale.
-
+\alpha \, \beta \, \mathbf{\Psi}(\alpha t, \beta \log_2 \lambda).
+\end{equation}
+In the equation above, $\alpha$ is a temporal modulation rate and $\beta$ is a temporal modulation scale, following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:related-work} section).
 While $\alpha$ is measured in Hertz and is strictly positive, $\beta$ is measured in cycles per octaves and may take positive as well as negative values.
@@
 
-% TODO define U2.
-% Briefly define S2.
+\begin{equation}
+\mathbf{U_2}\boldsymbol{x} (t, \lambda, \alpha, \beta) =
+\left\vert
+\iint_{\mathbb{R}^2}
+\mathrm{d}t^\prime \, \mathrm{d}u^\prime
+\right\vert
+\end{equation}
 
@@
 This figure is noticeably worse than joint time--frequency scattering ($99.0\% \pm 1$), all other things being equal.
-This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:soa} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
+This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
 

### 2020-05-17 paper/paper.tex
write equation for U_2
update definition of U1
update defitniion of Psi
finish writing on scattering

@@
 \int_{\mathbb{R}}
-\boldsymbol{x}(t^{\prime})
+\boldsymbol{x}(t - t^{\prime})
 \,
-\boldsymbol{\psi}_{\lambda} (t - t^{\prime})
+\boldsymbol{\psi}_{\lambda} (t^{\prime})
 \;
@@
 \right\vert,
+\label{eq:U1}
 \end{equation}
@@
 \Big[ \exp\big(2\pi \mathrm{i} (t + u)\big) - \kappa_{\Psi} \Big],
+\label{eq:Psi}
 \end{equation}
@@
 
-In the equation above, the Gaussian width $\sigma_{\Psi}$ is chosen so that the quality factor of the Morlet wavelet $\Psi$ is equal to $1$, both over the time dimension and over the log-frequency dimension.
+We choose the Gaussian width $\sigma_{\Psi}$ in Equation \ref{eq:Psi} such that the quality factor of the Morlet wavelet $\Psi$ is equal to $1$, both over the time dimension and over the log-frequency dimension.
 Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
@@
 \begin{equation}
-\mathbf{\Psi}_{\alpha,\beta} : (t, \log_2 \lambda) \longmapsto
-\alpha \, \beta \, \mathbf{\Psi}(\alpha t, \beta \log_2 \lambda).
+\mathbf{\Psi}_{\alpha,\beta} : (t, u) \longmapsto
+\alpha \, \beta \, \mathbf{\Psi}(\alpha t, \beta u).
 \end{equation}
@@
 Furthermore, the edge case $\beta=0$ corresponds to $\mathbf{\Psi}_{\alpha,\beta}$ being a Gaussian low-pass filter over the log-frequency dimension, while remaining a Morlet band-pass filter of center frequency $\alpha$ over the time dimension.
-
-\begin{equation}
-\mathbf{U_2}\boldsymbol{x} (t, \lambda, \alpha, \beta) =
+We denote this low-pass filter by $\boldsymbol{\phi}_F$ and set its width to $F=2$ octaves.
+
+Then, we convolve the scalogram $\mathbf{U_1}$ with all time--frequency wavelets $\mathbf{\Psi}_{\alpha,\beta}$ and apply the complex modulus, yielding the four-way tensor
+\begin{align}
+\mathbf{U_2}\boldsymbol{x} (t, \lambda, \alpha, \beta)
+=&
+\big \vert
+\mathbf{U_1}\boldsymbol{x} \circledast \boldsymbol{\Psi}_{\alpha,\beta}
+\big \vert (t, \lambda)
+\nonumber \\
+=&
 \left\vert
 \iint_{\mathbb{R}^2}
+\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)\,
+\boldsymbol{\Psi}_{\alpha,\beta}(t^\prime, u^\prime)\;
 \mathrm{d}t^\prime \, \mathrm{d}u^\prime
-\right\vert
-\end{equation}
-
-In comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
+\right\vert,
+\label{eq:U2}
+\end{align}
+where the circled asterisk operator $\circledast$ denotes a joint convolution over time and log-frequency.
+In the equation above, the sample rate of time $t$ is proportional to $\alpha$.
+Conversely, the sample rate of log-frequency $u = \log_2 \lambda$ is proportional to $\vert\beta\vert$ if $\beta \neq 0$ and proportional to $F^{-1}$ otherwise.
+
+Let $\boldsymbol{\phi}_T$ be a Gaussian low-pass filter.
+We define the joint time--frequency scattering coefficients of the signal $\boldsymbol{x}$ as the four-way tensor
+\begin{align}
+\mathbf{S_2}(t, \lambda, \alpha, \beta)
+=&
+\mathbf{U_2}\boldsymbol{x}
+\circledast
+\big(\boldsymbol{\phi}_T  \otimes \boldsymbol{\phi}_F\big)
+(t, \lambda)
+\nonumber \\
+=&
+\iint_{\mathbb{R}^2}
+\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime, \alpha, \beta)
+\,\boldsymbol{\phi}_T (t^\prime)
+\,\boldsymbol{\phi}_F (u^\prime)
+\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
+\end{align}
+where the symbol $\otimes$ denotes an outer product between two vectors.
+In the equation above, the sample rate of time $t$ is proportional to $T^{-1}$ and the sample rate of log-frequency $u=\log_2 \lambda$ is proportional to $F^{-1}$.
+Furthermore, the rate $\alpha$ spans along a geometric progression ranging from $T^{-1}$ to $\lambda/Q$.
+In the following, we set the time constant to $T=\SI{1000}{\milli\second}$ unless specified otherwise.
+
+The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of a spectrotemporal receptive field (STRF) at the rate $\alpha$ and the scale $\beta$.
+Nevertheless, in comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
+This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.
 We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.

### 2020-05-17 paper/paper.tex
update Equations 13 and 14

@@
 
-Let us recall the full definition of joint time--frequency scattering:
+Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
 \begin{equation}
-\mathbf{S_2^{\mathrm{joint}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
+\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
 \left(
@@
 \big\vert
-\overset{t}{\ast}
-\boldsymbol{\psi}_{\alpha}
-\overset{\log_2 \lambda}{\ast}
-\boldsymbol{\psi}_{\beta}
+\circledast
+\boldsymbol{\Psi}_{\alpha,\beta}
 \Big\vert
-\overset{t}{\ast} \boldsymbol{\phi}_{T}
-\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}\right)(t),
+\circledast
+\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
+\right)(t,\lambda),
 \label{eq:joint-scattering}
 \end{equation}
-where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
+where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
 Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
@@
 \big\vert
-\overset{t}{\ast}
+\ast
 \boldsymbol{\psi}_{\alpha}
 \Big\vert
-\overset{t}{\ast} \boldsymbol{\phi}_{T}
-\overset{\log_2 \lambda}{\ast}
-\boldsymbol{\psi}_{\beta}
+\circledast
+\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
 \bigg\vert
-\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}
-\right)(t),
+\circledast
+\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
+\right)(t, \lambda),
 \label{eq:separable-scattering}

### 2020-05-19 paper/paper.tex
improve scattering section

@@
 Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
-
-% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
-%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
 
@@
 \;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
+\label{eq:S2}
 \end{align}
@@
 This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.
+
+In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension, with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
+\begin{align}
+\mathbf{S_1}\boldsymbol{x}(t, \lambda, \alpha=0, \beta) &=
+\big\vert
+\mathbf{U_1}\boldsymbol{x} \circledast
+\big(\boldsymbol{\phi}_T \otimes \boldsymbol{\psi}_{\beta} \big)
+\big\vert (t, \lambda)
+\nonumber \\
+&=
+\left \vert
+\iint_{\mathbb{R}^2}
+\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
+\,\boldsymbol{\phi}_T (t^\prime)
+\,\boldsymbol{\psi}_{\beta} (u^\prime)
+\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime
+\right \vert,
+\label{eq:S1}
+\end{align}
+where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.
+Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
+Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\psi}_F$, comparably to Equation $\ref{eq:Psi}$.
+Lastly, we concatenate 
+\begin{equation}
+    \mathbf{S}\boldsymbol{x}(t, p) 
+\end{equation}
 We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.
@@
 % TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
+
+%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
 
@@
 \end{equation}
-which has been applied successfully to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.
+where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
+Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.
 

### 2020-05-19 paper/paper.tex
start writing on median-based log

@@
 where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.
+
 Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
 Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\psi}_F$, comparably to Equation $\ref{eq:Psi}$.
-Lastly, we concatenate 
-\begin{equation}
-    \mathbf{S}\boldsymbol{x}(t, p) 
-\end{equation}
 We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.
+
+\subsection*{Per-feature Gaussianization with median-based logarithmic compression}
+\label{sec:gaussianization}
+We adopt the multiindex notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
+The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) or second-order ($\mathbf{S_2}$) coefficients.
+Given an input waveform $\boldsymbol{x}$, we denote by $\mathbf{S}\boldsymbol{x}$ the feature vector resulting from the concatenation of $\mathbf{S_1}\boldsymbol{x}$ and $\mathbf{S_2}\boldsymbol{x}$:
+\begin{equation}
+\mathbf{S}\boldsymbol{x}\big(t, p=(\lambda, \alpha, \beta)\big)
+=\left\{
+\begin{array}{ll}
+\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha\neq0,\\
+\mathbf{S_{2}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{otherwise.}
+\end{array}\right.
+\end{equation}
+
+% TODO declare new math operator
+\begin{equation}
+\boldsymbol{\mu}(p) =
+\mathrm{median}
+\int_{\mathbb{R}} \mathbf{S}\boldsymbol{x}_n(t, p) \;\mathrm{d}t
+\end{equation}
 

### 2020-05-19 paper/paper.tex
write on the median scattering

@@
 \startlocaldefs
+
+\DeclareMathOperator*{\median}{median}
 
@@
 
-% TODO declare new math operator
+Given a collection of $N$ such waveforms, we number them arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
+Then, for every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we measure the median value taken by the average response of $\mathbf{S}\boldsymbol{x}_n$ across all samples $n$ from $1$ to $N$:
 \begin{equation}
 \boldsymbol{\mu}(p) =
-\mathrm{median}
-\int_{\mathbb{R}} \mathbf{S}\boldsymbol{x}_n(t, p) \;\mathrm{d}t
+\median_{1\leq n \leq N}
+\int_{\mathbb{R}} \mathbf{S}\boldsymbol{x}_n(t, p) \;\mathrm{d}t.
+\end{equation}
+
+\begin{equation}
+\mathbf{\widetilde{S}}\boldsymbol{x}_n (p) =
+\log \left(
+1 +
+\dfrac{
+\int_\mathbb{R} \mathbf{S}\boldsymbol{x}_n (t, p) \;\mathrm{d}t
+}{\varepsilon \times \boldsymbol{\mu}(p)}
+\right)
 \end{equation}

### 2020-05-19 paper/paper.tex
finish writing on compression

@@
 \end{equation}
-
+If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only, and re-apply it as a constant value on the test set.
+
+Then, to match a decibel-like perception of loudness, we apply the following adaptive transformation, which composes a median-based renormalization and a logarithmic compression:
 \begin{equation}
@@
 \right)
-\end{equation}
+\label{eq:S_tilde}
+\end{equation}
+where $\varepsilon$ is a predefined constant.
+The offset of $1$ before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps.
+Furthermore, a previous publication has shown empirically that Equation \ref{eq:S_tilde} brings, on environmental audio recordings, the histogram of $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ closer to a statistical Gaussianization of every feature $p$ \cite{lostanlen2018jasmp}.
+Since then, this finding has been confirmed in the case of musical sounds \cite{lostanlen2018extended}.
+
+Lastly, we standardize every feature $\mathbf{\widetilde{S}}\boldsymbol{x}_n$ to null mean and unit variance, across all $N$ samples $\boldsymbol{x}_1 \ldots \boldsymbol{x}_N$, independently for each scattering path $p$.
+Again, in cases where the dataset is split between training and test sets, we measure means and variances over the test set and propagate them as constants to the test set.
+With a slight abuse of notation, we still denote by $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ the standardized log-scattering features at path $p$ for sample $n$, even though its value differs from Equation \ref{eq:S_tilde} by an affine transformation.
 

### 2020-05-19 paper/paper.tex
write more on lmnn

@@
 \DeclareMathOperator*{\median}{median}
+\DeclareMathOperator*{\card}{card}
 
@@
 The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
-It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations.
+It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations \cite{anden2012scattering}.
 The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
@@
 The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
-The \lnameref{sec:survey} section describes our computer--human interface for annotating auditory similarities across playing techniques.
+The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
 The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
@@
 Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
-Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer.
+Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer \cite{loureiro2004ismir}.
 We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
@@
 \section*{\nmu Data collection}
-\label{sec:survey}
+\label{sec:data-collection}
 
@@
 %The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
-%The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:survey} section.
+%The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:data-collection} section.
 %This warping, or reweighting, of the space is performed using LMNN.
@@
 
-\subsection*{Per-feature Gaussianization with median-based logarithmic compression}
-\label{sec:gaussianization}
 We adopt the multiindex notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
@@
 
-Given a collection of $N$ such waveforms, we number them arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
-Then, for every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we measure the median value taken by the average response of $\mathbf{S}\boldsymbol{x}_n$ across all samples $n$ from $1$ to $N$:
+An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x})$ can be bounded as follows:
+\begin{equation}
+\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
+\leq
+\kappa_{\mathbf{S}} \times
+\Vert \boldsymbol{x} \Vert^2 \times
+\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
+\label{eq:lipschitz}
+\end{equation}
+where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
+In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
+In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.
+
+Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
+In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
+Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.
+
+\subsection*{Median-based logarithmic compression and affine standardization}
+\label{sec:gaussianization}
+
+It stems from the above that comparing Euclidean distances in the feature space of time--frequency scattering coefficients may serve as a computationally efficient surrogate for comparing Wasserstein distances in the scalogram, without need for solving an optimal transport problem.
+However, this surrogate is only valid if the underlying transport plan $\boldsymbol{\tau}$ is relatively small  in terms of supremum norm (see Equation \ref{eq:lipschitz}).
+This assumption may be true for certain local deformation of the time-frequency domain, \eg{} causing a pitch shift or a change in vibrato rate while leaving the instrument unchanged \cite{anden2012scattering}.
+However, it does not suffice in order to represent timbre similarity across multiple instruments and playing techniques.
+Indeed, the perception of timbre for extended playing techniques often involves long-range interactions in the time--frequency domain, which are not amenable to a local elastic deformation \cite{lostanlen2018extended}.
+
+To address this issue, we apply a metric learning algorithm, named large-margin nearest neighbors (LMNN), whose goal is to reweight and recombine scattering paths $p$ such that certain factors of variability in feature space $\mathbf{S}$ are enhanced while others are discarded.
+Before training this algorithm, we apply a pointwise nonlinear transformation on averaged joint time-frequency scattering coefficients.
+The role of this transformation, which is adapted to the dataset in an unsupervised way, is to Gaussianize the histogram of activations of each scattering path $p$.
+
+We consider a collection $\mathcal{X}$ of $N$ waveforms, numbered them arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
+For every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we measure the median value taken by the average response of $\mathbf{S}\boldsymbol{x}_n$ across all samples $n$ from $1$ to $N$:
 \begin{equation}
@@
 
-Lastly, we standardize every feature $\mathbf{\widetilde{S}}\boldsymbol{x}_n$ to null mean and unit variance, across all $N$ samples $\boldsymbol{x}_1 \ldots \boldsymbol{x}_N$, independently for each scattering path $p$.
-Again, in cases where the dataset is split between training and test sets, we measure means and variances over the test set and propagate them as constants to the test set.
+Lastly, we standardize every feature $\mathbf{\widetilde{S}}\boldsymbol{x}_n$ to null mean and unit variance, across the dataset $\mathcal{X} = \{\boldsymbol{x}_1 \ldots \boldsymbol{x}_N\}$, independently for each scattering path $p$.
+Again, if $\mathcal{X}$ is split between training and test sets, we measure means and variances over the training set only and propagate them as constants to the test set.
 With a slight abuse of notation, we still denote by $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ the standardized log-scattering features at path $p$ for sample $n$, even though its value differs from Equation \ref{eq:S_tilde} by an affine transformation.
@@
 \subsection*{Metric learning with large-margin nearest neighbors (LMNN)}
-\label{sec:weighting}
-
-The standard Euclidean distance on the raw scattering coefficients provides an instrument-agnostic representation of timbre dissimilarity.
-We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
-
-Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
+Let $\boldsymbol{x}$ some arbitrary audio sample in the dataset $\mathcal{X}$.
+Let $\mathcal{G}$ a cluster graph with $N = \card \mathcal{X}$ vertices and $C$ clusters in total.
+We denote by $\mathcal{G}(\boldsymbol{x})$ the cluster index to which the sample $\boldsymbol{x}$ belongs.
+Given another sample $\boldsymbol{y}$ in $\mathcal{X}$, $\boldsymbol{y}$ is similar to $\boldsymbol{x}$ if and only if belongs to the cluster $\mathcal{G}(\boldsymbol{x})$.
+Because $\mathcal{G}$ is a disjoint union of complete graphs, this relation is symmetric: $\boldsymbol{x} \in \mathcal{G}(\boldsymbol{y})$ is equivalent to $\boldsymbol{y} \in \mathcal{G}(\boldsymbol{x})$.
+
+In our protocol, $\boldsymbol{x}$ contains the sound of an isolated musical note and the cluster graph $\mathcal{G}$ encodes auditory similarities within the dataset $\mathcal{X}$.
+Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K=31$ subjects from our perceptual study (see \lnameref{sec:data-collection} section).
+%write on multiple subjects?
+
+%Furthermore, let us denote by $\mathcal{Y}_1 (\boldsymbol{x}) = \mathcal{X} \backslash \lbrace \boldsymbol{x} \rbrace$ the dataset of $(N-1)$ audio samples obtained from $\mathcal{X}$ after subtracting the singleton $\{\boldsymbol{x}\}$.
+
+We denote by $\mathbf{\widetilde{S}}\boldsymbol{x}$ the feature vector of joint time--frequency scattering resulting from $\boldsymbol{x}$.
+This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsections above).
+Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients, according to the Euclidean metric.
+Unlike cluster similarity, this relationship is not symmetric: $\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})$ does not necessarily imply $\boldsymbol{x}\in\mathcal{Y}_R (\boldsymbol{y})$.
+Note that the dependency of $\mathcal{Y}_R (\boldsymbol{x})$ upon the operator $\mathbf{S}$ is left implicit.
+In all of the following, we set the constant $R$ to $5$; this is in accordance with our chosen evaluation metric, i.e., average precision at rank 5 (AP@5, see \lnameref{sec:results} section).
+
+Let $P$ be the number of scattering paths in the operator $\mathbf{S}$.
+The large-margin nearest neighbor (LMNN) algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
+\begin{equation}
+\mathcal{E}(\mathbf{L}) = \frac{1}{2} \mathcal{E}_{\textrm{pull}} (\mathbf{L}) + \frac{1}{2} \mathcal{E}_{\textrm{push}} (\mathbf{L})
+\end{equation}
+Intuitively, the term $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space.
+Crucially, in the case of time--frequency scattering coefficients, these ``Euclidean neighborhoods'' corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
+The definition of $\mathcal{E}_{\textrm{pull}}$ is:
+\begin{equation}
+\mathcal{E}_{\textrm{pull}} (\mathbf{L}) =
+\sum_{\boldsymbol{x}\in\mathcal{X}}
+\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
+\big\Vert
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+\big\Vert^2,
+\end{equation}
+Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
+for pairs of samples belonging to the same cluster, while the term $\mathcal{E}_{\textrm{push}}$ tends to ensure that the distance between
+
+\begin{equation}
+\mathcal{E}_{\textrm{push}} (\mathbf{L}) =
+\sum_{\boldsymbol{x}\in\mathcal{X}}
+\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
+\sum_{\substack{\boldsymbol{z}\in\mathcal{Y}_R (\boldsymbol{x}) \\ \boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}}
+\rho
+\Big(
+1 +
+\big\Vert
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+\big\Vert^2
+-
+\big\Vert
+\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}
+\big\Vert^2
+\Big),
+\end{equation}
+where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU).
+
+%\begin{equation}
+%\mathbf{\Phi}_1 (\boldsymbol{x}) =
+%\arg \min_{\boldsymbol{y}\in \mathcal{Y}_1(\boldsymbol{x})}
+%\;
+%\big\Vert
+%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
+%-
+%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+%\big\Vert
+%\end{equation}
+% reference triplet loss audioset
+
+%The standard Euclidean distance on the raw scattering coefficients provides an instrument-agnostic representation of timbre dissimilarity.
+%We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
+
+%Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
 More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
 
-Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
-We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
-
-We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
-For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
-This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
+%Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
+%We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
+
+%We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
+%For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
+%This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
 % TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
-
 %We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
 
-$\mathcal{Y}_1 (\boldsymbol{x}) = \mathcal{Y} \backslash \lbrace \boldsymbol{x} \rbrace$.
-
-\begin{equation}
-\mathbf{\Phi}_1 (\boldsymbol{x}) =
-\arg \min_{\boldsymbol{y}\in \mathcal{Y}_1(\boldsymbol{x})}
-\;
-\big\Vert
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
--
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-\big\Vert
-\end{equation}
-
-$\mathcal{Y}_{r} (\boldsymbol{x}) = \mathcal{Y}_{(r-1)}(\boldsymbol{x}) \backslash \lbrace \mathbf{\Phi}_r (\boldsymbol{x}) \rbrace$.
-
-\begin{equation}
-\mathbf{\Phi}_r (\boldsymbol{x}) =
-\arg \min_{\boldsymbol{y}\in \mathcal{Y}_r(\boldsymbol{x})}
-\;
-\big\Vert
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
--
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-\big\Vert
-\end{equation}
+%$\mathcal{Y}_{r} (\boldsymbol{x}) = \mathcal{Y}_{(r-1)}(\boldsymbol{x}) \backslash \lbrace \mathbf{\Phi}_r (\boldsymbol{x}) \rbrace$.
+
+%\begin{equation}
+%\mathbf{\Phi}_r (\boldsymbol{x}) =
+%\arg \min_{\boldsymbol{y}\in \mathcal{Y}_r(\boldsymbol{x})}
+%\;
+%\big\Vert
+%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
+%-
+%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
+%\big\Vert
+%\end{equation}
 
@@
 We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
+In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
 However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
@@
 
-Given an audio stimulus $n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}$.
+Given an audio stimulus $\boldsymbol{x}_n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
 Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
@@
 \item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended},
-\item Per-feature affine scaling so that the dataset has null mean and unit variance,
+\item Affine standardization so that each feature has null mean and unit variance,
 \item Nearest-neighbor search according to priorly learned non-Euclidean metric.
@@
 
-For composers to personalize their search engine $\mathbf{\Phi}$, the quid pro quo is to define manually a cluster graph among $N=78$ audio samples, by means of the Cyberlioz annotation interface (see \lnameref{sec:survey} section).
+For composers to personalize their search engine $\mathbf{\Phi}$, the quid pro quo is to define manually a cluster graph among $N=78$ audio samples, by means of the Cyberlioz annotation interface (see \lnameref{sec:data-collection} section).
 This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
@@
 
-Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:survey} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
+Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
 That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.

### 2020-05-19 paper/paper.tex
cite jansen 2018 icassp

@@
 Intuitively, the term $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space.
-Crucially, in the case of time--frequency scattering coefficients, these ``Euclidean neighborhoods'' corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
+Crucially, in the case of time--frequency scattering coefficients, these Euclidean neighborhoods corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
 The definition of $\mathcal{E}_{\textrm{pull}}$ is:
@@
 Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
-for pairs of samples belonging to the same cluster, while the term $\mathcal{E}_{\textrm{push}}$ tends to ensure that the distance between
-
+
+Conversely, the term $\mathcal{E}_{\textrm{push}}$ tends to penalize small distances between samples that belong to different clusters in $\mathcal{G}$.
+More precisely, consider three samples $\boldsymbol{x}$, $\boldsymbol{y}$ and $\boldsymbol{z}$ in $\mathcal{X}$.
+Assume that the former belongs to the Euclidean neighborhood of $\boldsymbol{x}$ (i.e., $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$) while the latter belongs to a different cluster as $\boldsymbol{x}$ (i.e., $\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x}$)).
+The term $\mathcal{E}_{\textrm{push}}$ penalizes $\mathbf{L}$ unless the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}$ is smaller than the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}$ by a margin of at least $1$.
+The definition of $\mathcal{E}_{\textrm{push}}$ is:
 \begin{equation}
@@
 \sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
-\sum_{\substack{\boldsymbol{z}\in\mathcal{Y}_R (\boldsymbol{x}) \\ \boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}}
+\sum_{\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}
 \rho
@@
 \end{equation}
-where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU).
+where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU), also known as hinge loss.
+The cost function described in the equation above is known in deep learning as ``triplet loss'', and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
 

### 2020-05-21 paper/paper.tex
ml reads up to methods

@@
 
-The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain.
+The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
 Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
 The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
-Rather than to merely discriminate instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
+Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
 
@@
 The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
-The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/paperPlayingTechniquesSimilarityLearning}} %TODO: simplify this URL to lostanlen2020jasmp
+The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
 
@@
 The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
+\ml{preciser le contexte, un joueur de tabla ne serait pas d'accord, j'enleverai ces 2  phrases, ...
+and refer to ...}
 Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
@@
 The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
-The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a dense sampling of this combinatorial explosion, amounting to 25444 audio signals in total.
+The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a dense sampling of this \ml{repetition : combinatorial explosion, what would be the complete, give the percent achieved of SOL}, amounting to 25444 audio signals in total.
 We should note that this dense sampling is by no means comprehensive, as it erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
@@
 Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
-Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and efficient annotation interface.
+Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and the use of an efficient annotation interface.
 
@@
 In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
-Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments.
-
-Once both experts completed their annotation, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
+Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments \ml{available in the database}.
+
+Once both experts completed their annotation, we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
 At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
@@
 In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\times N$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
-In X 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
-We publicized this study on international mailing lists for research in music audio processing, such as X and Y. % TODO give mailing lists
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender? give age stats? musical experience?
+In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
+We publicized this study on \ml{internal student mailing list of the cnsmdp} international mailing lists for research in music audio processing, such as Auditory and Music-IR.
+% TODO give mailing lists
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender? give age stats? musical experience? did not kept this info sorry
 %Figure \ref{fig:xp2display}. % TODO

### 2020-05-21 paper/paper.tex
ml reading done

@@
 
+\ml{c'est bien ecrit, mais peut etre un peu trop detaillé, a voir avec Joakim}
+
 \subsection*{Median-based logarithmic compression and affine standardization}
@@
 We denote by $\mathbf{\widetilde{S}}\boldsymbol{x}$ the feature vector of joint time--frequency scattering resulting from $\boldsymbol{x}$.
-This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsections above).
+This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsectio\ml{s} above).
 Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients, according to the Euclidean metric.
@@
 In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
-However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
+However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
 
@@
 In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bambooo flute (\emph{dizi}).
+\ml{limiter les self citations a une occurence}
 However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
@@
 Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
-That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
+\ml{a deplacer ou a enlever : That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
 One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
 However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
-Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.
+Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
 
@@
 Thus, it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
+
+\ml{as the learned layer is relatively thin}
 
@@
 Because the runtime complexity of nearest-neighbor search grows quadratically with the dimensionality, i.e., is $O(d^2)$, replacing LMNN by LDA accelerates the system by a factor $(d/d_0)^2 \approx 3 \cdot 10^3$.
+
+\ml{I do not understand the coplexity argument here, not using anything is better, right ? Start with nothing then show lda, explain by reduced dimensionality of the lda output}
 
@@
 Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
-Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat/}}.
+Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}.
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
@@
 
-We see from the ablation study conducted above that each step of the proposed model is necessary for its high performance.
+We see from the ablation study conducted above that each step of the proposed model is necessary for its satisfactory performance.
 Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
@@
 
-\ml{je reecrirai le par du dessus en positif future work (TODO VL).}
-
 That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
-Indeed, the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed and work for more general classes of sounds.
+Indeed, \ml{we have strong assumptions} that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
+\ml{Though, future work should demonstrate that they} work for more general classes of sounds.
 To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
@@
 An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assigments.
+
+\ml{recaler ici la out of database query}
 

### 2020-05-22 paper/paper.tex
proofread run

--- 2020-05-19 paper/paper.tex
@@
 Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted orchestration \cite{maresz2013cmr}.
-In such a software environment, the composer queries the machine with an arbitrary audio signal; the outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
+In such context, the composer queries the software with an arbitrary audio signal; the outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
 Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range \cite{caetano2019swarm}.
@@
 The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
-It encodes patterns of spectrotemporal modulation in the query while offering numerical guarantees of stability to local deformations \cite{anden2012scattering}.
+It encodes patterns of spectrotemporal modulation in the acoustic query while offering numerical guarantees of stability to local deformations \cite{anden2012scattering}.
 The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
-It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from a collective consensus.
+It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from an intersubjective consensus.
 
@@
 Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
-Although scraping user-generated content from online music hosting platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
+Although scraping user-generated content from online music platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
 
@@
 Rather, each excerpt must be played back to multiple human listeners.
-Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, restricting the size of the musical corpus under study.
-Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in deep neural networks.
+Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, therefore restricting the size of the musical corpus under study.
+Such restriction hampers the applicability of iterative algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
 Nevertheless, while training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
@@
 Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
-Another appeal of the full cortical model is that the third-order $(\lambda, \alpha, \beta)$ tensor of frequencies, rates, and scales can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
+Another appeal of the full cortical model is that the three-way tensor of frequency $\lambda$, rate $\alpha$, and scale $\beta$ can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
 This is unlike fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for feature inversion \cite{mishra2018ismir}.
@@
 The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
-Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
+Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey), these are rarely shared with pitched instruments.
 In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
@@
 
-Once both experts completed their annotation, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
+Once both experts completed their annotations, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
 At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
@@
 In the case of $N=78$ IMTs, the task would amount to about 3k horizontal sliders, \ie{} several hours of cumbersome work for the human annotator.
+
 Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of continuous ratings.
@@
 \subsection*{Hypergraph partitioning}
-Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a similarity graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
+Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a cluster graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
 For a given $k$, the graph $\mathcal{G}_k$ contains $N=78$ vertices, each representing a different IMT in the corpus.
@@
 Otherwise, there is no edge connecting $m$ and $n$.
-This way, $\mathcal{G}_k$ contains as many connected components as the number of similarity clusters for the subject $k$, \ie{} the number of distinct colors on the Cyberlioz interface in the response of $k$.
+Thus, $\mathcal{G}_k$ contains as many connected components as the number of similarity clusters for the subject $k$, \ie{} the number of distinct colors on the Cyberlioz interface in the response of $k$.
 
@@
 Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_k$'s among all subject $k$.
-Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient,han1997scalable,strehl2002cluster}.
+Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient}.
 
@@
 Then, we run hypergraph partitioning on $\mathcal{H}$ to assign each vertex $i$ to one of the $C_0$ clusters in $\mathcal{G}_0$.
-Intuitively, hypergraph partitioning optimizes a tradeoff between two objectives: first, balancing the size of all clusters in terms of their respective numbers of vertices; and secondly, keeping most hyperedges enclosed within as few distinct clusters as possible.
+Intuitively, hypergraph partitioning optimizes a tradeoff between two objectives: first, balancing the size of all clusters in terms of their respective numbers of vertices; and secondly, keeping most hyperedges enclosed within as few distinct clusters as possible \cite{han1997scalable,strehl2002cluster}.
 
@@
 \begin{array}{ll}
-\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha\neq0,\\
+\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha=0,\\
 \mathbf{S_{2}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{otherwise.}
@@
 
-We consider a collection $\mathcal{X}$ of $N$ waveforms, numbered them arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
+We consider a collection $\mathcal{X}$ of $N$ waveforms, numbered arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
 For every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we measure the median value taken by the average response of $\mathbf{S}\boldsymbol{x}_n$ across all samples $n$ from $1$ to $N$:
@@
 \end{equation}
-Intuitively, the term $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space.
-Crucially, in the case of time--frequency scattering coefficients, these Euclidean neighborhoods corresponds to Wasserstein neighborhoods in the scalogram domain (see previous subsection).
+where, intuitively, $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space while $\mathcal{E}_{\textrm{push}}$ tends to penalize small distances between samples that belong to different clusters in $\mathcal{G}$.
+
 The definition of $\mathcal{E}_{\textrm{pull}}$ is:
@@
 \end{equation}
+Crucially, in the feature space of time--frequency scattering coefficients, the Euclidean neighborhood $\mathcal{Y}_R (\boldsymbol{x})$ corresponds to a Wasserstein neighborhood in the scalogram domain (see previous subsection).
 Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
 
-Conversely, the term $\mathcal{E}_{\textrm{push}}$ tends to penalize small distances between samples that belong to different clusters in $\mathcal{G}$.
-More precisely, consider three samples $\boldsymbol{x}$, $\boldsymbol{y}$ and $\boldsymbol{z}$ in $\mathcal{X}$.
-Assume that the former belongs to the Euclidean neighborhood of $\boldsymbol{x}$ (i.e., $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$) while the latter belongs to a different cluster as $\boldsymbol{x}$ (i.e., $\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x}$)).
-The term $\mathcal{E}_{\textrm{push}}$ penalizes $\mathbf{L}$ unless the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}$ is smaller than the distance between $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}$ and $\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}$ by a margin of at least $1$.
+While the the term $\mathcal{E}_{\textrm{pull}}$ operates on pairs of samples, the term $\mathcal{E}_{\textrm{push}}$, operates on triplets $(\boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z})\in\mathcal{X}^3$.
+The first sample, $\boldsymbol{x}$, is known as an ``anchor''.
+The second sample, $\boldsymbol{y}$, is known as a ``positive'', and is assumed to belong to the Euclidean neighborhood of the anchor: $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$.
+The third sample, $\boldsymbol{z}$, is known as a ``negative'', and is assumed to belong to a different similarity cluster as the anchor: $\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})$.
+The term $\mathcal{E}_{\textrm{push}}$ penalizes $\mathbf{L}$ unless the positive-to-anchor distance is smaller than the negative-to-anchor distance by a margin of at least $1$.
+
 The definition of $\mathcal{E}_{\textrm{push}}$ is:
@@
 In our case, the implicit similarity kernel is a homogeneous quadratic kernel.
-Despite this increase in representational power, we obtain a P@5 of $81.5\% \pm 7.3$, i.e. essentially the same as MFCC under a linear kernel.
+Despite this increase in representational power, we obtain a P@5 of $81.5\% \pm 7$, i.e. essentially the same as MFCC under a linear kernel.
 Therefore, it appears that the gap in performance between MFCC ($81.8\%\pm 7$) and joint time--frequency scattering ($99.0\% \pm 1$) is primarily attributable to the multiscale extraction of joint spectrotemporal modulations, whereas high-dimensional embedding likely plays a minor role.

### 2020-05-22 paper/paper.tex

@@
 The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
-Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey), these are rarely shared with pitched instruments.
+\ml{preciser le contexte, un joueur de tabla ne serait pas d'accord, j'enleverai ces 2  phrases, ...
+and refer to ...}
+Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
 In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
@@
 In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
-Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments.
-
-Once both experts completed their annotations, we retained all IMTs whose average relevance score was judged equal to 3 or higher.
+Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments \ml{available in the database}.
+
+Once both experts completed their annotation, we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
 At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.

### 2020-05-25 paper/paper.tex
shorten introduction of data collection (about perc)

@@
 In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
-The reason behind this choice is that most percussive instruments do not allow much variation in terms of pitch or intensity, if at all; thus reducing the amount of intra-technique variability in the time--frequency domain.
-\ml{preciser le contexte, un joueur de tabla ne serait pas d'accord, j'enleverai ces 2  phrases, ...
-and refer to ...}
-Furthermore, although percussive instruments certainly exhibit a rich diversity of playing techniques (we refer to \cite{peinkofer1976book} for a survey) these are rarely shared with pitched instruments.
-In comparison, techniques for sustained tones such as \emph{vibrato} or \emph{tremolo} permeate throughout a vast instrumentarium, \eg{} bowed strings and reed alike.
 We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.

### 2020-05-25 paper/paper.tex
commentaires
remover overset t in \ast
commit ML's conclusion

@@
 
+% Le numerique permet d'approcher la question de la notation de facon implicite
+% Cela ne retire pas l'interet de l'etude des graphemes ou du texte (glissando vs slide)
 The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
@@
 \big\vert
-\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
+\boldsymbol{x} \ast \boldsymbol{\psi}_{\lambda}
 \big\vert
@@
 That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
-Indeed, \ml{we have strong assumptions} that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
-\ml{Though, future work should demonstrate that they} work for more general classes of sounds.
+Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
+Though, future work should demonstrate that they work for more general classes of sounds.
 To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.

### 2020-05-30 paper/paper.tex
todo's
add musicography package

@@
 \usepackage{hyperref}
+\usepackage{musicography}
 \usepackage{relsize}
@@
 % Cela ne retire pas l'interet de l'etude des graphemes ou du texte (glissando vs slide)
+
 The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
-The authors of a recent publication \cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features \cite{anden2014tsp}, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
-Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be seen whether these findings can scale up from a single instrument to an entire orchestra.
-In addition, previous publications in time--frequency scattering, whether separable or joint, lack a human-centric evaluation, independently from any classification task.
+However, it remains to be seen whether joint time--frequency scattering is capable of fine-grained auditory categorization, involving variability in instrument, mute, and playing technique.
+In addition, previous publications on joint time--frequency scattering lack a human-centric evaluation, independently from any classification task.
 
@@
 However, the resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
-For example, the pitch of pedal tones in brass instruments is tied to the fundamental mode of the bore, \ie{} usually B-flat or F. % TODO check this, use special package for musical pitches
+For example, the pitch of pedal tones in brass instruments is tied to the fundamental mode of the bore, \ie{} usually $\mathsf{B}\musFlat$ or $\mathsf{F}$.
 Likewise, the intensity of key clicks in the oboe is necessarily \emph{pp}, while the intensity of snap pizzicato \emph{\`a la Bart\'ok} in plucked strings is necessarily \emph{ff}.
@@
 In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
-The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
-The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a dense sampling of this \ml{repetition : combinatorial explosion, what would be the complete, give the percent achieved of SOL}, amounting to 25444 audio signals in total.
-We should note that this dense sampling is by no means comprehensive, as it erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
+The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a joint sampling of these variables.
+The version of SOL that we use throughout this paper, named ``0.9 HQ'', amounts to 25444 audio signals in total.
+Beyond playing techniques, we should note that SOL erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
 Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.
@@
 Indeed, under first approximation, the perception of timbre is invariant to pitch and intensity.
-Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C (\ie{} C4) and \emph{mf}. % TODO use pitch spelling package
+Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C (\ie{} $\mathsf{C_4}$) and \emph{mf}.
 After this selection, every IMT triplet contains a single acoustic exemplar, regarded as canonical in the following.
@@
 One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in strings instruments.
-To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP), and explicitly gave them our definition of relevance.
+To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatoire (CNSMDP), and explicitly gave them our definition of relevance.
 Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
 In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
-Conversely, the value 7 (most relevant) denotes that the IMT under consideration has a strong similarity with IMTs from one or several other instruments \ml{available in the database}.
-
-Once both experts completed their annotation, we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
+Conversely, the value 7 (most relevant) denotes that the IMT under consideration bears a strong similarity with some other IMT from the corpus.
+
+Once both experts completed their annotation, we averaged their scores
+we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
 At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
@@
 Cyberlioz consists of a square panel on which is displayed a collection of circular grey dots, each of them corresponding to one of the $N=78$ IMTs, and initially distributed uniformly at random.
-Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} C4 and \emph{mf} in most cases. % TODO VL: use pitch notation library to express C4
+Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} $\textsf{C}_4$ and \emph{mf} in most cases.
 Furthermore, each dot can be freely placed on the screen by clicking, dragging, and dropping.
@@
 In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\times N$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
-In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones. % TODO give the exact two months.
-We publicized this study on \ml{internal student mailing list of the cnsmdp} international mailing lists for research in music audio processing, such as Auditory and Music-IR.
-% TODO give mailing lists
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task. % TODO break down by gender? give age stats? musical experience? did not kept this info sorry
-%Figure \ref{fig:xp2display}. % TODO
+In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
+To this end, we publicized this study on the internal mailing list of the Paris Conservatoire (CNSMDP) students, as well as two international mailing lists for research in music audio processing: ``AUDITORY''  and ``ISMIR Community''\footnote{
+For more information about these mailing lists, please visit:
+\url{http://www.auditory.org/} and \url{http://ismir.net/}}.
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.%Figure \ref{fig:xp2display}. % TODO
 
@@
 %We're no longer talking to biologists, so we don't need to be afraid.}
-% TODO: avoid citing the same papers over and over
 
@@
 We denote by $\mathbf{\widetilde{S}}\boldsymbol{x}$ the feature vector of joint time--frequency scattering resulting from $\boldsymbol{x}$.
-This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsectio\ml{s} above).
+This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsections above).
 Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients, according to the Euclidean metric.
@@
 The cost function described in the equation above is known in deep learning as ``triplet loss'', and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
+We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
 
@@
 %This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-% TODO: this section needs more detail. In particular, it is worth discussing the role of Euclidean distances as an "initial guess" for LMNN. (this is unlike MLR, for example).
 %We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
@@
 This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
-In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bambooo flute (\emph{dizi}).
+In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bamboo flute (\emph{dizi}).
 \ml{limiter les self citations a une occurence}
@@
 Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
+%TODO move to conclusion
 
@@
 
-\ml{as the learned layer is relatively thin}
+%\ml{as the learned layer is relatively thin}
 
@@
 \ml{I do not understand the coplexity argument here, not using anything is better, right ? Start with nothing then show lda, explain by reduced dimensionality of the lda output}
+% Move the part about complexity to the conclusion. Runtime complexity. How to scale this?
 

### 2020-06-01 paper/paper.tex
number equations A_T and A_I
move 1 par to conclusion

@@
 \url{http://www.auditory.org/} and \url{http://ismir.net/}}.
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.%Figure \ref{fig:xp2display}. % TODO
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.%Figure \ref{fig:xp2display}.
 
@@
 For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
-\[
+\begin{equation}
 \mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
@@
 \label{eq:instrument-similarity}
-\]
+\end{equation}
 
@@
 We define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinals:
-\[
+\begin{equation}
 \mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
@@
 },
-\]
+\end{equation}
 alike the definition of $\mathbf{A}_{\mathcal{I}}$.
@@
 This is the case, for example, of Orchidea\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}}, which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
-
 Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
-\ml{a deplacer ou a enlever : That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
-One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
-However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
-Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
-%TODO move to conclusion
 
@@
 \label{sec:conclusion}
+
+%\ml{a deplacer ou a enlever : That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
+%One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
+%However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
+%Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
+%TODO move to conclusion
 

### 2020-06-01 paper/paper.tex
caption inter-technique similarity

@@
 },
+\label{eq:technique-similarity}
 \end{equation}
@@
 \caption{
-%TODO caption
+Matrix $\mathbf{\widetilde{A}}_{\mathcal{I}}$ of perceived timbre similarities between instruments from our dataset of $N=78$. samples.
+Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:instrument-similarity}), obtained from a consensus of $K=31$ subjects.
+The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ were re-arranged according to Ward's minimum variance method.
+Observe that the block diagonal structure in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ reflects an organological taxonomy of musical instruments: woodwinds, brass, and strings.
+See \lnameref{sec:data-collection} section for details, including the meaning of each instrument abbreviation.
 }
@@
 \includegraphics[width = \textwidth]{figures/consensusVsPt.png}
-\caption{A description of the playing technique abbreviations in made available in the Appendix (Tables \ref{tab:list1} and \ref{tab:list2}).
-%TODO add a double arrow
+\caption{
+Matrix $\mathbf{\widetilde{A}}_{\mathcal{T}}$ of perceived timbre similarities between playing techniques from our dataset of $N=78$ samples.
+Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:technique-similarity}), obtained from a consensus of $K=31$ subjects.
+The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{T}}$ were re-arranged according to Ward's minimum variance method.
+See \lnameref{sec:data-collection} section for details,.
+See Tables \ref{tab:list1} and \ref{tab:list2} for an English description of each playing technique abbreviation.
+%TODO add a double arrow?
 % rough <-- soft --> rough

### 2020-06-05 paper/paper.tex
upload violin and flute figures

@@
 \usepackage{musicography}
+\usepackage{pdfpages}
 \usepackage{relsize}
@@
 \begin{figure}
-\caption{\csentence{Sample scattering}
-Same 3 sounds spectrogram + scattering
-% TODO figure + caption
-}
-\label{fig:sample-scattering}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-ord-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-tasto-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-trem-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-stacc-C4-mf_scalogram.pdf}
+
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-ord-C4-mf-4c_scattering-compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-tasto-C4-mf-4c_scattering-compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-trem-C4-mf-4c_scattering-compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/violin_scattering/Vn-stacc-C4-mf_scattering-compressed.pdf}
+\caption{\csentence{Violin scattering}}
+\label{fig:violin-scattering}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-ord-C4-mf_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-sfz-C4-fp_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scalogram.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-key-cl-C4-f_scalogram.pdf}
+
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-ord-C4-mf_scattering_compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-sfz-C4-fp_scattering_compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scattering_compressed.pdf}
+\includegraphics[width=0.23\textwidth]{figures/flute_scattering/Fl-key-cl-C4-f_scattering_compressed.pdf}
+\caption{\csentence{Flute scattering}}
+\label{fig:flute-scattering}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-ord-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-tasto-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-trem-C4-mf-4c_scalogram.pdf}
+
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-ord-C4-mf-4c_scattering-compressed.pdf}
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-tasto-C4-mf-4c_scattering-compressed.pdf}
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-trem-C4-mf-4c_scattering-compressed.pdf}
+
+\vspace{1cm}
+
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-stacc-C4-mf_scalogram.pdf}
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scalogram.pdf}
+\includegraphics[width=0.3\textwidth]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scalogram.pdf}
+
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-stacc-C4-mf_scattering-compressed.pdf}
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scattering-compressed.pdf}
+\hspace{0.05\textwidth}
+\includegraphics[width=0.25\textwidth]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scattering-compressed.pdf}
+
+\caption{\csentence{Violin scattering}}
+\label{fig:violin-scattering}
 \end{figure}

### 2020-06-05 paper/paper.tex
caption for violin figure

@@
 
+\csentence{(a)}\hspace{0.28\textwidth}
+\csentence{(b)}\hspace{0.28\textwidth}
+\csentence{(c)}
 \vspace{1cm}
@@
 
-\caption{\csentence{Violin scattering}}
+\csentence{(d)}\hspace{0.28\textwidth}
+\csentence{(e)}\hspace{0.28\textwidth}
+\csentence{(f)}
+
+\caption{\csentence{Six playing techniques of the violin.}
+Subfigures:
+(a) ordinario, (b) sul tasto, (c) tremolo, (d) staccato, (e) nonvibrato, and (f) sforzando.
+In each subfigure, the top image shows the wavelet scalogram, as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
+Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
+Darker shades denote greater values of acoustic energy.
+Note that the ordinary technique (a) contains vibrato.}
 \label{fig:violin-scattering}

### 2020-06-06 paper/paper.tex
finish figure 4

--- 2020-06-01 paper/paper.tex
@@
 \begin{figure}
-\caption{\csentence{Sample scattering}
-Same 3 sounds spectrogram + scattering
-% TODO figure + caption
-}
-\label{fig:sample-scattering}
+\includegraphics[height=4cm]{figures/violin_scattering/Vn-ord-C4-mf-4c_scalogram.pdf}
+\hspace{8mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scalogram.pdf}
+\hspace{12mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-trem-C4-mf-4c_scalogram.pdf}
+
+\hspace{2mm}
+\includegraphics[height=28mm]{figures/violin_scattering/Vn-ord-C4-mf-4c_scattering.pdf}
+\hspace{5mm}
+\includegraphics[height=28mm]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scattering.pdf}
+\hspace{5mm}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/violin_scattering/Vn-trem-C4-mf-4c_scattering.pdf}
+
+\csentence{(a) ordinario (with vibrato)}\hspace{18mm}
+\csentence{(b) nonvibrato}\hspace{24mm}
+\csentence{(c) tremolo}\hspace{8mm}
+\vspace{1cm}
+
+\includegraphics[height=4cm]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scalogram.pdf}
+\hspace{8mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-pizz-lv-C4-pp-4c_scalogram.pdf}
+\hspace{12mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-stacc-C4-mf_scalogram.pdf}
+
+\hspace{2mm}
+\includegraphics[height=28mm]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scattering.pdf}
+\hspace{5mm}
+\includegraphics[height=28mm]{figures/violin_scattering/Vn-pizz-lv-C4-pp-4c_scattering.pdf}
+\hspace{5mm}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/violin_scattering/Vn-stacc-C4-mf_scattering.pdf}
+
+\hspace{2mm}
+\csentence{(d) sforzando}\hspace{25mm}
+\csentence{(e) pizzicato}\hspace{25mm}
+\csentence{(f) staccato}\hspace{1mm}
+
+\caption{\csentence{Six playing techniques of the violin.}
+Subfigures:
+(a) ordinario (with vibrato), (b) nonvibrato, (c) tremolo, (d) sforzando, (e) pizzicato (\emph{laissez vibrer}, i.e., let ring), and (f) staccato.
+In each subfigure, the top image shows the wavelet scalogram, as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
+Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
+Darker shades denote greater values of acoustic energy.}
+\label{fig:violin-scattering}
 \end{figure}

### 2020-06-06 paper/paper.tex
finish figure 5

@@
 (a) ordinario (with vibrato), (b) nonvibrato, (c) tremolo, (d) sforzando, (e) pizzicato (\emph{laissez vibrer}, i.e., let ring), and (f) staccato.
-In each subfigure, the top image shows the wavelet scalogram, as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
+In each subfigure, the top image shows the wavelet scalogram as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
 Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
@@
 \label{fig:violin-scattering}
+\end{figure}
+
+\begin{figure}
+\includegraphics[height=4cm]{figures/flute_scattering/Fl-ord-C4-mf_scalogram.pdf}
+\hspace{8mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scalogram.pdf}
+\hspace{12mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-ply+sng-Csharp4+C4-mf_scalogram.pdf}
+
+\hspace{2mm}
+\includegraphics[height=28mm]{figures/flute_scattering/Fl-ord-C4-mf_scattering.pdf}
+\hspace{5mm}
+\includegraphics[height=28mm]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scattering.pdf}
+\hspace{5mm}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/flute_scattering/Fl-ply+sng-Csharp4+C4-mf_scattering.pdf}
+
+\hspace{12mm}
+\csentence{(a) ordinario}\hspace{22mm}
+\csentence{(b) trill $\mathsf{C_4} - \mathsf{D_4}$}\hspace{15mm}
+\csentence{(c) play $\mathsf{C\musSharp_4}$ while singing $\mathsf{C_4}$}\vspace{1cm}
+
+\includegraphics[height=4cm]{figures/flute_scattering/Fl-sfz-C4-fp_scalogram.pdf}
+\hspace{8mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-key-cl-C4-f_scalogram.pdf}
+\hspace{12mm}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-stacc-C4-mf_scalogram.pdf}
+
+\hspace{2mm}
+\includegraphics[height=28mm]{figures/flute_scattering/Fl-sfz-C4-fp_scattering.pdf}
+\hspace{5mm}
+\includegraphics[height=28mm]{figures/flute_scattering/Fl-key-cl-C4-f_scattering.pdf}
+\hspace{5mm}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/flute_scattering/Fl-stacc-C4-mf_scattering.pdf}
+
+\hspace{2mm}
+\csentence{(d) sforzando}\hspace{25mm}
+\csentence{(e) key click}\hspace{25mm}
+\csentence{(f) staccato}\hspace{1mm}
+
+\caption{\csentence{Six playing techniques of the flute.}
+Subfigures:
+(a) ordinario, (b) trill ($\mathsf{C_4} - \mathsf{D_4}$), (c) interference (play $\mathsf{C\musSharp_4}$ while singing $\mathsf{C_4}$), (d) sforzando, (e) key click, (f) staccato.
+In each subfigure, the top image shows the wavelet scalogram as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
+Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
+Darker shades denote greater values of acoustic energy.}
+\label{fig:flute}
 \end{figure}

### 2020-06-06 paper/paper.tex
cite calvo 2020 acm

@@
 
-% Le numerique permet d'approcher la question de la notation de facon implicite
-% Cela ne retire pas l'interet de l'etude des graphemes ou du texte (glissando vs slide)
-
-The prospect of modeling timbre perception thus necessarily exceeds the symbolic domain. \ml{even if research about its notation is of interest}
+Although a graphemic study of playing techniques in music has research potential in music information retrieval \cite{calvo2020acm}, the prospect of modeling timbre perception necessarily exceeds the symbolic domain.
 Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
@@
 where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
-Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.
+Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques from the Chinese bamboo flute \cite{wang2019ismir}.
 
@@
 In particular, a previous publication has shown that frequency-dependent time shifts affect joint time--frequency scattering coefficients while leaving separable time--frequency scattering coefficients almost unchanged \cite{anden2019tsp}.
-For this reason, separable time--frequency scattering coefficients can only serve to recognize periodic modulations (e.g.
+%For this reason, separable time--frequency scattering coefficients can only serve to recognize periodic modulations (e.g.
 vibrato), but joint time--frequency scattering is required to recognize pitch evolution-based techniques (PETs, e.g.
 glissando) \cite{wang2020icassp}.
-The same observation was made by \cite{schadler2015jasa} in the case of separable Gabor filterbank features (GBFB), which bear some resemblance with separable time--frequency scattering.
+The same observation was made by \cite{schadler2015jasa} in the case of joint and separable Gabor filterbank features (GBFB), which bear some resemblance with joint and separable time--frequency scattering coefficients respectively.
 

### 2020-06-06 paper/paper.tex
clarify score averaging

@@
 
-Once both experts completed their annotation, we averaged their scores
-we retained all IMTs whose \ml{relevance score averaged over experts} was judged equal to 3 or higher.
-At the end of this stage of pre-screening, the list of IMT reduces from 235 to $N=78$.
-Tables 1 and 2, in the Appendix section of this paper, give a full list of these 78 highly relevant IMTs.
-It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic, regardless of playing technique, to be relevant for this experiment.
+Once both experts completed their annotation, we ranked all 235 IMTs by decreasing average relevance score.
+Then, we retained all IMTs whose average score was judged equal to 3 or higher, thus resulting in a shortlist of $N=78$ IMTs (see Tables 1 and 2 in the Appendix).
+It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic to be relevant for this experiment, regardless of playing technique.
 Indeed, the accordion is the only instrument in the aforementioned list of

### 2020-06-06 paper/paper.tex
incorporate ML's feedback

@@
 
-\ml{c'est bien ecrit, mais peut etre un peu trop detaillé, a voir avec Joakim}
+%ML : c'est bien ecrit, mais peut etre un peu trop detaillé, a voir avec Joakim
 
@@
 In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bamboo flute (\emph{dizi}).
-\ml{limiter les self citations a une occurence}
 However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.

### 2020-06-06 paper/paper.tex
bis

@@
 
-%\ml{as the learned layer is relatively thin}
+%ML: as the learned layer is relatively thin
 
@@
 Replacing the large-margin nearest neighbors (LMNN) metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
-However, we should note that, from a computational standpoint, the nearest-neighbor search $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ is considerably faster with LDA than with LMNN.
-Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $d_0 = (C_0+1)$, where $C_0 = 19$ is the number of clusters in the consensus cluster graph $\mathcal{G}_0$.
-Because the runtime complexity of nearest-neighbor search grows quadratically with the dimensionality, i.e., is $O(d^2)$, replacing LMNN by LDA accelerates the system by a factor $(d/d_0)^2 \approx 3 \cdot 10^3$.
-
-\ml{I do not understand the coplexity argument here, not using anything is better, right ? Start with nothing then show lda, explain by reduced dimensionality of the lda output}
-% Move the part about complexity to the conclusion. Runtime complexity. How to scale this?
-
 Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all.
@@
 
-%\ml{a deplacer ou a enlever : That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
-%One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
-%However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
-%Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
-%TODO move to conclusion
-
 We see from the ablation study conducted above that each step of the proposed model is necessary for its satisfactory performance.
@@
 
-\ml{recaler ici la out of database query}
+%ML: recaler ici la out of database query}
+%That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
+%One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
+%However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
+%Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
+
+% Move the part about complexity to the conclusion. Runtime complexity. How to scale this?
+%However, we should note that, from a computational standpoint, the nearest-neighbor search $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ is considerably faster with LDA than with LMNN.
+%Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $d_0 = (C_0+1)$, where $C_0 = 19$ is the number of clusters in the consensus cluster graph $\mathcal{G}_0$.
+%Because the runtime complexity of nearest-neighbor search grows quadratically with the dimensionality, i.e., is $O(d^2)$, replacing LMNN by LDA accelerates the system by a factor $(d/d_0)^2 \approx 3 \cdot 10^3$.
 

### 2020-06-08 experiments/processResults_v2.m
adding ablation study figure

@@ -1,5 +1,6 @@
+addpath('libs/export_fig')
 clear all
-captions = {'full', 'split', 'random', 'lda', 'nolearning', '25ms', 'separable', 'mfcc', 'monomials'}
+captions = {'Proposed', 'train/test validation', 'randomized features', 'lda', 'no learning', 'reduce T to 25ms', 'separable scattering', 'mfcc', 'mfcc monomials'};
 ind = 1;
 
 iaInd = 2;
@@ -18,7 +19,10 @@ for k=1:length(captions)
     ind = ind+1;
 end
 
-
+reorder = [9 8 1 7 6 5 4 3 2];
+[~, reorder] = sort(reorder);
+p=p(reorder, :);
+captions = captions(reorder);
 clf();
 
 hold on;
@@ -40,7 +44,12 @@ xticks(-log10(custom_xticks(end:-1:1)));
 xticklabels(100-100*custom_xticks(end:-1:1))
 set(gca(), 'XGrid', 'on');
 
+set(gca, 'fontsize', 16)
 
-axis([-log10(custom_xticks(end)) -log10(custom_xticks(1)) 0.5 8.5])
+axis([-log10(custom_xticks(end)) -log10(custom_xticks(1)) 1.5 9.5])
 hold off;
+
+set(gcf, 'Color', 'None')
+savefig('../paper/figures/ablationStudyNoRand.fig')
+export_fig('../paper/figures/ablationStudyNoRand.png', '-transparent')
 

### 2020-06-08 paper/paper.tex
adding ablation study figure

@@
 We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
+Results of this ablation study is summarized in Figure \ref{fig:ablation}, where performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.
+It appears that the ranking over testing conditions is not strongly modified,  which demonstrate that for all conditions the reference clusterings which are hard to model (depicted by blue losanges) are hard to model for each conditions whereas the reference clusterings which are easier to model (depicted by yellow losanges) are easier to model for each conditions.
 
@@
 
+\begin{figure}[h!]
+\includegraphics[width=0.9\textwidth]{figures/ablationStudy}
+\caption{\csentence{Impact of different processing architecture or protocol designs.
+For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.
+The whiskers extend to the most extreme data points not considered outliers.
+Performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.}}
+\label{fig:ablation}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-09 paper/paper.tex
write on pareto efficiency

--- 2020-06-06 paper/paper.tex
@@
 We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
+
+Figure \ref{fig:ablation} summarizes our results.
+Interestingly, the system $\boldsymbol{\Phi}$ is not only best on average, but also best for every subject in the cohort: altering any design choice in $\boldsymbol{\Phi}$ (see subsections below) results in a P@5 below $\mathrm{P}_{\mathbf{\Phi}}(k, R)$ for every $k$.
+Borrowing from the terminology of welfare economics, $\boldsymbol{\Phi}$ can be said to be uniquely Pareto-efficient.
+This observation suggests that the increase in performance afforded by the state-of-the-art model with respect to the baseline does not come at the detriment of user fairness.
+%Results of this ablation study is summarized in Figure , where performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.
+It appears that the ranking over testing conditions is not strongly modified,  which demonstrate that for all conditions the reference clusterings which are hard to model (depicted by blue losanges) are hard to model for each conditions whereas the reference clusterings which are easier to model (depicted by yellow losanges) are easier to model for each conditions.
 

### 2020-06-14 paper/paper.tex
acknowledgment section

@@
 
-\section*{\nmu Author's contributions}
-    Text for this section \ldots
-
-\section*{\nmu Acknowledgements}
+%\section*{\nmu Author's contributions}
+%    Text for this section \ldots
+
+\section*{\nmu Acknowledgement}
+This work is partially supported by the European Research Council (ERC) award 320959 (InvariantClass).
+This work is partially supported by the National Science Foundation (NSF) award 1633259 (BIRDVOX).
 The Flatiron Institute is a division of the Simons Foundation.
+We wish to thank Philippe Brandeis, \'{E}tienne Graindorge, St\'{e}phane Mallat, Adrien Mamou-Mani, and Yan Maresz for contributing to the TICEL research project.
+We also wish to thank the students of the Paris Conservatory and all anonymous participants to our study.
 

### 2020-06-17 paper/paper.tex
update fig1 in paper
update andreux2020jmlr

@@
 Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
-On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
+On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2020jmlr} and invertibility \cite{lostanlen2019dafx}.
 Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
@@
 \begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v3.pdf}
-\caption{\csentence{Overview of the proposed approach}}
+\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v4.pdf}
+\caption{\csentence{Overview of the proposed approach. See \ref{sec:intro} for details.}}
 \label{fig:pipeline}

### 2020-06-17 paper/paper.tex
refs from figures to text in captions

@@
 \includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v4.pdf}
-\caption{\csentence{Overview of the proposed approach. See \ref{sec:intro} for details.}}
+\caption{\csentence{Overview of the proposed approach.} See \lnameref{sec:intro} for details.}
 \label{fig:pipeline}
@@
 Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
-Darker shades denote greater values of acoustic energy.}
+Darker shades denote greater values of acoustic energy.
+See \lnameref{sec:methods} section for details.}
 \label{fig:violin-scattering}
@@
 Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
-Darker shades denote greater values of acoustic energy.}
+Darker shades denote greater values of acoustic energy.
+See \lnameref{sec:methods} section for details.}
 \label{fig:flute}
@@
 \includegraphics[width=0.9\textwidth]{figures/ablationStudy}
-\caption{\csentence{Impact of different processing architecture or protocol designs.
-For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.
+\caption{\csentence{Impact of different processing architecture or protocol designs.} For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.
 The whiskers extend to the most extreme data points not considered outliers.
-Performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.}}
+Performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.
+See \lnameref{sec:discussion} section for details.}
 \label{fig:ablation}

### 2020-07-03 paper/paper.tex
arrange figs 2 and 3 on a single page
start writing on visualization

@@
 
-%ML : c'est bien ecrit, mais peut etre un peu trop detaillé, a voir avec Joakim
+\subsection*{Visualization}
+\label{sec:visualization}
+
+At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
+From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
+To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
+In accordance with their protocol, we compute the following matrix:
+
+\begin{equation}
+\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
+\iint_{\mathbb{R}^2}
+\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
+\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
+\end{equation}
+
+Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
+As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
+In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2002speechcomm}.
+
+Previous publications on STRF have focused on the visualization of synthetic stimuli and speech. % refs?
+%Although one publication has turned to the visualization of isolated musical notes, the scope of discussion on the effect of technique is
+
+Figures \ref{fig:violin-scattering} and \ref{fig:flute-scattering}.
 
@@
 This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
-In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bamboo flute (\emph{dizi}).
+In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bamboo flute \emph{(dizi)}.
 However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
@@
 \center
-\includegraphics[width = \textwidth]{figures/consensusVsI.png}
+\includegraphics[width = 0.85\textwidth]{figures/consensusVsI.png}
 \caption{
@@
 \center
-\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
+\includegraphics[width = 0.85\textwidth]{figures/consensusVsPt.png}
 \caption{
@@
 See \lnameref{sec:methods} section for details.}
-\label{fig:flute}
+\label{fig:flute-scattering}
 \end{figure}

### 2020-07-03 paper/paper.tex
cite prior work on strf visualization

@@
 As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2002speechcomm}.
-
-Previous publications on STRF have focused on the visualization of synthetic stimuli and speech. % refs?
+In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
+
+Previous publications have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
 %Although one publication has turned to the visualization of isolated musical notes, the scope of discussion on the effect of technique is

### 2020-07-03 paper/paper.tex
state that the goal is to fit human similarity judgments
remove double usepackage
write more on scale-rate visualization

@@
 This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
+Crucially, this dissimilarity is not evaluated in terms of acoustic tags, but in terms of \emph{ad hoc} clusters, as defined by a human consensus of auditory judgments.
 Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
@@
 
-Previous publications have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
-%Although one publication has turned to the visualization of isolated musical notes, the scope of discussion on the effect of technique is
-
-Figures \ref{fig:violin-scattering} and \ref{fig:flute-scattering}.
+Previous publications on STRF have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
+However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
+Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
+
+To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
 

### 2020-07-03 paper/paper.tex
finish section on scattering.m

@@
 To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
+By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
+For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
+A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
+As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
+Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
+
+The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
+Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
+These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
+These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
+Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
+This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
+
+Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
+Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
+Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
+We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
 

### 2020-07-03 paper/paper.tex
bis
todo reformulate contributions
wonderful title

@@
 
-\title{Learning Auditory Similarities Between Instrumental Playing Techniques}
+\title{Time--Frequency Scattering Accurately Models Auditory Similarities Between Instrumental Playing Techniques}
 
@@
 \subsection*{Contributions}
-This article gathers three novel contributions.
+This article gathers three contributions. % TODO: speak in terms of "intersection" of themes rather than novel contributions?
 First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
@@
 %They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-%\ja{Need to flesh out the above with some equations.
-%We're no longer talking to biologists, so we don't need to be afraid.}
 
@@
 If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only, and re-apply it as a constant value on the test set.
-
 Then, to match a decibel-like perception of loudness, we apply the following adaptive transformation, which composes a median-based renormalization and a logarithmic compression:

### 2020-07-09 paper/paper.tex
erudite -> classical

@@
 \begin{abstract}
-Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in erudite and folk contexts.
+Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in classical and folk contexts.
 However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique; use instrument identity as a proxy for timbre quality; and do not allow for rapid customization to the perceptual idiosyncracies of a newcoming subject.

### 2020-07-09 paper/paper.tex
edit abstract

@@
 Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in classical and folk contexts.
-However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique; use instrument identity as a proxy for timbre quality; and do not allow for rapid customization to the perceptual idiosyncracies of a newcoming subject.
-In this article, we ask 31 human subjects to organize 78 isolated notes into an unconstrained number of timbre clusters.
+However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique, use instrument identity as a proxy for timbre quality, and do not allow for customization to the perceptual idiosyncracies of a new subject.
+In this article, we ask 31 human subjects to organize 78 isolated notes into a set of timbre clusters.
 Analyzing their responses suggests that timbre perception operates within a more flexible taxonomy than those provided by instruments or playing techniques alone.
 In addition, we propose a machine listening model to recover the cluster graph of auditory similarities across instruments, mutes, and techniques.
-Our model relies on joint time--frequency scattering features to extract spectrotemporal modulations at multiple scales and rates.
+Our model relies on joint time--frequency scattering features to extract spectrotemporal modulations as acoustic features.
 Furthermore, it minimizes triplet loss in the cluster graph by means of the large-margin nearest neighbor (LMNN) metric learning algorithm.
-Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of $99.0\%\pm 1$.
+% \ja: "what?".
+VL: leaving it as is.
+i think it's important to use the words "triplet loss" in the abstract
+Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of 99.0\%\pm 1.
 An ablation study demonstrates that removing either the joint time--frequency scattering transform or the metric learning algorithm noticeably degrades performance.

### 2020-07-09 paper/paper.tex
edit intro

@@
 i think it's important to use the words "triplet loss" in the abstract
-Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of 99.0\%\pm 1.
+Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of $\mathsf{99.0\%\pm1}$.
 An ablation study demonstrates that removing either the joint time--frequency scattering transform or the metric learning algorithm noticeably degrades performance.
@@
 Music information retrieval (MIR) operates at two levels: symbolic and auditory \cite{downie2003mir}.
-By resting on a notation system, the symbolic level facilitates the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
+By relying on a notation system, the symbolic level allows the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
 Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
@@
 On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
-On the other hand, the sole mention of a playing technique does not determine any intended effect; as a result, it does not generalize to the entire instrumentarium \cite{kolozali2011ismir}.
+On the other hand, the sole mention of a playing technique does not specify its effect in terms of auditory perception.
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
-
-Although a graphemic study of playing techniques in music has research potential in music information retrieval \cite{calvo2020acm}, the prospect of modeling timbre perception necessarily exceeds the symbolic domain.
+Yet, such acoustical similarity at the scale of the entire instrumentarium is not directly reflected by the semantic similarity between playing technique denominations; and discovering it requires musical expertise \cite{kolozali2011ismir}.
+
+Although a notation-based study of playing techniques in music has research potential in music information retrieval \cite{calvo2020acm}, the prospect of modeling timbre perception necessarily exceeds the symbolic domain.
 Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.

### 2020-07-09 paper/paper.tex
edit related work

@@
 Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
-As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked \cite{kostka2016book}).
+As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked) \cite{lostanlen2018extended}.
 
@@
 Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
-The task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
-Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
+The simulation of this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates pairs of stimuli.
+Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
 
 Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted orchestration \cite{maresz2013cmr}.
-In such context, the composer queries the software with an arbitrary audio signal; the outcome of this query is an instrumental audio sample, alongside textual attributes of playing technique.
-Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of priorly specified range \cite{caetano2019swarm}.
+In such context, the composer queries the software with an arbitrary audio signal.
+The outcome is another audio signal which is selected from a database of instrumental samples and perceptually similar to the query.
+The advantage of this search is that, unlike the query, the retrieved sound is precisely encoded in terms of duration, pitch, intensity, instrument, and playing technique.
+Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of predefined range \cite{caetano2019swarm}.
 
 \subsection*{Goal}
-This article proposes a machine listening system which computes the dissimilarity in timbre between any two audio samples.
+This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.
 Crucially, this dissimilarity is not evaluated in terms of acoustic tags, but in terms of \emph{ad hoc} clusters, as defined by a human consensus of auditory judgments.
 Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
+The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
 It encodes patterns of spectrotemporal modulation in the acoustic query while offering numerical guarantees of stability to local deformations \cite{anden2012scattering}.
 The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
-It recombines scattering coefficients so as to fit a training set of human judgments \cite{mcadams1995psychres}, which may either stem from the perceptual idiosyncracies of a single subject or from an intersubjective consensus.
-
-Figure \ref{fig:pipeline} summarizes our experimental protocol: it illustrates how visual annotations (top) can inform feature extraction (center) to produce a nearest-neighbor search engine which is consistent with timbre similarity judgments (bottom).
+It reweights scattering coefficients so that pairwise distances between samples more accurately reflect human judgments on a training set.
+These human judgments may be sourced from a single subject or the intersubjective consensus of multiple subjects \cite{mcadams1995psychres}.
+
+Figure \ref{fig:pipeline} summarizes our experimental protocol: it illustrates how visual annotations (top) can inform feature extraction (center) to produce a nearest-neighbor search engine which is consistent with human judgments of timbre similarity (bottom).
 
 \subsection*{Contributions}
-This article gathers three contributions. % TODO: speak in terms of "intersection" of themes rather than novel contributions?
-First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production \cite{lostanlen2018extended}.
+The main contribution of this article can be formulated as the intersection between three topics.
+To the best of our knowledge, prior literature has addressed these topics separately, but never in combination.
+
+First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
+
 Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
 In other words, we rely on listeners, not performers, to define and evaluate the task at hand.
-A comprehensive ablation study of our system demonstrates that all design choices presented hereafter benefit the retrieval precision of our system.
-Thirdly, we supplement our quantitative benchmark with some visualizations of time--frequency scattering coefficients in the rate--scale domain, for various typical samples of realistic instrumental playing techniques.
-These visualizations confirm the theoretical predictions of previous research \cite{anden2019tsp}, which were founded on simple synthetic sounds such as exponential chirps.
-They are also in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.
+
+Thirdly, we supplement our quantitative benchmark with visualizations of time--frequency scattering coefficients in the rate--scale domain for various typical samples of instrumental playing techniques.
+These visualizations are in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.
 
@@
 Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more \cite{joder2009taslp}.
-Mapping this input to a feature space in which distances denote timbral dissimilarities requires a data-driven stage of dimensionality reduction.
-In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}.
+Mapping this input to a feature space in which distances denote timbral dissimilarity requires a data-driven stage of dimensionality reduction.
+In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of music information retrieval (MIR) encodes timbre under the form of ``audio tags'', music psychology measures timbre similarity directly from pairwise judgments.
 
@@
 For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
-Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from 8 instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
-
-Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures, and notably deep convolutional networks, in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
-Notwithstanding the growing accuracy of these architectures in the large data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably scarcer \cite{loureiro2004ismir}.
+% \ja wants 20k, 21k.
+Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from eight instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
+
+Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
+Notwithstanding the growing accuracy of these architectures in the large-scale data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably more scarce \cite{loureiro2004ismir}.
 We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
@@
 Rather, each excerpt must be played back to multiple human listeners.
-Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, therefore restricting the size of the musical corpus under study.
-Such restriction hampers the applicability of iterative algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
-Nevertheless, while training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}, thus yielding a useful form of inductive bias for downstream machine listening applications.
-
-The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
+Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, which restricts the size of the musical corpus under study.
+These small corpus sizes hampers the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
+
+While training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
+This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
+The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
 The simplest method to compute it in practice is by reverse correlation, i.e.\ by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
@@
 Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
+
 Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner--Ville distribution.
-In particular, a substantial portion of spectrographic STRFs exhibits a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain \cite{theunissen2000jneur}.
+In particular, a substantial portion of spectrographic STRFs exhibit a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain \cite{theunissen2000jneur}.
 This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
 
-Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the selected center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$.
-
-\subsection*{Spectrotemporal receptive fields (STRF)}
+Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the center time $t$ and center frequency $\lambda$.
+This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as the ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$ using geometric progressions.
+Because they do not require a data-driven training procedure, STRF yield a useful form of domain-specific knowledge for downstream machine listening applications, especially when the number of annotated samples is relatively small.
+
+\subsection*{Spectrotemporal receptive fields (STRFs) as a feature extractor}
 Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
 These biologically inspired features outperform the state of the art, especially in the small data regime where deep learning is inapplicable.
-Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments strongly correlates with the confusion matrix between a human listener and the ground truth.
+Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments is strongly correlated with the confusion matrix between a human listener and the ground truth.
 Another appeal of the full cortical model is that the three-way tensor of frequency $\lambda$, rate $\alpha$, and scale $\beta$ can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
@@
 Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
-On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2020jmlr} and invertibility \cite{lostanlen2019dafx}.
+In contrast, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2020jmlr} and invertibility \cite{lostanlen2019dafx}.
 Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
@@
 
-Previous publications on STRF have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
+Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
 However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.

### 2020-07-19 paper/paper.tex
clarify multirate scheme

@@
 Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
+This approach requires a uniform discretization of the scalogram, and thus an oversampling of the lower-frequency subbands to the Nyquist frequency of the original signal.
 In contrast, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2020jmlr} and invertibility \cite{lostanlen2019dafx}.
 Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
-This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
+In this multirate scheme, every subband is discretized at its critical sample rate, i.e., in proportion to its center frequency.
+As a by-product, the multirate approach draws an explicit connection between scattering networks and deep convolutional networks, because both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
 

### 2020-07-19 paper/paper.tex
remove dyadic

@@
 In this multirate scheme, every subband is discretized at its critical sample rate, i.e., in proportion to its center frequency.
-As a by-product, the multirate approach draws an explicit connection between scattering networks and deep convolutional networks, because both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
+As a by-product, the multirate approach draws an explicit connection between scattering networks and deep convolutional networks, because both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and pooling operations \cite{mallat2016philtrans}.
 

### 2020-07-19 paper/paper.tex
tractable -> efficient

@@
 Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
+% \ja wants 5,000
 However, it remains to be seen whether joint time--frequency scattering is capable of fine-grained auditory categorization, involving variability in instrument, mute, and playing technique.
@@
 The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
+From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
 Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.

### 2020-07-19 paper/paper.tex
formal -> theoretical

@@
 The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain.
+From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
 Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.

### 2020-07-19 paper/paper.tex
incorporate JA changes

@@
 From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being readily tailored to subjective preferences.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is also readily tailored to subjective preferences.
 
@@
 The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
-In this article, we focus on the former three, and leave the question of learning auditory similarities between percussion instruments to future research.
-We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition respectively.
+In this article, we focus on the first three, and leave the question of learning auditory similarities between percussion instruments to future research.
+We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition, respectively.
 

### 2020-07-19 paper/paper.tex
bis

@@
 We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-Among this list, the former six are strings, the next six are woodwind, and the last four are brass.
+Among this list, the first six are strings, the next six are woodwind, and the last four are brass.
 Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'', also known as \emph{harmon}, inside in the bell of a trumpet.
-Once augmented with mutes, the list of 16 instruments grows up to 33.
+Once augmented with mutes, the list of 16 instruments grows to 33.
 Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between 11 (for the accordion) and 41 (for the bass tuba).
@@
 One example of IMT is \texttt{TpC+S-ord}, \ie{} trumpet in C with a straight mute played in the ordinary technique.
-Another example of IMT is \texttt{Vn-pont}, \ie{} violin without any mute played in the \emph{sul ponticello} (bowing near the bridge) technique.
-
-Performers can play each IMT at various pitches, according to a tessitura that depends on the instrument but not on the choice of mute.
+Another example of IMT is \texttt{Vn-pont}, \ie{} violin without any mute played in the \emph{sul ponticello} technique (bowing near the bridge).
+
+Performers can play each IMT at various pitches according to the tessitura of their instrument.
+This tessitura may depend on the choice of playing technique but is independent of the choice of mute.
 Among the 16 instruments in this study, the two instruments with widest and narrowest tessituras, in their respective ordinary techniques, are the accordion (81 semitones) and the trumpet in C (32 semitones) respectively.
 Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzo forte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).
-However, the resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
+The resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
 For example, the pitch of pedal tones in brass instruments is tied to the fundamental mode of the bore, \ie{} usually $\mathsf{B}\musFlat$ or $\mathsf{F}$.
@@
 
-In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
+In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two quantitative variables (intensity and pitch).
 The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a joint sampling of these variables.

### 2020-07-19 paper/paper.tex
page 7

@@
 The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a joint sampling of these variables.
-The version of SOL that we use throughout this paper, named ``0.9 HQ'', amounts to 25444 audio signals in total.
-Beyond playing techniques, we should note that SOL erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
-Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.
-Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
+The version of SOL that we use throughout this paper, named ``0.9 HQ'', amounts to a total of 25444 audio signals.
+Beyond playing techniques, we should note that SOL erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics, which are all restricted to singletons.
+Addressing these factors of variability is beyond the scope of this paper, which focuses on the influence of playing technique.
+Despite this restriction, the SOL dataset remains impractically large for collecting human similarity judgments.
 Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and the use of an efficient annotation interface.
@@
 Indeed, under first approximation, the perception of timbre is invariant to pitch and intensity.
-Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C (\ie{} $\mathsf{C_4}$) and \emph{mf}.
+Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C ($\mathrm{C_4}$) and \emph{mf}.
 After this selection, every IMT triplet contains a single acoustic exemplar, regarded as canonical in the following.
 The number of canonical stimuli for the entire SOL dataset is equal to 235.
-We should note, however, that the proposed pitch--intensity cannot be strictly enforced across all IMTs.
+We should note, however, that the proposed pitch and intensity cannot be strictly enforced across all IMTs.
 Indeed, as explained above, a fraction of IMTs can only be achieved at restricted values of pitch and intensity parameters, \eg{} pedal tones or key clicks.
-Therefore, at a small cost of consistency, we only enforce the pitch--intensity reference (\ie{} C4 and \emph{mf}) when practically feasible, and fall back to other pitches and intensities if necessary.
+Therefore, at a small cost of consistency, we only enforce the pitch--intensity reference (\ie{} $\mathrm{C_4}$ and \emph{mf}) when practically feasible, and fall back to other pitches and intensities if necessary.
 
@@
 Here, we define the relevance of an IMT as the possibility of imitating it by means of another IMT from a different instrument.
-One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in strings instruments.
-To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatoire (CNSMDP), and explicitly gave them our definition of relevance.
+One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in string instruments.
+To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP\footnote{CNSMDP: Conservatoire National Sup\'erieur de Musique et de Danse de Paris.
+Official website: \url{https://www.conservatoiredeparis.fr}}).
 Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
-In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic of the instrument, and that therefore, it is unlikely that humans will pair it with IMTs from other instruments.
+In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic, and that therefore, it is unlikely that humans will pair it with other IMTs.
 Conversely, the value 7 (most relevant) denotes that the IMT under consideration bears a strong similarity with some other IMT from the corpus.
 
-Once both experts completed their annotation, we ranked all 235 IMTs by decreasing average relevance score.
-Then, we retained all IMTs whose average score was judged equal to 3 or higher, thus resulting in a shortlist of $N=78$ IMTs (see Tables 1 and 2 in the Appendix).
+Once both experts completed their annotations, we retained all IMTs whose average score was judged equal to 3 or higher, thus resulting in a shortlist of $N=78$ IMTs (see Tables 1 and 2 in the Appendix).
 It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic to be relevant for this experiment, regardless of playing technique.

### 2020-07-19 paper/paper.tex
page 8

@@
 Assuming that similarity is a symmetric quantity, collecting a dense matrix of continuously valued ratings of similarity among a dataset of $N$ items would require $\frac{1}{2}(N^2-N)$ Likert scales.
-In the case of $N=78$ IMTs, the task would amount to about 3k horizontal sliders, \ie{} several hours of cumbersome work for the human annotator.
-
-Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of continuous ratings.
-To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all $31$ participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}.
+In the case of $N=78$ IMTs, the task would amount to about 3k horizontal sliders, \ie{} several hours of cumbersome work for the human annotator. % \ja wants 3,000
+
+Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of quantitative ratings.
+To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity.\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all 31 participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}
 The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz.
@@
 
-Cyberlioz consists of a square panel on which is displayed a collection of circular grey dots, each of them corresponding to one of the $N=78$ IMTs, and initially distributed uniformly at random.
-Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} $\textsf{C}_4$ and \emph{mf} in most cases.
+Cyberlioz consists of a square panel on which is displayed a collection of circular grey dots, each of them corresponding to one of the IMTs, and initially distributed uniformly at random.
+Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} $\mathrm{C}_4$ and \emph{mf} in most cases.
 Furthermore, each dot can be freely placed on the screen by clicking, dragging, and dropping.
-Lastly, the user can assign a color to each dot by selecting it among a palette of 20 hues.
+Lastly, the user can assign a color to each dot among a palette of 20 hues.
 The goal of the Cyberlioz interface is to form clusters of timbre similarity between IMTs, expressed by sameness of color.
-Here, any two audio samples are defined as similar if and only if one can be replaced by the other in a context of spectralist orchestration.
-
-In comparison with a traditional Web-based form, Cyberlioz offers a more intuitive and playful user experience, while expediting the obtention of a dense matrix ($N\times N$) of similarity judgments within a moderate duration of $30$ to $60$ minutes for each participant.
+
+In comparison with a traditional web-based form, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
 In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
-To this end, we publicized this study on the internal mailing list of the Paris Conservatoire (CNSMDP) students, as well as two international mailing lists for research in music audio processing: ``AUDITORY''  and ``ISMIR Community''\footnote{
+To this end, we publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
 For more information about these mailing lists, please visit:
-\url{http://www.auditory.org/} and \url{http://ismir.net/}}.
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.%Figure \ref{fig:xp2display}.
+\url{http://www.auditory.org/} and \url{http://ismir.net/}}
+Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.
 

### 2020-07-19 paper/paper.tex
page 9

@@
 Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a cluster graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
-For a given $k$, the graph $\mathcal{G}_k$ contains $N=78$ vertices, each representing a different IMT in the corpus.
+For a given $k$, the graph $\mathcal{G}_k$ contains $N$ vertices, each representing a different IMT in the corpus.
 In $\mathcal{G}_k$, an edge connects any two vertices $m$ and $n$ if the corresponding dots in Cyberlioz have the same color.
@@
 We aggregate the similarity judgments from all $K$ subjects by embedding them into a hypergraph $\mathcal{H}$, that is, a graph whose edges may connect three or more vertices at once.
-Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_k$'s among all subject $k$.
-Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm algorithm known as hypergraph partitioning \cite{kernighan1970efficient}.
+Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$.
+Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm known as hypergraph partitioning \cite{kernighan1970efficient}.
 
 For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
-Across all $K=31$ subjects, the value of $C_k$ varies between $3$ and $19$, with a median value of $10$.
-To construct $\mathcal{G}_0$, we select a number of clusters $C_0$ that is equal to the maximal value of the $C_k$'s; that is,
-\begin{equation}
-C_0 = \max_{1\leq k \leq K} C_k = 19
-\end{equation}
+Across all $K=31$ subjects, $C_k$ varies between $3$ and $19$, with a median value of $10$.
+To construct $\mathcal{G}_0$, we select a number of clusters that is equal to the maximal value of the $C_k$'s; that is,$C_0 = 19$.
 Then, we run hypergraph partitioning on $\mathcal{H}$ to assign each vertex $i$ to one of the $C_0$ clusters in $\mathcal{G}_0$.
@@
 
-While the graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants, while discarding intersubjective variability.
+While the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants while discarding intersubjective variability.
 Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ subjects.
-This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a newcoming subject, after less than one hour of using Cyberlioz.
+This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
 
@@
 Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
-Specifically, we ask whether human listeners, when confronted with a diverse corpus of IMT samples, tend to cluster those samples in terms of an organological taxonomy.
+Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
 To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of pairs in $\mathcal{G}_0$ from instruments $i$ and $j$ respectively, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
+For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
 \begin{equation}

### 2020-07-19 paper/paper.tex
page 10

@@
 To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$:
+For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$.
+
+Let us denote the existence of an edge in $\mathcal{G}_0$ connecting two IMT samples $m$ and $n$ by the binary relation $m \overset{\mathcal{G}_0}{\sim} n$.
+The mathematical definition of the matrix $\mathbf{A}_{\mathcal{I}}$ is:
 \begin{equation}
 \mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
-2\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
 }{
@@
 
-Then, we run Ward's method to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
-This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$.
-Intuitively, for every instrument--instrument pair $(i, j) \in \mathcal{I}^2$, the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
+Then, we run Ward's method \cite{ward1963jasa} to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
+This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$ such that the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
 Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
+% \ja wants to remove this equation
 
@@
 From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
-We define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinals:
+As for $\mathbf{A}_{\mathcal{I}}$, we define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinalities:
 \begin{equation}
@@
 \Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
-},
+}.
 \label{eq:technique-similarity}
 \end{equation}
-alike the definition of $\mathbf{A}_{\mathcal{I}}$.
 Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
-Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$ as a result of such agglomerative hierarchical clustering procedure.
-
-Contrary to inter-instrument similarity (Figure \ref{fig:consensusVsI}), inter-technique similarity does not exhibit a clear segregation of instrument families.
-For example, the first two rows in $\widetilde{\mathbf{A}}_{\mathcal{T}}$ correspond to a oboe sample (\texttt{Ob-blow-no-reed-C4}, \ie{} blowing without the reed) and a cello sample (\texttt{Vc-legno-tratto-C4-mf-1c}, \ie{} drawing the wooden part of the bow across the string) respectively, despite the fact that the oboe is a woodwind instrument whereas the cello is a string instrument.
-
-More generally, even though the block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less prevalent than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$, playing technique clusters appear to transcend instrument categories.
+Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$.
+
+The block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less salient than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$.
 Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.

### 2020-07-19 paper/paper.tex
page 11

@@
 \label{sec:methods}
-The previous section described our protocol for collecting timbral similarity judgments between instrumental playing techniques at scale.
-In this section, we aim to recover these similarity judgments from digital audio recording, according to a paradigm of supervised metric learning.
-To this end, we present a new machine listening system, composing joint time--frequency scattering and large-margin nearest neighbors (LMNN).
-
-%Following the above discussed matters, we propose a ranking system that is based on a processing pipeline made of two main components, see Figure \ref{fig:pipeline}.
-%The first component maps the audio signal of the query and the items of the database into a high-dimensional timbre space using the scattering transform.
-%The second component warps this timbre space by taking into account perceptual judgments represented as a consensus clustering of the items in the dataset, as presented in the \nameref{sec:data-collection} section.
-%This warping, or reweighting, of the space is performed using LMNN.
+The previous section described our protocol for collecting timbral similarity judgments between instrumental playing techniques.
+In this section, we aim to recover these similarity judgments from digital audio recordings according to a paradigm of supervised metric learning.
+To this end, we present a machine listening system composing joint time--frequency scattering and large-margin nearest neighbors (LMNN).
 
@@
 
-Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ a Hilbert-analytic filter with null average, unit dimensionless center frequency, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
+Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ a complex-valued filter with zero average, dimensionless center frequency equal to one, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
 We define a constant-$Q$ wavelet filterbank as the family $\boldsymbol{\psi}_{\lambda} : t \mapsto \lambda \boldsymbol{\psi}(\lambda t)$.
@@
 \boldsymbol{\psi}:t \longmapsto \exp\left(-\dfrac{t^2}{2\sigma_{\psi}^2}\right)
-\big[ \exp\left(2\pi \mathrm{i}t\right) - \kappa_{\psi} \big],
+\big( \exp\left(2\pi \mathrm{i}t\right) - \kappa_{\psi} \big),
 \label{eq:psi}
 \end{equation}
-where the Gaussian width $\sigma_{\psi}$ grows in proportion with the quality factor $Q$ and the corrective term $\kappa_{\psi}$ ensures that $\boldsymbol{\psi}$ has a null average over $\mathbb{R}$.
+where the Gaussian width $\sigma_{\psi}$ grows in proportion with the quality factor $Q$ and the corrective term $\kappa_{\psi}$ ensures that $\boldsymbol{\psi}$ has a zero average.
 Moreover, we discretize the frequency variable $\lambda$ according to a geometric progression of common ratio $2^{\frac{1}{Q}}$.
@@
 
-Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R})$, followed by an application of the pointwise complex modulus, yields the wavelet scalogram
+Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R})$, followed by an application of the pointwise complex modulus yields the wavelet scalogram
 \begin{equation}
@@
 \end{equation}
-which is discretized according to a dyadic multirate scheme, similarly to the constant-$Q$ transform of \cite{schorkhuber2010smc}.
+which is discretized similarly to the constant-$Q$ transform of \cite{schorkhuber2010smc}.
 Then, we define a two-dimensional Morlet wavelet $\boldsymbol{\Psi}$ of the form
@@
 \exp\left(-\dfrac{t^2+u^2}{2\sigma_{\Psi}^2}\right)
-\Big[ \exp\big(2\pi \mathrm{i} (t + u)\big) - \kappa_{\Psi} \Big],
+\Big( \exp\big(2\pi \mathrm{i} (t + u)\big) - \kappa_{\Psi} \Big),
 \label{eq:Psi}
@@
 
-We choose the Gaussian width $\sigma_{\Psi}$ in Equation \ref{eq:Psi} such that the quality factor of the Morlet wavelet $\Psi$ is equal to $1$, both over the time dimension and over the log-frequency dimension.
+We choose the Gaussian width $\sigma_{\Psi}$ in Equation \ref{eq:Psi} such that the quality factor of the Morlet wavelet $\Psi$ is equal to one, both over the time dimension and over the log-frequency dimension.
 Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.

### 2020-07-19 paper/paper.tex
page 12

@@
 taking two real variables $t$ and $u$ as input.
+In the rest of this paper, we shall refer to $\boldsymbol{\Psi}$ as a time--frequency wavelet.
 The former is the time variable while the latter is the base-two logarithm of frequency: $u = \log_2 \lambda$.
@@
 We choose the Gaussian width $\sigma_{\Psi}$ in Equation \ref{eq:Psi} such that the quality factor of the Morlet wavelet $\Psi$ is equal to one, both over the time dimension and over the log-frequency dimension.
-Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a null average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
+Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a zero average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
 From $\mathbf{\Psi}$, we define a two-dimensional wavelet filterbank of the form:
@@
 While $\alpha$ is measured in Hertz and is strictly positive, $\beta$ is measured in cycles per octaves and may take positive as well as negative values.
-Both $\alpha$ and $\beta$ are discretized by geometric progressions of common ratio equal to $2$.
+Both $\alpha$ and $\beta$ are discretized by geometric progressions of common ratio equal to two.
 Furthermore, the edge case $\beta=0$ corresponds to $\mathbf{\Psi}_{\alpha,\beta}$ being a Gaussian low-pass filter over the log-frequency dimension, while remaining a Morlet band-pass filter of center frequency $\alpha$ over the time dimension.
@@
 
-Then, we convolve the scalogram $\mathbf{U_1}$ with all time--frequency wavelets $\mathbf{\Psi}_{\alpha,\beta}$ and apply the complex modulus, yielding the four-way tensor
+We now convolve the scalogram $\mathbf{U_1}$ with time--frequency wavelets $\mathbf{\Psi}_{\alpha,\beta}$ and apply the complex modulus, yielding the four-way tensor
 \begin{align}
@@
 \end{align}
-where the symbol $\otimes$ denotes an outer product between two vectors.
+where the symbol $\otimes$ denotes the outer product over time and log-frequency.
 In the equation above, the sample rate of time $t$ is proportional to $T^{-1}$ and the sample rate of log-frequency $u=\log_2 \lambda$ is proportional to $F^{-1}$.
@@
 
-The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of a spectrotemporal receptive field (STRF) at the rate $\alpha$ and the scale $\beta$.
-Nevertheless, in comparison with the representation known as ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
+The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of an STRF at the rate $\alpha$ and the scale $\beta$.
+Nevertheless, in comparison with the ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
 This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.
+% \ja says "where does this number come from"? Answer: from the Patil paper vs the scattering.m implementation
 

### 2020-07-19 paper/paper.tex
page 13

@@
 
-In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension, with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
+In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
 \begin{align}
@@
 Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
-Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\psi}_F$, comparably to Equation $\ref{eq:Psi}$.
+Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\phi}_F$.
 We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.
 
-We adopt the multiindex notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
-The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) or second-order ($\mathbf{S_2}$) coefficients.
+We adopt the multi-index notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
+The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) and second-order ($\mathbf{S_2}$) coefficients.
 Given an input waveform $\boldsymbol{x}$, we denote by $\mathbf{S}\boldsymbol{x}$ the feature vector resulting from the concatenation of $\mathbf{S_1}\boldsymbol{x}$ and $\mathbf{S_2}\boldsymbol{x}$:
@@
 
-An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x})$ can be bounded as follows:
-\begin{equation}
-\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
-\leq
-\kappa_{\mathbf{S}} \times
-\Vert \boldsymbol{x} \Vert^2 \times
-\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
-\label{eq:lipschitz}
-\end{equation}
-where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
-In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
-In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.
-
-Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
-In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
-Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.
+%An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{x}\circ\boldsymbol{\tau})$ can be bounded as follows:
+%\begin{equation}
+%\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
+%\leq
+%\kappa_{\mathbf{S}} \times
+%\Vert \boldsymbol{x} \Vert^2 \times
+%\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
+%\label{eq:lipschitz}
+%\end{equation}
+%where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
+%In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
+%In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.
+%
+%Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
+%In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
+%Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.
 

### 2020-07-19 paper/paper.tex
page 14

@@
 %In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
-%In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.
+%In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such a deformation.
 %
@@
 
-It stems from the above that comparing Euclidean distances in the feature space of time--frequency scattering coefficients may serve as a computationally efficient surrogate for comparing Wasserstein distances in the scalogram, without need for solving an optimal transport problem.
-However, this surrogate is only valid if the underlying transport plan $\boldsymbol{\tau}$ is relatively small  in terms of supremum norm (see Equation \ref{eq:lipschitz}).
-This assumption may be true for certain local deformation of the time-frequency domain, \eg{} causing a pitch shift or a change in vibrato rate while leaving the instrument unchanged \cite{anden2012scattering}.
-However, it does not suffice in order to represent timbre similarity across multiple instruments and playing techniques.
-Indeed, the perception of timbre for extended playing techniques often involves long-range interactions in the time--frequency domain, which are not amenable to a local elastic deformation \cite{lostanlen2018extended}.
-
-To address this issue, we apply a metric learning algorithm, named large-margin nearest neighbors (LMNN), whose goal is to reweight and recombine scattering paths $p$ such that certain factors of variability in feature space $\mathbf{S}$ are enhanced while others are discarded.
-Before training this algorithm, we apply a pointwise nonlinear transformation on averaged joint time-frequency scattering coefficients.
-The role of this transformation, which is adapted to the dataset in an unsupervised way, is to Gaussianize the histogram of activations of each scattering path $p$.
-
-We consider a collection $\mathcal{X}$ of $N$ waveforms, numbered arbitrarily from $\boldsymbol{x}_1$ to $\boldsymbol{x}_N$.
-For every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we measure the median value taken by the average response of $\mathbf{S}\boldsymbol{x}_n$ across all samples $n$ from $1$ to $N$:
+%It follows from the above that comparing Euclidean distances in the feature space of time--frequency scattering coefficients may serve as a computationally efficient surrogate for comparing Wasserstein distances in the scalogram, without need for solving an optimal transport problem.
+%However, this surrogate is only valid if the underlying transport plan $\boldsymbol{\tau}$ is relatively small  in terms of supremum norm (see Equation \ref{eq:lipschitz}).
+%This assumption may be true for certain local deformation of the time-frequency domain, \eg{} causing a pitch shift or a change in vibrato rate while leaving the instrument unchanged \cite{anden2012scattering}.
+%However, it does not suffice in order to represent timbre similarity across multiple instruments and playing techniques.
+%Indeed, the perception of timbre for extended playing techniques often involves long-range interactions in the time--frequency domain, which are not amenable to a local elastic deformation \cite{lostanlen2018extended}.
+
+%To address this issue, we apply a metric learning algorithm, named large-margin nearest neighbors (LMNN), whose goal is to reweight and recombine scattering paths $p$ such that certain factors of variability in feature space $\mathbf{S}$ are enhanced while others are discarded.
+Now, we apply a pointwise nonlinear transformation on averaged joint time-frequency scattering coefficients.
+The role of this transformation, which is adapted to the dataset in an unsupervised way, is to Gaussianize the histogram of amplitudes of each scattering path $p$.
+We consider a collection $\mathcal{X}$ of $N$ waveforms $\boldsymbol{x}_1,\ldots,\boldsymbol{x}_N$.
+For every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we average the response of each scattering coefficient $\mathbf{S}\boldsymbol{x}_n$ over time and take its median value across all samples $n$ from $1$ to $N$:
 \begin{equation}
@@
 \end{equation}
-If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only, and re-apply it as a constant value on the test set.
+If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only.
 Then, to match a decibel-like perception of loudness, we apply the following adaptive transformation, which composes a median-based renormalization and a logarithmic compression:
@@
 \int_\mathbb{R} \mathbf{S}\boldsymbol{x}_n (t, p) \;\mathrm{d}t
-}{\varepsilon \times \boldsymbol{\mu}(p)}
+}{\varepsilon\boldsymbol{\mu}(p)}
 \right)
@@
 where $\varepsilon$ is a predefined constant.
-The offset of $1$ before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps.
-Furthermore, a previous publication has shown empirically that Equation \ref{eq:S_tilde} brings, on environmental audio recordings, the histogram of $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ closer to a statistical Gaussianization of every feature $p$ \cite{lostanlen2018jasmp}.
-Since then, this finding has been confirmed in the case of musical sounds \cite{lostanlen2018extended}.
+The offset of one before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps.
+On a dataset of environmental audio recordings, a previous publication has shown empirically that Equation \ref{eq:S_tilde} brings the histogram of $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ closer to a Gaussian distribution \cite{lostanlen2018jasmp}.
+Since then, this finding has also been confirmed in the case of musical sounds \cite{lostanlen2018extended}.
 
@@
 In our protocol, $\boldsymbol{x}$ contains the sound of an isolated musical note and the cluster graph $\mathcal{G}$ encodes auditory similarities within the dataset $\mathcal{X}$.
-Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K=31$ subjects from our perceptual study (see \lnameref{sec:data-collection} section).
-%write on multiple subjects?
+Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K$ subjects from our perceptual study (see \lnameref{sec:data-collection} section).
 

### 2020-07-19 paper/paper.tex
page 15

@@
 Let $\mathcal{G}$ a cluster graph with $N = \card \mathcal{X}$ vertices and $C$ clusters in total.
-We denote by $\mathcal{G}(\boldsymbol{x})$ the cluster index to which the sample $\boldsymbol{x}$ belongs.
+We denote by $\mathcal{G}(\boldsymbol{x})$ the cluster to which the sample $\boldsymbol{x}$ belongs.
 Given another sample $\boldsymbol{y}$ in $\mathcal{X}$, $\boldsymbol{y}$ is similar to $\boldsymbol{x}$ if and only if belongs to the cluster $\mathcal{G}(\boldsymbol{x})$.
@@
 This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsections above).
-Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients, according to the Euclidean metric.
+Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients according to the Euclidean metric.
 Unlike cluster similarity, this relationship is not symmetric: $\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})$ does not necessarily imply $\boldsymbol{x}\in\mathcal{Y}_R (\boldsymbol{y})$.
-Note that the dependency of $\mathcal{Y}_R (\boldsymbol{x})$ upon the operator $\mathbf{S}$ is left implicit.
-In all of the following, we set the constant $R$ to $5$; this is in accordance with our chosen evaluation metric, i.e., average precision at rank 5 (AP@5, see \lnameref{sec:results} section).
+Note that the dependency of $\mathcal{Y}_R (\boldsymbol{x})$ upon the operator $\mathbf{S}$ is left implicit. % \ja wants to delete this sentence
+In all of the following, we set the constant $R$ to $5$; this is in accordance with our chosen evaluation metric, average precision at rank 5 (AP@5, see \lnameref{sec:results} section).
 
@@
 \end{equation}
-Crucially, in the feature space of time--frequency scattering coefficients, the Euclidean neighborhood $\mathcal{Y}_R (\boldsymbol{x})$ corresponds to a Wasserstein neighborhood in the scalogram domain (see previous subsection).
 Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
@@
 While the the term $\mathcal{E}_{\textrm{pull}}$ operates on pairs of samples, the term $\mathcal{E}_{\textrm{push}}$, operates on triplets $(\boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z})\in\mathcal{X}^3$.
-The first sample, $\boldsymbol{x}$, is known as an ``anchor''.
+The first sample, $\boldsymbol{x}$, is known as an ``anchor.''
 The second sample, $\boldsymbol{y}$, is known as a ``positive'', and is assumed to belong to the Euclidean neighborhood of the anchor: $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$.

### 2020-07-19 paper/paper.tex
page 16

@@
 where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU), also known as hinge loss.
-The cost function described in the equation above is known in deep learning as ``triplet loss'', and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
+The cost function described in the equation above is known in deep learning as ``triplet loss'' and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
 We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
@@
 We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
+In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
 However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
@@
 
-In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must withdraw from a diverse realm of instruments and techniques.
+In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must draw from a diverse realm of instruments and techniques.
 Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity level (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.
@@
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL (roughly one third of the entire dataset, i.e., $25444$ samples) for which the notion of timbre is perceptually encoded as a discrete category.
+This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
 
@@
 
-Let us denote by $\boldsymbol{x}_1 \ldots \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}=9346$ audio samples associated to our annotated dataset, after semi-supervised label propagation (see subsection above).
+Let us denote by $\boldsymbol{x}_1, \ldots, \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}$ audio samples associated with our annotated dataset after semi-supervised label propagation.
 Given a sample $n\leq N^{\prime}$ and a human subject $k\leq K$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.

### 2020-07-19 paper/paper.tex
page 17

@@
 Let us denote by $\boldsymbol{x}_1, \ldots, \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}$ audio samples associated with our annotated dataset after semi-supervised label propagation.
-Given a sample $n\leq N^{\prime}$ and a human subject $k\leq K$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.
+Given a sample $n$ and a human subject $k$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.
 Our machine listening system takes the waveform $\boldsymbol{x}_{n}$ as input and returns a ranked list of nearest neighbors: $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_2 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_3 (\boldsymbol{x}_n)$, and so forth.
@@
 Likewise, the second retrieved sample is relevant if and only if $\mathbf{\Phi}_{2}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
-To evaluate $\mathbf{\Phi}$ on the query $\boldsymbol{x}_n$, we measure the relevance of all nearest neighbors $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ up to some fixed rank $R$, and average the result:
+To evaluate $\mathbf{\Phi}$ on the query $\boldsymbol{x}_n$, we measure the relevance of all nearest neighbors $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ up to some fixed rank $R$ and average the result:
 \begin{equation}
@@
 \end{equation}
-In the equation above, the function $\mathds{1}$ converts Booleans to integers, i.e., $\mathds{1}(b)$ equals $1$ if $b$ is true and return $0$ if $b$ is false.
+In the equation above, the indicator function $\mathds{1}$ converts Booleans to integers, i.e., $\mathds{1}(b)$ returns one if $b$ is true and return zero if $b$ is false.
 Thus, the function $p_{\mathbf{\Phi}}$ takes fractional values between $0$ and $1$, which are typically expressed in percentage points.
 
-The precision at rank $R$ of the system $\mathbf{\Phi}$ is defined as the average value taken by the function $p_{\mathbf{\Phi}}$, for constant $R$, over the entire corpus of $N^{\prime}=9346$ audio samples:
+The precision at rank $R$ of the system $\mathbf{\Phi}$ is defined as the average value taken by the function $p_{\mathbf{\Phi}}$ over the entire corpus of $N^{\prime}$ audio samples:
 \begin{equation}
@@
 \end{equation}
-It appears from the above that an effective system $\mathbf{\Phi}$ should retrieve sounds whose IMT triplets are similar according to all of the $K=31$ cluster graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$.
-
-In the rest of this paper, we set the constant $R$ to $5$.
+It appears from the above that an effective system $\mathbf{\Phi}$ should retrieve sounds whose IMT triplets are similar according to all of the $K$ cluster graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$.
+
+In the rest of this paper, we set $R$ to $5$.
 This is in accordance with the protocol of \cite{lostanlen2018extended}, in which the authors trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
-Moreover, note that the value $R=5$ typically reflects the number of virtual panels simultaneously on display in a software environment for computer-assisted orchestration.
+%Moreover, note that the value $R=5$ typically reflects the number of virtual panels simultaneously on display in a software environment for computer-assisted orchestration.
 
 \subsection*{Best performing system}
-Our topline system comprises five computational blocks:
+Our best performing system comprises five computational blocks:
 \begin{enumerate}
-\item Joint time--frequency scattering up to a maximal time scale of $T=$\SI{1000}{\milli\second},
+\item Joint time--frequency scattering up to a maximal time scale of $T=$ \SI{1000}{\milli\second},
 \item Temporal averaging at the scale of the whole musical note,
-\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended},
-\item Affine standardization so that each feature has null mean and unit variance,
-\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
+\item median-based logarithmic compression,
+\item affine standardization so that each feature has zero mean and unit variance, and
+\item nearest-neighbor search according to priorly learned non-Euclidean metric.
 \end{enumerate}

### 2020-07-19 paper/paper.tex
page 18

@@
 Our best performing system comprises five computational blocks:
+
 \begin{enumerate}
@@
 \item Temporal averaging at the scale of the whole musical note,
-\item median-based logarithmic compression,
-\item affine standardization so that each feature has zero mean and unit variance, and
-\item nearest-neighbor search according to priorly learned non-Euclidean metric.
+\item Median-based logarithmic compression,
+\item Affine standardization so that each feature has zero mean and unit variance, and
+\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
 \end{enumerate}
-Note that the non-Euclidean metric is learned via large-margin nearest neighbors (LMNN, see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
+
+Note that the non-Euclidean metric is learned via LMNN (see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
 Therefore, the system $\mathbf{\Phi}$ performs timbre similarity retrieval in a user-agnostic way, and can serve as a convenient default algorithm for newcoming users.
 That being said, it is conceivable to replicate the five-stage protocol above on the cluster graph $\mathcal{G}_k$ of a specific user $k$, instead of the cluster graph $\mathcal{G}_0$.
-This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity, in comparison with the default setting $\mathcal{G}_0$.
-
-For composers to personalize their search engine $\mathbf{\Phi}$, the quid pro quo is to define manually a cluster graph among $N=78$ audio samples, by means of the Cyberlioz annotation interface (see \lnameref{sec:data-collection} section).
+This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity.
+
+Composers may personalize their search engine $\mathbf{\Phi}$ by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
 This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
@@
 However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
-In particular, it does not assume the ``metadata'' (instrument, mute, technique, pitch, dynamics, and so forth) of any musical sample $\boldsymbol{x}_n$ to be observable, in part or in full, at training time.
+In particular, it does not assume the metadata (instrument, mute, technique, pitch, dynamics, and so forth) of any musical sample $\boldsymbol{x}_n$ to be observable, in part or in full, at training time.
 
@@
 The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
-Over a cohort of $K=31$ subjects, we reported a precision at rank five (P@5) of $99.0\% \pm 1$.
+Over a cohort of $K=31$ subjects, we reported an average precision at rank five (AP@5) of $99.0\% \pm 1$.
 We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
@@
 Figure \ref{fig:ablation} summarizes our results.
-Interestingly, the system $\boldsymbol{\Phi}$ is not only best on average, but also best for every subject in the cohort: altering any design choice in $\boldsymbol{\Phi}$ (see subsections below) results in a P@5 below $\mathrm{P}_{\mathbf{\Phi}}(k, R)$ for every $k$.
-Borrowing from the terminology of welfare economics, $\boldsymbol{\Phi}$ can be said to be uniquely Pareto-efficient.
+Interestingly, the system $\boldsymbol{\Phi}$ is not only best on average, but also best for every subject in the cohort.
+Specifically, replacing $\boldsymbol{\Phi}$ by a simpler model $\boldsymbol{\Phi}^{\prime}$ (see subsections below for examples of such models) results in $\mathrm{P}_{\mathbf{\Phi^\prime}}(k, 5) < \mathrm{P}_{\mathbf{\Phi}}(k, 5)$ for every $k$.
+Borrowing from the terminology of welfare economics, $\boldsymbol{\Phi}$ can be said to be uniquely Pareto-efficient \cite{black2012dictionary}.
 This observation suggests that the increase in performance afforded by the state-of-the-art model with respect to the baseline does not come at the detriment of user fairness.
@@
 \subsection*{Is the dataset large enough for evaluation? Is the system overfitting?}
-First of all, it is worth noting that the presented P@5 figure of $99.0\% \pm 1$ does not abide to a conventional training set vs.
+First of all, it is worth noting that the presented AP@5 figure of $99.0\% \pm 1$ does not abide to a conventional training set vs.
 test set paradigm, as is most often done in machine learning research.
 Rather, the LMNN algorithm is trained on all available samples ($N^{\prime}=9346$ isolated notes) with the ``consensus'' cluster graph as training objective ($\mathcal{G}=\mathcal{G}_0$).
-Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$).
+Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$). %\ja wants to delete this "for k>=1" clause
 The reason behind this choice is that, in the context of computer-assisted orchestration, the collection of audio samples has a fixed size, as it is shipped alongside the software itself.
-This is the case, for example, of Orchidea\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}}, which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
+This is the case, for example, of Orchidea,\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}} which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
 Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).

### 2020-07-19 paper/paper.tex
page 17

@@
 Rather, the LMNN algorithm is trained on all available samples ($N^{\prime}=9346$ isolated notes) with the ``consensus'' cluster graph as training objective ($\mathcal{G}=\mathcal{G}_0$).
-Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$). %\ja wants to delete this "for k>=1" clause
+Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$).
 The reason behind this choice is that, in the context of computer-assisted orchestration, the collection of audio samples has a fixed size, as it is shipped alongside the software itself.
 This is the case, for example, of Orchidea,\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}} which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
-Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available ($N=78$, see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
+Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available (see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
 
 Despite this caveat, we may adopt a ``query-by-example'' (QbE) paradigm by  partitioning the database and audio samples in half ($\frac{1}{2}N^{\prime}=4673$), training the LMNN algorithm on the first half, and querying it with samples from the other half.
-In this evaluation framework, our system reaches a P@5 of $\textbf{96.2\%} \pm 2$.
-Interestingly, querying the system with samples from the training set leads to a P@5 of $\textbf{96.5\%} \pm 2$, i.e., roughly on par with the test set.
+In this evaluation framework, our system reaches an AP@5 of $\textbf{96.2\%} \pm 2$.
+Interestingly, querying the system with samples from the training set leads to an AP@5 of $\textbf{96.5\%} \pm 2$, i.e., roughly on par with the test set.
 Thus, it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
@@
 \subsection*{Is metric learning necessary?}
-Replacing the large-margin nearest neighbors (LMNN) metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
-Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all.
+Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
+Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all. % \ja wants to elaborate
 This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
-Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
+Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search. % \ja wants to delete this sentence
+We report an average precision at rank five (AP@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
 This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.

### 2020-07-19 paper/paper.tex
page 20

@@
 
-Let us recall the full definition of joint time--frequency scattering (see \lnameref{sec:methods} section):
+Let us recall the full definition of second-order joint time--frequency scattering (see \lnameref{sec:methods} section):
 \begin{equation}
@@
 This is because the temporal wavelet transform with $\boldsymbol{\psi}_\alpha$ and the frequential wavelet transform with $\boldsymbol{\psi}_\beta$ are separated by an operation of complex modulus.
-Hence, $\boldsymbol{\psi}_\beta$ operates on a real-valued input, and flipping the sign of the scale variable $\beta$ is no longer necessary.
+Hence, $\boldsymbol{\psi}_\beta$ operates on a real-valued input.
+Because separable time--frequency scattering cannot distinguish ascending chirps from descending chirps, flipping the sign of the scale variable $\beta$ is no longer necessary.
 Moreover, separable time--frequency scattering has a lower algorithmic complexity than joint time--frequency scattering.

### 2020-07-19 paper/paper.tex
page 20

@@
 Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
-This figure is noticeably worse than joint time--frequency scattering ($99.0\% \pm 1$), all other things being equal.
+This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
 This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
@@
 Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
-Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}.
+Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis.\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
@@
 \mathrm{MFCC}(\boldsymbol{x})(t, \alpha)
-\times
 \mathrm{MFCC}(\boldsymbol{x})(t, \beta)
@@
 Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
-This explains the relative success over such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity.
+This explains the relative success of such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity.
 

### 2020-07-19 paper/paper.tex
pages 22 and 23 :)

@@
 \address[id=ccm]{%
-  \orgname{Center for Computational Mathematics, Flatiron Institute},
+  \orgname{Department of Mathematics, KTH and Center for Computational Mathematics, Flatiron Institute},
   \street{162 5th Avenue},
@@
 Accessing the sounds in these databases therefore requires some knowledge of the taxonomy and keywords used.
-Furthermore, the user needs to have some idea of the particular IPT they are searching for.
+Furthermore, the user needs to have some idea of the particular playing technique they are searching for.
 
 As an alternative, content-based searches allow the user to identify sounds based on similarity to some given query sound.
-This query-by-example approach provides an opportunity to search for sounds without having a specific instrument of PT in mind, yielding a wider range of available sounds.
+This query-by-example approach provides an opportunity to search for sounds without having a specific instrument of playing technique in mind, yielding a wider range of available sounds.
 A composer with access to such a system would therefore be able to draw on a more diverse palette of musical timbre.
@@
 The computational model proposed in this work is well suited to such a query-by-example task.
-We have shown that it is able to approximate well the timbral judgments of a wide range of subjects included in our study.
+We have shown that it is able to adequately approximate the timbral judgments of a wide range of subjects included in our study.
 Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on the reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
@@
 Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
-Though, future work should demonstrate that they work for more general classes of sounds.
+Future work will demonstrate that they work for more general classes of sounds.
 To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.

### 2020-07-19 experiments/processResults_v2.m
update Figure 6 and its caption

@@ -1,6 +1,6 @@
 addpath('libs/export_fig')
 clear all
-captions = {'Proposed', 'train/test validation', 'randomized features', 'lda', 'no learning', 'reduce T to 25ms', 'separable scattering', 'mfcc', 'mfcc monomials'};
+captions = {'Proposed', 'train/test validation', 'random features', 'LDA', 'no learning', 'reduce T to 25ms', 'separable scattering', 'MFCC', 'MFCC monomials'};
 ind = 1;
 
 iaInd = 2;
@@ -50,6 +50,7 @@ axis([-log10(custom_xticks(end)) -log10(custom_xticks(1)) 1.5 9.5])
 hold off;
 
 set(gcf, 'Color', 'None')
+colormap(hsv());
 savefig('../paper/figures/ablationStudyNoRand.fig')
 export_fig('../paper/figures/ablationStudyNoRand.png', '-transparent')
 

### 2020-07-19 paper/paper.tex
update Figure 6 and its caption
bis

@@
 \caption{\csentence{Impact of different processing architecture or protocol designs.} For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.
-The whiskers extend to the most extreme data points not considered outliers.
-Performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.
+The performance achieved for each reference clustering is depicted by a lozenge whose color is chosen arbitrarily but consistently across conditions.
 See \lnameref{sec:discussion} section for details.}

### 2020-07-21 paper/paper.tex
final comments

@@
 Mapping this input to a feature space in which distances denote timbral dissimilarity requires a data-driven stage of dimensionality reduction.
-In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of music information retrieval (MIR) encodes timbre under the form of ``audio tags'', music psychology measures timbre similarity directly from pairwise judgments.
+In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of MIR mostly encodes timbre under the form of ``audio tags'', music psychology mostly measures timbre similarity directly from pairwise \ml{similarity} judgments.
 
@@
 \end{equation}
+\ml{why 2 ?}
 Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
@@
 We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
+In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
 However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
@@
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+This is a form of semi-supervised label propagation: from a limited amount of human annotation in order to curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
 
@@
 Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
-Future work will demonstrate that they work for more general classes of sounds.
+\ml{Future work will demonstrate that they work for more general classes of sounds}.
 To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.

### 2020-07-21 paper/paper.tex
fix 2 factor in technique-similarity
comment on line 936
final comments

@@
 Mapping this input to a feature space in which distances denote timbral dissimilarity requires a data-driven stage of dimensionality reduction.
-In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of MIR mostly encodes timbre under the form of ``audio tags'', music psychology mostly measures timbre similarity directly from pairwise \ml{similarity} judgments.
+In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of MIR mostly encodes timbre under the form of ``audio tags'', music psychology mostly measures timbre similarity directly from pairwise similarity judgments.
 
@@
 \mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
-2\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
 }{
@@
 \end{equation}
-\ml{why 2 ?}
+%\ml{why 2 ?} \vl: fixed
 Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
@@
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation in order to curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+% \vl: ?
 
@@
 That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
-Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend on the type of sound being processed.
-\ml{Future work will demonstrate that they work for more general classes of sounds}.
+Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend strongly on the type of sound being processed.
+Future work will investigate whether more general classes of sounds are also well modeled.
 To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.

### 2020-07-21 paper/paper.tex
Update affiliation for JA

@@
 \author[
-   addressref={ccm},
+   addressref={kth,ccm},
    email={janden@flatironinstitute.org}
@@
 }
+\address[id=kth]{%
+  \orgname{Department of Mathematics, KTH Royal Institute of Technology}
+  \street{Lindstedtsv\"{a}gen 25}
+  \city{SE-100 44, Stockholm}
+  \cny{Sweden}
+}
 \address[id=ccm]{%
-  \orgname{Department of Mathematics, KTH and Center for Computational Mathematics, Flatiron Institute},
+  \orgname{Center for Computational Mathematics, Flatiron Institute},
   \street{162 5th Avenue},

### 2020-07-21 paper/paper.tex
Spellcheck

@@
 Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in classical and folk contexts.
-However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique, use instrument identity as a proxy for timbre quality, and do not allow for customization to the perceptual idiosyncracies of a new subject.
+However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique, use instrument identity as a proxy for timbre quality, and do not allow for customization to the perceptual idiosyncrasies of a new subject.
 In this article, we ask 31 human subjects to organize 78 isolated notes into a set of timbre clusters.
@@
 This however, comes with additional difficulties in providing a consistent scale across subjects, but could be feasible if the goal is to only adapt the timbral space to a single individual.
-An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assigments.
+An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assignments.
 
@@
 
-\section*{\nmu Acknowledgement}
+\section*{\nmu Acknowledgment}
 This work is partially supported by the European Research Council (ERC) award 320959 (InvariantClass).

### 2020-07-21 paper/paper.tex
adaption path to graphics

@@
 
+\graphicspath{{figures/}{figures/violin_scattering/}{figures/flute_scattering/}{./}}
 %%% Put your definitions there:
@@
 %  (a) & (b) \\
-%  \includegraphics[width = \textwidth]{figures/nbc.png} &
-%  \includegraphics[width = \textwidth]{figures/sbc.png}
+%  \includegraphics[width = \textwidth]{nbc.png} &
+%  \includegraphics[width = \textwidth]{sbc.png}
 %\end{tabular}
@@
 \begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{figures/lostanlen2020jasmp_fig1_v4.pdf}
+\includegraphics[width=0.9\textwidth]{lostanlen2020jasmp_fig1_v4.pdf}
 \caption{\csentence{Overview of the proposed approach.} See \lnameref{sec:intro} for details.}
@@
 \center
-\includegraphics[width = 0.85\textwidth]{figures/consensusVsI.png}
+\includegraphics[width = 0.85\textwidth]{consensusVsI.png}
 \caption{
@@
 \center
-\includegraphics[width = 0.85\textwidth]{figures/consensusVsPt.png}
+\includegraphics[width = 0.85\textwidth]{consensusVsPt.png}
 \caption{
@@
 \begin{figure}
-\includegraphics[height=4cm]{figures/violin_scattering/Vn-ord-C4-mf-4c_scalogram.pdf}
+\includegraphics[height=4cm]{Vn-ord-C4-mf-4c_scalogram.pdf}
 \hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-nonvib-C4-mf-4c_scalogram.pdf}
 \hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-trem-C4-mf-4c_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-trem-C4-mf-4c_scalogram.pdf}
 
 \hspace{2mm}
-\includegraphics[height=28mm]{figures/violin_scattering/Vn-ord-C4-mf-4c_scattering.pdf}
+\includegraphics[height=28mm]{Vn-ord-C4-mf-4c_scattering.pdf}
 \hspace{5mm}
-\includegraphics[height=28mm]{figures/violin_scattering/Vn-nonvib-C4-mf-4c_scattering.pdf}
+\includegraphics[height=28mm]{Vn-nonvib-C4-mf-4c_scattering.pdf}
 \hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/violin_scattering/Vn-trem-C4-mf-4c_scattering.pdf}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{Vn-trem-C4-mf-4c_scattering.pdf}
 
@@
 
-\includegraphics[height=4cm]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scalogram.pdf}
+\includegraphics[height=4cm]{Vn-sfz-C4-fp-4c_scalogram.pdf}
 \hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-pizz-lv-C4-pp-4c_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-pizz-lv-C4-pp-4c_scalogram.pdf}
 \hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/violin_scattering/Vn-stacc-C4-mf_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-stacc-C4-mf_scalogram.pdf}
 
 \hspace{2mm}
-\includegraphics[height=28mm]{figures/violin_scattering/Vn-sfz-C4-fp-4c_scattering.pdf}
+\includegraphics[height=28mm]{Vn-sfz-C4-fp-4c_scattering.pdf}
 \hspace{5mm}
-\includegraphics[height=28mm]{figures/violin_scattering/Vn-pizz-lv-C4-pp-4c_scattering.pdf}
+\includegraphics[height=28mm]{Vn-pizz-lv-C4-pp-4c_scattering.pdf}
 \hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/violin_scattering/Vn-stacc-C4-mf_scattering.pdf}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{Vn-stacc-C4-mf_scattering.pdf}
 
@@
 \begin{figure}
-\includegraphics[height=4cm]{figures/flute_scattering/Fl-ord-C4-mf_scalogram.pdf}
+\includegraphics[height=4cm]{Fl-ord-C4-mf_scalogram.pdf}
 \hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-trill-maj2-C4-mf_scalogram.pdf}
 \hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-ply+sng-Csharp4+C4-mf_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-ply+sng-Csharp4+C4-mf_scalogram.pdf}
 
 \hspace{2mm}
-\includegraphics[height=28mm]{figures/flute_scattering/Fl-ord-C4-mf_scattering.pdf}
+\includegraphics[height=28mm]{Fl-ord-C4-mf_scattering.pdf}
 \hspace{5mm}
-\includegraphics[height=28mm]{figures/flute_scattering/Fl-trill-maj2-C4-mf_scattering.pdf}
+\includegraphics[height=28mm]{Fl-trill-maj2-C4-mf_scattering.pdf}
 \hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/flute_scattering/Fl-ply+sng-Csharp4+C4-mf_scattering.pdf}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{Fl-ply+sng-Csharp4+C4-mf_scattering.pdf}
 
@@
 
-\includegraphics[height=4cm]{figures/flute_scattering/Fl-sfz-C4-fp_scalogram.pdf}
+\includegraphics[height=4cm]{Fl-sfz-C4-fp_scalogram.pdf}
 \hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-key-cl-C4-f_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-key-cl-C4-f_scalogram.pdf}
 \hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{figures/flute_scattering/Fl-stacc-C4-mf_scalogram.pdf}
+\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-stacc-C4-mf_scalogram.pdf}
 
 \hspace{2mm}
-\includegraphics[height=28mm]{figures/flute_scattering/Fl-sfz-C4-fp_scattering.pdf}
+\includegraphics[height=28mm]{Fl-sfz-C4-fp_scattering.pdf}
 \hspace{5mm}
-\includegraphics[height=28mm]{figures/flute_scattering/Fl-key-cl-C4-f_scattering.pdf}
+\includegraphics[height=28mm]{Fl-key-cl-C4-f_scattering.pdf}
 \hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{figures/flute_scattering/Fl-stacc-C4-mf_scattering.pdf}
+\includegraphics[trim=0 0 77 0, clip, height=28mm]{Fl-stacc-C4-mf_scattering.pdf}
 
@@
 \begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{figures/ablationStudy}
+\includegraphics[width=0.9\textwidth]{ablationStudy}
 \caption{\csentence{Impact of different processing architecture or protocol designs.} For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.

### 2020-07-21 paper/submission/paper.tex
submitted version

--- 2020-07-21 paper/paper.tex
@@
 
-\graphicspath{{figures/}{figures/violin_scattering/}{figures/flute_scattering/}{./}}
+\graphicspath{{figures/}{figures/violin_scattering}{figures/flute_scattering}{./}}
 %%% Put your definitions there:
@@
 
-\bibliographystyle{bmc-mathphys}
-\bibliography{bib}
+% \bibliographystyle{bmc-mathphys}
+% \bibliography{bib}
+
+%% BioMed_Central_Bib_Style_v1.01
+
+\begin{thebibliography}{78}
+% BibTex style file: bmc-mathphys.bst (version 2.1), 2014-07-24
+\ifx \bisbn   \undefined \def \bisbn  #1{ISBN #1}\fi
+\ifx \binits  \undefined \def \binits#1{#1}\fi
+\ifx \bauthor  \undefined \def \bauthor#1{#1}\fi
+\ifx \batitle  \undefined \def \batitle#1{#1}\fi
+\ifx \bjtitle  \undefined \def \bjtitle#1{#1}\fi
+\ifx \bvolume  \undefined \def \bvolume#1{\textbf{#1}}\fi
+\ifx \byear  \undefined \def \byear#1{#1}\fi
+\ifx \bissue  \undefined \def \bissue#1{#1}\fi
+\ifx \bfpage  \undefined \def \bfpage#1{#1}\fi
+\ifx \blpage  \undefined \def \blpage #1{#1}\fi
+\ifx \burl  \undefined \def \burl#1{\textsf{#1}}\fi
+\ifx \doiurl  \undefined \def \doiurl#1{\textsf{#1}}\fi
+\ifx \betal  \undefined \def \betal{\textit{et al.}}\fi
+\ifx \binstitute  \undefined \def \binstitute#1{#1}\fi
+\ifx \binstitutionaled  \undefined \def \binstitutionaled#1{#1}\fi
+\ifx \bctitle  \undefined \def \bctitle#1{#1}\fi
+\ifx \beditor  \undefined \def \beditor#1{#1}\fi
+\ifx \bpublisher  \undefined \def \bpublisher#1{#1}\fi
+\ifx \bbtitle  \undefined \def \bbtitle#1{#1}\fi
+\ifx \bedition  \undefined \def \bedition#1{#1}\fi
+\ifx \bseriesno  \undefined \def \bseriesno#1{#1}\fi
+\ifx \blocation  \undefined \def \blocation#1{#1}\fi
+\ifx \bsertitle  \undefined \def \bsertitle#1{#1}\fi
+\ifx \bsnm \undefined \def \bsnm#1{#1}\fi
+\ifx \bsuffix \undefined \def \bsuffix#1{#1}\fi
+\ifx \bparticle \undefined \def \bparticle#1{#1}\fi
+\ifx \barticle \undefined \def \barticle#1{#1}\fi
+\ifx \bconfdate \undefined \def \bconfdate #1{#1}\fi
+\ifx \botherref \undefined \def \botherref #1{#1}\fi
+\ifx \url \undefined \def \url#1{\textsf{#1}}\fi
+\ifx \bchapter \undefined \def \bchapter#1{#1}\fi
+\ifx \bbook \undefined \def \bbook#1{#1}\fi
+\ifx \bcomment \undefined \def \bcomment#1{#1}\fi
+\ifx \oauthor \undefined \def \oauthor#1{#1}\fi
+\ifx \citeauthoryear \undefined \def \citeauthoryear#1{#1}\fi
+\ifx \endbibitem  \undefined \def \endbibitem {}\fi
+\ifx \bconflocation  \undefined \def \bconflocation#1{#1}\fi
+\ifx \arxivurl  \undefined \def \arxivurl#1{\textsf{#1}}\fi
+\csname PreBibitemsHook\endcsname
+
+%%% 1
+\bibitem{downie2003mir}
+\begin{barticle}
+\bauthor{\bsnm{Downie}, \binits{J.S.}}:
+\batitle{Music information retrieval}.
+\bjtitle{Annual review of information science and technology}
+\bvolume{37}(\bissue{1}),
+\bfpage{295}--\blpage{340}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 2
+\bibitem{siedenburg2019chapter}
+\begin{bbook}
+\bauthor{\bsnm{Siedenburg}, \binits{K.}},
+\bauthor{\bsnm{Saitis}, \binits{C.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+In: \beditor{\bsnm{Siedenburg}, \binits{K.}},
+\beditor{\bsnm{Saitis}, \binits{C.}},
+\beditor{\bsnm{McAdams}, \binits{S.}},
+\beditor{\bsnm{Popper}, \binits{A.N.}},
+\beditor{\bsnm{Fay}, \binits{R.R.}} (eds.)
+\bbtitle{The Present, Past, and Future of Timbre Research},
+pp. \bfpage{1}--\blpage{19}.
+\bpublisher{Springer},
+\blocation{Cham}
+(\byear{2019})
+\end{bbook}
+\endbibitem
+
+%%% 3
+\bibitem{faure1996icmpc}
+\begin{bchapter}
+\bauthor{\bsnm{Faure}, \binits{A.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}},
+\bauthor{\bsnm{Nosulenko}, \binits{V.}}:
+\bctitle{Verbal correlates of perceptual dimensions of timbre}.
+In: \bbtitle{Proceedings of the International Conference on Music Perception
+  and Cognition (ICMPC)}
+(\byear{1996})
+\end{bchapter}
+\endbibitem
+
+%%% 4
+\bibitem{lostanlen2018extended}
+\begin{bchapter}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{And{\'e}n}, \binits{J.}},
+\bauthor{\bsnm{Lagrange}, \binits{M.}}:
+\bctitle{Extended playing techniques: the next milestone in musical instrument
+  recognition}.
+In: \bbtitle{Proceedings of the International Conference on Digital Libraries
+  for Musicology (DLfM)},
+pp. \bfpage{1}--\blpage{10}
+(\byear{2018}).
+\bcomment{ACM}
+\end{bchapter}
+\endbibitem
+
+%%% 5
+\bibitem{antoine2018isma}
+\begin{bchapter}
+\bauthor{\bsnm{Antoine}, \binits{A.}},
+\bauthor{\bsnm{Miranda}, \binits{E.R.}}:
+\bctitle{Musical acoustics, timbre, and computer-aided orchestration
+  challenges}.
+In: \bbtitle{Proc. ISMA}
+(\byear{2018})
+\end{bchapter}
+\endbibitem
+
+%%% 6
+\bibitem{kolozali2011ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Kolozali}, \binits{S.}},
+\bauthor{\bsnm{Barthet}, \binits{M.}},
+\bauthor{\bsnm{Fazekas}, \binits{G.}},
+\bauthor{\bsnm{Sandler}, \binits{M.B.}}:
+\bctitle{Knowledge representation issues in musical instrument ontology
+  design.}
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2011})
+\end{bchapter}
+\endbibitem
+
+%%% 7
+\bibitem{calvo2020acm}
+\begin{botherref}
+\oauthor{\bsnm{Calvo-Zaragoza}, \binits{J.}},
+\oauthor{\bsnm{Haji\v{c}~Jr.}, \binits{J.}},
+\oauthor{\bsnm{Pacha}, \binits{A.}}:
+Understanding optical music recognition.
+ACM Computing Surveys
+(2020)
+\end{botherref}
+\endbibitem
+
+%%% 8
+\bibitem{erickson1975book}
+\begin{bbook}
+\bauthor{\bsnm{Erickson}, \binits{R.}}:
+\bbtitle{Sound Structure in Music}.
+\bpublisher{University of California Press}, \blocation{???}
+(\byear{1975})
+\end{bbook}
+\endbibitem
+
+%%% 9
+\bibitem{thoret2018jasa}
+\begin{barticle}
+\bauthor{\bsnm{Thoret}, \binits{E.}},
+\bauthor{\bsnm{Caramiaux}, \binits{B.}},
+\bauthor{\bsnm{Depalle}, \binits{P.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+\batitle{Human dissimilarity ratings of musical instrument timbre: A
+  computational meta-analysis}.
+\bjtitle{Journal of the Acoustical Society of America}
+\bvolume{143}(\bissue{3}),
+\bfpage{1745}--\blpage{1746}
+(\byear{2018})
+\end{barticle}
+\endbibitem
+
+%%% 10
+\bibitem{maresz2013cmr}
+\begin{barticle}
+\bauthor{\bsnm{Maresz}, \binits{Y.}}:
+\batitle{On computer-assisted orchestration}.
+\bjtitle{Contemporary Music Review}
+\bvolume{32}(\bissue{1}),
+\bfpage{99}--\blpage{109}
+(\byear{2013})
+\end{barticle}
+\endbibitem
+
+%%% 11
+\bibitem{caetano2019swarm}
+\begin{barticle}
+\bauthor{\bsnm{Caetano}, \binits{M.}},
+\bauthor{\bsnm{Zacharakis}, \binits{A.}},
+\bauthor{\bsnm{Barbancho}, \binits{I.}},
+\bauthor{\bsnm{Tard{\'o}n}, \binits{L.J.}}:
+\batitle{Leveraging diversity in computer-aided musical orchestration with an
+  artificial immune system for multi-modal optimization}.
+\bjtitle{Swarm and Evolutionary Computation}
+\bvolume{50},
+\bfpage{100484}
+(\byear{2019})
+\end{barticle}
+\endbibitem
+
+%%% 12
+\bibitem{anden2015mlsp}
+\begin{bchapter}
+\bauthor{\bsnm{And{\'e}n}, \binits{J.}},
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\bctitle{Joint time-frequency scattering for audio classification}.
+In: \bbtitle{Proceedings of the IEEE International Workshop on Machine Learning
+  for Signal Processing (MLSP)},
+pp. \bfpage{1}--\blpage{6}
+(\byear{2015}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 13
+\bibitem{anden2019tsp}
+\begin{barticle}
+\bauthor{\bsnm{And\'{e}n}, \binits{J.}},
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\batitle{Joint time--frequency scattering}.
+\bjtitle{IEEE Transactions on Signal Processing}
+\bvolume{67}(\bissue{14}),
+\bfpage{3704}--\blpage{3718}
+(\byear{2019})
+\end{barticle}
+\endbibitem
+
+%%% 14
+\bibitem{lostanlen2018jasmp}
+\begin{barticle}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Lafay}, \binits{G.}},
+\bauthor{\bsnm{And{\'e}n}, \binits{J.}},
+\bauthor{\bsnm{Lagrange}, \binits{M.}}:
+\batitle{Relevance-based quantization of scattering features for unsupervised
+  mining of environmental audio}.
+\bjtitle{EURASIP Journal on Audio, Speech, and Music Processing}
+\bvolume{2018}(\bissue{1}),
+\bfpage{15}
+(\byear{2018})
+\end{barticle}
+\endbibitem
+
+%%% 15
+\bibitem{anden2012scattering}
+\begin{bchapter}
+\bauthor{\bsnm{And\'{e}n}, \binits{J.}},
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\bctitle{Scattering representation of modulated sounds}.
+In: \bbtitle{Proceedings of the International Conference on Digital Audio
+  Effects (DAFx)}
+(\byear{2012})
+\end{bchapter}
+\endbibitem
+
+%%% 16
+\bibitem{weinberger2009distance}
+\begin{barticle}
+\bauthor{\bsnm{Weinberger}, \binits{K.Q.}},
+\bauthor{\bsnm{Saul}, \binits{L.K.}}:
+\batitle{Distance metric learning for large margin nearest neighbor
+  classification}.
+\bjtitle{Journal of Machine Learning Research}
+\bvolume{10}(\bissue{Feb}),
+\bfpage{207}--\blpage{244}
+(\byear{2009})
+\end{barticle}
+\endbibitem
+
+%%% 17
+\bibitem{mcadams1995psychres}
+\begin{barticle}
+\bauthor{\bsnm{McAdams}, \binits{S.}},
+\bauthor{\bsnm{Winsberg}, \binits{S.}},
+\bauthor{\bsnm{Donnadieu}, \binits{S.}},
+\bauthor{\bsnm{De~Soete}, \binits{G.}},
+\bauthor{\bsnm{Krimphoff}, \binits{J.}}:
+\batitle{Perceptual scaling of synthesized musical timbres: common dimensions,
+  specificities, and latent subject classes}.
+\bjtitle{Psychological research}
+\bvolume{58}(\bissue{3}),
+\bfpage{177}--\blpage{192}
+(\byear{1995})
+\end{barticle}
+\endbibitem
+
+%%% 18
+\bibitem{patil2012ploscompbiol}
+\begin{barticle}
+\bauthor{\bsnm{Patil}, \binits{K.}},
+\bauthor{\bsnm{Pressnitzer}, \binits{D.}},
+\bauthor{\bsnm{Shamma}, \binits{S.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\batitle{Music in our ears: {The} biological bases of musical timbre
+  perception}.
+\bjtitle{PLoS Computational Biology}
+\bvolume{8}(\bissue{11}),
+\bfpage{1002759}
+(\byear{2012})
+\end{barticle}
+\endbibitem
+
+%%% 19
+\bibitem{joder2009taslp}
+\begin{barticle}
+\bauthor{\bsnm{Joder}, \binits{C.}},
+\bauthor{\bsnm{Essid}, \binits{S.}},
+\bauthor{\bsnm{Richard}, \binits{G.}}:
+\batitle{Temporal integration for audio classification with application to
+  musical instrument classification}.
+\bjtitle{IEEE Transactions on Audio, Speech, and Language Processing}
+\bvolume{17}(\bissue{1}),
+\bfpage{174}--\blpage{186}
+(\byear{2009})
+\end{barticle}
+\endbibitem
+
+%%% 20
+\bibitem{siedenburg2016jnmr}
+\begin{barticle}
+\bauthor{\bsnm{Siedenburg}, \binits{K.}},
+\bauthor{\bsnm{Fujinaga}, \binits{I.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+\batitle{A comparison of approaches to timbre descriptors in music information
+  retrieval and music psychology}.
+\bjtitle{Journal of New Music Research}
+\bvolume{45}(\bissue{1}),
+\bfpage{27}--\blpage{41}
+(\byear{2016})
+\end{barticle}
+\endbibitem
+
+%%% 21
+\bibitem{martin1998asa}
+\begin{bchapter}
+\bauthor{\bsnm{Martin}, \binits{K.D.}},
+\bauthor{\bsnm{Kim}, \binits{Y.E.}}:
+\bctitle{Musical instrument identification: A pattern recognition approach}.
+In: \bbtitle{Proceedings of the Acoustical Society of America}
+(\byear{1998})
+\end{bchapter}
+\endbibitem
+
+%%% 22
+\bibitem{brown1999jasa}
+\begin{barticle}
+\bauthor{\bsnm{Brown}, \binits{J.C.}}:
+\batitle{Computer identification of musical instruments using pattern
+  recognition with cepstral coefficients as features}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{105}(\bissue{3}),
+\bfpage{1933}--\blpage{1941}
+(\byear{1999})
+\end{barticle}
+\endbibitem
+
+%%% 23
+\bibitem{eronen2000icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Eronen}, \binits{A.}},
+\bauthor{\bsnm{Klapuri}, \binits{A.}}:
+\bctitle{Musical instrument recognition using cepstral coefficients and
+  temporal features}.
+In: \bbtitle{Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)}
+(\byear{2000})
+\end{bchapter}
+\endbibitem
+
+%%% 24
+\bibitem{herrera2003jnmr}
+\begin{barticle}
+\bauthor{\bsnm{Herrera~Boyer}, \binits{P.}},
+\bauthor{\bsnm{Peeters}, \binits{G.}},
+\bauthor{\bsnm{Dubnov}, \binits{S.}}:
+\batitle{Automatic classification of musical instrument sounds}.
+\bjtitle{Journal of New Music Research}
+\bvolume{32}(\bissue{1}),
+\bfpage{3}--\blpage{21}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 25
+\bibitem{wieczorkowska2003jiis}
+\begin{barticle}
+\bauthor{\bsnm{Wieczorkowska}, \binits{A.A.}},
+\bauthor{\bsnm{{\.Z}ytkow}, \binits{J.M.}}:
+\batitle{Analysis of feature dependencies in sound description}.
+\bjtitle{Journal of Intelligent Information Systems}
+\bvolume{20}(\bissue{3}),
+\bfpage{285}--\blpage{302}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 26
+\bibitem{livshin2004dafx}
+\begin{bchapter}
+\bauthor{\bsnm{Livshin}, \binits{A.}},
+\bauthor{\bsnm{Rodet}, \binits{X.}}:
+\bctitle{Musical instrument identification in continuous recordings}.
+In: \bbtitle{Proceedings of the International Conference on Digital Audio
+  Effects (DAFx)}
+(\byear{2004})
+\end{bchapter}
+\endbibitem
+
+%%% 27
+\bibitem{krishna2004icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Krishna}, \binits{A.G.}},
+\bauthor{\bsnm{Sreenivas}, \binits{T.V.}}:
+\bctitle{Music instrument recognition: from isolated notes to solo phrases}.
+In: \bbtitle{Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)}
+(\byear{2004})
+\end{bchapter}
+\endbibitem
+
+%%% 28
+\bibitem{kaminskyj2005jiis}
+\begin{barticle}
+\bauthor{\bsnm{Kaminskyj}, \binits{I.}},
+\bauthor{\bsnm{Czaszejko}, \binits{T.}}:
+\batitle{{Automatic recognition of isolated monophonic musical instrument
+  sounds using kNNC}}.
+\bjtitle{Journal of Intelligent Information Systems}
+\bvolume{24}(\bissue{2-3}),
+\bfpage{199}--\blpage{221}
+(\byear{2005})
+\end{barticle}
+\endbibitem
+
+%%% 29
+\bibitem{benetos2006icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Benetos}, \binits{E.}},
+\bauthor{\bsnm{Kotti}, \binits{M.}},
+\bauthor{\bsnm{Kotropoulos}, \binits{C.}}:
+\bctitle{Musical instrument classification using non-negative matrix
+  factorization algorithms and subset feature selection}.
+In: \bbtitle{Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)}
+(\byear{2006})
+\end{bchapter}
+\endbibitem
+
+%%% 30
+\bibitem{bhalke2016jiis}
+\begin{barticle}
+\bauthor{\bsnm{Bhalke}, \binits{D.G.}},
+\bauthor{\bsnm{Rao}, \binits{C.B.R.}},
+\bauthor{\bsnm{Bormane}, \binits{D.S.}}:
+\batitle{{Automatic musical instrument classification using fractional Fourier
+  transform based-MFCC features and counter propagation neural network}}.
+\bjtitle{Journal of Intelligent Information Systems}
+\bvolume{46}(\bissue{3}),
+\bfpage{425}--\blpage{446}
+(\byear{2016})
+\end{barticle}
+\endbibitem
+
+%%% 31
+\bibitem{humphrey2018ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Humphrey}, \binits{E.}},
+\bauthor{\bsnm{Durand}, \binits{S.}},
+\bauthor{\bsnm{McFee}, \binits{B.}}:
+\bctitle{Openmic-2018: an open dataset for multiple instrument recognition}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2018})
+\end{bchapter}
+\endbibitem
+
+%%% 32
+\bibitem{mcfee2016ismir}
+\begin{bchapter}
+\bauthor{\bsnm{McFee}, \binits{B.}},
+\bauthor{\bsnm{Humphrey}, \binits{E.J.}},
+\bauthor{\bsnm{Urbano}, \binits{J.}}:
+\bctitle{{A plan for sustainable {MIR} evaluation}}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2016})
+\end{bchapter}
+\endbibitem
+
+%%% 33
+\bibitem{defferrard2017ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Defferrard}, \binits{M.}},
+\bauthor{\bsnm{Benzi}, \binits{K.}},
+\bauthor{\bsnm{Vandergheynst}, \binits{P.}},
+\bauthor{\bsnm{Bresson}, \binits{X.}}:
+\bctitle{{FMA: A dataset for music analysis}}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2017})
+\end{bchapter}
+\endbibitem
+
+%%% 34
+\bibitem{lostanlen2016ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Cella}, \binits{C.E.}}:
+\bctitle{Deep convolutional networks on the pitch spiral for musical instrument
+  recognition}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2016})
+\end{bchapter}
+\endbibitem
+
+%%% 35
+\bibitem{bittner2014ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Bittner}, \binits{R.M.}},
+\bauthor{\bsnm{Salamon}, \binits{J.}},
+\bauthor{\bsnm{Tierney}, \binits{M.}},
+\bauthor{\bsnm{Mauch}, \binits{M.}},
+\bauthor{\bsnm{Cannam}, \binits{C.}},
+\bauthor{\bsnm{Bello}, \binits{J.P.}}:
+\bctitle{{MedleyDB: A multitrack dataset for annotation-intensive MIR
+  research}}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2014})
+\end{bchapter}
+\endbibitem
+
+%%% 36
+\bibitem{mcfee2015ismir}
+\begin{bchapter}
+\bauthor{\bsnm{McFee}, \binits{B.}},
+\bauthor{\bsnm{Humphrey}, \binits{E.}},
+\bauthor{\bsnm{Bello}, \binits{J.}}:
+\bctitle{A software framework for musical data augmentation}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR)}
+(\byear{2015})
+\end{bchapter}
+\endbibitem
+
+%%% 37
+\bibitem{pons2017eusipco}
+\begin{bchapter}
+\bauthor{\bsnm{{Pons}}, \binits{J.}},
+\bauthor{\bsnm{{Slizovskaia}}, \binits{O.}},
+\bauthor{\bsnm{{Gong}}, \binits{R.}},
+\bauthor{\bsnm{{G\'{o}mez}}, \binits{E.}},
+\bauthor{\bsnm{{Serra}}, \binits{X.}}:
+\bctitle{Timbre analysis of music audio signals with convolutional neural
+  networks}.
+In: \bbtitle{25th European Signal Processing Conference (EUSIPCO)},
+pp. \bfpage{2744}--\blpage{2748}
+(\byear{2017})
+\end{bchapter}
+\endbibitem
+
+%%% 38
+\bibitem{gururani2018ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Gururani}, \binits{S.}},
+\bauthor{\bsnm{Summers}, \binits{C.}},
+\bauthor{\bsnm{Lerch}, \binits{A.}}:
+\bctitle{Instrument activity detection in polyphonic music using deep neural
+  networks.}
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2018})
+\end{bchapter}
+\endbibitem
+
+%%% 39
+\bibitem{loureiro2004ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Loureiro}, \binits{M.A.}},
+\bauthor{\bparticle{de} \bsnm{Paula}, \binits{H.B.}},
+\bauthor{\bsnm{Yehia}, \binits{H.C.}}:
+\bctitle{Timbre classification of a single musical instrument}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference}
+(\byear{2004})
+\end{bchapter}
+\endbibitem
+
+%%% 40
+\bibitem{han2017taslp}
+\begin{barticle}
+\bauthor{\bsnm{Han}, \binits{Y.}},
+\bauthor{\bsnm{Kim}, \binits{J.}},
+\bauthor{\bsnm{Lee}, \binits{K.}}:
+\batitle{Deep convolutional neural networks for predominant instrument
+  recognition in polyphonic music}.
+\bjtitle{IEEE Transactions on Audio, Speech, and Language Processing}
+\bvolume{25}(\bissue{1}),
+\bfpage{208}--\blpage{221}
+(\byear{2017})
+\end{barticle}
+\endbibitem
+
+%%% 41
+\bibitem{mcadams2009chapter}
+\begin{bbook}
+\bauthor{\bsnm{McAdams}, \binits{S.}},
+\bauthor{\bsnm{Giordano}, \binits{B.L.}}:
+\bbtitle{The perception of musical timbre},
+pp. \bfpage{72}--\blpage{80}
+(\byear{2009})
+\end{bbook}
+\endbibitem
+
+%%% 42
+\bibitem{depireux2001jneur}
+\begin{barticle}
+\bauthor{\bsnm{Depireux}, \binits{D.}},
+\bauthor{\bsnm{Simon}, \binits{J.}},
+\bauthor{\bsnm{Klein}, \binits{D.}},
+\bauthor{\bsnm{Shamma}, \binits{S.}}:
+\batitle{Spectro-temporal response field characterization with dynamic ripples
+  in ferret primary auditory cortex}.
+\bjtitle{Journal of Neurophysiology}
+\bvolume{85}(\bissue{3}),
+\bfpage{1220}--\blpage{1234}
+(\byear{2001})
+\end{barticle}
+\endbibitem
+
+%%% 43
+\bibitem{aertsen1981biolcyb}
+\begin{barticle}
+\bauthor{\bsnm{Aertsen}, \binits{A.}},
+\bauthor{\bsnm{Johannesma}, \binits{P.}}:
+\batitle{The spectro-temporal receptive field}.
+\bjtitle{Biological Cybernetics}
+\bvolume{42}(\bissue{2}),
+\bfpage{133}--\blpage{143}
+(\byear{1981})
+\end{barticle}
+\endbibitem
+
+%%% 44
+\bibitem{deboer1968biomed}
+\begin{botherref}
+\oauthor{\bsnm{De~Boer}, \binits{E.}},
+\oauthor{\bsnm{Kuyper}, \binits{P.}}:
+Triggered correlation.
+IEEE Transactions on Biomedical Engineering
+(3),
+169--179
+(1968)
+\end{botherref}
+\endbibitem
+
+%%% 45
+\bibitem{flandrin1998book}
+\begin{bbook}
+\bauthor{\bsnm{Flandrin}, \binits{P.}}:
+\bbtitle{Time-frequency/time-scale Analysis}.
+\bpublisher{Academic press}, \blocation{???}
+(\byear{1998})
+\end{bbook}
+\endbibitem
+
+%%% 46
+\bibitem{eggermont1993hearing}
+\begin{barticle}
+\bauthor{\bsnm{Eggermont}, \binits{J.}}:
+\batitle{Wiener and volterra analyses applied to the auditory system}.
+\bjtitle{Hearing research}
+\bvolume{66}(\bissue{2}),
+\bfpage{177}--\blpage{201}
+(\byear{1993})
+\end{barticle}
+\endbibitem
+
+%%% 47
+\bibitem{klein2000compneur}
+\begin{barticle}
+\bauthor{\bsnm{Klein}, \binits{D.}},
+\bauthor{\bsnm{Depireux}, \binits{D.}},
+\bauthor{\bsnm{Simon}, \binits{J.}},
+\bauthor{\bsnm{Shamma}, \binits{S.}}:
+\batitle{Robust spectrotemporal reverse correlation for the auditory system:
+  optimizing stimulus design}.
+\bjtitle{Journal of Computational Neuroscience}
+\bvolume{9}(\bissue{1}),
+\bfpage{85}--\blpage{111}
+(\byear{2000})
+\end{barticle}
+\endbibitem
+
+%%% 48
+\bibitem{theunissen2000jneur}
+\begin{barticle}
+\bauthor{\bsnm{Theunissen}, \binits{F.E.}},
+\bauthor{\bsnm{Sen}, \binits{K.}},
+\bauthor{\bsnm{Doupe}, \binits{A.J.}}:
+\batitle{Spectral-temporal receptive fields of nonlinear auditory neurons
+  obtained using natural sounds}.
+\bjtitle{Journal of Neuroscience}
+\bvolume{20}(\bissue{6}),
+\bfpage{2315}--\blpage{2331}
+(\byear{2000})
+\end{barticle}
+\endbibitem
+
+%%% 49
+\bibitem{chi2005jasa}
+\begin{barticle}
+\bauthor{\bsnm{Chi}, \binits{T.}},
+\bauthor{\bsnm{Ru}, \binits{P.}},
+\bauthor{\bsnm{Shamma}, \binits{S.A.}}:
+\batitle{Multiresolution spectrotemporal analysis of complex sounds}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{118}(\bissue{2}),
+\bfpage{887}--\blpage{906}
+(\byear{2005})
+\end{barticle}
+\endbibitem
+
+%%% 50
+\bibitem{patil2015eurasip}
+\begin{barticle}
+\bauthor{\bsnm{Patil}, \binits{K.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\batitle{Biomimetic spectro-temporal features for music instrument recognition
+  in isolated notes and solo phrases}.
+\bjtitle{EURASIP J. Audio Speech Music Process.}
+\bvolume{2015}(\bissue{1}),
+\bfpage{27}
+(\byear{2015})
+\end{barticle}
+\endbibitem
+
+%%% 51
+\bibitem{thoret2016jasa}
+\begin{barticle}
+\bauthor{\bsnm{Thoret}, \binits{E.}},
+\bauthor{\bsnm{Depalle}, \binits{P.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+\batitle{Perceptually salient spectrotemporal modulations for recognition of
+  sustained musical instruments}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{140}(\bissue{6}),
+\bfpage{478}--\blpage{483}
+(\byear{2016})
+\end{barticle}
+\endbibitem
+
+%%% 52
+\bibitem{mishra2018ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Mishra}, \binits{S.}},
+\bauthor{\bsnm{Sturm}, \binits{B.L.}},
+\bauthor{\bsnm{Dixon}, \binits{S.}}:
+\bctitle{Understanding a deep machine listening model through feature
+  inversion}.
+In: \bbtitle{Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference},
+pp. \bfpage{755}--\blpage{762}
+(\byear{2018})
+\end{bchapter}
+\endbibitem
+
+%%% 53
+\bibitem{hemery2015frontiers}
+\begin{botherref}
+\oauthor{\bsnm{Hemery}, \binits{E.}},
+\oauthor{\bsnm{Aucouturier}, \binits{J.}}:
+One hundred ways to process time, frequency, rate and scale in the central
+  auditory system: a pattern-recognition meta-analysis.
+Frontiers in computational neuroscience
+\textbf{9}
+(2015)
+\end{botherref}
+\endbibitem
+
+%%% 54
+\bibitem{andreux2020jmlr}
+\begin{botherref}
+\oauthor{\bsnm{Andreux}, \binits{M.}},
+\oauthor{\bsnm{Angles}, \binits{T.}},
+\oauthor{\bsnm{Exarchakis}, \binits{G.}},
+\oauthor{\bsnm{Leonarduzzi}, \binits{R.}},
+\oauthor{\bsnm{Rochette}, \binits{G.}},
+\oauthor{\bsnm{Thiry}, \binits{L.}},
+\oauthor{\bsnm{Zarka}, \binits{J.}},
+\oauthor{\bsnm{Mallat}, \binits{S.}},
+\oauthor{\bsnm{Belilovsky}, \binits{E.}},
+\oauthor{\bsnm{Bruna}, \binits{J.}}, et al.:
+Kymatio: Scattering transforms in python.
+Journal of Machine Learning Research
+\textbf{21}(60)
+(2020)
+\end{botherref}
+\endbibitem
+
+%%% 55
+\bibitem{lostanlen2019dafx}
+\begin{bchapter}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Hecker}, \binits{F.}}:
+\bctitle{The shape of remixxxes to come: Audio texture synthesis with
+  time--frequency scattering}.
+In: \bbtitle{Proceedings of the Digital Audio Effects Conference (DAFX)}
+(\byear{2019})
+\end{bchapter}
+\endbibitem
+
+%%% 56
+\bibitem{mallat2016philtrans}
+\begin{barticle}
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\batitle{Understanding deep convolutional networks}.
+\bjtitle{Philosophical Transactions of the Royal Society A: Mathematical,
+  Physical and Engineering Sciences}
+\bvolume{374}(\bissue{2065}),
+\bfpage{20150203}
+(\byear{2016})
+\end{barticle}
+\endbibitem
+
+%%% 57
+\bibitem{wu2018taslp}
+\begin{barticle}
+\bauthor{\bsnm{Wu}, \binits{C.-W.}},
+\bauthor{\bsnm{Dittmar}, \binits{C.}},
+\bauthor{\bsnm{Southall}, \binits{C.}},
+\bauthor{\bsnm{Vogl}, \binits{R.}},
+\bauthor{\bsnm{Widmer}, \binits{G.}},
+\bauthor{\bsnm{Hockman}, \binits{J.}},
+\bauthor{\bsnm{Muller}, \binits{M.}},
+\bauthor{\bsnm{Lerch}, \binits{A.}}:
+\batitle{A review of automatic drum transcription}.
+\bjtitle{IEEE Transactions on Audio, Speech, and Language Processing}
+\bvolume{26}(\bissue{9}),
+\bfpage{1457}--\blpage{1483}
+(\byear{2018})
+\end{barticle}
+\endbibitem
+
+%%% 58
+\bibitem{pearce2019appliedsciences}
+\begin{barticle}
+\bauthor{\bsnm{Pearce}, \binits{A.}},
+\bauthor{\bsnm{Brookes}, \binits{T.}},
+\bauthor{\bsnm{Mason}, \binits{R.}}:
+\batitle{Modelling timbral hardness}.
+\bjtitle{Applied Sciences}
+\bvolume{9}(\bissue{3}),
+\bfpage{466}
+(\byear{2019})
+\end{barticle}
+\endbibitem
+
+%%% 59
+\bibitem{kernighan1970efficient}
+\begin{barticle}
+\bauthor{\bsnm{Kernighan}, \binits{B.W.}},
+\bauthor{\bsnm{Lin}, \binits{S.}}:
+\batitle{An efficient heuristic procedure for partitioning graphs}.
+\bjtitle{The Bell system technical journal}
+\bvolume{49}(\bissue{2}),
+\bfpage{291}--\blpage{307}
+(\byear{1970})
+\end{barticle}
+\endbibitem
+
+%%% 60
+\bibitem{han1997scalable}
+\begin{bbook}
+\bauthor{\bsnm{Han}, \binits{E.-H.}},
+\bauthor{\bsnm{Karypis}, \binits{G.}},
+\bauthor{\bsnm{Kumar}, \binits{V.}}:
+\bbtitle{Scalable Parallel Data Mining for Association Rules}
+vol. \bseriesno{26}.
+\bpublisher{ACM}, \blocation{???}
+(\byear{1997})
+\end{bbook}
+\endbibitem
+
+%%% 61
+\bibitem{strehl2002cluster}
+\begin{barticle}
+\bauthor{\bsnm{Strehl}, \binits{A.}},
+\bauthor{\bsnm{Ghosh}, \binits{J.}}:
+\batitle{Cluster ensembles---a knowledge reuse framework for combining multiple
+  partitions}.
+\bjtitle{Journal of machine learning research}
+\bvolume{3}(\bissue{Dec}),
+\bfpage{583}--\blpage{617}
+(\byear{2002})
+\end{barticle}
+\endbibitem
+
+%%% 62
+\bibitem{ward1963jasa}
+\begin{barticle}
+\bauthor{\bsnm{Ward~Jr.}, \binits{J.H.}}:
+\batitle{Hierarchical grouping to optimize an objective function}.
+\bjtitle{Journal of the American Statistical Association}
+\bvolume{58}(\bissue{301}),
+\bfpage{236}--\blpage{244}
+(\byear{1963})
+\end{barticle}
+\endbibitem
+
+%%% 63
+\bibitem{schorkhuber2010smc}
+\begin{bchapter}
+\bauthor{\bsnm{Sch{\"o}rkhuber}, \binits{C.}},
+\bauthor{\bsnm{Klapuri}, \binits{A.}}:
+\bctitle{Constant-q transform toolbox for music processing}.
+In: \bbtitle{Proceedings of the Sound and Music Computing (SMC) Conference}
+(\byear{2010})
+\end{bchapter}
+\endbibitem
+
+%%% 64
+\bibitem{lostanlen2020icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Sridhar}, \binits{S.}},
+\bauthor{\bsnm{Farnsworth}, \binits{A.}},
+\bauthor{\bsnm{Bello}, \binits{J.P.}}:
+\bctitle{Learning the helix topology of musical pitch}.
+In: \bbtitle{Proceedings of the International Conference on Acoustics, Speech,
+  and Signal Processing (ICASSP)}
+(\byear{2020})
+\end{bchapter}
+\endbibitem
+
+%%% 65
+\bibitem{mallat2012cpam}
+\begin{barticle}
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\batitle{Group invariant scattering}.
+\bjtitle{Communications in Pure and Applied Mathematics}
+\bvolume{65}(\bissue{10}),
+\bfpage{1331}--\blpage{1398}
+(\byear{2012})
+\end{barticle}
+\endbibitem
+
+%%% 66
+\bibitem{elhilali2003speechcomm}
+\begin{barticle}
+\bauthor{\bsnm{Elhilali}, \binits{M.}},
+\bauthor{\bsnm{Chi}, \binits{T.}},
+\bauthor{\bsnm{Shamma}, \binits{S.A.}}:
+\batitle{A spectro-temporal modulation index (stmi) for assessment of speech
+  intelligibility}.
+\bjtitle{Speech communication}
+\bvolume{41}(\bissue{2-3}),
+\bfpage{331}--\blpage{348}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 67
+\bibitem{bellur2015ciss}
+\begin{bchapter}
+\bauthor{\bsnm{Bellur}, \binits{A.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\bctitle{Detection of speech tokens in noise using adaptive spectrotemporal
+  receptive fields}.
+In: \bbtitle{Proceedings of the Annual Conference on Information Sciences and
+  Systems (CISS)},
+pp. \bfpage{1}--\blpage{6}
+(\byear{2015}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 68
+\bibitem{emmanouilidou2012embs}
+\begin{bchapter}
+\bauthor{\bsnm{Emmanouilidou}, \binits{D.}},
+\bauthor{\bsnm{Patil}, \binits{K.}},
+\bauthor{\bsnm{West}, \binits{J.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\bctitle{A multiresolution analysis for detection of abnormal lung sounds}.
+In: \bbtitle{Proceedings of the International Conference of the IEEE
+  Engineering in Medicine and Biology Society (EMBS)},
+pp. \bfpage{3139}--\blpage{3142}
+(\byear{2012}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 69
+\bibitem{jansen2018icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Jansen}, \binits{A.}},
+\bauthor{\bsnm{Plakal}, \binits{M.}},
+\bauthor{\bsnm{Pandya}, \binits{R.}},
+\bauthor{\bsnm{Ellis}, \binits{D.P.}},
+\bauthor{\bsnm{Hershey}, \binits{S.}},
+\bauthor{\bsnm{Liu}, \binits{J.}},
+\bauthor{\bsnm{Moore}, \binits{R.C.}},
+\bauthor{\bsnm{Saurous}, \binits{R.A.}}:
+\bctitle{Unsupervised learning of semantic audio representations}.
+In: \bbtitle{Proceedings of the IEEE International Conference on Acoustics,
+  Speech and Signal Processing (ICASSP)},
+pp. \bfpage{126}--\blpage{130}
+(\byear{2018}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 70
+\bibitem{bellet2015book}
+\begin{bbook}
+\bauthor{\bsnm{Bellet}, \binits{A.}},
+\bauthor{\bsnm{Habrard}, \binits{A.}},
+\bauthor{\bsnm{Sebban}, \binits{M.}}:
+\bbtitle{{M}etric {L}earning}.
+\bpublisher{{M}organ \& {C}laypool {P}ublishers}, \blocation{???}
+(\byear{2015})
+\end{bbook}
+\endbibitem
+
+%%% 71
+\bibitem{wang2020icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Wang}, \binits{C.}},
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Benetos}, \binits{E.}},
+\bauthor{\bsnm{Chew}, \binits{E.}}:
+\bctitle{Playing technique recognition by joint time--frequency scattering}.
+(\byear{2020}).
+\bcomment{Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)}
+\end{bchapter}
+\endbibitem
+
+%%% 72
+\bibitem{black2012dictionary}
+\begin{bbook}
+\bauthor{\bsnm{Black}, \binits{J.}},
+\bauthor{\bsnm{Hashimzade}, \binits{N.}},
+\bauthor{\bsnm{Myles}, \binits{G.}}:
+\bbtitle{A Dictionary of Economics}.
+\bpublisher{Oxford university press}, \blocation{???}
+(\byear{2012})
+\end{bbook}
+\endbibitem
+
+%%% 73
+\bibitem{cella2020icmc}
+\begin{bchapter}
+\bauthor{\bsnm{Cella}, \binits{C.-E.}},
+\bauthor{\bsnm{Ghisi}, \binits{D.}},
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{L{\'e}vy}, \binits{F.}},
+\bauthor{\bsnm{Fineberg}, \binits{J.}},
+\bauthor{\bsnm{Maresz}, \binits{Y.}}:
+\bctitle{Orchideasol: A dataset of extended instrumental techniques for
+  computer-aided orchestration}.
+In: \bbtitle{Proceedings of the International Computer Music Conference (ICMC)}
+(\byear{2020})
+\end{bchapter}
+\endbibitem
+
+%%% 74
+\bibitem{lostanlen2019chapter}
+\begin{bchapter}
+\bauthor{\bsnm{Lostanlen}, \binits{V.}}:
+\bctitle{On time-frequency scattering and computer music}.
+In: \beditor{\bsnm{Schafhausen}, \binits{N.}},
+\beditor{\bsnm{M\"{u}ller}, \binits{V.J.}} (eds.)
+\bbtitle{Florian Hecker: Halluzination, Perspektive, Synthese}.
+\bpublisher{Sternberg Press},
+\blocation{Berlin}
+(\byear{2019})
+\end{bchapter}
+\endbibitem
+
+%%% 75
+\bibitem{bauge2013icassp}
+\begin{bchapter}
+\bauthor{\bsnm{Baug{\'e}}, \binits{C.}},
+\bauthor{\bsnm{Lagrange}, \binits{M.}},
+\bauthor{\bsnm{And{\'e}n}, \binits{J.}},
+\bauthor{\bsnm{Mallat}, \binits{S.}}:
+\bctitle{Representing environmental sounds using the separable scattering
+  transform}.
+In: \bbtitle{Proceedings of the IEEE International Conference on Acoustics,
+  Speech and Signal Processing},
+pp. \bfpage{8667}--\blpage{8671}
+(\byear{2013}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 76
+\bibitem{wang2019ismir}
+\begin{bchapter}
+\bauthor{\bsnm{Wang}, \binits{C.}},
+\bauthor{\bsnm{Benetos}, \binits{E.}},
+\bauthor{\bsnm{Lostanlen}, \binits{V.}},
+\bauthor{\bsnm{Chew}, \binits{E.}}:
+\bctitle{Adaptive time--frequency scattering for periodic modulation
+  recognition in music signals}.
+(\byear{2019}).
+\bcomment{Proceedings of the International Society for Music Information
+  Retrieval (ISMIR) Conference}
+\end{bchapter}
+\endbibitem
+
+%%% 77
+\bibitem{schadler2015jasa}
+\begin{barticle}
+\bauthor{\bsnm{Sch{\"a}dler}, \binits{M.R.}},
+\bauthor{\bsnm{Kollmeier}, \binits{B.}}:
+\batitle{Separable spectro-temporal gabor filter bank features: Reducing the
+  complexity of robust features for automatic speech recognition}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{137}(\bissue{4}),
+\bfpage{2047}--\blpage{2059}
+(\byear{2015})
+\end{barticle}
+\endbibitem
+
+%%% 78
+\bibitem{chang2010jmlr}
+\begin{barticle}
+\bauthor{\bsnm{Chang}, \binits{Y.-W.}},
+\bauthor{\bsnm{Hsieh}, \binits{C.-J.}},
+\bauthor{\bsnm{Chang}, \binits{K.-W.}},
+\bauthor{\bsnm{Ringgaard}, \binits{M.}},
+\bauthor{\bsnm{Lin}, \binits{C.-J.}}:
+\batitle{Training and testing low-degree polynomial data mappings via linear
+  svm}.
+\bjtitle{Journal of Machine Learning Research}
+\bvolume{11}(\bissue{Apr}),
+\bfpage{1471}--\blpage{1490}
+(\byear{2010})
+\end{barticle}
+\endbibitem
+
+\end{thebibliography}
+
+\newcommand{\BMCxmlcomment}[1]{}
+
+\BMCxmlcomment{
+
+<refgrp>
+
+<bibl id="B1">
+  <title><p>Music information retrieval</p></title>
+  <aug>
+    <au><snm>Downie</snm><fnm>JS</fnm></au>
+  </aug>
+  <source>Annual review of information science and technology</source>
+  <publisher>Wiley Online Library</publisher>
+  <pubdate>2003</pubdate>
+  <volume>37</volume>
+  <issue>1</issue>
+  <fpage>295</fpage>
+  <lpage>-340</lpage>
+</bibl>
+
+<bibl id="B2">
+  <title><p>The Present, Past, and Future of Timbre Research</p></title>
+  <aug>
+    <au><snm>Siedenburg</snm><fnm>K</fnm></au>
+    <au><snm>Saitis</snm><fnm>C</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Timbre: Acoustics, Perception, and Cognition</source>
+  <publisher>Cham: Springer International Publishing</publisher>
+  <editor>Siedenburg, Kai and Saitis, Charalampos and McAdams, Stephen and
+  Popper, Arthur N. and Fay, Richard R.</editor>
+  <pubdate>2019</pubdate>
+  <fpage>1</fpage>
+  <lpage>-19</lpage>
+</bibl>
+
+<bibl id="B3">
+  <title><p>Verbal correlates of perceptual dimensions of timbre</p></title>
+  <aug>
+    <au><snm>Faure</snm><fnm>A</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+    <au><snm>Nosulenko</snm><fnm>V</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference on Music Perception and
+  Cognition (ICMPC)</source>
+  <pubdate>1996</pubdate>
+</bibl>
+
+<bibl id="B4">
+  <title><p>Extended playing techniques: the next milestone in musical
+  instrument recognition</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>And{\'e}n</snm><fnm>J</fnm></au>
+    <au><snm>Lagrange</snm><fnm>M</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference on Digital Libraries for
+  Musicology (DLfM)</source>
+  <pubdate>2018</pubdate>
+  <fpage>1</fpage>
+  <lpage>-10</lpage>
+</bibl>
+
+<bibl id="B5">
+  <title><p>Musical Acoustics, Timbre, and Computer-Aided Orchestration
+  Challenges</p></title>
+  <aug>
+    <au><snm>Antoine</snm><fnm>A</fnm></au>
+    <au><snm>Miranda</snm><fnm>ER</fnm></au>
+  </aug>
+  <source>Proc. ISMA</source>
+  <pubdate>2018</pubdate>
+</bibl>
+
+<bibl id="B6">
+  <title><p>Knowledge Representation Issues in Musical Instrument Ontology
+  Design.</p></title>
+  <aug>
+    <au><snm>Kolozali</snm><fnm>S</fnm></au>
+    <au><snm>Barthet</snm><fnm>M</fnm></au>
+    <au><snm>Fazekas</snm><fnm>G</fnm></au>
+    <au><snm>Sandler</snm><fnm>MB</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2011</pubdate>
+</bibl>
+
+<bibl id="B7">
+  <title><p>Understanding optical music recognition</p></title>
+  <aug>
+    <au><snm>Calvo Zaragoza</snm><fnm>J</fnm></au>
+    <au><snm>Haji\v{c} Jr.</snm><fnm>J</fnm></au>
+    <au><snm>Pacha</snm><fnm>A</fnm></au>
+  </aug>
+  <source>ACM Computing Surveys</source>
+  <pubdate>2020</pubdate>
+</bibl>
+
+<bibl id="B8">
+  <title><p>Sound structure in music</p></title>
+  <aug>
+    <au><snm>Erickson</snm><fnm>R</fnm></au>
+  </aug>
+  <publisher>University of California Press</publisher>
+  <pubdate>1975</pubdate>
+</bibl>
+
+<bibl id="B9">
+  <title><p>Human dissimilarity ratings of musical instrument timbre: A
+  computational meta-analysis</p></title>
+  <aug>
+    <au><snm>Thoret</snm><fnm>E</fnm></au>
+    <au><snm>Caramiaux</snm><fnm>B</fnm></au>
+    <au><snm>Depalle</snm><fnm>P</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Journal of the Acoustical Society of America</source>
+  <publisher>ASA</publisher>
+  <pubdate>2018</pubdate>
+  <volume>143</volume>
+  <issue>3</issue>
+  <fpage>1745</fpage>
+  <lpage>-1746</lpage>
+</bibl>
+
+<bibl id="B10">
+  <title><p>On computer-assisted orchestration</p></title>
+  <aug>
+    <au><snm>Maresz</snm><fnm>Y</fnm></au>
+  </aug>
+  <source>Contemporary Music Review</source>
+  <pubdate>2013</pubdate>
+  <volume>32</volume>
+  <issue>1</issue>
+  <fpage>99</fpage>
+  <lpage>-109</lpage>
+</bibl>
+
+<bibl id="B11">
+  <title><p>Leveraging diversity in computer-aided musical orchestration with
+  an artificial immune system for multi-modal optimization</p></title>
+  <aug>
+    <au><snm>Caetano</snm><fnm>M</fnm></au>
+    <au><snm>Zacharakis</snm><fnm>A</fnm></au>
+    <au><snm>Barbancho</snm><fnm>I</fnm></au>
+    <au><snm>Tard{\'o}n</snm><fnm>LJ</fnm></au>
+  </aug>
+  <source>Swarm and Evolutionary Computation</source>
+  <publisher>Elsevier</publisher>
+  <pubdate>2019</pubdate>
+  <volume>50</volume>
+  <fpage>100484</fpage>
+</bibl>
+
+<bibl id="B12">
+  <title><p>Joint time-frequency scattering for audio
+  classification</p></title>
+  <aug>
+    <au><snm>And{\'e}n</snm><fnm>J</fnm></au>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Workshop on Machine Learning
+  for Signal Processing (MLSP)</source>
+  <pubdate>2015</pubdate>
+  <fpage>1</fpage>
+  <lpage>-6</lpage>
+</bibl>
+
+<bibl id="B13">
+  <title><p>Joint Time--Frequency Scattering</p></title>
+  <aug>
+    <au><snm>And\'{e}n</snm><fnm>J</fnm></au>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>IEEE Transactions on Signal Processing</source>
+  <pubdate>2019</pubdate>
+  <volume>67</volume>
+  <issue>14</issue>
+  <fpage>3704</fpage>
+  <lpage>-3718</lpage>
+</bibl>
+
+<bibl id="B14">
+  <title><p>Relevance-based quantization of scattering features for
+  unsupervised mining of environmental audio</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Lafay</snm><fnm>G</fnm></au>
+    <au><snm>And{\'e}n</snm><fnm>J</fnm></au>
+    <au><snm>Lagrange</snm><fnm>M</fnm></au>
+  </aug>
+  <source>EURASIP Journal on Audio, Speech, and Music Processing</source>
+  <pubdate>2018</pubdate>
+  <volume>2018</volume>
+  <issue>1</issue>
+  <fpage>15</fpage>
+</bibl>
+
+<bibl id="B15">
+  <title><p>Scattering Representation of Modulated Sounds</p></title>
+  <aug>
+    <au><snm>And\'{e}n</snm><fnm>J</fnm></au>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference on Digital Audio Effects
+  (DAFx)</source>
+  <pubdate>2012</pubdate>
+</bibl>
+
+<bibl id="B16">
+  <title><p>Distance metric learning for large margin nearest neighbor
+  classification</p></title>
+  <aug>
+    <au><snm>Weinberger</snm><fnm>KQ</fnm></au>
+    <au><snm>Saul</snm><fnm>LK</fnm></au>
+  </aug>
+  <source>Journal of Machine Learning Research</source>
+  <pubdate>2009</pubdate>
+  <volume>10</volume>
+  <issue>Feb</issue>
+  <fpage>207</fpage>
+  <lpage>-244</lpage>
+</bibl>
+
+<bibl id="B17">
+  <title><p>Perceptual scaling of synthesized musical timbres: common
+  dimensions, specificities, and latent subject classes</p></title>
+  <aug>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+    <au><snm>Winsberg</snm><fnm>S</fnm></au>
+    <au><snm>Donnadieu</snm><fnm>S</fnm></au>
+    <au><snm>De Soete</snm><fnm>G</fnm></au>
+    <au><snm>Krimphoff</snm><fnm>J</fnm></au>
+  </aug>
+  <source>Psychological research</source>
+  <publisher>Springer</publisher>
+  <pubdate>1995</pubdate>
+  <volume>58</volume>
+  <issue>3</issue>
+  <fpage>177</fpage>
+  <lpage>-192</lpage>
+</bibl>
+
+<bibl id="B18">
+  <title><p>Music in our ears: {The} biological bases of musical timbre
+  perception</p></title>
+  <aug>
+    <au><snm>Patil</snm><fnm>K</fnm></au>
+    <au><snm>Pressnitzer</snm><fnm>D</fnm></au>
+    <au><snm>Shamma</snm><fnm>S</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>PLoS Computational Biology</source>
+  <publisher>Public Library of Science</publisher>
+  <pubdate>2012</pubdate>
+  <volume>8</volume>
+  <issue>11</issue>
+  <fpage>e1002759</fpage>
+</bibl>
+
+<bibl id="B19">
+  <title><p>Temporal integration for audio classification with application to
+  musical instrument classification</p></title>
+  <aug>
+    <au><snm>Joder</snm><fnm>C</fnm></au>
+    <au><snm>Essid</snm><fnm>S</fnm></au>
+    <au><snm>Richard</snm><fnm>G</fnm></au>
+  </aug>
+  <source>IEEE Transactions on Audio, Speech, and Language Processing</source>
+  <publisher>IEEE</publisher>
+  <pubdate>2009</pubdate>
+  <volume>17</volume>
+  <issue>1</issue>
+  <fpage>174</fpage>
+  <lpage>-186</lpage>
+</bibl>
+
+<bibl id="B20">
+  <title><p>A comparison of approaches to timbre descriptors in music
+  information retrieval and music psychology</p></title>
+  <aug>
+    <au><snm>Siedenburg</snm><fnm>K</fnm></au>
+    <au><snm>Fujinaga</snm><fnm>I</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Journal of New Music Research</source>
+  <publisher>Taylor \& Francis</publisher>
+  <pubdate>2016</pubdate>
+  <volume>45</volume>
+  <issue>1</issue>
+  <fpage>27</fpage>
+  <lpage>-41</lpage>
+</bibl>
+
+<bibl id="B21">
+  <title><p>Musical instrument identification: A pattern recognition
+  approach</p></title>
+  <aug>
+    <au><snm>Martin</snm><fnm>KD</fnm></au>
+    <au><snm>Kim</snm><fnm>YE</fnm></au>
+  </aug>
+  <source>Proceedings of the Acoustical Society of America</source>
+  <pubdate>1998</pubdate>
+</bibl>
+
+<bibl id="B22">
+  <title><p>Computer identification of musical instruments using pattern
+  recognition with cepstral coefficients as features</p></title>
+  <aug>
+    <au><snm>Brown</snm><fnm>JC</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>ASA</publisher>
+  <pubdate>1999</pubdate>
+  <volume>105</volume>
+  <issue>3</issue>
+  <fpage>1933</fpage>
+  <lpage>-1941</lpage>
+</bibl>
+
+<bibl id="B23">
+  <title><p>Musical instrument recognition using cepstral coefficients and
+  temporal features</p></title>
+  <aug>
+    <au><snm>Eronen</snm><fnm>A</fnm></au>
+    <au><snm>Klapuri</snm><fnm>A</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)</source>
+  <pubdate>2000</pubdate>
+</bibl>
+
+<bibl id="B24">
+  <title><p>Automatic classification of musical instrument sounds</p></title>
+  <aug>
+    <au><snm>Herrera Boyer</snm><fnm>P</fnm></au>
+    <au><snm>Peeters</snm><fnm>G</fnm></au>
+    <au><snm>Dubnov</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Journal of New Music Research</source>
+  <publisher>Taylor \& Francis</publisher>
+  <pubdate>2003</pubdate>
+  <volume>32</volume>
+  <issue>1</issue>
+  <fpage>3</fpage>
+  <lpage>-21</lpage>
+</bibl>
+
+<bibl id="B25">
+  <title><p>Analysis of feature dependencies in sound description</p></title>
+  <aug>
+    <au><snm>Wieczorkowska</snm><fnm>AA</fnm></au>
+    <au><snm>{\.Z}ytkow</snm><fnm>JM</fnm></au>
+  </aug>
+  <source>Journal of Intelligent Information Systems</source>
+  <publisher>Springer</publisher>
+  <pubdate>2003</pubdate>
+  <volume>20</volume>
+  <issue>3</issue>
+  <fpage>285</fpage>
+  <lpage>-302</lpage>
+</bibl>
+
+<bibl id="B26">
+  <title><p>Musical instrument identification in continuous
+  recordings</p></title>
+  <aug>
+    <au><snm>Livshin</snm><fnm>A</fnm></au>
+    <au><snm>Rodet</snm><fnm>X</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference on Digital Audio Effects
+  (DAFx)</source>
+  <pubdate>2004</pubdate>
+</bibl>
+
+<bibl id="B27">
+  <title><p>Music instrument recognition: from isolated notes to solo
+  phrases</p></title>
+  <aug>
+    <au><snm>Krishna</snm><fnm>A.G.</fnm></au>
+    <au><snm>Sreenivas</snm><fnm>TV</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)</source>
+  <pubdate>2004</pubdate>
+</bibl>
+
+<bibl id="B28">
+  <title><p>{Automatic recognition of isolated monophonic musical instrument
+  sounds using kNNC}</p></title>
+  <aug>
+    <au><snm>Kaminskyj</snm><fnm>I</fnm></au>
+    <au><snm>Czaszejko</snm><fnm>T</fnm></au>
+  </aug>
+  <source>Journal of Intelligent Information Systems</source>
+  <publisher>Springer</publisher>
+  <pubdate>2005</pubdate>
+  <volume>24</volume>
+  <issue>2-3</issue>
+  <fpage>199</fpage>
+  <lpage>-221</lpage>
+</bibl>
+
+<bibl id="B29">
+  <title><p>Musical instrument classification using non-negative matrix
+  factorization algorithms and subset feature selection</p></title>
+  <aug>
+    <au><snm>Benetos</snm><fnm>E</fnm></au>
+    <au><snm>Kotti</snm><fnm>M</fnm></au>
+    <au><snm>Kotropoulos</snm><fnm>C</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Conference on Acoustics,
+  Speech, and Signal Processing (ICASSP)</source>
+  <pubdate>2006</pubdate>
+</bibl>
+
+<bibl id="B30">
+  <title><p>{Automatic musical instrument classification using fractional
+  Fourier transform based-MFCC features and counter propagation neural
+  network}</p></title>
+  <aug>
+    <au><snm>Bhalke</snm><fnm>D.G.</fnm></au>
+    <au><snm>Rao</snm><fnm>CR</fnm></au>
+    <au><snm>Bormane</snm><fnm>DS</fnm></au>
+  </aug>
+  <source>Journal of Intelligent Information Systems</source>
+  <publisher>Springer</publisher>
+  <pubdate>2016</pubdate>
+  <volume>46</volume>
+  <issue>3</issue>
+  <fpage>425</fpage>
+  <lpage>-446</lpage>
+</bibl>
+
+<bibl id="B31">
+  <title><p>OpenMIC-2018: an open dataset for multiple instrument
+  recognition</p></title>
+  <aug>
+    <au><snm>Humphrey</snm><fnm>E</fnm></au>
+    <au><snm>Durand</snm><fnm>S</fnm></au>
+    <au><snm>McFee</snm><fnm>B</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2018</pubdate>
+</bibl>
+
+<bibl id="B32">
+  <title><p>{A plan for sustainable {MIR} evaluation}</p></title>
+  <aug>
+    <au><snm>McFee</snm><fnm>B</fnm></au>
+    <au><snm>Humphrey</snm><fnm>EJ</fnm></au>
+    <au><snm>Urbano</snm><fnm>J</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2016</pubdate>
+</bibl>
+
+<bibl id="B33">
+  <title><p>{FMA: A dataset for music analysis}</p></title>
+  <aug>
+    <au><snm>Defferrard</snm><fnm>M</fnm></au>
+    <au><snm>Benzi</snm><fnm>K</fnm></au>
+    <au><snm>Vandergheynst</snm><fnm>P</fnm></au>
+    <au><snm>Bresson</snm><fnm>X</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2017</pubdate>
+</bibl>
+
+<bibl id="B34">
+  <title><p>Deep convolutional networks on the pitch spiral for musical
+  instrument recognition</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Cella</snm><fnm>CE</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2016</pubdate>
+</bibl>
+
+<bibl id="B35">
+  <title><p>{MedleyDB: A multitrack dataset for annotation-intensive MIR
+  research}</p></title>
+  <aug>
+    <au><snm>Bittner</snm><fnm>RM</fnm></au>
+    <au><snm>Salamon</snm><fnm>J</fnm></au>
+    <au><snm>Tierney</snm><fnm>M</fnm></au>
+    <au><snm>Mauch</snm><fnm>M</fnm></au>
+    <au><snm>Cannam</snm><fnm>C</fnm></au>
+    <au><snm>Bello</snm><fnm>JP</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2014</pubdate>
+</bibl>
+
+<bibl id="B36">
+  <title><p>A software framework for musical data augmentation</p></title>
+  <aug>
+    <au><snm>McFee</snm><fnm>B</fnm></au>
+    <au><snm>Humphrey</snm><fnm>EJ</fnm></au>
+    <au><snm>Bello</snm><fnm>JP</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR)</source>
+  <pubdate>2015</pubdate>
+</bibl>
+
+<bibl id="B37">
+  <title><p>Timbre analysis of music audio signals with convolutional neural
+  networks</p></title>
+  <aug>
+    <au><snm>{Pons}</snm><fnm>J.</fnm></au>
+    <au><snm>{Slizovskaia}</snm><fnm>O.</fnm></au>
+    <au><snm>{Gong}</snm><fnm>R.</fnm></au>
+    <au><snm>{G\'{o}mez}</snm><fnm>E.</fnm></au>
+    <au><snm>{Serra}</snm><fnm>X.</fnm></au>
+  </aug>
+  <source>25th European Signal Processing Conference (EUSIPCO)</source>
+  <pubdate>2017</pubdate>
+  <fpage>2744</fpage>
+  <lpage>-2748</lpage>
+</bibl>
+
+<bibl id="B38">
+  <title><p>Instrument Activity Detection in Polyphonic Music using Deep Neural
+  Networks.</p></title>
+  <aug>
+    <au><snm>Gururani</snm><fnm>S</fnm></au>
+    <au><snm>Summers</snm><fnm>C</fnm></au>
+    <au><snm>Lerch</snm><fnm>A</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2018</pubdate>
+</bibl>
+
+<bibl id="B39">
+  <title><p>Timbre Classification Of A Single Musical Instrument</p></title>
+  <aug>
+    <au><snm>Loureiro</snm><fnm>MA</fnm></au>
+    <au><snm>Paula</snm><fnm>HB</fnm></au>
+    <au><snm>Yehia</snm><fnm>HC</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2004</pubdate>
+</bibl>
+
+<bibl id="B40">
+  <title><p>Deep convolutional neural networks for predominant instrument
+  recognition in polyphonic music</p></title>
+  <aug>
+    <au><snm>Han</snm><fnm>Y</fnm></au>
+    <au><snm>Kim</snm><fnm>J</fnm></au>
+    <au><snm>Lee</snm><fnm>K</fnm></au>
+  </aug>
+  <source>IEEE Transactions on Audio, Speech, and Language Processing</source>
+  <publisher>IEEE Press</publisher>
+  <pubdate>2017</pubdate>
+  <volume>25</volume>
+  <issue>1</issue>
+  <fpage>208</fpage>
+  <lpage>-221</lpage>
+</bibl>
+
+<bibl id="B41">
+  <title><p>The perception of musical timbre</p></title>
+  <aug>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+    <au><snm>Giordano</snm><fnm>BL</fnm></au>
+  </aug>
+  <source>The {O}xford handbook of music psychology</source>
+  <pubdate>2009</pubdate>
+  <fpage>72</fpage>
+  <lpage>-80</lpage>
+</bibl>
+
+<bibl id="B42">
+  <title><p>Spectro-temporal response field characterization with dynamic
+  ripples in ferret primary auditory cortex</p></title>
+  <aug>
+    <au><snm>Depireux</snm><fnm>DA</fnm></au>
+    <au><snm>Simon</snm><fnm>JZ</fnm></au>
+    <au><snm>Klein</snm><fnm>DJ</fnm></au>
+    <au><snm>Shamma</snm><fnm>SA</fnm></au>
+  </aug>
+  <source>Journal of Neurophysiology</source>
+  <publisher>Am Physiological Soc</publisher>
+  <pubdate>2001</pubdate>
+  <volume>85</volume>
+  <issue>3</issue>
+  <fpage>1220</fpage>
+  <lpage>-1234</lpage>
+</bibl>
+
+<bibl id="B43">
+  <title><p>The spectro-temporal receptive field</p></title>
+  <aug>
+    <au><snm>Aertsen</snm><fnm>AMHJ</fnm></au>
+    <au><snm>Johannesma</snm><fnm>PIM</fnm></au>
+  </aug>
+  <source>Biological Cybernetics</source>
+  <publisher>Springer</publisher>
+  <pubdate>1981</pubdate>
+  <volume>42</volume>
+  <issue>2</issue>
+  <fpage>133</fpage>
+  <lpage>-143</lpage>
+</bibl>
+
+<bibl id="B44">
+  <title><p>Triggered correlation</p></title>
+  <aug>
+    <au><snm>De Boer</snm><fnm>E</fnm></au>
+    <au><snm>Kuyper</snm><fnm>P</fnm></au>
+  </aug>
+  <source>IEEE Transactions on Biomedical Engineering</source>
+  <publisher>IEEE</publisher>
+  <pubdate>1968</pubdate>
+  <issue>3</issue>
+  <fpage>169</fpage>
+  <lpage>-179</lpage>
+</bibl>
+
+<bibl id="B45">
+  <title><p>Time-frequency/time-scale analysis</p></title>
+  <aug>
+    <au><snm>Flandrin</snm><fnm>P</fnm></au>
+  </aug>
+  <publisher>Academic press</publisher>
+  <pubdate>1998</pubdate>
+</bibl>
+
+<bibl id="B46">
+  <title><p>Wiener and Volterra analyses applied to the auditory
+  system</p></title>
+  <aug>
+    <au><snm>Eggermont</snm><fnm>JJ</fnm></au>
+  </aug>
+  <source>Hearing research</source>
+  <publisher>Elsevier</publisher>
+  <pubdate>1993</pubdate>
+  <volume>66</volume>
+  <issue>2</issue>
+  <fpage>177</fpage>
+  <lpage>-201</lpage>
+</bibl>
+
+<bibl id="B47">
+  <title><p>Robust spectrotemporal reverse correlation for the auditory system:
+  optimizing stimulus design</p></title>
+  <aug>
+    <au><snm>Klein</snm><fnm>DJ</fnm></au>
+    <au><snm>Depireux</snm><fnm>DA</fnm></au>
+    <au><snm>Simon</snm><fnm>JZ</fnm></au>
+    <au><snm>Shamma</snm><fnm>SA</fnm></au>
+  </aug>
+  <source>Journal of Computational Neuroscience</source>
+  <pubdate>2000</pubdate>
+  <volume>9</volume>
+  <issue>1</issue>
+  <fpage>85</fpage>
+  <lpage>-111</lpage>
+</bibl>
+
+<bibl id="B48">
+  <title><p>Spectral-temporal receptive fields of nonlinear auditory neurons
+  obtained using natural sounds</p></title>
+  <aug>
+    <au><snm>Theunissen</snm><fnm>FE</fnm></au>
+    <au><snm>Sen</snm><fnm>K</fnm></au>
+    <au><snm>Doupe</snm><fnm>AJ</fnm></au>
+  </aug>
+  <source>Journal of Neuroscience</source>
+  <publisher>Soc. Neuroscience</publisher>
+  <pubdate>2000</pubdate>
+  <volume>20</volume>
+  <issue>6</issue>
+  <fpage>2315</fpage>
+  <lpage>-2331</lpage>
+</bibl>
+
+<bibl id="B49">
+  <title><p>Multiresolution spectrotemporal analysis of complex
+  sounds</p></title>
+  <aug>
+    <au><snm>Chi</snm><fnm>T</fnm></au>
+    <au><snm>Ru</snm><fnm>P</fnm></au>
+    <au><snm>Shamma</snm><fnm>SA</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>ASA</publisher>
+  <pubdate>2005</pubdate>
+  <volume>118</volume>
+  <issue>2</issue>
+  <fpage>887</fpage>
+  <lpage>-906</lpage>
+</bibl>
+
+<bibl id="B50">
+  <title><p>Biomimetic spectro-temporal features for music instrument
+  recognition in isolated notes and solo phrases</p></title>
+  <aug>
+    <au><snm>Patil</snm><fnm>K</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>EURASIP J. Audio Speech Music Process.</source>
+  <publisher>Nature Publishing Group</publisher>
+  <pubdate>2015</pubdate>
+  <volume>2015</volume>
+  <issue>1</issue>
+  <fpage>27</fpage>
+</bibl>
+
+<bibl id="B51">
+  <title><p>Perceptually salient spectrotemporal modulations for recognition of
+  sustained musical instruments</p></title>
+  <aug>
+    <au><snm>Thoret</snm><fnm>E</fnm></au>
+    <au><snm>Depalle</snm><fnm>P</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>ASA</publisher>
+  <pubdate>2016</pubdate>
+  <volume>140</volume>
+  <issue>6</issue>
+  <fpage>EL478</fpage>
+  <lpage>-EL483</lpage>
+</bibl>
+
+<bibl id="B52">
+  <title><p>Understanding a Deep Machine Listening Model Through Feature
+  Inversion</p></title>
+  <aug>
+    <au><snm>Mishra</snm><fnm>S</fnm></au>
+    <au><snm>Sturm</snm><fnm>BL</fnm></au>
+    <au><snm>Dixon</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Proceedings of the International Society on Music Information
+  Retrieval (ISMIR) Conference</source>
+  <pubdate>2018</pubdate>
+  <fpage>755</fpage>
+  <lpage>-762</lpage>
+</bibl>
+
+<bibl id="B53">
+  <title><p>One hundred ways to process time, frequency, rate and scale in the
+  central auditory system: a pattern-recognition meta-analysis</p></title>
+  <aug>
+    <au><snm>Hemery</snm><fnm>E</fnm></au>
+    <au><snm>Aucouturier</snm><fnm>JJ</fnm></au>
+  </aug>
+  <source>Frontiers in computational neuroscience</source>
+  <pubdate>2015</pubdate>
+  <volume>9</volume>
+</bibl>
+
+<bibl id="B54">
+  <title><p>Kymatio: Scattering Transforms in Python</p></title>
+  <aug>
+    <au><snm>Andreux</snm><fnm>M</fnm></au>
+    <au><snm>Angles</snm><fnm>T</fnm></au>
+    <au><snm>Exarchakis</snm><fnm>G</fnm></au>
+    <au><snm>Leonarduzzi</snm><fnm>R</fnm></au>
+    <au><snm>Rochette</snm><fnm>G</fnm></au>
+    <au><snm>Thiry</snm><fnm>L</fnm></au>
+    <au><snm>Zarka</snm><fnm>J</fnm></au>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+    <au><snm>Belilovsky</snm><fnm>E</fnm></au>
+    <au><snm>Bruna</snm><fnm>J</fnm></au>
+    <au><cnm>others</cnm></au>
+  </aug>
+  <source>Journal of Machine Learning Research</source>
+  <pubdate>2020</pubdate>
+  <volume>21</volume>
+  <issue>60</issue>
+</bibl>
+
+<bibl id="B55">
+  <title><p>The Shape of RemiXXXes to Come: Audio texture synthesis with
+  time--frequency scattering</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Hecker</snm><fnm>F</fnm></au>
+  </aug>
+  <source>Proceedings of the Digital Audio Effects Conference (DAFX)</source>
+  <pubdate>2019</pubdate>
+</bibl>
+
+<bibl id="B56">
+  <title><p>Understanding deep convolutional networks</p></title>
+  <aug>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Philosophical Transactions of the Royal Society A: Mathematical,
+  Physical and Engineering Sciences</source>
+  <publisher>The Royal Society Publishing</publisher>
+  <pubdate>2016</pubdate>
+  <volume>374</volume>
+  <issue>2065</issue>
+  <fpage>20150203</fpage>
+</bibl>
+
+<bibl id="B57">
+  <title><p>A review of automatic drum transcription</p></title>
+  <aug>
+    <au><snm>Wu</snm><fnm>CW</fnm></au>
+    <au><snm>Dittmar</snm><fnm>C</fnm></au>
+    <au><snm>Southall</snm><fnm>C</fnm></au>
+    <au><snm>Vogl</snm><fnm>R</fnm></au>
+    <au><snm>Widmer</snm><fnm>G</fnm></au>
+    <au><snm>Hockman</snm><fnm>J</fnm></au>
+    <au><snm>Muller</snm><fnm>M</fnm></au>
+    <au><snm>Lerch</snm><fnm>A</fnm></au>
+  </aug>
+  <source>IEEE Transactions on Audio, Speech, and Language Processing</source>
+  <publisher>IEEE Press</publisher>
+  <pubdate>2018</pubdate>
+  <volume>26</volume>
+  <issue>9</issue>
+  <fpage>1457</fpage>
+  <lpage>-1483</lpage>
+</bibl>
+
+<bibl id="B58">
+  <title><p>Modelling Timbral Hardness</p></title>
+  <aug>
+    <au><snm>Pearce</snm><fnm>A</fnm></au>
+    <au><snm>Brookes</snm><fnm>T</fnm></au>
+    <au><snm>Mason</snm><fnm>R</fnm></au>
+  </aug>
+  <source>Applied Sciences</source>
+  <publisher>Multidisciplinary Digital Publishing Institute</publisher>
+  <pubdate>2019</pubdate>
+  <volume>9</volume>
+  <issue>3</issue>
+  <fpage>466</fpage>
+</bibl>
+
+<bibl id="B59">
+  <title><p>An efficient heuristic procedure for partitioning
+  graphs</p></title>
+  <aug>
+    <au><snm>Kernighan</snm><fnm>BW</fnm></au>
+    <au><snm>Lin</snm><fnm>S</fnm></au>
+  </aug>
+  <source>The Bell system technical journal</source>
+  <publisher>Nokia Bell Labs</publisher>
+  <pubdate>1970</pubdate>
+  <volume>49</volume>
+  <issue>2</issue>
+  <fpage>291</fpage>
+  <lpage>-307</lpage>
+</bibl>
+
+<bibl id="B60">
+  <title><p>Scalable parallel data mining for association rules</p></title>
+  <aug>
+    <au><snm>Han</snm><fnm>EH</fnm></au>
+    <au><snm>Karypis</snm><fnm>G</fnm></au>
+    <au><snm>Kumar</snm><fnm>V</fnm></au>
+  </aug>
+  <publisher>ACM</publisher>
+  <pubdate>1997</pubdate>
+  <volume>26</volume>
+  <issue>2</issue>
+</bibl>
+
+<bibl id="B61">
+  <title><p>Cluster ensembles---a knowledge reuse framework for combining
+  multiple partitions</p></title>
+  <aug>
+    <au><snm>Strehl</snm><fnm>A</fnm></au>
+    <au><snm>Ghosh</snm><fnm>J</fnm></au>
+  </aug>
+  <source>Journal of machine learning research</source>
+  <pubdate>2002</pubdate>
+  <volume>3</volume>
+  <issue>Dec</issue>
+  <fpage>583</fpage>
+  <lpage>-617</lpage>
+</bibl>
+
+<bibl id="B62">
+  <title><p>Hierarchical grouping to optimize an objective function</p></title>
+  <aug>
+    <au><snm>Ward Jr.</snm><fnm>JH</fnm></au>
+  </aug>
+  <source>Journal of the American Statistical Association</source>
+  <publisher>Taylor \& Francis Group</publisher>
+  <pubdate>1963</pubdate>
+  <volume>58</volume>
+  <issue>301</issue>
+  <fpage>236</fpage>
+  <lpage>-244</lpage>
+</bibl>
+
+<bibl id="B63">
+  <title><p>Constant-Q transform toolbox for music processing</p></title>
+  <aug>
+    <au><snm>Sch{\"o}rkhuber</snm><fnm>C</fnm></au>
+    <au><snm>Klapuri</snm><fnm>A</fnm></au>
+  </aug>
+  <source>Proceedings of the Sound and Music Computing (SMC)
+  Conference</source>
+  <pubdate>2010</pubdate>
+</bibl>
+
+<bibl id="B64">
+  <title><p>Learning the helix topology of musical pitch</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Sridhar</snm><fnm>S</fnm></au>
+    <au><snm>Farnsworth</snm><fnm>A</fnm></au>
+    <au><snm>Bello</snm><fnm>JP</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference on Acoustics, Speech, and
+  Signal Processing (ICASSP)</source>
+  <pubdate>2020</pubdate>
+</bibl>
+
+<bibl id="B65">
+  <title><p>Group invariant scattering</p></title>
+  <aug>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Communications in Pure and Applied Mathematics</source>
+  <publisher>Wiley Online Library</publisher>
+  <pubdate>2012</pubdate>
+  <volume>65</volume>
+  <issue>10</issue>
+  <fpage>1331</fpage>
+  <lpage>-1398</lpage>
+</bibl>
+
+<bibl id="B66">
+  <title><p>A spectro-temporal modulation index (STMI) for assessment of speech
+  intelligibility</p></title>
+  <aug>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+    <au><snm>Chi</snm><fnm>T</fnm></au>
+    <au><snm>Shamma</snm><fnm>SA</fnm></au>
+  </aug>
+  <source>Speech communication</source>
+  <publisher>Elsevier</publisher>
+  <pubdate>2003</pubdate>
+  <volume>41</volume>
+  <issue>2-3</issue>
+  <fpage>331</fpage>
+  <lpage>-348</lpage>
+</bibl>
+
+<bibl id="B67">
+  <title><p>Detection of speech tokens in noise using adaptive spectrotemporal
+  receptive fields</p></title>
+  <aug>
+    <au><snm>Bellur</snm><fnm>A</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>Proceedings of the Annual Conference on Information Sciences and
+  Systems (CISS)</source>
+  <pubdate>2015</pubdate>
+  <fpage>1</fpage>
+  <lpage>-6</lpage>
+</bibl>
+
+<bibl id="B68">
+  <title><p>A multiresolution analysis for detection of abnormal lung
+  sounds</p></title>
+  <aug>
+    <au><snm>Emmanouilidou</snm><fnm>D</fnm></au>
+    <au><snm>Patil</snm><fnm>K</fnm></au>
+    <au><snm>West</snm><fnm>J</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference of the IEEE Engineering
+  in Medicine and Biology Society (EMBS)</source>
+  <pubdate>2012</pubdate>
+  <fpage>3139</fpage>
+  <lpage>-3142</lpage>
+</bibl>
+
+<bibl id="B69">
+  <title><p>Unsupervised learning of semantic audio representations</p></title>
+  <aug>
+    <au><snm>Jansen</snm><fnm>A</fnm></au>
+    <au><snm>Plakal</snm><fnm>M</fnm></au>
+    <au><snm>Pandya</snm><fnm>R</fnm></au>
+    <au><snm>Ellis</snm><fnm>DP</fnm></au>
+    <au><snm>Hershey</snm><fnm>S</fnm></au>
+    <au><snm>Liu</snm><fnm>J</fnm></au>
+    <au><snm>Moore</snm><fnm>RC</fnm></au>
+    <au><snm>Saurous</snm><fnm>RA</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Conference on Acoustics, Speech
+  and Signal Processing (ICASSP)</source>
+  <pubdate>2018</pubdate>
+  <fpage>126</fpage>
+  <lpage>-130</lpage>
+</bibl>
+
+<bibl id="B70">
+  <title><p>{M}etric {L}earning</p></title>
+  <aug>
+    <au><snm>Bellet</snm><fnm>A</fnm></au>
+    <au><snm>Habrard</snm><fnm>A</fnm></au>
+    <au><snm>Sebban</snm><fnm>M</fnm></au>
+  </aug>
+  <publisher>{M}organ \& {C}laypool {P}ublishers</publisher>
+  <pubdate>2015</pubdate>
+</bibl>
+
+<bibl id="B71">
+  <title><p>Playing Technique Recognition by Joint Time--Frequency
+  Scattering</p></title>
+  <aug>
+    <au><snm>Wang</snm><fnm>C</fnm></au>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Benetos</snm><fnm>E</fnm></au>
+    <au><snm>Chew</snm><fnm>E</fnm></au>
+  </aug>
+  <pubdate>2020</pubdate>
+</bibl>
+
+<bibl id="B72">
+  <title><p>A dictionary of economics</p></title>
+  <aug>
+    <au><snm>Black</snm><fnm>J</fnm></au>
+    <au><snm>Hashimzade</snm><fnm>N</fnm></au>
+    <au><snm>Myles</snm><fnm>G</fnm></au>
+  </aug>
+  <publisher>Oxford university press</publisher>
+  <pubdate>2012</pubdate>
+</bibl>
+
+<bibl id="B73">
+  <title><p>OrchideaSOL: A Dataset of Extended Instrumental Techniques for
+  Computer-aided Orchestration</p></title>
+  <aug>
+    <au><snm>Cella</snm><fnm>CE</fnm></au>
+    <au><snm>Ghisi</snm><fnm>D</fnm></au>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>L{\'e}vy</snm><fnm>F</fnm></au>
+    <au><snm>Fineberg</snm><fnm>J</fnm></au>
+    <au><snm>Maresz</snm><fnm>Y</fnm></au>
+  </aug>
+  <source>Proceedings of the International Computer Music Conference
+  (ICMC)</source>
+  <pubdate>2020</pubdate>
+</bibl>
+
+<bibl id="B74">
+  <title><p>On Time-frequency Scattering and Computer Music</p></title>
+  <aug>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+  </aug>
+  <source>Florian Hecker: Halluzination, Perspektive, Synthese</source>
+  <publisher>Berlin: Sternberg Press</publisher>
+  <editor>Schafhausen, Nicolaus and M\"{u}ller, Vanessa Joan</editor>
+  <pubdate>2019</pubdate>
+</bibl>
+
+<bibl id="B75">
+  <title><p>Representing environmental sounds using the separable scattering
+  transform</p></title>
+  <aug>
+    <au><snm>Baug{\'e}</snm><fnm>C</fnm></au>
+    <au><snm>Lagrange</snm><fnm>M</fnm></au>
+    <au><snm>And{\'e}n</snm><fnm>J</fnm></au>
+    <au><snm>Mallat</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Proceedings of the IEEE International Conference on Acoustics, Speech
+  and Signal Processing</source>
+  <pubdate>2013</pubdate>
+  <fpage>8667</fpage>
+  <lpage>-8671</lpage>
+</bibl>
+
+<bibl id="B76">
+  <title><p>Adaptive Time--Frequency Scattering for Periodic Modulation
+  Recognition in Music Signals</p></title>
+  <aug>
+    <au><snm>Wang</snm><fnm>C</fnm></au>
+    <au><snm>Benetos</snm><fnm>E</fnm></au>
+    <au><snm>Lostanlen</snm><fnm>V</fnm></au>
+    <au><snm>Chew</snm><fnm>E</fnm></au>
+  </aug>
+  <pubdate>2019</pubdate>
+</bibl>
+
+<bibl id="B77">
+  <title><p>Separable spectro-temporal Gabor filter bank features: Reducing the
+  complexity of robust features for automatic speech recognition</p></title>
+  <aug>
+    <au><snm>Sch{\"a}dler</snm><fnm>MR</fnm></au>
+    <au><snm>Kollmeier</snm><fnm>B</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>Acoustical Society of America</publisher>
+  <pubdate>2015</pubdate>
+  <volume>137</volume>
+  <issue>4</issue>
+  <fpage>2047</fpage>
+  <lpage>-2059</lpage>
+</bibl>
+
+<bibl id="B78">
+  <title><p>Training and testing low-degree polynomial data mappings via linear
+  SVM</p></title>
+  <aug>
+    <au><snm>Chang</snm><fnm>YW</fnm></au>
+    <au><snm>Hsieh</snm><fnm>CJ</fnm></au>
+    <au><snm>Chang</snm><fnm>KW</fnm></au>
+    <au><snm>Ringgaard</snm><fnm>M</fnm></au>
+    <au><snm>Lin</snm><fnm>CJ</fnm></au>
+  </aug>
+  <source>Journal of Machine Learning Research</source>
+  <pubdate>2010</pubdate>
+  <volume>11</volume>
+  <issue>Apr</issue>
+  <fpage>1471</fpage>
+  <lpage>-1490</lpage>
+</bibl>
+
+</refgrp>
+} % end of \BMCxmlcomment
 

### 2020-07-23 paper/paper.tex
declaration spection

@@
 In this section, we aim to recover these similarity judgments from digital audio recordings according to a paradigm of supervised metric learning.
-To this end, we present a machine listening system composing joint time--frequency scattering and large-margin nearest neighbors (LMNN).
+To this end, we present a machine listening system composing joint time--frequency scattering and LMNN.
 
@@
 As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
+In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
 
@@
 
-Let $P$ be the number of scattering paths in the operator $\mathbf{S}$.
-The large-margin nearest neighbor (LMNN) algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
+Let $P$ be the number of scattering paths in the operator $\mathbf{S}$. The LMNN algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
 \begin{equation}
@@
 Given an audio stimulus $\boldsymbol{x}_n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
-Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
+Then, we identify its IMT triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
@@
 This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
-This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
+This gap in performance, together with the theory of STRFs in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
 
 \subsection*{What is the baseline performance?}
-Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
+Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by MFCCs.
 Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis.\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}
@@
 
-\section*{\nmu Competing interests}
+\section*{Abbreviations}
+
+\begin{itemize}
+  \item AP@5: average precision at rank 5
+  \item CNSMDP: conservatoire national sup\'erieur de musique et de danse de Paris
+  \item COSMIR: community for open and sustainable music information research
+  \item DCT: discrete cosine transform
+  \item FMA: free music archive
+  \item GBFB: Gabor filterbank features
+  \item IMT: instrument--mute--technique
+  \item LMNN: large-margin nearest neighbor
+  \item MFCC: mel-frequency cepstral coefficients
+  \item MIR: music information retrieval
+  \item P@5: precision at rank five
+  \item ReLU: rectified linear unit
+  \item STRF: spectrotemporal receptive fields
+  \item SOL: studio On Line dataset
+\end{itemize}
+
+\subsection{Musical instruments}
+
+\begin{itemize}
+  \item Vn: violin,
+  \item Va: viola,
+  \item Vc: celloVc,
+  \item Cb: contrabass,
+  \item Hp: concert harp,
+  \item Gtr: Spanish guitar,
+  \item Acc: accordion,
+  \item Fl: flute,
+  \item BbCl: soprano clarinet,
+  \item ASax: alto saxophone,
+  \item Ob: oboe,
+  \item Bn: bassoon,
+  \item TpC: trumpet in C,
+  \item Hn: French horn,
+  \item TTbn: tenor trombone,
+  \item BBTb: bass tuba.
+\end{itemize}
+
+\subsection{Musical nuances}
+
+\begin{itemize}
+  \item \emph{pp}: pianissimo,
+  \item \emph{p}: piano,
+  \item \emph{mf}: mezzo forte,
+  \item \emph{f}: forte,
+  \item \emph{ff}: fortissimo.
+\end{itemize}
+
+\section*{Declarations}
+
+\subsection*{\nmu Availability of data and materials}
+
+The processing code, the cyberlioz interface and an anonimized version of the  perceptual judgments gathered using this interface are available at the companion website: \url{https://github.com/mathieulagrange/lostanlen2020jasmp}.
+
+The SOL dataset is available online: \url{https://forum.ircam.fr/projects/detail/fullsol}
+
+\subsection*{\nmu Competing interests}
   The authors declare that they have no competing interests.
@@
 
-\section*{\nmu Acknowledgment}
+\subsection*{\nmu Funding}
+This work is partially supported by the Paris sciences et lettres (PSL) TICEL project.
 This work is partially supported by the European Research Council (ERC) award 320959 (InvariantClass).
 This work is partially supported by the National Science Foundation (NSF) award 1633259 (BIRDVOX).
-The Flatiron Institute is a division of the Simons Foundation.
+This work is partially supported by the Flatiron Institute, a division of the Simons Foundation.
+
+\subsection*{\nmu Acknowledgments}
+
 We wish to thank Philippe Brandeis, \'{E}tienne Graindorge, St\'{e}phane Mallat, Adrien Mamou-Mani, and Yan Maresz for contributing to the TICEL research project.
 We also wish to thank the students of the Paris Conservatory and all anonymous participants to our study.
+
+\subsection*{Authors' contributions}
+
+VL provided guidance in the design of the computational experiments and was a major contributor in writing the manuscript.
+GEH conducted part of the computational experiments.
+MR designed and implemented the listening tests.
+GL designed the listening tests and provided guidance in the design of the computational experiments.
+JA provided guidance in the design of the computational experiments, participated to the data analysis and  to the writing of the manuscript.
+ML designed and conducted the computational experiments, the data analysis and participated to the writing of the paper.
 

### 2020-07-23 paper/paper.tex
p@5 -> ap@5

@@
 i think it's important to use the words "triplet loss" in the abstract
-Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of $\mathsf{99.0\%\pm1}$.
+Over a dataset of 9346 isolated notes, we report a state-of-the-art average precision at rank five (AP@5) of $\mathsf{99.0\%\pm1}$.
 An ablation study demonstrates that removing either the joint time--frequency scattering transform or the metric learning algorithm noticeably degrades performance.
@@
 \subsection*{Is metric learning necessary?}
-Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
+Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to an AP@5 of $\textbf{76.6}\% \pm 11$.
 Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all. % \ja wants to elaborate
@@
 As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
+Over our cohort of $K=31$ subjects, we report an AP@5 of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
 This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.
@@
 
-Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
+Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves n AP@5 of $\textbf{91.9\%} \pm 4$.
 This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
@@
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{81.8\%} \pm 7$.
-
-Arguably, this figure is not directly comparable with our best performing system (P@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
+Over our cohort of $K=31$ subjects, we report an AP@5 of $\textbf{81.8\%} \pm 7$.
+
+Arguably, this figure is not directly comparable with our best performing system (AP@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
 In order to clarify the role of feature dimensionality in our computational pipeline, we apply a feature engineering technique involving multiplicative combinations of MFCC.
@@
 In our case, the implicit similarity kernel is a homogeneous quadratic kernel.
-Despite this increase in representational power, we obtain a P@5 of $81.5\% \pm 7$, i.e. essentially the same as MFCC under a linear kernel.
+Despite this increase in representational power, we obtain an AP@5 of $81.5\% \pm 7$, i.e. essentially the same as MFCC under a linear kernel.
 Therefore, it appears that the gap in performance between MFCC ($81.8\%\pm 7$) and joint time--frequency scattering ($99.0\% \pm 1$) is primarily attributable to the multiscale extraction of joint spectrotemporal modulations, whereas high-dimensional embedding likely plays a minor role.
@@
 
-\subsection{Musical instruments}
+\subsection*{Musical instruments}
 
@@
 
-\subsection{Musical nuances}
+\subsection*{Musical nuances}
 

### 2020-10-12 paper/paper.tex
update list of keywords
rephrase paragraph 2 from intro

@@
 \kwd{audio databases}
+\kwd{audio similarity}
 \kwd{continuous wavelet transform}
@@
 For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
-Yet, such acoustical similarity at the scale of the entire instrumentarium is not directly reflected by the semantic similarity between playing technique denominations; and discovering it requires musical expertise \cite{kolozali2011ismir}.
+Yet, in a diverse instrumentarium, the semantic similarity between playing technique denominations does not reflect such acoustical similarity \cite{kolozali2011ismir}.
 
@@
 This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
-% \vl: ?
 

### 2020-10-12 paper/paper.tex
more minor changes from R1

@@
 
-First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
+First, our dataset encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
 Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
@@
 Timbre involves multiple time scales in conjunction, from a few microseconds for an attack transient to several seconds for a sustained tone.
-Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more \cite{joder2009taslp}.
+Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window \cite{joder2009taslp}.
 Mapping this input to a feature space in which distances denote timbral dissimilarity requires a data-driven stage of dimensionality reduction.

### 2020-10-12 paper/paper.tex
rephrase "however accurate" (requested by R1)
philharmonic orchestra (R1)

@@
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
-In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
+In this setting, prior knowledge of instrumentation, however accurate, does not suffice to conduct a study on timbre perception.
 Rather, each excerpt must be played back to multiple human listeners.
@@
 
-The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
+The philharmonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
 In this article, we focus on the first three, and leave the question of learning auditory similarities between percussion instruments to future research.

### 2020-10-12 paper/paper.tex
give an equation for Lipschitz stability (R1)

@@
 where $\varepsilon$ is a predefined constant.
-The offset of one before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps.
+The offset of one before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps: there exists a constant $c$ such that
+\begin{equation}
+\big \Vert
+\mathbf{\widetilde{S}}\boldsymbol{x}_m
+-
+\mathbf{\widetilde{S}}\boldsymbol{x}_{n}
+\big \Vert
+\leq
+c
+\big \Vert
+\mathbf{S}\boldsymbol{x}_m
+-
+\mathbf{S}\boldsymbol{x}_{n}
+\big \Vert
+\end{equation}
+for every pair of samples $(\boldsymbol{x}_m, \boldsymbol{x}_n)$.
 On a dataset of environmental audio recordings, a previous publication has shown empirically that Equation \ref{eq:S_tilde} brings the histogram of $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ closer to a Gaussian distribution \cite{lostanlen2018jasmp}.

### 2020-10-12 paper/paper.tex
let x be. priorly -> previously

@@
 
-Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ a complex-valued filter with zero average, dimensionless center frequency equal to one, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
+Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ be a complex-valued filter with zero average, dimensionless center frequency equal to one, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
 We define a constant-$Q$ wavelet filterbank as the family $\boldsymbol{\psi}_{\lambda} : t \mapsto \lambda \boldsymbol{\psi}(\lambda t)$.
@@
 \subsection*{Metric learning with large-margin nearest neighbors (LMNN)}
-Let $\boldsymbol{x}$ some arbitrary audio sample in the dataset $\mathcal{X}$.
-Let $\mathcal{G}$ a cluster graph with $N = \card \mathcal{X}$ vertices and $C$ clusters in total.
+Let $\boldsymbol{x}$ be some arbitrary audio sample in the dataset $\mathcal{X}$.
+Let $\mathcal{G}$ be a cluster graph with $N = \card \mathcal{X}$ vertices and $C$ clusters in total.
 We denote by $\mathcal{G}(\boldsymbol{x})$ the cluster to which the sample $\boldsymbol{x}$ belongs.
@@
 \item Affine standardization so that each feature has zero mean and unit variance, and
-\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
+\item Nearest-neighbor search according to previously learned non-Euclidean metric.
 \end{enumerate}

### 2020-10-12 paper/paper.tex
cite Brown and Eronen re: 23ms MFCC

@@
 This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
-Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$.
+Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$ \cite{brown1999jasa,eronen2000icassp}.
 

### 2020-10-12 paper/paper.tex
satisfactory performance -> high performance
GEH -> CEH in author initials

@@
 
-We see from the ablation study conducted above that each step of the proposed model is necessary for its satisfactory performance.
+We see from the ablation study conducted above that each step of the proposed model is necessary for its high performance.
 Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
@@
 VL provided guidance in the design of the computational experiments and was a major contributor in writing the manuscript.
-GEH conducted part of the computational experiments.
+CEH conducted part of the computational experiments.
 MR designed and implemented the listening tests.

### 2020-10-24 paper/paper.tex
rename Contributions to Approach
comment out Outline, move footnote

@@
 \subsection*{Goal}
-This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.
+This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.\footnote{
+For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
 Crucially, this dissimilarity is not evaluated in terms of acoustic tags, but in terms of \emph{ad hoc} clusters, as defined by a human consensus of auditory judgments.
@@
 
-\subsection*{Contributions}
+\subsection*{Approach}
 The main contribution of this article can be formulated as the intersection between three topics.
@@
 
-\subsection*{Outline}
-The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
-The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
-The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-The \lnameref{sec:results} section describes our experimental benchmark.
-The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
-The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
+From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is readily tailored to subjective preferences.
+
+%\subsection*{Outline}
+%The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
+%The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
+%The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+%The \lnameref{sec:results} section describes our experimental benchmark.
+%The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
+%The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.
 

### 2020-10-24 paper/paper.tex
integrate claim of originality to contributions

@@
 In addition, previous publications on joint time--frequency scattering lack a human-centric evaluation, independently from any classification task.
-
-\subsection*{Claim of originality}
-The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is also readily tailored to subjective preferences.
 

### 2020-10-24 paper/paper.tex
move metric to methods, visualisation to results
minor change to list of contributions
rename sections. move inter-* to results

@@
 
+Out paper strives to fill the gap in scholarship between MIR and music cognition in the context of extended playing techniques.
 From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
@@
 
-\section*{\nmu Data collection}
+\section*{\nmu Perceptual data collection}
 \label{sec:data-collection}
@@
 This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
-
-\subsection*{Inter-instrument similarity}
-
-Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
-Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
-To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$.
-
-Let us denote the existence of an edge in $\mathcal{G}_0$ connecting two IMT samples $m$ and $n$ by the binary relation $m \overset{\mathcal{G}_0}{\sim} n$.
-The mathematical definition of the matrix $\mathbf{A}_{\mathcal{I}}$ is:
-\begin{equation}
-\mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Instrument}(n) \in \{ i, j \} \big\}
-}.
-\label{eq:instrument-similarity}
-\end{equation}
-
-Then, we run Ward's method \cite{ward1963jasa} to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
-This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$ such that the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
-Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
-% \ja wants to remove this equation
-
-Interestingly, the matrix $\widetilde{\mathbf{A}}_{\mathcal{I}}$ exhibits a block diagonal structure, which roughly reflects the classical taxonomy of musical instruments: four woodwinds (BbCl, Fl, ASax, Ob) are clustered together, followed by a cluster containing one woodwind (Bn) and four brass (BBTb, Hn, TTbn, and TpC), followed by a cluster of all strings (Cb, Va, Vn, Vc, Gtr, Hp).
-Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
-
-\subsection*{Inter-technique similarity}
-From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
-As for $\mathbf{A}_{\mathcal{I}}$, we define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinalities:
-\begin{equation}
-\mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
-}.
-\label{eq:technique-similarity}
-\end{equation}
-%\ml{why 2 ?} \vl: fixed
-Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
-Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$.
-
-The block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less salient than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$.
-Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
-Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
-This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{} an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
-
-\section*{\nmu Methods}
+\section*{\nmu Machine listening methods}
 \label{sec:methods}
@@
 
-\subsection*{Visualization}
-\label{sec:visualization}
-
-At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
-From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
-To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
-In accordance with their protocol, we compute the following matrix:
-
-\begin{equation}
-\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
-\iint_{\mathbb{R}^2}
-\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
-\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
-\end{equation}
-
-Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
-As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
-
-Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
-However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
-Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
-
-To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
-By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
-For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
-A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
-As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
-Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
-
-The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
-Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
-These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
-These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
-Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
-This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
-
-Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
-Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
-Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
-We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
-
 \subsection*{Median-based logarithmic compression and affine standardization}
@@
 \end{equation}
-If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only.
+If the collection is split between a training set and a test set (see \lnameref{sec:results} section), we compute $\boldsymbol{\mu}$ on the training set only.
 Then, to match a decibel-like perception of loudness, we apply the following adaptive transformation, which composes a median-based renormalization and a logarithmic compression:
@@
 
-\section*{\nmu Results}
-\label{sec:results}
-
-The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
-We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
-However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
-
 \subsection*{Semi-supervised label propagation}
@@
 
+\section*{\nmu Results}
+\label{sec:results}
+
+The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
+We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
+In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
+However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
+
+\subsection*{Inter-instrument similarity}
+
+Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
+Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
+To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
+For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$.
+
+Let us denote the existence of an edge in $\mathcal{G}_0$ connecting two IMT samples $m$ and $n$ by the binary relation $m \overset{\mathcal{G}_0}{\sim} n$.
+The mathematical definition of the matrix $\mathbf{A}_{\mathcal{I}}$ is:
+\begin{equation}
+\mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
+\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+}{
+\Card \big\{n \,\vert\, \mathrm{Instrument}(n) \in \{ i, j \} \big\}
+}.
+\label{eq:instrument-similarity}
+\end{equation}
+
+Then, we run Ward's method \cite{ward1963jasa} to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
+This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$ such that the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
+Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
+% \ja wants to remove this equation
+
+Interestingly, the matrix $\widetilde{\mathbf{A}}_{\mathcal{I}}$ exhibits a block diagonal structure, which roughly reflects the classical taxonomy of musical instruments: four woodwinds (BbCl, Fl, ASax, Ob) are clustered together, followed by a cluster containing one woodwind (Bn) and four brass (BBTb, Hn, TTbn, and TpC), followed by a cluster of all strings (Cb, Va, Vn, Vc, Gtr, Hp).
+Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
+
+\subsection*{Inter-technique similarity}
+From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
+As for $\mathbf{A}_{\mathcal{I}}$, we define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinalities:
+\begin{equation}
+\mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
+\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
+}{
+\Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
+}.
+\label{eq:technique-similarity}
+\end{equation}
+%\ml{why 2 ?} \vl: fixed
+Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
+Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$.
+
+The block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less salient than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$.
+Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
+Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
+This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{} an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
+
 \subsection*{Best performing system}
@@
 
-\section*{\nmu Discussion}
-\label{sec:discussion}
-
-The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
-Over a cohort of $K=31$ subjects, we reported an average precision at rank five (AP@5) of $99.0\% \pm 1$.
+\subsection*{Visualization of joint--time frequency scattering coefficients}
+\label{sec:visualization}
+
+At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
+From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
+To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
+In accordance with their protocol, we compute the following matrix:
+
+\begin{equation}
+\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
+\iint_{\mathbb{R}^2}
+\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
+\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
+\end{equation}
+
+Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
+As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
+In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
+
+Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
+However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
+Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
+
+To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
+By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
+For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
+A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
+As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
+Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
+
+The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
+Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
+These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
+These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
+Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
+This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
+
+Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
+Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
+Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
+We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
+
+\subsection*{Ablation study}
 We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
@@
 
-\subsection*{Is the dataset large enough for evaluation? Is the system overfitting?}
+\subsection*{Role of dataset size}
 First of all, it is worth noting that the presented AP@5 figure of $99.0\% \pm 1$ does not abide to a conventional training set vs.
@@
 
-\subsection*{Is metric learning necessary?}
+\subsection*{Role of metric learning}
 Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to an AP@5 of $\textbf{76.6}\% \pm 11$.
@@
 
-\subsection*{How important is the amount of temporal context $T$?}
+\subsection*{Role of temporal context $T$}
 
@@
 
-\subsection*{Is joint time--frequency scattering necessary?}
+\subsection*{Role of joint time--frequency scattering}
 
@@
 
-\subsection*{What is the baseline performance?}
+\subsection*{Comparison with mel--frequency cepstral coefficients (MFCC) baseline}
 Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by MFCCs.

### 2020-10-26 paper/paper.tex
remove clustering figures and sections

@@
 However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
-
-\subsection*{Inter-instrument similarity}
-
-Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
-Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
-To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$.
-
-Let us denote the existence of an edge in $\mathcal{G}_0$ connecting two IMT samples $m$ and $n$ by the binary relation $m \overset{\mathcal{G}_0}{\sim} n$.
-The mathematical definition of the matrix $\mathbf{A}_{\mathcal{I}}$ is:
-\begin{equation}
-\mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Instrument}(n) \in \{ i, j \} \big\}
-}.
-\label{eq:instrument-similarity}
-\end{equation}
-
-Then, we run Ward's method \cite{ward1963jasa} to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
-This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$ such that the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
-Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
-% \ja wants to remove this equation
-
-Interestingly, the matrix $\widetilde{\mathbf{A}}_{\mathcal{I}}$ exhibits a block diagonal structure, which roughly reflects the classical taxonomy of musical instruments: four woodwinds (BbCl, Fl, ASax, Ob) are clustered together, followed by a cluster containing one woodwind (Bn) and four brass (BBTb, Hn, TTbn, and TpC), followed by a cluster of all strings (Cb, Va, Vn, Vc, Gtr, Hp).
-Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
-
-\subsection*{Inter-technique similarity}
-From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
-As for $\mathbf{A}_{\mathcal{I}}$, we define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinalities:
-\begin{equation}
-\mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
-}.
-\label{eq:technique-similarity}
-\end{equation}
-%\ml{why 2 ?} \vl: fixed
-Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
-Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$.
-
-The block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less salient than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$.
-Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
-Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
-This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{} an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
 
@@
 \begin{figure}
-\center
-\includegraphics[width = 0.85\textwidth]{consensusVsI.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{I}}$ of perceived timbre similarities between instruments from our dataset of $N=78$. samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:instrument-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ were re-arranged according to Ward's minimum variance method.
-Observe that the block diagonal structure in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ reflects an organological taxonomy of musical instruments: woodwinds, brass, and strings.
-See \lnameref{sec:data-collection} section for details, including the meaning of each instrument abbreviation.
-}
-\label{fig:consensusVsI}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = 0.85\textwidth]{consensusVsPt.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{T}}$ of perceived timbre similarities between playing techniques from our dataset of $N=78$ samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:technique-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{T}}$ were re-arranged according to Ward's minimum variance method.
-See \lnameref{sec:data-collection} section for details,.
-See Tables \ref{tab:list1} and \ref{tab:list2} for an English description of each playing technique abbreviation.
-%TODO add a double arrow?
-% rough <-- soft --> rough
-% sustained <-- "ordinary" --> percussive
-}
-\label{fig:consensusVsPt}
-\end{figure}
-
-\begin{figure}
 \includegraphics[height=4cm]{Vn-ord-C4-mf-4c_scalogram.pdf}

### 2020-10-26 paper/paper.tex
re-title subsections in related work section

@@
 
-\subsection*{Timbre modeling in music information retrieval (MIR)}
+\subsection*{Automatic classification of musical instruments and playing techniques}
 On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
@@
 
-\subsection*{Timbre modeling in music cognition research}
+\subsection*{Spectrotemporal receptive fields (STRF) in music cognition}
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.

### 2020-10-28 paper/paper.tex
subjects --> participants

@@
 However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique, use instrument identity as a proxy for timbre quality, and do not allow for customization to the perceptual idiosyncrasies of a new subject.
-In this article, we ask 31 human subjects to organize 78 isolated notes into a set of timbre clusters.
+In this article, we ask 31 human participants to organize 78 isolated notes into a set of timbre clusters.
 Analyzing their responses suggests that timbre perception operates within a more flexible taxonomy than those provided by instruments or playing techniques alone.
@@
 It reweights scattering coefficients so that pairwise distances between samples more accurately reflect human judgments on a training set.
-These human judgments may be sourced from a single subject or the intersubjective consensus of multiple subjects \cite{mcadams1995psychres}.
+These human judgments may be sourced from a single subject or the intersubjective consensus of multiple participants \cite{mcadams1995psychres}.
 
@@
 \url{http://www.auditory.org/} and \url{http://ismir.net/}}
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.
+Within two months, $K=31$ participants accessed Cyberlioz and completed the task.
+We did not collect personal information on the age, gender, nor musical background of participants.
 
@@
 
+For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
+Figure \ref{fig:cluster-histograms} (a) shows the histogram of $C_k$ across the cohort of $K=31$ participants.
+We observe that the number of clusters varies between $3$ and $19$ with a median value of $10$.
+Accordingly, the number of samples belonging to a cluster varies between $1$ (the most frequent value) and $50$, as shown in Figure \ref{fig:cluster-histograms} (b).
 %Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
@@
 
-We aggregate the similarity judgments from all $K$ subjects by embedding them into a hypergraph $\mathcal{H}$, that is, a graph whose edges may connect three or more vertices at once.
+We aggregate the similarity judgments from all $K$ participants by embedding them into a hypergraph $\mathcal{H}$, that is, a graph whose edges may connect three or more vertices at once.
 Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$.
@@
 
-For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
-Across all $K=31$ subjects, $C_k$ varies between $3$ and $19$, with a median value of $10$.
-To construct $\mathcal{G}_0$, we select a number of clusters that is equal to the maximal value of the $C_k$'s; that is,$C_0 = 19$.
+To construct $\mathcal{G}_0$, we select a number of clusters that is equal to the maximal value of the $C_k$'s; that is, $C_0 = 19$.
 Then, we run hypergraph partitioning on $\mathcal{H}$ to assign each vertex $i$ to one of the $C_0$ clusters in $\mathcal{G}_0$.
@@
 While the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants while discarding intersubjective variability.
-Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ subjects.
+Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ participants.
 This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
@@
 In our protocol, $\boldsymbol{x}$ contains the sound of an isolated musical note and the cluster graph $\mathcal{G}$ encodes auditory similarities within the dataset $\mathcal{X}$.
-Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K$ subjects from our perceptual study (see \lnameref{sec:data-collection} section).
+Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K$ participants from our perceptual study (see \lnameref{sec:data-collection} section).
 
@@
 
-Lastly, the ``average precision at rank $R$'' (henceforth, AP@$R$) is the average value of $\mathrm{P}_{\mathbf{\Phi}}$, for constant $R$, across all $K=31$ subjects from our perceptual study:
+Lastly, the ``average precision at rank $R$'' (henceforth, AP@$R$) is the average value of $\mathrm{P}_{\mathbf{\Phi}}$, for constant $R$, across all $K=31$ participants from our perceptual study:
 \begin{equation}
@@
 
-Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ subjects of the order of $1\%$.
+Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ participants of the order of $1\%$.
 This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
@@
 As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
-Over our cohort of $K=31$ subjects, we report an AP@5 of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
+Over our cohort of $K=31$ participants, we report an AP@5 of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
 This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.
@@
 
-Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves n AP@5 of $\textbf{91.9\%} \pm 4$.
+Over our cohort of $K=31$ participants, separable time--frequency scattering achieves n AP@5 of $\textbf{91.9\%} \pm 4$.
 This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
@@
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
-Over our cohort of $K=31$ subjects, we report an AP@5 of $\textbf{81.8\%} \pm 7$.
+Over our cohort of $K=31$ participants, we report an AP@5 of $\textbf{81.8\%} \pm 7$.
 
@@
 The computational model proposed in this work is well suited to such a query-by-example task.
-We have shown that it is able to adequately approximate the timbral judgments of a wide range of subjects included in our study.
+We have shown that it is able to adequately approximate the timbral judgments of a wide range of participants included in our study.
 Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on the reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
@@
 A more nuanced similarity judgment would ask users to rate the similarity of a pair of IPTs on a more graduated scale, which would yield a finer, or \emph{soft}, assignment.
-This however, comes with additional difficulties in providing a consistent scale across subjects, but could be feasible if the goal is to only adapt the timbral space to a single individual.
+This however, comes with additional difficulties in providing a consistent scale across participants, but could be feasible if the goal is to only adapt the timbral space to a single individual.
 An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assignments.
@@
 \begin{figure}
+\center
+\begin{tabular}{cc}
+  (a) & (b) \\
+  \includegraphics[
+  	width = 0.47\textwidth, trim = 10mm 0 5mm 0, clip]
+	  {nbc.png} &
+  \includegraphics[
+    width = 0.47\textwidth, trim = 10mm 0 5mm 0, clip]
+      {sbc.png}
+\end{tabular}
+\caption{\csentence{Intersubjective}.
+Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects. %% TODO UPDATE
+See \lnameref{sec:data-collection} for details.}
+\label{fig:cluster-histograms}
+\end{figure}
+
+\begin{figure}
 \includegraphics[height=4cm]{Vn-ord-C4-mf-4c_scalogram.pdf}

### 2020-10-31 paper/paper.tex
write on participant demographics

@@
 Within two months, $K=31$ participants accessed Cyberlioz and completed the task.
-We did not collect personal information on the age, gender, nor musical background of participants.
+
+We did not collect personal information on the age, sex, nor musical background of participants.
+This is because the goal of our perceptual study is to build a consensus of similarity judgments, rather than to compare demographic subgroups.
+We note that the relatively small size of our cohort ($K=31$) lacks the statistical power to carry out such a comparison.
+In particular, we leave the important question of the effect of musical training on the  as future work.
 

### 2020-11-01 paper/paper.tex
finish paragraph on musical training
write on the influence of pitch on timbre perception

@@
 We note that the relatively small size of our cohort ($K=31$) lacks the statistical power to carry out such a comparison.
-In particular, we leave the important question of the effect of musical training on the  as future work.
+In particular, we leave the important question of the effect of musical training on the perception of auditory similarities between playing techniques as future work.
+% TODO revise the paragraph above
 
@@
 This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+
+Underlying our label propagation procedure is the assumption that the perception of timbre is fully invariant to frequency transposition as well as changes in dynamics.
+This assumption coincides with the commonplace definition of timbre as an autonomous parameter of musical sound.
+Previous studies on real-world musical sounds have confirmed that listeners are able to ignore salient pitch differences while rating timbre similarity \cite{handel2001musicperception,marozeau2003dependency}, insofar as these differences do not exceed one octave.
+Since then, another study has shown that trained musicians can identify the similarity within pairs of notes from two instruments (horn and bassoon) spanning a range of $2.5$ octave with an accuracy of $80\%$ \cite{steele2006musicperception}.
+In comparison, nonmusicians has great difficulty identifying whether two notes were produced by the same instrument or not when the notes were separated by one octave.
+Thus, the influence of pitch on timbre perception appears to depend upon musical training.
+With this important caveat in mind, we leave as future work the question of disentangling the effect of pitch and musical training in the modeling of auditory similarities between instrumental playing techniques.
 

### 2020-11-01 paper/paper.tex
cite elliot, giordano, siedenburg

@@
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
-In this setting, prior knowledge of instrumentation, however accurate, does not suffice to conduct a study on timbre perception.
-Rather, each excerpt must be played back to multiple human listeners.
+In this setting, prior knowledge of instrumentation, however accurate, does not suffice to conduct a study on timbre perception: rather, the timbre perception relies on an interplay of acoustic and categorical information \cite{siedenburg2016frontiers}.
 Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, which restricts the size of the musical corpus under study.
@@
 
-Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of quantitative ratings.
+Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of continuously valued ratings.
 To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity.\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all 31 participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}
@@
 
-In comparison with a traditional web-based form, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
+Cyberlioz implements a data collection procedure known as ``free sorting''.
+In comparison with the direct collection of timbre dissimilarity ratings, free sorting is more efficient yet less accurate \cite{giordano2011multivariate}.
+We refer to \cite{elliott2013jasa} for an example protocol in which timbre similarity judgments rely on stimuli pairs rather than on a free sorting task.
+
+In comparison with Web-based forms, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
+Another advantage of Cyberlioz is that it allows to present all stimuli at once rather than according to a randomized sequence.
+
 In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
@@
 Thus, the influence of pitch on timbre perception appears to depend upon musical training.
-With this important caveat in mind, we leave as future work the question of disentangling the effect of pitch and musical training in the modeling of auditory similarities between instrumental playing techniques.
+With this important caveat in mind, we leave as future work the question of disentangling the effects of pitch and musical training in the modeling of auditory similarities between instrumental playing techniques.
 
@@
 
-On the other hand, we note that the complexity of the learned model does not have to be high.
+On the other hand, we note that the complexity of the learned model does not need to be high.
 Indeed, for this task, a linear model on the scattering coefficients is sufficient.
@@
 \end{tabular}
-\caption{\csentence{Intersubjective}.
+\caption{\csentence{Inter-subject variability}.
 Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects. %% TODO UPDATE

### 2020-11-01 paper/paper.tex
cite caetano
cite siedenburg 2019 jasa

@@
 In addition, previous publications on joint time--frequency scattering lack a human-centric evaluation, independently from any classification task.
+Beyond the case of STRF, we refer to \cite{caetano2019chapter} for a detailed review of the state of the art on audio descriptors of timbre.
 
@@
 Thus, it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
+
+These findings are in line with a previous publication \cite{patil2012ploscompbiol} which modeled perceived dissimilarity between musical sounds by means of STRF features.
+Likewise, a recent publication has observed similar performance in instrument identification of a human and machine classifier using separable Gabor filterbank (GBFB) features \cite{siedenburg2019jasa}.
+We note that both STRF and GBFB bear a strong computational resemblance with time--frequency scattering.
 

### 2020-11-01 paper/paper.tex
feedback from Reviewer 2

@@
 On the other hand, the sole mention of a playing technique does not specify its effect in terms of auditory perception.
-For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
+For instance, although the term \emph{breathy} alludes to a playing technique that is specific to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
 Yet, in a diverse instrumentarium, the semantic similarity between playing technique denominations does not reflect such acoustical similarity \cite{kolozali2011ismir}.
@@
 Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
+For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music excerpts spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
 % \ja wants 20k, 21k.
@@
 Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, which restricts the size of the musical corpus under study.
-These small corpus sizes hampers the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
+These small corpus sizes hamper the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
 
@@
 \end{equation}
-In the equation above, $\alpha$ is a temporal modulation rate and $\beta$ is a temporal modulation scale, following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:related-work} section).
+In the equation above, $\alpha$ is a temporal modulation rate and $\beta$ is a frequential modulation scale, following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:related-work} section).
 While $\alpha$ is measured in Hertz and is strictly positive, $\beta$ is measured in cycles per octaves and may take positive as well as negative values.
@@
 
-While the the term $\mathcal{E}_{\textrm{pull}}$ operates on pairs of samples, the term $\mathcal{E}_{\textrm{push}}$, operates on triplets $(\boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z})\in\mathcal{X}^3$.
+While the term $\mathcal{E}_{\textrm{pull}}$ operates on pairs of samples, the term $\mathcal{E}_{\textrm{push}}$, operates on triplets $(\boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z})\in\mathcal{X}^3$.
 The first sample, $\boldsymbol{x}$, is known as an ``anchor.''
@@
 
-\subsection*{Semi-supervised label propagation}
+\subsection*{Extension to diverse pitches and dynamics}
 
@@
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
-
-Underlying our label propagation procedure is the assumption that the perception of timbre is fully invariant to frequency transposition as well as changes in dynamics.
+Thus, from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+
+In doing so, we assume that the perception of timbre is fully invariant to frequency transposition as well as changes in dynamics.
 This assumption coincides with the commonplace definition of timbre as an autonomous parameter of musical sound.
@@
 
-Let us denote by $\boldsymbol{x}_1, \ldots, \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}$ audio samples associated with our annotated dataset after semi-supervised label propagation.
+Let us denote by $\boldsymbol{x}_1, \ldots, \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}$ audio samples associated with our annotated dataset.
 Given a sample $n$ and a human subject $k$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.
@@
 Composers may personalize their search engine $\mathbf{\Phi}$ by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
-This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
+This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after extension to all available pitches and dynamics: i.e., roughly three hours of audio.
 

### 2020-11-02 paper/paper.tex
add use case subsection
write on personalization in use case

@@
 
+\subsection*{Use case}
+
 Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted orchestration \cite{maresz2013cmr}.
@@
 The outcome is another audio signal which is selected from a database of instrumental samples and perceptually similar to the query.
+
 The advantage of this search is that, unlike the query, the retrieved sound is precisely encoded in terms of duration, pitch, intensity, instrument, and playing technique.
 Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of predefined range \cite{caetano2019swarm}.
+
+Composers may personalize their search engine by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
+This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in the database: i.e., roughly three hours of audio.
 
@@
 This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity.
-
-Composers may personalize their search engine $\mathbf{\Phi}$ by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
-This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after extension to all available pitches and dynamics: i.e., roughly three hours of audio.
 

### 2020-11-02 future/paper.tex
create future/paper.tex

--- 2020-07-21 paper/paper.tex
@@
-\documentclass{bmcart}
+\documentclass{article}
 
@@
 \usepackage{hyperref}
-\usepackage{musicography}
 \usepackage{pdfpages}
@@
 
-\graphicspath{{figures/}{figures/violin_scattering/}{figures/flute_scattering/}{./}}
-%%% Put your definitions there:
-\startlocaldefs
+\graphicspath{{figures/}{./}}
 
@@
 
-\newcommand{\ipt}{IPT\xspace}
-\newcommand{\ipts}{IPTs\xspace}
-
-\newcommand{\patk}{p\mathrm{@}k}
-
 \newcommand{\Card}{\mathrm{card}}
@@
 
-\newcommand{\lnameref}[1]{%
-\bgroup
-\let\nmu\MakeLowercase
-\nameref{#1}\egroup}
-\newcommand{\fnameref}[1]{%
-\bgroup
-\def\nmu{\let\nmu\MakeLowercase}%
-\nameref{#1}\egroup}
+\title{Judgments of timbre similarity \\ across instruments and playing techniques}
+\author{Mathieu Lagrange, Vincent Lostanlen, Joakim And\'{e}n}
+\date{\today}
 
-\newcommand{\nmu}{}
-
-\endlocaldefs
-
-%%% Begin ...
 \begin{document}
 
-%%% Start of article front matter
-\begin{frontmatter}
+\maketitle
 
-\begin{fmbox}
-\dochead{Research}
+\begin{figure}
+\center
+\includegraphics[width = 0.85\textwidth]{consensusVsI.png}
+\caption{
+Matrix $\mathbf{\widetilde{A}}_{\mathcal{I}}$ of perceived timbre similarities between instruments from our dataset of $N=78$. samples.
+Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:instrument-similarity}), obtained from a consensus of $K=31$ subjects.
+The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ were re-arranged according to Ward's minimum variance method.
+Observe that the block diagonal structure in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ reflects an organological taxonomy of musical instruments: woodwinds, brass, and strings.
+}
+\label{fig:consensusVsI}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Time--Frequency Scattering Accurately Models Auditory Similarities Between Instrumental Playing Techniques}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Specify information, if available,       %%
-%% in the form:                             %%
-%%   <key>={<id1>,<id2>}                    %%
-%%   <key>=                                 %%
-%% Comment or delete the keys which are     %%
-%% not used. Repeat \author command as much %%
-%% as required.                             %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-                        % id of corresponding address, if any
-%   noteref={n1},                        % id's of article notes, if any
-   email={vincent.lostanlen@nyu.edu}            % email address
-]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
-\author[
-   addressref={aff2},
-   email={christian.elhajj@ls2n.fr}
-]{\inits{CE}\fnm{Christian} \snm{El-Hajj}}
-\author[
-   addressref={aff3},
-   email={mathias.rossignol@lonofi.com}
-]{\inits{MR}\fnm{Mathias} \snm{Rossignol}}
-\author[
-   addressref={aff3},
-   email={gregoire.lafay@lonofi.com}
-]{\inits{GL}\fnm{Gr\'egoire} \snm{Lafay}}
-\author[
-   addressref={kth,ccm},
-   email={janden@flatironinstitute.org}
-]{\inits{JA}\fnm{Joakim} \snm{And\'en}}
-\author[
-  corref={aff2},
-  addressref={aff2},
-  email={mathieu.lagrange@ls2n.fr}
-]{\inits{ML}\fnm{Mathieu} \snm{Lagrange}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address[id=aff1]{%                           % unique id
-  \orgname{Music and Audio Research Lab, New York University}, % university, etc
-  \street{370 Jay Street},                     %
-  %\postcode{}                                % post or zip code
-  \city{New York, NY},                              % city
-  \cny{USA}                                    % country
+\begin{figure}
+\center
+\includegraphics[width = 0.85\textwidth]{consensusVsPt.png}
+\caption{
+Matrix $\mathbf{\widetilde{A}}_{\mathcal{T}}$ of perceived timbre similarities between playing techniques from our dataset of $N=78$ samples.
+Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:technique-similarity}), obtained from a consensus of $K=31$ subjects.
+The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{T}}$ were re-arranged according to Ward's minimum variance method.
+%TODO add a double arrow?
+% rough <-- soft --> rough
+% sustained <-- "ordinary" --> percussive
 }
-\address[id=aff2]{%
-  \orgname{LS2N, CNRS, Centrale Nantes, Nantes University},
-  \street{1, rue de la Noe},
-  \postcode{44000},
-  \city{Nantes},
-  \cny{France}
-}
-\address[id=aff3]{%
-  \orgname{Lonofi},
-  \street{57 rue Letort},
-  \postcode{75018},
-  \city{Paris},
-  \cny{France}
-}
-\address[id=kth]{%
-  \orgname{Department of Mathematics, KTH Royal Institute of Technology}
-  \street{Lindstedtsv\"{a}gen 25}
-  \city{SE-100 44, Stockholm}
-  \cny{Sweden}
-}
-\address[id=ccm]{%
-  \orgname{Center for Computational Mathematics, Flatiron Institute},
-  \street{162 5th Avenue},
-  \city{New York, NY 10010},
-  \cny{USA}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{artnotes}
-%\note[id=n1]{Equal contributor}
-\end{artnotes}
-
-\end{fmbox}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% Please refer to the Instructions for     %%
-%% authors on http://www.biomedcentral.com  %%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract}
-Instrumental playing techniques such as vibratos, glissandos, and trills often denote musical expressivity, both in classical and folk contexts.
-However, most existing approaches to music similarity retrieval fail to describe timbre beyond the so-called ``ordinary'' technique, use instrument identity as a proxy for timbre quality, and do not allow for customization to the perceptual idiosyncrasies of a new subject.
-In this article, we ask 31 human subjects to organize 78 isolated notes into a set of timbre clusters.
-Analyzing their responses suggests that timbre perception operates within a more flexible taxonomy than those provided by instruments or playing techniques alone.
-In addition, we propose a machine listening model to recover the cluster graph of auditory similarities across instruments, mutes, and techniques.
-Our model relies on joint time--frequency scattering features to extract spectrotemporal modulations as acoustic features.
-Furthermore, it minimizes triplet loss in the cluster graph by means of the large-margin nearest neighbor (LMNN) metric learning algorithm.
-% \ja: "what?".
-VL: leaving it as is.
-i think it's important to use the words "triplet loss" in the abstract
-Over a dataset of 9346 isolated notes, we report a state-of-the-art precision at rank five (P@5) of $\mathsf{99.0\%\pm1}$.
-An ablation study demonstrates that removing either the joint time--frequency scattering transform or the metric learning algorithm noticeably degrades performance.
-\end{abstract}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{audio databases}
-\kwd{continuous wavelet transform}
-\kwd{demodulation}
-\kwd{distance learning}
-\kwd{human--computer interaction}
-\kwd{music information retrieval.}
-\end{keyword}
-
-\end{abstractbox}
-
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% Please refer to the instructions for     %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Introduction}
-\label{sec:intro}
-
-Music information retrieval (MIR) operates at two levels: symbolic and auditory \cite{downie2003mir}.
-By relying on a notation system, the symbolic level allows the comparison of musical notes in terms of quantitative attributes, such as duration, pitch, and intensity at the source.
-Timbre, in contrast, is a qualitative attribute of music, and is thus not reducible to a one-dimensional axis \cite{siedenburg2019chapter}.
-As a result, symbolic representations describe timbre indirectly, either via visuotactile metaphors (\eg{} bright, rough, and so forth \cite{faure1996icmpc}) or via an instrumental playing technique (\eg{} bowed or plucked) \cite{lostanlen2018extended}.
-
-Despite their widespread use, purely linguistic references to timbre fail to convey the intention of the composer.
-On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them \cite{antoine2018isma}.
-On the other hand, the sole mention of a playing technique does not specify its effect in terms of auditory perception.
-For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
-Yet, such acoustical similarity at the scale of the entire instrumentarium is not directly reflected by the semantic similarity between playing technique denominations; and discovering it requires musical expertise \cite{kolozali2011ismir}.
-
-Although a notation-based study of playing techniques in music has research potential in music information retrieval \cite{calvo2020acm}, the prospect of modeling timbre perception necessarily exceeds the symbolic domain.
-Instead, it involves a cognitive process which arises from the subjective experience of listening \cite{erickson1975book}.
-The simulation of this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates pairs of stimuli.
-Rather than merely discriminating instruments as mutually exclusive categories, this function must reflect judgments of acoustic dissimilarity, all other parameters---duration, pitch, and intensity---being equal \cite{thoret2018jasa}.
-
-Behind the overarching challenge of coming up with a robust predictive model for listening behaviors in humans, the main practical application of timbre similarity retrieval lies in the emerging topic of computer-assisted orchestration \cite{maresz2013cmr}.
-In such context, the composer queries the software with an arbitrary audio signal.
-The outcome is another audio signal which is selected from a database of instrumental samples and perceptually similar to the query.
-The advantage of this search is that, unlike the query, the retrieved sound is precisely encoded in terms of duration, pitch, intensity, instrument, and playing technique.
-Thus, following the aesthetic tradition of spectralism in contemporary music creation, the computer serves as a bridge from the auditory level to the symbolic level, \ie{} from a potentially infinite realm of timbral sensations to a musical score of predefined range \cite{caetano2019swarm}.
-
-\subsection*{Goal}
-This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.
-Crucially, this dissimilarity is not evaluated in terms of acoustic tags, but in terms of \emph{ad hoc} clusters, as defined by a human consensus of auditory judgments.
-Our system consists of two stages: unsupervised feature extraction and supervised metric learning.
-The feature extraction stage is a nonlinear map which relies on the joint time--frequency scattering transform \cite{anden2015mlsp,anden2019tsp}, followed by per-feature Gaussianization \cite{lostanlen2018jasmp}.
-It encodes patterns of spectrotemporal modulation in the acoustic query while offering numerical guarantees of stability to local deformations \cite{anden2012scattering}.
-The metric learning stage is a linear map, optimized via large-margin nearest neighbors (LMNN) \cite{weinberger2009distance}.
-It reweights scattering coefficients so that pairwise distances between samples more accurately reflect human judgments on a training set.
-These human judgments may be sourced from a single subject or the intersubjective consensus of multiple subjects \cite{mcadams1995psychres}.
-
-Figure \ref{fig:pipeline} summarizes our experimental protocol: it illustrates how visual annotations (top) can inform feature extraction (center) to produce a nearest-neighbor search engine which is consistent with human judgments of timbre similarity (bottom).
-
-\subsection*{Contributions}
-The main contribution of this article can be formulated as the intersection between three topics.
-To the best of our knowledge, prior literature has addressed these topics separately, but never in combination.
-
-First, our model encompasses a broad range of extended playing techniques, well beyond the so-called ``ordinary'' mode of acoustic production.
-Specifically, we fit pairwise judgments for 78 different techniques arising from 16 instruments, some of which include removable timbre-altering devices such as mutes.
-
-Secondly, we purposefully disregard the playing technique metadata underlying each audio sample during the training phase of our model.
-In other words, we rely on listeners, not performers, to define and evaluate the task at hand.
-
-Thirdly, we supplement our quantitative benchmark with visualizations of time--frequency scattering coefficients in the rate--scale domain for various typical samples of instrumental playing techniques.
-These visualizations are in line with visualizations of the modulation power spectrum in auditory neurophysiology \cite{patil2012ploscompbiol}, while offering an accelerated algorithm for scalable feature extraction.
-
-\subsection*{Outline}
-The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
-The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
-The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-The \lnameref{sec:results} section describes our experimental benchmark.
-The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
-The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
-
-\section*{\nmu Related work}
-\label{sec:related-work}
-
-Timbre involves multiple time scales in conjunction, from a few microseconds for an attack transient to several seconds for a sustained tone.
-Therefore, computational models of timbre perception must summarize acoustic information over a long analysis window, typically amounting to $10^{4}$ digital audio samples or more \cite{joder2009taslp}.
-Mapping this input to a feature space in which distances denote timbral dissimilarity requires a data-driven stage of dimensionality reduction.
-In this respect, the scientific literature exhibits a methodological divide as regards the collection of human-annotated data \cite{siedenburg2016jnmr}: while the field of MIR mostly encodes timbre under the form of ``audio tags'', music psychology mostly measures timbre similarity directly from pairwise similarity judgments.
-
-\subsection*{Timbre modeling in music information retrieval (MIR)}
-On the one hand, most publications in music information retrieval cast timbre modeling as an audio classification problem  \cite{martin1998asa,brown1999jasa,eronen2000icassp,herrera2003jnmr,wieczorkowska2003jiis,livshin2004dafx,krishna2004icassp,kaminskyj2005jiis,benetos2006icassp,bhalke2016jiis}.
-In this context, the instrumentation of each musical excerpt serves as an unstructured set of ``audio tags,'' encoded as binary outputs within some predefined label space.
-Because such tags often belong to the metadata of music releases, the process of curating a training set for musical instrument classification requires little or no human intervention.
-Although scraping user-generated content from online music platforms may not always reflect the true instrumentation with perfect accuracy, it offers a scalable and ecologically valid insight onto the acoustic underpinnings of musical timbre.
-
-Furthermore, supplementing user-generated content with the outcome of a crowdsourced annotation campaign allows an explicit verification of instrument tags.
-For instance, the Open-MIC dataset \cite{humphrey2018ismir}, maintained by the Community for Open and Sustainable Music Information Research (COSMIR) \cite{mcfee2016ismir}, comprises a vast corpus of 20k polyphonic music recordings spanning 20 instruments as a derivative of the Free Music Archive (FMA) dataset \cite{defferrard2017ismir}.
-% \ja wants 20k, 21k.
-Another example is the Medley-solos-DB dataset \cite{lostanlen2016ismir}, which comprises 21k monophonic excerpts from eight instruments as a derivative of the MedleyDB dataset of multitrack music \cite{bittner2014ismir}.
-
-Over the past decade, the availability of large digital audio collections, together with the democratization of high-performance computing on dedicated hardware, has spurred the development of deep learning architectures in music instrument recognition \cite{mcfee2015ismir,pons2017eusipco,gururani2018ismir}.
-Notwithstanding the growing accuracy of these architectures in the large-scale data regime, it remains unclear how to extend them from musical instrument recognition to playing technique recognition, where labeled samples are considerably more scarce \cite{loureiro2004ismir}.
-We refer to \cite{han2017taslp} for a recent review of the state of the art in this domain.
-
-\subsection*{Timbre modeling in music cognition research}
-On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
-In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
-Rather, each excerpt must be played back to multiple human listeners.
-Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, which restricts the size of the musical corpus under study.
-These small corpus sizes hampers the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
-
-While training artificial neurons is prone to statistical overfitting, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
-This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
-The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
-The simplest method to compute it in practice is by reverse correlation, i.e.\ by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
-Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but eliciting cross-term interferences \cite{eggermont1993hearing}.
-Since then, the STRF of a neuron was redefined as a spectrographic representation of its spike-triggered average \cite{klein2000compneur}.
-
-Although this new definition is necessarily tied to a choice of spectrogram parameters, it yields more interpretable patterns than a Wigner--Ville distribution.
-In particular, a substantial portion of spectrographic STRFs exhibit a ripple-like response around a given region $(t, \lambda)$ of the time--frequency domain \cite{theunissen2000jneur}.
-This response can be approximately described by a pair of scalar values: a temporal modulation rate $\alpha$ in Hertz and a frequential modulation rate $\beta$ in cycles per octave.
-
-Interestingly, both $\alpha$ and $\beta$ appear to be arranged in a geometric series and independent from the center time $t$ and center frequency $\lambda$.
-This observation has led auditory neuroscientists to formulate an idealized computational model for STRF, known as the ``full cortical model'' \cite{chi2005jasa}, which densely covers the rate--scale domain $(\alpha, \beta)$ using geometric progressions.
-Because they do not require a data-driven training procedure, STRF yield a useful form of domain-specific knowledge for downstream machine listening applications, especially when the number of annotated samples is relatively small.
-
-\subsection*{Spectrotemporal receptive fields (STRFs) as a feature extractor}
-Over recent years, several publications have employed the full cortical model as a feature extractor for a task of musical instrument classification, both in isolated recordings \cite{patil2012ploscompbiol} and in solo phrases \cite{patil2015eurasip}.
-These biologically inspired features outperform the state of the art, especially in the small data regime where deep learning is inapplicable.
-Furthermore, the confusion matrix of the full cortical model in the label space of musical instruments is strongly correlated with the confusion matrix between a human listener and the ground truth.
-Another appeal of the full cortical model is that the three-way tensor of frequency $\lambda$, rate $\alpha$, and scale $\beta$ can be segmented into contiguous regions of maximal perceptual relevance for each instrument \cite{thoret2016jasa}.
-This is unlike fully end-to-end learning architectures, whose post hoc interpretability requires advanced techniques for feature inversion \cite{mishra2018ismir}.
-Lastly, beyond the realm of supervised classification, a previous publication \cite{hemery2015frontiers} has shown that query-by-example search with STRFs allows to discriminate categories of environmental soundscapes, even after temporal integration and unsupervised dimensionality reduction.
-
-The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
-Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
-Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
-This approach requires a uniform discretization of the scalogram, and thus an oversampling of the lower-frequency subbands to the Nyquist frequency of the original signal.
-In contrast, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2020jmlr} and invertibility \cite{lostanlen2019dafx}.
-Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
-In this multirate scheme, every subband is discretized at its critical sample rate, i.e., in proportion to its center frequency.
-As a by-product, the multirate approach draws an explicit connection between scattering networks and deep convolutional networks, because both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and pooling operations \cite{mallat2016philtrans}.
-
-Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class \cite{anden2019tsp}.
-% \ja wants 5,000
-However, it remains to be seen whether joint time--frequency scattering is capable of fine-grained auditory categorization, involving variability in instrument, mute, and playing technique.
-In addition, previous publications on joint time--frequency scattering lack a human-centric evaluation, independently from any classification task.
-
-\subsection*{Claim of originality}
-The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is also readily tailored to subjective preferences.
-
-\section*{\nmu Data collection}
-\label{sec:data-collection}
-
-The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
-In this article, we focus on the first three, and leave the question of learning auditory similarities between percussion instruments to future research.
-We refer to \cite{wu2018taslp} and \cite{pearce2019appliedsciences} for reviews of the recent literature on the timbre modeling of percussive instruments, from the standpoints of MIR and music cognition, respectively.
-
-\subsection*{Dataset}
-We consider a list of 16 instruments: violin (Vn), viola (Va), cello (Vc), contrabass (Cb), concert harp (Hp), Spanish guitar (Gtr), accordion (Acc), flute (Fl), soprano clarinet (BbCl), alto saxophone (ASax), oboe (Ob), bassoon (Bn), trumpet in C (TpC), French horn (Hn), tenor trombone (TTbn), and bass tuba (BBTb).
-Among this list, the first six are strings, the next six are woodwind, and the last four are brass.
-Some of these instruments may be temporarily equipped with timbre-altering mutes, such as a rubber sordina on the bridge of a violin or an aluminium ``wah-wah'', also known as \emph{harmon}, inside in the bell of a trumpet.
-Once augmented with mutes, the list of 16 instruments grows to 33.
-Furthermore, every instrument, whether equipped with a mute or not, affords a panel of playing techniques ranging in size between 11 (for the accordion) and 41 (for the bass tuba).
-In the rest of this paper, we abbreviate instrument--mute--technique by means of the acronym ``IMT''.
-One example of IMT is \texttt{TpC+S-ord}, \ie{} trumpet in C with a straight mute played in the ordinary technique.
-Another example of IMT is \texttt{Vn-pont}, \ie{} violin without any mute played in the \emph{sul ponticello} technique (bowing near the bridge).
-
-Performers can play each IMT at various pitches according to the tessitura of their instrument.
-This tessitura may depend on the choice of playing technique but is independent of the choice of mute.
-Among the 16 instruments in this study, the two instruments with widest and narrowest tessituras, in their respective ordinary techniques, are the accordion (81 semitones) and the trumpet in C (32 semitones) respectively.
-Lastly, each IMT may be played at up to five intensity dynamics, ranging from quietest to loudest as: pianissimo (\emph{pp}), piano (\emph{p}), mezzo forte (\emph{mf}), forte (\emph{f}), and fortissimo (\emph{ff}).
-The resort to a non-ordinary playing technique may restrict both the tessitura and the dynamics range of the instrument--mute pair under consideration.
-For example, the pitch of pedal tones in brass instruments is tied to the fundamental mode of the bore, \ie{} usually $\mathsf{B}\musFlat$ or $\mathsf{F}$.
-Likewise, the intensity of key clicks in the oboe is necessarily \emph{pp}, while the intensity of snap pizzicato \emph{\`a la Bart\'ok} in plucked strings is necessarily \emph{ff}.
-
-In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two quantitative variables (intensity and pitch).
-The Studio On Line dataset (SOL), recorded at Ircam in 1998, offers a joint sampling of these variables.
-The version of SOL that we use throughout this paper, named ``0.9 HQ'', amounts to a total of 25444 audio signals.
-Beyond playing techniques, we should note that SOL erases other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics, which are all restricted to singletons.
-Addressing these factors of variability is beyond the scope of this paper, which focuses on the influence of playing technique.
-Despite this restriction, the SOL dataset remains impractically large for collecting human similarity judgments.
-Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and the use of an efficient annotation interface.
-
-\subsection*{Disentanglement of factors}
-First, we purposefully disentangle categorical variables (IMTs) from continuous variables (pitch and intensity) in the SOL dataset.
-Indeed, under first approximation, the perception of timbre is invariant to pitch and intensity.
-Therefore, we select auditory stimuli according to a reference pitch and a reference intensity; in our case, middle C ($\mathrm{C_4}$) and \emph{mf}.
-After this selection, every IMT triplet contains a single acoustic exemplar, regarded as canonical in the following.
-The number of canonical stimuli for the entire SOL dataset is equal to 235.
-We should note, however, that the proposed pitch and intensity cannot be strictly enforced across all IMTs.
-Indeed, as explained above, a fraction of IMTs can only be achieved at restricted values of pitch and intensity parameters, \eg{} pedal tones or key clicks.
-Therefore, at a small cost of consistency, we only enforce the pitch--intensity reference (\ie{} $\mathrm{C_4}$ and \emph{mf}) when practically feasible, and fall back to other pitches and intensities if necessary.
-
-\subsection*{Expert pre-screening}
-Secondly, we reduce the number of IMTs in our study by focusing on those which are deemed to be most relevant.
-Here, we define the relevance of an IMT as the possibility of imitating it by means of another IMT from a different instrument.
-One example of such imitation is the acoustic similarity between slap tonguing in reed instruments and a snap pizzicato in string instruments.
-To collect perceptual ratings of relevance, we recruited two professors in music composition at the Paris Conservatory (CNSMDP\footnote{CNSMDP: Conservatoire National Sup\'erieur de Musique et de Danse de Paris.
-Official website: \url{https://www.conservatoiredeparis.fr}}).
-Each of them inspected the entire corpus of $235$ IMTs and annotated them in terms of relevance according to a Likert scale with seven ticks.
-In this Likert scale, the value 1 (least relevant) denotes that the IMT under consideration has a timbre that is idiosyncratic, and that therefore, it is unlikely that humans will pair it with other IMTs.
-Conversely, the value 7 (most relevant) denotes that the IMT under consideration bears a strong similarity with some other IMT from the corpus.
-
-Once both experts completed their annotations, we retained all IMTs whose average score was judged equal to 3 or higher, thus resulting in a shortlist of $N=78$ IMTs (see Tables 1 and 2 in the Appendix).
-It is worth noting that, according to both experts, the timbre of the accordion was judged too idiosyncratic to be relevant for this experiment, regardless of playing technique.
-Indeed, the accordion is the only instrument in the aforementioned list of
-instrument to have free reeds, keyboard-based actuation, or handheld airflow.
-Consequently, regardless of mute and technique, the set of instruments $\mathcal{I}$ in our study contains 15 elements.
-
-\subsection*{Efficient annotation interface}
-Thirdly, we design a graphical user interface for partitioning a corpus of short audio samples.
-The need for such an interface arises from the unscalability of Likert scales in the context of pairwise similarity judgments.
-Assuming that similarity is a symmetric quantity, collecting a dense matrix of continuously valued ratings of similarity among a dataset of $N$ items would require $\frac{1}{2}(N^2-N)$ Likert scales.
-In the case of $N=78$ IMTs, the task would amount to about 3k horizontal sliders, \ie{} several hours of cumbersome work for the human annotator. % \ja wants 3,000
-
-Engaging as many participants as possible in our study called for a more streamlined form of human--computer interaction, even if it sacrificed the availability of quantitative ratings.
-To this end, we implemented a web application, named Cyberlioz, in which the user can spontaneously listen and arrange sounds into clusters of timbre similarity.\footnote{The Web application for efficient audio annotation, as well as the raw anonymized responses of all 31 participants to our study, is available at: \url{https://soundthings.org/research/cyberlioz/}}
-The name Cyberlioz is a portmanteau between the prefix cyber- and the French composer Hector Berlioz.
-The choice is by no means coincidental: Berlioz is famous for having, in his \emph{Treatise on Orchestration} (1844), shed a particular focus on the role of timbre as a parameter for musical expression.
-
-Cyberlioz consists of a square panel on which is displayed a collection of circular grey dots, each of them corresponding to one of the IMTs, and initially distributed uniformly at random.
-Hovering the screen pointer onto each dot results in a playback of a representative audio sample of this IMT, \ie{} $\mathrm{C}_4$ and \emph{mf} in most cases.
-Furthermore, each dot can be freely placed on the screen by clicking, dragging, and dropping.
-Lastly, the user can assign a color to each dot among a palette of 20 hues.
-The goal of the Cyberlioz interface is to form clusters of timbre similarity between IMTs, expressed by sameness of color.
-
-In comparison with a traditional web-based form, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
-In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
-To this end, we publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
-For more information about these mailing lists, please visit:
-\url{http://www.auditory.org/} and \url{http://ismir.net/}}
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.
-
-\subsection*{Hypergraph partitioning}
-Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a cluster graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
-For a given $k$, the graph $\mathcal{G}_k$ contains $N$ vertices, each representing a different IMT in the corpus.
-In $\mathcal{G}_k$, an edge connects any two vertices $m$ and $n$ if the corresponding dots in Cyberlioz have the same color.
-Otherwise, there is no edge connecting $m$ and $n$.
-Thus, $\mathcal{G}_k$ contains as many connected components as the number of similarity clusters for the subject $k$, \ie{} the number of distinct colors on the Cyberlioz interface in the response of $k$.
-
-%Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
-The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
-We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
-
-We aggregate the similarity judgments from all $K$ subjects by embedding them into a hypergraph $\mathcal{H}$, that is, a graph whose edges may connect three or more vertices at once.
-Specifically, $\mathcal{H}$ contains $N$ vertices, each representing an IMT; and each ``hyperedge'' in $\mathcal{H}$ corresponds to some connected component in one of the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$.
-Then, we convert the hypergraph $\mathcal{H}$ back into a conventional graph $\mathcal{G}_0$ by means of a combinatorial optimization algorithm known as hypergraph partitioning \cite{kernighan1970efficient}.
-
-For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
-Across all $K=31$ subjects, $C_k$ varies between $3$ and $19$, with a median value of $10$.
-To construct $\mathcal{G}_0$, we select a number of clusters that is equal to the maximal value of the $C_k$'s; that is,$C_0 = 19$.
-Then, we run hypergraph partitioning on $\mathcal{H}$ to assign each vertex $i$ to one of the $C_0$ clusters in $\mathcal{G}_0$.
-Intuitively, hypergraph partitioning optimizes a tradeoff between two objectives: first, balancing the size of all clusters in terms of their respective numbers of vertices; and secondly, keeping most hyperedges enclosed within as few distinct clusters as possible \cite{han1997scalable,strehl2002cluster}.
-
-While the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants while discarding intersubjective variability.
-Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ subjects.
-This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
+\label{fig:consensusVsPt}
+\end{figure}
 
@@
 
-Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
-Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
+We ask whether human listeners tend to cluster audio samples by instrument.
 To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
@@
 \end{equation}
-%\ml{why 2 ?} \vl: fixed
 Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
@@
 
-\section*{\nmu Methods}
-\label{sec:methods}
-The previous section described our protocol for collecting timbral similarity judgments between instrumental playing techniques.
-In this section, we aim to recover these similarity judgments from digital audio recordings according to a paradigm of supervised metric learning.
-To this end, we present a machine listening system composing joint time--frequency scattering and large-margin nearest neighbors (LMNN).
-
-\subsection*{Joint time--frequency scattering transform}
-\label{sec:scattering}
-
-%While cortical representations have been successful in distinguishing properties of audio recordings (including similarity analysis and classification tasks), they remain models for the electrophysiological signals recorded in the audio cortex.
-%As such, their calculation is not always computationally efficient and their mathematical properties may be difficult to analyze directly.
-
-%A closely related representation is that of the time scattering transform, introduced by And\'{e}n and Mallat \cite{anden2011multiscale,anden2014tsp}.
-%Given an audio recording, it decomposes the signal using an analytic wavelet transform and computes the modulus, obtaining the \emph{scalogram} of the signal.
-%The scalogram is similar to the spectrogram, but has a time--frequency resolution that varies with frequency: at low frequency, the filters are narrowly tuned in frequency (and hence wide in time), while at high frequency, the filters are wide in frequency (and hence narrowly concentrated in time).
-%This representation contains a large amount of information, but not in a form suitable for building statistical models.
-%In particular, the scalogram is not \emph{stable} in the sense that small deformations to the original signal may induce large changes to the scalogram.
-%It also lacks \emph{invariance} to time-shifting, which does not affect our perception of a sound, and should therefore not affect its timbre model.
-%One way to resolve this is to average the scalogram in time to obtain a spectral profile of the signal, yielding the first-order scattering coefficients.
-%These possess both invariance to time-shifting and stability to time-warping deformations \cite{anden2014tsp}.
-
-%Since these first-order coefficients only include spectral information, they cannot accurately characterize more sophisticated structure in the signal.
-%One way to augment them is to calculate another wavelet decomposition in time, this time on each frequency channel of the scalogram.
-%Again, we compute the modulus of the wavelet coefficients and average in time.
-%The resulting coefficients are known as the second-order time scattering coefficients.
-%These also possess the desired invariance and stability properties, but capture more information on the temporal structure of the scalogram \cite{anden2014tsp}.
-
-%While richer than a simple spectral decomposition, such as MFCCs, the second-order time scattering coefficients described above primarily capture structure in time.
-%Indeed, they do not explicitly encode frequency structure in an invariant and stable manner.
-%If a signal is transposed or warped in frequency by a small amount, we do not expect its timbre to change significantly.
-%We would therefore like our representation to have the same invariance and stability properties along the frequency axis.
-%This is achieved by taking first- and second-order scattering coefficients and computing a second scattering transform, this time along the log-frequency axis.
-%The resulting coefficients are known as separable time and frequency scattering coefficients \cite{anden2014tsp}.
-
-%This representation does possess the desired invariance and stability properties, but its separable nature limits its power to adequately represent time--frequency structures that do not separate cleanly in time and frequency, such as chirps.
-%To remedy this, another approach was suggested by And\'{e}n et al.
-\cite{anden2015mlsp,anden2019tsp} wherein the temporal wavelet decomposition of the scalogram is replaced with a two-dimensional, spectro-temporal decomposition.
-%Similarly to the cortical representation, the result is a four-dimensional array indexed by time, acoustic frequency, modulation frequency (or \emph{rate}), and spectral modulation frequency, known as \emph{quefrency} (or \emph{scale}).
-%As before, we compute the modulus and average in time, obtaining a set of coefficients known as second-order joint time--frequency scattering coefficients.
-
-%The definition of the scattering transform in terms of wavelet decompositions and modulus nonlinearities presents several advantages.
-%On the computational side, the critical bandwidth guarantees of wavelet transforms allows us to judiciously subsample the output, resulting in a lower-dimensional representation compared to the cortical representations.
-%The scattering transform also satisfies the aforementioned stability conditions, providing a guarantee that slightly deforming a signal only results in a negligible change in its scattering transform.
-%Furthermore, we may calculate the scattering transforms of several model signals, including amplitude-modulated harmonic sounds and noise excitations \cite{anden2012scattering,anden2014tsp}, frequency-modulated sinusoids (e.g., chirps) \cite{anden2012scattering,anden2019tsp}, and beating phenomena \cite{anden2014tsp}.
-%They have also enjoyed success in several audio classification and similarity retrieval tasks \cite{anden2011multiscale,anden2014tsp,anden2019tsp,lostanlen2018jasmp,lostanlen2018extended}.
-
-Let $\boldsymbol{\psi} \in \mathbf{L}^2(\mathbb{R}, \mathbb{C})$ a complex-valued filter with zero average, dimensionless center frequency equal to one, and an equivalent rectangular bandwidth (ERB) equal to $1/Q$.
-We define a constant-$Q$ wavelet filterbank as the family $\boldsymbol{\psi}_{\lambda} : t \mapsto \lambda \boldsymbol{\psi}(\lambda t)$.
-Each wavelet $\boldsymbol{\psi}_{\lambda}$ has a center frequency of $\lambda$, an ERB of $\lambda/Q$, and an effective receptive field of $(2\pi Q/\lambda)$ in the time domain.
-In practice, we define $\boldsymbol{\psi}$ as a Morlet wavelet:
-\begin{equation}
-\boldsymbol{\psi}:t \longmapsto \exp\left(-\dfrac{t^2}{2\sigma_{\psi}^2}\right)
-\big( \exp\left(2\pi \mathrm{i}t\right) - \kappa_{\psi} \big),
-\label{eq:psi}
-\end{equation}
-where the Gaussian width $\sigma_{\psi}$ grows in proportion with the quality factor $Q$ and the corrective term $\kappa_{\psi}$ ensures that $\boldsymbol{\psi}$ has a zero average.
-Moreover, we discretize the frequency variable $\lambda$ according to a geometric progression of common ratio $2^{\frac{1}{Q}}$.
-Thus, the base-two logarithm of center frequency, denoted by $\log_2 \lambda$, follows an arithmetic progression.
-We set the constant quality factor of the wavelet filterbank $(\boldsymbol{\psi}_\lambda)_\lambda$ to $Q=12$, thus matching twelve-tone equal temperament in music.
-
-Convolving the wavelets in this filterbank with an input waveform $\boldsymbol{x}\in\mathbf{L}^2(\mathbb{R})$, followed by an application of the pointwise complex modulus yields the wavelet scalogram
-\begin{equation}
-\mathbf{U_1}\boldsymbol{x}(t,\lambda) =
-\big\vert
-\boldsymbol{x}
-\ast
-\boldsymbol{\psi}_{\lambda}
-\big\vert(t)
-=
-\left\vert
-\int_{\mathbb{R}}
-\boldsymbol{x}(t - t^{\prime})
-\,
-\boldsymbol{\psi}_{\lambda} (t^{\prime})
-\;
-\mathrm{d}{t^\prime}
-\right\vert,
-\label{eq:U1}
-\end{equation}
-which is discretized similarly to the constant-$Q$ transform of \cite{schorkhuber2010smc}.
-Then, we define a two-dimensional Morlet wavelet $\boldsymbol{\Psi}$ of the form
-\begin{equation}
-\Psi : (t, u) \longmapsto
-\exp\left(-\dfrac{t^2+u^2}{2\sigma_{\Psi}^2}\right)
-\Big( \exp\big(2\pi \mathrm{i} (t + u)\big) - \kappa_{\Psi} \Big),
-\label{eq:Psi}
-\end{equation}
-taking two real variables $t$ and $u$ as input.
-In the rest of this paper, we shall refer to $\boldsymbol{\Psi}$ as a time--frequency wavelet.
-The former is the time variable while the latter is the base-two logarithm of frequency: $u = \log_2 \lambda$.
-Note that $u$ roughly corresponds to the human perception of relative pitch \cite{lostanlen2020icassp}.
-
-We choose the Gaussian width $\sigma_{\Psi}$ in Equation \ref{eq:Psi} such that the quality factor of the Morlet wavelet $\Psi$ is equal to one, both over the time dimension and over the log-frequency dimension.
-Furthermore, the corrective term $\kappa_{\Psi}$ ensures that $\boldsymbol{\Psi}$ has a zero average over $\mathbb{R}^2$, similarly to Equation \ref{eq:psi}.
-From $\mathbf{\Psi}$, we define a two-dimensional wavelet filterbank of the form:
-\begin{equation}
-\mathbf{\Psi}_{\alpha,\beta} : (t, u) \longmapsto
-\alpha \, \beta \, \mathbf{\Psi}(\alpha t, \beta u).
-\end{equation}
-In the equation above, $\alpha$ is a temporal modulation rate and $\beta$ is a temporal modulation scale, following the terminology of spectrotemporal receptive fields (STRF, see \lnameref{sec:related-work} section).
-While $\alpha$ is measured in Hertz and is strictly positive, $\beta$ is measured in cycles per octaves and may take positive as well as negative values.
-Both $\alpha$ and $\beta$ are discretized by geometric progressions of common ratio equal to two.
-Furthermore, the edge case $\beta=0$ corresponds to $\mathbf{\Psi}_{\alpha,\beta}$ being a Gaussian low-pass filter over the log-frequency dimension, while remaining a Morlet band-pass filter of center frequency $\alpha$ over the time dimension.
-We denote this low-pass filter by $\boldsymbol{\phi}_F$ and set its width to $F=2$ octaves.
-
-We now convolve the scalogram $\mathbf{U_1}$ with time--frequency wavelets $\mathbf{\Psi}_{\alpha,\beta}$ and apply the complex modulus, yielding the four-way tensor
-\begin{align}
-\mathbf{U_2}\boldsymbol{x} (t, \lambda, \alpha, \beta)
-=&
-\big \vert
-\mathbf{U_1}\boldsymbol{x} \circledast \boldsymbol{\Psi}_{\alpha,\beta}
-\big \vert (t, \lambda)
-\nonumber \\
-=&
-\left\vert
-\iint_{\mathbb{R}^2}
-\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)\,
-\boldsymbol{\Psi}_{\alpha,\beta}(t^\prime, u^\prime)\;
-\mathrm{d}t^\prime \, \mathrm{d}u^\prime
-\right\vert,
-\label{eq:U2}
-\end{align}
-where the circled asterisk operator $\circledast$ denotes a joint convolution over time and log-frequency.
-In the equation above, the sample rate of time $t$ is proportional to $\alpha$.
-Conversely, the sample rate of log-frequency $u = \log_2 \lambda$ is proportional to $\vert\beta\vert$ if $\beta \neq 0$ and proportional to $F^{-1}$ otherwise.
-
-Let $\boldsymbol{\phi}_T$ be a Gaussian low-pass filter.
-We define the joint time--frequency scattering coefficients of the signal $\boldsymbol{x}$ as the four-way tensor
-\begin{align}
-\mathbf{S_2}(t, \lambda, \alpha, \beta)
-=&
-\mathbf{U_2}\boldsymbol{x}
-\circledast
-\big(\boldsymbol{\phi}_T  \otimes \boldsymbol{\phi}_F\big)
-(t, \lambda)
-\nonumber \\
-=&
-\iint_{\mathbb{R}^2}
-\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime, \alpha, \beta)
-\,\boldsymbol{\phi}_T (t^\prime)
-\,\boldsymbol{\phi}_F (u^\prime)
-\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
-\label{eq:S2}
-\end{align}
-where the symbol $\otimes$ denotes the outer product over time and log-frequency.
-In the equation above, the sample rate of time $t$ is proportional to $T^{-1}$ and the sample rate of log-frequency $u=\log_2 \lambda$ is proportional to $F^{-1}$.
-Furthermore, the rate $\alpha$ spans along a geometric progression ranging from $T^{-1}$ to $\lambda/Q$.
-In the following, we set the time constant to $T=\SI{1000}{\milli\second}$ unless specified otherwise.
-
-The tensor $\mathbf{S_2}$ bears a strong resemblance with the idealized response of an STRF at the rate $\alpha$ and the scale $\beta$.
-Nevertheless, in comparison with the ``full cortical model'' \cite{patil2012ploscompbiol}, joint time--frequency scattering enjoys a thirty-fold reduction in dimensionality while covering a time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
-This is due to the multirate discretization scheme applied throughout the application of wavelet convolutions and pointwise modulus nonlinearities.
-% \ja says "where does this number come from"? Answer: from the Patil paper vs the scattering.m implementation
-
-In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
-\begin{align}
-\mathbf{S_1}\boldsymbol{x}(t, \lambda, \alpha=0, \beta) &=
-\big\vert
-\mathbf{U_1}\boldsymbol{x} \circledast
-\big(\boldsymbol{\phi}_T \otimes \boldsymbol{\psi}_{\beta} \big)
-\big\vert (t, \lambda)
-\nonumber \\
-&=
-\left \vert
-\iint_{\mathbb{R}^2}
-\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
-\,\boldsymbol{\phi}_T (t^\prime)
-\,\boldsymbol{\psi}_{\beta} (u^\prime)
-\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime
-\right \vert,
-\label{eq:S1}
-\end{align}
-where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.
-
-Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
-Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\phi}_F$.
-We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.
-
-We adopt the multi-index notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
-The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) and second-order ($\mathbf{S_2}$) coefficients.
-Given an input waveform $\boldsymbol{x}$, we denote by $\mathbf{S}\boldsymbol{x}$ the feature vector resulting from the concatenation of $\mathbf{S_1}\boldsymbol{x}$ and $\mathbf{S_2}\boldsymbol{x}$:
-\begin{equation}
-\mathbf{S}\boldsymbol{x}\big(t, p=(\lambda, \alpha, \beta)\big)
-=\left\{
-\begin{array}{ll}
-\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha=0,\\
-\mathbf{S_{2}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{otherwise.}
-\end{array}\right.
-\end{equation}
-
-%An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{x}\circ\boldsymbol{\tau})$ can be bounded as follows:
-%\begin{equation}
-%\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
-%\leq
-%\kappa_{\mathbf{S}} \times
-%\Vert \boldsymbol{x} \Vert^2 \times
-%\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
-%\label{eq:lipschitz}
-%\end{equation}
-%where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
-%In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
-%In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such a deformation.
-%
-%Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
-%In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
-%Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.
-
-\subsection*{Visualization}
-\label{sec:visualization}
-
-At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
-From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
-To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
-In accordance with their protocol, we compute the following matrix:
-
-\begin{equation}
-\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
-\iint_{\mathbb{R}^2}
-\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
-\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
-\end{equation}
-
-Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
-As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
-
-Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
-However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
-Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
-
-To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
-By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
-For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
-A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
-As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
-Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
-
-The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
-Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
-These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
-These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
-Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
-This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
-
-Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
-Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
-Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
-We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
-
-\subsection*{Median-based logarithmic compression and affine standardization}
-\label{sec:gaussianization}
-
-%It follows from the above that comparing Euclidean distances in the feature space of time--frequency scattering coefficients may serve as a computationally efficient surrogate for comparing Wasserstein distances in the scalogram, without need for solving an optimal transport problem.
-%However, this surrogate is only valid if the underlying transport plan $\boldsymbol{\tau}$ is relatively small  in terms of supremum norm (see Equation \ref{eq:lipschitz}).
-%This assumption may be true for certain local deformation of the time-frequency domain, \eg{} causing a pitch shift or a change in vibrato rate while leaving the instrument unchanged \cite{anden2012scattering}.
-%However, it does not suffice in order to represent timbre similarity across multiple instruments and playing techniques.
-%Indeed, the perception of timbre for extended playing techniques often involves long-range interactions in the time--frequency domain, which are not amenable to a local elastic deformation \cite{lostanlen2018extended}.
-
-%To address this issue, we apply a metric learning algorithm, named large-margin nearest neighbors (LMNN), whose goal is to reweight and recombine scattering paths $p$ such that certain factors of variability in feature space $\mathbf{S}$ are enhanced while others are discarded.
-Now, we apply a pointwise nonlinear transformation on averaged joint time-frequency scattering coefficients.
-The role of this transformation, which is adapted to the dataset in an unsupervised way, is to Gaussianize the histogram of amplitudes of each scattering path $p$.
-We consider a collection $\mathcal{X}$ of $N$ waveforms $\boldsymbol{x}_1,\ldots,\boldsymbol{x}_N$.
-For every path $p$ in the joint time--frequency scattering transform operator $\mathbf{S}$, we average the response of each scattering coefficient $\mathbf{S}\boldsymbol{x}_n$ over time and take its median value across all samples $n$ from $1$ to $N$:
-\begin{equation}
-\boldsymbol{\mu}(p) =
-\median_{1\leq n \leq N}
-\int_{\mathbb{R}} \mathbf{S}\boldsymbol{x}_n(t, p) \;\mathrm{d}t.
-\end{equation}
-If the collection is split between a training set and a test set (see \lnameref{sec:discussion} section), we compute $\boldsymbol{\mu}$ on the training set only.
-Then, to match a decibel-like perception of loudness, we apply the following adaptive transformation, which composes a median-based renormalization and a logarithmic compression:
-\begin{equation}
-\mathbf{\widetilde{S}}\boldsymbol{x}_n (p) =
-\log \left(
-1 +
-\dfrac{
-\int_\mathbb{R} \mathbf{S}\boldsymbol{x}_n (t, p) \;\mathrm{d}t
-}{\varepsilon\boldsymbol{\mu}(p)}
-\right)
-\label{eq:S_tilde}
-\end{equation}
-where $\varepsilon$ is a predefined constant.
-The offset of one before the application of the pointwise logarithm ensures that the transformation is nonexpansive in the sense of Lipschitzian maps.
-On a dataset of environmental audio recordings, a previous publication has shown empirically that Equation \ref{eq:S_tilde} brings the histogram of $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ closer to a Gaussian distribution \cite{lostanlen2018jasmp}.
-Since then, this finding has also been confirmed in the case of musical sounds \cite{lostanlen2018extended}.
-
-Lastly, we standardize every feature $\mathbf{\widetilde{S}}\boldsymbol{x}_n$ to null mean and unit variance, across the dataset $\mathcal{X} = \{\boldsymbol{x}_1 \ldots \boldsymbol{x}_N\}$, independently for each scattering path $p$.
-Again, if $\mathcal{X}$ is split between training and test sets, we measure means and variances over the training set only and propagate them as constants to the test set.
-With a slight abuse of notation, we still denote by $\mathbf{\widetilde{S}}\boldsymbol{x}_n (p)$ the standardized log-scattering features at path $p$ for sample $n$, even though its value differs from Equation \ref{eq:S_tilde} by an affine transformation.
-
-\subsection*{Metric learning with large-margin nearest neighbors (LMNN)}
-Let $\boldsymbol{x}$ some arbitrary audio sample in the dataset $\mathcal{X}$.
-Let $\mathcal{G}$ a cluster graph with $N = \card \mathcal{X}$ vertices and $C$ clusters in total.
-We denote by $\mathcal{G}(\boldsymbol{x})$ the cluster to which the sample $\boldsymbol{x}$ belongs.
-Given another sample $\boldsymbol{y}$ in $\mathcal{X}$, $\boldsymbol{y}$ is similar to $\boldsymbol{x}$ if and only if belongs to the cluster $\mathcal{G}(\boldsymbol{x})$.
-Because $\mathcal{G}$ is a disjoint union of complete graphs, this relation is symmetric: $\boldsymbol{x} \in \mathcal{G}(\boldsymbol{y})$ is equivalent to $\boldsymbol{y} \in \mathcal{G}(\boldsymbol{x})$.
-
-In our protocol, $\boldsymbol{x}$ contains the sound of an isolated musical note and the cluster graph $\mathcal{G}$ encodes auditory similarities within the dataset $\mathcal{X}$.
-Moreover, we take $\mathcal{G}$ to be the equal to the ``consensus'' cluster graph $\mathcal{G}_0$, i.e., arising from the partition of a hypergraph $\mathcal{H}$ which contains the judgments of all $K$ subjects from our perceptual study (see \lnameref{sec:data-collection} section).
-
-%Furthermore, let us denote by $\mathcal{Y}_1 (\boldsymbol{x}) = \mathcal{X} \backslash \lbrace \boldsymbol{x} \rbrace$ the dataset of $(N-1)$ audio samples obtained from $\mathcal{X}$ after subtracting the singleton $\{\boldsymbol{x}\}$.
-
-We denote by $\mathbf{\widetilde{S}}\boldsymbol{x}$ the feature vector of joint time--frequency scattering resulting from $\boldsymbol{x}$.
-This vector includes both first-order and second-order scattering coefficients, after median-based logarithmic compression and affine standardization (see subsections above).
-Furthermore, we denote by $\mathcal{Y}_R (\boldsymbol{x})$ the list of $R$ nearest neighbors to $\mathbf{\widetilde{S}}\boldsymbol{x}$ in the feature space of joint time--frequency scattering coefficients according to the Euclidean metric.
-Unlike cluster similarity, this relationship is not symmetric: $\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})$ does not necessarily imply $\boldsymbol{x}\in\mathcal{Y}_R (\boldsymbol{y})$.
-Note that the dependency of $\mathcal{Y}_R (\boldsymbol{x})$ upon the operator $\mathbf{S}$ is left implicit. % \ja wants to delete this sentence
-In all of the following, we set the constant $R$ to $5$; this is in accordance with our chosen evaluation metric, average precision at rank 5 (AP@5, see \lnameref{sec:results} section).
-
-Let $P$ be the number of scattering paths in the operator $\mathbf{S}$.
-The large-margin nearest neighbor (LMNN) algorithm learns a matrix $\mathbf{L}$ with $P$ rows and $P$ columns by minimizing an error function of the form:
-\begin{equation}
-\mathcal{E}(\mathbf{L}) = \frac{1}{2} \mathcal{E}_{\textrm{pull}} (\mathbf{L}) + \frac{1}{2} \mathcal{E}_{\textrm{push}} (\mathbf{L})
-\end{equation}
-where, intuitively, $\mathcal{E}_{\textrm{pull}}$ tends to shrink local Euclidean neighborhoods in feature space while $\mathcal{E}_{\textrm{push}}$ tends to penalize small distances between samples that belong to different clusters in $\mathcal{G}$.
-
-The definition of $\mathcal{E}_{\textrm{pull}}$ is:
-\begin{equation}
-\mathcal{E}_{\textrm{pull}} (\mathbf{L}) =
-\sum_{\boldsymbol{x}\in\mathcal{X}}
-\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
-\big\Vert
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-\big\Vert^2,
-\end{equation}
-Note that the error term $\mathcal{E}_\textrm{pull}$ is unsupervised, in the sense that it does not depend on the cluster assignments of $\boldsymbol{x}$ and $\boldsymbol{y}$ in $\mathcal{G}$.
-
-While the the term $\mathcal{E}_{\textrm{pull}}$ operates on pairs of samples, the term $\mathcal{E}_{\textrm{push}}$, operates on triplets $(\boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z})\in\mathcal{X}^3$.
-The first sample, $\boldsymbol{x}$, is known as an ``anchor.''
-The second sample, $\boldsymbol{y}$, is known as a ``positive'', and is assumed to belong to the Euclidean neighborhood of the anchor: $\boldsymbol{y} \in \mathcal{Y}_R (\boldsymbol{x})$.
-The third sample, $\boldsymbol{z}$, is known as a ``negative'', and is assumed to belong to a different similarity cluster as the anchor: $\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})$.
-The term $\mathcal{E}_{\textrm{push}}$ penalizes $\mathbf{L}$ unless the positive-to-anchor distance is smaller than the negative-to-anchor distance by a margin of at least $1$.
-
-The definition of $\mathcal{E}_{\textrm{push}}$ is:
-\begin{equation}
-\mathcal{E}_{\textrm{push}} (\mathbf{L}) =
-\sum_{\boldsymbol{x}\in\mathcal{X}}
-\sum_{\boldsymbol{y}\in\mathcal{Y}_R (\boldsymbol{x})}
-\sum_{\boldsymbol{z}\not\in\mathcal{G}(\boldsymbol{x})}
-\rho
-\Big(
-1 +
-\big\Vert
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-\big\Vert^2
--
-\big\Vert
-\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x} - \mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{z}
-\big\Vert^2
-\Big),
-\end{equation}
-where the function $\rho : u \mapsto \max (0, u)$ denotes the activation function of the rectified linear unit (ReLU), also known as hinge loss.
-The cost function described in the equation above is known in deep learning as ``triplet loss'' and has recently been applied to train large-vocabulary audio classifiers in an unsupervised way \cite{jansen2018icassp}.
-We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
-
-%\begin{equation}
-%\mathbf{\Phi}_1 (\boldsymbol{x}) =
-%\arg \min_{\boldsymbol{y}\in \mathcal{Y}_1(\boldsymbol{x})}
-%\;
-%\big\Vert
-%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
-%-
-%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-%\big\Vert
-%\end{equation}
-% reference triplet loss audioset
-
-%The standard Euclidean distance on the raw scattering coefficients provides an instrument-agnostic representation of timbre dissimilarity.
-%We now consider a transformation of this distance, in effect warping the timbre space, in order to more closely mimic a particular set of timbre similarity judgments.
-
-%Due to the size of corpus, we restrict ourselves to a simple linear transform of the scattering coefficients.
-More flexible and powerful methods could be brought to bear on this problem, such as deep neural networks, but with their flexibility comes an increase of the needed size of training data.
-
-%Specifically, we construct a positive semidefinite matrix $L$ that we apply to the scattering coefficients $x$ so that the Euclidean distance between $Lx$ and $Ly$ best approximates the consensus clustering obtained above.
-%We expect the resulting representation $Lx$ to provide a more accurate model for timbre perception.
-
-%We now use the consensus clustering to reweight the similarity between feature vectors (either MFCCs or scattering transforms).
-%For this, we use the \emph{large margin nearest neighbor} (LMNN) metric learning algorithm \cite{weinberger2006distance, weinberger2009distance}.
-%This approach constructs a positive semi-definite weighting matrix $L$ such that the distance $\|Lx - Ly\|$ best classifies a set of points into a fixed clustering assignment.
-%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
-
-%$\mathcal{Y}_{r} (\boldsymbol{x}) = \mathcal{Y}_{(r-1)}(\boldsymbol{x}) \backslash \lbrace \mathbf{\Phi}_r (\boldsymbol{x}) \rbrace$.
-
-%\begin{equation}
-%\mathbf{\Phi}_r (\boldsymbol{x}) =
-%\arg \min_{\boldsymbol{y}\in \mathcal{Y}_r(\boldsymbol{x})}
-%\;
-%\big\Vert
-%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{x}
-%-
-%\mathbf{L}\mathbf{\widetilde{S}}\boldsymbol{y}
-%\big\Vert
-%\end{equation}
-
-\section*{\nmu Results}
-\label{sec:results}
-
-The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
-We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
-However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
-
-\subsection*{Semi-supervised label propagation}
-
-In order to suit the practical needs of contemporary music composers, computer-assisted orchestration must draw from a diverse realm of instruments and techniques.
-Therefore, whereas our data collection procedure for timbre similarity judgments focused on a single pitch (middle C) and a single intensity level (\emph{mf}), we formulate our machine listening experiment on an expanded dataset of audio samples, containing variations in pitch and dynamics.
-
-Given an audio stimulus $\boldsymbol{x}_n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
-Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
-We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
-% \vl: ?
-
-\subsection*{Evaluation metric}
-
-Let us denote by $\boldsymbol{x}_1, \ldots, \boldsymbol{x}_{{N}^{\prime}}$ the $N^{\prime}$ audio samples associated with our annotated dataset after semi-supervised label propagation.
-Given a sample $n$ and a human subject $k$, we denote by $\mathcal{G}_k$ the cluster graph associated to the subject $k$, and by $\mathcal{G}_k (n)$ the cluster to which the sample $\boldsymbol{x}_{n}$ belongs.
-Our machine listening system takes the waveform $\boldsymbol{x}_{n}$ as input and returns a ranked list of nearest neighbors: $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_2 (\boldsymbol{x}_n)$, $\mathbf{\Phi}_3 (\boldsymbol{x}_n)$, and so forth.
-
-In the context of browsing an audio collection by timbre similarity, $\boldsymbol{x}_n$ is a user-defined query while the function $\mathbf{\Phi}$ plays the role of a search engine.
-We consider the first retrieved sample, $\mathbf{\Phi}_1 (\boldsymbol{x}_n)$, to be relevant to user $k$ if and only if it belongs to the same cluster as $\boldsymbol{x}_n$ in the cluster graph $\mathcal{G}_k$; hence the Boolean condition $\mathbf{\Phi}_{1}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
-Likewise, the second retrieved sample is relevant if and only if $\mathbf{\Phi}_{2}(\boldsymbol{x}_{n}) \in \mathcal{G}_k (n)$.
-To evaluate $\mathbf{\Phi}$ on the query $\boldsymbol{x}_n$, we measure the relevance of all nearest neighbors $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ up to some fixed rank $R$ and average the result:
-\begin{equation}
-p_{\mathbf{\Phi}}(n, k, R) =
-    \dfrac{1}{R}
-    \sum_{r=1}^{R}
-    \mathlarger{\mathds{1}}
-    \big(
-        \mathbf{\Phi}_r(\boldsymbol{x}_n)
-        \in
-        \mathcal{G}_k (n)
-    \big).
-\end{equation}
-In the equation above, the indicator function $\mathds{1}$ converts Booleans to integers, i.e., $\mathds{1}(b)$ returns one if $b$ is true and return zero if $b$ is false.
-Thus, the function $p_{\mathbf{\Phi}}$ takes fractional values between $0$ and $1$, which are typically expressed in percentage points.
-
-The precision at rank $R$ of the system $\mathbf{\Phi}$ is defined as the average value taken by the function $p_{\mathbf{\Phi}}$ over the entire corpus of $N^{\prime}$ audio samples:
-\begin{equation}
-\mathrm{P}_{\mathbf{\Phi}}(k, R) =
-\dfrac{1}{N^{\prime}}
-\sum_{n=1}^{N^{\prime}}
-p_{\mathbf{\Phi}}(n, k, R)
-\end{equation}
-
-Lastly, the ``average precision at rank $R$'' (henceforth, AP@$R$) is the average value of $\mathrm{P}_{\mathbf{\Phi}}$, for constant $R$, across all $K=31$ subjects from our perceptual study:
-\begin{equation}
-\mathrm{AP}_{\mathbf{\Phi}}(R) =
-\dfrac{1}{K}
-\sum_{k=1}^{K}
-\mathrm{P}_{\mathbf{\Phi}}(k, R)
-\end{equation}
-It appears from the above that an effective system $\mathbf{\Phi}$ should retrieve sounds whose IMT triplets are similar according to all of the $K$ cluster graphs $\mathcal{G}_1 \ldots \mathcal{G}_K$.
-
-In the rest of this paper, we set $R$ to $5$.
-This is in accordance with the protocol of \cite{lostanlen2018extended}, in which the authors trained a metric learning algorithm on the SOL dataset to search for similar instruments and playing techniques, yet without the intervention of a human subject.
-%Moreover, note that the value $R=5$ typically reflects the number of virtual panels simultaneously on display in a software environment for computer-assisted orchestration.
-
-\subsection*{Best performing system}
-Our best performing system comprises five computational blocks:
-
-\begin{enumerate}
-\item Joint time--frequency scattering up to a maximal time scale of $T=$ \SI{1000}{\milli\second},
-\item Temporal averaging at the scale of the whole musical note,
-\item Median-based logarithmic compression,
-\item Affine standardization so that each feature has zero mean and unit variance, and
-\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
-\end{enumerate}
-
-Note that the non-Euclidean metric is learned via LMNN (see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
-Therefore, the system $\mathbf{\Phi}$ performs timbre similarity retrieval in a user-agnostic way, and can serve as a convenient default algorithm for newcoming users.
-That being said, it is conceivable to replicate the five-stage protocol above on the cluster graph $\mathcal{G}_k$ of a specific user $k$, instead of the cluster graph $\mathcal{G}_0$.
-This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity.
-
-Composers may personalize their search engine $\mathbf{\Phi}$ by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
-This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
-
-Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ subjects of the order of $1\%$.
-This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
-In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bamboo flute \emph{(dizi)}.
-However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
-In particular, it does not assume the metadata (instrument, mute, technique, pitch, dynamics, and so forth) of any musical sample $\boldsymbol{x}_n$ to be observable, in part or in full, at training time.
-
-\section*{\nmu Discussion}
-\label{sec:discussion}
-
-The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
-Over a cohort of $K=31$ subjects, we reported an average precision at rank five (AP@5) of $99.0\% \pm 1$.
-We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
-
-Figure \ref{fig:ablation} summarizes our results.
-Interestingly, the system $\boldsymbol{\Phi}$ is not only best on average, but also best for every subject in the cohort.
-Specifically, replacing $\boldsymbol{\Phi}$ by a simpler model $\boldsymbol{\Phi}^{\prime}$ (see subsections below for examples of such models) results in $\mathrm{P}_{\mathbf{\Phi^\prime}}(k, 5) < \mathrm{P}_{\mathbf{\Phi}}(k, 5)$ for every $k$.
-Borrowing from the terminology of welfare economics, $\boldsymbol{\Phi}$ can be said to be uniquely Pareto-efficient \cite{black2012dictionary}.
-This observation suggests that the increase in performance afforded by the state-of-the-art model with respect to the baseline does not come at the detriment of user fairness.
-%Results of this ablation study is summarized in Figure , where performance achieved for each reference clustering is depicted by a losange whose color is chosen arbitrarily but consistently across conditions.
-It appears that the ranking over testing conditions is not strongly modified,  which demonstrate that for all conditions the reference clusterings which are hard to model (depicted by blue losanges) are hard to model for each conditions whereas the reference clusterings which are easier to model (depicted by yellow losanges) are easier to model for each conditions.
-
-\subsection*{Is the dataset large enough for evaluation? Is the system overfitting?}
-First of all, it is worth noting that the presented AP@5 figure of $99.0\% \pm 1$ does not abide to a conventional training set vs.
-test set paradigm, as is most often done in machine learning research.
-Rather, the LMNN algorithm is trained on all available samples ($N^{\prime}=9346$ isolated notes) with the ``consensus'' cluster graph as training objective ($\mathcal{G}=\mathcal{G}_0$).
-Then, it is evaluated on the same samples with individual cluster graphs as ground truth ($\mathcal{G}=\mathcal{G}_k$ for $k\geq1$).
-The reason behind this choice is that, in the context of computer-assisted orchestration, the collection of audio samples has a fixed size, as it is shipped alongside the software itself.
-This is the case, for example, of Orchidea,\footnote{Link to Orchidea software and OrchideaSOL dataset: \url{www.orch-idea.org}} which comes paired with its own version of SOL, named OrchideaSOL \cite{cella2020icmc}.
-Henceforth, our main goal was to evaluate the generalization ability of our metric learning algorithm beyond the restricted set of samples for which human annotations are directly available (see \lnameref{sec:data-collection} section); that is, beyond one pitch class (middle C) and one intensity level (\emph{mf}).
-
-Despite this caveat, we may adopt a ``query-by-example'' (QbE) paradigm by  partitioning the database and audio samples in half ($\frac{1}{2}N^{\prime}=4673$), training the LMNN algorithm on the first half, and querying it with samples from the other half.
-In this evaluation framework, our system reaches an AP@5 of $\textbf{96.2\%} \pm 2$.
-Interestingly, querying the system with samples from the training set leads to an AP@5 of $\textbf{96.5\%} \pm 2$, i.e., roughly on par with the test set.
-Thus, it appears that the gap in performance between the evaluation presented in the \lnameref{sec:results} section ($99.0\%\pm 1$) and query-by-example evaluation ($96.2\% \pm 2$) is primarily attributable to a reduction of the size of training set by half, whereas statistical overfitting of the training set with respect to the test set likely plays a minor role.
-
-%ML: as the learned layer is relatively thin
-
-\subsection*{Is metric learning necessary?}
-Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
-Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all. % \ja wants to elaborate
-This corresponds to using a Euclidean distance to compare scattering coefficients, i.e., to set $\mathbf{L}$ to the identity matrix.
-Note that the runtime complexity of Euclidean nearest-neighbor search is identical to LMNN search. % \ja wants to delete this sentence
-We report an average precision at rank five (AP@5) of $\textbf{92.9\%} \pm 3$, which is noticeably worse than the best performing system.
-This gap in performance demonstrates the importance of complementing unsupervised feature extraction by supervised metric learning in the design  of computational models for timbre similarity between instrumental playing techniques.
-
-\subsection*{How important is the amount of temporal context $T$?}
-
-Our best performing system operates with joint time--frequency scattering coefficients as spectrotemporal modulation features.
-These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
-This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
-Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$.
-
-As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
-This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.
-
-\subsection*{Is joint time--frequency scattering necessary?}
-
-Let us recall the full definition of second-order joint time--frequency scattering (see \lnameref{sec:methods} section):
-\begin{equation}
-\mathbf{S_2}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
-\left(
-\Big\vert
-\big\vert
-\boldsymbol{x} \ast \boldsymbol{\psi}_{\lambda}
-\big\vert
-\circledast
-\boldsymbol{\Psi}_{\alpha,\beta}
-\Big\vert
-\circledast
-\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_{F})
-\right)(t,\lambda),
-\label{eq:joint-scattering}
-\end{equation}
-where the $\boldsymbol{\psi}_{\lambda}$ denotes a Morlet wavelet of center frequency $\lambda$ and resp. $\boldsymbol{\phi}_T$ denotes a Gaussian low-pass filter of width $T$.
-Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
-\begin{equation}
-\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) =
-\left(
-\bigg\vert
-\Big\vert
-\big\vert
-\boldsymbol{x} \ast \boldsymbol{\psi}_{\lambda}
-\big\vert
-\ast
-\boldsymbol{\psi}_{\alpha}
-\Big\vert
-\circledast
-\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\psi}_{\beta}\big)
-\bigg\vert
-\circledast
-\big(\boldsymbol{\phi}_{T} \otimes \boldsymbol{\phi}_F\big)
-\right)(t, \lambda),
-\label{eq:separable-scattering}
-\end{equation}
-where the wavelet $\boldsymbol{\psi}_{\lambda}$ has a quality factor of $Q=12$ whereas the wavelets $\boldsymbol{\psi}_{\alpha}$ and $\boldsymbol{\psi}_{\beta}$ have a quality factor of one.
-Previous publications have successfully applied separable time--frequency scattering in order to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques from the Chinese bamboo flute \cite{wang2019ismir}.
-
-In comparison with its joint counterpart, separable time--frequency scattering contains about half as many coefficients.
-This is because the temporal wavelet transform with $\boldsymbol{\psi}_\alpha$ and the frequential wavelet transform with $\boldsymbol{\psi}_\beta$ are separated by an operation of complex modulus.
-Hence, $\boldsymbol{\psi}_\beta$ operates on a real-valued input.
-Because separable time--frequency scattering cannot distinguish ascending chirps from descending chirps, flipping the sign of the scale variable $\beta$ is no longer necessary.
-Moreover, separable time--frequency scattering has a lower algorithmic complexity than joint time--frequency scattering.
-Indeed, in Equation \ref{eq:separable-scattering}, the frequential wavelet transform with $\boldsymbol{\psi}_{\beta}$ operates on a tensor whose time axis is subsampled at a fixed rate $T^{-1}$, thus allowing vectorized computations.
-Conversely, in Equation \ref{eq:joint-scattering}, the frequential wavelet transform must operate on a multiresolution input, whose sample rate varies  depending on $\alpha$: it ranges between $T^{-1}$ and the sample rate of $\boldsymbol{x}$ itself.
-
-Yet, despite its greater simplicity, separable time--frequency scattering suffers from known weaknesses in its ability to represent spectrotemporal modulations.
-In particular, a previous publication has shown that frequency-dependent time shifts affect joint time--frequency scattering coefficients while leaving separable time--frequency scattering coefficients almost unchanged \cite{anden2019tsp}.
-%For this reason, separable time--frequency scattering coefficients can only serve to recognize periodic modulations (e.g.
-vibrato), but joint time--frequency scattering is required to recognize pitch evolution-based techniques (PETs, e.g.
-glissando) \cite{wang2020icassp}.
-The same observation was made by \cite{schadler2015jasa} in the case of joint and separable Gabor filterbank features (GBFB), which bear some resemblance with joint and separable time--frequency scattering coefficients respectively.
-
-Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
-This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
-This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
-
-\subsection*{What is the baseline performance?}
-Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
-Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis.\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}
-This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{81.8\%} \pm 7$.
-
-Arguably, this figure is not directly comparable with our best performing system (P@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
-In order to clarify the role of feature dimensionality in our computational pipeline, we apply a feature engineering technique involving multiplicative combinations of MFCC.
-We construct the following Gram matrix:
-\begin{equation}
-\mathbf{G} \boldsymbol{x}(\alpha, \beta) =
-\int_0^{+\infty}
-\mathrm{MFCC}(\boldsymbol{x})(t, \alpha)
-\mathrm{MFCC}(\boldsymbol{x})(t, \beta)
-\;\mathrm{d}t,
-\end{equation}
-where $\alpha$ and $\beta$ represent different dimensions (``quefrencies'') of the MFCC feature vector.
-The symmetric matrix $\mathbf{G}\boldsymbol{x}$ contains $40$ rows and $40$ columns, hence $800$ unique coefficients.
-Concatenating these coefficients to the $40$ averaged MFCC features results in a feature vector of $840$ coefficients.
-This dimension is of the same order of magnitude as the dimension of our joint time--frequency scattering representation ($d=1180$, see \lnameref{sec:methods} section).
-
-Training the LMNN algorithm on this 840-dimensional representation is analogous to a ``kernel trick'' in support vector machines \cite{chang2010jmlr}.
-In our case, the implicit similarity kernel is a homogeneous quadratic kernel.
-Despite this increase in representational power, we obtain a P@5 of $81.5\% \pm 7$, i.e. essentially the same as MFCC under a linear kernel.
-Therefore, it appears that the gap in performance between MFCC ($81.8\%\pm 7$) and joint time--frequency scattering ($99.0\% \pm 1$) is primarily attributable to the multiscale extraction of joint spectrotemporal modulations, whereas high-dimensional embedding likely plays a minor role.
-
-%In order to quantify the importance of the information carried by the feature set in the final performance, the numerical values of the features are replaced with random numbers sampled from a uniform distribution.
-%In this case, the performance drops to $23.6\% \pm 10.6$.
-%Replacing the learned linear mapping with the identity gives a $ap@5$ of $23.8\%$ for these features.
-%It can be concluded that the learning block is not able to perform effectively if no information is carried by the input features.
-%The slight decrease of performance when considering LMNN may be explained by the fact that the learning algorithm optimizes the projection using correlations in the training set that are not found in the testing set.
-
-\section*{\nmu Conclusion}
-\label{sec:conclusion}
-
-We see from the ablation study conducted above that each step of the proposed model is necessary for its satisfactory performance.
-Among other things, it indicates the joint time--frequency scattering transform itself should not be used directly for similarity comparisons, but is best augmented by a learned stage.
-This explains the relative success of such models \cite{lostanlen2018jasmp,lostanlen2018extended} over others where distances on raw scattering coefficients were used for judging audio similarity.
-
-On the other hand, we note that the complexity of the learned model does not have to be high.
-Indeed, for this task, a linear model on the scattering coefficients is sufficient.
-There is no need for deep networks with large numbers of parameters to accurately represent the similarity information.
-In other words, the scattering transform parametrizes the signal structure in a way that many relevant quantities (such as timbre similarity) can be extracted through a simple linear mapping.
-This is in line with other classification results, where linear support vector machines applied to scattering coefficients have achieved significant success \cite{anden2015mlsp,anden2019tsp}.
-
-We also see the necessity of a fully \emph{joint} time--frequency model for accessing timbre, as opposed to a purely spectral model or one that treats the two axes in a separable manner.
-This fact has also been observed in other contexts, such as the work of Patil et al. \cite{patil2012ploscompbiol}.
-A related observation is the need for capturing large-scale structure.
-Indeed, reducing the window size to $25~\mathrm{ms}$ means that we lose a great deal of time--frequency structure, bringing results closer to that of the separable model.
-
-The success of the above system in identifying timbral similarities has immediate applications in browsing music databases.
-These are typically organized based on instrumental and playing techniques taxonomies, with additional keywords offering a more flexible organization.
-Accessing the sounds in these databases therefore requires some knowledge of the taxonomy and keywords used.
-Furthermore, the user needs to have some idea of the particular playing technique they are searching for.
-
-As an alternative, content-based searches allow the user to identify sounds based on similarity to some given query sound.
-This query-by-example approach provides an opportunity to search for sounds without having a specific instrument of playing technique in mind, yielding a wider range of available sounds.
-A composer with access to such a system would therefore be able to draw on a more diverse palette of musical timbre.
-
-The computational model proposed in this work is well suited to such a query-by-example task.
-We have shown that it is able to adequately approximate the timbral judgments of a wide range of subjects included in our study.
-Not only that, but the system can be easily retrained to approximate an individual user's timbre perception by having that user perform the clustering task on the reduced set of $78$ IPTs and running the LMNN training step on those clustering assignments.
-This model can then be applied to new data, or, alternatively, be retrained with these new examples if the existing model proves unsatisfactory.
-
-We shall note, however, that the current model has several drawbacks.
-First, it is only applied to instrumental sounds.
-While this has the advantage of simplifying the interpretation of the results, the range of timbre under consideration is necessarily limited (although less restricted than only considering \emph{ordinario} PTs).
-This also makes applications such as \emph{query-by-humming} difficult, since we cannot guarantee that the timbral similarity measure is accurate for such sounds.
-
-That being said, the above model is general enough to encompass a wide variety of recordings, not just instrumental sounds.
-Indeed, we have strong assumptions that the tools used (scattering transforms and LMNN weighting matrices) do not depend strongly on the type of sound being processed.
-Future work will investigate whether more general classes of sounds are also well modeled.
-To extend the model, it is only necessary to retrain the LMNN weighting matrix by supplying it with new cluster assignments.
-These can again be obtained by performing a new clustering experiment with one or more human subjects.
-
-Another aspect is the granularity of the similarity judgments.
-In the above method, we have used \emph{hard} clustering assignments to build our model.
-A more nuanced similarity judgment would ask users to rate the similarity of a pair of IPTs on a more graduated scale, which would yield a finer, or \emph{soft}, assignment.
-This however, comes with additional difficulties in providing a consistent scale across subjects, but could be feasible if the goal is to only adapt the timbral space to a single individual.
-An approach not based on clustering would also have to replace the LMNN algorithm with one that accounts for such soft assignments.
-
-%ML: recaler ici la out of database query}
-%That being said, a potential future application of our work would be to use the pre-trained metric learning system $\mathbf{\Phi}$ to approximate out-of-sample targets \cite{carpentier2012cmj}.
-%One of such target could be a vocal sketch \cite{cartwright2015acm}, a sound event recorded from the environment, or a computer-generated waveform \cite{lemaitre2016plosone}.
-%However, evaluating the accuracy of a query-by-singing system for audio retrieval is incompatible with the clustering paradigm of the Cyberlioz annotation interface, in which all audio samples are assumed to be recorded ahead of human--computer interaction.
-%Therefore, we focus on modeling similarity in timbre across instruments and playing techniques, and leave the study of non-instrumental targets to future work.}
-
-% Move the part about complexity to the conclusion. Runtime complexity. How to scale this?
-%However, we should note that, from a computational standpoint, the nearest-neighbor search $\mathbf{\Phi}_r (\boldsymbol{x}_n)$ is considerably faster with LDA than with LMNN.
-%Indeed, while LMNN operates in the original dimensionality of joint time--frequency scattering features ($d=1180$), LDA projects these features onto a subspace whose dimensionality is equal to $d_0 = (C_0+1)$, where $C_0 = 19$ is the number of clusters in the consensus cluster graph $\mathcal{G}_0$.
-%Because the runtime complexity of nearest-neighbor search grows quadratically with the dimensionality, i.e., is $O(d^2)$, replacing LMNN by LDA accelerates the system by a factor $(d/d_0)^2 \approx 3 \cdot 10^3$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{\nmu Competing interests}
-  The authors declare that they have no competing interests.
-
-%\section*{\nmu Author's contributions}
-%    Text for this section \ldots
-
-\section*{\nmu Acknowledgment}
-This work is partially supported by the European Research Council (ERC) award 320959 (InvariantClass).
-This work is partially supported by the National Science Foundation (NSF) award 1633259 (BIRDVOX).
-The Flatiron Institute is a division of the Simons Foundation.
-We wish to thank Philippe Brandeis, \'{E}tienne Graindorge, St\'{e}phane Mallat, Adrien Mamou-Mani, and Yan Maresz for contributing to the TICEL research project.
-We also wish to thank the students of the Paris Conservatory and all anonymous participants to our study.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{bmc-mathphys}
-\bibliography{bib}
-
-%% Appendices
-
-% \section*{\nmu Dataset}
-% \label{sec:appendix-dataset}
-%
-%
-% The dataset considered in this paper as been recorded at Ircam and is composed of audio recording of $16$ different musical instruments played with different PTs, which leads to $235$ different IPT pairs.
-% In addition, the pitch and nuance is varie for each pair (when applicable), yielding a total of $25444$ audio recordings.
-% For each couple, the pitch and intonation is varied leading to 25444 audio samples.
-%
-% The recorded instruments belongs to the wind and string classes.
-% For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
-% For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
-% \ml{Better to display these in a table? VL todo table}
-%
-% Those instruments are recorded for different nuance and pitch if the latter is relevant, but more importantly, several playing technique are considered.
-% For the second experiment described in this paper, the playing techniques considered are: artificial harmonic (art), blow (blow), exploding slap (explo), harmonic (harm), key-click (key), col legno (legno), non vibrato (nonvib), ordinario (ord), pedal tone (pdl), pizzicato (pizz), ponticello (pont), slap (slap), (tasto), tongue-ram (tng), and xylophonic (xyl).
-%
-% \vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
-% The abbreviations can be removed. @data @writing}
-% \ml{}
-%
-% \vl{A bubble plot of number of notes per instrument (x axis) and playing technique
-% (y axis) would be extremely useful. @data @figure}
-% \ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}
-% \vl{Les donn\'{e} de l'\'{e}tude, relatives aux Figures 5 et 6.}
-
-%\section*{\nmu Human timbre judgments}
-%\label{sec:subjective}
-
-%The dataset considered in this study is taken
-%from the studio-on-line (SOL) music library  \cite{peeters2000instrument}.
-%It has audio recordings of 16 different musical instruments played
-%with different playing techniques,
-%which leads to $235$ different couples of instrument / playing technique.
-%For each couple, the pitch and intonation is varied leading to $25444$ recordings.
-%The way the dataset is structured is explained in more details in Appendix \ref{sec:appendix-dataset}.
-
-%Gathering the similarity judgments between each pair of samples would be a daunting task.
-Indeed, studies about perceptual similarity are hindered with a dimensional problem as the number of pairs to be evaluated grows with the square of the number of items considered.
-
-%One reasonable assumption that we choose to consider here is that the pitch and nuance do not alter significantly the perception of timbral similarity.
-This is in line with the ANSI negative definition of timbre \vl{expliquer, citer, et modérer le propos en donnant des exemples dans lequel le pitch ou la nuance change le timbre}.
-
-%\section*{\nmu Reducing Expert judgments of technique peculiarities}
-
-%These IPTs were selected from a larger set of $235$ IPTs by two experts at the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP).
-%The selection criteria were based on the likelihood of an IPT being similar to some other IPT, ensuring that IPTs that were too singular were excluded from the survey.
-
-%Even so, considering every pairs of IPTs is not practical. A selection
-%experiment is thus conducted where the subject is asked to give his opinion on which
-%\ipts are relevant to study. More precisely, an \ipt is said to be relevant to study
-%if it is likely to be associated to another \ipt of another instrument.
-%Interest is rated on a 7 ticks scale. The subject can listen to all the different
-%nuance and pitch samples of each \ipt. The experiment is over when all the
-%\ipts are rated.
-
-%Rating guidelines are provided :
-%\begin{itemize}
-%  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt of another instrument.
-%  \item Four stars: there is a proximity on one aspect of the sound between this
-%  \ipt and of an \ipt of another instrument,
-%  but this one is neither decisive nor obvious.
-%  \item Seven stars: there is a large similarity between this \ipt and an \ipt of another instrument.
-%\end{itemize}
-
-%Two music composition professors of the CNSMDP performed the test, that is rating each $235$ \ipt.
-
-\begin{table}
-\caption{Full list of audio stimuli (1/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
-\label{tab:list1}
-      \begin{tabular}{llll}
-          name & instrument  & mute   &  playing technique \\ \hline \hline
-        \texttt{ASax-key-cl-C4-p} & Alto saxophone & & key click \\
-        \texttt{ASax-ord-C4-mf} & Alto saxophone &  &  \\
-        \texttt{ASax-ord-hi-reg-C6-mf} & Alto saxophone &   & \\
-        \texttt{ASax-slap-C4-mf} & Alto saxophone & & slap tonguing \\
-        \texttt{ASax-slap-unp-C4-p} & Alto saxophone & & unpitched slap tonguing \\ \hline
-        \texttt{BbCl-key-cl-C4-pp} & Soprano clarinet & & key click \\
-        \texttt{BbCl-ord-hi-reg-A6-ff} & Soprano clarinet & & \\ \hline
-        \texttt{BBTb-explo-slap-C\#1-mf} & Bass tuba &  & explosive slap \\
-		\texttt{BBTb-pdl-tone-F1-mf} & Bass tuba & & pedal tone \\
-		\texttt{BBTb-slap-F1-mf} & Bass tuba & & slap tonguing \\
-		\texttt{BBTb-slap-unp-mf-1} & Bass tuba & & unpitched slap tonguing \\ \hline
-		\texttt{Bn-key-cl-C3-mf} & Bassoon & & key click \\
-		\texttt{Bn-ord-C4-mf} & Bassoon & & \\ \hline
-		\texttt{Cb-pizz-bartok-C4-ff-1c} & Contrabass & & pizzicato \`a la Bart\'ok (snap) \\
-		\texttt{Cb-pizz-lv-C4-mf-1c} & Contrabass & & pizzicato \& \emph{laissez vibrer} \\
-        \texttt{Cb-pizz-sec-C4-mf-1c} & Contrabass & & pizzicato secco \\
-        \texttt{Cb-pont-C4-mf-1c} & Contrabass & & sul ponticello \\ \hline
-        \texttt{Fl-key-cl-C4-f} & Flute & & key click \\
-        \texttt{Fl-ord-C4-mf} & Flute & & \\
-        \texttt{Fl-tng-ram-C4-mf} & Flute & & tongue ram \\ \hline
-        \texttt{Gtr-ord-C4-mf-2c} & Spanish guitar & & \\
-        \texttt{Gtr-ord-hi-reg-E5-mf-3c} & Spanish guitar & & \\
-        \texttt{Gtr-pizz-bartok-C4-ff-2c} & Spanish guitar & & pizzicato Bart\'ok\\
-        \texttt{Gtr-pizz-C4-mf-2c} & Spanish guitar &  & pizzicato \\ \hline
-        \texttt{Hn-ord-C4-mf} & French horn & & \\
-        \texttt{Hn-slap-C4-mf} & French horn & & slap tonguing \\ \hline
-        \texttt{Hp-harm-fngr-C4-f} & Harp & & harmonic fingering \\
-        \texttt{Hp-ord-C4-m4} & Harp & & \\
-        \texttt{Hp-pizz-bartok-C4-mf} & Harp & & pizzicato Bart\'ok \\
-        \texttt{Hp-xyl-C4-p} & Harp & & xylophonic \\ \hline
-        \texttt{Ob-blow-no-reed-C4} & Oboe & & blow without reed \\
-        \texttt{Ob-key-cl-C4-pp} & Oboe & & key click \\
-        \texttt{Ob-ord-C4-mf} & Oboe & & \\ \hline
-        \texttt{TpC-ord-C4-mf} & Trumpet in C & & \\
-        \texttt{TpC-pdl-tone-F3-mf} & Trumpet in C & & pedal tone \\
-        \texttt{TpC-slap-C4-p} & Trumpet in C & & slap tonguing \\
-        \texttt{TpC+C-ord-C4-mf} & Trumpet in C & cup & \\
-        \texttt{TpC+H-ord-C4-mf} & Trumpet in C & harmon & \\
-        \texttt{TpC+S-ord-C4-mf} & Trumpet in C & straight & \\
-        \texttt{TpC+W-ord-closed-C4-mf} & Trumpet in C & wah (closed) & \\
-        \texttt{TpC+W-ord-open-C4-mf} & Trumpet in C & wah (open) & \\ \hline
-        \texttt{TTbn-ord-C4-mf} & Tenor trombone & &  \\
-        \texttt{TTbn+C-ord-C4-mf} & Tenor trombone & cup & \\
-        \texttt{TTbn+H-ord-C4-mf} & Tenor trombone & harmon & \\
-        \texttt{TTbn+S-ord-C4-mf} & Tenor trombone & straight & \\
-        \texttt{TTbn+W-ord-closed-C4-mf} & Tenor trombone & wah (closed) & \\
-        \texttt{TTbn+W-ord-open-C4-mf} & Tenor trombone & wah (open) & \\ \hline
-        \texttt{Va-art-harm-C5-mf-4c} & Viola & & artificial harmonic \\
-        \texttt{Va-legno-batt-C4-mf-3c} & Viola & & col legno battuto \\
-        \texttt{Va-ord-C4-mf-3c} & Viola & & \\
-        \texttt{Va-pizz-bartok-C4-ff-3c} & Viola & & pizzicato Bart\'ok \\
-        \texttt{Va-pizz-lv-C4-mf-3c} & Viola & & pizzicato \& \emph{laissez vibrer} \\
-        \texttt{Va-pizz-sec-C4-mf-3c} & Viola & & pizzicato secco \\
-        \texttt{Va-pont-C4-mf-3c} & Viola & & sul ponticello \\
-        \texttt{Va+S-ord-C3-mf-3c} & Viola & sordina & \\
-        \texttt{Va+SP-ord-D4-mf-2c} & Viola & piombo & \\ \hline
-        \texttt{Vc-art-harm-C4-mf} & Cello & & artificial harmonic \\
-        \texttt{Vc-legno-batt-C4-mf-1c} & Cello & & col legno battuto \\
-        \texttt{Vc-legno-tratto-C4-mf-1c} & Cello & & col legno tratto \\
-        \texttt{Vc-nonvib-C4-mf-1c} & Cello & & nonvibrato \\
-        \texttt{Vc-ord-C4-mf-1c} & Cello & & \\
-        \texttt{Vc-pizz-bartok-C4-ff-1c} & Cello & & pizzicato Bart\'ok \\
-        \texttt{Vc-pizz-lv-C4-mf-1c} & Cello & & pizzicato \& \emph{laissez vibrer} \\
-        \texttt{Vc-pizz-sec-C4-mf-1c} & Cello & & pizzicato secco \\
-        \texttt{Vc-pont-C4-mf-2c} & Cello & & sul ponticello \\
-        \texttt{Vc-tasto-C4-mf-1c} & Cello & & sul tasto \\
-        \texttt{Vc+S-ord-C4-mf-1c} & Cello & sordina & \\
-        \texttt{Vc+SP-ord-C4-mf-1c} & Cello & piombo & \\ \hline
-    \end{tabular}
-\end{table}
-
-\begin{table*}
-\caption{Full list of audio stimuli (2/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
-\label{tab:list2}
-\begin{tabular}{llll}
-          name & instrument  & mute   &  playing technique \\ \hline \hline
-          \texttt{Vn-art-harm-G5-mf-4c} & Violin & & artificial harmonic \\
-          \texttt{Vn-legno-batt-C4-mf-4c} & Violin & & col legno battuto \\
-          \texttt{Vn-nonvib-C4-mf-4c} & Violin & & nonvibrato \\
-          \texttt{Vn-ord-C4-mf-4c} & Violin & & \\
-          \texttt{Vn-pizz-bartok-C4-ff-4c} & Violin & & pizzicato Bart\'ok \\
-          \texttt{Vn-pizz-lv-C4-mf-4c} & Violin & pizzicato \& \emph{laissez vibrer} \\
-          \texttt{Vn-pizz-sec-C4-mf-4c} & Violin & pizzicato secco \\
-          \texttt{Vn-pont-C4-mf-4c} & Violin & & sul ponticello \\
-          \texttt{Vn+S-ord-C4-mf-4c} & Violin & sordina & \\
-          \texttt{Vn+SP-ord-C4-mf-4c} & Violin & piombo & \\ \hline
-\end{tabular}
-\end{table*}
-
-%Even if the hypothesis that the distributions of ratings comes from independent random samples from normal distributions with equal means and equal but unknown variances, is verified using the two-sample t-test, inspection of the respective histograms reveals that averaging the ratings of the two raters is reasonable.
-
-%Keeping in mind that the gathering of the perceptual judgments of \ipt similarities shall remain practical, the number of \ipts considered shall be lower than $100$.
-We thus select the \ipts with an average rating higher than three, which brings us 78 \ipts to study which are listed in Tables \ref{tab:list1} and \ref{tab:list2}.
-
-%%%%%%%%%%%%%% CROWDSOURCED JUDGMENTS OF PAIRWISE SIMILARITY %%%%%%%%%%%%%%%%%%%%
-%\section*{\nmu Crowdsourced judgments of pairwise similarity}
-
-%\ja{A lot of this is covered in the main text.
-%Not sure what we need to add in the appendix.}
-
-%The second experiment aims at gathering spontaneous judgment of similarity for those 78 \ipts using a canonical free sorting task.
-This type of task is chosen among others for two main reasons.
-First, it gives the ability to the subject to listen to the whole set of \ipts before performing the task.
-Second, it gives more freedom to the subject, thus potentially more interest to perform the task than a lengthy series of XXY dissimilarity ratings.
-
-%31 subjects completed the experiment.
-An example of resulting organization of \ipt is given on Figure \ref{fig:xp2display}.
-\footnote{The resulting organization for each subject are available for reader's inspection using the full feature interface provided to the subjects for performing the experiment \url{https://mathieulagrange.github.io/paperSpontaneousSimilarity/demo}}.
-
-%Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
-The spatial organization of the dots representing the \ipt could provide information about the similarity, but this would implicitly enforce the fact that the timbral similarity space is two dimensional.
-We therefore now only study the properties of the clusterings performed by the subject prior to converting them into an overall similarity matrix.
-
-%This results in $31$ clusterings that can be considered for defining a perceptual timbre space of \ipts.
-Among the 20 colors available, the subjects used on average $10.2 \pm  4.1$ different colors to group \ipts.
-The distribution of the number of groups can be seen on Figure \ref{fig:xp2nbSize}(a).
-The size of the groups is on average $7.7 \pm   7.2$.
-The distribution of the size of groups can be seen on Figure \ref{fig:xp2nbSize}(b).
-The large number of small groups indicates that many subjects considered that a few of the \ipt were very different from all the remaining \ipt.
-
-%\begin{figure}
-%\center
-%\begin{tabular}{cc}
-%  (a) & (b) \\
-%  \includegraphics[width = \textwidth]{nbc.png} &
-%  \includegraphics[width = \textwidth]{sbc.png}
-%\end{tabular}
-%\caption{Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects.}
-%\label{fig:xp2nbSize}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Figures}
-
-\begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{lostanlen2020jasmp_fig1_v4.pdf}
-\caption{\csentence{Overview of the proposed approach.} See \lnameref{sec:intro} for details.}
-\label{fig:pipeline}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = 0.85\textwidth]{consensusVsI.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{I}}$ of perceived timbre similarities between instruments from our dataset of $N=78$. samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:instrument-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ were re-arranged according to Ward's minimum variance method.
-Observe that the block diagonal structure in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ reflects an organological taxonomy of musical instruments: woodwinds, brass, and strings.
-See \lnameref{sec:data-collection} section for details, including the meaning of each instrument abbreviation.
-}
-\label{fig:consensusVsI}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = 0.85\textwidth]{consensusVsPt.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{T}}$ of perceived timbre similarities between playing techniques from our dataset of $N=78$ samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:technique-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{T}}$ were re-arranged according to Ward's minimum variance method.
-See \lnameref{sec:data-collection} section for details,.
-See Tables \ref{tab:list1} and \ref{tab:list2} for an English description of each playing technique abbreviation.
-%TODO add a double arrow?
-% rough <-- soft --> rough
-% sustained <-- "ordinary" --> percussive
-}
-\label{fig:consensusVsPt}
-\end{figure}
-
-\begin{figure}
-\includegraphics[height=4cm]{Vn-ord-C4-mf-4c_scalogram.pdf}
-\hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-nonvib-C4-mf-4c_scalogram.pdf}
-\hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-trem-C4-mf-4c_scalogram.pdf}
-
-\hspace{2mm}
-\includegraphics[height=28mm]{Vn-ord-C4-mf-4c_scattering.pdf}
-\hspace{5mm}
-\includegraphics[height=28mm]{Vn-nonvib-C4-mf-4c_scattering.pdf}
-\hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{Vn-trem-C4-mf-4c_scattering.pdf}
-
-\csentence{(a) ordinario (with vibrato)}\hspace{18mm}
-\csentence{(b) nonvibrato}\hspace{24mm}
-\csentence{(c) tremolo}\hspace{8mm}
-\vspace{1cm}
-
-\includegraphics[height=4cm]{Vn-sfz-C4-fp-4c_scalogram.pdf}
-\hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-pizz-lv-C4-pp-4c_scalogram.pdf}
-\hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Vn-stacc-C4-mf_scalogram.pdf}
-
-\hspace{2mm}
-\includegraphics[height=28mm]{Vn-sfz-C4-fp-4c_scattering.pdf}
-\hspace{5mm}
-\includegraphics[height=28mm]{Vn-pizz-lv-C4-pp-4c_scattering.pdf}
-\hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{Vn-stacc-C4-mf_scattering.pdf}
-
-\hspace{2mm}
-\csentence{(d) sforzando}\hspace{25mm}
-\csentence{(e) pizzicato}\hspace{25mm}
-\csentence{(f) staccato}\hspace{1mm}
-
-\caption{\csentence{Six playing techniques of the violin.}
-Subfigures:
-(a) ordinario (with vibrato), (b) nonvibrato, (c) tremolo, (d) sforzando, (e) pizzicato (\emph{laissez vibrer}, i.e., let ring), and (f) staccato.
-In each subfigure, the top image shows the wavelet scalogram as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
-Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
-Darker shades denote greater values of acoustic energy.
-See \lnameref{sec:methods} section for details.}
-\label{fig:violin-scattering}
-\end{figure}
-
-\begin{figure}
-\includegraphics[height=4cm]{Fl-ord-C4-mf_scalogram.pdf}
-\hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-trill-maj2-C4-mf_scalogram.pdf}
-\hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-ply+sng-Csharp4+C4-mf_scalogram.pdf}
-
-\hspace{2mm}
-\includegraphics[height=28mm]{Fl-ord-C4-mf_scattering.pdf}
-\hspace{5mm}
-\includegraphics[height=28mm]{Fl-trill-maj2-C4-mf_scattering.pdf}
-\hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{Fl-ply+sng-Csharp4+C4-mf_scattering.pdf}
-
-\hspace{12mm}
-\csentence{(a) ordinario}\hspace{22mm}
-\csentence{(b) trill $\mathsf{C_4} - \mathsf{D_4}$}\hspace{15mm}
-\csentence{(c) play $\mathsf{C\musSharp_4}$ while singing $\mathsf{C_4}$}\vspace{1cm}
-
-\includegraphics[height=4cm]{Fl-sfz-C4-fp_scalogram.pdf}
-\hspace{8mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-key-cl-C4-f_scalogram.pdf}
-\hspace{12mm}
-\includegraphics[trim=35 0 0 0, clip, height=4cm]{Fl-stacc-C4-mf_scalogram.pdf}
-
-\hspace{2mm}
-\includegraphics[height=28mm]{Fl-sfz-C4-fp_scattering.pdf}
-\hspace{5mm}
-\includegraphics[height=28mm]{Fl-key-cl-C4-f_scattering.pdf}
-\hspace{5mm}
-\includegraphics[trim=0 0 77 0, clip, height=28mm]{Fl-stacc-C4-mf_scattering.pdf}
-
-\hspace{2mm}
-\csentence{(d) sforzando}\hspace{25mm}
-\csentence{(e) key click}\hspace{25mm}
-\csentence{(f) staccato}\hspace{1mm}
-
-\caption{\csentence{Six playing techniques of the flute.}
-Subfigures:
-(a) ordinario, (b) trill ($\mathsf{C_4} - \mathsf{D_4}$), (c) interference (play $\mathsf{C\musSharp_4}$ while singing $\mathsf{C_4}$), (d) sforzando, (e) key click, (f) staccato.
-In each subfigure, the top image shows the wavelet scalogram as a function of time $t$ (in seconds) and frequency $\lambda$ (in Hertz).
-Conversely, the bottom image shows the average time--frequency scattering coefficients, as a function of temporal modulation rate $\alpha$ (in Hertz) and frequential modulation scale $\beta$ (in cycles per octave).
-Darker shades denote greater values of acoustic energy.
-See \lnameref{sec:methods} section for details.}
-\label{fig:flute-scattering}
-\end{figure}
-
-\begin{figure}[h!]
-\includegraphics[width=0.9\textwidth]{ablationStudy}
-\caption{\csentence{Impact of different processing architecture or protocol designs.} For each condition, the central mark indicates the median, and the bottom and top edges of the box indicate the 25th and 75th percentiles, respectively.
-The performance achieved for each reference clustering is depicted by a lozenge whose color is chosen arbitrarily but consistently across conditions.
-See \lnameref{sec:discussion} section for details.}
-\label{fig:ablation}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \section*{Tables}
-% \begin{table}[h!]
-% \caption{Sample table title. This is where the description of the table should go.}
-%       \begin{tabular}{cccc}
-%         \hline
-%            & B1  &B2   & B3\\ \hline
-%         A1 & 0.1 & 0.2 & 0.3\\
-%         A2 & ... & ..  & .\\
-%         A3 & ..  & .   & .\\ \hline
-%       \end{tabular}
-% \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section*{Additional Files}
-%  \subsection*{Additional file 1 --- Sample additional file title}
-%    Additional file descriptions text (including details of how to
-%    view the file, if it is in a non-standard format or the file extension).  This might
-%    refer to a multi-page table or a figure.
-
-%  \subsection*{Additional file 2 --- Sample additional file title}
-%    Additional file descriptions text.
-
-\end{backmatter}
-
 \end{document}

### 2020-11-04 paper/paper.tex
capitalization
create future/paper.tex
Minor modifications

@@
 
-Out paper strives to fill the gap in scholarship between MIR and music cognition in the context of extended playing techniques.
+Our paper strives to fill the gap in scholarship between MIR and music cognition in the context of extended playing techniques.
 From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
@@
 
-In comparison with Web-based forms, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
+In comparison with web-based forms, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
 Another advantage of Cyberlioz is that it allows to present all stimuli at once rather than according to a randomized sequence.
@@
 For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
-Figure \ref{fig:cluster-histograms} (a) shows the histogram of $C_k$ across the cohort of $K=31$ participants.
+Figure \ref{fig:cluster-histograms}(a) shows the histogram of $C_k$ across the cohort of $K=31$ participants.
 We observe that the number of clusters varies between $3$ and $19$ with a median value of $10$.
-Accordingly, the number of samples belonging to a cluster varies between $1$ (the most frequent value) and $50$, as shown in Figure \ref{fig:cluster-histograms} (b).
+Accordingly, the number of samples belonging to a cluster varies between $1$ (the most frequent value) and $50$, as shown in Figure \ref{fig:cluster-histograms}(b).
 %Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
@@
 This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
+
 \section*{\nmu Machine listening methods}
@@
 Previous studies on real-world musical sounds have confirmed that listeners are able to ignore salient pitch differences while rating timbre similarity \cite{handel2001musicperception,marozeau2003dependency}, insofar as these differences do not exceed one octave.
-Since then, another study has shown that trained musicians can identify the similarity within pairs of notes from two instruments (horn and bassoon) spanning a range of $2.5$ octave with an accuracy of $80\%$ \cite{steele2006musicperception}.
+Since then, another study has shown that trained musicians can identify the similarity within pairs of notes from two instruments (horn and bassoon) spanning a range of $2.5$ octaves with an accuracy of $80\%$ \cite{steele2006musicperception}.
 In comparison, nonmusicians has great difficulty identifying whether two notes were produced by the same instrument or not when the notes were separated by one octave.
@@
 
-At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
+In the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
 From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
@@
 As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
+In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is known as the ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
 
@@
 By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
-For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
+For example, Figure \ref{fig:violin-scattering}(a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
 A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
-As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
-Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
+As seen in Figure \ref{fig:violin-scattering}(b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
+Furthermore, Figure \ref{fig:violin-scattering}(c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
 
 The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
-Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
+Figures \ref{fig:violin-scattering}(d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
 These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
@@
 Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
-Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
-Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
-We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
+Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering}(b)) or a beating tone (Figure \ref{fig:flute-scattering}(c)), cause local energy maxima whose rate $\alpha$ is physically interpretable.
+Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering}(d)), key click (Figure \ref{fig:flute-scattering}(e)), and vibrato (Figure \ref{fig:flute-scattering}(f)) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
+We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license.\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}
 
@@
 These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
-This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
+This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCCs).
 Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$ \cite{brown1999jasa,eronen2000icassp}.
@@
 
-\subsection*{Comparison with mel--frequency cepstral coefficients (MFCC) baseline}
+\subsection*{Comparison with mel--frequency cepstral coefficient (MFCC) baseline}
 Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by MFCCs.
@@
   \item STRF: spectrotemporal receptive fields
-  \item SOL: studio On Line dataset
+  \item SOL: Studio On Line dataset
 \end{itemize}

### 2020-11-04 paper/paper.tex
Spellcheck

@@
 
-The processing code, the cyberlioz interface and an anonimized version of the  perceptual judgments gathered using this interface are available at the companion website: \url{https://github.com/mathieulagrange/lostanlen2020jasmp}.
+The processing code, the Cyberlioz interface and an anonymized version of the  perceptual judgments gathered using this interface are available at the companion website: \url{https://github.com/mathieulagrange/lostanlen2020jasmp}.
 

### 2020-11-04 paper/paper.tex
celloVc -> cello
nonmusicians -> non-musicians

@@
 Since then, another study has shown that trained musicians can identify the similarity within pairs of notes from two instruments (horn and bassoon) spanning a range of $2.5$ octaves with an accuracy of $80\%$ \cite{steele2006musicperception}.
-In comparison, nonmusicians has great difficulty identifying whether two notes were produced by the same instrument or not when the notes were separated by one octave.
+In comparison, non-musicians has great difficulty identifying whether two notes were produced by the same instrument or not when the notes were separated by one octave.
 Thus, the influence of pitch on timbre perception appears to depend upon musical training.
@@
   \item Va: viola,
-  \item Vc: celloVc,
+  \item Vc: cello,
   \item Cb: contrabass,

### 2020-11-04 paper/paper.tex
ready to second submission

@@
 In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
-To this end, we publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
+The subjects were asked to ‘cluster sounds into groups by assigning the same color to the corresponding dots according to how similar the sounds are’.
+
+We publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
 For more information about these mailing lists, please visit:
@@
 
-We did not collect personal information on the age, sex, nor musical background of participants.
-This is because the goal of our perceptual study is to build a consensus of similarity judgments, rather than to compare demographic subgroups.
-We note that the relatively small size of our cohort ($K=31$) lacks the statistical power to carry out such a comparison.
+Personal information on the age, sex, nor musical background of participants is not collected, because the goal of our perceptual study is to build a consensus of similarity judgments, rather than to compare demographic subgroups.
+% We note that the relatively small size of our cohort ($K=31$) lacks the statistical power to carry out such a comparison.
+
 In particular, we leave the important question of the effect of musical training on the perception of auditory similarities between playing techniques as future work.
@@
 \subsection*{Hypergraph partitioning}
+
 Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a cluster graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
@@
 The performance achieved for each reference clustering is depicted by a lozenge whose color is chosen arbitrarily but consistently across conditions.
-See \lnameref{sec:discussion} section for details.}
+See \lnameref{sec:results} section for details.}
 \label{fig:ablation}

### 2020-11-04 paper/revision/paper.tex
self content tex file

--- 2020-07-21 paper/submission/paper.tex
@@
 
-\graphicspath{{figures/}{figures/violin_scattering}{figures/flute_scattering}{./}}
+\graphicspath{{figures/}{figures/violin_scattering/}{figures/flute_scattering/}{./}}
 %%% Put your definitions there:
@@
 \author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   addressref={aff2},                   % id's of addresses, e.g. {aff1,aff2}
                         % id of corresponding address, if any
 %   noteref={n1},                        % id's of article notes, if any
-   email={vincent.lostanlen@nyu.edu}            % email address
+   email={vincent.lostanlen@ls2n.fr}            % email address
 ]{\inits{VL}\fnm{Vincent} \snm{Lostanlen}}
@@
    addressref={kth,ccm},
-   email={janden@flatironinstitute.org}
+   email={janden@kth.se}
 ]{\inits{JA}\fnm{Joakim} \snm{And\'en}}
@@
 On the other hand, the sole mention of a playing technique does not specify its effect in terms of auditory perception.
-For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
-Yet, such acoustical similarity at the scale of the entire instrumentarium is not directly reflected by the semantic similarity between playing technique denominations; and discovering it requires musical expertise \cite{kolozali2011ismir}.
+For instance, although the term \emph{breathy} alludes to a playing technique that is specific to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.
+Yet, in a diverse instrumentarium, the semantic similarity between playing technique denominations does not reflect such acoustical similarity \cite{kolozali2011ismir}.
 
@@
 The outcome is another audio signal which is selected from a database of instrumental samples and perceptually similar to the query.
+
 The advantage of this search is that, unlike the query, the retrieved sound is precisely encoded in terms of duration, pitch, intensity, instrument, and playing technique.
@@
 
+Composers may personalize their search engine by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
+This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in the database: i.e., roughly three hours of audio.
+
 \subsection*{Goal}
-This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.
+This article proposes a machine listening system which computes the dissimilarity in timbre between two audio samples.\footnote{
+For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
 Crucially, this dissimilarity is not evaluated in terms of acoustic tags, but in terms of \emph{ad hoc} clusters, as defined by a human consensus of auditory judgments.
@@
 
-\subsection*{Outline}
-The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
-The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
-The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
-The \lnameref{sec:results} section describes our experimental benchmark.
-The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
-The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.\footnote{For the sake of research reproducibility, the source code for the experimental protocol of this paper is available online, alongside anonymized data from human subjects:  \url{https://github.com/mathieulagrange/lostanlen2020jasmp}}
+Our paper strives to fill the gap in scholarship between MIR and music cognition in the context of extended playing techniques.
+From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
+Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is readily tailored to subjective preferences.
+
+%\subsection*{Outline}
+%The \lnameref{sec:related-work} section reviews recent research in the computational modeling of musical timbre.
+%The \lnameref{sec:data-collection} section describes our computer--human interface for annotating auditory similarities across playing techniques.
+%The \lnameref{sec:methods} section presents the technical components of our machine listening system, namely joint time--frequency scattering and large-margin nearest neighbors.
+%The \lnameref{sec:results} section describes our experimental benchmark.
+%The \lnameref{sec:discussion} section conducts an ablation study of our best performing system.
+%The \lnameref{sec:conclusion} summarizes the implications of our findings for future research in music information retrieval.
 
@@
 
-\subsection*{Timbre modeling in music cognition research}
+\subsection*{Spectrotemporal receptive fields (STRF) in music cognition}
 On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
-In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
-Rather, each excerpt must be played back to multiple human listeners.
+In this setting, prior knowledge of instrumentation, however accurate, does not suffice to conduct a study on timbre perception: rather, the timbre perception relies on an interplay of acoustic and categorical information \cite{siedenburg2016frontiers}.
 Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, which restricts the size of the musical corpus under study.
-These small corpus sizes hampers the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
+These small corpus sizes hamper the applicability of optimization algorithms for representation learning, such as stochastic gradient descent in deep neural networks.
 
@@
 In addition, previous publications on joint time--frequency scattering lack a human-centric evaluation, independently from any classification task.
-
-\subsection*{Claim of originality}
-The contributions of this paper strive to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
-From the standpoint of MIR, the model presented here offers an efficient and generic multidimensional representation for timbre similarity, alongside theoretical guarantees of robustness to elastic deformations in the time--frequency domain.
-Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, which is also readily tailored to subjective preferences.
-
-\section*{\nmu Data collection}
+Beyond the case of STRF, we refer to \cite{caetano2019chapter} for a detailed review of the state of the art on audio descriptors of timbre.
+
+\section*{\nmu Perceptual data collection}
 \label{sec:data-collection}
 
-The European symphonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
+The philharmonic orchestra encompasses four families of instruments: strings, woodwinds, brass, and percussion.
 In this article, we focus on the first three, and leave the question of learning auditory similarities between percussion instruments to future research.
@@
 
-In comparison with a traditional web-based form, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
+Cyberlioz implements a data collection procedure known as ``free sorting''.
+In comparison with the direct collection of timbre dissimilarity ratings, free sorting is more efficient yet less accurate \cite{giordano2011multivariate}.
+We refer to \cite{elliott2013jasa} for an example protocol in which timbre similarity judgments rely on stimuli pairs rather than on a free sorting task.
+
+In comparison with web-based forms, Cyberlioz offers a more intuitive and playful user experience, while limiting the acquisition of similarity judgments to a moderate duration of $30$ to $60$ minutes for each participant.
+Another advantage of Cyberlioz is that it allows to present all stimuli at once rather than according to a randomized sequence.
+
 In May and June 2016, we recruited volunteers to use Cyberlioz on their own computers, via a web browser, and equipped with a pair of earphones.
-To this end, we publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
+The subjects were asked to ‘cluster sounds into groups by assigning the same color to the corresponding dots according to how similar the sounds are’.
+
+We publicized this study on the internal mailing list of students at CNSMDP, as well as two international mailing lists for research in music audio processing: AUDITORY  and ISMIR Community.\footnote{
 For more information about these mailing lists, please visit:
 \url{http://www.auditory.org/} and \url{http://ismir.net/}}
-Within two months, $K=31$ subjects accessed Cyberlioz and completed the task.
+Within two months, $K=31$ participants accessed Cyberlioz and completed the task.
+
+Personal information on the age, sex, nor musical background of participants is not collected, because the goal of our perceptual study is to build a consensus of similarity judgments, rather than to compare demographic subgroups.
+% We note that the relatively small size of our cohort ($K=31$) lacks the statistical power to carry out such a comparison.
+
+In particular, we leave the important question of the effect of musical training on the perception of auditory similarities between playing techniques as future work.
+% TODO revise the paragraph above
 
 \subsection*{Hypergraph partitioning}
+
 Once the data collection campaign was complete, we analyzed the color assignments of each subject $k$ and converted them into a cluster graph $\mathcal{G}_k$, where the integer $k$ is an anonymized subject index, ranging between $1$ and $K$.
@@
 
+For a particular subject $k$, let us denote by $C_k$ the number of clusters in the graph $\mathcal{G}_k$.
+Figure \ref{fig:cluster-histograms}(a) shows the histogram of $C_k$ across the cohort of $K=31$ participants.
+We observe that the number of clusters varies between $3$ and $19$ with a median value of $10$.
+Accordingly, the number of samples belonging to a cluster varies between $1$ (the most frequent value) and $50$, as shown in Figure \ref{fig:cluster-histograms}(b).
 %Only the clustering given by the subjects using the color labels is considered to estimate the spontaneous judgments of similarity among \ipt.
@@
 While the graphs $\mathcal{G}_1, \ldots, \mathcal{G}_K$ encode the subjective similarity judgments of participants $1$ to $K$, the graph $\mathcal{G}_0$ represents a form of consensual judgment that is shared across all participants while discarding intersubjective variability.
-Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ subjects.
+Although the rest of our paper focuses on the consensus $\mathcal{G}_0$, it is worth pointing out that the same technical framework could apply to a single subject $k$, or to a subgroup of the $K=31$ participants.
 This remark emphasizes the potential of our similarity learning method as a customizable tool for visualizing and extrapolating the timbre similarity space of a new subject.
 
-\subsection*{Inter-instrument similarity}
-
-Before addressing the question of modeling timbre in the audio domain, we conduct an exploratory study on the structure of similarity judgments themselves.
-Specifically, given a corpus of IMT samples, we ask whether human listeners tend to cluster these samples by instrument.
-To answer this question, we derive from the consensus clustering graph $\mathcal{G}_0$ an instrument-wise similarity matrix $\mathbf{A}_{\mathcal{I}}$ whose rows and columns are defined on the finite set $\mathcal{I}$ of instruments.
-For every instrument--instrument pair $(i,j)$, we set the value of $\mathbf{A}_{\mathcal{I}}$ at row $i$ and column $j$ equal to the number of edges in $\mathcal{G}_0$ connecting IMTs from instruments $i$ and $j$, renormalized by the number of IMT samples from either instrument $i$ or instrument $j$.
-
-Let us denote the existence of an edge in $\mathcal{G}_0$ connecting two IMT samples $m$ and $n$ by the binary relation $m \overset{\mathcal{G}_0}{\sim} n$.
-The mathematical definition of the matrix $\mathbf{A}_{\mathcal{I}}$ is:
-\begin{equation}
-\mathbf{A}_{\mathcal{I}}(i,j) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Instrument}(m)=i ; \mathrm{Instrument}(n)=j ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Instrument}(n) \in \{ i, j \} \big\}
-}.
-\label{eq:instrument-similarity}
-\end{equation}
-
-Then, we run Ward's method \cite{ward1963jasa} to cluster instruments in $\mathcal{I}$ according to the similarity matrix $\mathbf{A}_{\mathcal{I}}$.
-This method yields a permutation $\sigma_{\mathcal{I}}: \mathcal{I} \rightarrow \mathcal{I}$ of the rows and columns in $\mathbf{A}_{\mathcal{I}}$ such that the distance $\vert \sigma_{\mathcal{I}}(i) - \sigma_{\mathcal{I}}(j) \vert$ after permutation is small if and only if the similarity $\mathbf{A}_{\mathcal{I}}(i, j)$ is large.
-Figure \ref{fig:consensusVsI} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{I}} : (i,j)\in \mathcal{I}^2 \mapsto \mathbf{A}_{\mathcal{I}}(\sigma_{\mathcal{I}}^{-1}(i), \sigma_{\mathcal{I}}^{-1}(j))$ as a result of such agglomerative hierarchical clustering procedure.
-% \ja wants to remove this equation
-
-Interestingly, the matrix $\widetilde{\mathbf{A}}_{\mathcal{I}}$ exhibits a block diagonal structure, which roughly reflects the classical taxonomy of musical instruments: four woodwinds (BbCl, Fl, ASax, Ob) are clustered together, followed by a cluster containing one woodwind (Bn) and four brass (BBTb, Hn, TTbn, and TpC), followed by a cluster of all strings (Cb, Va, Vn, Vc, Gtr, Hp).
-Furthermore, the largest cross-instrument similarity arises for the only pair of purely plucked instruments, namely, harp and guitar.
-
-\subsection*{Inter-technique similarity}
-From the consensus clustering graph $\mathcal{G}_0$, we also derive a technique-wise similarity matrix $\mathbf{A}_{\mathcal{T}}$ whose rows and columns are defined on the finite set $\mathcal{T}$ of playing techniques.
-As for $\mathbf{A}_{\mathcal{I}}$, we define the similarity between any two playing techniques $u$ and $v$ in $\mathcal{T}$ as the following ratio of set cardinalities:
-\begin{equation}
-\mathbf{A}_{\mathcal{T}}(u,v) = \dfrac{
-\,\Card \big\{ (m, n) \,\vert\, \mathrm{Technique}(m)=u ; \mathrm{Technique}(n)=v ; m \overset{\mathcal{G}_0}{\sim} n \big\}
-}{
-\Card \big\{n \,\vert\, \mathrm{Technique}(n) \in \{ u, v \} \big\}
-}.
-\label{eq:technique-similarity}
-\end{equation}
-%\ml{why 2 ?} \vl: fixed
-Again, we run Ward's method to cluster playing techniques in $\mathcal{T}$ according to their similarity $\mathbf{A}_{\mathcal{T}}$.
-Figure \ref{fig:consensusVsPt} displays the rearranged similarity matrix $\widetilde{\mathbf{A}}_{\mathcal{T}} : (u,v)\in \mathcal{T}^2 \mapsto \mathbf{A}_{\mathcal{T}}(\sigma_{\mathcal{T}}^{-1}(u), \sigma_{\mathcal{T}}^{-1}(v))$.
-
-The block diagonal structure of the matrix $\widetilde{\mathbf{A}}_{\mathcal{T}}$ is less salient than in $\widetilde{\mathbf{A}}_{\mathcal{I}}$.
-Nevertheless, these clusters correspond to some basic attributes of qualitative timbre: by and large, the top-left region of the matrix contains sustained sounds whereas the bottom-right region contains percussive sounds.
-Interestingly, the ordinary technique (\texttt{ord}) appears at the center of the matrix, \ie{} at the intersection between sustained sounds and percussive sounds.
-This is because the notion of ``ordinariness'' does not prescribe the same gesture for all instruments: \eg{} an ordinary guitar sound is expected to be percussive whereas an ordinary flute sound is expected to be sustained.
-
-\section*{\nmu Methods}
+\section*{\nmu Machine listening methods}
 \label{sec:methods}
@@
 
-\subsection*{Visualization}
-\label{sec:visualization}
-
-At the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
-From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
-To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
-In accordance with their protocol, we compute the following matrix:
-
-\begin{equation}
-\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
-\iint_{\mathbb{R}^2}
-\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
-\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
-\end{equation}
-
-Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
-As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
-In the scientific literature on spectrotemporal receptive fields (STRF), the matrix $\mathbf{V}\boldsymbol{x}$ is called ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
-
-Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
-However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
-Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
-
-To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
-By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
-For example, Figure \ref{fig:violin-scattering} (a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
-A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
-As seen in Figure \ref{fig:violin-scattering} (b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
-Furthermore, Figure \ref{fig:violin-scattering} (c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
-
-The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
-Figures \ref{fig:violin-scattering} (d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
-These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
-These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
-Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
-This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
-
-Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
-Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering} b) or a beating tone (c), cause local energy maxima whose rate $\alpha$ is physically interpretable.
-Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering} d), key click (e), and vibrato (f) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
-We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}.
-
 \subsection*{Median-based logarithmic compression and affine standardization}
@@
 
-\section*{\nmu Results}
-\label{sec:results}
-
-The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
-We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
-In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
-However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
-
-\subsection*{Semi-supervised label propagation}
+\subsection*{Extension to diverse pitches and dynamics}
 
@@
 Given an audio stimulus $\boldsymbol{x}_n$ from our perceptual study, we seek its position in the cluster graph $\mathcal{G}_0$.
-Then, we identify its instrument--mute--technique (IMT) triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
+Then, we identify its IMT triplet, scrape for audio samples in SOL matching this triplet, and assign them all to the same cluster as the original audio stimulus.
 We repeat the same procedure for all $N=78$ nodes in $\mathcal{G}_0$, resulting in $N^\prime = 9346$ samples in total.
-This is a form of semi-supervised label propagation: from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
-% \vl: ?
+Thus, from a limited amount of human annotation, we curate a relatively large subset of SOL, amounting to about one third of the entire dataset ($9346$ out of $25444$ samples).
+
+In doing so, we assume that the perception of timbre is fully invariant to frequency transposition as well as changes in dynamics.
+This assumption coincides with the commonplace definition of timbre as an autonomous parameter of musical sound.
+Previous studies on real-world musical sounds have confirmed that listeners are able to ignore salient pitch differences while rating timbre similarity \cite{handel2001musicperception,marozeau2003dependency}, insofar as these differences do not exceed one octave.
+Since then, another study has shown that trained musicians can identify the similarity within pairs of notes from two instruments (horn and bassoon) spanning a range of $2.5$ octaves with an accuracy of $80\%$ \cite{steele2006musicperception}.
+In comparison, non-musicians has great difficulty identifying whether two notes were produced by the same instrument or not when the notes were separated by one octave.
+Thus, the influence of pitch on timbre perception appears to depend upon musical training.
+With this important caveat in mind, we leave as future work the question of disentangling the effects of pitch and musical training in the modeling of auditory similarities between instrumental playing techniques.
 
@@
 
+\section*{\nmu Results}
+\label{sec:results}
+
+The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
+We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
+In practice, for training purposes, the cluster graph $\mathcal{G}$ represents the consensus of the $K=31$ clustering provided by the users interacting with the Cyberlioz web application, which was described in the \lnameref{sec:data-collection} section ($\mathcal{G}=\mathcal{G}_0$).
+However, for evaluation purposes, this cluster graph corresponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
+
 \subsection*{Best performing system}
@@
 
-Composers may personalize their search engine $\mathbf{\Phi}$ by manually defining a cluster graph via the Cyberlioz interface (see \lnameref{sec:data-collection} section).
-This preliminary annotation lasts $30$ to $60$ minutes, which is relatively short in comparison with the duration of the $N^\prime = 9346$ audio samples in $\mathcal{G}_k$ after label propagation: i.e., roughly three hours of audio.
-
-Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ subjects of the order of $1\%$.
+Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ participants of the order of $1\%$.
 This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
@@
 
-\section*{\nmu Discussion}
-\label{sec:discussion}
-
-The previous section described our dataset of audio samples with diverse instruments, mutes, and playing techniques, as well as the method we adopt to evaluate a given system for timbre similarity retrieval.
-Over a cohort of $K=31$ subjects, we reported an average precision at rank five (AP@5) of $99.0\% \pm 1$.
+\subsection*{Visualization of joint--time frequency scattering coefficients}
+\label{sec:visualization}
+
+In the second order, joint time--frequency scattering coefficients depend upon four variables: time $t$, log-frequency $\lambda$, temporal modulation rate $\alpha$ in Hertz, and frequential modulation rate $\beta$ in cycles per octave.
+From a data visualization standpoint, rendering the four-dimensional tensor $\mathbf{S_2}\boldsymbol{x}(t, \lambda, \alpha, \beta)$ is impossible.
+To address this limitation, a recent publication has projected this tensor into a two-dimensional ``slice'', thus yielding an image raster \cite{anden2019tsp}.
+In accordance with their protocol, we compute the following matrix:
+
+\begin{equation}
+\mathbf{V}\boldsymbol{x}(\alpha, \beta) =
+\iint_{\mathbb{R}^2}
+\mathbf{U_2}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
+\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime,
+\end{equation}
+
+Observe that the equation above is a limit case of $\mathbf{S_2}$ (Equation \ref{eq:S2}) in which the constants $T$ and $F$ tend towards infinity.
+As a result, $\mathbf{V}\boldsymbol{x}$ depends solely upon scale $\alpha$ and rate $\beta$.
+In the scientific literature on STRF, the matrix $\mathbf{V}\boldsymbol{x}$ is known as the ``cortical output collapsed on the rate-scale axes'' \cite{elhilali2003speechcomm}.
+
+Previous publications on STRFs have demonstrated the interest of visualizing the slice $\mathbf{V}\boldsymbol{x}$ in the case of speech \cite{bellur2015ciss}, lung sounds \cite{emmanouilidou2012embs}, and music \cite{patil2012ploscompbiol}.
+However, the visualization of musical sounds has been restricted to some of most common playing techniques, i.e., piano played staccato and violin played pizzicato.
+Furthermore, prior publications on time--frequency scattering have displayed slices of $\mathbf{S_2}\boldsymbol{x}$ in the case of synthetic signals; but there is a gap in literature as regards the interpretability of the scale--rate domain in the case of real-world signals.
+
+To remedy this gap, we select twelve isolated notes from the SOL dataset from two instruments: violin (Figure \ref{fig:violin-scattering}) and flute (Figure \ref{fig:flute-scattering}).
+By and large, we find that joint time--frequency scattering produces comparable patterns in the scale--rate domain as the ``cortical output'' of the STRF.
+For example, Figure \ref{fig:violin-scattering}(a) shows that a violin note played in the \emph{ordinario} technique has a local energy maximum at the rate $\alpha = \SI{6}{\hertz}$.
+A visual inspection of $\mathbf{U_1}$ demonstrates that this rate coincides with the rate of vibrato of the left hand in the violin note.
+As seen in Figure \ref{fig:violin-scattering}(b), this local energy maximum is absent when the playing technique is denoted as \emph{nonvibrato}.
+Furthermore, Figure \ref{fig:violin-scattering}(c) shows that the local energy maximum is displaced to a higher rate ($\alpha = \SI{12}{\hertz}$) when the vibrato is replaced by a tremolo.
+
+The visual interpretation of playing techniques in terms of their joint time--frequency scattering coefficients is not restricted to periodic modulation, such as vibrato or tremolo: rather, it also encompasses the analysis of attack transients.
+Figures \ref{fig:violin-scattering}(d), (e), and (f) show the matrix $\mathbf{V}\boldsymbol{x}$ for three instances of impulsive violin sounds: sforzando, pizzicato, and staccato respectively.
+These three techniques create ridges in the scale--rate domain $(\alpha, \beta)$, where the cutoff rate $\alpha$ is lowest with sforzando and highest with staccato.
+These variations in cutoff rate coincide with perceptual variations in ``hardness'', i.e. impulsivity, of the violin sound.
+Moreover, in the case of staccato, we observe a slight asymmetry in the frequential scale parameter $\beta$.
+This asymmetry could be due to the fact that higher-order harmonics decay faster than the fundamental, thus yielding a triangular shape in the time--frequency domain.
+
+Figure \ref{fig:flute-scattering} shows six playing techniques of the flute.
+Similarly to the violin (Figure \ref{fig:violin-scattering}), we observe that periodic modulations, such as a trill (Figure \ref{fig:flute-scattering}(b)) or a beating tone (Figure \ref{fig:flute-scattering}(c)), cause local energy maxima whose rate $\alpha$ is physically interpretable.
+Likewise, impulsive flute sounds such as sforzando (Figure \ref{fig:flute-scattering}(d)), key click (Figure \ref{fig:flute-scattering}(e)), and vibrato (Figure \ref{fig:flute-scattering}(f)) create ridges in the scale--rate domain of varying cutoff rates $\alpha$.
+We distribute the implementation of these figures as part of the MATLAB library \emph{scattering.m}, which is released under the MIT license.\footnote{Link to download the \emph{scattering.m} library: \url{https://github.com/lostanlen/scattering.m}}
+
+\subsection*{Ablation study}
 We now turn to alter certain key choices in the design of the above-described computational blocks, and discuss their respective impacts on downstream performance.
@@
 
+These findings are in line with a previous publication \cite{patil2012ploscompbiol} which modeled perceived dissimilarity between musical sounds by means of STRF features.
+Likewise, a recent publication has observed similar performance in instrument identification of a human and machine classifier using separable Gabor filterbank (GBFB) features \cite{siedenburg2019jasa}.
+We note that both STRF and GBFB bear a strong computational resemblance with time--frequency scattering.
+
 %ML: as the learned layer is relatively thin
 
-\subsection*{Is metric learning necessary?}
-Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to a P@5 of $\textbf{76.6}\% \pm 11$.
+\subsection*{Role of metric learning}
+Replacing the LMNN metric learning algorithm by linear discriminant analysis (LDA) leads to an AP@5 of $\textbf{76.6}\% \pm 11$.
 Moreover, we evaluate the nearest neighbor algorithm $\mathbf{\Phi}$ in the absence of any metric learning at all. % \ja wants to elaborate
@@
 These features are extracted within a temporal context of duration equal to $T=\SI{1000}{\milli\second}$.
-This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCC).
-Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$.
+This value is considerably larger than the frame size of purely spectral features, such as spectral centroid, spectral flux, or mel-frequency cepstral coefficients (MFCCs).
+Indeed, the frame size of spectral features for machine listening is typically set to $T=\SI{23}{\milli\second}$, i.e., $2^{10}=1024$ samples at a sampling rate of $\SI{44,1}{\kilo\hertz}$ \cite{brown1999jasa,eronen2000icassp}.
 
 As a point of comparison, we set the maximum time scale of joint time--frequency scattering coefficients to $T=\SI{25}{\milli\second}$, hence a $40$-fold reduction in context size.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
+Over our cohort of $K=31$ participants, we report an AP@5 of $\textbf{90.9\%} \pm 4$, which is noticeably worse than the best performing system.
 This gap in performance extends the findings of a previous publication \cite{lostanlen2018extended}, which reported that metric learning with temporal scattering coefficients tends to improve with growing values of $T$, until reaching a plateau of performance around $T\sim\SI{500}{\milli\second}$.
 
-\subsection*{Is joint time--frequency scattering necessary?}
+\subsection*{Role of joint time--frequency scattering}
 
@@
 
-Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
+Over our cohort of $K=31$ participants, separable time--frequency scattering achieves n AP@5 of $\textbf{91.9\%} \pm 4$.
 This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
-This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
-
-\subsection*{What is the baseline performance?}
-Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
+This gap in performance, together with the theory of STRFs in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
+
+\subsection*{Comparison with mel--frequency cepstral coefficient (MFCC) baseline}
+Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by MFCCs.
 Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis.\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{81.8\%} \pm 7$.
-
-Arguably, this figure is not directly comparable with our best performing system (P@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
+Over our cohort of $K=31$ participants, we report an AP@5 of $\textbf{81.8\%} \pm 7$.
+
+Arguably, this figure is not directly comparable with our best performing system (AP@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
 In order to clarify the role of feature dimensionality in our computational pipeline, we apply a feature engineering technique involving multiplicative combinations of MFCC.
@@
 
-\section*{\nmu Competing interests}
+\section*{Abbreviations}
+
+\begin{itemize}
+  \item AP@5: average precision at rank 5
+  \item CNSMDP: conservatoire national sup\'erieur de musique et de danse de Paris
+  \item COSMIR: community for open and sustainable music information research
+  \item DCT: discrete cosine transform
+  \item FMA: free music archive
+  \item GBFB: Gabor filterbank features
+  \item IMT: instrument--mute--technique
+  \item LMNN: large-margin nearest neighbor
+  \item MFCC: mel-frequency cepstral coefficients
+  \item MIR: music information retrieval
+  \item P@5: precision at rank five
+  \item ReLU: rectified linear unit
+  \item STRF: spectrotemporal receptive fields
+  \item SOL: Studio On Line dataset
+\end{itemize}
+
+\subsection*{Musical instruments}
+
+\begin{itemize}
+  \item Vn: violin,
+  \item Va: viola,
+  \item Vc: cello,
+  \item Cb: contrabass,
+  \item Hp: concert harp,
+  \item Gtr: Spanish guitar,
+  \item Acc: accordion,
+  \item Fl: flute,
+  \item BbCl: soprano clarinet,
+  \item ASax: alto saxophone,
+  \item Ob: oboe,
+  \item Bn: bassoon,
+  \item TpC: trumpet in C,
+  \item Hn: French horn,
+  \item TTbn: tenor trombone,
+  \item BBTb: bass tuba.
+\end{itemize}
+
+\subsection*{Musical nuances}
+
+\begin{itemize}
+  \item \emph{pp}: pianissimo,
+  \item \emph{p}: piano,
+  \item \emph{mf}: mezzo forte,
+  \item \emph{f}: forte,
+  \item \emph{ff}: fortissimo.
+\end{itemize}
+
+\section*{Declarations}
+
+\subsection*{\nmu Availability of data and materials}
+
+The processing code, the Cyberlioz interface and an anonymized version of the  perceptual judgments gathered using this interface are available at the companion website: \url{https://github.com/mathieulagrange/lostanlen2020jasmp}.
+
+The SOL dataset is available online: \url{https://forum.ircam.fr/projects/detail/fullsol}
+
+\subsection*{\nmu Competing interests}
   The authors declare that they have no competing interests.
@@
 
-\section*{\nmu Acknowledgment}
+\subsection*{\nmu Funding}
+This work is partially supported by the Paris sciences et lettres (PSL) TICEL project.
 This work is partially supported by the European Research Council (ERC) award 320959 (InvariantClass).
 This work is partially supported by the National Science Foundation (NSF) award 1633259 (BIRDVOX).
-The Flatiron Institute is a division of the Simons Foundation.
+This work is partially supported by the Flatiron Institute, a division of the Simons Foundation.
+
+\subsection*{\nmu Acknowledgments}
+
 We wish to thank Philippe Brandeis, \'{E}tienne Graindorge, St\'{e}phane Mallat, Adrien Mamou-Mani, and Yan Maresz for contributing to the TICEL research project.
 We also wish to thank the students of the Paris Conservatory and all anonymous participants to our study.
+
+\subsection*{Authors' contributions}
+
+VL provided guidance in the design of the computational experiments and was a major contributor in writing the manuscript.
+CEH conducted part of the computational experiments.
+MR designed and implemented the listening tests.
+GL designed the listening tests and provided guidance in the design of the computational experiments.
+JA provided guidance in the design of the computational experiments, participated to the data analysis and  to the writing of the manuscript.
+ML designed and conducted the computational experiments, the data analysis and participated to the writing of the paper.
 
@@
 
-% \bibliographystyle{bmc-mathphys}
-% \bibliography{bib}
-
 %% BioMed_Central_Bib_Style_v1.01
 
-\begin{thebibliography}{78}
+\begin{thebibliography}{85}
 % BibTex style file: bmc-mathphys.bst (version 2.1), 2014-07-24
@@
 %%% 42
+\bibitem{siedenburg2016frontiers}
+\begin{barticle}
+\bauthor{\bsnm{Siedenburg}, \binits{K.}},
+\bauthor{\bsnm{Jones-Mollerup}, \binits{K.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+\batitle{Acoustic and categorical dissimilarity of musical timbre: Evidence
+  from asymmetries between acoustic and chimeric sounds}.
+\bjtitle{Frontiers in Psychology}
+\bvolume{6},
+\bfpage{1977}
+(\byear{2016})
+\end{barticle}
+\endbibitem
+
+%%% 43
 \bibitem{depireux2001jneur}
@@
 
-%%% 43
+%%% 44
 \bibitem{aertsen1981biolcyb}
@@
 
-%%% 44
+%%% 45
 \bibitem{deboer1968biomed}
@@
 
-%%% 45
+%%% 46
 \bibitem{flandrin1998book}
@@
 
-%%% 46
+%%% 47
 \bibitem{eggermont1993hearing}
@@
 
-%%% 47
+%%% 48
 \bibitem{klein2000compneur}
@@
 
-%%% 48
+%%% 49
 \bibitem{theunissen2000jneur}
@@
 
-%%% 49
+%%% 50
 \bibitem{chi2005jasa}
@@
 
-%%% 50
+%%% 51
 \bibitem{patil2015eurasip}
@@
 
-%%% 51
+%%% 52
 \bibitem{thoret2016jasa}
@@
 
-%%% 52
+%%% 53
 \bibitem{mishra2018ismir}
@@
 
-%%% 53
+%%% 54
 \bibitem{hemery2015frontiers}
@@
 
-%%% 54
+%%% 55
 \bibitem{andreux2020jmlr}
@@
 
-%%% 55
+%%% 56
 \bibitem{lostanlen2019dafx}
@@
 
-%%% 56
+%%% 57
 \bibitem{mallat2016philtrans}
@@
 
-%%% 57
+%%% 58
+\bibitem{caetano2019chapter}
+\begin{bchapter}
+\bauthor{\bsnm{Caetano}, \binits{M.}},
+\bauthor{\bsnm{Saitis}, \binits{C.}},
+\bauthor{\bsnm{Siedenburg}, \binits{K.}}:
+\bctitle{Audio content descriptors of timbre}.
+In: \bbtitle{Timbre: Acoustics, Perception, and Cognition},
+pp. \bfpage{297}--\blpage{333}.
+\bpublisher{Springer}, \blocation{???}
+(\byear{2019})
+\end{bchapter}
+\endbibitem
+
+%%% 59
 \bibitem{wu2018taslp}
@@
 
-%%% 58
+%%% 60
 \bibitem{pearce2019appliedsciences}
@@
 
-%%% 59
+%%% 61
+\bibitem{giordano2011multivariate}
+\begin{barticle}
+\bauthor{\bsnm{Giordano}, \binits{B.L.}},
+\bauthor{\bsnm{Guastavino}, \binits{C.}},
+\bauthor{\bsnm{Murphy}, \binits{E.}},
+\bauthor{\bsnm{Ogg}, \binits{M.}},
+\bauthor{\bsnm{Smith}, \binits{B.K.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}}:
+\batitle{Comparison of methods for collecting and modeling dissimilarity data:
+  Applications to complex sound stimuli}.
+\bjtitle{Multivariate behavioral research}
+\bvolume{46}(\bissue{5}),
+\bfpage{779}--\blpage{811}
+(\byear{2011})
+\end{barticle}
+\endbibitem
+
+%%% 62
+\bibitem{elliott2013jasa}
+\begin{barticle}
+\bauthor{\bsnm{Elliott}, \binits{T.M.}},
+\bauthor{\bsnm{Hamilton}, \binits{L.S.}},
+\bauthor{\bsnm{Theunissen}, \binits{F.E.}}:
+\batitle{Acoustic structure of the five perceptual dimensions of timbre in
+  orchestral instrument tones}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{133}(\bissue{1}),
+\bfpage{389}--\blpage{404}
+(\byear{2013})
+\end{barticle}
+\endbibitem
+
+%%% 63
 \bibitem{kernighan1970efficient}
@@
 
-%%% 60
+%%% 64
 \bibitem{han1997scalable}
@@
 
-%%% 61
+%%% 65
 \bibitem{strehl2002cluster}
@@
 
-%%% 62
-\bibitem{ward1963jasa}
-\begin{barticle}
-\bauthor{\bsnm{Ward~Jr.}, \binits{J.H.}}:
-\batitle{Hierarchical grouping to optimize an objective function}.
-\bjtitle{Journal of the American Statistical Association}
-\bvolume{58}(\bissue{301}),
-\bfpage{236}--\blpage{244}
-(\byear{1963})
-\end{barticle}
-\endbibitem
-
-%%% 63
+%%% 66
 \bibitem{schorkhuber2010smc}
@@
 
-%%% 64
+%%% 67
 \bibitem{lostanlen2020icassp}
@@
 
-%%% 65
+%%% 68
 \bibitem{mallat2012cpam}
@@
 \end{barticle}
-\endbibitem
-
-%%% 66
-\bibitem{elhilali2003speechcomm}
-\begin{barticle}
-\bauthor{\bsnm{Elhilali}, \binits{M.}},
-\bauthor{\bsnm{Chi}, \binits{T.}},
-\bauthor{\bsnm{Shamma}, \binits{S.A.}}:
-\batitle{A spectro-temporal modulation index (stmi) for assessment of speech
-  intelligibility}.
-\bjtitle{Speech communication}
-\bvolume{41}(\bissue{2-3}),
-\bfpage{331}--\blpage{348}
-(\byear{2003})
-\end{barticle}
-\endbibitem
-
-%%% 67
-\bibitem{bellur2015ciss}
-\begin{bchapter}
-\bauthor{\bsnm{Bellur}, \binits{A.}},
-\bauthor{\bsnm{Elhilali}, \binits{M.}}:
-\bctitle{Detection of speech tokens in noise using adaptive spectrotemporal
-  receptive fields}.
-In: \bbtitle{Proceedings of the Annual Conference on Information Sciences and
-  Systems (CISS)},
-pp. \bfpage{1}--\blpage{6}
-(\byear{2015}).
-\bcomment{IEEE}
-\end{bchapter}
-\endbibitem
-
-%%% 68
-\bibitem{emmanouilidou2012embs}
-\begin{bchapter}
-\bauthor{\bsnm{Emmanouilidou}, \binits{D.}},
-\bauthor{\bsnm{Patil}, \binits{K.}},
-\bauthor{\bsnm{West}, \binits{J.}},
-\bauthor{\bsnm{Elhilali}, \binits{M.}}:
-\bctitle{A multiresolution analysis for detection of abnormal lung sounds}.
-In: \bbtitle{Proceedings of the International Conference of the IEEE
-  Engineering in Medicine and Biology Society (EMBS)},
-pp. \bfpage{3139}--\blpage{3142}
-(\byear{2012}).
-\bcomment{IEEE}
-\end{bchapter}
 \endbibitem
@@
 %%% 71
+\bibitem{handel2001musicperception}
+\begin{barticle}
+\bauthor{\bsnm{Handel}, \binits{S.}},
+\bauthor{\bsnm{Erickson}, \binits{M.L.}}:
+\batitle{A rule of thumb: The bandwidth for timbre invariance is one octave}.
+\bjtitle{Music Perception}
+\bvolume{19}(\bissue{1}),
+\bfpage{121}--\blpage{126}
+(\byear{2001})
+\end{barticle}
+\endbibitem
+
+%%% 72
+\bibitem{marozeau2003dependency}
+\begin{barticle}
+\bauthor{\bsnm{Marozeau}, \binits{J.}},
+\bauthor{\bparticle{de} \bsnm{Cheveign{\'e}}, \binits{A.}},
+\bauthor{\bsnm{McAdams}, \binits{S.}},
+\bauthor{\bsnm{Winsberg}, \binits{S.}}:
+\batitle{The dependency of timbre on fundamental frequency}.
+\bjtitle{The Journal of the Acoustical Society of America}
+\bvolume{114}(\bissue{5}),
+\bfpage{2946}--\blpage{2957}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 73
+\bibitem{steele2006musicperception}
+\begin{barticle}
+\bauthor{\bsnm{Steele}, \binits{K.M.}},
+\bauthor{\bsnm{Williams}, \binits{A.K.}}:
+\batitle{Is the bandwidth for timbre invariance only one octave?}
+\bjtitle{Music Perception}
+\bvolume{23}(\bissue{3}),
+\bfpage{215}--\blpage{220}
+(\byear{2006})
+\end{barticle}
+\endbibitem
+
+%%% 74
 \bibitem{wang2020icassp}
@@
 
-%%% 72
+%%% 75
+\bibitem{elhilali2003speechcomm}
+\begin{barticle}
+\bauthor{\bsnm{Elhilali}, \binits{M.}},
+\bauthor{\bsnm{Chi}, \binits{T.}},
+\bauthor{\bsnm{Shamma}, \binits{S.A.}}:
+\batitle{A spectro-temporal modulation index (stmi) for assessment of speech
+  intelligibility}.
+\bjtitle{Speech communication}
+\bvolume{41}(\bissue{2-3}),
+\bfpage{331}--\blpage{348}
+(\byear{2003})
+\end{barticle}
+\endbibitem
+
+%%% 76
+\bibitem{bellur2015ciss}
+\begin{bchapter}
+\bauthor{\bsnm{Bellur}, \binits{A.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\bctitle{Detection of speech tokens in noise using adaptive spectrotemporal
+  receptive fields}.
+In: \bbtitle{Proceedings of the Annual Conference on Information Sciences and
+  Systems (CISS)},
+pp. \bfpage{1}--\blpage{6}
+(\byear{2015}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 77
+\bibitem{emmanouilidou2012embs}
+\begin{bchapter}
+\bauthor{\bsnm{Emmanouilidou}, \binits{D.}},
+\bauthor{\bsnm{Patil}, \binits{K.}},
+\bauthor{\bsnm{West}, \binits{J.}},
+\bauthor{\bsnm{Elhilali}, \binits{M.}}:
+\bctitle{A multiresolution analysis for detection of abnormal lung sounds}.
+In: \bbtitle{Proceedings of the International Conference of the IEEE
+  Engineering in Medicine and Biology Society (EMBS)},
+pp. \bfpage{3139}--\blpage{3142}
+(\byear{2012}).
+\bcomment{IEEE}
+\end{bchapter}
+\endbibitem
+
+%%% 78
 \bibitem{black2012dictionary}
@@
 
-%%% 73
+%%% 79
 \bibitem{cella2020icmc}
@@
 
-%%% 74
+%%% 80
+\bibitem{siedenburg2019jasa}
+\begin{barticle}
+\bauthor{\bsnm{Siedenburg}, \binits{K.}},
+\bauthor{\bsnm{Sch{\"a}dler}, \binits{M.R.}},
+\bauthor{\bsnm{H{\"u}lsmeier}, \binits{D.}}:
+\batitle{Modeling the onset advantage in musical instrument recognition}.
+\bjtitle{The journal of the Acoustical Society of America}
+\bvolume{146}(\bissue{6}),
+\bfpage{523}--\blpage{529}
+(\byear{2019})
+\end{barticle}
+\endbibitem
+
+%%% 81
 \bibitem{lostanlen2019chapter}
@@
 
-%%% 75
+%%% 82
 \bibitem{bauge2013icassp}
@@
 
-%%% 76
+%%% 83
 \bibitem{wang2019ismir}
@@
 
-%%% 77
+%%% 84
 \bibitem{schadler2015jasa}
@@
 
-%%% 78
+%%% 85
 \bibitem{chang2010jmlr}
@@
 <bibl id="B42">
+  <title><p>Acoustic and Categorical Dissimilarity of Musical Timbre: Evidence
+  from Asymmetries Between Acoustic and Chimeric Sounds</p></title>
+  <aug>
+    <au><snm>Siedenburg</snm><fnm>K</fnm></au>
+    <au><snm>Jones Mollerup</snm><fnm>K</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Frontiers in Psychology</source>
+  <pubdate>2016</pubdate>
+  <volume>6</volume>
+  <fpage>1977</fpage>
+</bibl>
+
+<bibl id="B43">
   <title><p>Spectro-temporal response field characterization with dynamic
@@
 
-<bibl id="B43">
+<bibl id="B44">
   <title><p>The spectro-temporal receptive field</p></title>
@@
 
-<bibl id="B44">
+<bibl id="B45">
   <title><p>Triggered correlation</p></title>
@@
 
-<bibl id="B45">
+<bibl id="B46">
   <title><p>Time-frequency/time-scale analysis</p></title>
@@
 
-<bibl id="B46">
+<bibl id="B47">
   <title><p>Wiener and Volterra analyses applied to the auditory
@@
 
-<bibl id="B47">
+<bibl id="B48">
   <title><p>Robust spectrotemporal reverse correlation for the auditory system:
@@
 
-<bibl id="B48">
+<bibl id="B49">
   <title><p>Spectral-temporal receptive fields of nonlinear auditory neurons
@@
 
-<bibl id="B49">
+<bibl id="B50">
   <title><p>Multiresolution spectrotemporal analysis of complex
@@
 
-<bibl id="B50">
+<bibl id="B51">
   <title><p>Biomimetic spectro-temporal features for music instrument
@@
 
-<bibl id="B51">
+<bibl id="B52">
   <title><p>Perceptually salient spectrotemporal modulations for recognition of
@@
 
-<bibl id="B52">
+<bibl id="B53">
   <title><p>Understanding a Deep Machine Listening Model Through Feature
@@
 
-<bibl id="B53">
+<bibl id="B54">
   <title><p>One hundred ways to process time, frequency, rate and scale in the
@@
 
-<bibl id="B54">
+<bibl id="B55">
   <title><p>Kymatio: Scattering Transforms in Python</p></title>
@@
 
-<bibl id="B55">
+<bibl id="B56">
   <title><p>The Shape of RemiXXXes to Come: Audio texture synthesis with
@@
 
-<bibl id="B56">
+<bibl id="B57">
   <title><p>Understanding deep convolutional networks</p></title>
@@
 
-<bibl id="B57">
+<bibl id="B58">
+  <title><p>Audio content descriptors of timbre</p></title>
+  <aug>
+    <au><snm>Caetano</snm><fnm>M</fnm></au>
+    <au><snm>Saitis</snm><fnm>C</fnm></au>
+    <au><snm>Siedenburg</snm><fnm>K</fnm></au>
+  </aug>
+  <source>Timbre: Acoustics, perception, and cognition</source>
+  <publisher>Springer</publisher>
+  <pubdate>2019</pubdate>
+  <fpage>297</fpage>
+  <lpage>-333</lpage>
+</bibl>
+
+<bibl id="B59">
   <title><p>A review of automatic drum transcription</p></title>
@@
 
-<bibl id="B58">
+<bibl id="B60">
   <title><p>Modelling Timbral Hardness</p></title>
@@
 
-<bibl id="B59">
+<bibl id="B61">
+  <title><p>Comparison of methods for collecting and modeling dissimilarity
+  data: Applications to complex sound stimuli</p></title>
+  <aug>
+    <au><snm>Giordano</snm><fnm>BL</fnm></au>
+    <au><snm>Guastavino</snm><fnm>C</fnm></au>
+    <au><snm>Murphy</snm><fnm>E</fnm></au>
+    <au><snm>Ogg</snm><fnm>M</fnm></au>
+    <au><snm>Smith</snm><fnm>BK</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+  </aug>
+  <source>Multivariate behavioral research</source>
+  <publisher>Taylor \& Francis</publisher>
+  <pubdate>2011</pubdate>
+  <volume>46</volume>
+  <issue>5</issue>
+  <fpage>779</fpage>
+  <lpage>-811</lpage>
+</bibl>
+
+<bibl id="B62">
+  <title><p>Acoustic structure of the five perceptual dimensions of timbre in
+  orchestral instrument tones</p></title>
+  <aug>
+    <au><snm>Elliott</snm><fnm>TM</fnm></au>
+    <au><snm>Hamilton</snm><fnm>LS</fnm></au>
+    <au><snm>Theunissen</snm><fnm>FE</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>Acoustical Society of America</publisher>
+  <pubdate>2013</pubdate>
+  <volume>133</volume>
+  <issue>1</issue>
+  <fpage>389</fpage>
+  <lpage>-404</lpage>
+</bibl>
+
+<bibl id="B63">
   <title><p>An efficient heuristic procedure for partitioning
@@
 
-<bibl id="B60">
+<bibl id="B64">
   <title><p>Scalable parallel data mining for association rules</p></title>
@@
 
-<bibl id="B61">
+<bibl id="B65">
   <title><p>Cluster ensembles---a knowledge reuse framework for combining
@@
 
-<bibl id="B62">
-  <title><p>Hierarchical grouping to optimize an objective function</p></title>
-  <aug>
-    <au><snm>Ward Jr.</snm><fnm>JH</fnm></au>
-  </aug>
-  <source>Journal of the American Statistical Association</source>
-  <publisher>Taylor \& Francis Group</publisher>
-  <pubdate>1963</pubdate>
-  <volume>58</volume>
-  <issue>301</issue>
-  <fpage>236</fpage>
-  <lpage>-244</lpage>
-</bibl>
-
-<bibl id="B63">
+<bibl id="B66">
   <title><p>Constant-Q transform toolbox for music processing</p></title>
@@
 
-<bibl id="B64">
+<bibl id="B67">
   <title><p>Learning the helix topology of musical pitch</p></title>
@@
 
-<bibl id="B65">
+<bibl id="B68">
   <title><p>Group invariant scattering</p></title>
@@
   <lpage>-1398</lpage>
-</bibl>
-
-<bibl id="B66">
-  <title><p>A spectro-temporal modulation index (STMI) for assessment of speech
-  intelligibility</p></title>
-  <aug>
-    <au><snm>Elhilali</snm><fnm>M</fnm></au>
-    <au><snm>Chi</snm><fnm>T</fnm></au>
-    <au><snm>Shamma</snm><fnm>SA</fnm></au>
-  </aug>
-  <source>Speech communication</source>
-  <publisher>Elsevier</publisher>
-  <pubdate>2003</pubdate>
-  <volume>41</volume>
-  <issue>2-3</issue>
-  <fpage>331</fpage>
-  <lpage>-348</lpage>
-</bibl>
-
-<bibl id="B67">
-  <title><p>Detection of speech tokens in noise using adaptive spectrotemporal
-  receptive fields</p></title>
-  <aug>
-    <au><snm>Bellur</snm><fnm>A</fnm></au>
-    <au><snm>Elhilali</snm><fnm>M</fnm></au>
-  </aug>
-  <source>Proceedings of the Annual Conference on Information Sciences and
-  Systems (CISS)</source>
-  <pubdate>2015</pubdate>
-  <fpage>1</fpage>
-  <lpage>-6</lpage>
-</bibl>
-
-<bibl id="B68">
-  <title><p>A multiresolution analysis for detection of abnormal lung
-  sounds</p></title>
-  <aug>
-    <au><snm>Emmanouilidou</snm><fnm>D</fnm></au>
-    <au><snm>Patil</snm><fnm>K</fnm></au>
-    <au><snm>West</snm><fnm>J</fnm></au>
-    <au><snm>Elhilali</snm><fnm>M</fnm></au>
-  </aug>
-  <source>Proceedings of the International Conference of the IEEE Engineering
-  in Medicine and Biology Society (EMBS)</source>
-  <pubdate>2012</pubdate>
-  <fpage>3139</fpage>
-  <lpage>-3142</lpage>
 </bibl>
@@
 <bibl id="B71">
+  <title><p>A rule of thumb: The bandwidth for timbre invariance is one
+  octave</p></title>
+  <aug>
+    <au><snm>Handel</snm><fnm>S</fnm></au>
+    <au><snm>Erickson</snm><fnm>ML</fnm></au>
+  </aug>
+  <source>Music Perception</source>
+  <publisher>University of California Press Journals</publisher>
+  <pubdate>2001</pubdate>
+  <volume>19</volume>
+  <issue>1</issue>
+  <fpage>121</fpage>
+  <lpage>-126</lpage>
+</bibl>
+
+<bibl id="B72">
+  <title><p>The dependency of timbre on fundamental frequency</p></title>
+  <aug>
+    <au><snm>Marozeau</snm><fnm>J</fnm></au>
+    <au><snm>Cheveign{\'e}</snm><fnm>A</fnm></au>
+    <au><snm>McAdams</snm><fnm>S</fnm></au>
+    <au><snm>Winsberg</snm><fnm>S</fnm></au>
+  </aug>
+  <source>The Journal of the Acoustical Society of America</source>
+  <publisher>ASA</publisher>
+  <pubdate>2003</pubdate>
+  <volume>114</volume>
+  <issue>5</issue>
+  <fpage>2946</fpage>
+  <lpage>-2957</lpage>
+</bibl>
+
+<bibl id="B73">
+  <title><p>Is the bandwidth for timbre invariance only one octave?</p></title>
+  <aug>
+    <au><snm>Steele</snm><fnm>KM</fnm></au>
+    <au><snm>Williams</snm><fnm>AK</fnm></au>
+  </aug>
+  <source>Music Perception</source>
+  <publisher>JSTOR</publisher>
+  <pubdate>2006</pubdate>
+  <volume>23</volume>
+  <issue>3</issue>
+  <fpage>215</fpage>
+  <lpage>-220</lpage>
+</bibl>
+
+<bibl id="B74">
   <title><p>Playing Technique Recognition by Joint Time--Frequency
@@
 
-<bibl id="B72">
+<bibl id="B75">
+  <title><p>A spectro-temporal modulation index (STMI) for assessment of speech
+  intelligibility</p></title>
+  <aug>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+    <au><snm>Chi</snm><fnm>T</fnm></au>
+    <au><snm>Shamma</snm><fnm>SA</fnm></au>
+  </aug>
+  <source>Speech communication</source>
+  <publisher>Elsevier</publisher>
+  <pubdate>2003</pubdate>
+  <volume>41</volume>
+  <issue>2-3</issue>
+  <fpage>331</fpage>
+  <lpage>-348</lpage>
+</bibl>
+
+<bibl id="B76">
+  <title><p>Detection of speech tokens in noise using adaptive spectrotemporal
+  receptive fields</p></title>
+  <aug>
+    <au><snm>Bellur</snm><fnm>A</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>Proceedings of the Annual Conference on Information Sciences and
+  Systems (CISS)</source>
+  <pubdate>2015</pubdate>
+  <fpage>1</fpage>
+  <lpage>-6</lpage>
+</bibl>
+
+<bibl id="B77">
+  <title><p>A multiresolution analysis for detection of abnormal lung
+  sounds</p></title>
+  <aug>
+    <au><snm>Emmanouilidou</snm><fnm>D</fnm></au>
+    <au><snm>Patil</snm><fnm>K</fnm></au>
+    <au><snm>West</snm><fnm>J</fnm></au>
+    <au><snm>Elhilali</snm><fnm>M</fnm></au>
+  </aug>
+  <source>Proceedings of the International Conference of the IEEE Engineering
+  in Medicine and Biology Society (EMBS)</source>
+  <pubdate>2012</pubdate>
+  <fpage>3139</fpage>
+  <lpage>-3142</lpage>
+</bibl>
+
+<bibl id="B78">
   <title><p>A dictionary of economics</p></title>
@@
 
-<bibl id="B73">
+<bibl id="B79">
   <title><p>OrchideaSOL: A Dataset of Extended Instrumental Techniques for
@@
 
-<bibl id="B74">
+<bibl id="B80">
+  <title><p>Modeling the onset advantage in musical instrument
+  recognition</p></title>
+  <aug>
+    <au><snm>Siedenburg</snm><fnm>K</fnm></au>
+    <au><snm>Sch{\"a}dler</snm><fnm>MR</fnm></au>
+    <au><snm>H{\"u}lsmeier</snm><fnm>D</fnm></au>
+  </aug>
+  <source>The journal of the Acoustical Society of America</source>
+  <publisher>Acoustical Society of America</publisher>
+  <pubdate>2019</pubdate>
+  <volume>146</volume>
+  <issue>6</issue>
+  <fpage>EL523</fpage>
+  <lpage>-EL529</lpage>
+</bibl>
+
+<bibl id="B81">
   <title><p>On Time-frequency Scattering and Computer Music</p></title>
@@
 
-<bibl id="B75">
+<bibl id="B82">
   <title><p>Representing environmental sounds using the separable scattering
@@
 
-<bibl id="B76">
+<bibl id="B83">
   <title><p>Adaptive Time--Frequency Scattering for Periodic Modulation
@@
 
-<bibl id="B77">
+<bibl id="B84">
   <title><p>Separable spectro-temporal Gabor filter bank features: Reducing the
@@
 
-<bibl id="B78">
+<bibl id="B85">
   <title><p>Training and testing low-degree polynomial data mappings via linear
@@
 \center
-\includegraphics[width = 0.85\textwidth]{consensusVsI.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{I}}$ of perceived timbre similarities between instruments from our dataset of $N=78$. samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:instrument-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ were re-arranged according to Ward's minimum variance method.
-Observe that the block diagonal structure in $\mathbf{\widetilde{A}}_{\mathcal{I}}$ reflects an organological taxonomy of musical instruments: woodwinds, brass, and strings.
-See \lnameref{sec:data-collection} section for details, including the meaning of each instrument abbreviation.
-}
-\label{fig:consensusVsI}
-\end{figure}
-
-\begin{figure}
-\center
-\includegraphics[width = 0.85\textwidth]{consensusVsPt.png}
-\caption{
-Matrix $\mathbf{\widetilde{A}}_{\mathcal{T}}$ of perceived timbre similarities between playing techniques from our dataset of $N=78$ samples.
-Darker shades indicate higher frequencies of co-occurrence in the cluster graph $\mathcal{G}_0$ (see Equation \ref{eq:technique-similarity}), obtained from a consensus of $K=31$ subjects.
-The rows and columns in $\mathbf{\widetilde{A}}_{\mathcal{T}}$ were re-arranged according to Ward's minimum variance method.
-See \lnameref{sec:data-collection} section for details,.
-See Tables \ref{tab:list1} and \ref{tab:list2} for an English description of each playing technique abbreviation.
-%TODO add a double arrow?
-% rough <-- soft --> rough
-% sustained <-- "ordinary" --> percussive
-}
-\label{fig:consensusVsPt}
+\begin{tabular}{cc}
+  (a) & (b) \\
+  \includegraphics[
+  	width = 0.47\textwidth, trim = 10mm 0 5mm 0, clip]
+	  {nbc.png} &
+  \includegraphics[
+    width = 0.47\textwidth, trim = 10mm 0 5mm 0, clip]
+      {sbc.png}
+\end{tabular}
+\caption{\csentence{Inter-subject variability}.
+Histogram of the number of clusters (a) and the size of the clusters (b) defined by the $31$ subjects. %% TODO UPDATE
+See \lnameref{sec:data-collection} for details.}
+\label{fig:cluster-histograms}
 \end{figure}

### 2020-11-18 proofs/report/lagrangeTechReport.tex
create proofs/ folder

--- 2018-02-13 paper/lagrangeSpontaneousSimilarity2017.tex
@@
 Vincent Lostanlen,
-Mathias Rossignol,
+Mathias Rossignol, \\
 Gr\'egoire Lafay,
-Mathieu Lagrange}
+and Mathieu Lagrange}
 
@@
 
-Musical timbre is a multi-faceted notion that have been extensively studied,
-mostly by focusing on its spectral aspect.
-The aim of this paper is to better understand the influence
-of the playing technique on the perception of musical timbre.
-We do so by studying the spontaneous judgments of similarity between several
-instruments played with a diverse set of playing techniques,
+Musical timbre is a multi-faceted notion that have been extensively studied, mostly by focusing on its spectral aspect.
+The aim of this paper is to better understand the influence of the playing technique on the perception of musical timbre.
+We do so by studying the spontaneous judgments of similarity between several instruments played with a diverse set of playing techniques,
 
-A first experiment is conducted by music experts to identify
-which couple of instrument / playing technique in a given dataset
-is worth comparing to other couples of instrument / playing technique.
-In a second experiment, spontaneous similarity
-judgments among the retained couples are collected
-using a canonic free sorting task experiment design.
+A first experiment is conducted by music experts to identify which couple of instrument / playing technique in a given dataset is worth comparing to other couples of instrument / playing technique.
+In a second experiment, spontaneous similarity judgments among the retained couples are collected using a canonic free sorting task experiment design.
 
-To further study the outcomes of this experiment, we assume a
-two-step computational model of human perception, where the acoustic signal
-is fed to a statically designed processing unit that accounts for frequency
-and temporal modulations.
-The resulting features are then projected in a supervised manner so that
-the resulting description space matches the perceptual similarity judgments
-gathered in the second experiment.
-Numerical experiments show that the induced perceptual space
-is able to approximate perceptual data with satisfying accuracy.
+To further study the outcomes of this experiment, we assume a two-step computational model of human perception, where the acoustic signal is fed to a statically designed processing unit that accounts for frequency and temporal modulations.
+The resulting features are then projected in a supervised manner so that the resulting description space matches the perceptual similarity judgments gathered in the second experiment.
+Numerical experiments show that the induced perceptual space is able to approximate perceptual data with satisfying accuracy.
 
@@
 
+% wastebasket notion
+
+\vl{below is the original introduction}
+
 Being perceptively defined, musical timbre is an interesting notion to study to better understand human auditory perception \cite{grey1977multidimensional}.
@@
 In order to further restrict this diversity, we focus in this paper on the influence of the playing technique used for the production of a single tone on the perception of timbre.
-A popular view of this kind of stimuli is that they can be described by a distribution of energy that evolve through time and across frequency.
+A popular view of this kind of stimuli is that they can be described by a distribution of energy that evolves through time and across frequency.
 By following this trend, most researchers translate the exclusion definition of timbre \cite{marozeau2003dependency} into a more specified definition.
@@
 Agus \& al effectively showed that instrument recognition can be achieved with very little listening time \cite{agus2012fast}.
-However, instrument recognition is only part of musical timbre perception.
-We propose in this paper to study the influence of the playing technique on musical timbre percept.
+However, instrument recognition is only part of musical timbre perception. We propose in this paper to study the influence of the playing technique on musical timbre percept.
 Evidently, the influence of the playing technique can be measured in terms of energy distribution along the frequency axis but also with respect to time on longer time scales than usually considered.
@@
 pp. 27?41. @biblio}
+
+\vl{Automatic taxonomies: Essid et al. TASLP 2009.}
 
@@
 \center
-\includegraphics[width = \textwidth]{figures/AllClassInstruments33}
+%\includegraphics[width = \textwidth]{figures/AllClassInstruments33}
 \caption{Mds projection with instruments as labels.}
@@
 \center
-\includegraphics[width = \textwidth]{figures/AllClassConsensus33}
+%\includegraphics[width = \textwidth]{figures/AllClassConsensus33}
 \caption{Mds projection with consensus clustering as labels.}

### 2020-11-18 proofs/submission/paperV1.tex
create proofs/ folder

--- 2020-07-21 paper/submission/paper.tex
@@
 
-Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves a precision at rank five (P@5) of $\textbf{91.9\%} \pm 4$.
+Over our cohort of $K=31$ subjects, separable time--frequency scattering achieves n AP@5 of $\textbf{91.9\%} \pm 4$.
 This figure is noticeably worse than joint time--frequency scattering ($\textbf{99.0}\% \pm 1$), all other things being equal.
-This gap in performance, together with the theory of spectrotemporal receptive fields (STRF) in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
+This gap in performance, together with the theory of STRFs in auditory neuroscience (see \lnameref{sec:related-work} section), demonstrates the importance of joint spectrotemporal modulations in the modeling of timbre similarity across instrumental playing techniques.
 
 \subsection*{What is the baseline performance?}
-Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by mel-frequency cepstral coefficients (MFCC).
+Lastly, we train a baseline system in which joint time--frequency scattering coefficients are replaced by MFCCs.
 Specifically, we extract a 40-band mel-frequency spectrogram by means of the RASTAMAT library; apply the pointwise logarithm; and compute a discrete cosine transform (DCT) over the mel-frequency axis.\footnote{Link to RASTAMAT library: \url{http://www.ee.columbia.edu/~dpwe/resources/matlab/rastamat}}
 This operation results in a 40-dimensional feature vector over frames of duration $T=\SI{25}{\milli\second}$.
-Over our cohort of $K=31$ subjects, we report a precision at rank five (P@5) of $\textbf{81.8\%} \pm 7$.
-
-Arguably, this figure is not directly comparable with our best performing system (P@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
+Over our cohort of $K=31$ subjects, we report an AP@5 of $\textbf{81.8\%} \pm 7$.
+
+Arguably, this figure is not directly comparable with our best performing system (AP@5 of $99.0\%\pm 1$), due to the mismatch in dimensionality between MFCC and joint time--frequency scattering coefficients.
 In order to clarify the role of feature dimensionality in our computational pipeline, we apply a feature engineering technique involving multiplicative combinations of MFCC.
@@
 
-\section*{\nmu Competing interests}
+\section*{Abbreviations}
+
+\begin{itemize}
+  \item AP@5: average precision at rank 5
+  \item CNSMDP: conservatoire national sup\'erieur de musique et de danse de Paris
+  \item COSMIR: community for open and sustainable music information research
+  \item DCT: discrete cosine transform
+  \item FMA: free music archive
+  \item GBFB: Gabor filterbank features
+  \item IMT: instrument--mute--technique
+  \item LMNN: large-margin nearest neighbor
+  \item MFCC: mel-frequency cepstral coefficients
+  \item MIR: music information retrieval
+  \item P@5: precision at rank five
+  \item ReLU: rectified linear unit
+  \item STRF: spectrotemporal receptive fields
+  \item SOL: studio On Line dataset
+\end{itemize}
+
+\subsection*{Musical instruments}
+
+\begin{itemize}
+  \item Vn: violin,
+  \item Va: viola,
+  \item Vc: celloVc,
+  \item Cb: contrabass,
+  \item Hp: concert harp,
+  \item Gtr: Spanish guitar,
+  \item Acc: accordion,
+  \item Fl: flute,
+  \item BbCl: soprano clarinet,
+  \item ASax: alto saxophone,
+  \item Ob: oboe,
+  \item Bn: bassoon,
+  \item TpC: trumpet in C,
+  \item Hn: French horn,
+  \item TTbn: tenor trombone,
+  \item BBTb: bass tuba.
+\end{itemize}
+
+\subsection*{Musical nuances}
+
+\begin{itemize}
+  \item \emph{pp}: pianissimo,
+  \item \emph{p}: piano,
+  \item \emph{mf}: mezzo forte,
+  \item \emph{f}: forte,
+  \item \emph{ff}: fortissimo.
+\end{itemize}
+
+\section*{Declarations}
+
+\subsection*{\nmu Availability of data and materials}
+
+The processing code, the cyberlioz interface and an anonimized version of the  perceptual judgments gathered using this interface are available at the companion website: \url{https://github.com/mathieulagrange/lostanlen2020jasmp}.
+
+The SOL dataset is available online: \url{https://forum.ircam.fr/projects/detail/fullsol}
+
+\subsection*{\nmu Competing interests}
   The authors declare that they have no competing interests.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \bibliographystyle{bmc-mathphys}
-% \bibliography{bib}
 
